### 2020-03-14 04-CCNA.tex
Add files via upload

% !TEX root = ./main_RS2L_LNCS.tex

\section{Chained \CNA \ (c\CNA)}
\label{sec:ccna}

In this section we introduce  the syntax and operational semantics
of a variant of the {\tt link}-calculus~\cite{BodeiBB12}, the c\CNA\  (chained \CNA)
where the prefixes are link chains.
%In this section we just focus on the basic ideas of link
% interaction, while we shall enhance the model with name mobility in
% the next section.

\paragraph{Link Chains.}
Let $\mathcal{C}$ be the set of channels, ranged over by $a,b,...$, and 
let $\mathcal{A} = \mathcal{C} \cup \setof{\silent} \cup \setof{{\noacts}}$ be the set of actions, ranged over by $\alpha,\beta,...$,
where the symbol $\silent$ denotes a \emph{silent} action, while the symbol $\noact$ denotes a \emph{virtual} (non-specified) action.
A \emph{link} is a pair $\ell=\link{\alpha}{\beta}$;
it is \emph{solid} if $\alpha,\beta \neq \ \noacts$; 
the link $\link{\noact}{\noact}$ is called \emph{virtual}.
A link is \emph{valid} if it is solid or virtual.
We let $\mathcal{L}$ be the set of valid links.
%
A \emph{link chain} is a 
finite sequence $v = \ell_{1}...\ell_{n}$ of (valid) links  $\ell_{i} = \link{\alpha_{i}}{\beta_{i}}$ such that:
\begin{enumerate}
\item for any $i\in [1,n-1]$, 
$\left\{\begin{array}{ll}
\beta_{i},\alpha_{i+1}\in \mathcal{C} & \mbox{ implies } \beta_{i} = \alpha_{i+1}\\
\beta_{i}=\silent & \mbox{ iff } \alpha_{i+1}=\silent
\end{array}\right.
$
\item  $\exists i \in [1,n]. \ \ell_i \neq \link{\noact}{\noact}$.
\end{enumerate}

Virtual links represent missing elements of a chain. 
The equivalence $\blackstretcheq$ models expansion/contraction of virtual links to adjust the length of a link chain.

\begin{definition}[Equivalence $\blackstretcheq$]\label{def:black}
We let $\blackstretcheq$ be the least equivalence relation 
over link chains closed under the axioms (whenever both sides are well defined):
\[
\begin{array}{rclcrcl}
v\link{\noact}{\noact} & \blackstretcheq &  v & \qquad &
v_1 \startchain{\noact}\chainedlink{\noact}{\noact}\chainend{\noact}v_2 & \blackstretcheq & v_1 \link{\noact}{\noact} v_2\\
\link{\noact}{\noact}v & \blackstretcheq & v & &
v_1 \startchain{\alpha}\chainedlink{a}{a}\chainend{\beta} v_2
& \blackstretcheq & 
v_1 \startchain{\alpha}\chainedlink{a}{\noact}\chainedlink{\noact}{a}\chainend{\beta}v_2 
\end{array}
\]
\end{definition}

Two link chains of equal length can be merged whenever each position occupied by a solid link in one chain is occupied by a virtual link in the other chain and solid links in adjacent positions match. Positions occupied by virtual links in both chains remain virtual. Merging is denoted by $\merges{v_1}{v_2}$.
%
For example, given $v_1 = \startchain{a}\chainedlink{b}{\noact}\chainedlink{\noact}{\noact}\chainend{\noact}$ and $v_2 = \startchain{\noact}\chainedlink{\noact}{b}\chainedlink{c}{\noact}\chainend{\noact}$ we have $\merges{v_1}{v_2} = \startchain{a}\chainedlink{b}{b}\chainedlink{c}{\noact}\chainend{\noact}$.

Some names in a link chain can be restricted as non observable and transformed into silent actions $\silent$. This is possible only if they are matched by some adjacent link. Restriction is denoted by $\restrict{a}v$.
%
For example, given $v = \startchain{a}\chainedlink{b}{b}\chainedlink{c}{\noact}\chainend{\noact}$ we have 
$\restrict{b}v = \startchain{a}\chainedlink{\silent}{\silent}\chainedlink{c}{\noact}\chainend{\noact}$.

\paragraph{Syntax.}
%\begin{definition}
The c\CNA\ processes, denoted as $\cal{P}$, are generated by the following grammar:
\[
\begin{array}{rcl}
P,Q & ::= &
%\nil \prodsep
%X \prodsep
%s.P  \prodsep
\sum_{i\in I} \upsilon_i.P_i \prodsep
P|Q \prodsep
\restrict{a} P \prodsep
P[\phi] \prodsep
A
\end{array}
\]
\noindent
where $\upsilon_i$ is a  link chain,
%(i.e.\  $\upsilon_i =
%\startchain{\alpha}\chainedlink{\beta}{\noacts}\chainedlink{\noacts}{\alpha'}\chainend{\beta'}$
%with $\alpha,\alpha',\beta,\beta'\neq \ \noacts$),
$\phi$ is a channel renaming function,
and $A$ is a process identifier for which we assume a definition $A
\defeq P$ is available in a given set $\Delta$ of (possibly
recursive) process definitions. We let $\nil$, the inactive process,
denote the empty summation.
%\end{definition}

The syntax of c\CNA\  extends that of \CNA~\cite{BBB17} by allowing to use link chains as prefixes instead of links.
For the rest it features nondeterministic choice, parallel composition, restriction, relabelling and possibly recursive definitions. Here we do not consider name mobility, which is present instead in the {\tt link}-calculus.
\paragraph{Semantics.}

The operational semantics of c\CNA\ is defined in the SOS style by the inference rules in Fig.\ref{fig:cnasos}. The rules are reminiscent of those for Milner's CCS and they essentially coincide with those of \CNA\ in~\cite{BBB17}. The only difference is due to the presence of prefixes that are link chains.
Briefly: rule (\textit{Sum}) selects one alternative and puts as label a possible contraction/expansion of the link chain in the selected prefix; rule (\textit{Ide}) selects one transition of the defining process for a constant; rule (\textit{Res}) restricts some names in the label (it cannot be applied when $\restrict{a}v$ is not defined); rules (\textit{Lpar}) and (\textit{Rpar}) account for interleaving in parallel composition; rule (\textit{Com}) 
synchronises interactions (it cannot be applied when $\merges{v}{v'}$ is not defined).

Analogously to \CNA, the operational semantics of c\CNA\ satisfies the so called Accordion Lemma: whenever $P \xrightarrow{v} P'$ and $v'\blackstretcheq v$ then $P \xrightarrow{v'} P'$.

%\begin{figure}[t]
%\begin{center} 
%\begin{prooftree} 
%\AxiomC{$\upsilon' \blackstretcheq \upsilon $} %\mbox{($\ell$ {\em only} solid link in $s$)}
%\RightLabel{\scriptsize(\textit{Act})}
%\UnaryInfC{$\upsilon.P \xrightarrow{\upsilon'} P$} 
%% \DisplayProof
%\end{prooftree}
%\end{center}
%\caption{SOS semantic rules of the c\CNA. Only rule $(Act)$ is
%changed. The others are the same as in Figure~\ref{fig:cnasos}, in the Appendix.}
%\label{fig:cnasos}
%\end{figure}

\begin{figure}[t]
\begin{center} 
\begin{prooftree} 
\AxiomC{$\upsilon \blackstretcheq \upsilon_j\quad j\in I$} %\mbox{($\ell$ {\em only} solid link in $s$)}
\RightLabel{\scriptsize(\textit{Sum})}
\UnaryInfC{$\sum_{i\in I} \upsilon_i.P_i \xrightarrow{\upsilon} P_j$} 
\DisplayProof
\
\AxiomC{$P \xrightarrow{\upsilon} P'$}
 \AxiomC{$(A \defeq P)\in\Delta$}
\RightLabel{\scriptsize(\textit{Ide})}
\BinaryInfC{$A \xrightarrow{\upsilon} P'$} 
%\AxiomC{$P \xrightarrow{v} P'$} 
%\RightLabel{\scriptsize(\textit{Lsum})}
%\UnaryInfC{$P+Q \xrightarrow{v} P'$} 
%\DisplayProof
%\
%\AxiomC{$Q \xrightarrow{s} Q'$} 
%\RightLabel{\scriptsize(Rsum)}
%\UnaryInfC{$P+Q \xrightarrow{s} Q'$} 
\end{prooftree} 
\end{center}

\begin{center} 
\begin{prooftree} 
\AxiomC{$P \xrightarrow{\upsilon} P'$} 
\RightLabel{\scriptsize(\textit{Res})}
\UnaryInfC{$\restrict{a}P \xrightarrow{\restrict{a}\upsilon} \restrict{a}P'$} 
\DisplayProof
%\,
%\AxiomC{$P \xrightarrow{v} P'$} 
%\RightLabel{\scriptsize(\textit{Rel})}
%\UnaryInfC{$P[\phi] \xrightarrow{v[\phi]} P'[\phi]$} 
%\DisplayProof
\AxiomC{$P \xrightarrow{v} P'$} 
\RightLabel{\scriptsize(\textit{Lpar})}
\UnaryInfC{$P|Q \xrightarrow{\upsilon} P'|Q$} 
\DisplayProof
%\,
%\AxiomC{$Q \xrightarrow{s} Q'$} 
%\RightLabel{\scriptsize(Rpar)}
%\UnaryInfC{$P|Q \xrightarrow{s} P|Q'$} 
%\end{prooftree} 
%\end{center}
%\begin{center} 
%\begin{prooftree} 
\AxiomC{$P \xrightarrow{\upsilon'} P'$} 
\AxiomC{$Q \xrightarrow{\upsilon} Q'$} 
\RightLabel{\scriptsize(\textit{Com})}
\BinaryInfC{$P|Q \xrightarrow{\merges{\upsilon}{\upsilon'}} P'|Q'$} 
%\DisplayProof
\end{prooftree} 
\end{center}
\caption{SOS semantics of the c\CNA\  (rules (\textit{Rel}) and (\textit{Rpar}) omitted).}
\label{fig:cnasos}
\end{figure}

\subsection{Notation for link chains}

Hereafter we make use of some new notations for link chains that will facilitate  
the presentation of our translation.
%First, we define the juxtaposition of link or link chains as
\begin{definition}[Replication]
Let $\upsilon$ be a link chain. Its $n$ times replication $\upsilon^n$ is defined recursively by letting
$\upsilon^0 = \epsilon$ (i.e. the empty chain) and $\upsilon^n = \upsilon^{n-1}\upsilon$,
%\[
%\begin{array}{c@{\qquad}c}
%%\ell ^1 = \ell & \ell^n = \ell^{n-1}\ell\\
%\upsilon^1 = \upsilon& \upsilon^n = \upsilon^{n-1}\upsilon
%\end{array}
%\]
with the hypothesis that all the links in the resulting link chains match.
%, and that $\upsilon^0 = \epsilon$ is the empty chain.
\end{definition}

For example,  the expression $(\startchain{a}\chainedlink{b}{\noact}\chainend{\noact})^3$  denotes the chain $\startchain{a}\chainedlink{b}{\noact}\chainedlink{\noact}{a}\chainedlink{b}{\noact}\chainedlink{\noact}{a}\chainedlink{b}{\noact}\chainend{\noact}$.

 \noindent
We introduce the \emph{half link} that will be used in conjunction with the \emph{open block of chain} to form regular link chains.

\begin{definition}[Half links]
Let $a$ be a channel name, we define the \emph{half left link}:  $\link{a}{}$,   and the  \emph{half right link}: $\link{}{a}$.
\end{definition}

\begin{definition}[Open block]
Let $R$ be a set of names. We 
define an \emph{open block} as $\blockchain{a_i}{a_o}{a \in R}$\ , where $a_i$ and $a_o$ are annotated version of the name $a$, as
\[
\begin{array}{l@{\quad}l@{\quad}l}
\mbox{set}& \mbox{block of chain}& \mbox{result}\\
R= \emptyset & \blockchain{a_i}{a_o}{a \in R}   &\epsilon\\
R = \{ b\}    &\blockchain{a_i}{a_o}{a \in R}  & {}_{b_i}^{\noact}\chainedlink{\noact}{b_o}\\
R = \{b\} \cup R' & \blockchain{a_i}{a_o}{a \in R} & {}_{b_i}^{\noact}\chainedlink{\noact}{b_o}\link{}{}\blockchain{a_i}{a_o}{a \in R'}
\end{array}
\]
\end{definition}
%Please note that, channel names of juxtaposed open blocks of chain are not required to match.

%\noindent
We then combine half links and  open blocks to form regular link chains.
%
%  $\halfl{a}\blockchain{c_i}{c_o}{c \in R}\halfr{b}$ to form a regular link chain; or we can put together half links and open blocks of chain:  $\halfl{a}\blockchain{c_i}{c_o}{c \in R}\blockchain{c_i}{c_o}{c \in R'}\dots \blockchain{c_i}{c_o}{c \in R^{''}}\halfr{b}$ to still form a regular  link chain.
% Again,  all the channel names appearing in the half links or in the open blocks of chain are not required to match.
 For example, for $R = \{a,b\}$ the expression $ \blockchain{c_i}{c_o}{c \in R} $\;
denotes the block of chains ${}_{a_i}^{\noact}\chainedlink{\noact}{a_o} \backslash{}_{b_i}^{\noact}\chainedlink{\noact}{b_o}$; and the expression $\link{r_1}{} \blockchain{c_i}{c_o}{c \in R}\;\link{}{r_2} $ denotes the chain $ \startchain{r_1}\chainedlink{a_i}{\noact}\chainedlink{\noact}{a_o}\chainedlink{b_i}{\noact}\chainedlink{\noact}{b_o}\chainend{r_2}$.

### 2020-03-14 05-translation.tex
Add files via upload

% !TEX root = ./main_RS2L_LNCS.tex

\section{From Reaction Systems to c\CNA}
\label{sec:trans}
Here we present a translation from Reaction Systems to c\CNA.
The idea is to define separated processes for representing the behaviour of each entity, each reaction, 
and for the provisioning of each entity by the context.

\paragraph{Processes for entities.}
Given an entity $s \in S$, we exploit four different names for the interactions over $s$:
names $s_i$, $s_o$ are used to test the presence of $s$ in the system; names $\hat{s}_i$, $\hat{s}_o$ are used to test the provisioning of $s$ from the context;
names $\tilde{s}_i$, $\tilde{s}_o$  are used to test the production of $s$ by some reaction; names $\overline{s}_i$, $\overline{s}_o$ are used to test the absence of $s$ from the context; and names $\underline{s}_i$, $\underline{s}_o$ are used to test the absence of $s$ in the system.
We let $P_s$ be the process implementing the  presence of $s$ in the system, and $\overline{P_s}$ be the one for its absence.
They can be seen as instances of the same template, which is given below.

\[
\begin{array}{lcl@{\hskip 1cm}lcl}
 P_s &\defeq & P(s,\tilde{s},\hat{s},\underline{s}) 
 &
 \overline{P_s}& \defeq & P(\overline{s},\tilde{s},\hat{s},\underline{s})
\end{array}
\]
%{\color{red}
%\[
%\begin{array}{rcl}
% P(s,\tilde{s},\hat{s},\underline{s}) & \defeq &\sum_{h_j,k_j\in\{0,1\}}
%  (\startchain{s_i}\chainedlink{s_o}{\noact}\chainend{\noact})^{h_j} \ (\startchain{\tilde{s}_i}\chainedlink{\tilde{s}_o}{\noact}\chainend{\noact})^{k_j}\ \link{\hat{s}_i}{\hat{s}_o}.P_s\\
%&&\ + \\
%&&
%\sum_{\scriptsize \begin{array}{l}h_j,k_j, \\ h'_j\in\{0,1\}\end{array}}
%  (\startchain{s_i}\chainedlink{s_o}{\noact}\chainend{\noact})^{h_j} \ (\startchain{\tilde{s}_i}\chainedlink{\tilde{s}_o}{\noact}\chainend{\noact})^{k_j} \  \startchain{\tilde{s}_i}\chainedlink{\tilde{s}_o}{\noact}\chainend{\noact} \    (\startchain{s_i}\chainedlink{s_o}{\noact}\chainend{\noact})^{h'_j} \ \link{\underline{s}_i}{\underline{s}_o}.P_s \\
%
%%\sum_{R_1+R_2\geq 0}(\sum_{h,k \in[0,1]} 
%%  (\startchain{s_i}\chainedlink{s_o}{\noact}\chainend{\noact})^h \ (\startchain{\tilde{s}_i}\chainedlink{\tilde{s}_o}{\noact}\chainend{\noact})^k)^{R_1} \startchain{\tilde{s}_i}\chainedlink{\tilde{s}_o}{\noact}\chainend{\noact} (\sum_{h ,k\in[0,1]}
%%  (\startchain{s_i}\chainedlink{s_o}{\noact}\chainend{\noact})^h \ (\startchain{\tilde{s}_i}\chainedlink{\tilde{s}_o}{\noact}\chainend{\noact})^k)^{R_2} \link{\underline{s}_i}{\underline{s}_o}.P_s\\
%&&\ +\\
%&&\sum_{j\geq 0}(
%  \startchain{s_i}\chainedlink{s_o}{\noact}\chainend{\noact})^j\ \ \link{\underline{s}_i}{\underline{s}_o}.\overline{P_s}
%\end{array}
%\]
%}
\[
\begin{array}{rcl}
 P(s,\tilde{s},\hat{s},\underline{s}) & \defeq &\sum_{h,k\geq 0}  
  (\startchain{s_i}\chainedlink{s_o}{\noact}\chainend{\noact})^h \ \startchain{\hat{s}_i}\chainedlink{\hat{s}_o}{\noact}\chainend{\noact} (\startchain{\tilde{s}_i}\chainedlink{\tilde{s}_o}{\noact}\chainend{\noact})^k.P_s\\
&&\ + \\
&&
\sum_{h\geq 0,k\geq 1}
 (\startchain{s_i}\chainedlink{s_o}{\noact}\chainend{\noact})^h \ \startchain{\underline{s}_i}\chainedlink{\underline{s}_o}{\noact}\chainend{\noact}(\startchain{\tilde{s}_i}\chainedlink{\tilde{s}_o}{\noact}\chainend{\noact})^k.P_s\\
&&\ +\\
&&\sum_{h\geq 0}(\startchain{s_i}\chainedlink{s_o}{\noact}\chainend{\noact})^h \ \link{\underline{s}_i}{\underline{s}_o}.\overline{P_s}
\end{array}
\]
The first line of $ P(s,\tilde{s},\hat{s},\underline{s})$ accounts for the case where $s$ is tested for presence by $h$ reactions and produced by $k$ reactions, while being provided by the context ($\link{\hat{s}_i}{\hat{s}_o}$).
Thus, $s$ will be present at the next step (the continuation is $P_s$). Here $h$ and $k$ are not known a priori and therefore any
combination is possible. By knowing the number of reactions that test $s$, we can bound the maximum 
values of $h$ and $k$.
The second line accounts for the analogous case where $s$ is not provided by the context ($\link{\underline{s}_i}{\underline{s}_o}$). 
The condition $k\geq 1$ guarantees that $s$ will remain present (the continuation is $P_s$).
%Here taking
%$h = \sum_{j\geq0} h_j +h'_j$, $k = (\sum_{j \geq 0} k_j )  + 1$ guarantees that $s$ will be present in the next step. Please note that 
The third line accounts for the case where $s$ is tested for presence, but it is neither produced nor provided by  the context. Therefore, in the next step $s$ will be absent in the system (the continuation is $\overline{P_s}$).
Note that in the case of $\overline{P_s}$ the test for presence of $s$ in the system is just replaced by the test for its absence.
%Process $P_s$ (and $\overline{P_s}$) offers link chain prefixes for all the possible
% permutations of the solid links; nevertheless if the reactions, and their enumeration, is known
% the code of $P_s$ can be drastically simplified.

\paragraph{Processes for reactions.}
We assume that each reaction $a$ is assigned a progressive number $j$.
The process for reaction $aj=(R_j,I_j,P_j)$ must assert either the possibility to apply the reaction or its impossibility.
The first case happens when all its reactants are present (the link $\link{s_i}{s_o}$ is requested for any $s\in R_j$) and all its inhibitors are absent (the link $\link{\overline{e}_i}{\overline{e}_o}$ is requested for any $e\in I_j$), then the product set is released (the link $\link{\tilde{c}_i}{\tilde{c}_o}$ is requested for any $c\in P_j$).
The next case can happen for two reasons: one of the reactants is absent (the link $\link{\overline{s}_i}{\overline{s}_o}$ is requested for some $s\in R_j$) or one of the inhibitors is present (the link $\link{e_i}{e_o}$ is requested for some $e\in I_j$).
The process is recursive so that reactions can be applied at any step.
\[
\begin{array}{lcl@{\quad}l}
P_{aj}& \defeq & 
 ^{r_j} \backslash  \blockchain{s_i}{s_o}{s \in R_j} \, \link{}{} \blockchain{\overline{e}_i}{\overline{e}_o}{e \in I_j} \, \chainedlink{r_{j+1}}{\noact}\chainedlink{\noact}{p_j}\link{}{} \blockchain{\tilde{c}_i}{\tilde{c}_o}{c \in P_j}\, \backslash_{p_{j+1}}.P_{aj} & \mbox{\small  \{$aj$ is applicable\}}\\
&& + \ \\
&& \sum_{s\in R_j}\startchain{r_j}\chainedlink{\overline{s}_i}{\noact}\chainedlink{\noact}{\overline{s}_o} \chainedlink{r_{j+1}}{\noact}\chainedlink{\noact}{p_{j}}\chainend{p_{j+1}}.P_{aj}  & \mbox{\small  \{$aj$ is not applicable\}}\\
&&+\\
&& \sum_{e\in I_j}\startchain{r_j}\chainedlink{e_i}{\noact} \chainedlink{\noact}{e_o} \chainedlink{r_{i+1}}{\noact}\chainedlink{\noact}{p_{j}}\chainend{p_{j+1}}.P_{aj}  & \mbox{\small \{$aj$ is not applicable\}}\\
\end{array}
\]

\noindent
We exploit names $r_j,p_j$ to join the chains provided by the application of all the reactions. 
Channels $r_j$ and $r_{j+1}$ enclose the enabling/disabling condition of reaction $aj$.
Channels $p_i$ and $p_{j+1}$ enclose the links related to the entities produced by $aj$.
We will see that all the link chain labels of transitions follow the same schema: first we find all the reactions limited to the reactants and inhibitors (chained using $r_j$ channels), then all the supplies by the contexts (chained using $cxt_j$ channels, to be introduced next), and finally the products for all the reactions (chained using $p_j$ channels). 
In the following there is an example explaining this schema.
%\[
%\begin{array}{lcl@{\qquad}l}
%P_{aj}& \defeq & 
% ^{r_j} \backslash  \blockchain{s_i}{s_o}{s \in R_j} \, \link{}{} \blockchain{\overline{e}_i}{\overline{e}_o}{e \in I_j} \, \link{}{}
% \blockchain{\tilde{c}_i}{\tilde{c}_o}{c \in P_j}\, \backslash_{r_{j+1}}.P_{aj} & \mbox{\small  \{$aj$ is applicable\}}\\
%&& + \ \\
%&& \sum_{s\in R_j}\startchain{r_j}\chainedlink{\overline{s}_i}{\noact}\chainedlink{\noact}{\overline{s}_o} \chainend{r_{j+1}}.P_{aj}  & \mbox{\small  \{$aj$ is not applicable\}}\\
%&&+\\
%&& \sum_{e\in I_j}\startchain{r_j}\chainedlink{e_i}{\noact} \chainedlink{\noact}{e_o} \chainend{r_{i+1}}.P_{aj}  & \mbox{\small \{$aj$ is not applicable\}}\\
%\end{array}
%\]

\paragraph{Processes for contexts.}
For each entity $s \in S$, we introduce another process $Cxt_s$,  participating in each transition and saying if, the entity $s$ is provided by the context or not.
As done for the reactions, we assume that entities are enumerated and use the names $cxt_j$ to concatenate
the chains formed by the application of all the contexts. 
For each entity $s$ with number $j$, at step $n >0$  there are two possible behaviours: 
\[
\begin{array}{lcl}
Cxt^n &\defeq &
%\left \{
\begin{array}{ll}
\sum _{j}^{cxt} \backslash  \blockchain{s_i}{s_o}{s \in Ij_n} \, \link{}{} \blockchain{\overline{e}_i}{\overline{e}_o}{e \in Oj_n} \,  \backslash_{p_1}. Cxt^{n+1}\\
%+ \\
% ^{cxt} \backslash  \blockchain{s_i}{s_o}{s \in I2_n} \, \link{}{} \blockchain{\overline{e}_i}{\overline{e}_o}{e \in Oj_n} \, \backslash_{p_1}
% \chainedlink{\hat{s}_i}{\noact}\chainedlink{\noact}{\hat{s}_o}\chainend{p_1}. Cxt_{s}^{n+1} & \ \mbox{if the context provides $s$  at the $n$-th step}\\[5pt] 
% \startchain{cxt_j}\chainedlink{\underline{s}_i}{\noact}\chainedlink{\noact}{\underline{s}_o}\chainend{cxt_{j+1}}.Cxt_s^{n+1} & \ \mbox{otherwise}
%\end{array}\right.\\
%Cxt_s & \defeq & Cxt_s^1\\
\end{array}
%\right.\\
\end{array}
\]

%\[
%\begin{array}{lcl}
%Cxt_{s}^n &\defeq &
%\left \{\begin{array}{ll}
% \startchain{cxt_j}\chainedlink{\hat{s}_i}{\noact}\chainedlink{\noact}{\hat{s}_o}\chainend{cxt_{j+1}}. Cxt_{s}^{n+1} & \ \mbox{if the context provides $s$  at the $n$-th step}\\[5pt] 
% \startchain{cxt_j}\chainedlink{\underline{s}_i}{\noact}\chainedlink{\noact}{\underline{s}_o}\chainend{cxt_{j+1}}.Cxt_s^{n+1} & \ \mbox{otherwise}
%\end{array}\right.\\
%Cxt_s & \defeq & Cxt_s^1\\
%\end{array}
%\]

We only consider $Cxt^n$ with $n >0$, as the entities  that are present at step zero  are considered to be present in the initial system (process $P_s$ instead of $\overline{P_s}$).

\begin{definition}[Translation]
\label{def:trans}
Let ${\cal A} = (S, A)$ be a RS, and let $\pi=(\gamma,\delta)$ be an extended interactive process in ${\cal A}$, with  $\gamma=\{C_i\}_{i\in\mathbb{N}}$. 
We define its c\CNA~translation $\llbracket{\cal A},\gamma \rrbracket$ as follows: 
\begin{center}$\llbracket{\cal A},\gamma \rrbracket = \restrict{reacts,ctxs,ents,prods}(\Pi_{s\in C_0}P_{s} | \Pi_{s \notin C_0}\overline{P_s} | \Pi_{a\in A} P_a| \Pi_{s\in S} Cxt_s)$,
\end{center} 
with  $reacts$ be the set of reaction names $r_j $, $cxts$ the set of context names $cxt_j$,  $ents$ the set of decorated entity names  $\{s_i,s_o,\hat{s}_i,\hat{s}_o,\tilde{s}_i,\tilde{s}_o,\overline{s}_i,\overline{s}_o,\underline{s}_i,\underline{s}_o| s \in S\}$, and  $prods$ be the set of names $p_j$ associated to each reaction. 
In the following, we set $names = reacts\, \cup\, ctxs\, \cup\, ents\, \cup \, prods$.
%For notational convenience, we fix that $r_1 = \silent$,  $r_{u+1} = cxt_1$ for $u$ the number of reacts, and $cxt_{w+1} = \silent$ for $w$ the number of entities.
For notational convenience, we fix that $r_1 = \silent$,  $r_{u+1} = cxt_1$ for $u$ the number of reacts, and $cxt_{w+1} = p_1$ $p_{u+1}=\silent$ for $w$ the number of entities.
\end{definition}

%can be expressed depending on the actual behaviour of the system we want
%simulate: it could be  recursive, mutually recursive, etc..
%In the general case we can use a counter, $N$, that allows us to specify the behaviour of the context for the first $N$ steps of the system, i.e.  $CXT_s \defeq CXT_s(N)$ and the counter at each step decrements its values, until it arrive to a stable behaviour.

It is important to observe that,   for each transition, our c\CNA \ encoding requires  all the processes $P_a$, with $a \in A$, and  $Cxt_s$ and  $P_s$, with $s \in S$,  be interacting in that transition.
This is due to the fact that  all the channels $r_j$, $p_j$, $cxt_h$, and $s_{hi}$, and $s_{ho}$ are  restricted.
Each reaction defines a pattern to be satisfied, i.e. each reaction inserts as many virtual links  as  the number of reactants, inhibitors, and products, as required by the corresponding reaction.

 \begin{restatable}[]{lemma}{lemmastruct}
 \label{lem:struct}
%\begin{lemma}
 %\label{lem:struct}
  Let ${\cal A} = (S, A)$ be a RS and let $\pi=(\gamma,\delta)$ be an extended  interactive process in ${\cal A}$. Let $P=\llbracket{\cal A},\gamma\rrbracket$ its c\CNA~translation. If $exists$ $P'$ such that $t= (P \xrightarrow{\restrict{names}\upsilon} P')$ is a transition of $P$, then
  \begin{enumerate}
  \item for each reaction  $a_j \in A$, the corresponding channels $r_j$ and $p_j$ appear in  $\upsilon$; for each entity $s_h \in S$ (where $h$ is the identifying number of $s$), the corresponding channel $s_h$ (suitably decorated), and the corresponding channel $cxt_h$ appear in $\upsilon$; 
  %processes $P_s$ (or $\overline{P_s}$) 
%  for each  $Cxt_s$ are
 %involved in the transition $t$;
  \item for each reaction $a \in A$ and each entity $s \in S$, each virtual link offered by processes $P_a$ and $Cxt_s$ is  overlapped by exactly one solid link offered by  processes representing entities.
  %each process $P_s $ (or its alternative version $\overline{P}_s$), with $s \in S$, participates at most  with one solid link for each process $P_a$, and with exactly one solid link for each process $Cxt_s$.%, if $s$ is a \emph{required element} of $A_j$ in $t$.

  \end{enumerate}
  \end{restatable}

 \begin{example}\label{ex:backbone}
 Let ${\cal A}$ be a RS whose specification contains two entities, $s1$ and $s2$, and, among the others, the reaction $a=(s1,\ , s1)$ that guarantees the presence of the  $s1$ in the system. Then, we assume an  extended interactive process $\pi=(\gamma,\delta)$ where the context $\gamma$ provides  $s1$ and $s2$.
 %, for the first $n$ steps.
 Our translation includes the processes:\\
 \[
 \begin{array}{lcllcl}
 P_a &\defeq& \startchain{\silent}\chainedlink{s1_i}{\noact}\chainedlink{\noact}{s1_o}\chainedlink{r_2}{\noact} \chainedlink{\noact}{p_1}\chainedlink{\tilde{s1}_i}{\noact}\chainedlink{\noact}{\tilde{s1}_o}\chainend{p_2}.P_a + \dots; \\
 P_{s1} &\defeq& \startchain{s1_i}\chainedlink{s1_o}{\noact}\chainedlink{\noact}{\hat{s1}_i}\chainedlink{\hat{s1}_o}{\noact}\chainedlink{\noact}{\tilde{s1}_i}\chainedlink{\tilde{s1}_o}.P_{s1} + \dots; &
  P_{s2} &\defeq& \link{\hat{s2}_i}{\hat{s2}_o}.P_{s2} + \dots;\\
 Cxt_{s1}& \defeq & \startchain{cxt_1}\chainedlink{\hat{s1}_i}{\noact}\chainedlink{\noact}{\hat{s1}_o}\chainend{cxt_2}.Cxt_{s1} &
  Cxt _{s2} &\defeq& \startchain{cxt_2}\chainedlink{\hat{s2}_i}{\noact}\chainedlink{\noact}{\hat{s2}_o}\chainend{p_1}. Cxt_{s2}
\end{array}
\]  \\
   Now, we assume that $s1$ is in the initial state of ${\cal A}$, and in Figure~\ref{fig:backbone} we show the  structure of a link chain label related to the execution of a transition of the c\CNA \ system: $\restrict{names} (P_{s1}|P_{s2}| P_a|\dots|Cxt_{s1}|Cxt_{s2})$. The yellow blocks are referred to the processes encoding the reactions ($P_a$, in our case) and the contexts ($Cxt_{s1}$ and $Cxt_{s2}$). As the figure puts in evidence, these two kinds of processes determine the structure of the link chain, from end to end, i.e. from the left $\silent$ to the right one. We could say that these processes form the \emph{backbone} of the interaction.
   In contrast, the processes encoding the entities ($P_{s1}$, and $P_{s2}$, in our case) provides the
    solid links to overlap the virtual links   of the backbone.

  \begin{figure}[t]
  \begin{tikzpicture}
  \draw (0,1) -- (3,1) ;
  \draw (4,1) -- (5,1) ;
  \draw (0.5,0) -- (3.5,0);
  \draw (4.5,0) -- (5.5,0) ;
  \draw (0,1)--(0,0.5) -- (0.5,0.5) -- (0.5,0) -- (1.5,0) ;
  \draw (0.25,0.75)  node[minimum size=0.495cm,fill=yellow] {${ \silent}$};
  \draw (0.75,0.75)  node[minimum size=0.5cm,fill=yellow] {$\ $};
  \draw (0.76,0.25)  node[minimum size=0.5cm,fill=yellow] {$\ $};
  \draw (1.21,0.25)  node[minimum size=0.5cm,fill=yellow] {${\scriptscriptstyle s1_i}$};

  \draw (2,1) -- (2,0.5) -- (2.5,0.5) -- (2.5,0) -- (1.5,0)  -- (1.5,0.5) -- (1,0.5) -- (1,1);

 \draw (1.25,0.75) node{$\noact$};%node[minimum size=0.5cm,draw,draw=white] {$\noact$};
 \draw (2.25,0.25) node{$\noact$};
 \draw (2.31,0.75)  node[minimum size=0.47cm,draw,fill=yellow, draw=yellow] {${\scriptscriptstyle s1_o}$};
  \draw (2.75,0.75)  node[minimum size=0.48cm,draw,fill=yellow,yellow] {$\ $};
  \draw (2.75,0.26)  node[minimum size=0.49cm,draw,fill=yellow,yellow] {$\ $};

  \draw (3.245,0.25)  node[minimum size=0.475cm,draw,fill=yellow,draw=yellow] {${\scriptscriptstyle   r_2}$};

  \draw (3.75,0.5) node{$\dots$};

  \draw(3.5,0.5) -- (3.5,0);
%%     \draw (3.25,0.75)  node[minimum size=0.5cm,draw,draw=white] {$\noact$};
%%  \draw (3.75,0.75)  node[minimum size=0.5cm,draw,white] {$\ $};
%%  \draw (3.75,0.25)  node[minimum size=0.5cm,draw,white] {$\ $};
%%  \draw (4.25,0.25)  node[minimum size=0.5cm,draw,draw=white] {$\noact$};
\draw  (4,1) --  (4,0.5) -- (4.5,0.5) -- (4.5,0)   (3.5,0.5) -- (3,0.5) -- (3,1);
% \draw (3.25,0.75) node{$\noact$};%node[minimum size=0.5cm,draw,draw=white] {$\noact$};
%  \draw (4.25,0.25) node{$\noact$};
\draw (4.75,0.745)  node[minimum size=0.485cm,draw,fill=yellow,yellow] {$\ $};
  \draw (4.38,0.75) node[minimum size=0.47cm,draw,fill=yellow, draw=yellow] 
  {${\scriptscriptstyle  cxt_1}$};

  \draw (4.75,0.25)  node[minimum size=0.48cm,draw,fill=yellow,yellow] {$\ $};
  \draw (5.2,0.255)  node[minimum size=0.485cm,draw,fill=yellow,draw=yellow] {${\scriptscriptstyle  \hat{s1}_i}$};
  \draw(5,1) -- (5,0.5) -- (5.5,0.5) -- (5.5,0);

   \draw (5.75,0.5) node{$\dots$};

    \draw (6,1) -- (11.5,1) ;
    \draw (7,1) -- (7,0.5) -- (7.5,0.5) -- (7.5,0);
    \draw (6.5,0) -- (11.5,0);
    \draw   (6,1)-- (6,0.5) -- (6.5,0.5) -- (6.5,0.5) -- (6.5,0) ;
     \draw (6.75,0.75)  node[minimum size=0.47cm,fill=yellow, draw=yellow] {$\ $};
 \draw (6.36,0.75) node[minimum size=0.475cm,draw,fill=yellow,draw=yellow] 
 {${\scriptscriptstyle cxt_2}$};%node[minimum size=0.5cm,draw,draw=white] {$\noact$};
   \draw (6.75,0.26)  node[minimum size=0.485cm,fill=yellow, draw=yellow] {$\ $};

 \draw (7.20,0.25) node[minimum size=0.475cm,fill=yellow, draw=yellow] {${\scriptscriptstyle \hat{s2}_i}$};

     \draw (7.3,0.75)   node{$\noact$};
  \draw (7.75,0.75) node{$\ $};
  \draw (7.76,0.26) node {$\ $};
  \draw (8.25,0.26)  node {$\noact$ };

  \draw (8.3,0.75)  node[minimum size=0.48cm,draw,fill=yellow, draw=yellow] {${\scriptscriptstyle \hat{s2}_o}$};
  \draw (8,1)--(8,0.5)--(8.5,0.5) -- (8.5,0);
 % \draw(9.5,0.5) --(9.5,0);
  \draw (8.745,0.75)  node[minimum size=0.48cm,draw,fill=yellow,yellow] {$\ $};
  \draw (8.75,0.26)  node[minimum size=0.48cm,draw,fill=yellow,yellow] {$\ $};
 \draw (9.24,0.26)  node[minimum size=0.48cm,draw,fill=yellow,draw=yellow] {${\scriptscriptstyle p_1}$};
 %\draw(9.5,0.5) --(9.5,0);
    % \draw (9,1) -- (10,1) -- (10,0.5) -- (10.5,0.5) --  (10.5,0) -- (9.5,0)  (9.5,0.5) -- (9,0.5) -- (9,1);
 \draw (9.24,0.75) node[minimum size=0.48cm,draw,fill=yellow,draw=yellow]{${\scriptscriptstyle p_1}$};%node[minimum size=0.5cm,draw,draw=white] {$\noact$};
 \draw (10.22,0.26) node [minimum size=0.48cm,draw,fill=yellow,draw=yellow]{${\scriptscriptstyle \tilde{s1}_i}$};
 \draw (9.74,0.745) node [minimum size=0.485cm,draw,fill=yellow,draw=yellow]{$\ $};
 \draw (9.735,0.26) node [minimum size=0.48cm,draw,fill=yellow,draw=yellow]{$\ $};
 \draw (10,1) -- (10,0.5) -- (10.5,0.5)--(10.5,0);
  \draw (10.31,0.75)  node{$\noact$};
  \draw (10.75,0.75)  node {$\ $};
  \draw (10.75,0.26)  node{$\ $};
 \draw (11.25,0.25)  node {$\noact$};
   \draw (11,1) -- (11,0.5) -- (11.5,0.5) -- (11.5,0);% -- (10,0) -- (9,0) -- (9,0.5) -- (8.5,0.5) -- (8.5,1);
    \draw (11,1) -- (11.5,1) -- (12,1) -- (12,0.5) -- (12.5,0.5)--(12.5,0)--(11.5,0);
   %\draw (11.25,0.25)  node[minimum size=0.48cm,draw,fill=yellow,draw=yellow] {$\noact$};
    \draw (11.31,0.75) node[minimum size=0.48cm,draw,fill=yellow,draw=yellow]{${\scriptscriptstyle \tilde{s1}_o}$};
    \draw (11.75,0.26)node[minimum size=0.48cm,draw,fill=yellow,draw=yellow]{$\ $};
      \draw (11.75,0.75)node[minimum size=0.48cm,draw,fill=yellow,draw=yellow]{$\ $};
     \draw (12.2,0.25) node[minimum size=0.478cm,draw,fill=yellow,draw=yellow]{\;$\silent$\;  };
  \end{tikzpicture}
  \caption{The link chain structure arising from reactions and context processes.}
   \label{fig:backbone}
 \end{figure}
 \end{example}

% \begin{figure}[t]
% \begin{tikzpicture}
%   \draw (0,1) -- (5,1) ;
%    \draw (0.5,0) -- (5.5,0) ;
%   \draw (0,1)--(0,0.5) -- (0.5,0.5) -- (0.5,0) -- (1.5,0) ;
%     \draw (0.25,0.75)  node[minimum size=0.495cm,fill=yellow] {${ \silent}$};
% \draw (0.75,0.75)  node[minimum size=0.5cm,fill=yellow] {$\ $};
%  \draw (0.76,0.25)  node[minimum size=0.5cm,fill=yellow] {$\ $};
%  \draw (1.21,0.25)  node[minimum size=0.5cm,fill=yellow] {${\scriptscriptstyle s1_i}$};
%%    \draw (1.25,0.75)  node[minimum size=0.5cm,draw,draw=white] {$\noact$};
%%  \draw (1.75,0.75)  node[minimum size=0.5cm,draw,white] {$\ $};
%%  \draw (1.75,0.25)  node[minimum size=0.5cm,draw,white] {$\ $};
%%  \draw (2.25,0.25)  node[minimum size=0.5cm,draw,draw=white] {$\noact$};
%  \draw (2,1) -- (2,0.5) -- (2.5,0.5) -- (2.5,0) -- (1.5,0)  -- (1.5,0.5) -- (1,0.5) -- (1,1);
%
% \draw (1.25,0.75) node{$\noact$};%node[minimum size=0.5cm,draw,draw=white] {$\noact$};
%  \draw (2.25,0.25) node{$\noact$};
%   \draw (2.31,0.75)  node[minimum size=0.47cm,draw,fill=yellow, draw=yellow] {${\scriptscriptstyle s1_o}$};
%  \draw (2.75,0.75)  node[minimum size=0.48cm,draw,fill=yellow,yellow] {$\ $};
%  \draw (2.75,0.26)  node[minimum size=0.49cm,draw,fill=yellow,yellow] {$\ $};
%
%  \draw (3.20,0.25)  node[minimum size=0.47cm,draw,fill=yellow,draw=yellow] {${\scriptscriptstyle  \tilde{s1}_i}$};
%  
%  
%  \draw(3.5,0.5) -- (3.5,0);
%%     \draw (3.25,0.75)  node[minimum size=0.5cm,draw,draw=white] {$\noact$};
%%  \draw (3.75,0.75)  node[minimum size=0.5cm,draw,white] {$\ $};
%%  \draw (3.75,0.25)  node[minimum size=0.5cm,draw,white] {$\ $};
%%  \draw (4.25,0.25)  node[minimum size=0.5cm,draw,draw=white] {$\noact$};
%\draw (3,1) -- (4,1) -- (4,0.5) -- (4.5,0.5) -- (4.5,0) -- (3.5,0)  (3.5,0.5) -- (3,0.5) -- (3,1);
% \draw (3.25,0.75) node{$\noact$};%node[minimum size=0.5cm,draw,draw=white] {$\noact$};
%  \draw (4.25,0.25) node{$\noact$};
%  \draw (4.31,0.75)  node[minimum size=0.47cm,draw,fill=yellow, draw=yellow] {${ \scriptscriptstyle \tilde{s1}_o}$};
%  \draw (4.75,0.745)  node[minimum size=0.48cm,draw,fill=yellow,yellow] {$\ $};
%  \draw (4.75,0.25)  node[minimum size=0.48cm,draw,fill=yellow,yellow] {$\ $};
%  \draw (5.24,0.255)  node[minimum size=0.485cm,draw,fill=yellow,draw=yellow] {${\scriptscriptstyle  r_2}$};
%  \draw(5,1) -- (5,0.5) -- (5.5,0.5) -- (5.5,0);
%   \draw (6,0.5) node{$\dots$};
%    \draw (6.5,1) -- (11.5,1) ;
%    \draw (7.5,1) -- (7.5,0.5) -- (8,0.5) -- (8,0);
%    \draw (7,0) -- (12,0);
%    \draw   (6.5,1)-- (6.5,0.5) -- (7,0.5) -- (7,0.5) -- (7,0) ;
% \draw (6.77,0.75) node{$\noact$};%node[minimum size=0.5cm,draw,draw=white] {$\noact$};
% \draw (7.77,0.25) node{$\noact$};
%
%     \draw (7.80,0.75)  node[minimum size=0.485cm,draw,fill=yellow, draw=yellow] {${\scriptscriptstyle  \tilde{s1}_o}$};
%  \draw (8.25,0.75)  node[minimum size=0.485cm,draw,fill=yellow,yellow] {$\ $};
%  \draw (8.26,0.26)  node[minimum size=0.495cm,draw,fill=yellow,yellow] {$\ $};
%  \draw (8.75,0.26)  node[minimum size=0.49cm,draw,fill=yellow,draw=yellow] {${\scriptscriptstyle cxt_2}$ };
%
%  \draw (8.75,0.75)  node[minimum size=0.48cm,draw,fill=yellow, draw=yellow] {${\scriptscriptstyle cxt_2}$};
%  \draw (9.25,0.75)  node[minimum size=0.48cm,draw,fill=yellow,yellow] {$\ $};
%  \draw (9.25,0.26)  node[minimum size=0.48cm,draw,fill=yellow,yellow] {$\ $};
% \draw (9.71,0.25)  node[minimum size=0.47cm,draw,fill=yellow,draw=yellow] {${\scriptscriptstyle \tilde{s2}_i}$};
% \draw(10,0.5) --(10,0);
%     \draw (9.5,1) -- (10.5,1) -- (10.5,0.5) -- (11,0.5) --  (11,0) -- (10,0)  (10,0.5) -- (9.5,0.5) -- (9.5,1);
% \draw (9.75,0.75) node{$\noact$};%node[minimum size=0.5cm,draw,draw=white] {$\noact$};
% \draw (10.75,0.25) node{$\noact$};
%  \draw (10.81,0.75)  node[minimum size=0.47cm,draw,fill=yellow, draw=yellow] {${\scriptscriptstyle \tilde{s2}_o}$};
%  \draw (11.25,0.75)  node[minimum size=0.48cm,draw,fill=yellow,yellow] {$\ $};
%  \draw (11.25,0.26)  node[minimum size=0.49cm,draw,fill=yellow,yellow] {$\ $};
% \draw (11.75,0.25)  node[minimum size=0.48cm,draw,fill=yellow,draw=yellow] {$\silent$};
%   \draw (11.5,1) -- (11.5,0.5) -- (12,0.5) -- (12,0);% -- (10,0) -- (9,0) -- (9,0.5) -- (8.5,0.5) -- (8.5,1);
%  \end{tikzpicture}
%  \caption{The link chain structure arising from reactions and context processes.}
%   \label{fig:backbone}
% \end{figure}
% \end{example}

Example~\ref{ex:backbone}   outlines two different roles of the processes defining the translation of an interactive process: those processes encoding the reactions and the context provide the backbone of each transition, whereas the processes encoding the entities provide the resources needed for the communication to take place.

With the next proposition, we analyse the structure of a c\CNA \ 
process encoding of  a reactive process after one step transition.
In the following four statements, for brevity, we let ${\cal A} = (S, A)$ be a RS, and let  $\pi=(\gamma,\delta)$ be an extended interactive process in $A$, with $\gamma=\{C_i\}_{i\in\mathbb{N}}$ and $\delta=\{D_i\}_{i\in\mathbb{N}}$. Moreover, we denote by $\pi^j$ the shift of $\pi$ starting at the $j$-th state sequence; formally we let $\pi^j=(\gamma^j,\delta^j)$ with
$\gamma^j=\{C_i'\}_{i\in\mathbb{N}}$, $\delta^j=\{D_i'\}_{i\in\mathbb{N}}$ with 
  $C_0'=C_j \cup D_j$, and   $C_i'=C_{i+j}$, $D_i'= D_{i+j}$ for any $i\geq 1$.

 \begin{restatable}[Correctness 1]{proposition}{propcorruno}
% \begin{proposition}(correctness 1)\\
 \label{prop:corr1}
% We define $\pi' = (\gamma',\delta')$ with $\gamma' = C_0',\dots,C_{n-1}'$ and $\delta' = D_0',\dots,D_{n-1}'$ as $C_0' = C_1 \cup D_1$,  and
% $C_i' = C_{i+1}$, $D_{i}' = D_{i+1}$ with $ 1 \leq i < n$.
 Let 
 $P=\llbracket{\cal A},\gamma\rrbracket$ with\\
$P  = \restrict{names} (\Pi_{a \in A}P_a| \Pi_{s \in S}Cxt_s | \Pi_{s \in C_0}P_s| \Pi_{s \notin C_0}\overline{P}_s)$.\\
 If there exists  $P'$  such that $P \xrightarrow{v}P'$,  it holds that:\\
  $v=\link{\silent}{\silent}\dots\link{\silent}{\silent}$ and \\
 $P'= \restrict{names} (\Pi_{a \in A}P_a| \Pi_{s \in S}Cxt_s | \Pi_{s \in C_1 \cup D_1}P_s| \Pi_{s \notin C_1 \cup D_1}\overline{P}_s)$.\\
Moreover, given $\pi^1=(\gamma^1,\delta^1)$, we have $P' = \llbracket {\cal A},\gamma^1\rrbracket$.
 %\end{proposition}
\end{restatable}

% \begin{lemma}[correctness 1 {\color{red} Ma serve ??}]
% \label{cor:complet1}
% Let ${\cal A} = (S, A)$ be a rs, and let  $\pi=(\gamma,\delta)$ be an extended interactive process in $A$, with
% $\gamma = C_0,\dots,C_t$, $\delta = D_0,\dots,D_t$, with $n \geq 0$. Let   $P=[|{\cal A},\gamma|]$ be  its c\CNA~translation. We define $\pi' = (\gamma',\delta')$ with $\gamma' = C_0',\dots,C_{n-1}'$ and $\delta' = D_0',\dots,D_{n-1}'$ as $C_0' = C_1 \cup D_1$,  and
% $C_i' = C_{i+1}$, $D_{i}' = D_{i+1}$ with $ 1 \leq i < n$. Then,
% if $\exists$ $P'$, $P''$ such that $P \xrightarrow{\link{\silent}{\silent}} P'$ and $P \xrightarrow{\link{\silent}{\silent}} P''$, then $P' = P''$.
%\end{lemma}

% \begin{proposition}[correctness2]
% \label{prop:correctness1}
% Let $P$ be a c\CNA \ process such that $\exists$ a $rs$ ${\cal A}$ and a $n$-step interactive process $\pi=(\gamma,\delta)$ with $\gamma =C_0,\dots,C_n$ and $\delta = D_0,\dots,D_n$ such that $P= [|{\cal A},\gamma|]$. 
% %We set $\pi'=(\gamma',\delta')$ with $\gamma'=C_0',\dots,C_{n-1}'$,  $\delta'=D_0',\dots,D_{n-1}$,  where $C_0'= C_1 \cup D_1$, $C_i'= C_{i+1}$, $D_i=D_{i+1}'$,  and $1 \leq i < n$.
%Then,   $\exists$ $P'$ and $\pi'=(\gamma',\delta')$ such that $P \xrightarrow{\link{\silent}{\silent}}P'$, $P' =[|{\cal A},\gamma'|]$.
% \end{proposition} 
\noindent
Now, we extend the previous result to a series of transitions.

 \begin{restatable}[Correctness 2]{corollary}{corrcorrdue} 
% \begin{corollary}[correcteness2]
 \label{corr:corr2}
  Let $P = \llbracket{\cal A},\gamma\rrbracket$ and $j\geq 1$.
%  Let $P$ a  c\CNA process such that exists a rs ${\cal A}$ and an extended  interactive process $\pi=(\gamma,\delta)$ with $\gamma =C_0,\dots,C_t$, $\delta = D_0,\dots,D_t$, with $t \geq 0$, such that $P= [|{\cal A},\gamma|]$. 
If there exists  $P''$ such that $P \xrightarrow{\link{\silent}{\silent}\dots \link{\silent}{\silent}}^{j}P''$, then letting  $\pi^j=(\gamma^j,\delta^j)$ we have
$P'' =\llbracket{\cal A},\gamma^j\rrbracket$.
%such that $P=[|{\cal A}|]$, then if $P \rightarrow^* P'$, then $\exists$ ${\cal A'}$ such that ${\cal A}$  evolves in ${\cal A'}$ and $P'=[|{\cal A'}|]$.
 \end{restatable}

\noindent
With the following propositions, we prove that, given a RS ${\cal A} = (S, A)$ and an extended  interactive process $\pi=(\gamma,\delta)$, then the $c\CNA$ process $\llbracket{\cal A},\gamma\rrbracket$ can simulate all the evolutions  of $\pi$.

 \begin{restatable}[Completeness 1]{proposition}{propcompluno} 
% \begin{corollary}[correcteness2]
 \label{prop:compl1}
% \begin{proposition}[completeness1]
 %\label{lem:compl2}
Let   $P=\llbracket{\cal A},\gamma\rrbracket$ and $\pi^1 = (\gamma^1,\delta^1)$.
% be the extended interactive process starting  at the next state sequence. 
Then, 
 % following holds:
 %\begin{enumerate} 
 %\item  
  $P \xrightarrow{\link{\silent}{\silent}\dots \link{\silent}{\silent}} P'=\llbracket{\cal A},\gamma^1\rrbracket$.
 %\item if $\exists$ $P'$, $P''$ such that $P \xrightarrow{\link{\silent}{\silent}} P'$ and $P \xrightarrow{\link{\silent}{\silent}} P''$, then $P' = P''$.
 %\end{enumerate}
% \end{proposition} 
\end{restatable}

\noindent
Now, we extend the previous result to a series of transitions.

\begin{restatable}[Completeness 2]{corollary}{corrcompldue} 
% \begin{corollary}[correcteness2]
 \label{corr:compl2}
% \begin{corollary}[completeness2]
 %\label{prop:completeness}
Let $P=\llbracket{\cal A},\gamma\rrbracket$ and  $\pi^j= (\gamma^j,\delta^j)$.  
Then, $P \xrightarrow{\link{\silent}{\silent}\dots\link{\silent}{\silent}}^j P''=\llbracket{\cal A},\gamma^j\rrbracket$.  
 \end{restatable}

% 
% \begin{proposition}[correctness]
% Let $P$ a c\CNA\  process such that exists a reaction system ${\cal A}$ such that $P=[|{\cal A}|]$, then if $P \rightarrow P'$, then $\exists$ ${\cal A'}$ such that ${\cal A}$  evolves in ${\cal A'}$ in one step and $P'=[|{\cal A'}|]$.
% \end{proposition}

### 2020-03-14 06b-bisimulation.tex
Add files via upload

% !TEX root = ./main_RS2L_LNCS.tex
\section{Bio-simulation}
\label{sec:biosimulation}

The classical notion of bisimulation for process algebra equates two processes
when one process can simulate all the instructions executed by the other one and viceversa.
In its weak formulation, internal instructions, i.e. non visible by external observers,
are abstracted away. 
There are many variants of the bisimulation for process algebras,  for example the
barbed bisimulation~\cite{10.1007/3-540-55719-9_114} only consider the execution of invisible actions, and then equates two processes when the expose the same prefixes; for the mobile ambients~\cite{CardelliG00}, a process algebra equipped with a reduction semantics, a notion of
behavioural equivalence equates two processes when they expose the same ambients\cite{GC03}.

There are some previous works based on bisimulation applied to models for biological systems. Barbuti et al \cite{BMMT08} define a classical setting for bisimulation for two formalisms: the Calculus of Looping Sequences, which is a rewriting system, and the Brane Calculi, which is based on process calculi.
Bisimulation is used to verify properties of the regulation of lactose degradation in
Escherichia coli and the EGF signalling pathway. These calculi allow the authors to model membranes' behaviour.
\cite{CTTV15} presents two quantitative behavioral equivalences over species of a 
chemical reaction network with semantics based on ordinary differential equations.
Bisimulation identifies a partition where each equivalence class represents the exact sum of the concentrations of the species belonging to that class.
Bisimulation also relates species that have identical solutions at all time points when starting from the same initial conditions.
Both the mentioned formalisms \cite{BMMT08,CTTV15} adopt a classical approach to bisimulation. Moreover our formalisms is different, for different applications and purposes.

Albeit the bisimulation is a powerful tool for verifying if the behaviour of two different software 
programs is indistinguishable, in the case of biological systems the classical bisimulation seems to be inappropriate, as it considers too many details.
In fact, in a biological soup, a high number of interactions occur every seconds, and generally, biologists
are only interested to analyse a small subset of them.

For this reason, we propose an alternative notion of bisimulation, that hereafter we call biosimulation,
that allows us  to compare two biological systems by restricting the observation to only the limited events of interest.

The transition labels of our systems record detailed information about all the reactions that have been applied in one transition, about the elements that acted as reagents, as inhibitors or that have been produced, or that have been provided by the context.

In a way, we want to query our transition labels to get only the information we are interested in.
To this goal, we introduce a simple language
%simple assertion language 
that allows us to formulate detailed and partial queries about what happened in a single transition.

First, we need to introduce the $flat$ function that takes a link chain and returns a simple string
by eliminating all the channel matching pairs leaving just one channel per each pair:\\
\[
\begin{array}{lcl@{\hspace{0.5cm}}lcl@{\hspace{0.5cm}}lcl@{\hspace{0.5cm}}lcl}
flat(\epsilon)&\defeq & \epsilon &
flat(\link{\silent}{\beta}) & \defeq & \beta &
flat(\link{\alpha}{\silent})& \defeq & \epsilon &
flat(\link{\alpha}{\beta}) & \defeq &\beta \\
\end{array}\]
\[
\begin{array}{lcl}
flat(\link{\alpha}{\beta}\upsilon)& \defeq & \beta :: flat(\upsilon)\\
\end{array}\\
\]
where  $::$ is the string concatenation operator.

\noindent
Now, we can introduce an assertion language that operates on simple strings
and that combines regular expression operators with the classical $and$ and $or$ logical
operators.

\begin{definition}[Assertion language]
\label{def:assetionl}
Assertions are built from: 
$$
%B ::=  s~\mid~ \tilde{s}~\mid~ \hat{s} ~\mid~ \overline{s} ~\mid~  \underline{s} ~\mid~ F \oplus F  
\begin{array}{lcl}
\zeta' &::= & \epsilon  ~\mid~ \zeta::\zeta'\\ 
re&::=& . ~\mid ~   (\zeta \textrm{-}\zeta)  ~\mid ~ \zeta' ~\mid ~ re::re\\
re' &::= & 
%(re')|(re')  ~\mid~
[\zeta\dots\zeta]~\mid~ [\hat{}  re] ~\mid ~ re* ~\mid ~ re+ \\
F &::=& re' ~\mid ~  F \vee F ~\mid ~  F \wedge F\\
\end{array}
$$
%\noindent where $s \in {\cal C}$, and  $\oplus \in \{\wedge, \vee, \to \}$. 
where $\zeta \in \{s, \tilde{s}, \hat{s}, \overline{s}, \underline{s} \}$ with  $s \in {\cal C}$.
%, and  $\oplus \in \{\wedge, \vee \}$. 
% \{\wedge, \vee, \to \}$. 
\end{definition}

$\zeta'$ is the set of  strings composed by characters in the set ${\cal C}$.
Then, the dot "$.$" stands for any character, $\zeta_1 \textrm{-}\zeta_2$ is a shorthand for the string
starting with $\zeta_1$ and ending with $\zeta_2$ and that in the  middle includes all the characters
following the alphanumeric order.
%The round brackets $(\zeta')$ build a single string  $\zeta'$, i.e. a non-separable sequence of characters.
 The square brackets match one of the characters they contain.
 The symbol $[\hat{} \dots]$, inside the square brackets excludes any of the characters they contain.
 The star $*$ allows the previous character to be replicated zero or more times.
 The plus symbol $+$ allows the previous character to be replicated at least one or more times.
Then, we can combine the patterns build with the above constructors with the logic operators $\vee$ and
$\wedge$.

\begin{definition}[Semantics]\label{sec:semantics}
Let  $\upsilon$ be  a transition label, and $F$ be an assertion.
We define $\upsilon \entails F$ (read as the transition label $v$ satisfies the formula $F$) as: 
\begin{itemize}
 \item   $\upsilon \entails \zeta $ iff $\zeta \in flat(\upsilon)$. 
 \item   $\upsilon \entails F_1 \wedge F_2$ iff   $\upsilon \entails F_1$  $\wedge$  $\upsilon \entails F_2$.
\item   $ \upsilon \entails F_1 \vee F_2$ iff   $\upsilon \entails F_1$  $\vee$  $\upsilon \entails F_2$.
\end{itemize}
If it is not the case that $\upsilon \entails  F $, then we say that $F$ does not hold at $v$ and we write $\upsilon \not\entails F$. 
\end{definition}

%\begin{definition}[Semantics]\label{sec:semantics}
%Let $\pi$ be a trace of transition labels, $v$ be  a transition label, and $F$ be an assertion.
%We inductively define $\pi,v_i \entails F$ (read as the transition label $v$, which is at position $i$ of $\pi$ satisfies the formula $F$) as: 
%\begin{itemize}
% \item   $\pi, v_i \entails \zeta $ if $\zeta \in v_i$. 
% \item   $\pi, v_i \entails F_1 \wedge F_2$ if   $\pi, v_i \entails F_1$  $\wedge$  $\pi, v_i \entails F_2$.
%\item   $\pi, v_i \entails F_1 \vee F_2$ if   $\pi, v_i \entails F_1$  $\vee$  $\pi, v_i \entails F_2$.
%\item   $\pi, v_i \entails F_1 \to F_2$ if   $\pi, v_i \entails F_1$  $\to$  $\pi, v_i \entails F_2$. 
%\end{itemize}
%If it is not the case vbthat $\pi, v_i \entails  F $, then we say that $F$ does not hold at $v_i$ and we write $\pi,v_i \not\entails F$. 
%\end{definition}
Our  language allows us to check if the simulation of a system (i.e. the transition labels, or in other words the trace of a simulation) 
satisfies the properties of the following types:% that can be expressed using our assertion language. Some simple examples  follow:
\begin{enumerate}
\item Has the entity $s$ been used by rule $r$  as reagent ?
\item Has the entity $s$ blocked the application of  rule $r$ ?
\item Has the entity $s$ been produced by some rule ?
\item Has the entity $s$ been produced by rule $p$ ?
\item Has the entity $s$ been provided by the context ? 
\item Has the rule $i$ been applied ? 
%\item Has the entity $s$ been used by some rule as reagent ?
%\item Has the entity $s$ blocked  the application of some rules ?
%\item Has the entity $s$ been produced ?
%\item Has the entity $s$ been provided by the context ?
%\item Has the rule $i$ been applied ?
\end{enumerate}
The corresponding formulas are as follows:\\

$
\begin{array}{cl}
1. & r[  s1 \dots sn ] *s[  s1 \dots sn ]* \\
2. & r [ s \overline{s}]  r' \\
3. & \tilde{s}\\
4. & p[  \tilde{s1} \dots \tilde{sn} ] *\tilde{s}[  \tilde{s1} \dots \tilde{sn} ]* \\
5. & \hat{s}\\
6. &DA DECIDERE come
\end{array}
$\\
 where $r'$ is next rule.

Now, we introduce a slight modified version of the Hennessy Milner Logic (the Hennessy Milner link logic, HM-linkL)~\cite{}; due to the reasons we explained above, we do not want to look at the complete transition labels, thus we rely on our simple assertion language:
\begin{definition}[HM-linkL]
Let   $F$ be an assertion, then 
the set of  HM-linkL formulae ${\cal G}$are built by the following syntax:\\
$$
G,H ::= {\tt t} ~\mid~ {\tt f} ~\mid~ G\wedge H ~\mid~ G \vee H ~\mid~ \langle F \rangle G ~\mid~ [F]G
$$
\end{definition}

\begin{definition}[semantics of $G$]
 We define $\llbracket G\rrbracket \subseteq {\cal P}$ by induction on the structure of $G \in {\cal G}$:%(read as the transition label $v$ satisfies the formula $F$) as: 
\[
\begin{array}{ccl@{\hspace{1cm}}ccl}
 \llbracket {\tt t}\rrbracket &\defeq& {\cal P} &  \llbracket G\wedge H\rrbracket &\defeq&  \llbracket G\rrbracket  \cap \llbracket H\rrbracket  \\
\llbracket {\tt f}\rrbracket& \defeq& \emptyset &
  \llbracket G\vee H\rrbracket &\defeq&  \llbracket G\rrbracket  \cup \llbracket H\rrbracket \\
%\item   $ v \entails F_1 \vee F_2$ iff   $v \entails F_1$  $\vee$  $v \entails F_2$.
\end{array}
\]
\[
\begin{array}{lcl}
 \llbracket  \langle F \rangle G \rrbracket &\defeq& \{P \in {\cal P}: \exists P', v.P\xrightarrow{v}P' \mbox{ with } v \entails F \mbox{ and } P' \in  \llbracket  G \rrbracket\} \\
 \llbracket [F]G \rrbracket &\defeq& \{P \in {\cal P}: \forall P',v.P\xrightarrow{v}P' \mbox{ implies } v \entails F \mbox{ and } P' \in  \llbracket  G \rrbracket\}
\end{array}
\]
We write $P \vdash G$ ($P$ satisfies $G$) if and only if $P \in  \llbracket G \rrbracket$.
\end{definition}

Please recall that we have defined the behaviour of the contest in a non determinist way, thus 
at each step, different possible sets of entities can be provided to the system. 
%situations where in the same conditions, two alternative (mutually exclusive) behaviors can happen.

%With an example, we  illustrate a possible (interesting) application of the implication in a query.
%Let us consider the following rule that acts as follows: if reagent $a$ is present and inhibitor $b$ is absent,  either $c$ or $z$ can be produced. 
%Typically, a non-deterministic model reflects incomplete knowledge of the real system.
%
%
%
%{\small
%\[
%\begin{array}{lcl}
%P & \defeq& \startchain{\silent}\chainedlink{a_i}{\noact}\chainedlink{\noact}{a_o}\chainedlink{\overline{b}_i}{\noact}\chainedlink{\noact}{\overline{b}_o}
%\chainedlink{r_2}{\noact}\chainedlink{\noact}{p_1}
%\chainedlink{\widetilde{c}_i}{\noact}\chainedlink{\noact}{\widetilde{c}_o}\chainend{p_2}.P  \\
%&&+\\
% && \startchain{\silent}\chainedlink{a_i}{\noact}\chainedlink{\noact}{a_o}\chainedlink{\overline{b}_i}{\noact}\chainedlink{\noact}{\overline{b}_o}
%\chainedlink{r_2}{\noact}\chainedlink{\noact}{p_1}
%\chainedlink{\widetilde{z}_i}{\noact}\chainedlink{\noact}{\widetilde{z}_o}\chainend{p_2}.P  \\
%&&+\\
%&&\dots\\
%\end{array}
%\]
%}
%In this situation, a comparison between two models including the above rule,  based on the query $\tilde{z} \vee \underline{b} \rightarrow x$  can happen that one system always satisfies the query, and the other one not.
%This difference could reveals that in the first system there is another rule producing $z$, i.e. biologically there is another path to produce $z$ that is activated in the actual system configuration:
%{\small
%\[
%\begin{array}{lcl}
%P & \defeq& \dots \\
%&&+\\
% && \startchain{\silent}\chainedlink{a_i}{\noact}\chainedlink{\noact}{a_o}\chainedlink{z_i}{\noact}\chainedlink{\noact}{z_o}\chainedlink{\overline{b}_i}{\noact}\chainedlink{\noact}{\overline{b}_o}
%\chainedlink{r_2}{\noact}\chainedlink{\noact}{p_1}
%\chainedlink{\widetilde{z}_i}{\noact}\chainedlink{\noact}{\widetilde{z}_o}\chainend{p_2}.P  \\
%&&+\\
%&&\dots\\
%\end{array}
%\]
%}

\begin{definition}[Biosimilarity  $ \sim_F $]
A \emph{biosimulation} $\mathbf{R}_F$, with respect to the predicate $F$, is a binary relation over {\tt link}-calculus processes such that, if $P \mathrel{\mathbf{R}_F} Q$ then:
%\begin{itemize}
%\item 
 \begin{center} %s$\exists \, v,v'$ s.t. 
 $P \xrightarrow{v} P'$ with $v \entails F$ \, iff  \, $Q  \xrightarrow{v'} Q'$ with $v' \entails F$ and $P' \mathrel{\mathbf{R}_F} Q'$.
 \end{center}
%\item i$Q \xrightarrow{v} Q'$ with $v \entails F$ iff $\exists v',P'$ s.t.~$v \stretcheq v'$, $P  \xrightarrow{v'} P'$ with $v \entails F$. and $P' \mathrel{\mathbf{R}_{F}} Q'$.
%\end{itemize}
\noindent
We let $ \sim_F $ denote the largest  biosimulation and we say that $P$ is \emph{biosimilar} to $Q$, with respect to $F$, if $P \sim_F Q$.
\end{definition}

### 2020-03-27 06-examples.tex
Add files via upload

% !TEX root = ./main_RS2L_LNCS.tex
\section{Examples}
\label{ex:examples}

\subsection{Labelled transition system}
This example is inspired by the example in~\cite{BEMR11}, where
a deterministic transition system is coded in the reaction system
framework.
Here we consider the minimal deterministic transition system in Figure~\ref{fig:lts}.
%following the reaction system approach, to code it in the reaction system framework, we have to add
%extra labels, as in Figure~\ref{fig:ndlts2}.
\begin{figure}
\[
\xymatrix{
*+[o][F-]{q} \ar@/ ^/[r]^a \ar@(l,u)[]^b &
*+[o][F-]{w} \ar@/^/[l]^b \ar@(r,u)[]_a
}
\]
\caption{Minimal deterministic labelled transition system.}
\label{fig:lts}
\end{figure}
\noindent
In our chained {\tt link}-calculus encoding, $q$, $w$, $a$, and $b$ became the $entities$, the context can only provide $a$ and $b$ and the reaction rules are as follows:\\
\[
\begin{array}{cl@{\hspace{1cm}} cl}
1 \ &(\{q,a\},\{b\},q) & 2\ &(\{q,b\},\{a\},w)\\
3 \  &(\{w,a\},\{b\},q) & 4\ &(\{w,b\},\{a\},w)\\ 
\end{array}
\]
\paragraph{Encoding of the rules.}
The encoding of the rules for reactions is  given in a parametric way:
\[
\begin{array}{lcl}
P_n(q,b,w,a,q) & \defeq & \startchain{r_n}\chainedlink{q_i}{\noact}\chainedlink{\noact}{q_o}
                                                    \chainedlink{b_i}{\noact}\chainedlink{\noact}{b_o}
                                                     \chainedlink{\overline{w}_i}{\noact}\chainedlink{\noact}{\overline{w}_o}
					        \chainedlink{\overline{a}_i}{\noact}\chainedlink{\noact}{\overline{a}_o}
					        \chainedlink{r_n}{\noact}\chainedlink{\noact}{p_n}
					        \chainedlink{\tilde{q}_i}{\noact}\chainedlink{\noact}{\tilde{q}_o}
			\chainend{p_{n+1}}.P_n(q,b,w,a,q)  \\
			&& +\\
			&&P_n'(\overline{q}) \ +\ P_n'(\overline{b}) \ + \ P_n'(w)\ +\ P_n'(a)\\
			\end{array}
			\]
			\noindent
			where
			\[
\begin{array}{lcl}
P_n'(w) &\defeq &  \startchain{r_n}\chainedlink{w_i}{\noact}\chainedlink{\noact}{w_o}\chainedlink{r_{n+1}}{\noact}\chainedlink{\noact}{p_n} \chainend{p_{n+1}}.P_n'(w) 
% \ + \   \startchain{r_n}\chainedlink{w_i}{\noact}\chainedlink{\noact}{w_o}\chainedlink{r_{n+1}}{\noact}\chainedlink{\noact}{p_n} \chainend{p_{n+1}}.P_n(q,b,w,a,q) \\
%			&&+\\
%			&& \startchain{r_n}\chainedlink{\overline{b}_i}{\noact}\chainedlink{\noact}{\overline{b}_o}\chainedlink{r_{n+1}}{\noact}\chainedlink{\noact}{p_n}\chainend{p_{n+1}}.P_n(q,b,w,a,q) \ + \  \startchain{r_n}\chainedlink{a_i}{\noact}\chainedlink{\noact}{a_o}\chainedlink{r_{n+1}}{\noact}\chainedlink{\noact}{p_n} \chainend{p_{n+1}}.P_n(q,b,w,a,q) 
\end{array}
\]
\noindent
Then we have 
\[
\begin{array}{lcl @{\hspace{0.5cm}}lcl  }
P_1 & \defeq & P_1(q,a,w,b,q) & P_3 & \defeq & P_3(w,a,q,b,w)  \\
P_2 & \defeq & P_2(q,b,w,a,w) &P_4 & \defeq & P_4(w,b,q,a,q) 
\end{array}
\]
\noindent
and we put $r_1 = \silent$, $r_5 = cxt$ and  $p_5 = \silent$.

\paragraph{Encoding of the entities.}
As for reactions, also the encoding of the entities is given in a parametric way.
Here we differentiate the encoding for the entities that are not provided by the context and that can be produced by the reactions, and the ones that can be provided by the context and that are not produced by the reactions.\\
Here, the entities $p$ and $w$ that are not provided by the context:
\[
\begin{array}{lcl@{\hspace{1cm}}lcl}
P (q)&\defeq & \sum_{h=0}^1 (\startchain{q_i}\chainedlink{q_o}{\noact}\chainend{\noact})^h\ \link{\tilde{q}_i}{\tilde{q}_o}.P(q) &
\overline{P} (q)&\defeq & \sum_{h=1}^2 (\startchain{\overline{q}_i}\chainedlink{\overline{q}_o}{\noact}\chainend{\noact})^h\ \link{\tilde{q}_i}{\tilde{q}_o}.P(q)\\
&& + & && +\\
&& \link{q_i}{q_o}.\overline{P}(q) &
&& \sum_{h=1}^2 (\startchain{\overline{q}_i}\chainedlink{\overline{q}_o}{\noact}\chainend{\noact})^h.\overline{P}(q)\\
\end{array}
\]
Then, we have
\[
\begin{array}{lcl @{\hspace{1cm}}lcl}
P_q & \defeq & P(q) & P_w& \defeq & P(w)\\
\end{array}
\]
Here, the entities that are provided by the context:
\[
\begin{array}{lcl @{\hspace{1cm}}lcl}
E(a) & \defeq &  \sum_{h=0}^1 (\startchain{a_i}\chainedlink{a_o}{\noact}\chainend{\noact})^h\ \link{\hat{a}_i}{\hat{a}_o}.E(a) &  \overline{E} (a)&\defeq & \sum_{h=1}^2 (\startchain{\overline{a}_i}\chainedlink{\overline{a}_o}{\noact}\chainend{\noact})^h\ \link{\hat{a}_i}{\hat{a}_o}.E(a)\\
&& + &  && +\\
&& \sum_{h=0}^1 (\startchain{a_i}\chainedlink{a_o}{\noact}\chainend{\noact})^h\ \link{\underline{a}_i}{\underline{a}_o}.\overline{E}(a) &
&& \sum_{h=1}^2 (\startchain{\overline{a}_i}\chainedlink{\overline{a}_o}{\noact}\chainend{\noact})^h\ \link{\underline{a}_i}{\underline{a}_o}.\overline{E}(q)\\
\end{array}
\]
Then, we have
\[
\begin{array}{lcl @{\hspace{1cm}}lcl}
P_a & \defeq & E(a) & P_b& \defeq & E(b)\\
\end{array}
\]
For the context, the encoding follows:
\[
\begin{array}{lcl @{\hspace{0.5cm}}c@{\hspace{0.5cm}} l}
Cxt &\defeq &  \startchain{cxt}\chainedlink{\hat{a}_i}{\noact}\chainedlink{\noact}{\hat{a}_o}\chainedlink{\underline{b}_i}{\noact}\chainedlink{\noact}{\underline{b}_i}\chainend{p_1}.Cxt
&+
&\startchain{cxt}\chainedlink{\hat{b}_i}{\noact}\chainedlink{\noact}{\hat{b}_o}\chainedlink{\underline{a}_i}{\noact}\chainedlink{\noact}{\underline{a}_i}\chainend{p_1}.Cxt\\
\end{array}
\]
We model the context to always offer either  $a$ or $b$, and no both the entities. The reason is that in
the other cases (providing both $a$ and $b$ or neither of them) leads the system to be stuck.\\
Now, we assume that we have an initial configuration containing the entities $q$ and $b$ (with$\tilde{q},\tilde{w},\tilde{a},\tilde{b}$ grouping all the possibile decorations for these names):
\[
\begin{array}{lcl}
Sys  \defeq \ \restrict{r1,r2,r3,r4,\tilde{q},\tilde{w},\tilde{a},\tilde{b}} (P_q \mid \overline{P}_w  \mid  P_a  \mid \overline{P}_b  \mid  P_1  \mid  P_2  \mid  P_3  \mid P_4  \mid Cxt).
\end{array}
\] 
Then only the second reaction can be applied, and the label transition follows
{\small
\[
\begin{array}{l}
\restrict{r1,r2,r3,r4,\tilde{q},\tilde{w},\tilde{a},\tilde{b}}\\
 (\startchain{\silent}\chainedlink{\overline{a}_i}{\bf  \overline{a}_i}\chainedlink{\bf \overline{a}_o}{\overline{a}_o}\chainedlink{r_2}{r_2}\chainedlink{q_i}{\bf q_i}\chainedlink{\bf q_i}{q_i}\chainedlink{b_i}{\bf b_i}\chainedlink{\bf b_i}{b_i}\chainedlink{\overline{w}_i}{\bf \overline{w}_i}\chainedlink{\bf \overline{w}_o}{\overline{w}_o}\chainedlink{\overline{a}_i}{\bf \overline{a}_i}\chainedlink{\bf \overline{a}_o}{\overline{a}_o}\chainedlink{r_3}{r_3}\chainedlink{\overline{a}_i}{\bf \overline{a}_i}\chainedlink{\bf \overline{a}_o}{\overline{a}_o}\chainedlink{r_4}{r_4}\chainedlink{\overline{w}_i}{\bf \overline{w}_i}\chainedlink{\bf \overline{w}_o}{\overline{w}_o}\chainedlink{cxt}{cxt}\chainedlink{\hat{a}_i}{\bf \hat{a}_i}\chainedlink{\bf \hat{a}_o}{\hat{a}_o}\chainedlink{\underline{b}_i}{\bf \underline{b}_i}\chainedlink{\bf \underline{b}_o}{\underline{b}_o}\chainedlink{p_1}{p_1}\chainedlink{p_2}{p_2}\chainedlink{\tilde{w}_i}{\bf \tilde{w}_i}\chainedlink{\bf \tilde{w}_o}{\tilde{w}_o}\chainedlink{p_3}{p_3}\chainedlink{p_4}{p_4}\chainend{\silent})
\end{array}
\]}
The parts in bold are provided by the entity processes, the other parts are provided by the processes encoding the reactions and by the process encoding the context (starting at $cxt$ and ended at $p_1$.
In the label we can read that the rules $1$ and $4$ have been not executed as the entity $a$ is absent, 
the rule $3$ has been not applied as the entity $w$ is absent, then only rule $2$ has been applied, and then it has produced entity $w$. Also, in the next state the context will provide entity $a$ and will not entity $b$.

\subsection{A  biological toy example}
Here we consider a biological toy example where a gene $a$  codes for a protein $T$ when molecules 
$G$ is present and $C$ is absent, and in the opposite situation $a$ codes for protein $T'$.
This behavior is encoded in rules $1$ and $2$.
Then, rule three codes for the production of $C$ when proteins $T$ and $F$ are present, and $T'$ absent;
rule four codes for the production of $G$ when  proteins $T'$ is present and $F$ is absent.
\paragraph{Encoding of the rules.}
The encoding of the rules for reactions is  given in a parametric way:
\[
\begin{array}{lcl}
P_n(a,G,C,T) & \defeq &\startchain{r_n}\chainedlink{a_i}{\noact}\chainedlink{\noact}{a_o}
                                                    \chainedlink{G_i}{\noact}\chainedlink{\noact}{G_o}
                                                     \chainedlink{\overline{C}_i}{\noact}\chainedlink{\noact}{\overline{C}_o}
					        \chainedlink{r_{n+1}}{\noact}\chainedlink{\noact}{p_n}
					        \chainedlink{\tilde{T}_i}{\noact}\chainedlink{\noact}{\tilde{T}_o}
			\chainend{p_{n+1}}.P_n(a,G,C,T)  \\
				&& +\\
			&&P_n'(\overline{a}) \ +\ P_n'(\overline{G}) \ + \ P_n'(C)\\
			\end{array}
			\]
			\noindent
			where
			\[
\begin{array}{lcl}
P_n'(C) &\defeq &  \startchain{r_n}\chainedlink{C_i}{\noact}\chainedlink{\noact}{C_o}\chainedlink{r_{n+1}}{\noact}\chainedlink{\noact}{p_n} \chainend{p_{n+1}}.P_n'(C) 
% \ + \   \startchain{r_n}\chainedlink{w_i}{\noact}\chainedlink{\noact}{w_o}\chainedlink{r_{n+1}}{\noact}\chainedlink{\noact}{p_n} \chainend{p_{n+1}}.P_n(q,b,w,a,q) \\
%			&&+\\
%			&& \startchain{r_n}\chainedlink{\overline{b}_i}{\noact}\chainedlink{\noact}{\overline{b}_o}\chainedlink{r_{n+1}}{\noact}\chainedlink{\noact}{p_n}\chainend{p_{n+1}}.P_n(q,b,w,a,q) \ + \  \startchain{r_n}\chainedlink{a_i}{\noact}\chainedlink{\noact}{a_o}\chainedlink{r_{n+1}}{\noact}\chainedlink{\noact}{p_n} \chainend{p_{n+1}}.P_n(q,b,w,a,q) 
\end{array}
\]

Then we have 
\[
\begin{array}{lcl @{\hspace{0.5cm}}lcl @{\hspace{0.5cm}}lcl }
P_1 & \defeq & P_1(a,G,C,T) &   P_2 & \defeq & P_2(a,C,G,T') &   P_3 & \defeq & P_3(F,T,T',C)  \\
\end{array}
\]
\[
\begin{array}{lcl}
P_4 & \defeq &\startchain{r_4}\chainedlink{T'_i}{\noact}\chainedlink{\noact}{T'_o}
                                                     \chainedlink{\overline{F}_i}{\noact}\chainedlink{\noact}{\overline{F}_o}
					        \chainedlink{cxt}{\noact}\chainedlink{\noact}{p_4}
					        \chainedlink{\tilde{G}_i}{\noact}\chainedlink{\noact}{\tilde{G}_o}
			\chainend{\silent}.P_4  \\
				&& +\\
&&\startchain{r_4}\chainedlink{\overline{T'}_i}{\noact}\chainedlink{\noact}{\overline{T'}_o}\chainedlink{cxt}{\noact}\chainedlink{\noact}{p_4} \chainend{\silent}.P_4 \ +\  \startchain{r_4}\chainedlink{F_i}{\noact}\chainedlink{\noact}{F_o}\chainedlink{cxt}{\noact}\chainedlink{\noact}{p_4} \chainend{\silent}.P_4
\end{array}
\]
\noindent
and we put $r_1 = \silent$.
\paragraph{Encoding of the entities.}
As for reactions, also the encoding of the entities is given in a parametric way.
Here we differentiate three types of encodings: (1) for the entities that are not  provided by the context and  can be produced by the reactions; (2) for the entities that can be provided by the context and can be  produced by the reactions; (3) for the entities that are only provided by the context.\\
Here, the entities $T$ and $T'$ that can be produced by the reactions and that are not provided by the context:
\[
\begin{array}{lcllcl}
P (T,\tilde{T})&\defeq & \sum_{h=0}^1 (\startchain{T_i}\chainedlink{T_o}{\noact}\chainend{\noact})^h\ \link{\tilde{T}_i}{\tilde{T}_o}.P (T,\tilde{T})
& +  && \link{T_i}{T_o}.P (\overline{T},\tilde{T}) \\
%\overline{P} (q)&\defeq &  \sum_{h=0}^1 (\startchain{\overline{T}_i}\chainedlink{\overline{T}_o}{\noact}\chainend{\noact})^h\link{\tilde{T}_i}{\tilde{T}_o}.P(q)\\
%&& + & && +\\
%&& \link{T_i}{T_o}.\overline{P}(T) &
%&& \link{\overline{T}_i}{\overline{T}_o}.\overline{P}(T)\\
\end{array}
\]
Then, we have
\[
\begin{array}{lcl @{\hspace{0.5cm}}lcl @{\hspace{0.5cm}}lcl @{\hspace{0.5cm}}lcl}
P_T & \defeq & P(T,\tilde{T})& \overline{P}_T & \defeq & P(\overline{T},\tilde{T}) & P_{T'}& \defeq & P(T',\tilde{T'}) & \overline{P}_{T',\tilde{T'}}& \defeq & P(\overline{T'},\tilde{T'})\\\
\end{array}
\]
Here, the entities that can be produced by the reactions and that can be  provided by the context:
\[
\begin{array}{lcl @{\hspace{1cm}}lcl}
P(C,\hat{C},\underline{C},\tilde{C}) & \defeq &  \sum_{h=0}^1 (\startchain{C_i}\chainedlink{C_o}{\noact}\chainend{\noact})^h\ \link{\hat{C}_i}{\hat{C}_o} (\startchain{\noact}\chainedlink{\noact}{\tilde{C}_i}\chainend{\tilde{C}_o})^h.   P(C,\hat{C},\underline{C},\tilde{C})\\
&&+\\
&& \sum_{h=0}^1 (\startchain{C_i}\chainedlink{C_o}{\noact}\chainend{\noact})^h\ \link{\underline{C}_i}{\underline{C}_o}.P(\overline{C},\hat{C},\underline{C},\tilde{C})\\
&&+\\
 && \sum_{h=0}^1 (\startchain{C_i}\chainedlink{C_o}{\noact}\chainend{\noact})^h\ \link{\underline{C}_i}{\underline{C}_o} \startchain{\noact}\chainedlink{\noact}{\hat{C}_i}\chainend{\hat{C}_o}.   P(C,\hat{C},\underline{C},\tilde{C}) \\
 %&  \overline{E} (a)&\defeq & \sum_{h=1}^2 (\startchain{\overline{a}_i}\chainedlink{\overline{a}_o}{\noact}\chainend{\noact})^h\ \link{\hat{a}_i}{\hat{a}_o}.E(a)\\
%&&  &  && +\\
%&& 
%&
%&& \sum_{h=1}^2 (\startchain{\overline{a}_i}\chainedlink{\overline{a}_o}{\noact}\chainend{\noact})^h\ \link{\underline{a}_i}{\underline{a}_o}.\overline{E}(q)\\
\end{array}
\]
Then, we have
\[
\begin{array}{lcl @{\hspace{1cm}}lcl@{\hspace{0.4cm}}lcl@{\hspace{0.4cm}}lcl}
P_C & \defeq &  P(C,\hat{C},\underline{C},\tilde{C})&  P_G& \defeq &   P(G,\hat{G},\underline{G},\tilde{G}) \\
\overline{P}_C & \defeq &  P(\overline{C},\hat{C},\underline{C},\tilde{C}) & 
\overline{P}_G& \defeq &   P(\overline{G},\hat{G},\underline{G},\tilde{G})\\
\end{array}
\]

Here, the encoding of the entity $F$ that can be provided by the context:
\[
\begin{array}{lcllcl}
P_F&\defeq & \sum_{h=0}^1 (\startchain{F_i}\chainedlink{F_o}{\noact}\chainend{\noact})^h\ \link{\hat{F}_i}{\hat{F}_o}.P_F
& +  && \sum_{h=0}^1 (\startchain{F_i}\chainedlink{F_o}{\noact}\chainend{\noact})^h. \link{\underline{F}_i}{\underline{F}_o}.\overline{P}_F \\
\overline{P}_F&\defeq & \sum_{h=0}^1 (\startchain{\overline{F}_i}\chainedlink{\overline{F}_o}{\noact}\chainend{\noact})^h\ \link{\hat{F}_i}{\hat{F}_o}.P_F
& +  && \sum_{h=0}^1 (\startchain{\overline{F}_i}\chainedlink{\overline{F}_o}{\noact}\chainend{\noact})^h. \link{\underline{F}_i}{\underline{F}_o}.\overline{P}_F \\
%\overline{P} (q)&\defeq &  \sum_{h=0}^1 (\startchain{\overline{T}_i}\chainedlink{\overline{T}_o}{\noact}\chainend{\noact})^h\link{\tilde{T}_i}{\tilde{T}_o}.P(q)\\
%&& + & && +\\
%&& \link{T_i}{T_o}.\overline{P}(T) &
%&& \link{\overline{T}_i}{\overline{T}_o}.\overline{P}(T)\\
\end{array}
\]
Now, the context follow can non deterministically provide the entities: $C$, $G$ $F$:\\
\[
\begin{array}{lcl}
 Cxt & \defeq &
 \sum_{\tiny \begin{array}{l}C^{\star} \in \{C,\overline{C}\}\\G^{\star} \in \{G,\overline{G}\}\\
 F^{\star} \in \{F,\overline{F}\}
  \end{array}}
 \startchain{cxt} \chainedlink{C_i^{\star}}{\noact}\chainedlink{\noact}{C_o^{\star}}\chainedlink{G_i^{\star}}{\noact}\chainedlink{\noact}{G_o^{\star}}
 \chainedlink{F_i^{\star}}{\noact}\chainedlink{\noact}{F_o^{\star}}\chainend{p_1}.Cxt\\
\end{array}
\]

Now, to show a possible composition of a transition label, we assume an system where the entities $a$, $T'$, and $G$ are present:
\[
\begin{array}{lcl}
Sys &\defeq &\restrict{\tilde{C},\tilde{G},\tilde{F}, \tilde{T},\tilde{T'}}( P_a\mid P_C\mid P_G\mid \overline{P}_F \mid \overline{P}_T \mid P_{T'} \mid P_1 \mid P_2 \mid P_3 \mid P_4)
\end{array}
\]
In the above configuration, reactions $1$ and $3$ can be applied, and also we assume that the context will provide the entity $F$, that will be available in the target configuration.
The transition label of

### 2020-03-29 06b-bisimulation.tex
Update 06b-bisimulation.tex

@@
 barbed bisimulation~\cite{10.1007/3-540-55719-9_114} only consider the execution of invisible actions, and then equates two processes when the expose the same prefixes; for the mobile ambients~\cite{CardelliG00}, a process algebra equipped with a reduction semantics, a notion of
-behavioural equivalence equates two processes when they expose the same ambients\cite{GC03}.
+behavioural equivalence equates two processes when they expose the same ambients~\cite{GC03}.
 
 There are some previous works based on bisimulation applied to models for biological systems.
-Barbuti et al \cite{BMMT08} define a classical setting for bisimulation for two formalisms: the Calculus of Looping Sequences, which is a rewriting system, and the Brane Calculi, which is based on process calculi.
+Barbuti et al~\cite{BMMT08} define a classical setting for bisimulation for two formalisms: the Calculus of Looping Sequences, which is a rewriting system, and the Brane Calculi, which is based on process calculi.
 Bisimulation is used to verify properties of the regulation of lactose degradation in
 Escherichia coli and the EGF signalling pathway. These calculi allow the authors to model membranes' behaviour.
-\cite{CTTV15} presents two quantitative behavioral equivalences over species of a 
+Cardelli et al~\cite{CTTV15} present two quantitative behavioral equivalences over species of a 
 chemical reaction network with semantics based on ordinary differential equations.
@@
 Bisimulation also relates species that have identical solutions at all time points when starting from the same initial conditions.
-Both the mentioned formalisms \cite{BMMT08,CTTV15} adopt a classical approach to bisimulation. Moreover our formalisms is different, for different applications and purposes.
-
+Both the mentioned formalisms~\cite{BMMT08,CTTV15} adopt a classical approach to bisimulation. 
 Albeit the bisimulation is a powerful tool for verifying if the behaviour of two different software 
@@
 
-For this reason, we propose an alternative notion of bisimulation, that hereafter we call biosimulation,
+For this reason, we propose an alternative notion of bisimulation, that hereafter we call \emph{biosimulation},
 that allows us  to compare two biological systems by restricting the observation to only the limited events of interest.
+This allows one to tailor the equivalence to different applications and purposes.
 
 The transition labels of our systems record detailed information about all the reactions that have been applied in one transition, about the elements that acted as reagents, as inhibitors or that have been produced, or that have been provided by the context.
+All these information are stored in the label because they are necessary to compose a transition in a modular way.
+Depending on the application, only a suitable abstraction over the label can be of interest.
 
-In a way, we want to query our transition labels to get only the information we are interested in.
+In a way, at each step of the bisimulation game, we want to query our transition labels to get only the information we care about.
 To this goal, we introduce a simple language
@@
 
-First, we need to introduce the $flat$ function that takes a link chain and returns a simple string
-by eliminating all the channel matching pairs leaving just one channel per each pair:\\
+For example we would like to express properties about each step of the simulation of a system like the following:
+\begin{enumerate}
+\item Has the entity $s_i$ been used by rule $r_j$ as reagent?
+\item Has the entity $s_i$ blocked the application of  rule $r_j$?
+\item Has the entity $s_i$ been produced by rule $r_j$?
+\item Has the entity $s_i$ been produced by some rule?
+\item Has the entity $s_i$ been provided by the context? 
+\item Has the rule $r_j$ not been applied? 
+%\item Has the entity $s$ been used by some rule as reagent ?
+%\item Has the entity $s$ blocked  the application of some rules ?
+%\item Has the entity $s$ been produced ?
+%\item Has the entity $s$ been provided by the context ?
+%\item Has the rule $i$ been applied ?
+\end{enumerate}
+
+As detailed before, in the following we assume that: (i) the context can be non-deterministic, otherwise it makes little sense to rely on bisimulation to observe the branching structure of system dynamics; (ii) we are interested in observing the names of the entities involved in the transitions and also the rules that have been applied, thus we assume top level restrictions are absent and rely on solid transitions only (with leftmost and rightmost silent actions).
+
+\subsection{Assertion language}
+
+To simplify the query language, we first introduce a function that takes a solid link chain and returns a simple string by eliminating all the channel matching pairs leaving just one channel per each pair.
+This transformation is harmless, in the sense that it retains all the information in the chain, because it is applied to solid chains only.
+The function $\mathit{flat}$ is defined inductively as follows:
 \[
 \begin{array}{lcl@{\hspace{0.5cm}}lcl@{\hspace{0.5cm}}lcl@{\hspace{0.5cm}}lcl}
-flat(\epsilon)&\defeq & \epsilon &
-flat(\link{\silent}{\beta}) & \defeq & \beta &
-flat(\link{\alpha}{\silent})& \defeq & \epsilon &
-flat(\link{\alpha}{\beta}) & \defeq &\beta \\
+\mathit{flat}(\epsilon)&\defeq & \epsilon &
+\mathit{flat}(\link{\silent}{\beta}) & \defeq & \beta &
+\mathit{flat}(\link{\alpha}{\silent})& \defeq & \epsilon &
+\mathit{flat}(\link{\alpha}{\beta}) & \defeq &\beta \\
 \end{array}\]
@@
 \begin{array}{lcl}
-flat(\link{\alpha}{\beta}\upsilon)& \defeq & \beta :: flat(\upsilon)\\
+\mathit{flat}(\link{\alpha}{\beta}\upsilon)& \defeq & \beta :: \mathit{flat}(\upsilon)\\
 \end{array}\\
 \]
-where  $::$ is the string concatenation operator.
+where  $::$ is the usual string concatenation operator.
 
-\noindent
-Now, we can introduce an assertion language that operates on simple strings
-and that combines regular expression operators with the classical $and$ and $or$ logical
-operators.
+Next, we introduce an assertion language that operates on simple strings
+and that combines regular expression operators with conjunction and disjunction.
 
@@
 \begin{array}{lcl}
-\zeta' &::= & \epsilon  ~\mid~ \zeta::\zeta'\\ 
-re&::=& . ~\mid ~   (\zeta \textrm{-}\zeta)  ~\mid ~ \zeta' ~\mid ~ re::re\\
-re' &::= & 
-%(re')|(re')  ~\mid~
-[\zeta\dots\zeta]~\mid~ [\hat{}  re] ~\mid ~ re* ~\mid ~ re+ \\
-F &::=& re' ~\mid ~  F \vee F ~\mid ~  F \wedge F\\
+\zeta & ::= & 
+\alpha ~\mid ~ 
+? ~\mid ~ 
+[N]
+\\
+F & ::= & 
+\epsilon ~\mid ~ 
+\zeta ~\mid ~ 
+F::F ~\mid ~ 
+F^+ ~\mid ~ 
+F^* ~\mid ~ 
+F \vee F ~\mid ~ 
+F \wedge F
 \end{array}
@@
 %\noindent where $s \in {\cal C}$, and  $\oplus \in \{\wedge, \vee, \to \}$. 
-where $\zeta \in \{s, \tilde{s}, \hat{s}, \overline{s}, \underline{s} \}$ with  $s \in {\cal C}$.
+where $\alpha \in \mathit{names}$ and $N\subseteq \mathit{names}$.
 %, and  $\oplus \in \{\wedge, \vee \}$. 
@@
 
-$\zeta'$ is the set of  strings composed by characters in the set ${\cal C}$.
-Then, the dot "$.$" stands for any character, $\zeta_1 \textrm{-}\zeta_2$ is a shorthand for the string
-starting with $\zeta_1$ and ending with $\zeta_2$ and that in the  middle includes all the characters
-following the alphanumeric order.
-%The round brackets $(\zeta')$ build a single string  $\zeta'$, i.e. a non-separable sequence of characters.
- The square brackets match one of the characters they contain.
- The symbol $[\hat{} \dots]$, inside the square brackets excludes any of the characters they contain.
- The star $*$ allows the previous character to be replicated zero or more times.
- The plus symbol $+$ allows the previous character to be replicated at least one or more times.
-Then, we can combine the patterns build with the above constructors with the logic operators $\vee$ and
-$\wedge$.
+Roughly, a \emph{singleton} assertion $\zeta$ denotes either a string composed by a single symbol $\alpha$ (one of the symbols in the set $\mathit{names}$ for denoting a particular entity, rule, production or context), or
+the wildcard $?$ that stands for any symbol, or the pattern $[N]$ that stands for any of the strings composed by a single symbol in the set $N$.
+Clearly $?$  is just a shorthand for $[\mathit{names}]$.
+We write $\mathbf{0}$ for $[\emptyset]$ and $[s_1,...,s_n]$ instead of $[\{s_1,...,s_n\}]$.
 
-\begin{definition}[Semantics]\label{sec:semantics}
+An \emph{assertion} $F$ is either the empty string $\epsilon$, a singleton assertion $\zeta$, the concatenation of two assertions $F_1::F_2$, the replication of $F$ for 1 or more times $F^+$, the replication of $F$ for 0 or more times $F^*$, the disjunction of two assertions $F_1 \vee F_2$ or their conjunction $F_1 \wedge F_2$.
+We write $\star$ as an abbreviation of $?^*$.
+
+An assertion denotes a set of strings over the alphabet $\mathit{names}$ as expected.
+
+\begin{definition}[Semantics of assertions]
+ We define $\llbracket F\rrbracket \subseteq\wp(\mathit{names}^*)$ by induction on the structure of $F$: 
+\[
+\begin{array}{rcl}
+ \llbracket \alpha \rrbracket &\defeq&  \{\alpha\} \\
+ \llbracket ? \rrbracket &\defeq&   \mathit{names}\\
+ \llbracket [N]\rrbracket &\defeq&   N\\
+ \llbracket \epsilon \rrbracket &\defeq&  \{\epsilon\} \\
+ \llbracket F_1::F_2 \rrbracket &\defeq& \{\omega_1 :: \omega_2~\mid~ \omega_1 \in \llbracket F_1 \rrbracket \wedge \omega_2 \in \llbracket F_2 \rrbracket\}  \\
+ \llbracket F^+ \rrbracket &\defeq& \llbracket F \rrbracket^+  \\
+ \llbracket F^* \rrbracket &\defeq& \llbracket F \rrbracket^*  \\
+ \llbracket  F_1 \vee F_2 \rrbracket &\defeq& \llbracket  F_1\rrbracket \cup \llbracket  F_2\rrbracket  \\
+ \llbracket  F_1 \wedge F_2 \rrbracket &\defeq& \llbracket F_1 \rrbracket \wedge \llbracket F_2\rrbracket 
+\end{array}
+\]
+\end{definition}
+
+\begin{definition}[Satisfaction as membership]\label{sec:semantics}
 Let  $\upsilon$ be  a transition label, and $F$ be an assertion.
-We define $\upsilon \entails F$ (read as the transition label $v$ satisfies the formula $F$) as: 
-\begin{itemize}
- \item   $\upsilon \entails \zeta $ iff $\zeta \in flat(\upsilon)$. 
- \item   $\upsilon \entails F_1 \wedge F_2$ iff   $\upsilon \entails F_1$  $\wedge$  $\upsilon \entails F_2$.
-\item   $ \upsilon \entails F_1 \vee F_2$ iff   $\upsilon \entails F_1$  $\vee$  $\upsilon \entails F_2$.
-\end{itemize}
-If it is not the case that $\upsilon \entails  F $, then we say that $F$ does not hold at $v$ and we write $\upsilon \not\entails F$. 
+We write $\upsilon \entails F$ (read as the transition label $v$ satisfies the assertion $F$) 
+if $\mathit{flat}(\upsilon)\in \llbracket F\rrbracket$, otherwise we write $\upsilon \not\entails F$ and say that $F$ does not hold at $v$. 
 \end{definition}
+
+Given two transition labels $v,w$ we write $v\equiv_F w$ if $v \entails F\ \Leftrightarrow\ w \entails F$, i.e. if both $v,w$ satisfy $F$ or they do not.
 
@@
 %\end{definition}
-Our  language allows us to check if the simulation of a system (i.e. the transition labels, or in other words the trace of a simulation) 
-satisfies the properties of the following types:% that can be expressed using our assertion language. Some simple examples  follow:
+
+The corresponding formulas to the sample queries listed above are as follows:\\
+
 \begin{enumerate}
-\item Has the entity $s$ been used by rule $r$  as reagent ?
-\item Has the entity $s$ blocked the application of  rule $r$ ?
-\item Has the entity $s$ been produced by some rule ?
-\item Has the entity $s$ been produced by rule $p$ ?
-\item Has the entity $s$ been provided by the context ? 
-\item Has the rule $i$ been applied ? 
-%\item Has the entity $s$ been used by some rule as reagent ?
-%\item Has the entity $s$ blocked  the application of some rules ?
-%\item Has the entity $s$ been produced ?
-%\item Has the entity $s$ been provided by the context ?
-%\item Has the rule $i$ been applied ?
+\item
+$\star :: r_j :: [ s_1,...,s_n ]^* :: s_i :: [  s_1,...,s_n ]^* :: [  \overline{s}_1,...,\overline{s}_n ]^+ :: \star$
+\item
+$\star :: r_j :: [s_i, \overline{s}_i] :: r_{j+1} :: \star$
+\item
+$\star :: p_j :: [  \mathit{ents} ]^* :: \tilde{s}_i :: \star$
+\item
+$\star :: \tilde{s}_i :: \star$
+\item
+$\star :: \hat{s}_i :: \star$
+\item
+$\star :: r_j :: ? :: r_{j+1} :: \star$
 \end{enumerate}
-The corresponding formulas are as follows:\\
+where in 1, 2, 6 we exploit the fact that in a reaction $(R,I,P)$ the sets $R$ of reactants and  $I$ of inhibitors are non empty, so that if there is only one symbol between the occurrence of $r_j$ and $r_{j+1}$ it means the reaction $r_j$ has not been applied.
+Viceversa, if the reaction $r_j$ has been applied the occurrence of $r_j$ must be followed by at least one of the symbols in $\{s_1,...,s_n\}$ and then by at least one of the symbols in $\{\overline{s}_1,...,\overline{s}_n\}$.
 
-$
-\begin{array}{cl}
-1. & r[  s1 \dots sn ] *s[  s1 \dots sn ]* \\
-2. & r [ s \overline{s}]  r' \\
-3. & \tilde{s}\\
-4. & p[  \tilde{s1} \dots \tilde{sn} ] *\tilde{s}[  \tilde{s1} \dots \tilde{sn} ]* \\
-5. & \hat{s}\\
-6. &DA DECIDERE come
-\end{array}
-$\\
- where $r'$ is next rule.
+The notion of biosimulation builds on the above language of assertions to parameterize the induced equivalence on the property of interest. 
+Please recall that we have defined the behaviour of the contest in a non determinist way, thus 
+at each step, different possible sets of entities can be provided to the system and different sets of reaction can be enabled/disabled. 
+Biosimulation can thus be used to compare the behaviour of different systems that share some of the reactions or entities or also to compare the behaviour of the same set of reaction rules when different contexts are provided.
 
-Now, we introduce a slight modified version of the Hennessy Milner Logic (the Hennessy Milner link logic, HM-linkL)~\cite{}; due to the reasons we explained above, we do not want to look at the complete transition labels, thus we rely on our simple assertion language:
-\begin{definition}[HM-linkL]
-Let   $F$ be an assertion, then 
-the set of  HM-linkL formulae ${\cal G}$are built by the following syntax:\\
-$$
-G,H ::= {\tt t} ~\mid~ {\tt f} ~\mid~ G\wedge H ~\mid~ G \vee H ~\mid~ \langle F \rangle G ~\mid~ [F]G
-$$
-\end{definition}
-
-\begin{definition}[semantics of $G$]
- We define $\llbracket G\rrbracket \subseteq {\cal P}$ by induction on the structure of $G \in {\cal G}$:%(read as the transition label $v$ satisfies the formula $F$) as: 
-\[
-\begin{array}{ccl@{\hspace{1cm}}ccl}
- \llbracket {\tt t}\rrbracket &\defeq& {\cal P} &  \llbracket G\wedge H\rrbracket &\defeq&  \llbracket G\rrbracket  \cap \llbracket H\rrbracket  \\
-\llbracket {\tt f}\rrbracket& \defeq& \emptyset &
-  \llbracket G\vee H\rrbracket &\defeq&  \llbracket G\rrbracket  \cup \llbracket H\rrbracket \\
-%\item   $ v \entails F_1 \vee F_2$ iff   $v \entails F_1$  $\vee$  $v \entails F_2$.
-\end{array}
-\]
-\[
-\begin{array}{lcl}
- \llbracket  \langle F \rangle G \rrbracket &\defeq& \{P \in {\cal P}: \exists P', v.P\xrightarrow{v}P' \mbox{ with } v \entails F \mbox{ and } P' \in  \llbracket  G \rrbracket\} \\
- \llbracket [F]G \rrbracket &\defeq& \{P \in {\cal P}: \forall P',v.P\xrightarrow{v}P' \mbox{ implies } v \entails F \mbox{ and } P' \in  \llbracket  G \rrbracket\}
-\end{array}
-\]
-We write $P \vdash G$ ($P$ satisfies $G$) if and only if $P \in  \llbracket G \rrbracket$.
-\end{definition}
-
-Please recall that we have defined the behaviour of the contest in a non determinist way, thus 
-at each step, different possible sets of entities can be provided to the system. 
 %situations where in the same conditions, two alternative (mutually exclusive) behaviors can happen.
@@
 \begin{definition}[Biosimilarity  $ \sim_F $]
-A \emph{biosimulation} $\mathbf{R}_F$, with respect to the predicate $F$, is a binary relation over {\tt link}-calculus processes such that, if $P \mathrel{\mathbf{R}_F} Q$ then:
-%\begin{itemize}
-%\item 
- \begin{center} %s$\exists \, v,v'$ s.t. 
- $P \xrightarrow{v} P'$ with $v \entails F$ \, iff  \, $Q  \xrightarrow{v'} Q'$ with $v' \entails F$ and $P' \mathrel{\mathbf{R}_F} Q'$.
- \end{center}
+Given an assertion $F$, a \emph{biosimulation} $\mathbf{R}_F$ that respects $F$ is a binary relation over {\tt link}-calculus processes such that, if $P \mathrel{\mathbf{R}_F} Q$ then:
+ \begin{itemize} 
+ \item
+ for any $v,P'$ such that 
+ $P \xrightarrow{v} P'$ then there exist $v',Q'$ such that $Q  \xrightarrow{w} Q'$ with $v\equiv_F w$ and $P' \mathrel{\mathbf{R}_F} Q'$.
+\item
+ for any $w,Q'$ such that 
+ $Q \xrightarrow{w} Q'$ then there exist $v,P'$ such that $P  \xrightarrow{v} P'$ with $v\equiv_F w$ and $P' \mathrel{\mathbf{R}_F} Q'$.
+\end{itemize}
 %\item i$Q \xrightarrow{v} Q'$ with $v \entails F$ iff $\exists v',P'$ s.t.~$v \stretcheq v'$, $P  \xrightarrow{v'} P'$ with $v \entails F$. and $P' \mathrel{\mathbf{R}_{F}} Q'$.
 %\end{itemize}
-\noindent
 We let $ \sim_F $ denote the largest  biosimulation and we say that $P$ is \emph{biosimilar} to $Q$, with respect to $F$, if $P \sim_F Q$.
 \end{definition}
+
+\begin{remark}
+An alternative way to look at a biosimulation that respects $F$ is to define it as an ordinary bisimulation over the transition system labelled over $\{F,\neg F\}$ obtained by transforming each transition $P \xrightarrow{v} P'$ such that $v\entails F$ into $P \xrightarrow{F} P'$ and  each transition $P \xrightarrow{v} P'$ such that $v\not\entails F$ into $P \xrightarrow{\neg F} P'$.
+\end{remark}
+
+Now, we introduce a slightly modified version of the Hennessy Milner Logic~\cite{}, called bioHML; due to the reasons we explained above, we do not want to look at the complete transition labels, thus we rely on our simple assertion language to make it parametric to the assertion $F$ of interest:
+\begin{definition}[bioHML]
+Let   $F$ be an assertion, then 
+the set of bioHML formulas $G$ that respects $F$ are built by the following syntax:
+$$
+\begin{array}{rcl}
+\chi & ::= & F~\mid~ \neg F\\
+G &::= &{\tt t} ~\mid~ {\tt f} ~\mid~ G\wedge G ~\mid~ G \vee G ~\mid~ \langle \chi \rangle G ~\mid~ [\chi]G
+\end{array}
+$$
+\end{definition}
+
+\begin{remark}
+An alternative way to look at bioHML formulas is as ordinary HML formulas over the set of labels $\{F,\neg F\}$.
+\end{remark}
+
+As usual, the semantics of a bioHML formula is the set of processes that satisfy it.
+
+\begin{definition}[semantics of $G$]
+ We define $\llbracket G\rrbracket \subseteq {\cal P}$ by induction on the structure of $G$:
+ %(read as the transition label $v$ satisfies the formula $F$) as: 
+\[
+\begin{array}{ccl@{\hspace{1cm}}ccl}
+ \llbracket {\tt t}\rrbracket &\defeq& {\cal P} &  \llbracket G\wedge H\rrbracket &\defeq&  \llbracket G\rrbracket  \cap \llbracket H\rrbracket  \\
+\llbracket {\tt f}\rrbracket& \defeq& \emptyset &
+  \llbracket G\vee H\rrbracket &\defeq&  \llbracket G\rrbracket  \cup \llbracket H\rrbracket \\
+%\item   $ v \entails F_1 \vee F_2$ iff   $v \entails F_1$  $\vee$  $v \entails F_2$.
+\end{array}
+\]
+\[
+\begin{array}{lcl}
+ \llbracket  \langle \chi \rangle G \rrbracket &\defeq& \{P \in {\cal P}: \exists v,P'.\ P\xrightarrow{v}P' \mbox{ with } v \entails \chi \mbox{ and } P' \in  \llbracket  G \rrbracket\} \\
+ \llbracket [\chi]G \rrbracket &\defeq& \{P \in {\cal P}: \forall v,P'.\ P\xrightarrow{v}P' \mbox{ implies } v \entails \chi \mbox{ and } P' \in  \llbracket  G \rrbracket\}
+\end{array}
+\]
+We write $P \entails G$ ($P$ satisfies $G$) if and only if $P \in  \llbracket G \rrbracket$.
+\end{definition}
+
+We let $\mathcal{L}_F$ be the set of all bioHML formulas that respects $F$ and 
+we say that two processes $P,Q$ are bio-logically equivalent w.r.t.
+$F$, written $P \equiv_{\mathcal{L}_F}Q$, when $P$ and $Q$ satisfy the exactly the same bioHML formulas in $\mathcal{L}_F$, i.e.
+when for any $G\in \mathcal{L}_F$ we have $P \entails G\ \Leftrightarrow Q \entails G$.
+
+Finally, we extend the classical result establishing the correspondence between the logical equivalence induced by HML with bisimilarity for proving that biosimilarity coincides with bio-logical equivalence.
+
+\begin{theorem}
+$\sim_F\ =\ \equiv_{\mathcal{L}_F}$
+\end{theorem}

### 2020-03-29 00-abstract.tex
nuovo commit abs intro concl biblio
nuovo commit

% !TEX root =  main_RS2L_LNCS.tex

%Natural computing is an emerging area of research for modeling biological 
%computations by means of computer science formal methods. Reaction systems 
%have been defined in this area as a qualitative abstraction inspired by the 
%functioning of living cells, suitable to model the main
%mechanisms of biochemical reactions.
%Natural computing is an emerging area of research for modeling biological 
%computations by means of computer science formal methods. 
In the area of Natural Computing, reaction systems are
a qualitative abstraction inspired by the 
functioning of living cells, suitable to model the main
mechanisms of biochemical reactions.
This model has already been applied
and extended successfully to various areas of research. Reaction 
systems interact with the environment represented by the context, and
pose problems of implementation, as it is a new computation model. 
In this paper we consider the {\tt link}-calculus, which allows to model 
multiparty interaction in concurrent systems, and show that it allows to
embed reaction systems, by representing the behaviour of each entity and
preserving faithfully their features. 
We show the correctness and completeness 
of our embedding.
We illustrate our framework by showing the embedding of 
a few examples expressing computer science
and biological applications.
%how to embed a \emph{lac} operon 
%regulatory network. 
We then modify the classical notion (in concurrency) of bisimulation
making it more suitable for proving properties of reaction systems, in our
framework.
We define a new assertion language, which allows to specify
the properties to be verified, and show that it extends Hennessy-Milner logic
to our framework.
Finally, our methodology can contribute to increase the expressiveness
of reaction systems, by exploiting the interaction among 
different reaction systems.
%entities and the context.

### 2020-03-29 01-introduction.tex
nuovo commit abs intro concl biblio
nuovo commit

% !TEX root = ./main_RS2L_LNCS.tex

\section{Introduction}

Natural Computing is an emerging area of research which has two main 
aspects: human designed computing inspired by nature, and computation 
performed in nature. Reaction Systems (RSs)~\cite{BEMR11} are 
% model of
% formal computation 
a rewriting formalism
%model 
inspired by the way biochemical reactions take place in living 
cells.  
This theory has already shown to be relevant in several different 
fields, such as computer science~\cite{MPR15}, 
biology~\cite{ABP14,CMMBM12,Az17,BarbutiGLM16}, 
molecular chemistry~\cite{OY16}.
%,
% and it
%is continuously developing, with a growing number of workshops and
%international meetings. 
Reaction Systems formalise the mechanisms of biochemical systems, 
such as {\em facilitation} and {\em inhibition}. 
As a qualitative approximation of the real biochemical reactions, they
% do not take
%into account the quantity of reagents effectively present. They
consider if a necessary reagent is or not present, and likewise they
consider if an inhibiting molecule is or not present. 
The  possible reactants and inhibitors are called `entities'.
RSs model in a direct way the interaction of a living cell
with the environment (called `context'). However, two RSs are seen
as independent models and do not interact.

%Formally, Reaction Systems are defined as a rewriting formalism.
In this paper, we present an encoding from RSs, 
to the open multiparty process algebra c\CNA,\footnote{After `chained Core Network 
Algebra'.} a variant of the {\tt link}-calculus~\cite{BodeiBB12,BBB17} without name mobility.
This formalism allows several processes to synchronise and 
communicate altogether, at the same time, with a new communicating
mechanism based on links and link chains. 
Our initial motivation for introducing this mechanism was to encode
Mobile Ambients~\cite{CardelliG00}, getting a much stronger operational
correspondence than any available in the literature, such as the one in~\cite{B16}.
 This allowed us to easily encode calculi for biology equipped with membranes, as in~\cite{BodeiBBC14}.
Process calculi have been used successfully to model
biological processes, see~\cite{BBDFH18} for a recent survey.
We illustrate our embedding by means of some examples coming
both from the computer science and the biological field.
%and then we consider a more complex example, by modeling a RS 
%representing a regulatory network for \emph{lac} operon, presented in~\cite{CMMBM12}. 
We also show that our embedding preserves the main
features of RSs, and prove its correctness and completeness.
Then, we present a methodology for verifying formally
properties of reaction systems. The classical notion of bisimulation for 
process algebras allows to consider equivalent two processes
where one process can simulate all the actions executed 
by the other one and viceversa. In this paper we define a new
notion of bisimilarity which takes into account the characteristics
of biological systems. We define a new {\em bio-simulation} having in mind
the possibility that two interacting systems may be compared w.r.t.
a subset of the possible biological actions. This is useful for concentrating 
on the sub-model that one may need to study for a specific study or application,
without getting lost in the complexity of the full biological system, or network.
Besides bio-simulation, we define a completely new and simple assertion language,
which allows to specify properties of the reaction systems to be verified.
The classical logic of Hennessy-Milner~\cite{HM80} is a powerful
formalism for specifying properties of labeled transition systems, 
and we show that our assertions allow us to specify formulas 
that extend this logic to the case of complex labels.
Our formulas allow to abstract from some of the labels.

Our main contributions are as follows:
\begin{itemize}
\item the behaviour of the context, for each single 
entity,
%\footnote{Entities that belong to the `domain' of a RS.}
can be
specified in a recursive way as an ordinary process;
\item we can express the behaviour of entity mutation,
%change the behaviour of  each single entity $s$  in a
%way that a mutation $s'$ can be expressed; 
in such a way that the mutated entity
$s'$ can  take part to only a subset of rules requiring entity $s$;
\item with a little coding effort, two RSs can
communicate; i.e.  a subset of those entities that the context  can
provide, are then provided by a second RS;
\item as our translation results in  a  c\CNA \ system, from
each state only one transition can be generated, thus the c\CNA \
computation is fully deterministic.
\item we define an assertion language to specify local properties of 
reaction systems
\item we show that our assertions extend to our framework the Hennessy-Milner logic.
\end{itemize}

The main drawback of our proposal, is that the c\CNA \ translation is
verbose. Nevertheless it is clear that our
translation can be automatised by means of a proper front-end in
an implementation of the {\tt link}-calculus.

As we have remarked, in our translation, Reaction Systems 
get the ability to interact between them in a synchronized manner. 
This interaction is not foreseen in
the basic RS framework, as it can only happen with the context.
By exploiting recursion, the kind of interactions which can be 
defined can be complex and expressive.
Example~\ref{ex:twoRS} and more in general the discussion in 
Section~\ref{sec:discussion} show that 
the interaction between RSs can help to model new scenarios.
\paragraph{Related work}
We are not aware of any inference rules for the RS; it seems not the case
that a deterministic behaviour, as the one of the RS, once the initial state is 
fixed, could be defined by inference rules that classically are applied 
to define non-deterministic transition systems.
The reversible computation paradigm for RS extends the RS framework  by
allowing  backward computations as well as forward computations; for this 
goal a set of inference rules for rewriting logic has been 
defined in~\cite{10.1007/978-3-319-73359-3_3}
to exactly keep trace of those elements that  dissolve in the next 
computation step.

This paper is an extended version of paper~\cite{BBF19}.
There are several major differences w.r.t. paper cite{BBF19}, as follows:
\begin{itemize}
\item here we present several new examples to illustrate our framework,
and give more detailed explanations
\item we define the bio-simulation of two reaction systems
\item we define an assertion language to specify the properties of reaction systems
\item we show that our assertion language allows to specify properties extending to our
framework the Hennessy-Milner logic
\item we give full proofs of our results
\end{itemize}

\paragraph{Structure of the paper.} Section \ref{sec:rs} 
describes RSs and their semantics (interactive processes).
Section \ref{sec:ccna} describes briefly the c\CNA \ process algebra and
its operational semantics.
Section \ref{sec:trans} defines the embedding of RSs in
c\CNA \ processes and shows some simple examples to illustrate 
it.
%Section \ref{ex:lactose} shows a more complex example taken from
%the literature on RSs and illustrating a \emph{lac} operon.
Section~\ref{Bio-simulation} presents a methodology for verifying formally
properties of reaction systems, presents an assertion language
to express properties and shows the relationship
with the logic of Hennessy and Milner.
Section~\ref{sec:discussion} presents some features and 
advantages of our embedding for the compositionality of RSs.
Finally, Section~\ref{sec:conclusion}
discusses future work, and concludes.

### 2020-03-29 08-conclusion.tex
nuovo commit abs intro concl biblio
nuovo commit

% !TEX root = ./main_RS2L_LNCS.tex

\section{Conclusion}\label{sec:conclusion}

In this paper we have introduced a variant of the {\tt link}-calculus
where prefixes are link chains and no more single links, as it was 
briefly described in the future work section in~\cite{BBB17}.
This variant allowed us to define 
%shown 
an elegant embedding of reaction systems,
an emerging formalism to model computationally biochemical systems.
%into multilink processes.  
This translation shows several benefits.
For instance, the context behaviour can also be expressed recursively. 
Entity mutations can be expressed easily. Reaction systems can
communicate between them.

We have then modified the classical notion (in concurrency) of bisimulation
making it more suitable for proving properties of reaction systems in our
framework.
We have defined a new assertion language, which allows us to specify
the properties to be verified, and have shown that it extends Hennessy-Milner logic
to our framework.

We believe that our embedding can contribute to extend the applications
of reaction systems to diverse fields of computer science, and life
sciences.
As we have already mentioned, the evolution of each process resulting from our embedding 
is deterministic, thus we do not have the problem of having infinitely many transitions
in the produced labelled transition system. In any case, we can exploit the implementation of  the symbolic semantics of the {\tt link}-calculus~\cite{BrodoO17} that can be found in~\cite{tool}.

As future work, we plan to implement a prototype of our framework,
with an automatic translation from RSs to  {\tt link}-calculus. 
We believe that our work can also help to extend the framework
of RSs towards a model which can improve the communication
between different RSs. We also believe that our work can make 
possible to investigate how to apply formal techniques to prove 
properties of the
modeled systems \cite{CFHOT15,OCHF16,BBGLBH2017}.

### 2020-03-30 06-examples.tex
examples + flat fucntion

@@
 \begin{array}{lcl}
-Sys  \defeq \ \restrict{r1,r2,r3,r4,\tilde{q},\tilde{w},\tilde{a},\tilde{b}} (P_q \mid \overline{P}_w  \mid  P_a  \mid \overline{P}_b  \mid  P_1  \mid  P_2  \mid  P_3  \mid P_4  \mid Cxt).
+Sys  \defeq \ %\restrict{r1,r2,r3,r4,\tilde{q},\tilde{w},\tilde{a},\tilde{b}}
+ P_q \mid \overline{P}_w  \mid  P_a  \mid \overline{P}_b  \mid  P_1  \mid  P_2  \mid  P_3  \mid P_4  \mid Cxt.
 \end{array}
@@
 \begin{array}{l}
-\restrict{r1,r2,r3,r4,\tilde{q},\tilde{w},\tilde{a},\tilde{b}}\\
- (\startchain{\silent}\chainedlink{\overline{a}_i}{\bf  \overline{a}_i}\chainedlink{\bf \overline{a}_o}{\overline{a}_o}\chainedlink{r_2}{r_2}\chainedlink{q_i}{\bf q_i}\chainedlink{\bf q_i}{q_i}\chainedlink{b_i}{\bf b_i}\chainedlink{\bf b_i}{b_i}\chainedlink{\overline{w}_i}{\bf \overline{w}_i}\chainedlink{\bf \overline{w}_o}{\overline{w}_o}\chainedlink{\overline{a}_i}{\bf \overline{a}_i}\chainedlink{\bf \overline{a}_o}{\overline{a}_o}\chainedlink{r_3}{r_3}\chainedlink{\overline{a}_i}{\bf \overline{a}_i}\chainedlink{\bf \overline{a}_o}{\overline{a}_o}\chainedlink{r_4}{r_4}\chainedlink{\overline{w}_i}{\bf \overline{w}_i}\chainedlink{\bf \overline{w}_o}{\overline{w}_o}\chainedlink{cxt}{cxt}\chainedlink{\hat{a}_i}{\bf \hat{a}_i}\chainedlink{\bf \hat{a}_o}{\hat{a}_o}\chainedlink{\underline{b}_i}{\bf \underline{b}_i}\chainedlink{\bf \underline{b}_o}{\underline{b}_o}\chainedlink{p_1}{p_1}\chainedlink{p_2}{p_2}\chainedlink{\tilde{w}_i}{\bf \tilde{w}_i}\chainedlink{\bf \tilde{w}_o}{\tilde{w}_o}\chainedlink{p_3}{p_3}\chainedlink{p_4}{p_4}\chainend{\silent})
+%\restrict{r1,r2,r3,r4,\tilde{q},\tilde{w},\tilde{a},\tilde{b}}\\
+ \startchain{\silent}\chainedlink{\overline{a}_i}{\bf  \overline{a}_i}\chainedlink{\bf \overline{a}_o}{\overline{a}_o}\chainedlink{r_2}{r_2}\chainedlink{q_i}{\bf q_i}\chainedlink{\bf q_i}{q_i}\chainedlink{b_i}{\bf b_i}\chainedlink{\bf b_i}{b_i}\chainedlink{\overline{w}_i}{\bf \overline{w}_i}\chainedlink{\bf \overline{w}_o}{\overline{w}_o}\chainedlink{\overline{a}_i}{\bf \overline{a}_i}\chainedlink{\bf \overline{a}_o}{\overline{a}_o}\chainedlink{r_3}{r_3}\chainedlink{\overline{a}_i}{\bf \overline{a}_i}\chainedlink{\bf \overline{a}_o}{\overline{a}_o}\chainedlink{r_4}{r_4}\chainedlink{\overline{w}_i}{\bf \overline{w}_i}\chainedlink{\bf \overline{w}_o}{\overline{w}_o}\chainedlink{cxt}{cxt}\chainedlink{\hat{a}_i}{\bf \hat{a}_i}\chainedlink{\bf \hat{a}_o}{\hat{a}_o}\chainedlink{\underline{b}_i}{\bf \underline{b}_i}\chainedlink{\bf \underline{b}_o}{\underline{b}_o}\chainedlink{p_1}{p_1}\chainedlink{p_2}{p_2}\chainedlink{\tilde{w}_i}{\bf \tilde{w}_i}\chainedlink{\bf \tilde{w}_o}{\tilde{w}_o}\chainedlink{p_3}{p_3}\chainedlink{p_4}{p_4}\chainend{\silent}
 \end{array}
@@
 
-\subsection{A  biological toy example}
+\subsection{A  biological toy example}\label{subsec:toy}
+Hereafter, we only take into account transitions whose labels are solid, i.e. they do not contain 
+virtual links, and   complete, i.e. thy starts and ends with the $\silent$ symbol.\\
+
 Here we consider a biological toy example where a gene $a$  codes for a protein $T$ when molecules 
@@
 \begin{array}{lcl}
-Sys &\defeq &\restrict{\tilde{C},\tilde{G},\tilde{F}, \tilde{T},\tilde{T'}}( P_a\mid P_C\mid P_G\mid \overline{P}_F \mid \overline{P}_T \mid P_{T'} \mid P_1 \mid P_2 \mid P_3 \mid P_4)
+Sys &\defeq &%\restrict{\tilde{C},\tilde{G},\tilde{F}, \tilde{T},\tilde{T'}}(
+P_a\mid \overline{P}_C\mid P_G\mid \overline{P}_F \mid \overline{P}_T \mid P_{T'} \mid P_1 \mid P_2 \mid P_3 \mid P_4
 \end{array}
@@
 In the above configuration, reactions $1$ and $3$ can be applied, and also we assume that the context will provide the entity $F$, that will be available in the target configuration.
-The transition label of
+The transition label appears as:
+\[
+\tiny
+\begin{array}{l}
+%\restrict{r1,r2,r3,r4,a,\tilde{C},\tilde{G},\tilde{F},\tilde{T},\tilde{T'}}\\
+ \startchain{\silent}\chainedlink{a_i}{\bf  a_i}\chainedlink{\bf a_o}{a_o}
+ \chainedlink{G_i}{\bf  G_i}\chainedlink{\bf G_o}{G_o}
+ \chainedlink{\overline{C}_i}{\bf \overline{C}_i}\chainedlink{\bf \overline{C}_o}{\overline{C}_o}
+ \chainedlink{r_2}{r_2}\chainedlink{G_i}{\bf G_i}\chainedlink{\bf G_o}{G_o}
+\chainedlink{r_3}{r_3}\chainedlink{T'_i}{\bf T'_i}\chainedlink{\bf T'_o}{T'_o}\chainedlink{r_4}{r_4}\chainedlink{T'_i}{\bf T'_i}\chainedlink{\bf T'_o}{T'_o}\chainedlink{\overline{F}_i}{\bf \overline{F}_i}\chainedlink{\bf \overline{F}_o}{\overline{F}_o}\chainedlink{cxt}{cxt}
+\chainedlink{\underline{C}_i}{\bf \underline{C}_i}\chainedlink{\bf \underline{C}_o}{\underline{C}_o}
+\chainedlink{\underline{G}_i}{\bf \underline{G}_i}\chainedlink{\bf \underline{G}_o}{\underline{G}_o}
+\chainedlink{\hat{F}_i}{\bf \hat{F}_i}\chainedlink{\bf \hat{F}_o}{\hat{F}_o}\chainedlink{p_1}{p_1} 
+ \chainedlink{\tilde{T}_i}{\bf \tilde{T}_i}\chainedlink{\bf \tilde{T}_o}{\tilde{T}_o}
+  \chainedlink{p_2}{p_2}\chainedlink{p_3}{p_3}\chainedlink{p_4}{p_4} \chainedlink{\tilde{G}_i}{\bf \tilde{G}_i}\chainedlink{\bf \tilde{G}_o}{\tilde{G}_o} \chainend{\silent}
+\end{array}
+\]

### 2020-03-30 00-abstract.tex
revisde abstract

@@
 multiparty interaction in concurrent systems, and show that it allows to
-embed reaction systems, by representing the behaviour of each entity and
+embed reaction systems in a modular way, by representing the behaviour of each entity and
 preserving faithfully their features. 
@@
 %regulatory network. 
-We then modify the classical notion (in concurrency) of bisimulation
-making it more suitable for proving properties of reaction systems, in our
-framework.
-We define a new assertion language, which allows to specify
-the properties to be verified, and show that it extends Hennessy-Milner logic
+On top of the LTS semantics for reaction systems provided by our framework, 
+we then adapt the classical notion (in concurrency) of bisimulation
+to make it more suitable for studying properties of reaction systems.
+In particular, we define a new assertion language based on regular expressions, 
+which allows to specify the properties of interest, and use it to extend Hennessy-Milner logic
 to our framework.
+Interestingly, the novel notion of bisimilarity and logical equivalence that are defined parametrically on some assertion of interest are proved to coincide, like in the classical case.
 Finally, our methodology can contribute to increase the expressiveness

### 2020-03-30 01-introduction.tex
Update 01-introduction.tex

@@
 This formalism allows several processes to synchronise and 
-communicate altogether, at the same time, with a new communicating
+communicate altogether, at the same time, with a new interaction
 mechanism based on links and link chains. 
-Our initial motivation for introducing this mechanism was to encode
+The initial motivation for introducing this mechanism was to encode
 Mobile Ambients~\cite{CardelliG00}, getting a much stronger operational
 correspondence than any available in the literature, such as the one in~\cite{B16}.
- This allowed us to easily encode calculi for biology equipped with membranes, as in~\cite{BodeiBBC14}.
-Process calculi have been used successfully to model
-biological processes, see~\cite{BBDFH18} for a recent survey.
+Later it was shown that the {\tt link}-calculus allowed one to easily encode calculi for biology equipped with membranes, as in~\cite{BodeiBBC14}.
+
 We illustrate our embedding by means of some examples coming
@@
 features of RSs, and prove its correctness and completeness.
+
 Then, we present a methodology for verifying formally
@@
 without getting lost in the complexity of the full biological system, or network.
-Besides bio-simulation, we define a completely new and simple assertion language,
-which allows to specify properties of the reaction systems to be verified.
-The classical logic of Hennessy-Milner~\cite{HM80} is a powerful
+The notion of bio-simulation relies on a new and simple assertion language,
+which allows to focus on some properties of the reaction systems to be verified.
+In fact, bio-similarity is parametric on a given assertion of interest.
+The we prove that the well-known logical characterization of bisimilarity in terms of Hennessy-Milner Logic~\cite{HM80} (HML)
+can be extended to the case of bio-similarity by tailoring HML formulas to
+the same assertion of interest used in bio-similarity.  As HML is a powerful
 formalism for specifying properties of labeled transition systems, 
-and we show that our assertions allow us to specify formulas 
-that extend this logic to the case of complex labels.
-Our formulas allow to abstract from some of the labels.
+its variant considered here looks a suitable option to specify formulas 
+over complex labels by abstracting from unnecessary details.
 
@@
 specified in a recursive way as an ordinary process;
-\item we can express the behaviour of entity mutation,
+\item along the same lines, non-deterministic contexts can be seamlessly expressed in our setting; 
+\item when deterministic contexts are considered, the c\CNA \
+computation is deterministic.
+\item we define an assertion language to specify local properties of 
+reaction systems and exploit it to define a novel notion of behavioural equivalence for reaction systems, called bio-similarity;
+\item we show that our assertions extend can be used to extend the Hennessy-Milner logic to our framework, in such a way that logical equivalence of processes coincide with bio-similarity;
+\item we sketch how one can express the behaviour of entity mutation,
 %change the behaviour of  each single entity $s$  in a
@@
 $s'$ can  take part to only a subset of rules requiring entity $s$;
-\item with a little coding effort, two RSs can
+\item we show that with a little coding effort, two RSs can
 communicate; i.e.  a subset of those entities that the context  can
 provide, are then provided by a second RS;
-\item as our translation results in  a  c\CNA \ system, from
-each state only one transition can be generated, thus the c\CNA \
-computation is fully deterministic.
-\item we define an assertion language to specify local properties of 
-reaction systems
-\item we show that our assertions extend to our framework the Hennessy-Milner logic.
 \end{itemize}
@@
 the interaction between RSs can help to model new scenarios.
+
 \paragraph{Related work}
+Process calculi have been used successfully to model
+biological processes, see~\cite{BBDFH18} for a recent survey.
 We are not aware of any inference rules for the RS; it seems not the case

### 2020-03-30 01-introduction.tex
Update 01-introduction.tex

@@
 communicate; i.e.  a subset of those entities that the context  can
-provide, are then provided by a second RS;
+provide, are then provided by a second RS.
 \end{itemize}
@@
 
-\paragraph{Related work}
+\paragraph{Related work.}
+
 Process calculi have been used successfully to model
@@
 
-This paper is an extended version of paper~\cite{BBF19}.
-There are several major differences w.r.t. paper cite{BBF19}, as follows:
+A preliminary version of this paper appeared in~\cite{BBF19}.
+There are several major differences w.r.t. the paper~\cite{BBF19}, which is here extended as follows:
 \begin{itemize}
 \item here we present several new examples to illustrate our framework,
-and give more detailed explanations
-\item we define the bio-simulation of two reaction systems
-\item we define an assertion language to specify the properties of reaction systems
+and give more detailed explanations about its use;
+\item we introduce an assertion language to specify the properties of reaction systems;
+\item we define the notion of bio-simulation to relate reaction systems and compare their behaviours;
 \item we show that our assertion language allows to specify properties extending to our
-framework the Hennessy-Milner logic
-\item we give full proofs of our results
+framework the Hennessy-Milner logic;
+\item proofs of main results are included here.
 \end{itemize}
 
-\paragraph{Structure of the paper.} Section \ref{sec:rs} 
+\paragraph{Structure of the paper.}
+
+Section \ref{sec:rs} 
 describes RSs and their semantics (interactive processes).
@@
 %the literature on RSs and illustrating a \emph{lac} operon.
-Section~\ref{Bio-simulation} presents a methodology for verifying formally
-properties of reaction systems, presents an assertion language
-to express properties and shows the relationship
-with the logic of Hennessy and Milner.
+Section~\ref{sec:biosimulation} presents a methodology for the formal verification of
+properties of reaction systems that are expressed in a novel assertion language.
 Section~\ref{sec:discussion} presents some features and 

### 2020-03-30 02-ReactionSystem.tex
Update 02-ReactionSystem.tex
cambiato leggermente notazione e corretto qualche typo
(W invece di T, \mathit{en}, \mathit{res}, \pi^\infty)

% !TEX root = ./main_RS2L_LNCS.tex

\section{Reaction Systems}
\label{sec:rs}

Natural Computing is concerned with human-designed computing inspired by 
nature as well as with computation taking place in nature.
The theory of Reaction Systems~\cite{BEMR11} 
was born in the field of Natural Computing 
to model the behaviour of biochemical reactions taking place in living cells. 
Despite its initial aim, this formalism has shown to be quite useful 
not only for modeling biological phenomena, but also for
the contributions which is giving to computer science~\cite{MPR15}, 
theory of computing, 
mathematics, biology~\cite{ABP14,CMMBM12,Az17,BarbutiGLM16}, 
and molecular chemistry~\cite{OY16}.
%
Here we briefly review the basic notions of RSs, see~\cite{BEMR11} for more details.

The mechanisms that are at the basis of biochemical reactions and thus 
regulate the functioning of a living cell, are 
{\em facilitation} and {\em inhibition}. These mechanisms are 
reflected in the basic definitions of Reaction Systems.

\begin{definition}[Reaction]
A reaction is a triplet $a = (R,I,P)$, where $R, I, P$ are finite,  
non empty sets  and
$R \cap I = \emptyset$. If $S$ is a set such that  $R, I, P \subseteq S$, 
then $a$ is a reaction in $S$.
\end{definition}

The sets $R, I, P$ are also written $R_a, I_a, P_a$ and called the 
\emph{reactant set} of $a$, the  
\emph{inhibitor set} of $a$, and the \emph{product set} of $a$, respectively. 
All reactants are needed for the reaction to take place.
Any inhibitor blocks the reaction if it is present. Products are the outcome of the reaction.
Also,  $R_a \cup I_a$ is the set of the resources of $a$ and $ rac(S)$
denotes the set of all reactions in $S$.
%
Because $R$ and $I$ are non empty, all products are produced from at least one reactant and every reaction can be inhibited in some way. 
Sometimes artificial inhibitors are used that are never produced by any reaction.
For the sake of simplicity, in some examples, we will allow $I$ to be empty.
%Note that since $R_a$ and $I_a$ are both nonempty, we always 
%have $|M_a| \geq 2$.

\begin{definition}[Reaction System]
A Reaction System (RS) is an ordered pair ${\cal A} = (S, A)$ 
such that $S$ is a finite set, and $A \subseteq rac(S)$.
%The set $S$ is called the background (set) of $A$.\\
%The main notion concerning the functioning of a reaction system is defined 
%as follows.
\end{definition}

The set $S$ is called the \emph{background set} of ${\cal A}$, its elements are called \emph{entities},
they represent molecular substances (e.g., atoms, ions, molecules) that may be present
in the states of a biochemical system.
% modeled by $A$. 
The set $A$ is the set of \emph{reactions} of $\mathcal{A}$.
Since $S$ is finite, so is $A$: 
we denote by $|A|$ the number of reactions in $A$.

\begin{definition}[Reaction Result]
Let $W$ be a finite subset of $S$.
\begin{enumerate}
\item Let $a$ be a reaction in $S$. Then $a$ is enabled by $W$, denoted by $\mathit{en}_a(W)$, 
if $R_a \subseteq W$
and $I_a \cap W = \emptyset$. The result of $a$ on $W$, denoted by $\mathit{res}_a(W)$, 
is defined by:
$\mathit{res}_a(W) =P_a$, if $\mathit{en}_a(W)$, and $\mathit{res}_a(W) = \emptyset$ otherwise.
\item Let $A$ be a finite set of reactions. The result of $A$ on $W$, denoted 
by $\mathit{res}_A(W)$,  is defined
by: $\mathit{res}_A(W) = \bigcup_{a \in A} \mathit{res}_a(W)$.
\end{enumerate}
\end{definition}

%\medskip
%\noindent
%{\bf Important Assumptions}\\
The theory of Reaction Systems is based on the following assumptions.
\begin{itemize}
\item {\bf No permanency.} An entity of a set $W$ vanishes unless it is sustained 
by a reaction. This reflects the fact that a living cell would die for lack 
of energy, without chemical reactions.
\item {\bf No counting.} The basic model of RSs is very abstract and 
qualitative, i.e. the quantity of entities that are present 
in a cell is not taken into account. 
%It is just analyzed the behavior of 
%the system based on the reactions that take place.
\item {\bf Threshold nature of resources.} From the 
previous item,
we assume that either an entity is available and there is enough of it 
(i.e. there are no conflicts), or it is not available at all.
\end{itemize}

The dynamic behaviour of a RS is formalized in terms of 
{\em interactive processes}.

\begin{definition}[Interactive Process]
Let ${\cal A} = (S, A)$ be a RS and let $n \geq 0$. 
An $n$-step
\emph{interactive process} in ${\cal A}$ is a pair $\pi = (\gamma, \delta)$ of 
finite sequences s.t.
%
%\begin{center}
$ \gamma=\{C_i\}_{i\in[0,n]}$  and $\delta=\{D_i\}_{i\in[0,n]} $
%$\gamma = C_0, \dots , C_n$ and $\delta = D_0, \dots , D_n$, 
%\end{center}
%\noindent
where 
$ C_{i}, D_{i}  \subseteq S$ for any $i\in[0,n]$, $D_0 = \emptyset$, and 
%$C_0,\dots, C_n, D_0,\dots, D_n  \subseteq S$, $D_0 = \emptyset$, and 
$D_i =  \mathit{res}_{A}(D_{i-1} \cup C_{i-1})$ for any $i \in [1,n]$.
% $i \in \{1,\dots, n\}$.
\end{definition}

Living cells are seen as open systems that continuously react with 
the external environment, in discrete steps. 
The sequence $\gamma$ is the {\em context sequence} of $\pi$ and 
represents the influence of the environment on the Reaction System.
The sequence $\delta$ is
the {\em result sequence} of $\pi$ and it is entirely determined by $\gamma$ and $A$.
%
The sequence $\tau = W_{0}, \ldots, W_{n}$ with $W_{i} = C_{i} \cup D_{i}$, 
for any $i \in [0, n]$
%, and $D_{0} = \emptyset$ 
is called a \emph{state sequence}.
%
Each state $W_{i}$ in a state sequence
is the union of two sets: the context $C_{i}$
at step $i$ and the result $D_i=\mathit{res}_{A}(W_{i-1})$ from the previous step.

%\begin{definition}[State Sequence]
%\label{def:statesequence}
% A {\em state sequence} $\tau$ of an interactive process $\pi$:
%
%\begin{center}
%
%\end{center}
%\end{definition}

%\begin{definition} 
%A process $\pi$, or a state 
%sequence $\tau = W_{0}, \ldots, W_{n}$, is said to be 
%{\em context-independent} when $C_{i} 
%\subseteq D_{i}$, $\forall i \in \{i, \ldots, n \}$.
  %\end{definition}

For technical reasons, we extend the notion of an interactive process to deal with 
infinite sequences.

\begin{definition}[extended interactive process]
Let ${\cal A}=(S,A)$ be a RS, and let $\pi =(\gamma,\delta)$  be an  $n$-step interactive process,
with $ \gamma=\{C_i\}_{i\in[0,n]}$ and  $\delta=\{D_i\}_{i\in[0,n]}$
%, $\delta= D_0, \dots D_n$.
Then, we let $\pi^\infty= (\gamma^\infty,\delta^\infty)$ be the extended interactive process of  $\pi$,
defined as $  \gamma^\infty=\{C_i'\}_{i\in\mathbb{N}}$, $\delta^\infty=\{D_i'\}_{i\in\mathbb{N}}$, where:
$$
C'_j = \left\{
\begin{array}{ll}
C_j & \mbox{if $j\in[0,n]$}\\
\emptyset & \mbox{otherwise ($j>n$)}
\end{array}
\right.
\qquad
D'_j = \left\{
\begin{array}{ll}
D_0 & \mbox{if $j=0$}\\
\mathit{res}_A(D_{j-1}' \cup C_{j-1}') & \mbox{otherwise ($j\geq 1$)}
\end{array}
\right.
$$
% $C_j' = C_j$ for $j\in[0,n]$ and $C_j' =\emptyset$ for $j>n$,   $D_0' = D_0$ and  $D_j' = \mathit{res}_A(D_{j-1}' \cup C_{j-1}')$ for $j\geq 1$.
\end{definition}

### 2020-03-30 04-CCNA.tex
Update 04-CCNA.tex

@@
 %
-For example, given $v_1 = \startchain{a}\chainedlink{b}{\noact}\chainedlink{\noact}{\noact}\chainend{\noact}$ and $v_2 = \startchain{\noact}\chainedlink{\noact}{b}\chainedlink{c}{\noact}\chainend{\noact}$ we have $\merges{v_1}{v_2} = \startchain{a}\chainedlink{b}{b}\chainedlink{c}{\noact}\chainend{\noact}$.
+For example, given $v_1 = \startchain{\silent}\chainedlink{a}{\noact}\chainedlink{\noact}{\noact}\chainend{\noact}$, $v_2 = \startchain{\noact}\chainedlink{\noact}{a}\chainedlink{b}{\noact}\chainend{\noact}$ and $v = \startchain{\silent}\chainedlink{a}{a}\chainedlink{b}{\noact}\chainend{\noact}$
+we have $\merges{v_1}{v_2} = v$, whereas $\merges{v_1}{v}$ is not defined. Notably the merge operation is commutative and associative.
 
@@
 %
-For example, given $v = \startchain{a}\chainedlink{b}{b}\chainedlink{c}{\noact}\chainend{\noact}$ we have 
-$\restrict{b}v = \startchain{a}\chainedlink{\silent}{\silent}\chainedlink{c}{\noact}\chainend{\noact}$.
+For example, given $v = \startchain{\silent}\chainedlink{a}{a}\chainedlink{b}{\noact}\chainend{\noact}$ as above, we have 
+$\restrict{a}v = \startchain{\silent}\chainedlink{\silent}{\silent}\chainedlink{b}{\noact}\chainend{\noact}$,
+whereas $\restrict{b}v$ is not defined.
 
@@
 %\begin{definition}
-The c\CNA\ processes, denoted as $\cal{P}$, are generated by the following grammar:
+The set of c\CNA\ processes, denoted as $\cal{P}$ and ranged over by $P,Q$, is defined by the following grammar:
 \[
@@
 The syntax of c\CNA\  extends that of \CNA~\cite{BBB17} by allowing to use link chains as prefixes instead of links.
-For the rest it features nondeterministic choice, parallel composition, restriction, relabelling and possibly recursive definitions.
+For the rest it features nondeterministic (guarded) choice, parallel composition, restriction, relabelling and possibly recursive definitions.
 Here we do not consider name mobility, which is present instead in the {\tt link}-calculus.
@@
 synchronises interactions (it cannot be applied when $\merges{v}{v'}$ is not defined).
+
+\begin{example}
+As some simple examples, consider the recursive process definitions $H \defeq \link{a}{b}.\link{a}{c}.H$ and $K \defeq \link{a}{b}.K + \link{a}{c}.K$: the former is some sort of forwarder that alternates between providing a link from $a$ to $b$ and to $c$ ; the latter is a nondeterministic forwarder that at each step can link $a$ to $b$ or to $c$.
+\end{example}
 
@@
 the presentation of our translation.
-%First, we define the juxtaposition of link or link chains as
+
 \begin{definition}[Replication]
-Let $\upsilon$ be a link chain. Its $n$ times replication $\upsilon^n$ is defined recursively by letting
-$\upsilon^0 = \epsilon$ (i.e. the empty chain) and $\upsilon^n = \upsilon^{n-1}\upsilon$,
+Let $\upsilon$ be a link chain such that $\upsilon\upsilon$ is also a valid link chain. Its $n$ times replication $\upsilon^n$ is defined recursively by letting
+$\upsilon^0 = \epsilon$ (i.e. the empty chain) and $\upsilon^n = \upsilon\upsilon^{n-1}$.
 %\[
@@
 %\]
-with the hypothesis that all the links in the resulting link chains match.
+%with the hypothesis that all the links in the resulting link chains match.
 %, and that $\upsilon^0 = \epsilon$ is the empty chain.
@@
 For example,  the expression $(\startchain{a}\chainedlink{b}{\noact}\chainend{\noact})^3$  denotes the chain $\startchain{a}\chainedlink{b}{\noact}\chainedlink{\noact}{a}\chainedlink{b}{\noact}\chainedlink{\noact}{a}\chainedlink{b}{\noact}\chainend{\noact}$.
+Instead the expression $(\link{a}{b})^2$ is ill-defined because $a$ does not match with $b$.
 
- \noindent
-We introduce the \emph{half link} that will be used in conjunction with the \emph{open block of chain} to form regular link chains.
+Then, we introduce the notation for \emph{half links} that will be used in conjunction with the \emph{open block of chain} to form regular link chains.
 
 \begin{definition}[Half links]
-Let $a$ be a channel name, we define the \emph{half left link}:  $\link{a}{}$,   and the  \emph{half right link}: $\link{}{a}$.
+Let $a$ be a channel name, we define the \emph{half left link}  $\link{a}{}$,   and the  \emph{half right link} $\link{}{a}$.
 \end{definition}

### 2020-03-30 04-CCNA.tex
fino a sect4 e flat
update fino a sezione 4 inclusa e def di flat

@@
 \begin{example}
-As some simple examples, consider the recursive process definitions $H \defeq \link{a}{b}.\link{a}{c}.H$ and $K \defeq \link{a}{b}.K + \link{a}{c}.K$: the former is some sort of forwarder that alternates between providing a link from $a$ to $b$ and to $c$ ; the latter is a nondeterministic forwarder that at each step can link $a$ to $b$ or to $c$.
+As some simple examples, consider the recursive process definitions $H \defeq \link{a}{b}.\link{a}{c}.H$ and $K \defeq \link{a}{b}.K + \link{a}{c}.K$: the former is some sort of forwarder that alternates between providing a link from $a$ to $b$ and to $c$; the latter is a nondeterministic forwarder that at each step can link $a$ to $b$ or to $c$.
 \end{example}

### 2020-03-30 05-translation.tex
fino a sect4 e flat
update fino a sezione 4 inclusa e def di flat

@@
 Given an entity $s \in S$, we exploit four different names for the interactions over $s$:
-names $s_i$, $s_o$ are used to test the presence of $s$ in the system; names $\hat{s}_i$, $\hat{s}_o$ are used to test the provisioning of $s$ from the context;
-names $\tilde{s}_i$, $\tilde{s}_o$  are used to test the production of $s$ by some reaction; names $\overline{s}_i$, $\overline{s}_o$ are used to test the absence of $s$ from the context; and names $\underline{s}_i$, $\underline{s}_o$ are used to test the absence of $s$ in the system.
+names $s_i$, $s_o$ are used to test the presence of $s$ in the system; names $\widehat{s}_i$, $\widehat{s}_o$ are used to test the provisioning of $s$ from the context;
+names $\widetilde{s}_i$, $\widetilde{s}_o$  are used to test the production of $s$ by some reaction; names $\overline{s}_i$, $\overline{s}_o$ are used to test the absence of $s$ in the system; and names $\underline{s}_i$, $\underline{s}_o$ are used to test the absence of $s$ from the context.
 We let $P_s$ be the process implementing the  presence of $s$ in the system, and $\overline{P_s}$ be the one for its absence.
@@
 \begin{array}{lcl@{\hskip 1cm}lcl}
- P_s &\defeq & P(s,\tilde{s},\hat{s},\underline{s}) 
+ P_s &\defeq & P(s,\widetilde{s},\widehat{s},\underline{s}) 
  &
- \overline{P_s}& \defeq & P(\overline{s},\tilde{s},\hat{s},\underline{s})
+ \overline{P_s}& \defeq & P(\overline{s},\widetilde{s},\widehat{s},\underline{s})
 \end{array}
@@
 \begin{array}{rcl}
- P(s,\tilde{s},\hat{s},\underline{s}) & \defeq &\sum_{h,k\geq 0}  
-  (\startchain{s_i}\chainedlink{s_o}{\noact}\chainend{\noact})^h \ \startchain{\hat{s}_i}\chainedlink{\hat{s}_o}{\noact}\chainend{\noact} (\startchain{\tilde{s}_i}\chainedlink{\tilde{s}_o}{\noact}\chainend{\noact})^k.P_s\\
+ P(s,\widetilde{s},\widehat{s},\underline{s}) & \defeq &\sum_{h,k\geq 0}  
+  (\startchain{s_i}\chainedlink{s_o}{\noact}\chainend{\noact})^h \ \startchain{\widehat{s}_i}\chainedlink{\widehat{s}_o}{\noact}\chainend{\noact} (\startchain{\widetilde{s}_i}\chainedlink{\widetilde{s}_o}{\noact}\chainend{\noact})^k.P_s\\
 &&\ + \\
@@
 \sum_{h\geq 0,k\geq 1}
- (\startchain{s_i}\chainedlink{s_o}{\noact}\chainend{\noact})^h \ \startchain{\underline{s}_i}\chainedlink{\underline{s}_o}{\noact}\chainend{\noact}(\startchain{\tilde{s}_i}\chainedlink{\tilde{s}_o}{\noact}\chainend{\noact})^k.P_s\\
+ (\startchain{s_i}\chainedlink{s_o}{\noact}\chainend{\noact})^h \ \startchain{\underline{s}_i}\chainedlink{\underline{s}_o}{\noact}\chainend{\noact}(\startchain{\widetilde{s}_i}\chainedlink{\widetilde{s}_o}{\noact}\chainend{\noact})^k.P_s\\
 &&\ +\\
@@
 \]
-The first line of $ P(s,\tilde{s},\hat{s},\underline{s})$ accounts for the case where $s$ is tested for presence by $h$ reactions and produced by $k$ reactions, while being provided by the context ($\link{\hat{s}_i}{\hat{s}_o}$).
+The first line of $ P(s,\widetilde{s},\widehat{s},\underline{s})$ accounts for the case where $s$ is tested for presence by $h$ reactions and produced by $k$ reactions, while being provided by the context ($\link{\widehat{s}_i}{\widehat{s}_o}$).
 Thus, $s$ will be present at the next step (the continuation is $P_s$). Here $h$ and $k$ are not known a priori and therefore any
@@
 The process for reaction $aj=(R_j,I_j,P_j)$ must assert either the possibility to apply the reaction or its impossibility.
-The first case happens when all its reactants are present (the link $\link{s_i}{s_o}$ is requested for any $s\in R_j$) and all its inhibitors are absent (the link $\link{\overline{e}_i}{\overline{e}_o}$ is requested for any $e\in I_j$), then the product set is released (the link $\link{\tilde{c}_i}{\tilde{c}_o}$ is requested for any $c\in P_j$).
+The first case happens when all its reactants are present (the link $\link{s_i}{s_o}$ is requested for any $s\in R_j$) and all its inhibitors are absent (the link $\link{\overline{e}_i}{\overline{e}_o}$ is requested for any $e\in I_j$), then the product set is released (the link $\link{\widetilde{c}_i}{\widetilde{c}_o}$ is requested for any $c\in P_j$).
 The next case can happen for two reasons: one of the reactants is absent (the link $\link{\overline{s}_i}{\overline{s}_o}$ is requested for some $s\in R_j$) or one of the inhibitors is present (the link $\link{e_i}{e_o}$ is requested for some $e\in I_j$).
@@
 P_{aj}& \defeq & 
- ^{r_j} \backslash  \blockchain{s_i}{s_o}{s \in R_j} \, \link{}{} \blockchain{\overline{e}_i}{\overline{e}_o}{e \in I_j} \, \chainedlink{r_{j+1}}{\noact}\chainedlink{\noact}{p_j}\link{}{} \blockchain{\tilde{c}_i}{\tilde{c}_o}{c \in P_j}\, \backslash_{p_{j+1}}.P_{aj} & \mbox{\small  \{$aj$ is applicable\}}\\
+ ^{r_j} \backslash  \blockchain{s_i}{s_o}{s \in R_j} \, \link{}{} \blockchain{\overline{e}_i}{\overline{e}_o}{e \in I_j} \, \chainedlink{r_{j+1}}{\noact}\chainedlink{\noact}{p_j}\link{}{} \blockchain{\widetilde{c}_i}{\widetilde{c}_o}{c \in P_j}\, \backslash_{p_{j+1}}.P_{aj} & \mbox{\small  \{$aj$ is applicable\}}\\
 && + \ \\
@@
 Channels $p_i$ and $p_{j+1}$ enclose the links related to the entities produced by $aj$.
-We will see that all the link chain labels of transitions follow the same schema: first we find all the reactions limited to the reactants and inhibitors (chained using $r_j$ channels), then all the supplies by the contexts (chained using $cxt_j$ channels, to be introduced next), and finally the products for all the reactions (chained using $p_j$ channels).
+We will see that all the link chain labels of transitions follow the same schema: first we find all the reactions limited to the reactants and inhibitors (chained using $r_j$ channels), then all the supplies by the contexts (chained using the channel $\mathit{cxt}$, to be introduced next), and finally the products for all the reactions (chained using $p_j$ channels).
 In the following there is an example explaining this schema.
@@
 \paragraph{Processes for contexts.}
-For each entity $s \in S$, we introduce another process $Cxt_s$,  participating in each transition and saying if, the entity $s$ is provided by the context or not.
-As done for the reactions, we assume that entities are enumerated and use the names $cxt_j$ to concatenate
-the chains formed by the application of all the contexts. 
-For each entity $s$ with number $j$, at step $n >0$  there are two possible behaviours: 
+The context can vary at each step $n$ and for each entity $s \in S$, the context must say if the entity is provided or not.
+Let us denote by $C_n$ the context at step $n$, i.e.
+the set of entities that are made available.
+Correspondingly, we introduce another process $\mathit{Cxt}^n$ defined as follows:
 \[
 \begin{array}{lcl}
-Cxt^n &\defeq &
+\mathit{Cxt}^n &\defeq &
 %\left \{
-\begin{array}{ll}
-\sum _{j}^{cxt} \backslash  \blockchain{s_i}{s_o}{s \in Ij_n} \, \link{}{} \blockchain{\overline{e}_i}{\overline{e}_o}{e \in Oj_n} \,  \backslash_{p_1}. Cxt^{n+1}\\
+^{cxt} \backslash  \blockchain{s_i}{s_o}{s \in C_n} \, \link{}{} \blockchain{\overline{e}_i}{\overline{e}_o}{e \not\in C_n} \,  \backslash_{p_1}. \mathit{Cxt}^{n+1}\\
 %+ \\
@@
 %Cxt_s & \defeq & Cxt_s^1\\
-\end{array}
+%\end{array}
 %\right.\\
@@
 \]
+
+%,  participating in each transition and saying if, the entity $s$ is provided by the context or not.
+%As done for the reactions, we assume that entities are enumerated and use the names $cxt_j$ to concatenate
+%the chains formed by the application of all the contexts. 
+%For each entity $s$ with number $j$, at step $n >0$  there are two possible behaviours:
 
@@
 
-We only consider $Cxt^n$ with $n >0$, as the entities  that are present at step zero  are considered to be present in the initial system (process $P_s$ instead of $\overline{P_s}$).
+We only consider $\mathit{Cxt}^n$ with $n >0$, as the entities  that are present at step zero  are considered to be present in the initial system (if $s\in C_0$ the process $P_s$ will be present initially, otherwise $\overline{P_s}$ will be present).
 
@@
 We define its c\CNA~translation $\llbracket{\cal A},\gamma \rrbracket$ as follows: 
-\begin{center}$\llbracket{\cal A},\gamma \rrbracket = \restrict{reacts,ctxs,ents,prods}(\Pi_{s\in C_0}P_{s} | \Pi_{s \notin C_0}\overline{P_s} | \Pi_{a\in A} P_a| \Pi_{s\in S} Cxt_s)$,
-\end{center} 
-with  $reacts$ be the set of reaction names $r_j $, $cxts$ the set of context names $cxt_j$,  $ents$ the set of decorated entity names  $\{s_i,s_o,\hat{s}_i,\hat{s}_o,\tilde{s}_i,\tilde{s}_o,\overline{s}_i,\overline{s}_o,\underline{s}_i,\underline{s}_o| s \in S\}$, and  $prods$ be the set of names $p_j$ associated to each reaction.
-In the following, we set $names = reacts\, \cup\, ctxs\, \cup\, ents\, \cup \, prods$.
+$$
+\llbracket{\cal A},\gamma \rrbracket 
+\triangleq
+ \restrict{\mathit{names}}(\Pi_{s\in C_0}P_{s} ~|~ \Pi_{s \notin C_0}\overline{P_s} ~|~ \Pi_{a\in A} P_a ~|~  \mathit{Cxt}^1)
+$$
+where $\mathit{names} = \mathit{reacts}\, \cup\, \mathit{ctxs}\, \cup\, \mathit{ents}\, \cup \, \mathit{prods}$, with  $\mathit{reacts}$ be the set of reaction names $r_j $, $\mathit{cxts}$ the set of context names $\{\mathit{cxt}\}$,  $\mathit{ents}$ the set of decorated entity names  $\{s_i,s_o,\widehat{s}_i,\widehat{s}_o,\widetilde{s}_i,\widetilde{s}_o,\overline{s}_i,\overline{s}_o,\underline{s}_i,\underline{s}_o| s \in S\}$, and  $\mathit{prods}$ be the set of names for production $p_j$ associated to each reaction.
 %For notational convenience, we fix that $r_1 = \silent$,  $r_{u+1} = cxt_1$ for $u$ the number of reacts, and $cxt_{w+1} = \silent$ for $w$ the number of entities.
-For notational convenience, we fix that $r_1 = \silent$,  $r_{u+1} = cxt_1$ for $u$ the number of reacts, and $cxt_{w+1} = p_1$ $p_{u+1}=\silent$ for $w$ the number of entities.
+For notational convenience, we fix that $r_1 = \silent$,  $r_{u+1} = \mathit{cxt}$ and $p_{u+1}=\silent$  for $u$ the number of reacts.
 \end{definition}
@@
 
-It is important to observe that,   for each transition, our c\CNA \ encoding requires  all the processes $P_a$, with $a \in A$, and  $Cxt_s$ and  $P_s$, with $s \in S$,  be interacting in that transition.
-This is due to the fact that  all the channels $r_j$, $p_j$, $cxt_h$, and $s_{hi}$, and $s_{ho}$ are  restricted.
+It is important to observe that,   for each transition, our c\CNA \ encoding requires  all the processes 
+%$P_a$, with $a \in A$, and  $\mathit{Cxt}^n$ and  $P_s$, with $s \in S$,  
+running in parallel to  interact in that transition.
+This is due to the fact that  all the channels $r_j$, $p_j$, $\mathit{cxt}$, including those for decorated names $s_{i},s_{o},\overline{s}_i,\overline{s}_o$ are  restricted.
 Each reaction defines a pattern to be satisfied, i.e. each reaction inserts as many virtual links  as  the number of reactants, inhibitors, and products, as required by the corresponding reaction.
@@
 Let $P=\llbracket{\cal A},\gamma\rrbracket$ its c\CNA~translation.
-If $exists$ $P'$ such that $t= (P \xrightarrow{\restrict{names}\upsilon} P')$ is a transition of $P$, then
+If $exists$ $P'$ such that $t= (P \xrightarrow{\restrict{\mathit{names}}\upsilon} P')$ is a transition of $P$, then
   \begin{enumerate}
-  \item for each reaction  $a_j \in A$, the corresponding channels $r_j$ and $p_j$ appear in  $\upsilon$; for each entity $s_h \in S$ (where $h$ is the identifying number of $s$), the corresponding channel $s_h$ (suitably decorated), and the corresponding channel $cxt_h$ appear in $\upsilon$;
+  \item for each reaction  $a_j \in A$, the corresponding channels $r_j$ and $p_j$ appear in  $\upsilon$; for each entity $s \in S$, the corresponding channel $s$ (suitably decorated) appear in $\upsilon$;
+  the channel $\mathit{cxt}$ appears in $\upsilon$;
   %processes $P_s$ (or $\overline{P_s}$) 
@@
  %involved in the transition $t$;
-  \item for each reaction $a \in A$ and each entity $s \in S$, each virtual link offered by processes $P_a$ and $Cxt_s$ is  overlapped by exactly one solid link offered by  processes representing entities.
+  \item for each reaction $a \in A$ and each entity $s \in S$, each virtual link offered by processes $P_a$ and $\mathit{Cxt}^n$ is  overlapped by exactly one solid link offered by  processes representing entities.
   %each process $P_s $ (or its alternative version $\overline{P}_s$), with $s \in S$, participates at most  with one solid link for each process $P_a$, and with exactly one solid link for each process $Cxt_s$.%, if $s$ is a \emph{required element} of $A_j$ in $t$.
@@
 $\gamma^j=\{C_i'\}_{i\in\mathbb{N}}$, $\delta^j=\{D_i'\}_{i\in\mathbb{N}}$ with 
-  $C_0'=C_j \cup D_j$, and   $C_i'=C_{i+j}$, $D_i'= D_{i+j}$ for any $i\geq 1$.
+  $C_0'=C_j \cup D_j$, $D'_0 = \emptyset$, and   $C_i'=C_{i+j}$, $D_i'= D_{i+j}$ for any $i\geq 1$.
 
@@
  Let 
- $P=\llbracket{\cal A},\gamma\rrbracket$ with\\
-$P  = \restrict{names} (\Pi_{a \in A}P_a| \Pi_{s \in S}Cxt_s | \Pi_{s \in C_0}P_s| \Pi_{s \notin C_0}\overline{P}_s)$.\\
- If there exists  $P'$  such that $P \xrightarrow{v}P'$,  it holds that:\\
-  $v=\link{\silent}{\silent}\dots\link{\silent}{\silent}$ and \\
- $P'= \restrict{names} (\Pi_{a \in A}P_a| \Pi_{s \in S}Cxt_s | \Pi_{s \in C_1 \cup D_1}P_s| \Pi_{s \notin C_1 \cup D_1}\overline{P}_s)$.\\
+ $P=\llbracket{\cal A},\gamma\rrbracket$ with
+$$P  = \restrict{\mathit{names}} (\Pi_{a \in A}P_a ~|~ \mathit{Cxt}^1 ~|~ \Pi_{s \in C_0}P_s ~|~ \Pi_{s \notin C_0}\overline{P}_s).
+$$
+ If there exists  $P'$  such that $P \xrightarrow{v}P'$,  it holds that:
+ \begin{enumerate}
+ \item
+  $v=\link{\silent}{\silent}\dots\link{\silent}{\silent}$, and
+  \item 
+ $P'= \restrict{\mathit{names}} (\Pi_{a \in A}P_a~|~ \mathit{Cxt}^2 ~|~ \Pi_{s \in C_1 \cup D_1}P_s| \Pi_{s \notin C_1 \cup D_1}\overline{P}_s)$.
+ \end{enumerate}
 Moreover, given $\pi^1=(\gamma^1,\delta^1)$, we have $P' = \llbracket {\cal A},\gamma^1\rrbracket$.
@@
 
-\noindent
 With the following propositions, we prove that, given a RS ${\cal A} = (S, A)$ and an extended  interactive process $\pi=(\gamma,\delta)$, then the $c\CNA$ process $\llbracket{\cal A},\gamma\rrbracket$ can simulate all the evolutions  of $\pi$.

### 2020-03-31 05-translation.tex
commit

@@
 $$
-where $\mathit{names} = \mathit{reacts}\, \cup\, \mathit{ctxs}\, \cup\, \mathit{ents}\, \cup \, \mathit{prods}$, with  $\mathit{reacts}$ be the set of reaction names $r_j $, $\mathit{cxts}$ the set of context names $\{\mathit{cxt}\}$,  $\mathit{ents}$ the set of decorated entity names  $\{s_i,s_o,\widehat{s}_i,\widehat{s}_o,\widetilde{s}_i,\widetilde{s}_o,\overline{s}_i,\overline{s}_o,\underline{s}_i,\underline{s}_o| s \in S\}$, and  $\mathit{prods}$ be the set of names for production $p_j$ associated to each reaction.
+where $\mathit{names} = \mathit{reacts}\, \cup\, \mathit{ents}\, \cup \, \mathit{prods}\, \cup\, \{cxt\}$, with  $\mathit{reacts}$ be the set of reaction names $r_j $, $\mathit{cxts}$ the set of context names $\{\mathit{cxt}\}$,  $\mathit{ents}$ the set of decorated entity names  $\{s_i,s_o,\widehat{s}_i,\widehat{s}_o,\widetilde{s}_i,\widetilde{s}_o,\overline{s}_i,\overline{s}_o,\underline{s}_i,\underline{s}_o| s \in S\}$, and  $\mathit{prods}$ be the set of names for production $p_j$ associated to each reaction.
 %For notational convenience, we fix that $r_1 = \silent$,  $r_{u+1} = cxt_1$ for $u$ the number of reacts, and $cxt_{w+1} = \silent$ for $w$ the number of entities.
-For notational convenience, we fix that $r_1 = \silent$,  $r_{u+1} = \mathit{cxt}$ and $p_{u+1}=\silent$  for $u$ the number of reacts.
+{\color{red} For technical reasons, we introduce the process $Init \defeq \link{\silent}{r_1}.Init$
+to allow the name $r_1$ to be matched; also, 
+for notational convenience, we fix that 
+%$r_1 = \silent$,
+$r_{u+1} = \mathit{cxt}$ and $p_{u+1}=\silent$,  for $u$ the number of reacts.}
 \end{definition}
@@
 
- \begin{example}\label{ex:backbone}
- Let ${\cal A}$ be a RS whose specification contains two entities, $s1$ and $s2$, and, among the others, the reaction $a=(s1,\ , s1)$ that guarantees the presence of the  $s1$ in the system.
+{\color{red} \begin{example}\label{ex:backbone}
+ Let ${\cal A}$ be a RS whose specification contains two entities, $s1$ and $s2$, and the reactions $r_1=(s1,\ , s2)$ and $r_2 = (s2, \ , s1)$ that produce $s2$ if $s1$ is present and $s1$ if $s2$ is present.
+For simplicity we do not consider inhibitors.
 Then, we assume an  extended interactive process $\pi=(\gamma,\delta)$ where the context $\gamma$ provides  $s1$ and $s2$.
  %, for the first $n$ steps.
- Our translation includes the processes:\\
+ Our translation follows:\\% includes the processes:\\
  \[
- \begin{array}{lcllcl}
- P_a &\defeq& \startchain{\silent}\chainedlink{s1_i}{\noact}\chainedlink{\noact}{s1_o}\chainedlink{r_2}{\noact} \chainedlink{\noact}{p_1}\chainedlink{\tilde{s1}_i}{\noact}\chainedlink{\noact}{\tilde{s1}_o}\chainend{p_2}.P_a + \dots; \\
- P_{s1} &\defeq& \startchain{s1_i}\chainedlink{s1_o}{\noact}\chainedlink{\noact}{\hat{s1}_i}\chainedlink{\hat{s1}_o}{\noact}\chainedlink{\noact}{\tilde{s1}_i}\chainedlink{\tilde{s1}_o}.P_{s1} + \dots; &
-  P_{s2} &\defeq& \link{\hat{s2}_i}{\hat{s2}_o}.P_{s2} + \dots;\\
- Cxt_{s1}& \defeq & \startchain{cxt_1}\chainedlink{\hat{s1}_i}{\noact}\chainedlink{\noact}{\hat{s1}_o}\chainend{cxt_2}.Cxt_{s1} &
-  Cxt _{s2} &\defeq& \startchain{cxt_2}\chainedlink{\hat{s2}_i}{\noact}\chainedlink{\noact}{\hat{s2}_o}\chainend{p_1}. Cxt_{s2}
+ \begin{array}{l@{\hspace{-0.009cm}}c@{\hspace{-0.009cm}}ll@{\hspace{-0.009cm}}c@{\hspace{-0.009cm}}l}
+ P_{r_1} &\defeq& \startchain{r_1}\chainedlink{s1_i}{\noact}\chainedlink{\noact}{s1_o}\chainedlink{r_2}{\noact} \chainedlink{\noact}{p_1}\chainedlink{\tilde{s2}_i}{\noact}\chainedlink{\noact}{\tilde{s2}_o}\chainend{p_2}.P_{r_1} + \dots; &
+   P_{s2} &\defeq& \startchain{\hat{s2}_i}\chainedlink{\hat{s2}_o}{\noact}
+   \chainedlink{\noact} {\tilde{s2}_i}\chainend{\tilde{s2}_o}.P_{s2} + \dots;\\
+  P_{r_2} &\defeq& \startchain{r_2}\chainedlink{\overline{s2}_i}{\noact}\chainedlink{\noact}{\overline{s2}_o}\chainedlink{cxt}{\noact} \chainedlink{\noact}{p_2}\chainend{\silent}.P_{r_2} + \dots; &
+ P_{s1} &\defeq& \startchain{s1_i}\chainedlink{s1_o}{\noact}\chainedlink{\noact}{\hat{s1}_i}\chainend{\hat{s1}_o}
+ %{\noact}\chainedlink{\noact}{\tilde{s1}_i}\chainedlink{\tilde{s1}_o}
+ .P_{s1} + \dots; \\
+ Cxt& \defeq & \startchain{cxt}\chainedlink{\hat{s1}_i}{\noact}\chainedlink{\noact}{\hat{s1}_o}
+  \chainedlink{\hat{s2}_i}{\noact}\chainedlink{\noact}{\hat{s2}_o}\chainend{p_1}.Cxt\\
+
+  %Cxt& \defeq & \startchain{cxt}\chainedlink{\hat{s1}_i}{\noact}\chainedlink{\noact}{\hat{s1}_o}
+  %\chainedlink{\hat{s2}_i}{\noact}\chainedlink{\noact}{\hat{s2}_o}\chainend{p_1}.Cxt
+ %Cxt_{s1}& \defeq & \startchain{cxt_1}\chainedlink{\hat{s1}_i}{\noact}\chainedlink{\noact}{\hat{s1}_o}\chainend{cxt_2}.Cxt_{s1} &
+ % Cxt _{s2} &\defeq& \startchain{cxt_2}\chainedlink{\hat{s2}_i}{\noact}\chainedlink{\noact}{\hat{s2}_o}\chainend{p_1}. Cxt_{s2}
 \end{array}
 \]  \\
-   Now, we assume that $s1$ is in the initial state of ${\cal A}$, and in Figure~\ref{fig:backbone} we show the  structure of a link chain label related to the execution of a transition of the c\CNA \ system: $\restrict{names} (P_{s1}|P_{s2}| P_a|\dots|Cxt_{s1}|Cxt_{s2})$.
-The yellow blocks are referred to the processes encoding the reactions ($P_a$, in our case) and the contexts ($Cxt_{s1}$ and $Cxt_{s2}$).
+   Now, we assume that $s1$ only is in the initial state of ${\cal A}$; for clarity we partially specify the {\tt link}-calculus code of the processes.
+In Figure~\ref{fig:backbone} we show the  structure of a link chain label related to the execution of a transition of the c\CNA \ system: $\restrict{names} (Init | P_{s1}|P_{s2}| P_a|\dots|Cxt)$.
+The yellow blocks are referred to process $Init$, and the processes encoding the reactions ($P_{r_1}$, in our case) and the context ($Cxt$).
 As the figure puts in evidence, these two kinds of processes determine the structure of the link chain, from end to end, i.e.
@@
   \draw (4.38,0.75) node[minimum size=0.47cm,draw,fill=yellow, draw=yellow] 
-  {${\scriptscriptstyle  cxt_1}$};
+  {${\scriptscriptstyle  cxt \ }$};
 
@@
 
-    \draw (6,1) -- (11.5,1) ;
-    \draw (7,1) -- (7,0.5) -- (7.5,0.5) -- (7.5,0);
-    \draw (6.5,0) -- (11.5,0);
+    \draw (6,1) -- (10,1) ;
+ %   \draw (7,1) -- (7,0.5) -- (7.5,0.5) -- (7.5,0);
+    \draw (6.5,0) -- (10.5,0);
     \draw   (6,1)-- (6,0.5) -- (6.5,0.5) -- (6.5,0.5) -- (6.5,0) ;
@@
  \draw (6.36,0.75) node[minimum size=0.475cm,draw,fill=yellow,draw=yellow] 
- {${\scriptscriptstyle cxt_2}$};%node[minimum size=0.5cm,draw,draw=white] {$\noact$};
+ {${\scriptscriptstyle \hat{s2}_o \ }$};%node[minimum size=0.5cm,draw,draw=white] {$\noact$};
    \draw (6.75,0.26)  node[minimum size=0.485cm,fill=yellow, draw=yellow] {$\ $};
 
- \draw (7.20,0.25) node[minimum size=0.475cm,fill=yellow, draw=yellow] {${\scriptscriptstyle \hat{s2}_i}$};
-
-     \draw (7.3,0.75)   node{$\noact$};
-  \draw (7.75,0.75) node{$\ $};
-  \draw (7.76,0.26) node {$\ $};
-  \draw (8.25,0.26)  node {$\noact$ };
-
-  \draw (8.3,0.75)  node[minimum size=0.48cm,draw,fill=yellow, draw=yellow] {${\scriptscriptstyle \hat{s2}_o}$};
-  \draw (8,1)--(8,0.5)--(8.5,0.5) -- (8.5,0);
- % \draw(9.5,0.5) --(9.5,0);
-  \draw (8.745,0.75)  node[minimum size=0.48cm,draw,fill=yellow,yellow] {$\ $};
-  \draw (8.75,0.26)  node[minimum size=0.48cm,draw,fill=yellow,yellow] {$\ $};
- \draw (9.24,0.26)  node[minimum size=0.48cm,draw,fill=yellow,draw=yellow] {${\scriptscriptstyle p_1}$};
+% \draw (7.20,0.25) node[minimum size=0.475cm,fill=yellow, draw=yellow] {${\scriptscriptstyle{p_1}} \ 
+\draw (7.20,0.25)  node[minimum size=0.475cm,draw,fill=yellow,draw=yellow] {${\scriptscriptstyle p_1}$};
+\draw (7.2,0.74) node[minimum size=0.48cm,draw,fill=yellow,draw=yellow]{${\scriptscriptstyle p_1}$};
+\draw (7.6,0.74)  node[minimum size=0.48cm,draw,fill=yellow, draw=yellow] {$ \ $};
+\draw (7.6,0.25)  node[minimum size=0.48cm,draw,fill=yellow, draw=yellow] {$ \ $};
+\draw (8.14,0.25)  node[minimum size=0.475cm,draw,fill=yellow,draw=yellow] {${\scriptscriptstyle \tilde{s1}_i}$};
+ \draw (8.14,0.75) node{$\noact$};%node[minimum size=0.5cm,draw,draw=white] {$\noact$};
+ \draw (9.24,0.25) node{$\noact$};
+  \draw (7.85,1) -- (7.85,0.5) ;
+   \draw (7.85,0.5) -- (8.45,0.5) ;
+      \draw  (8.45,0.5) --(8.45,0) ;
+       \draw  (8.93,1) --(8.93,0.5) ;
+        \draw  (8.93,0.5) --(9.5,0.5) ;
+             \draw  (9.49,0.5) --(9.49,0) ;
+         \draw  (9.49,0.5) --(9.49,0) ;
+         \draw(10.0,1) -- (10.0,0.5);
+          \draw(10,0.52) -- (10.48,0.52);
+          \draw(10.48,0.52)--(10.49,0);
+%  $};
+%
+%%     \draw (7.3,0.75)   node{$\noact$};
+%%  \draw (7.75,0.75) node{$\ $};
+%%  \draw (7.76,0.26) node {$\ $};
+%%  \draw (8.25,0.26)  node {$\noact$ };
+%
+%  \draw (8.3,0.75)  node[minimum size=0.48cm,draw,fill=yellow, draw=yellow] {${\scriptscriptstyle \hat{s2}_o}$};
+%  \draw (8,1)--(8,0.5)--(8.5,0.5) -- (8.5,0);
+% % \draw(9.5,0.5) --(9.5,0);
+%  \draw (8.745,0.75)  node[minimum size=0.48cm,draw,fill=yellow,yellow] {$\ $};
+%  \draw (8.75,0.26)  node[minimum size=0.48cm,draw,fill=yellow,yellow] {$\ $};
+ %\draw (9.24,0.26)  node[minimum size=0.48cm,draw,fill=yellow,draw=yellow] {${\scriptscriptstyle p_1}$};
  %\draw(9.5,0.5) --(9.5,0);
     % \draw (9,1) -- (10,1) -- (10,0.5) -- (10.5,0.5) --  (10.5,0) -- (9.5,0)  (9.5,0.5) -- (9,0.5) -- (9,1);
- \draw (9.24,0.75) node[minimum size=0.48cm,draw,fill=yellow,draw=yellow]{${\scriptscriptstyle p_1}$};%node[minimum size=0.5cm,draw,draw=white] {$\noact$};
- \draw (10.22,0.26) node [minimum size=0.48cm,draw,fill=yellow,draw=yellow]{${\scriptscriptstyle \tilde{s1}_i}$};
+ \draw (9.24,0.75) node[minimum size=0.48cm,draw,fill=yellow,draw=yellow]{${\scriptscriptstyle \tilde{s1}_o}$};%node[minimum size=0.5cm,draw,draw=white] {$\noact$};
+
+ \draw (10.22,0.26) node [minimum size=0.48cm,draw,fill=yellow,draw=yellow]{$\silent $};
+%{${\scriptscriptstyle \tilde{s1}_i}$};
  \draw (9.74,0.745) node [minimum size=0.485cm,draw,fill=yellow,draw=yellow]{$\ $};
  \draw (9.735,0.26) node [minimum size=0.48cm,draw,fill=yellow,draw=yellow]{$\ $};
- \draw (10,1) -- (10,0.5) -- (10.5,0.5)--(10.5,0);
-  \draw (10.31,0.75)  node{$\noact$};
-  \draw (10.75,0.75)  node {$\ $};
-  \draw (10.75,0.26)  node{$\ $};
- \draw (11.25,0.25)  node {$\noact$};
-   \draw (11,1) -- (11,0.5) -- (11.5,0.5) -- (11.5,0);% -- (10,0) -- (9,0) -- (9,0.5) -- (8.5,0.5) -- (8.5,1);
-    \draw (11,1) -- (11.5,1) -- (12,1) -- (12,0.5) -- (12.5,0.5)--(12.5,0)--(11.5,0);
+ %  \draw (10.8,0.5) node{$\dots$};
+% \draw (10,1) -- (10,0.5) -- (10.5,0.5)--(10.5,0);
+  %\draw (10.31,0.75)  node{$\noact$};
+%  \draw (10.75,0.75)  node {$\ $};
+%  \draw (10.75,0.26)  node{$\ $};
+% \draw (11.25,0.25)  node {$\noact$};
+%   \draw (11,1) -- (11,0.5) -- (11.5,0.5) -- (11.5,0);% -- (10,0) -- (9,0) -- (9,0.5) -- (8.5,0.5) -- (8.5,1);
+%    \draw (11,1) -- (11.5,1) -- (12,1) -- (12,0.5) -- (12.5,0.5)--(12.5,0)--(11.5,0);
    %\draw (11.25,0.25)  node[minimum size=0.48cm,draw,fill=yellow,draw=yellow] {$\noact$};
-    \draw (11.31,0.75) node[minimum size=0.48cm,draw,fill=yellow,draw=yellow]{${\scriptscriptstyle \tilde{s1}_o}$};
-    \draw (11.75,0.26)node[minimum size=0.48cm,draw,fill=yellow,draw=yellow]{$\ $};
-      \draw (11.75,0.75)node[minimum size=0.48cm,draw,fill=yellow,draw=yellow]{$\ $};
-     \draw (12.2,0.25) node[minimum size=0.478cm,draw,fill=yellow,draw=yellow]{\;$\silent$\;  };
+%    \draw (11.31,0.75) node[minimum size=0.48cm,draw,fill=yellow,draw=yellow]{${\scriptscriptstyle \tilde{s1}_o}$};
+%    \draw (11.75,0.26)node[minimum size=0.48cm,draw,fill=yellow,draw=yellow]{$\ $};
+%      \draw (11.75,0.75)node[minimum size=0.48cm,draw,fill=yellow,draw=yellow]{$\ $};
+%     \draw (12.2,0.25) node[minimum size=0.478cm,draw,fill=yellow,draw=yellow]{\;$\silent$\;  };
   \end{tikzpicture}
@@
  \end{example}
-
-% \begin{figure}[t]
-% \begin{tikzpicture}
-%   \draw (0,1) -- (5,1) ;
-%    \draw (0.5,0) -- (5.5,0) ;
-%   \draw (0,1)--(0,0.5) -- (0.5,0.5) -- (0.5,0) -- (1.5,0) ;
-%     \draw (0.25,0.75)  node[minimum size=0.495cm,fill=yellow] {${ \silent}$};
-% \draw (0.75,0.75)  node[minimum size=0.5cm,fill=yellow] {$\ $};
-%  \draw (0.76,0.25)  node[minimum size=0.5cm,fill=yellow] {$\ $};
-%  \draw (1.21,0.25)  node[minimum size=0.5cm,fill=yellow] {${\scriptscriptstyle s1_i}$};
-%%    \draw (1.25,0.75)  node[minimum size=0.5cm,draw,draw=white] {$\noact$};
-%%  \draw (1.75,0.75)  node[minimum size=0.5cm,draw,white] {$\ $};
-%%  \draw (1.75,0.25)  node[minimum size=0.5cm,draw,white] {$\ $};
-%%  \draw (2.25,0.25)  node[minimum size=0.5cm,draw,draw=white] {$\noact$};
-%  \draw (2,1) -- (2,0.5) -- (2.5,0.5) -- (2.5,0) -- (1.5,0)  -- (1.5,0.5) -- (1,0.5) -- (1,1);
-%
-% \draw (1.25,0.75) node{$\noact$};%node[minimum size=0.5cm,draw,draw=white] {$\noact$};
-%  \draw (2.25,0.25) node{$\noact$};
-%   \draw (2.31,0.75)  node[minimum size=0.47cm,draw,fill=yellow, draw=yellow] {${\scriptscriptstyle s1_o}$};
-%  \draw (2.75,0.75)  node[minimum size=0.48cm,draw,fill=yellow,yellow] {$\ $};
-%  \draw (2.75,0.26)  node[minimum size=0.49cm,draw,fill=yellow,yellow] {$\ $};
-%
-%  \draw (3.20,0.25)  node[minimum size=0.47cm,draw,fill=yellow,draw=yellow] {${\scriptscriptstyle  \tilde{s1}_i}$};
-%  
-%  
-%  \draw(3.5,0.5) -- (3.5,0);
-%%     \draw (3.25,0.75)  node[minimum size=0.5cm,draw,draw=white] {$\noact$};
-%%  \draw (3.75,0.75)  node[minimum size=0.5cm,draw,white] {$\ $};
-%%  \draw (3.75,0.25)  node[minimum size=0.5cm,draw,white] {$\ $};
-%%  \draw (4.25,0.25)  node[minimum size=0.5cm,draw,draw=white] {$\noact$};
-%\draw (3,1) -- (4,1) -- (4,0.5) -- (4.5,0.5) -- (4.5,0) -- (3.5,0)  (3.5,0.5) -- (3,0.5) -- (3,1);
-% \draw (3.25,0.75) node{$\noact$};%node[minimum size=0.5cm,draw,draw=white] {$\noact$};
-%  \draw (4.25,0.25) node{$\noact$};
-%  \draw (4.31,0.75)  node[minimum size=0.47cm,draw,fill=yellow, draw=yellow] {${ \scriptscriptstyle \tilde{s1}_o}$};
-%  \draw (4.75,0.745)  node[minimum size=0.48cm,draw,fill=yellow,yellow] {$\ $};
-%  \draw (4.75,0.25)  node[minimum size=0.48cm,draw,fill=yellow,yellow] {$\ $};
-%  \draw (5.24,0.255)  node[minimum size=0.485cm,draw,fill=yellow,draw=yellow] {${\scriptscriptstyle  r_2}$};
-%  \draw(5,1) -- (5,0.5) -- (5.5,0.5) -- (5.5,0);
-%   \draw (6,0.5) node{$\dots$};
-%    \draw (6.5,1) -- (11.5,1) ;
-%    \draw (7.5,1) -- (7.5,0.5) -- (8,0.5) -- (8,0);
-%    \draw (7,0) -- (12,0);
-%    \draw   (6.5,1)-- (6.5,0.5) -- (7,0.5) -- (7,0.5) -- (7,0) ;
-% \draw (6.77,0.75) node{$\noact$};%node[minimum size=0.5cm,draw,draw=white] {$\noact$};
-% \draw (7.77,0.25) node{$\noact$};
-%
-%     \draw (7.80,0.75)  node[minimum size=0.485cm,draw,fill=yellow, draw=yellow] {${\scriptscriptstyle  \tilde{s1}_o}$};
-%  \draw (8.25,0.75)  node[minimum size=0.485cm,draw,fill=yellow,yellow] {$\ $};
-%  \draw (8.26,0.26)  node[minimum size=0.495cm,draw,fill=yellow,yellow] {$\ $};
-%  \draw (8.75,0.26)  node[minimum size=0.49cm,draw,fill=yellow,draw=yellow] {${\scriptscriptstyle cxt_2}$ };
-%
-%  \draw (8.75,0.75)  node[minimum size=0.48cm,draw,fill=yellow, draw=yellow] {${\scriptscriptstyle cxt_2}$};
-%  \draw (9.25,0.75)  node[minimum size=0.48cm,draw,fill=yellow,yellow] {$\ $};
-%  \draw (9.25,0.26)  node[minimum size=0.48cm,draw,fill=yellow,yellow] {$\ $};
-% \draw (9.71,0.25)  node[minimum size=0.47cm,draw,fill=yellow,draw=yellow] {${\scriptscriptstyle \tilde{s2}_i}$};
-% \draw(10,0.5) --(10,0);
-%     \draw (9.5,1) -- (10.5,1) -- (10.5,0.5) -- (11,0.5) --  (11,0) -- (10,0)  (10,0.5) -- (9.5,0.5) -- (9.5,1);
-% \draw (9.75,0.75) node{$\noact$};%node[minimum size=0.5cm,draw,draw=white] {$\noact$};
-% \draw (10.75,0.25) node{$\noact$};
-%  \draw (10.81,0.75)  node[minimum size=0.47cm,draw,fill=yellow, draw=yellow] {${\scriptscriptstyle \tilde{s2}_o}$};
-%  \draw (11.25,0.75)  node[minimum size=0.48cm,draw,fill=yellow,yellow] {$\ $};
-%  \draw (11.25,0.26)  node[minimum size=0.49cm,draw,fill=yellow,yellow] {$\ $};
-% \draw (11.75,0.25)  node[minimum size=0.48cm,draw,fill=yellow,draw=yellow] {$\silent$};
-%   \draw (11.5,1) -- (11.5,0.5) -- (12,0.5) -- (12,0);% -- (10,0) -- (9,0) -- (9,0.5) -- (8.5,0.5) -- (8.5,1);
-%  \end{tikzpicture}
-%  \caption{The link chain structure arising from reactions and context processes.}
-%   \label{fig:backbone}
-% \end{figure}
-% \end{example}
+}
 
@@
   \item 
- $P'= \restrict{\mathit{names}} (\Pi_{a \in A}P_a~|~ \mathit{Cxt}^2 ~|~ \Pi_{s \in C_1 \cup D_1}P_s| \Pi_{s \notin C_1 \cup D_1}\overline{P}_s)$.
+ $P'= \restrict{\mathit{names}} (Init | \Pi_{a \in A}P_a~|~ \mathit{Cxt}^2 ~|~ \Pi_{s \in C_1 \cup D_1}P_s| \Pi_{s \notin C_1 \cup D_1}\overline{P}_s)$.
  \end{enumerate}

### 2020-04-01 05-translation.tex
just typos

@@
 
-\paragraph{Processes for entities.}
+\paragraph{Processes for entities}
 Given an entity $s \in S$, we exploit four different names for the interactions over $s$:
@@
 
-\paragraph{Processes for reactions.}
+\paragraph{Processes for reactions}
 We assume that each reaction $a$ is assigned a progressive number $j$.
@@
 
-\paragraph{Processes for contexts.}
+\paragraph{Processes for contexts}
 The context can vary at each step $n$ and for each entity $s \in S$, the context must say if the entity is provided or not.

### 2020-04-01 06-examples.tex
commit
just typos

@@
 \label{ex:examples}
+Hereafter, we only take into account transitions whose labels are solid, i.e. they do not contain 
+virtual links, and   complete, i.e. thy starts and ends with the $\silent$ symbol.\\
 
@@
 \]
-\paragraph{Encoding of the rules.}
+\paragraph{Encoding of the rules}
 The encoding of the rules for reactions is  given in a parametric way:
@@
 
-\paragraph{Encoding of the entities.}
+\paragraph{Encoding of the entities}
 As for reactions, also the encoding of the entities is given in a parametric way.
@@
 the other cases (providing both $a$ and $b$ or neither of them) leads the system to be stuck.\\
-Now, we assume that we have an initial configuration containing the entities $q$ and $b$ (with$\tilde{q},\tilde{w},\tilde{a},\tilde{b}$ grouping all the possibile decorations for these names):
+Now, we assume that we have an initial configuration containing the entities $q$ and $b$ (with $\tilde{q},\tilde{w},\tilde{a},\tilde{b}$ grouping all the possibile decorations for these names):
 \[
@@
 \subsection{A  biological toy example}\label{subsec:toy}
-Hereafter, we only take into account transitions whose labels are solid, i.e. they do not contain 
-virtual links, and   complete, i.e. thy starts and ends with the $\silent$ symbol.\\
 
@@
 rule four codes for the production of $G$ when  proteins $T'$ is present and $F$ is absent.
-\paragraph{Encoding of the rules.}
+
+\paragraph{Encoding of the rules}
+
 The encoding of the rules for reactions is  given in a parametric way:
@@
 and we put $r_1 = \silent$.
-\paragraph{Encoding of the entities.}
+
+\paragraph{Encoding of the entities}
+
 As for reactions, also the encoding of the entities is given in a parametric way.

### 2020-04-01 06b-bisimulation.tex
Update 06b-bisimulation.tex
examples + flat fucntion
fino a sect4 e flat
update fino a sezione 4 inclusa e def di flat
just typos

@@
 
-For this reason, we propose an alternative notion of bisimulation, that hereafter we call \emph{biosimulation},
+For this reason, we propose an alternative notion of bisimulation, that hereafter we call \emph{bio-simulation},
 that allows us  to compare two biological systems by restricting the observation to only the limited events of interest.
@@
 This transformation is harmless, in the sense that it retains all the information in the chain, because it is applied to solid chains only.
+Below we denote by $\mathit{decs} = \{ s, \overline{s}, \widetilde{s},\widehat{s},\underline{s}~|~ s\in S\}$ the set of decorated names (without subscripts $_i$ and $_o$).
 The function $\mathit{flat}$ is defined inductively as follows:
+%\[
+%\begin{array}{lcl@{\hspace{0.5cm}}lcl@{\hspace{0.5cm}}lcl@{\hspace{0.5cm}}lcl}
+%\mathit{flat}(\epsilon)&\defeq & \epsilon &
+%\mathit{flat}(\link{\gamma}{\gamma'}) &  \defeq & \left \{ \begin{array}{cl} \epsilon & \mathrm{if } (\gamma = \beta_i \wedge \gamma = \beta_o)  \vee \gamma = \silent\\
+%\gamma' & \mathrm{otw }  \end{array} \right. &
+%%\mathit{flat}(\link{\alpha}{\silent})& \defeq & \epsilon &
+%%\mathit{flat}(\link{\alpha}{\beta}) & \defeq &\beta \\
+%\end{array}\]
+
 \[
-\begin{array}{lcl@{\hspace{0.5cm}}lcl@{\hspace{0.5cm}}lcl@{\hspace{0.5cm}}lcl}
-\mathit{flat}(\epsilon)&\defeq & \epsilon &
-\mathit{flat}(\link{\silent}{\beta}) & \defeq & \beta &
-\mathit{flat}(\link{\alpha}{\silent})& \defeq & \epsilon &
-\mathit{flat}(\link{\alpha}{\beta}) & \defeq &\beta \\
-\end{array}\]
+\mathit{flat}(\epsilon)\defeq \epsilon 
+\qquad
+\mathit{flat}(\link{\alpha}{\beta}) \defeq 
+\left \{ \begin{array}{ll} 
+\beta & \mbox{if $\beta\in \mathit{reacts}\cup\{\mathit{cxt}\}\cup\mathit{prods}$}\\
+\gamma & \mbox{if $\beta=\gamma_i$ with $\gamma\in\mathit{decs}$}\\
+\epsilon & \mbox{otherwise}  
+\end{array} \right. 
+\]
+
 \[
 \begin{array}{lcl}
-\mathit{flat}(\link{\alpha}{\beta}\upsilon)& \defeq & \beta :: \mathit{flat}(\upsilon)\\
+\mathit{flat}(\link{\alpha}{\beta}\upsilon)& \defeq & \mathit{flat}(\link{\alpha}{\beta}):: \mathit{flat}(\upsilon)\\
 \end{array}\\
@@
  \llbracket  F_1 \vee F_2 \rrbracket &\defeq& \llbracket  F_1\rrbracket \cup \llbracket  F_2\rrbracket  \\
- \llbracket  F_1 \wedge F_2 \rrbracket &\defeq& \llbracket F_1 \rrbracket \wedge \llbracket F_2\rrbracket 
+ \llbracket  F_1 \wedge F_2 \rrbracket &\defeq& \llbracket F_1 \rrbracket \cap \llbracket F_2\rrbracket 
 \end{array}
@@
 
-The notion of biosimulation builds on the above language of assertions to parameterize the induced equivalence on the property of interest. 
+\subsection{Bio-similarity and bio-logical equivalence}
+
+The notion of bio-simulation builds on the above language of assertions to parameterize the induced equivalence on the property of interest. 
 Please recall that we have defined the behaviour of the contest in a non determinist way, thus 
 at each step, different possible sets of entities can be provided to the system and different sets of reaction can be enabled/disabled. 
-Biosimulation can thus be used to compare the behaviour of different systems that share some of the reactions or entities or also to compare the behaviour of the same set of reaction rules when different contexts are provided.
+Bio-simulation can thus be used to compare the behaviour of different systems that share some of the reactions or entities or also to compare the behaviour of the same set of reaction rules when different contexts are provided.
 
@@
 
-\begin{definition}[Biosimilarity  $ \sim_F $]
-Given an assertion $F$, a \emph{biosimulation} $\mathbf{R}_F$ that respects $F$ is a binary relation over {\tt link}-calculus processes such that, if $P \mathrel{\mathbf{R}_F} Q$ then:
+\begin{definition}[Bio-similarity  $ \sim_F $]
+Given an assertion $F$, a \emph{bio-simulation} $\mathbf{R}_F$ that respects $F$ is a binary relation over {\tt link}-calculus processes such that, if $P \mathrel{\mathbf{R}_F} Q$ then:
  \begin{itemize} 
@@
 %\end{itemize}
-We let $ \sim_F $ denote the largest  biosimulation and we say that $P$ is \emph{biosimilar} to $Q$, with respect to $F$, if $P \sim_F Q$.
+We let $ \sim_F $ denote the largest  bio-simulation and we say that $P$ is \emph{bio-similar} to $Q$, with respect to $F$, if $P \sim_F Q$.
 \end{definition}
@@
 \begin{remark}
-An alternative way to look at a biosimulation that respects $F$ is to define it as an ordinary bisimulation over the transition system labelled over $\{F,\neg F\}$ obtained by transforming each transition $P \xrightarrow{v} P'$ such that $v\entails F$ into $P \xrightarrow{F} P'$ and  each transition $P \xrightarrow{v} P'$ such that $v\not\entails F$ into $P \xrightarrow{\neg F} P'$.
+An alternative way to look at a bio-simulation that respects $F$ is to define it as an ordinary bisimulation over the transition system labelled over $\{F,\neg F\}$ obtained by transforming each transition $P \xrightarrow{v} P'$ such that $v\entails F$ into $P \xrightarrow{F} P'$ and  each transition $P \xrightarrow{v} P'$ such that $v\not\entails F$ into $P \xrightarrow{\neg F} P'$.
 \end{remark}
@@
 
-Finally, we extend the classical result establishing the correspondence between the logical equivalence induced by HML with bisimilarity for proving that biosimilarity coincides with bio-logical equivalence.
+Finally, we extend the classical result establishing the correspondence between the logical equivalence induced by HML with bisimilarity for proving that bio-similarity coincides with bio-logical equivalence.
 

### 2020-04-01 05-translation.tex
flat function

--- 2020-03-31 05-translation.tex
@@
 \triangleq
- \restrict{\mathit{names}}(\Pi_{s\in C_0}P_{s} ~|~ \Pi_{s \notin C_0}\overline{P_s} ~|~ \Pi_{a\in A} P_a ~|~  \mathit{Cxt}^1)
+ \restrict{\mathit{names}}(I~|~\Pi_{s\in C_0}P_{s} ~|~ \Pi_{s \notin C_0}\overline{P_s} ~|~ \Pi_{a\in A} P_a ~|~  \mathit{Cxt}^1)
 $$
@@
 %For notational convenience, we fix that $r_1 = \silent$,  $r_{u+1} = cxt_1$ for $u$ the number of reacts, and $cxt_{w+1} = \silent$ for $w$ the number of entities.
-{\color{red} For technical reasons, we introduce the process $Init \defeq \link{\silent}{r_1}.Init$
+{\color{red} For technical reasons, we introduce the init process $I \defeq \link{\silent}{r_1}.I$
 to allow the name $r_1$ to be matched; also, 
@@
    Now, we assume that $s1$ only is in the initial state of ${\cal A}$; for clarity we partially specify the {\tt link}-calculus code of the processes.
-In Figure~\ref{fig:backbone} we show the  structure of a link chain label related to the execution of a transition of the c\CNA \ system: $\restrict{names} (Init | P_{s1}|P_{s2}| P_a|\dots|Cxt)$.
-The yellow blocks are referred to process $Init$, and the processes encoding the reactions ($P_{r_1}$, in our case) and the context ($Cxt$).
+In Figure~\ref{fig:backbone} we show the  structure of a link chain label related to the execution of a transition of the c\CNA \ system: $\restrict{names} (I | P_{s1}|P_{s2}| P_a|\dots|Cxt)$.
+The yellow blocks are referred to init process $I$, and the processes encoding the reactions, $P_{r_1}$ and $P_{r_2}$, and the context $Cxt$.
 As the figure puts in evidence, these two kinds of processes determine the structure of the link chain, from end to end, i.e.
@@
 We could say that these processes form the \emph{backbone} of the interaction.
-   In contrast, the processes encoding the entities ($P_{s1}$, and $P_{s2}$, in our case) provides the
+   In contrast, the processes encoding the entities, $P_{s1}$,  $P_{s2}$, provides the
     solid links to overlap the virtual links   of the backbone.
@@
     \draw (6.5,0) -- (10.5,0);
-    \draw   (6,1)-- (6,0.5) -- (6.5,0.5) -- (6.5,0.5) -- (6.5,0) ;
+    \draw   (5.99,1)-- (5.99,0.5) -- (6.5,0.5) -- (6.5,0.5) -- (6.5,0) ;
+    \
      \draw (6.75,0.75)  node[minimum size=0.47cm,fill=yellow, draw=yellow] {$\ $};
  \draw (6.36,0.75) node[minimum size=0.475cm,draw,fill=yellow,draw=yellow] 
- {${\scriptscriptstyle \hat{s2}_o \ }$};%node[minimum size=0.5cm,draw,draw=white] {$\noact$};
+ {${\scriptscriptstyle \hat{s2}_o \ }$};
+ %node[minimum size=0.5cm,draw,draw=white] {$\noact$};
    \draw (6.75,0.26)  node[minimum size=0.485cm,fill=yellow, draw=yellow] {$\ $};
@@
  $P=\llbracket{\cal A},\gamma\rrbracket$ with
-$$P  = \restrict{\mathit{names}} (\Pi_{a \in A}P_a ~|~ \mathit{Cxt}^1 ~|~ \Pi_{s \in C_0}P_s ~|~ \Pi_{s \notin C_0}\overline{P}_s).
+$$P  = \restrict{\mathit{names}} (I~|~\Pi_{a \in A}P_a ~|~ \mathit{Cxt}^1 ~|~ \Pi_{s \in C_0}P_s ~|~ \Pi_{s \notin C_0}\overline{P}_s).
 $$
@@
   \item 
- $P'= \restrict{\mathit{names}} (Init | \Pi_{a \in A}P_a~|~ \mathit{Cxt}^2 ~|~ \Pi_{s \in C_1 \cup D_1}P_s| \Pi_{s \notin C_1 \cup D_1}\overline{P}_s)$.
+ $P'= \restrict{\mathit{names}} (I~|~\Pi_{a \in A}P_a~|~ \mathit{Cxt}^2 ~|~ \Pi_{s \in C_1 \cup D_1}P_s| \Pi_{s \notin C_1 \cup D_1}\overline{P}_s)$.
  \end{enumerate}
@@
 
+\paragraph{The flat function}
+For technical reasons, our transition labels are quite verbose; then, to simplify their processing, we introduce a function that takes a solid link chain and returns a simple string by eliminating all the channel matching pairs leaving just one channel per each pair.
+This transformation is harmless, in the sense that it retains all the information in the chain, because it is applied to solid chains only.
+Below we denote by $\mathit{decs} = \{ s, \overline{s}, \widetilde{s},\widehat{s},\underline{s}~|~ s\in S\}$ the set of decorated names (without subscripts $_i$ and $_o$).
+The function $\mathit{flat}$ is defined inductively as follows:
+%\[
+%\begin{array}{lcl@{\hspace{0.5cm}}lcl@{\hspace{0.5cm}}lcl@{\hspace{0.5cm}}lcl}
+%\mathit{flat}(\epsilon)&\defeq & \epsilon &
+%\mathit{flat}(\link{\gamma}{\gamma'}) &  \defeq & \left \{ \begin{array}{cl} \epsilon & \mathrm{if } (\gamma = \beta_i \wedge \gamma = \beta_o)  \vee \gamma = \silent\\
+%\gamma' & \mathrm{otw }  \end{array} \right. &
+%%\mathit{flat}(\link{\alpha}{\silent})& \defeq & \epsilon &
+%%\mathit{flat}(\link{\alpha}{\beta}) & \defeq &\beta \\
+%\end{array}\]
+
+\[
+\mathit{flat}(\epsilon)\defeq \epsilon 
+\qquad
+\mathit{flat}(\link{\alpha}{\beta}) \defeq 
+\left \{ \begin{array}{ll} 
+\beta & \mbox{if $\beta\in \mathit{reacts}\cup\{\mathit{cxt}\}\cup\mathit{prods}$}\\
+\gamma & \mbox{if $\beta=\gamma_i$ with $\gamma\in\mathit{decs}$}\\
+\epsilon & \mbox{otherwise}  
+\end{array} \right. 
+\]
+
+\[
+\begin{array}{lcl}
+\mathit{flat}(\link{\alpha}{\beta}\upsilon)& \defeq & \mathit{flat}(\link{\alpha}{\beta}):: \mathit{flat}(\upsilon)\\
+\end{array}\\
+\]
+where  $::$ is the usual string concatenation operator.
+
 % 

### 2020-04-01 06-examples.tex
flat function

--- 2020-03-27 06-examples.tex
@@
 \label{ex:examples}
+Hereafter, we only take into account transitions whose labels are solid, i.e. they do not contain 
+virtual links, and   complete, i.e. they starts and ends with the $\silent$ symbol.\\
 
@@
 \noindent
-and we put $r_1 = \silent$, $r_5 = cxt$ and  $p_5 = \silent$.
+and we put 
+%$r_1 = \silent$, 
+$r_5 = cxt$ and  $p_5 = \silent$.
 
@@
 \begin{array}{lcl}
-Sys  \defeq \ \restrict{r1,r2,r3,r4,\tilde{q},\tilde{w},\tilde{a},\tilde{b}} (P_q \mid \overline{P}_w  \mid  P_a  \mid \overline{P}_b  \mid  P_1  \mid  P_2  \mid  P_3  \mid P_4  \mid Cxt).
+Sys  \defeq \ %\restrict{r1,r2,r3,r4,\tilde{q},\tilde{w},\tilde{a},\tilde{b}}
+ I\mid P_q \mid \overline{P}_w  \mid  P_a  \mid \overline{P}_b  \mid  P_1  \mid  P_2  \mid  P_3  \mid P_4  \mid Cxt.
 \end{array}
 \] 
-Then only the second reaction can be applied, and the label transition follows
-{\small
+Then, only the second reaction can be applied, and the label transition $\ell$ follows
+{\tiny
 \[
 \begin{array}{l}
-\restrict{r1,r2,r3,r4,\tilde{q},\tilde{w},\tilde{a},\tilde{b}}\\
- (\startchain{\silent}\chainedlink{\overline{a}_i}{\bf  \overline{a}_i}\chainedlink{\bf \overline{a}_o}{\overline{a}_o}\chainedlink{r_2}{r_2}\chainedlink{q_i}{\bf q_i}\chainedlink{\bf q_i}{q_i}\chainedlink{b_i}{\bf b_i}\chainedlink{\bf b_i}{b_i}\chainedlink{\overline{w}_i}{\bf \overline{w}_i}\chainedlink{\bf \overline{w}_o}{\overline{w}_o}\chainedlink{\overline{a}_i}{\bf \overline{a}_i}\chainedlink{\bf \overline{a}_o}{\overline{a}_o}\chainedlink{r_3}{r_3}\chainedlink{\overline{a}_i}{\bf \overline{a}_i}\chainedlink{\bf \overline{a}_o}{\overline{a}_o}\chainedlink{r_4}{r_4}\chainedlink{\overline{w}_i}{\bf \overline{w}_i}\chainedlink{\bf \overline{w}_o}{\overline{w}_o}\chainedlink{cxt}{cxt}\chainedlink{\hat{a}_i}{\bf \hat{a}_i}\chainedlink{\bf \hat{a}_o}{\hat{a}_o}\chainedlink{\underline{b}_i}{\bf \underline{b}_i}\chainedlink{\bf \underline{b}_o}{\underline{b}_o}\chainedlink{p_1}{p_1}\chainedlink{p_2}{p_2}\chainedlink{\tilde{w}_i}{\bf \tilde{w}_i}\chainedlink{\bf \tilde{w}_o}{\tilde{w}_o}\chainedlink{p_3}{p_3}\chainedlink{p_4}{p_4}\chainend{\silent})
+\ell = \\
+%\restrict{r1,r2,r3,r4,\tilde{q},\tilde{w},\tilde{a},\tilde{b}}\\
+ \startchain{\silent}\chainedlink{r_1}{r_1}
+ \chainedlink{\overline{a}_i}{\bf  \overline{a}_i}\chainedlink{\bf \overline{a}_o}{\overline{a}_o}\chainedlink{r_2}{r_2}\chainedlink{q_i}{\bf q_i}\chainedlink{\bf q_i}{q_i}\chainedlink{b_i}{\bf b_i}\chainedlink{\bf b_i}{b_i}\chainedlink{\overline{w}_i}{\bf \overline{w}_i}\chainedlink{\bf \overline{w}_o}{\overline{w}_o}\chainedlink{\overline{a}_i}{\bf \overline{a}_i}\chainedlink{\bf \overline{a}_o}{\overline{a}_o}\chainedlink{r_3}{r_3}\chainedlink{\overline{a}_i}{\bf \overline{a}_i}\chainedlink{\bf \overline{a}_o}{\overline{a}_o}\chainedlink{r_4}{r_4}\chainedlink{\overline{w}_i}{\bf \overline{w}_i}\chainedlink{\bf \overline{w}_o}{\overline{w}_o}\chainedlink{cxt}{cxt}\chainedlink{\hat{a}_i}{\bf \hat{a}_i}\chainedlink{\bf \hat{a}_o}{\hat{a}_o}\chainedlink{\underline{b}_i}{\bf \underline{b}_i}\chainedlink{\bf \underline{b}_o}{\underline{b}_o}\chainedlink{p_1}{p_1}\chainedlink{p_2}{p_2}\chainedlink{\tilde{w}_i}{\bf \tilde{w}_i}\chainedlink{\bf \tilde{w}_o}{\tilde{w}_o}\chainedlink{p_3}{p_3}\chainedlink{p_4}{p_4}\chainend{\silent}
 \end{array}
@@
 the rule $3$ has been not applied as the entity $w$ is absent, then only rule $2$ has been applied, and then it has produced entity $w$.
-Also, in the next state the context will provide entity $a$ and will not entity $b$.
+Also,  the context provides entity $a$, that will be available in the next state, and not entity $b$.
+Now, to let the label more readable, we apply the function $flat(\ell)$ to it, obtaining a more readable and easily processing character sequence:
+\[
+r_1~\overline{a}~r_2~q~b~\overline{w}~\overline{a}~r_3~\overline{a}~r_4~\overline{w}~cxt~\hat{a}~\underline{b}~p_1~p_2~\tilde{w}~p_3~p_4~\silent\]
 
-\subsection{A  biological toy example}
-Here we consider a biological toy example where a gene $a$  codes for a protein $T$ when molecules 
+\subsection{A  biological toy example}\label{subsec:toy}
+
+\noindent
+We consider a biological toy example where a gene $a$  codes for a protein $T$ when molecules 
 $G$ is present and $C$ is absent, and in the opposite situation $a$ codes for protein $T'$.
@@
 \]
-
+\noindent
 Then we have 
@@
 \noindent
-and we put $r_1 = \silent$.
+%and we put $r_1 = \silent$.
 \paragraph{Encoding of the entities.}
@@
 \begin{array}{lcl}
-Sys &\defeq &\restrict{\tilde{C},\tilde{G},\tilde{F}, \tilde{T},\tilde{T'}}( P_a\mid P_C\mid P_G\mid \overline{P}_F \mid \overline{P}_T \mid P_{T'} \mid P_1 \mid P_2 \mid P_3 \mid P_4)
+Sys &\defeq &%\restrict{\tilde{C},\tilde{G},\tilde{F}, \tilde{T},\tilde{T'}}(
+I\mid P_a\mid \overline{P}_C\mid P_G\mid \overline{P}_F \mid \overline{P}_T \mid P_{T'} \mid P_1 \mid P_2 \mid P_3 \mid P_4 \mid Cxt
 \end{array}
@@
 In the above configuration, reactions $1$ and $3$ can be applied, and also we assume that the context will provide the entity $F$, that will be available in the target configuration.
-The transition label of
+Instead of showing the transition label, we give give the string generated by the function $flat$:\
+%The transition label appears as:
+\[
+\tiny
+\begin{array}{l}
+%\restrict{r1,r2,r3,r4,a,\tilde{C},\tilde{G},\tilde{F},\tilde{T},\tilde{T'}}\\
+% \startchain{\silent}\chainedlink{a_i}{\bf  a_i}\chainedlink{\bf a_o}{a_o}
+% \chainedlink{G_i}{\bf  G_i}\chainedlink{\bf G_o}{G_o}
+% \chainedlink{\overline{C}_i}{\bf \overline{C}_i}\chainedlink{\bf \overline{C}_o}{\overline{C}_o}
+% \chainedlink{r_2}{r_2}\chainedlink{G_i}{\bf G_i}\chainedlink{\bf G_o}{G_o}
+%\chainedlink{r_3}{r_3}\chainedlink{T'_i}{\bf T'_i}\chainedlink{\bf T'_o}{T'_o}\chainedlink{r_4}{r_4}\chainedlink{T'_i}{\bf T'_i}\chainedlink{\bf T'_o}{T'_o}\chainedlink{\overline{F}_i}{\bf \overline{F}_i}\chainedlink{\bf \overline{F}_o}{\overline{F}_o}\chainedlink{cxt}{cxt}
+%\chainedlink{\underline{C}_i}{\bf \underline{C}_i}\chainedlink{\bf \underline{C}_o}{\underline{C}_o}
+%\chainedlink{\underline{G}_i}{\bf \underline{G}_i}\chainedlink{\bf \underline{G}_o}{\underline{G}_o}
+%\chainedlink{\hat{F}_i}{\bf \hat{F}_i}\chainedlink{\bf \hat{F}_o}{\hat{F}_o}\chainedlink{p_1}{p_1} 
+% \chainedlink{\tilde{T}_i}{\bf \tilde{T}_i}\chainedlink{\bf \tilde{T}_o}{\tilde{T}_o}
+%  \chainedlink{p_2}{p_2}\chainedlink{p_3}{p_3}\chainedlink{p_4}{p_4} \chainedlink{\tilde{G}_i}{\bf \tilde{G}_i}\chainedlink{\bf \tilde{G}_o}{\tilde{G}_o} \chainend{\silent}
+r_1~ a~G~\overline{C}~ r_2~ G~ r_3~T'~r_4~T'~\overline{F}~cxt~\underline{C}~\underline{G}~\hat{F}~p_1~p_2~p_3~\tilde{G}~\silent
+
+\end{array}
+\]

### 2020-04-01 06b-bisimulation.tex
flat function

--- 2020-03-29 06b-bisimulation.tex
@@
 
-To simplify the query language, we first introduce a function that takes a solid link chain and returns a simple string by eliminating all the channel matching pairs leaving just one channel per each pair.
+%To simplify the query language, we first introduce a function that takes a solid link chain and returns a simple string by eliminating all the channel matching pairs leaving just one channel per each pair.
 This transformation is harmless, in the sense that it retains all the information in the chain, because it is applied to solid chains only.
-The function $\mathit{flat}$ is defined inductively as follows:
-\[
-\begin{array}{lcl@{\hspace{0.5cm}}lcl@{\hspace{0.5cm}}lcl@{\hspace{0.5cm}}lcl}
-\mathit{flat}(\epsilon)&\defeq & \epsilon &
-\mathit{flat}(\link{\silent}{\beta}) & \defeq & \beta &
-\mathit{flat}(\link{\alpha}{\silent})& \defeq & \epsilon &
-\mathit{flat}(\link{\alpha}{\beta}) & \defeq &\beta \\
-\end{array}\]
-\[
-\begin{array}{lcl}
-\mathit{flat}(\link{\alpha}{\beta}\upsilon)& \defeq & \beta :: \mathit{flat}(\upsilon)\\
-\end{array}\\
-\]
-where  $::$ is the usual string concatenation operator.
+Below we denote by $\mathit{decs} = \{ s, \overline{s}, \widetilde{s},\widehat{s},\underline{s}~|~ s\in S\}$ the set of decorated names (without subscripts $_i$ and $_o$).
+%The function $\mathit{flat}$ is defined inductively as follows:
+%%\[
+%%\begin{array}{lcl@{\hspace{0.5cm}}lcl@{\hspace{0.5cm}}lcl@{\hspace{0.5cm}}lcl}
+%%\mathit{flat}(\epsilon)&\defeq & \epsilon &
+%%\mathit{flat}(\link{\gamma}{\gamma'}) &  \defeq & \left \{ \begin{array}{cl} \epsilon & \mathrm{if } (\gamma = \beta_i \wedge \gamma = \beta_o)  \vee \gamma = \silent\\
+%%\gamma' & \mathrm{otw }  \end{array} \right. &
+%%%\mathit{flat}(\link{\alpha}{\silent})& \defeq & \epsilon &
+%%%\mathit{flat}(\link{\alpha}{\beta}) & \defeq &\beta \\
+%%\end{array}\]
+%
+%\[
+%\mathit{flat}(\epsilon)\defeq \epsilon 
+%\qquad
+%\mathit{flat}(\link{\alpha}{\beta}) \defeq 
+%\left \{ \begin{array}{ll} 
+%\beta & \mbox{if $\beta\in \mathit{reacts}\cup\{\mathit{cxt}\}\cup\mathit{prods}$}\\
+%\gamma & \mbox{if $\beta=\gamma_i$ with $\gamma\in\mathit{decs}$}\\
+%\epsilon & \mbox{otherwise}  
+%\end{array} \right. 
+%\]
+%
+%\[
+%\begin{array}{lcl}
+%\mathit{flat}(\link{\alpha}{\beta}\upsilon)& \defeq & \mathit{flat}(\link{\alpha}{\beta}):: \mathit{flat}(\upsilon)\\
+%\end{array}\\
+%\]
+%where  $::$ is the usual string concatenation operator.
+%
 
@@
  for any $v,P'$ such that 
- $P \xrightarrow{v} P'$ then there exist $v',Q'$ such that $Q  \xrightarrow{w} Q'$ with $v\equiv_F w$ and $P' \mathrel{\mathbf{R}_F} Q'$.
+ $P \xrightarrow{v} P'$ then there exist $w,Q'$ such that $Q  \xrightarrow{w} Q'$ with $v\equiv_F w$ and $P' \mathrel{\mathbf{R}_F} Q'$.
 \item

### 2020-04-01 04-CCNA.tex

@@
 
-\paragraph{Link Chains.}
+\paragraph{Link Chains}
 Let $\mathcal{C}$ be the set of channels, ranged over by $a,b,...$, and 
@@
 
-\paragraph{Syntax.}
+\paragraph{Syntax}
 %\begin{definition}
@@
 Here we do not consider name mobility, which is present instead in the {\tt link}-calculus.
-\paragraph{Semantics.}
+
+\paragraph{Semantics}
 

### 2020-04-01 05-translation.tex
Update 05-translation.tex

@@
 
+In the following use the following conventions for denoting different categories of names:
+\begin{itemize}
+\item
+$\mathit{decs} \defeq \{ s, \overline{s}, \widetilde{s},\widehat{s},\underline{s}~|~ s\in S\}$ is the set of names for decorated entities;
+\item
+$\mathit{ents} \defeq \{ d_i, d_o~|~ d\in \mathit{decs}\}$ is the set names for entities;
+\item
+$\mathit{reacts}\defeq \{r_1,...,r_u\}$ is the set of reaction names associated with each reaction $aj$;
+\item 
+$\mathit{prods}\defeq \{p_1,...,p_u\}$ is the set of names for production $p_j$, associated with  reaction $aj$.
+\end{itemize}
+
 \begin{definition}[Translation]
@@
 $$
-where $\mathit{names} = \mathit{reacts}\, \cup\, \mathit{ents}\, \cup \, \mathit{prods}\, \cup\, \{cxt\}$, with  $\mathit{reacts}$ be the set of reaction names $r_j $, $\mathit{cxts}$ the set of context names $\{\mathit{cxt}\}$,  $\mathit{ents}$ the set of decorated entity names  $\{s_i,s_o,\widehat{s}_i,\widehat{s}_o,\widetilde{s}_i,\widetilde{s}_o,\overline{s}_i,\overline{s}_o,\underline{s}_i,\underline{s}_o| s \in S\}$, and  $\mathit{prods}$ be the set of names for production $p_j$ associated to each reaction.
+where $\mathit{names} = \mathit{reacts}\, \cup\, \mathit{ents}\, \cup \, \mathit{prods}\, \cup\, \{cxt\}$. 
 %For notational convenience, we fix that $r_1 = \silent$,  $r_{u+1} = cxt_1$ for $u$ the number of reacts, and $cxt_{w+1} = \silent$ for $w$ the number of entities.
-{\color{red} For technical reasons, we introduce the init process $I \defeq \link{\silent}{r_1}.I$
-to allow the name $r_1$ to be matched; also, 
+For technical reasons, we introduce the init process $I \defeq \link{\silent}{r_1}.I$
+to allow the name $r_1$ to be matched at the start of any chain; also, 
 for notational convenience, we fix that 
 %$r_1 = \silent$,
-$r_{u+1} = \mathit{cxt}$ and $p_{u+1}=\silent$,  for $u$ the number of reacts.}
+$r_{u+1} = \mathit{cxt}$ and $p_{u+1}=\silent$,  for $u$ the cardinality of $\mathit{reacts}$.
 \end{definition}
@@
  %involved in the transition $t$;
-  \item for each reaction $a \in A$ and each entity $s \in S$, each virtual link offered by processes $P_a$ and $\mathit{Cxt}^n$ is  overlapped by exactly one solid link offered by  processes representing entities.
+  \item for each reaction $a \in A$ and each entity $s \in S$, each virtual link offered by processes $P_a$ and $\mathit{Cxt}^1$ is  overlapped by exactly one solid link offered by  processes representing entities.
   %each process $P_s $ (or its alternative version $\overline{P}_s$), with $s \in S$, participates at most  with one solid link for each process $P_a$, and with exactly one solid link for each process $Cxt_s$.%, if $s$ is a \emph{required element} of $A_j$ in $t$.
@@
 
-{\color{red} \begin{example}\label{ex:backbone}
+\begin{example}\label{ex:backbone}
  Let ${\cal A}$ be a RS whose specification contains two entities, $s1$ and $s2$, and the reactions $r_1=(s1,\ , s2)$ and $r_2 = (s2, \ , s1)$ that produce $s2$ if $s1$ is present and $s1$ if $s2$ is present.
-For simplicity we do not consider inhibitors.
-Then, we assume an  extended interactive process $\pi=(\gamma,\delta)$ where the context $\gamma$ provides  $s1$ and $s2$.
- %, for the first $n$ steps.
- Our translation follows:\\% includes the processes:\\
+For simplicity we do not consider inhibitors, but the reader can assume a void inhibitor is present in both reactions.
+Then, we assume an  extended interactive process $\pi=(\gamma,\delta)$ where the context $\gamma$ provides  $s1$ and $s2$ at every step, but we assume that only $s1$ is initially present.
+The corresponding c\CNA \ process is:
+$$\restrict{\mathit{names}} (~I ~|~ P_{s1}~|~\overline{P}_{s2}~|~ P_{r1}~|~P_{r2}~|~\mathit{Cxt})$$
+ where:
  \[
  \begin{array}{l@{\hspace{-0.009cm}}c@{\hspace{-0.009cm}}ll@{\hspace{-0.009cm}}c@{\hspace{-0.009cm}}l}
- P_{r_1} &\defeq& \startchain{r_1}\chainedlink{s1_i}{\noact}\chainedlink{\noact}{s1_o}\chainedlink{r_2}{\noact} \chainedlink{\noact}{p_1}\chainedlink{\tilde{s2}_i}{\noact}\chainedlink{\noact}{\tilde{s2}_o}\chainend{p_2}.P_{r_1} + \dots; &
-   P_{s2} &\defeq& \startchain{\hat{s2}_i}\chainedlink{\hat{s2}_o}{\noact}
-   \chainedlink{\noact} {\tilde{s2}_i}\chainend{\tilde{s2}_o}.P_{s2} + \dots;\\
-  P_{r_2} &\defeq& \startchain{r_2}\chainedlink{\overline{s2}_i}{\noact}\chainedlink{\noact}{\overline{s2}_o}\chainedlink{cxt}{\noact} \chainedlink{\noact}{p_2}\chainend{\silent}.P_{r_2} + \dots; &
- P_{s1} &\defeq& \startchain{s1_i}\chainedlink{s1_o}{\noact}\chainedlink{\noact}{\hat{s1}_i}\chainend{\hat{s1}_o}
+ P_{r_1} &\defeq& \startchain{r_1}\chainedlink{s1_i}{\noact}\chainedlink{\noact}{s1_o}\chainedlink{r_2}{\noact} \chainedlink{\noact}{p_1}\chainedlink{\widetilde{s2}_i}{\noact}\chainedlink{\noact}{\widetilde{s2}_o}\chainend{p_2}.P_{r_1} + \dots; &
+   P_{s1} &\defeq& \startchain{s1_i}\chainedlink{s1_o}{\noact}\chainedlink{\noact}{\widehat{s1}_i}\chainend{\widehat{s1}_o}
  %{\noact}\chainedlink{\noact}{\tilde{s1}_i}\chainedlink{\tilde{s1}_o}
  .P_{s1} + \dots; \\   
- Cxt& \defeq & \startchain{cxt}\chainedlink{\hat{s1}_i}{\noact}\chainedlink{\noact}{\hat{s1}_o}
-  \chainedlink{\hat{s2}_i}{\noact}\chainedlink{\noact}{\hat{s2}_o}\chainend{p_1}.Cxt\\
+  P_{r_2} &\defeq& \dots + \startchain{r_2}\chainedlink{\overline{s2}_i}{\noact}\chainedlink{\noact}{\overline{s2}_o}\chainedlink{cxt}{\noact} \chainedlink{\noact}{p_2}\chainend{\silent}.P_{r_2} + \dots; &
+  \overline{P}_{s2} &\defeq& \startchain{\overline{s2}_i}\chainedlink{\overline{s2}_o}{\noact}
+   \chainedlink{\noact} {\widetilde{s2}_i}\chainend{\widetilde{s2}_o}.P_{s2} + \dots;\\
+ Cxt& \defeq & \startchain{cxt}\chainedlink{\widehat{s1}_i}{\noact}\chainedlink{\noact}{\widehat{s1}_o}
+  \chainedlink{\widehat{s2}_i}{\noact}\chainedlink{\noact}{\widehat{s2}_o}\chainend{p_1}.Cxt\\
 
@@
 \]  \\
-   Now, we assume that $s1$ only is in the initial state of ${\cal A}$; for clarity we partially specify the {\tt link}-calculus code of the processes.
-In Figure~\ref{fig:backbone} we show the  structure of a link chain label related to the execution of a transition of the c\CNA \ system: $\restrict{names} (I | P_{s1}|P_{s2}| P_a|\dots|Cxt)$.
-The yellow blocks are referred to init process $I$, and the processes encoding the reactions, $P_{r_1}$ and $P_{r_2}$, and the context $Cxt$.
+For clarity we show the code of the processes just in part, to focus on the prefixes that will be involved in the first transition of the system.
+In Figure~\ref{fig:backbone} we show the  structure of a link chain label related to the execution of such a transition.
+The yellow blocks are referred to init process $I$, to the processes encoding the reactions, $P_{r_1}$ and $P_{r_2}$, and to the context $\mathit{Cxt}$.
 As the figure puts in evidence, these two kinds of processes determine the structure of the link chain, from end to end, i.e.
@@
 We could say that these processes form the \emph{backbone} of the interaction.
-   In contrast, the processes encoding the entities, $P_{s1}$,  $P_{s2}$, provides the
-    solid links to overlap the virtual links   of the backbone.
+   In contrast, the processes encoding the entities, $P_{s1}$,  $\overline{P}_{s2}$, provides the
+    solid links to be merged with the virtual links of the backbone (i.e. to be plugged in the backbone).
 
@@
  \end{example}
-}
 

### 2020-04-01 02-ReactionSystem.tex
major update
introdotto notazione, spostato flat, riveduto esempio 13

@@
 \end{definition}
+
+Given an extended interactive process $\pi =(\gamma,\delta)$, we denote by $\pi^k$ the shift of $\pi$ starting at the $k$-th state sequence; formally we let $\pi^k=(\gamma^k,\delta^k)$ with
+$\gamma^k=\{C_i'\}_{i\in\mathbb{N}}$, $\delta^k=\{D_i'\}_{i\in\mathbb{N}}$ with 
+  $C_0'=C_k \cup D_k$, $D'_0 = \emptyset$, and   $C_i'=C_{i+k}$, $D_i'= D_{i+k}$ for any $i\geq 1$.

### 2020-04-01 05-translation.tex
major update
introdotto notazione, spostato flat, riveduto esempio 13

@@
 \item
-$\mathit{decs} \defeq \{ s, \overline{s}, \widetilde{s},\widehat{s},\underline{s}~|~ s\in S\}$ is the set of names for decorated entities;
+$\mathit{decs} \defeq \{ s, \overline{s}, \widetilde{s},\widehat{s},\underline{s}~|~ s\in S\}$ is the set of names for decorated entities (without subscripts $_i$ and $_o$);
 \item
@@
   %each process $P_s $ (or its alternative version $\overline{P}_s$), with $s \in S$, participates at most  with one solid link for each process $P_a$, and with exactly one solid link for each process $Cxt_s$.%, if $s$ is a \emph{required element} of $A_j$ in $t$.
-
   \end{enumerate}
   \end{restatable}
+
+The topmost restriction $\restrict{\mathit{names}}$ appearing in the process $\llbracket{\cal A},\gamma\rrbracket$ serves to guarantee that all names appearing in a link of the chain labelling a transition are matched.
+Since all names appearing in any prefix of $\llbracket{\cal A},\gamma\rrbracket$ are restricted, in the transition $\llbracket{\cal A},\gamma\rrbracket \xrightarrow{\restrict{\mathit{names}}\upsilon} P'$ it means that the observation $\restrict{\mathit{names}}\upsilon$ has the form $\link{\silent}{\silent}\dots\link{\silent}{\silent}$, i.e., it is silent, and that $\upsilon$ is solid.
+Later on we will be interested in reasoning about the actual chain $\upsilon$ used in the transition.
+It has the peculiarity to start and end with silent actions and to include all names in $\mathit{reacts}\cup\mathit{prods}\cup\{\mathit{cxt}\}$.
+As a matter of notation we call such chain $\upsilon$ \emph{complete} (solid and with $\silent$ actions at the extremes).
+We will then use $\llparenthesis{\cal A},\gamma\rrparenthesis$ to refer to the encoding without topmost name restrictions, i.e.,
+$$
+\llparenthesis{\cal A},\gamma\rrparenthesis \defeq 
+I~|~\Pi_{s\in C_0}P_{s} ~|~ \Pi_{s \notin C_0}\overline{P_s} ~|~ \Pi_{a\in A} P_a ~|~  \mathit{Cxt}^1
+$$
+and will write $P\xRightarrow{\upsilon} P'$ to mean that $P\xrightarrow{\upsilon} P'$ with $\upsilon$ complete.
 
@@
 Then, we assume an  extended interactive process $\pi=(\gamma,\delta)$ where the context $\gamma$ provides  $s1$ and $s2$ at every step, but we assume that only $s1$ is initially present.
-The corresponding c\CNA \ process is:
-$$\restrict{\mathit{names}} (~I ~|~ P_{s1}~|~\overline{P}_{s2}~|~ P_{r1}~|~P_{r2}~|~\mathit{Cxt})$$
+The corresponding c\CNA \ process is $\llbracket{\cal A},\gamma\rrbracket \defeq \restrict{\mathit{names}} \llparenthesis{\cal A},\gamma\rrparenthesis$, with
+$$
+\llparenthesis{\cal A},\gamma\rrparenthesis \defeq 
+I ~|~ P_{s1}~|~\overline{P}_{s2}~|~ P_{r1}~|~P_{r2}~|~\mathit{Cxt}
+$$
  where:
@@
  .P_{s1} + \dots; \\   
-  P_{r_2} &\defeq& \dots + \startchain{r_2}\chainedlink{\overline{s2}_i}{\noact}\chainedlink{\noact}{\overline{s2}_o}\chainedlink{cxt}{\noact} \chainedlink{\noact}{p_2}\chainend{\silent}.P_{r_2} + \dots; &
+  P_{r_2} &\defeq& \dots + \startchain{r_2}\chainedlink{\overline{s2}_i}{\noact}\chainedlink{\noact}{\overline{s2}_o}\chainedlink{\mathit{cxt}}{\noact} \chainedlink{\noact}{p_2}\chainend{\silent}.P_{r_2} + \dots; &
   \overline{P}_{s2} &\defeq& \startchain{\overline{s2}_i}\chainedlink{\overline{s2}_o}{\noact}
-   \chainedlink{\noact} {\widetilde{s2}_i}\chainend{\widetilde{s2}_o}.P_{s2} + \dots;\\
- Cxt& \defeq & \startchain{cxt}\chainedlink{\widehat{s1}_i}{\noact}\chainedlink{\noact}{\widehat{s1}_o}
-  \chainedlink{\widehat{s2}_i}{\noact}\chainedlink{\noact}{\widehat{s2}_o}\chainend{p_1}.Cxt\\
-
+   \chainedlink{\noact} {\widehat{s2}_i}\chainedlink{\widehat{s2}_o}{\noact}\chainedlink{\noact} {\widetilde{s2}_i}\chainend{\widetilde{s2}_o}.P_{s2} + \dots;\\
+ \mathit{Cxt}& \defeq & \startchain{\mathit{cxt}}\chainedlink{\widehat{s1}_i}{\noact}\chainedlink{\noact}{\widehat{s1}_o}
+  \chainedlink{\widehat{s2}_i}{\noact}\chainedlink{\noact}{\widehat{s2}_o}\chainend{p_1}.\mathit{Cxt}
   %Cxt& \defeq & \startchain{cxt}\chainedlink{\hat{s1}_i}{\noact}\chainedlink{\noact}{\hat{s1}_o}
@@
 \]  \\
-For clarity we show the code of the processes just in part, to focus on the prefixes that will be involved in the first transition of the system.
+For clarity of exposition, we show the code of the processes just in part, to focus on the prefixes that will be involved in the first transition of the system.
 In Figure~\ref{fig:backbone} we show the  structure of a link chain label related to the execution of such a transition.
@@
     solid links to be merged with the virtual links of the backbone (i.e. to be plugged in the backbone).
+ Formally, we have
+ $$
+ \llparenthesis{\cal A},\gamma\rrparenthesis \xRightarrow{
+ \startchain{\silent}
+ \chainedlink{r_1}{r_1}
+ \chainedlink{s1_i}{s1_i}
+ \chainedlink{s1_o}{s1_o}
+ \chainedlink{r_2}{r_2}
+ \chainedlink{\overline{s2}_i}{\overline{s2}_i}
+ \chainedlink{\overline{s2}_o}{\overline{s2}_o}
+ \chainedlink{\mathit{cxt}}{\mathit{cxt}}
+ \chainedlink{\widehat{s1}_i}{\widehat{s1}_i}
+ \chainedlink{\widehat{s1}_o}{\widehat{s1}_o}
+  \chainedlink{\widehat{s2}_i}{\widehat{s2}_i}
+  \chainedlink{\widehat{s2}_o}{\widehat{s2}_o}
+  \chainedlink{p_1}{p_1}
+  \chainedlink{\widetilde{s2}_i}{\widetilde{s2}_i}
+  \chainedlink{\widetilde{s2}_o}{\widetilde{s2}_o}
+  \chainedlink{p_2}{p_2}
+  \chainend{\silent}
+ }
+ P'
+ $$
 
@@
 
+\paragraph{The flat function}
+For technical reasons, our transition labels are quite verbose; then, to simplify their processing, we introduce a function that takes a solid link chain and returns a simple string by eliminating all the channel matching pairs leaving just one placeholder for them.
+This transformation is harmless, in the sense that it retains all the information in the chain, because it is applied to complete chains only.
+The function $\mathit{flat}$ is defined inductively as follows:
+%\[
+%\begin{array}{lcl@{\hspace{0.5cm}}lcl@{\hspace{0.5cm}}lcl@{\hspace{0.5cm}}lcl}
+%\mathit{flat}(\epsilon)&\defeq & \epsilon &
+%\mathit{flat}(\link{\gamma}{\gamma'}) &  \defeq & \left \{ \begin{array}{cl} \epsilon & \mathrm{if } (\gamma = \beta_i \wedge \gamma = \beta_o)  \vee \gamma = \silent\\
+%\gamma' & \mathrm{otw }  \end{array} \right. &
+%%\mathit{flat}(\link{\alpha}{\silent})& \defeq & \epsilon &
+%%\mathit{flat}(\link{\alpha}{\beta}) & \defeq &\beta \\
+%\end{array}\]
+
+\[
+\mathit{flat}(\epsilon)\defeq \epsilon 
+\qquad
+\mathit{flat}(\link{\alpha}{\beta}) \defeq 
+\left \{ \begin{array}{ll} 
+\beta & \mbox{if $\beta\in \mathit{reacts}\cup\{\mathit{cxt}\}\cup\mathit{prods}$}\\
+d & \mbox{if $\beta=d_i$ with $d\in\mathit{decs}$}\\
+\epsilon & \mbox{otherwise}  
+\end{array} \right. 
+\]
+
+\[
+\begin{array}{lcl}
+\mathit{flat}(\link{\alpha}{\beta}\upsilon)& \defeq & \mathit{flat}(\link{\alpha}{\beta}):: \mathit{flat}(\upsilon)\\
+\end{array}\\
+\]
+where  $::$ is the usual string concatenation operator.
+
+For example, if we consider again the complete label
+$$
+\upsilon =  \startchain{\silent}
+ \chainedlink{r_1}{r_1}
+ \chainedlink{s1_i}{s1_i}
+ \chainedlink{s1_o}{s1_o}
+ \chainedlink{r_2}{r_2}
+ \chainedlink{\overline{s2}_i}{\overline{s2}_i}
+ \chainedlink{\overline{s2}_o}{\overline{s2}_o}
+ \chainedlink{\mathit{cxt}}{\mathit{cxt}}
+ \chainedlink{\widehat{s1}_i}{\widehat{s1}_i}
+ \chainedlink{\widehat{s1}_o}{\widehat{s1}_o}
+  \chainedlink{\widehat{s2}_i}{\widehat{s2}_i}
+  \chainedlink{\widehat{s2}_o}{\widehat{s2}_o}
+  \chainedlink{p_1}{p_1}
+  \chainedlink{\widetilde{s2}_i}{\widetilde{s2}_i}
+  \chainedlink{\widetilde{s2}_o}{\widetilde{s2}_o}
+  \chainedlink{p_2}{p_2}
+  \chainend{\silent}
+$$
+from Example~\ref{ex:backbone}, we have
+$$
+\mathit{flat}(\upsilon) = 
+r_1\ 
+s1\ 
+r_2\
+\overline{s2}\ 
+\mathit{cxt}\ 
+\widehat{s1}\ 
+\widehat{s2}\ 
+p_1\ 
+\widetilde{s2}\ 
+p_2.
+$$
+
+Viceversa, it is immediate to define the function $\mathit{unflat}$ to rebuild the complete label from the compact string (here we exploit again the half link and block notation):
+$$
+\mathit{unflat}(x) \defeq\left\{
+\begin{array}{ll}
+^{x}_{x} & \mbox{if $x\in \mathit{reacts}\cup\{\mathit{cxt}\}\cup\mathit{prods}$}\\
+^{x_i}_{x_i}\backslash^{x_o}_{x_o} & \mbox{if $x\in \mathit{decs}$}\\
+\end{array}
+\right.
+$$
+$$
+\mathit{unflat}(x_1\dots x_n) \defeq\ ^{\silent}\backslash \mathit{unflat}(x_1) \backslash \dots \backslash \mathit{unflat}(x_1) \backslash_{\silent}
+$$
+
+It is immediate to check that for any complete label $\upsilon$ of our processes we have
+$\upsilon = \mathit{unflat}(\mathit{flat}(\upsilon))$.
+
 With the next proposition, we analyse the structure of a c\CNA \ 
@@
 In the following four statements, for brevity, we let ${\cal A} = (S, A)$ be a RS, and let  $\pi=(\gamma,\delta)$ be an extended interactive process in $A$, with $\gamma=\{C_i\}_{i\in\mathbb{N}}$ and $\delta=\{D_i\}_{i\in\mathbb{N}}$.
-Moreover, we denote by $\pi^j$ the shift of $\pi$ starting at the $j$-th state sequence; formally we let $\pi^j=(\gamma^j,\delta^j)$ with
-$\gamma^j=\{C_i'\}_{i\in\mathbb{N}}$, $\delta^j=\{D_i'\}_{i\in\mathbb{N}}$ with 
-  $C_0'=C_j \cup D_j$, $D'_0 = \emptyset$, and   $C_i'=C_{i+j}$, $D_i'= D_{i+j}$ for any $i\geq 1$.
 
@@
 
-\paragraph{The flat function}
-For technical reasons, our transition labels are quite verbose; then, to simplify their processing, we introduce a function that takes a solid link chain and returns a simple string by eliminating all the channel matching pairs leaving just one channel per each pair.
-This transformation is harmless, in the sense that it retains all the information in the chain, because it is applied to solid chains only.
-Below we denote by $\mathit{decs} = \{ s, \overline{s}, \widetilde{s},\widehat{s},\underline{s}~|~ s\in S\}$ the set of decorated names (without subscripts $_i$ and $_o$).
-The function $\mathit{flat}$ is defined inductively as follows:
-%\[
-%\begin{array}{lcl@{\hspace{0.5cm}}lcl@{\hspace{0.5cm}}lcl@{\hspace{0.5cm}}lcl}
-%\mathit{flat}(\epsilon)&\defeq & \epsilon &
-%\mathit{flat}(\link{\gamma}{\gamma'}) &  \defeq & \left \{ \begin{array}{cl} \epsilon & \mathrm{if } (\gamma = \beta_i \wedge \gamma = \beta_o)  \vee \gamma = \silent\\
-%\gamma' & \mathrm{otw }  \end{array} \right. &
-%%\mathit{flat}(\link{\alpha}{\silent})& \defeq & \epsilon &
-%%\mathit{flat}(\link{\alpha}{\beta}) & \defeq &\beta \\
-%\end{array}\]
-
-\[
-\mathit{flat}(\epsilon)\defeq \epsilon 
-\qquad
-\mathit{flat}(\link{\alpha}{\beta}) \defeq 
-\left \{ \begin{array}{ll} 
-\beta & \mbox{if $\beta\in \mathit{reacts}\cup\{\mathit{cxt}\}\cup\mathit{prods}$}\\
-\gamma & \mbox{if $\beta=\gamma_i$ with $\gamma\in\mathit{decs}$}\\
-\epsilon & \mbox{otherwise}  
-\end{array} \right. 
-\]
-
-\[
-\begin{array}{lcl}
-\mathit{flat}(\link{\alpha}{\beta}\upsilon)& \defeq & \mathit{flat}(\link{\alpha}{\beta}):: \mathit{flat}(\upsilon)\\
-\end{array}\\
-\]
-where  $::$ is the usual string concatenation operator.
-
 % 

### 2020-04-01 05-translation.tex
major rewriting of examples
check carefully!

@@
 \begin{array}{lcl@{\hskip 1cm}lcl}
- P_s &\defeq & P(s,\widetilde{s},\widehat{s},\underline{s}) 
+ P_s &\defeq & E(s,\widetilde{s},\widehat{s},\underline{s}) 
  &
- \overline{P_s}& \defeq & P(\overline{s},\widetilde{s},\widehat{s},\underline{s})
+ \overline{P_s}& \defeq & E(\overline{s},\widetilde{s},\widehat{s},\underline{s})
 \end{array}
@@
 \begin{array}{rcl}
- P(s,\widetilde{s},\widehat{s},\underline{s}) & \defeq &\sum_{h,k\geq 0}  
+ E(s,\widetilde{s},\widehat{s},\underline{s}) & \defeq &\sum_{h,k\geq 0}  
   (\startchain{s_i}\chainedlink{s_o}{\noact}\chainend{\noact})^h \ \startchain{\widehat{s}_i}\chainedlink{\widehat{s}_o}{\noact}\chainend{\noact} (\startchain{\widetilde{s}_i}\chainedlink{\widetilde{s}_o}{\noact}\chainend{\noact})^k.P_s\\
@@
 \]
-The first line of $ P(s,\widetilde{s},\widehat{s},\underline{s})$ accounts for the case where $s$ is tested for presence by $h$ reactions and produced by $k$ reactions, while being provided by the context ($\link{\widehat{s}_i}{\widehat{s}_o}$).
+The first line of $ E(s,\widetilde{s},\widehat{s},\underline{s})$ accounts for the case where $s$ is tested for presence by $h$ reactions and produced by $k$ reactions, while being provided by the context ($\link{\widehat{s}_i}{\widehat{s}_o}$).
 Thus, $s$ will be present at the next step (the continuation is $P_s$). Here $h$ and $k$ are not known a priori and therefore any
@@
 \triangleq
- \restrict{\mathit{names}}(I~|~\Pi_{s\in C_0}P_{s} ~|~ \Pi_{s \notin C_0}\overline{P_s} ~|~ \Pi_{a\in A} P_a ~|~  \mathit{Cxt}^1)
+ \restrict{\mathit{names}}\left(I ~|~ \prod_{a\in A} P_a ~|~  \mathit{Cxt}^1 ~|~\prod_{s\in C_0}P_{s} ~|~ \prod_{s \notin C_0}\overline{P_s}\right)
 $$
@@
 \llparenthesis{\cal A},\gamma\rrparenthesis \defeq 
-I~|~\Pi_{s\in C_0}P_{s} ~|~ \Pi_{s \notin C_0}\overline{P_s} ~|~ \Pi_{a\in A} P_a ~|~  \mathit{Cxt}^1
+I~|~ \prod_{a\in A} P_a ~|~  \mathit{Cxt}^1~|~\prod_{s\in C_0}P_{s} ~|~ \prod_{s \notin C_0}\overline{P_s} 
 $$
@@
  $P=\llbracket{\cal A},\gamma\rrbracket$ with
-$$P  = \restrict{\mathit{names}} (I~|~\Pi_{a \in A}P_a ~|~ \mathit{Cxt}^1 ~|~ \Pi_{s \in C_0}P_s ~|~ \Pi_{s \notin C_0}\overline{P}_s).
+$$P  = \restrict{\mathit{names}} \left(I~|~\prod_{a \in A}P_a ~|~ \mathit{Cxt}^1 ~|~ \prod_{s \in C_0}P_s ~|~ \prod_{s \notin C_0}\overline{P}_s\right).
 $$
@@
   \item 
- $P'= \restrict{\mathit{names}} (I~|~\Pi_{a \in A}P_a~|~ \mathit{Cxt}^2 ~|~ \Pi_{s \in C_1 \cup D_1}P_s| \Pi_{s \notin C_1 \cup D_1}\overline{P}_s)$.
+ $P'= \restrict{\mathit{names}} \left(I~|~\prod_{a \in A}P_a~|~ \mathit{Cxt}^2 ~|~ \prod_{s \in C_1 \cup D_1}P_s| \prod_{s \notin C_1 \cup D_1}\overline{P}_s\right)$.
  \end{enumerate}

### 2020-04-01 06-examples.tex
major update
introdotto notazione, spostato flat, riveduto esempio 13
major rewriting of examples
check carefully!

@@
 \label{ex:examples}
-Hereafter, we only take into account transitions whose labels are solid, i.e. they do not contain 
-virtual links, and   complete, i.e. they starts and ends with the $\silent$ symbol.\\
+
+Hereafter, we shall often omit topmost restrictions but shall take into account only transitions whose labels are complete, i.e. they do not contain 
+virtual links and start/end with the $\silent$ action symbol.
 
@@
 \noindent
-In our chained {\tt link}-calculus encoding, $q$, $w$, $a$, and $b$ became the $entities$, the context can only provide $a$ and $b$ and the reaction rules are as follows:\\
+In our $c\CNA$ encoding, $q$, $w$, $a$, and $b$ became the $entities$, the context can only provide $a$ and $b$ and the reaction rules are as follows:\\
 \[
 \begin{array}{cl@{\hspace{1cm}} cl}
-1 \ &(\{q,a\},\{b\},q) & 2\ &(\{q,b\},\{a\},w)\\
-3 \  &(\{w,a\},\{b\},q) & 4\ &(\{w,b\},\{a\},w)\\ 
+1 \ &(\{q,a\},\{w,b\},\{q\}) & 2\ &(\{q,b\},\{w,a\},\{w\})\\
+3 \  &(\{w,a\},\{q,b\},\{q\}) & 4\ &(\{w,b\},\{q,a\},\{w\})\\ 
 \end{array}
@@
 \begin{array}{lcl}
-P_n(q,b,w,a,q) & \defeq & \startchain{r_n}\chainedlink{q_i}{\noact}\chainedlink{\noact}{q_o}
-                                                    \chainedlink{b_i}{\noact}\chainedlink{\noact}{b_o}
-                                                     \chainedlink{\overline{w}_i}{\noact}\chainedlink{\noact}{\overline{w}_o}
-					        \chainedlink{\overline{a}_i}{\noact}\chainedlink{\noact}{\overline{a}_o}
-					        \chainedlink{r_n}{\noact}\chainedlink{\noact}{p_n}
-					        \chainedlink{\tilde{q}_i}{\noact}\chainedlink{\noact}{\tilde{q}_o}
-			\chainend{p_{n+1}}.P_n(q,b,w,a,q)  \\
+P_n(q,b,w,a,q) & \defeq & \pi_n(q,b,w,a,q).P_n(q,b,w,a,q)  \\
 			&& +\\
-			&&P_n'(\overline{q}) \ +\ P_n'(\overline{b}) \ + \ P_n'(w)\ +\ P_n'(a)\\
+			&& \sum_{x\in\{\overline{q},\overline{b},w,a\}}\ \pi'_n(x).P_n(q,b,w,a,q)\\
+%			&& +\ \pi'(\overline{b}).P_n(q,b,w,a,q)\\
+%			&& +\ \pi'(w).P_n(q,b,w,a,q)\\ 
+%			&& +\ \pi'(a).P_n(q,b,w,a,q)\\
 			\end{array}
@@
 \begin{array}{lcl}
-P_n'(w) &\defeq &  \startchain{r_n}\chainedlink{w_i}{\noact}\chainedlink{\noact}{w_o}\chainedlink{r_{n+1}}{\noact}\chainedlink{\noact}{p_n} \chainend{p_{n+1}}.P_n'(w) 
+\pi_n(q,b,w,a,q) &\defeq & \startchain{r_n}\chainedlink{q_i}{\noact}\chainedlink{\noact}{q_o}
+                                                    \chainedlink{b_i}{\noact}\chainedlink{\noact}{b_o}
+                                                     \chainedlink{\overline{w}_i}{\noact}\chainedlink{\noact}{\overline{w}_o}
+					        \chainedlink{\overline{a}_i}{\noact}\chainedlink{\noact}{\overline{a}_o}
+					        \chainedlink{r_{n+1}}{\noact}\chainedlink{\noact}{p_n}
+					        \chainedlink{\widetilde{q}_i}{\noact}\chainedlink{\noact}{\widetilde{q}_o}
+			\chainend{p_{n+1}}\\
+\pi'_n(x) &\defeq &  \startchain{r_n}\chainedlink{x_i}{\noact}\chainedlink{\noact}{x_o}\chainedlink{r_{n+1}}{\noact}\chainedlink{\noact}{p_n} \chainend{p_{n+1}} 
 % \ + \   \startchain{r_n}\chainedlink{w_i}{\noact}\chainedlink{\noact}{w_o}\chainedlink{r_{n+1}}{\noact}\chainedlink{\noact}{p_n} \chainend{p_{n+1}}.P_n(q,b,w,a,q) \\
@@
 \noindent
-and we put 
+and we put, as usual,
 %$r_1 = \silent$, 
-$r_5 = cxt$ and  $p_5 = \silent$.
+$r_5 = \mathit{cxt}$ and  $p_5 = \silent$.
 
-\paragraph{Encoding of the entities.}
+\paragraph{Encoding of the entities}
 As for reactions, also the encoding of the entities is given in a parametric way.
-Here we differentiate the encoding for the entities that are not provided by the context and that can be produced by the reactions, and the ones that can be provided by the context and that are not produced by the reactions.\\
-Here, the entities $p$ and $w$ that are not provided by the context:
+Here we differentiate the encoding for the entities that are not provided by the context and that can be produced by the reactions, and the ones that can be provided by the context and that are not produced by the reactions.
+
+Here, for the entities $q$ and $w$ that are not provided by the context, we let:
 \[
 \begin{array}{lcl@{\hspace{1cm}}lcl}
-P (q)&\defeq & \sum_{h=0}^1 (\startchain{q_i}\chainedlink{q_o}{\noact}\chainend{\noact})^h\ \link{\tilde{q}_i}{\tilde{q}_o}.P(q) &
-\overline{P} (q)&\defeq & \sum_{h=1}^2 (\startchain{\overline{q}_i}\chainedlink{\overline{q}_o}{\noact}\chainend{\noact})^h\ \link{\tilde{q}_i}{\tilde{q}_o}.P(q)\\
-&& + & && +\\
-&& \link{q_i}{q_o}.\overline{P}(q) &
-&& \sum_{h=1}^2 (\startchain{\overline{q}_i}\chainedlink{\overline{q}_o}{\noact}\chainend{\noact})^h.\overline{P}(q)\\
+P_q & \defeq & E(q,\widetilde{q}) & \overline{P}_q & \defeq & E(\overline{q},\widetilde{q}) \\
+P_w& \defeq & E(w,\widetilde{w}) & \overline{P}_w & \defeq & E(\overline{w},\widetilde{w})
 \end{array}
 \]
-Then, we have
+where:
 \[
-\begin{array}{lcl @{\hspace{1cm}}lcl}
-P_q & \defeq & P(q) & P_w& \defeq & P(w)\\
+\begin{array}{lcl}
+E(q,\widetilde{q})&\defeq & \sum_{h=1}^3 (\startchain{q_i}\chainedlink{q_o}{\noact}\chainend{\noact})^h\ \link{\widetilde{q}_i}{\widetilde{q}_o}.P_q\\
+&& +\\
+&& \sum_{h=1}^3 (\startchain{q_i}\chainedlink{q_o}{\noact}\chainend{\noact})^h.\overline{P}_q\\
 \end{array}
 \]
-Here, the entities that are provided by the context:
+In fact the presence/absence of $q$ and $w$ will be exploited by at least one rule and as most three rules.
+
+Here, for the entities $a$ and $b$ that can be provided by the context but not produced by the rules, we let:
 \[
-\begin{array}{lcl @{\hspace{1cm}}lcl}
-E(a) & \defeq &  \sum_{h=0}^1 (\startchain{a_i}\chainedlink{a_o}{\noact}\chainend{\noact})^h\ \link{\hat{a}_i}{\hat{a}_o}.E(a) &  \overline{E} (a)&\defeq & \sum_{h=1}^2 (\startchain{\overline{a}_i}\chainedlink{\overline{a}_o}{\noact}\chainend{\noact})^h\ \link{\hat{a}_i}{\hat{a}_o}.E(a)\\
-&& + &  && +\\
-&& \sum_{h=0}^1 (\startchain{a_i}\chainedlink{a_o}{\noact}\chainend{\noact})^h\ \link{\underline{a}_i}{\underline{a}_o}.\overline{E}(a) &
-&& \sum_{h=1}^2 (\startchain{\overline{a}_i}\chainedlink{\overline{a}_o}{\noact}\chainend{\noact})^h\ \link{\underline{a}_i}{\underline{a}_o}.\overline{E}(q)\\
+\begin{array}{lcl@{\hspace{1cm}}lcl}
+P_a & \defeq & E(a,\widehat{a},\underline{a}) & 
+\overline{P}_a & \defeq & E(\overline{a},\widehat{a},\underline{a}) \\
+P_b& \defeq & E(b,\widehat{b},\underline{b})&
+\overline{P}_b& \defeq & E(\overline{b},\widehat{b},\underline{b})
 \end{array}
 \]
-Then, we have
+where:
 \[
-\begin{array}{lcl @{\hspace{1cm}}lcl}
-P_a & \defeq & E(a) & P_b& \defeq & E(b)\\
+\begin{array}{rcl}
+ E(a,\widehat{a},\underline{a}) & \defeq &\sum_{h=1}^3  
+  (\startchain{a_i}\chainedlink{a_o}{\noact}\chainend{\noact})^h \ \startchain{\widehat{a}_i}\chainend{\widehat{a}_o}.P_a\\
+&&\ + \\
+&&
+\sum_{h=1}^3
+ (\startchain{a_i}\chainedlink{a_o}{\noact}\chainend{\noact})^h \ \startchain{\underline{a}_i}\chainend{\underline{a}_o}.\overline{P}_a
 \end{array}
 \]
-For the context, the encoding follows:
+
+Finally, for the context, the encoding follows:
 \[
 \begin{array}{lcl @{\hspace{0.5cm}}c@{\hspace{0.5cm}} l}
-Cxt &\defeq &  \startchain{cxt}\chainedlink{\hat{a}_i}{\noact}\chainedlink{\noact}{\hat{a}_o}\chainedlink{\underline{b}_i}{\noact}\chainedlink{\noact}{\underline{b}_i}\chainend{p_1}.Cxt
+\mathit{Cxt} &\defeq &  \startchain{\mathit{cxt}}\chainedlink{\widehat{a}_i}{\noact}\chainedlink{\noact}{\widehat{a}_o}\chainedlink{\underline{b}_i}{\noact}\chainedlink{\noact}{\underline{b}_o}\chainend{p_1}.\mathit{Cxt}
 &+
-&\startchain{cxt}\chainedlink{\hat{b}_i}{\noact}\chainedlink{\noact}{\hat{b}_o}\chainedlink{\underline{a}_i}{\noact}\chainedlink{\noact}{\underline{a}_i}\chainend{p_1}.Cxt\\
+&\startchain{\mathit{cxt}}\chainedlink{\widehat{b}_i}{\noact}\chainedlink{\noact}{\widehat{b}_o}\chainedlink{\underline{a}_i}{\noact}\chainedlink{\noact}{\underline{a}_o}\chainend{p_1}.\mathit{Cxt}\\
 \end{array}
 \]
-We model the context to always offer either  $a$ or $b$, and no both the entities. The reason is that in
-the other cases (providing both $a$ and $b$ or neither of them) leads the system to be stuck.\\
-Now, we assume that we have an initial configuration containing the entities $q$ and $b$ (with $\tilde{q},\tilde{w},\tilde{a},\tilde{b}$ grouping all the possibile decorations for these names):
+
+Notice that we exploit here the capabilities of the process algebraic framework to define a nondeterministic, recursive context:
+We model the context to always offer either $a$ or $b$, but never both the entities together. The reason is that in
+the other cases (providing both $a$ and $b$ or neither of them) would lead the system to be stuck because of the simplifications we have adopted in the other processes.
+
+Now, we assume that we have an initial configuration containing the entities $q$ and $b$:
+% (with $\widetilde{q},\widetilde{w},\widetilde{a},\widetilde{b}$ grouping all the possibile decorations for these names):
 \[
 \begin{array}{lcl}
-Sys  \defeq \ %\restrict{r1,r2,r3,r4,\tilde{q},\tilde{w},\tilde{a},\tilde{b}}
- I\mid P_q \mid \overline{P}_w  \mid  P_a  \mid \overline{P}_b  \mid  P_1  \mid  P_2  \mid  P_3  \mid P_4  \mid Cxt.
+\mathit{Sys}  \defeq \ %\restrict{r1,r2,r3,r4,\widetilde{q},\widetilde{w},\widetilde{a},\widetilde{b}}
+ I\mid P_q \mid \overline{P}_w  \mid  \overline{P}_a  \mid P_b  \mid  P_1  \mid  P_2  \mid  P_3  \mid P_4  \mid \mathit{Cxt}.
 \end{array}
 \] 
-Then, only the second reaction can be applied, and the label transition $\ell$ follows
+Then, only the second reaction can be applied, and the transition carries the complete label $\upsilon$ below
 {\tiny
 \[
-\begin{array}{l}
-\ell = \\
-%\restrict{r1,r2,r3,r4,\tilde{q},\tilde{w},\tilde{a},\tilde{b}}\\
- \startchain{\silent}\chainedlink{r_1}{r_1}
- \chainedlink{\overline{a}_i}{\bf  \overline{a}_i}\chainedlink{\bf \overline{a}_o}{\overline{a}_o}\chainedlink{r_2}{r_2}\chainedlink{q_i}{\bf q_i}\chainedlink{\bf q_i}{q_i}\chainedlink{b_i}{\bf b_i}\chainedlink{\bf b_i}{b_i}\chainedlink{\overline{w}_i}{\bf \overline{w}_i}\chainedlink{\bf \overline{w}_o}{\overline{w}_o}\chainedlink{\overline{a}_i}{\bf \overline{a}_i}\chainedlink{\bf \overline{a}_o}{\overline{a}_o}\chainedlink{r_3}{r_3}\chainedlink{\overline{a}_i}{\bf \overline{a}_i}\chainedlink{\bf \overline{a}_o}{\overline{a}_o}\chainedlink{r_4}{r_4}\chainedlink{\overline{w}_i}{\bf \overline{w}_i}\chainedlink{\bf \overline{w}_o}{\overline{w}_o}\chainedlink{cxt}{cxt}\chainedlink{\hat{a}_i}{\bf \hat{a}_i}\chainedlink{\bf \hat{a}_o}{\hat{a}_o}\chainedlink{\underline{b}_i}{\bf \underline{b}_i}\chainedlink{\bf \underline{b}_o}{\underline{b}_o}\chainedlink{p_1}{p_1}\chainedlink{p_2}{p_2}\chainedlink{\tilde{w}_i}{\bf \tilde{w}_i}\chainedlink{\bf \tilde{w}_o}{\tilde{w}_o}\chainedlink{p_3}{p_3}\chainedlink{p_4}{p_4}\chainend{\silent}
-\end{array}
+%\begin{array}{l}
+%\upsilon = \\
+%\restrict{r1,r2,r3,r4,\widetilde{q},\widetilde{w},\widetilde{a},\widetilde{b}}\\
+ \startchain{\silent}
+ \chainedlink{r_1}{r_1}
+ \chainedlink{\overline{a}_i}{\bf  \overline{a}_i}\chainedlink{\bf \overline{a}_o}{\overline{a}_o}
+ \chainedlink{r_2}{r_2}
+ \chainedlink{q_i}{\bf q_i}\chainedlink{\bf q_o}{q_o}
+ \chainedlink{b_i}{\bf b_i}\chainedlink{\bf b_o}{b_io}
+ \chainedlink{\overline{w}_i}{\bf \overline{w}_i}\chainedlink{\bf \overline{w}_o}{\overline{w}_o}
+ \chainedlink{\overline{a}_i}{\bf \overline{a}_i}\chainedlink{\bf \overline{a}_o}{\overline{a}_o}
+ \chainedlink{r_3}{r_3}
+ \chainedlink{\overline{a}_i}{\bf \overline{a}_i}\chainedlink{\bf \overline{a}_o}{\overline{a}_o}
+ \chainedlink{r_4}{r_4}
+ \chainedlink{\overline{w}_i}{\bf \overline{w}_i}\chainedlink{\bf \overline{w}_o}{\overline{w}_o}
+ \chainedlink{\mathit{cxt}}{\mathit{cxt}}
+ \chainedlink{\widehat{a}_i}{\bf \widehat{a}_i}\chainedlink{\bf \widehat{a}_o}{\widehat{a}_o}
+ \chainedlink{\underline{b}_i}{\bf \underline{b}_i}\chainedlink{\bf \underline{b}_o}{\underline{b}_o}
+ \chainedlink{p_1}{p_1}
+ \chainedlink{p_2}{p_2}
+ \chainedlink{\widetilde{w}_i}{\bf \widetilde{w}_i}\chainedlink{\bf \widetilde{w}_o}{\widetilde{w}_o}
+ \chainedlink{p_3}{p_3}
+ \chainedlink{p_4}{p_4}
+ \chainend{\silent}
+%\end{array}
 \]}
-The parts in bold are provided by the entity processes, the other parts are provided by the processes encoding the reactions and by the process encoding the context (starting at $cxt$ and ended at $p_1$.
-In the label we can read that the rules $1$ and $4$ have been not executed as the entity $a$ is absent, 
-the rule $3$ has been not applied as the entity $w$ is absent, then only rule $2$ has been applied, and then it has produced entity $w$.
-Also,  the context provides entity $a$, that will be available in the next state, and not entity $b$.
-Now, to let the label more readable, we apply the function $flat(\ell)$ to it, obtaining a more readable and easily processing character sequence:
+The parts in bold are provided by the entity processes, the other parts are provided by the processes encoding the reactions and by the process encoding the context (starting at $\mathit{cxt}$ and ended at $p_1$.
+In the label we can read that the rules $1$ and $4$ have been not executed because the entity $a$ is absent, 
+the rule $3$ has been not applied because the entity $w$ is absent, then only rule $2$ has been applied, and it has produced the entity $w$.
+Also,  the context provides entity $a$, that will be available in the next state, and not the entity $b$.
+Now, to let the label more readable, we show the result of the application of the function $flat(\cdot)$ to it:
 \[
-r_1~\overline{a}~r_2~q~b~\overline{w}~\overline{a}~r_3~\overline{a}~r_4~\overline{w}~cxt~\hat{a}~\underline{b}~p_1~p_2~\tilde{w}~p_3~p_4~\silent\]
+r_1~\overline{a}~r_2~q~b~\overline{w}~\overline{a}~r_3~\overline{a}~r_4~\overline{w}~\mathit{cxt}~\widehat{a}~\underline{b}~p_1~p_2~\widetilde{w}~p_3~p_4~\silent\]
 
@@
 This behavior is encoded in rules $1$ and $2$.
-Then, rule three codes for the production of $C$ when proteins $T$ and $F$ are present, and $T'$ absent;
-rule four codes for the production of $G$ when  proteins $T'$ is present and $F$ is absent.
-\paragraph{Encoding of the rules.}
+Then, rule 3 codes for the production of $C$ when proteins $T$ and $F$ are present, and $T'$ absent;
+rule 4 codes for the production of $G$ when  proteins $T'$ is present and $F$ is absent.
+
+\paragraph{Encoding of the rules}
+
 The encoding of the rules for reactions is  given in a parametric way:
@@
 \begin{array}{lcl}
-P_n(a,G,C,T) & \defeq &\startchain{r_n}\chainedlink{a_i}{\noact}\chainedlink{\noact}{a_o}
+P_n(a,G,C,T) & \defeq &\pi_n(a,G,C,T).P_n(a,G,C,T)  \\
+				&& +\\
+			&&\sum_{x\in \{a,G,C\}} \pi'_n(x).P_n(a,G,C,T)
+\end{array}
+\]
+where
+\[
+\begin{array}{lcl}
+\pi_n(a,G,C,T) & \defeq & \startchain{r_n}\chainedlink{a_i}{\noact}\chainedlink{\noact}{a_o}
                                                     \chainedlink{G_i}{\noact}\chainedlink{\noact}{G_o}
@@
 					        \chainedlink{r_{n+1}}{\noact}\chainedlink{\noact}{p_n}
-					        \chainedlink{\tilde{T}_i}{\noact}\chainedlink{\noact}{\tilde{T}_o}
-			\chainend{p_{n+1}}.P_n(a,G,C,T)  \\
-				&& +\\
-			&&P_n'(\overline{a}) \ +\ P_n'(\overline{G}) \ + \ P_n'(C)\\
-			\end{array}
-			\]
-			\noindent
-			where
-			\[
-\begin{array}{lcl}
-P_n'(C) &\defeq &  \startchain{r_n}\chainedlink{C_i}{\noact}\chainedlink{\noact}{C_o}\chainedlink{r_{n+1}}{\noact}\chainedlink{\noact}{p_n} \chainend{p_{n+1}}.P_n'(C) 
+					        \chainedlink{\widetilde{T}_i}{\noact}\chainedlink{\noact}{\widetilde{T}_o}
+			\chainend{p_{n+1}}\\
+\pi'_n(x) &\defeq &  \startchain{r_n}\chainedlink{x_i}{\noact}\chainedlink{\noact}{x_o}\chainedlink{r_{n+1}}{\noact}\chainedlink{\noact}{p_n} \chainend{p_{n+1}} 
 % \ + \   \startchain{r_n}\chainedlink{w_i}{\noact}\chainedlink{\noact}{w_o}\chainedlink{r_{n+1}}{\noact}\chainedlink{\noact}{p_n} \chainend{p_{n+1}}.P_n(q,b,w,a,q) \\
@@
                                                      \chainedlink{\overline{F}_i}{\noact}\chainedlink{\noact}{\overline{F}_o}
-					        \chainedlink{cxt}{\noact}\chainedlink{\noact}{p_4}
-					        \chainedlink{\tilde{G}_i}{\noact}\chainedlink{\noact}{\tilde{G}_o}
+					        \chainedlink{\mathit{cxt}}{\noact}\chainedlink{\noact}{p_4}
+					        \chainedlink{\widetilde{G}_i}{\noact}\chainedlink{\noact}{\widetilde{G}_o}
 			\chainend{\silent}.P_4  \\
 				&& +\\
-&&\startchain{r_4}\chainedlink{\overline{T'}_i}{\noact}\chainedlink{\noact}{\overline{T'}_o}\chainedlink{cxt}{\noact}\chainedlink{\noact}{p_4} \chainend{\silent}.P_4 \ +\  \startchain{r_4}\chainedlink{F_i}{\noact}\chainedlink{\noact}{F_o}\chainedlink{cxt}{\noact}\chainedlink{\noact}{p_4} \chainend{\silent}.P_4
+&&\startchain{r_4}\chainedlink{\overline{T'}_i}{\noact}\chainedlink{\noact}{\overline{T'}_o}\chainedlink{cxt}{\noact}\chainedlink{\noact}{p_4} \chainend{\silent}.P_4 \ +\  \startchain{r_4}\chainedlink{F_i}{\noact}\chainedlink{\noact}{F_o}\chainedlink{\mathit{cxt}}{\noact}\chainedlink{\noact}{p_4} \chainend{\silent}.P_4
 \end{array}
@@
 %and we put $r_1 = \silent$.
-\paragraph{Encoding of the entities.}
+
+\paragraph{Encoding of the entities}
+
 As for reactions, also the encoding of the entities is given in a parametric way.
-Here we differentiate three types of encodings: (1) for the entities that are not  provided by the context and  can be produced by the reactions; (2) for the entities that can be provided by the context and can be  produced by the reactions; (3) for the entities that are only provided by the context.\\
+Here we differentiate three types of encodings: (1)~for the entities that are not  provided by the context and  can be produced by the reactions; (2)~for the entities that can be provided by the context and can be  produced by the reactions; (3)~for the entities that are only provided by the context.
+
 Here, the entities $T$ and $T'$ that can be produced by the reactions and that are not provided by the context:
@@
 \begin{array}{lcllcl}
-P (T,\tilde{T})&\defeq & \sum_{h=0}^1 (\startchain{T_i}\chainedlink{T_o}{\noact}\chainend{\noact})^h\ \link{\tilde{T}_i}{\tilde{T}_o}.P (T,\tilde{T})
-& +  && \link{T_i}{T_o}.P (\overline{T},\tilde{T}) \\
-%\overline{P} (q)&\defeq &  \sum_{h=0}^1 (\startchain{\overline{T}_i}\chainedlink{\overline{T}_o}{\noact}\chainend{\noact})^h\link{\tilde{T}_i}{\tilde{T}_o}.P(q)\\
+P (T,\widetilde{T})&\defeq & \sum_{h=0}^1 (\startchain{T_i}\chainedlink{T_o}{\noact}\chainend{\noact})^h\ \link{\widetilde{T}_i}{\widetilde{T}_o}.P (T,\widetilde{T})
+& +  && \link{T_i}{T_o}.P (\overline{T},\widetilde{T}) \\
+%\overline{P} (q)&\defeq &  \sum_{h=0}^1 (\startchain{\overline{T}_i}\chainedlink{\overline{T}_o}{\noact}\chainend{\noact})^h\link{\widetilde{T}_i}{\widetilde{T}_o}.P(q)\\
 %&& + & && +\\
@@
 \begin{array}{lcl @{\hspace{0.5cm}}lcl @{\hspace{0.5cm}}lcl @{\hspace{0.5cm}}lcl}
-P_T & \defeq & P(T,\tilde{T})& \overline{P}_T & \defeq & P(\overline{T},\tilde{T}) & P_{T'}& \defeq & P(T',\tilde{T'}) & \overline{P}_{T',\tilde{T'}}& \defeq & P(\overline{T'},\tilde{T'})\\\
+P_T & \defeq & P(T,\widetilde{T})& \overline{P}_T & \defeq & P(\overline{T},\widetilde{T}) & P_{T'}& \defeq & P(T',\widetilde{T'}) & \overline{P}_{T',\widetilde{T'}}& \defeq & P(\overline{T'},\widetilde{T'})\\\
 \end{array}
@@
 \begin{array}{lcl @{\hspace{1cm}}lcl}
-P(C,\hat{C},\underline{C},\tilde{C}) & \defeq &  \sum_{h=0}^1 (\startchain{C_i}\chainedlink{C_o}{\noact}\chainend{\noact})^h\ \link{\hat{C}_i}{\hat{C}_o} (\startchain{\noact}\chainedlink{\noact}{\tilde{C}_i}\chainend{\tilde{C}_o})^h.
-P(C,\hat{C},\underline{C},\tilde{C})\\
+P(C,\widehat{C},\underline{C},\widetilde{C}) & \defeq &  \sum_{h=0}^1 (\startchain{C_i}\chainedlink{C_o}{\noact}\chainend{\noact})^h\ \link{\widehat{C}_i}{\widehat{C}_o} (\startchain{\noact}\chainedlink{\noact}{\widetilde{C}_i}\chainend{\widetilde{C}_o})^h.
+P(C,\widehat{C},\underline{C},\widetilde{C})\\
 &&+\\
-&& \sum_{h=0}^1 (\startchain{C_i}\chainedlink{C_o}{\noact}\chainend{\noact})^h\ \link{\underline{C}_i}{\underline{C}_o}.P(\overline{C},\hat{C},\underline{C},\tilde{C})\\
+&& \sum_{h=0}^1 (\startchain{C_i}\chainedlink{C_o}{\noact}\chainend{\noact})^h\ \link{\underline{C}_i}{\underline{C}_o}.P(\overline{C},\widehat{C},\underline{C},\widetilde{C})\\
 &&+\\
- && \sum_{h=0}^1 (\startchain{C_i}\chainedlink{C_o}{\noact}\chainend{\noact})^h\ \link{\underline{C}_i}{\underline{C}_o} \startchain{\noact}\chainedlink{\noact}{\hat{C}_i}\chainend{\hat{C}_o}.
-P(C,\hat{C},\underline{C},\tilde{C}) \\
- %&  \overline{E} (a)&\defeq & \sum_{h=1}^2 (\startchain{\overline{a}_i}\chainedlink{\overline{a}_o}{\noact}\chainend{\noact})^h\ \link{\hat{a}_i}{\hat{a}_o}.E(a)\\
+ && \sum_{h=0}^1 (\startchain{C_i}\chainedlink{C_o}{\noact}\chainend{\noact})^h\ \link{\underline{C}_i}{\underline{C}_o} \startchain{\noact}\chainedlink{\noact}{\widehat{C}_i}\chainend{\widehat{C}_o}.
+P(C,\widehat{C},\underline{C},\widetilde{C}) \\
+ %&  \overline{E} (a)&\defeq & \sum_{h=1}^2 (\startchain{\overline{a}_i}\chainedlink{\overline{a}_o}{\noact}\chainend{\noact})^h\ \link{\widehat{a}_i}{\widehat{a}_o}.E(a)\\
 %&&  &  && +\\
@@
 \begin{array}{lcl @{\hspace{1cm}}lcl@{\hspace{0.4cm}}lcl@{\hspace{0.4cm}}lcl}
-P_C & \defeq &  P(C,\hat{C},\underline{C},\tilde{C})&  P_G& \defeq &   P(G,\hat{G},\underline{G},\tilde{G}) \\
-\overline{P}_C & \defeq &  P(\overline{C},\hat{C},\underline{C},\tilde{C}) & 
-\overline{P}_G& \defeq &   P(\overline{G},\hat{G},\underline{G},\tilde{G})\\
+P_C & \defeq &  P(C,\widehat{C},\underline{C},\widetilde{C})&  P_G& \defeq &   P(G,\widehat{G},\underline{G},\widetilde{G}) \\
+\overline{P}_C & \defeq &  P(\overline{C},\widehat{C},\underline{C},\widetilde{C}) & 
+\overline{P}_G& \defeq &   P(\overline{G},\widehat{G},\underline{G},\widetilde{G})\\
 \end{array}
@@
 \begin{array}{lcllcl}
-P_F&\defeq & \sum_{h=0}^1 (\startchain{F_i}\chainedlink{F_o}{\noact}\chainend{\noact})^h\ \link{\hat{F}_i}{\hat{F}_o}.P_F
+P_F&\defeq & \sum_{h=0}^1 (\startchain{F_i}\chainedlink{F_o}{\noact}\chainend{\noact})^h\ \link{\widehat{F}_i}{\widehat{F}_o}.P_F
 & +  && \sum_{h=0}^1 (\startchain{F_i}\chainedlink{F_o}{\noact}\chainend{\noact})^h. \link{\underline{F}_i}{\underline{F}_o}.\overline{P}_F \\
-\overline{P}_F&\defeq & \sum_{h=0}^1 (\startchain{\overline{F}_i}\chainedlink{\overline{F}_o}{\noact}\chainend{\noact})^h\ \link{\hat{F}_i}{\hat{F}_o}.P_F
+\overline{P}_F&\defeq & \sum_{h=0}^1 (\startchain{\overline{F}_i}\chainedlink{\overline{F}_o}{\noact}\chainend{\noact})^h\ \link{\widehat{F}_i}{\widehat{F}_o}.P_F
 & +  && \sum_{h=0}^1 (\startchain{\overline{F}_i}\chainedlink{\overline{F}_o}{\noact}\chainend{\noact})^h. \link{\underline{F}_i}{\underline{F}_o}.\overline{P}_F \\
-%\overline{P} (q)&\defeq &  \sum_{h=0}^1 (\startchain{\overline{T}_i}\chainedlink{\overline{T}_o}{\noact}\chainend{\noact})^h\link{\tilde{T}_i}{\tilde{T}_o}.P(q)\\
+%\overline{P} (q)&\defeq &  \sum_{h=0}^1 (\startchain{\overline{T}_i}\chainedlink{\overline{T}_o}{\noact}\chainend{\noact})^h\link{\widetilde{T}_i}{\widetilde{T}_o}.P(q)\\
 %&& + & && +\\
@@
 \]
-Now, the context follow can non deterministically provide the entities: $C$, $G$ $F$:\\
+
+Also in this example we account for a nondeterministic context that can (nondeterministically) provide any combination of the entities: $C$, $G$ $F$:
 \[
@@
 
-Now, to show a possible composition of a transition label, we assume an system where the entities $a$, $T'$, and $G$ are present:
+Now, to show a possible composition of a transition label, we assume a system where only the entities $a$, $T'$, and $G$ are present:
 \[
 \begin{array}{lcl}
-Sys &\defeq &%\restrict{\tilde{C},\tilde{G},\tilde{F}, \tilde{T},\tilde{T'}}(
-I\mid P_a\mid \overline{P}_C\mid P_G\mid \overline{P}_F \mid \overline{P}_T \mid P_{T'} \mid P_1 \mid P_2 \mid P_3 \mid P_4 \mid Cxt
+\mathit{Sys} &\defeq &%\restrict{\widetilde{C},\widetilde{G},\widetilde{F}, \widetilde{T},\widetilde{T'}}(
+I\mid P_a\mid \overline{P}_C\mid P_G\mid \overline{P}_F \mid \overline{P}_T \mid P_{T'} \mid P_1 \mid P_2 \mid P_3 \mid P_4 \mid \mathit{Cxt}
 \end{array}
 \]
-In the above configuration, reactions $1$ and $3$ can be applied, and also we assume that the context will provide the entity $F$, that will be available in the target configuration.
-Instead of showing the transition label, we give give the string generated by the function $flat$:\
+In the above configuration, reactions $1$ and $4$ can be applied, and also we assume that the context will provide the entity $F$, that will be available in the target configuration.
+Instead of showing the complete transition label, we give its flattened version obtained by applying the function $\mathit{flat}(\cdot)$:
 %The transition label appears as:
 \[
-\tiny
-\begin{array}{l}
-%\restrict{r1,r2,r3,r4,a,\tilde{C},\tilde{G},\tilde{F},\tilde{T},\tilde{T'}}\\
+%\restrict{r1,r2,r3,r4,a,\widetilde{C},\widetilde{G},\widetilde{F},\widetilde{T},\widetilde{T'}}\\
 % \startchain{\silent}\chainedlink{a_i}{\bf  a_i}\chainedlink{\bf a_o}{a_o}
@@
 %\chainedlink{\underline{G}_i}{\bf \underline{G}_i}\chainedlink{\bf \underline{G}_o}{\underline{G}_o}
-%\chainedlink{\hat{F}_i}{\bf \hat{F}_i}\chainedlink{\bf \hat{F}_o}{\hat{F}_o}\chainedlink{p_1}{p_1} 
-% \chainedlink{\tilde{T}_i}{\bf \tilde{T}_i}\chainedlink{\bf \tilde{T}_o}{\tilde{T}_o}
-%  \chainedlink{p_2}{p_2}\chainedlink{p_3}{p_3}\chainedlink{p_4}{p_4} \chainedlink{\tilde{G}_i}{\bf \tilde{G}_i}\chainedlink{\bf \tilde{G}_o}{\tilde{G}_o} \chainend{\silent}
-r_1~ a~G~\overline{C}~ r_2~ G~ r_3~T'~r_4~T'~\overline{F}~cxt~\underline{C}~\underline{G}~\hat{F}~p_1~p_2~p_3~\tilde{G}~\silent
-
-\end{array}
+%\chainedlink{\widehat{F}_i}{\bf \widehat{F}_i}\chainedlink{\bf \widehat{F}_o}{\widehat{F}_o}\chainedlink{p_1}{p_1} 
+% \chainedlink{\widetilde{T}_i}{\bf \widetilde{T}_i}\chainedlink{\bf \widetilde{T}_o}{\widetilde{T}_o}
+%  \chainedlink{p_2}{p_2}\chainedlink{p_3}{p_3}\chainedlink{p_4}{p_4} \chainedlink{\widetilde{G}_i}{\bf \widetilde{G}_i}\chainedlink{\bf \widetilde{G}_o}{\widetilde{G}_o} \chainend{\silent}
+r_1~ a~G~\overline{C}~ r_2~ G~ r_3~T'~r_4~T'~\overline{F}~cxt~\underline{C}~\underline{G}~\widehat{F}~p_1~\widetilde{T}~p_2~p_3~p_4~\widetilde{G}~\silent
 \]
+The original label can then be reconstructed just applying the function $\mathit{unflat}(\cdot)$ to the string above.

### 2020-04-02 05-translation.tex
proof of Lemma 12 revised
and Corollary added

@@
   \begin{enumerate}
-  \item for each reaction  $a_j \in A$, the corresponding channels $r_j$ and $p_j$ appear in  $\upsilon$; for each entity $s \in S$, the corresponding channel $s$ (suitably decorated) appear in $\upsilon$;
+  \item for each reaction  $aj \in A$, the corresponding channels $r_j$ and $p_j$ appear in  $\upsilon$; for each entity $s \in S$, the corresponding channel $s$ (suitably decorated) appear in $\upsilon$;
   the channel $\mathit{cxt}$ appears in $\upsilon$;
@@
  %involved in the transition $t$;
-  \item for each reaction $a \in A$ and each entity $s \in S$, each virtual link offered by processes $P_a$ and $\mathit{Cxt}^1$ is  overlapped by exactly one solid link offered by  processes representing entities.
+  \item for each reaction $aj \in A$ and each entity $s \in S$, each virtual link offered by processes $P_a$ and $\mathit{Cxt}$ is  overlapped by exactly one solid link offered by  processes representing entities.
   %each process $P_s $ (or its alternative version $\overline{P}_s$), with $s \in S$, participates at most  with one solid link for each process $P_a$, and with exactly one solid link for each process $Cxt_s$.%, if $s$ is a \emph{required element} of $A_j$ in $t$.
@@
 and will write $P\xRightarrow{\upsilon} P'$ to mean that $P\xrightarrow{\upsilon} P'$ with $\upsilon$ complete.
+
+The following Corollary immediately follows from Lemma~\ref{lem:struct}.
+
+\begin{corollary}
+Let ${\cal A} = (S, A)$ be a RS and let $\pi=(\gamma,\delta)$ be an extended  interactive process in ${\cal A}$. 
+Let $Q=\llparenthesis{\cal A},\gamma\rrparenthesis$ and $P= \llbracket{\cal A},\gamma\rrbracket = \restrict{\mathit{names}}Q$. 
+Then $P \xrightarrow{\restrict{\mathit{names}}\upsilon} P'$ iff $Q \xRightarrow{\upsilon} Q'$ and $P'=\restrict{\mathit{names}}Q'$.
+\end{corollary}
 

### 2020-04-02 09-appendix.tex
proof of Lemma 12 revised
and Corollary added

% !TEX root = ./main_RS2L_LNCS.tex

%\section{Background on Core Network Algebra (CNA)}
%We briefly report the basic concepts of the \CNA~calculus.
%
%\paragraph{Links and link chains}
%A link is a pair $\link{\alpha}{\beta}$
%that records the source and the target sites 
%of a communication,  meaning that the input available 
%at the source site $\alpha$ can be forwarded to the target one $\beta$.  
%Links are suitably combined in link chains to describe 
%how information can be routed across processes before 
%arriving at their destination.
%
%More formally, let $\mathcal{C}$ be the set of channels, ranged over by $a,b,c,...$, and 
%let $\mathcal{A} = \mathcal{C} \cup \setof{\silent} \cup \setof{{\noacts}}$ be the set of actions, ranged over by $\alpha,\beta,\gamma,...$,
%where the symbol $\silent$ denotes a \emph{silent} action, while the symbol $\noact$ denotes a \emph{virtual} (non-specified) action.
%A \emph{link} is a pair $\ell=\link{\alpha}{\beta}$;
% it can be read as forwarding the input available on $\alpha$ to $\beta$, and we call $\alpha$ the \emph{source site} of $\ell$ and $\beta$ the \emph{target site} of $\ell$.
%%
%A link $\link{\alpha}{\beta}$ is \emph{solid} if $\alpha,\beta \neq \ \noacts$; 
%the link $\link{\noact}{\noact}$ is called \emph{virtual}.
%A link is \emph{valid} if it is solid or virtual.
%We let $\mathcal{L}$ be the set of valid links.
%% ranged over by $p,q,r,...$.
%
%All the  processes that want to participate to a single interaction, must offer
% a link chain composed the same number of links: each link chain must
% be formed by only one solid link and all the other ones must be virtual.
% The positions of the solid links in each chain must be different and 
% sibling links must be such that the 
%right hand channel of the previous link must match with the left hand channel of the
%next link.  Please note that virtual links match with any channel name.
%
%More formally, a \emph{link chain} is a 
%finite sequence $s = \ell_{1}...\ell_{n}$ of (valid) links  $\ell_{i} = \link{\alpha_{i}}{\beta_{i}}$ such that:
%\begin{enumerate}
%
%\item for any $i\in [1,n-1]$, 
%$\left\{\begin{array}{ll}
%\beta_{i},\alpha_{i+1}\in \mathcal{C} & \mbox{ implies } \beta_{i} = \alpha_{i+1}\\
%\beta_{i}=\silent & \mbox{ iff } \alpha_{i+1}=\silent
%\end{array}\right.
%$
%\item  $\exists i \in [1,n]. \ \ell_i \neq \link{\noact}{\noact}$.
%\end{enumerate}
%
%
%\paragraph{Syntax and semantics}
%
%The syntax of \CNA is as follow:
%
%
%\begin{definition}
%The \CNA\ processes are generated by the following grammar:
%\[
%\begin{array}{rcl}
%P,Q & ::= &
%\nil \prodsep
%%X \prodsep
%\ell. P  \prodsep
%P+Q \prodsep
%P|Q \prodsep
%\restrict{a} P \prodsep
%P[\phi] \prodsep
%A
%\end{array}
%\]
%\noindent
%where $\ell$ is a solid link ( i.e.\  $\ell = \link{\alpha}{\beta}$ with $\alpha,\beta\neq \ \noacts$) and $A$ is a process identifier for which we assume a definition $A \defeq P$ is available in a given set $\Delta$ of (possibly recursive) process definitions.
%\end{definition}
%
%The operational semantics is defined in terms of a Labelled Transition System, in which states are \CNA\ processes, labels are link chains, and transitions are generated by the SOS rules in Figure~\ref{fig:cnasos}.  The semantics makes use of the an equivalence relation that allow rule  $(Act)$ 
%of the semantics to decide the dimention (i.e. the maximum number of the participants) of the
%interaction a process want to take part in.
%
%\begin{definition}[Equivalence $\blackstretcheq$]\label{def:black}
%We let $\blackstretcheq$ be the least equivalence relation 
%%(i.e.~closed under reflexivity, symmetry and transitivity) 
%over link chains closed under the axioms (whenever both sides are well defined link chains):
%\[
%\begin{array}{rclcrcl}
%s\link{\noact}{\noact} & \blackstretcheq &  s & \qquad &
%s_1 \startchain{\noact}\chainedlink{\noact}{\noact}\chainend{\noact}s_2 & \blackstretcheq & s_1 \link{\noact}{\noact} s_2\\
%\link{\noact}{\noact}s & \blackstretcheq & s & &
%s_1 \startchain{\alpha}\chainedlink{a}{a}\chainend{\beta} s_2
%& \blackstretcheq & 
%s_1 \startchain{\alpha}\chainedlink{a}{\noact}\chainedlink{\noact}{a}\chainend{\beta}s_2 
%\end{array}
%\]
%\end{definition}
%
%
%
%\begin{figure}[t]
%\begin{center} 
%\begin{prooftree} 
%\AxiomC{$s \blackstretcheq \ell $} %\mbox{($\ell$ {\em only} solid link in $s$)}
%\RightLabel{\scriptsize(\textit{Act})}
%\UnaryInfC{$\ell.P \xrightarrow{s} P$} 
%\DisplayProof
%\
%\AxiomC{$P \xrightarrow{s} P'$} 
%\RightLabel{\scriptsize(\textit{Lsum})}
%\UnaryInfC{$P+Q \xrightarrow{s} P'$} 
%\DisplayProof
%\
%\AxiomC{$P \xrightarrow{s} P'$} 
%\RightLabel{\scriptsize(\textit{Res})}
%\UnaryInfC{$\restrict{a}P \xrightarrow{\restrict{a}s} \restrict{a}P'$} 
%%\DisplayProof
%%\,
%%\AxiomC{$Q \xrightarrow{s} Q'$} 
%%\RightLabel{\scriptsize(Rsum)}
%%\UnaryInfC{$P+Q \xrightarrow{s} Q'$} 
%\end{prooftree} 
%\end{center}
%
%\begin{center} 
%\begin{prooftree} 
%\AxiomC{$P \xrightarrow{s} P'$} 
%\RightLabel{\scriptsize(\textit{Rel})}
%\UnaryInfC{$P[\phi] \xrightarrow{s[\phi]} P'[\phi]$} 
%\DisplayProof
%\
%\AxiomC{$P \xrightarrow{s} P'$} 
%\RightLabel{\scriptsize(\textit{Lpar})}
%\UnaryInfC{$P|Q \xrightarrow{s} P'|Q$} 
%\DisplayProof
%\ 
%%\AxiomC{$Q \xrightarrow{s} Q'$} 
%%\RightLabel{\scriptsize(Rpar)}
%%\UnaryInfC{$P|Q \xrightarrow{s} P|Q'$} 
%%\end{prooftree} 
%%\end{center}
%%\begin{center} 
%%\begin{prooftree} 
%\AxiomC{$P \xrightarrow{s} P'$} 
%\AxiomC{$Q \xrightarrow{s'} Q'$} 
%\RightLabel{\scriptsize(\textit{Com})}
%\BinaryInfC{$P|Q \xrightarrow{\merges{s}{s'}} P'|Q'$} 
%\DisplayProof
%\end{center}
%\begin{center}
%%\AxiomC{$P[\subst{\recproc{X}{P}}{X}] \xrightarrow{s} P'$} 
%%\RightLabel{\scriptsize(Rec)}
%%\UnaryInfC{$\recproc{X}{P} \xrightarrow{s} P'$} 
%\AxiomC{$P \xrightarrow{s} P'$}
% \AxiomC{$(A \defeq P)\in\Delta$}
%\RightLabel{\scriptsize(\textit{Ide})}
%\BinaryInfC{$A \xrightarrow{s} P'$} 
%\end{prooftree} 
%\end{center}
%\caption{SOS semantics of the \CNA\  (rules (\textit{Rsum}) and (\textit{Rpar}) are omitted).}
%\label{fig:cnasos}
%\end{figure}
%

\newpage

\section{Omitted Proofs}
In this section we report the proofs for the results in Section~\ref{sec:trans}.
%\noindent
%We recall Lemma~\ref{lem:struct} (the original appears on p.\,\pageref{lem:struct}).

\lemmastruct*
\begin{proof}
  We prove the two items separately:
  \begin{enumerate}
\item by Definition~\ref{def:trans}, all the names that appear in the prefixes of any subprocess are in the set $\mathit{names}$ and thus restricted. 
They include all the reaction names $r_j$, all the production names $p_j$, all the entity names $s_i,s_o$, in all their decorated versions, and the name $\mathit{cxt}$. 
Therefore the chain $\upsilon$ must start and end with a $\silent$ action and cannot contain virtual links.
The only prefix that starts with $\silent$ is the one of the recursive init process $I$ (prefix $\link{\silent}{r_1}$) and the only prefixes that end with $\silent$ are those associated to reaction $a_u$ (as we have assumed that $p_{u+1} = \silent$, where $u$ is the number of reactions).
Then each prefix that starts with $r_j$ involves $p_j$ and $r_{j+1}$.
Thus all the prefixes associated with reactions must be concatenated and also the prefix associated with the context (remember that $r_{u+1} = \mathit{cxt}$), forming the backbone of the label.
Since the context processes is involved, then all entities processes are also involved.
Then, all the processes $I$, $P_a$, $P_s$ (or $\overline{P_s}$), and $\mathit{Cxt}$  must participate to each transition.
 \item for each reaction $aj \in A$, the c\CNA \ code of  $P_{aj}$ leaves one virtual link between two solid links of the types $\link{r_j}{\dots} \chainedlink{s_{i}}{\noact}\chainedlink{\noact}{s_{o}}\link{}{\dots}\link{}{r_{j+1}} ^{\dots}\link{p_j}{\dots}\link{}{p_{j+1}}$. Then, it derives that the process $P_{s}$, encoding the behaviour of entity $s$, can participate by filling the virtual link in the above transition by only offering one solid link of the form $\link{s_{i}}{s_{o}}$. In fact, there is no other way to generate a solid chain from $s_{i}$ to $s_{o}$.
 The same reasoning holds for the processes $\mathit{Cxt}$ and for all the decorated versions of $s_i,s_o$.
% $P_{s_1}$ can not participate with a link chain, as by juxtaposing  $\link{s_{1i}}{s_{1o}}$, not valid
 %link chains would be generated.
   \end{enumerate}
  \end{proof}

%\noindent
%We recall Proposition~\ref{prop:corr1}
%(the original appears on p.\,\pageref{prop:corr1}).
\propcorruno*

 %\begin{proposition}(correctness 1)\\
 %\label{lem:struct}
%  Let ${\cal A} = (S, A)$ be a rs, and let  $\pi=(\gamma,\delta)$ be a $n$-step interactive process in $A$, with
% $\gamma = C_0,\dots,C_n$ and $\delta = D_0,\dots,D_n$.
%% We define $\pi' = (\gamma',\delta')$ with $\gamma' = C_0',\dots,C_{n-1}'$ and $\delta' = D_0',\dots,D_{n-1}'$ as $C_0' = C_1 \cup D_1$,  and
%% $C_i' = C_{i+1}$, $D_{i}' = D_{i+1}$ with $ 1 \leq i < n$.
%  Let   $P=[|{\cal A},\gamma|]$ be  its c\CNA~translation, and let $P'$ be such that $P \xrightarrow{\link{\silent}{\silent}}P'$.% then $P' = [|{\cal A},\gamma' |]$. 
%Then it holds that\\
% $P= \restrict{reacts,ctxs,ents} (\Pi_{a \in A}P_a| \Pi_{s \in S}CXT_s | \Pi_{s \in C_0}P_s| \Pi_{s \notin C_0}\overline{P}_s)$\\
% $P'= \restrict{reacts,ctxs,ents} (\Pi_{a \in A}P_a| \Pi_{s \in S}CXT_s | \Pi_{s \in C_1 \cup D_1}P_s| \Pi_{s \notin C_1 \cup D_1}\overline{P}_s)$
 %\end{proposition}
 \begin{proof}
 First, we note that all the channels in the system are restricted, see Def.~\ref{def:trans}, then
 it holds that the transition labels are of the form  $v=\link{\silent}{\silent}\dots\link{\silent}{\silent}$.\\
 Now,
by Definition~\ref{def:trans} and by Lemma~\ref{lem:struct}.1, all the  channels $r_j,p_j$, with
 $j \in [1,\dots,u]$, and $cxt_h$, with $h \in [1, \dots, w]$, and all the annotated versions of 
 $s_i$, $s_o$ are restricted. Also, processes $Cxt_s$ always requires the 
 interaction with $P_s$ on either on channels $\hat{s}_i$, $\hat{s}_o$ or on channels 
 $\underline{s}_i$, $\underline{s}_o$.
 It derives that all the processes: $P_a$  (coding the behaviour of reaction $a \in A$), $P_s$ (coding the behaviour of entity $s \in S$), and $Cxt_s$ (coding the behaviour of the context regarding the entity $s$) have been involved in the transition. 
We have the following cases:
\begin{description}
\item[(a)] processes $P_a$ encoding a reaction $a$,  with serial  number $j$, producing the entity $s$ provide a code of this type:
%By Lemma~\ref{lem:compl1}, all the processes $A_j$, and $Cxt_h$ participated in the tr
%The code executed by any process encoding rule $R_i$ producing $e$ is 
$P_a\defeq \link{r_j}{\dots}\chainedlink{r_{j+1}}{\noact}\chainedlink{\noact}{p_j}\link{}{\dots}\chainedlink{\tilde{s}_i}{\noact}\chainedlink{\noact}{\dots}\link{\tilde{s}_o}{\dots} \chainend{p_{j+1}}.P_a $;
\item[(b)] processes $P_a$ encoding a reaction $a$, with serial number $j$, consuming the entity $s$
% without producing it, 
provide a code of this type: 
$P_a\defeq \link{r_j}{\dots}\chainedlink{s_i}{\noact}\chainedlink{\noact}{}\link{s_o}{\dots} \chainedlink{r_{j+1}}{\noact}\chainedlink{\noact}{p_j}\link{}{\dots}\link{}{p_{j+1}}.P_a $;
\item[(c)]  processes $P_a$ encoding a reaction $a$, with serial number $j$, requiring the absence of the entity $s$ provide a code of this type: \\
$P_a\defeq \link{r_j}{\dots}\chainedlink{\overline{s}_i}{\noact}\chainedlink{\noact}{}\link{\overline{s}_o}{\dots} \chainedlink{r_{j+1}}{\noact}\chainedlink{\noact}{p_j}\link{}{\dots}\link{}{p_{j+1}}.P_a $;
\item[(d)] processes $P_a$ encoding a reaction $a$ (that cannot be applied), with serial number $j$, 
execute a code  capturing either the absence of one of its reactants (case 1), or the presence  of one of its inhibitors (case 2):
\begin{enumerate}
\item $P_a\defeq \link{r_j}{\dots}\chainedlink{\overline{s}_i}{\noact}\chainedlink{\noact}{}\link{\overline{s}_o}{\dots} \chainedlink{r_{j+1}}{\noact}\chainedlink{\noact}{p_j}\link{}{\dots}\link{}{p_{j+1}}.P_a $.
\item $P_a\defeq \link{r_j}{\dots}\chainedlink{s_i}{\noact}\chainedlink{\noact}{}\link{s_o}{\dots} \chainedlink{r_{j+1}}{\noact}\chainedlink{\noact}{p_j}\link{}{\dots}\link{}{p_{j+1}} .P_a $.
\end{enumerate}
\end{description}
%It is worth nothing that there could be other processes $P_a$ encoding  a reaction $a \in A$ that do not
%involve the entity $s$.\\
Now, we consider the structure of the process $Cxt_s = Cxt^0_s$. By Definition~\ref{def:trans}, $Cxt^0_s$ is the unique process encoding the behaviour of the context regulating the supply of $s$. 
The code of the process $Cxt^t_s$, with $t \geq 0$, has the following structure: 
\begin{description}
\item[(e)] $Cxt^i_s \defeq  \startchain{cxt_h}\chainedlink{\hat{s}_i}{\noact}\chainedlink{\noact}{\hat{s}_o}\chainend{cxt_{h+1}}. Cxt^{i+1}_{s}$, if $s \in C_{i+1}$;
\item[(f)] $Cxt^i_s \defeq  \startchain{cxt_h}\chainedlink{\underline{s}_i}{\noact}\chainedlink{\noact}{\underline{s}_o}\chainend{cxt_{h+1}}. Cxt^{i+1}_{s}$, if $s \notin C_{i+1}$;
%\item[(e)] $Cxt_{s}^1 \defeq \startchain{cxt_h}\chainedlink{\hat{s}_i}{\noact}\chainedlink{\noact}{\hat{s}_o}\chainend{cxt_{h+1}}. Cxt_{s}^{2}$, if  $s \in C_1$ (and hence $s \in C_0')$;
%\item[(f)] $Cxt_{s}^1 \defeq \startchain{cxt_h}\chainedlink{\underline{s}_i}{\noact}\chainedlink{\noact}{\underline{s}_o}\chainend{cxt_{h+1}}. Cxt_{s}^{2}$, if  $s \notin C_1$.
\end{description}

%$R_j \defeq \startchain{rule_j}\chainedlink{e_i}{\noact}\chainedlink{\noact}{e_o}\chainedlink{\overline{e'}_i}{\noact} \chainedlink{\noact}{\overline{e'}_o}\chainedlink{\tilde{e''}_i}{\noact}\chainedlink{\noact}{\tilde{e''}_o} \chainend{rule_{j+1}}.R_j + \dots$, with $e',e''\neq e $.\\
%The code executed by any rule $j$ not requiring $e$ is $R_j \defeq \startchain{rule_j}\chainedlink{e_i}{\noact}\chainedlink{\noact}{e_o} \chainend{rule_{j+1}}.R_j + \dots$.\\
\noindent
The code executed by  $P_s$ has the following structure:
\begin{description}
\item[(g)] $P_s \defeq  \sum_{h,k\geq0}(\startchain{s_i}\chainedlink{s_o}{\noact}\chainend{\noact}) ^h\, \startchain{\hat{s}_i}\chainedlink{\hat{s}_o}{\noact}\chainend{\noact}
 (\startchain{\tilde{s}_i}\chainedlink{\tilde{s}_o}{\noact}\chainend{\noact})^k\,.P_s $, if $s \in C_{i+1}$;
% \underbrace{ \startchain{s_i}\chainedlink{s_o}{\noact}\chainedlink{\noact}{s_i}\chainedlink{s_o}{\noact}\chainend{\noact}...}_{v\ times}\underbrace{ \startchain{\tilde{s}_i}\chainedlink{\tilde{s}_o}{\noact}\chainedlink{\noact}{\tilde{s}_i}\chainedlink{\tilde{s}_o}{\noact}\chainend{\noact}...}_{z\ times}\link{\hat{s}_i}{\hat{s}_o}.P_s $, if $s \in C_{i+1}$;
%\item[()] $P_s \defeq  \underbrace{ \startchain{s_i}\chainedlink{s_o}{\noact}\chainedlink{\noact}{s_i}\chainedlink{s_o}{\noact}\chainend{\noact}...}_{v\ times}\link{\hat{s}_i}{\hat{s}_o}.P_s $, if $s \in C_1$ %(and hence $s \in C_0'$)
\item[(h)] $P_s \defeq \sum_{h \geq 0,k\geq 1}(\startchain{s_i}\chainedlink{s_o}{\noact}\chainend{\noact})^h\ \startchain{\underline{s}_i}\chainedlink{\underline{s}_o}{\noact}\chainend{\noact} (\startchain{\tilde{s}_i}\chainedlink{\tilde{s}_o}{\noact}\chainend{\noact})^k
.P_s$, if $s \not\in C_{i+1}$
% \underbrace{ \startchain{s_i}\chainedlink{s_o}{\noact}\chainedlink{\noact}{s_i}\chainedlink{s_o}{\noact}\chainend{\noact}...}_{v\ times}\underbrace{ \startchain{\tilde{s}_i}\chainedlink{\tilde{s}_o}{\noact}\chainedlink{\noact}{\tilde{s}_i}\chainedlink{\tilde{s}_o}{\noact}\chainend{\noact}...}_{z >0 \ times}\link{\underline{s}_i}{\underline{s}_o}.P_s$, if $s \in C_{i+1}$
\item[(i)] $P_s \defeq  \sum_{h\geq0} (\startchain{s_i}\chainedlink{s_o}{\noact}\chainend{\noact})^h
%{s_i}\chainedlink{s_o}{\noact}\chainend{\noact}...}_{v\ times}
\link{\underline{s}_i}{\underline{s}_o}.\overline{P}_s$, if $s \notin C_{i+1}$
\end{description}
where, by Lemma~\ref{lem:struct}.2, $h$ is the number of reactions requiring the presence of $s$ plus possibly some reactions not requiring $s$; and $k$ is the number of reactions producing $s$.
The code executed by  $\overline{P}_s$ has the following structure:
\begin{description}
\item[(g')] $\overline{P}_s \defeq  \sum_{h,k\geq 0}(\startchain{\overline{s}_i}\chainedlink{\overline{s}_o}{\noact}\chainend{\noact})^h  \, \startchain{\hat{s}_i}\chainedlink{\hat{s}_o}{\noact}\chainend{\noact}
 (\startchain{\tilde{s}_i}\chainedlink{\tilde{s}_o}{\noact}\chainend{\noact})^k\,.P_s $, if $s \in C_{i+1}$;
% \underbrace{ \startchain{\overline{s}_i}\chainedlink{\overline{s}_o}{\noact}\chainedlink{\noact}{\overline{s}_i}\chainedlink{\overline{s}_o}{\noact}\chainend{\noact}...}_{v\ times}\underbrace{ \startchain{\tilde{s}_i}\chainedlink{\tilde{s}_o}{\noact}\chainedlink{\noact}{\tilde{s}_i}\chainedlink{\tilde{s}_o}{\noact}\chainend{\noact}...}_{z\ times}\link{\hat{s}_i}{\hat{s}_o}.P_s $, if $s \in C_{i+1}$;
%\item[()] $P_s \defeq  \underbrace{ \startchain{s_i}\chainedlink{s_o}{\noact}\chainedlink{\noact}{s_i}\chainedlink{s_o}{\noact}\chainend{\noact}...}_{v\ times}\link{\hat{s}_i}{\hat{s}_o}.P_s $, if $s \in C_1$ %(and hence $s \in C_0'$)
\item[(h')] $\overline{P}_s \defeq  \sum_{h \geq 0,k\geq 1} (\startchain{\overline{s}_i}\chainedlink{\overline{s}_o}{\noact}\chainend{\noact})^h\ \startchain{\underline{s}_i}\chainedlink{\underline{s}_o}{\noact}\chainend{\noact} (\startchain{\tilde{s}_i}\chainedlink{\tilde{s}_o}{\noact}\chainend{\noact})^k.P_s$, if $s \notin C_{i+1}$

% \underbrace{ \startchain{\overline{s}_i}\chainedlink{\overline{s}_o}{\noact}\chainedlink{\noact}{\overline{s}_i}\chainedlink{\overline{s}_o}{\noact}\chainend{\noact}...}_{v\ times}\underbrace{ \startchain{\tilde{s}_i}\chainedlink{\tilde{s}_o}{\noact}\chainedlink{\noact}{\tilde{s}_i}\chainedlink{\tilde{s}_o}{\noact}\chainend{\noact}...}_{z >0 \ times}\link{\underline{s}_i}{\underline{s}_o}.P_s$, if $s \in C_{i+1}$
\item[(i')] $\overline{P}_s \defeq  \sum_{h \geq 0}(\startchain{\overline{s}_i}\chainedlink{\overline{s}_o}{\noact}\chainend{\noact})^h\,\link{\underline{s}_i}{\underline{s}_o}.\overline{P}_s$, if $s \notin C_{i+1}$
\end{description}
 \noindent
where, by Lemma~\ref{lem:struct}.2, $h$ is the number of reactions requiring the absence of $s$ plus possibly some reactions  requiring $s$; and $k$ is the number of reactions producing $s$.\\

\noindent 
It is worth nothing that, depending on the presence ($P_s$) or the absence ($\overline{P}_s$) of each entity $s$, for each process $P_a$ (encoding a reaction $a$) the choice between the execution of the reaction code (points {\bf (a)}, {\bf (b)}, {\bf (c)}) or  the code expressing that reaction $a$ is not applicable (point {\bf (d)}) is deterministic.
Also, the building of the code of processes $Cxt_s$ (points {\bf (e)}, {\bf (f)}), follows the evolution of $\gamma$.
It derives that the trend followed by the processes $P_s$ (or $\overline{P}_s$) is also deterministic (points {\bf (g)}, {\bf (h)}, {\bf (i)} or {\bf (g')}, {\bf (h')}, {\bf (i')}), leading to $P' = \llbracket {\cal A},\gamma^1\rrbracket$.
 \end{proof}

%  \begin{corollary}[correctness 1]
% \label{cor:complet1}
% Let ${\cal A} = (S, A)$ be a rs, and let  $\pi=(\gamma,\delta)$ be a $n$-step interactive process in $A$, with
% $\gamma = C_0,\dots,C_n$ and $\delta = D_0,\dots,D_n$. Let   $P=[|{\cal A},\gamma|]$ be  its c\CNA~translation. We define $\pi' = (\gamma',\delta')$ with $\gamma' = C_0',\dots,C_{n-1}'$ and $\delta' = D_0',\dots,D_{n-1}'$ as $C_0' = C_1 \cup D_1$,  and
% $C_i' = C_{i+1}$, $D_{i}' = D_{i+1}$ with $ 1 \leq i < n$. Then,
% if $\exists$ $P'$, $P''$ such that $P \xrightarrow{\link{\silent}{\silent}} P'$ and $P \xrightarrow{\link{\silent}{\silent}} P''$, then $P' = P''$.
%\end{corollary}
% \begin{proof}
% %\noindent
%%{\bf Point 2.}\\
%By Definition~\ref{def:trans} if follows that:
%\begin{enumerate}
%\item for the processes $P_a$, the choice  between applying the behaviour
%corresponding of applying reaction $a$ or one of its negations is deterministic, as it depends on the presence of the required reactants ($P_s$) and the absence of its inhibitors ($\overline{P}_s$);
%\item the behaviour of processes $Cxt_s$ are deterministic, as their  behaviour is driven by the component 
%$\gamma$ of the corresponding reactive system.
%\end{enumerate}
%It derives that for each process $P_s$, its continuation ($P_s$ or $\overline{P}_s$) is deterministic.
%And we are done.
%
% \end{proof}

% \begin{proposition}[correctness2]
% \label{prop:correctness1}
% Let $P$ be a c\CNA \ process such that $\exists$ a $rs$ ${\cal A}$ and a $n$-step interactive process $\pi=(\gamma,\delta)$ with $\gamma =C_0,\dots,C_n$ and $\delta = D_0,\dots,D_n$ such that $P= [|{\cal A},\gamma|]$. 
% %We set $\pi'=(\gamma',\delta')$ with $\gamma'=C_0',\dots,C_{n-1}'$,  $\delta'=D_0',\dots,D_{n-1}$,  where $C_0'= C_1 \cup D_1$, $C_i'= C_{i+1}$, $D_i=D_{i+1}'$,  and $1 \leq i < n$.
%Then,   $\exists$ $P'$ and $\pi'=(\gamma',\delta')$ such that $P \xrightarrow{\link{\silent}{\silent}}P'$, $P' =[|{\cal A},\gamma'|]$.
% \end{proposition} 
% \begin{proof}
% By Lemma~\ref{lem:compl2},  $\exists !$ $P'$ such that $P \xrightarrow{\link{\silent}{\silent}}P'$.\\
% By Definition~\ref{def:trans}, $P= \restrict{reacts,ctxs,ents}(\Pi_{a \in A}P_a | \Pi_{s \in S}CXT_s|\Pi_{s\in C_0} P_s | \Pi_{s \notin C_0}\overline{P}_s)$.\\
% By Lemma~\ref{lem:struct}, \\$P' =  \restrict{reacts,ctxs,ents}(\Pi_{a \in A}P_a | \Pi_{s \in S}CXT_s|\Pi_{s\in C_1 \cup D_1} P_s | \Pi_{s \notin C_1 \cup D_1}\overline{P}_s)$ 
%  Then, we set $\gamma' = C_0',\dots,C_{n-1}'$ and $\delta'=D_0',\dots,D_n'$ as
%  We set $C_0'=C_1 \cup D_1$, $C_i' = C_{i+1}$, and $D_i' = D_{i+1}$, with $1 \leq  i < n-1$.
%  And we are done.
% \end{proof}
%
% 
%

%\noindent
% We recall Corollary~\ref{corr:corr2}
%(the original appears on p.\,\pageref{corr:corr2}).
\corrcorrdue*

% \begin{corollary}[correcteness2]
% %\label{prop:correctness2}
%  Let $P$ a  c\CNA process such that exists a rs ${\cal A}$ and a $n$-step  interactive process $\pi=(\gamma,\delta)$ with $\gamma =C_0,\dots,C_n$ and $\delta = D_0,\dots,D_n$ such that $P= [|{\cal A},\gamma|]$. 
% %We set $\pi'=(\gamma',\delta')$ with $\gamma'=C_0',\dots,C_{n-1}'$,  $\delta'=D_0',\dots,D_{n-1}$,  where $C_0'= C_1 \cup D_1$, $C_i'= C_{i+1}$, $D_i=D_{i+1}'$,  and $1 \leq i < n$.
%Then,   $\exists$ $P''$ and $\pi''=(\gamma'',\delta'')$ such that $P \xrightarrow{\link{\silent}{\silent}}^{w-1}P''$, $P'' =[|{\cal A},\gamma''|]$.
%%such that $P=[|{\cal A}|]$, then if $P \rightarrow^* P'$, then $\exists$ ${\cal A'}$ such that ${\cal A}$  evolves in ${\cal A'}$ and $P'=[|{\cal A'}|]$.
% \end{corollary}
\begin{proof}
We proceed by induction on the transition number $j \geq 0$.\\
% case $j=0$\\
% It is enough to set $C_0' = C_0 \cup D_0$ and $P' =P$.\\
 case $j=1$\\
 This case falls into the case of Proposition~\ref{prop:corr1}.\\
 case $j>1$\\
By inductive hypothesis, it holds that $\exists$ $P'$ such that $P \xrightarrow{\link{\silent}{\silent}\dots \link{\silent}{\silent}}^{j-1} P'$ and $P' = \llbracket{\cal A},\gamma^{j-1}\rrbracket$.\\
%, with $1 \leq  i < t-1$.\\
As $P'$ is the encoding of an extended interactive process, by Proposition~\ref{prop:corr1}, it exists $P''$  such that $P' \xrightarrow{\link{\silent}{\silent}\dots\link{\silent}{\silent}} P''$,
 and $P'' = \llbracket{\cal A},\gamma^j\rrbracket$. 
% By Lemma~\ref{lem:compl2}.2, if $\exists$ $P'$ such that $P'' \xrightarrow{\link{\silent}{\silent}} P'$, then $P'$ is 
 %unique, and by Lemma~\ref{lem:compl2}.1 $P' = [|{\cal A},\gamma'|]$. 
%By Proposition~\ref{prop:corr1}, we know that  $C_0''=C_1' \cup D_1'$, $C_i''= C_{i+1}'$,
% and $D_i'' = D_{i+1}'$.  And we are done, as $C_0''=C_1' \cup D_1' = C_j \cup D_j$, $C_i''= C_{i+1}' = C_{i+j}$,  and $D_i'' = D_{i+1}'= D_{i+j}$.
\end{proof}

%\noindent
%We recall Proposition~\ref{prop:compl1}
%(the original appears on p.\,\pageref{prop:compl1}).
\propcompluno*

% \begin{proposition}[completeness1]
% \label{lem:compl2}
% Let ${\cal A} = (S, A)$ be a rs, and let  $\pi=(\gamma,\delta)$ be a $n$-step interactive process in $A$, with
% $\gamma = C_0,\dots,C_n$ and $\delta = D_0,\dots,D_n$. Let   $P=[|{\cal A},\gamma|]$ be  its c\CNA~translation. We define $\pi' = (\gamma',\delta')$ with $\gamma' = C_0',\dots,C_{n-1}'$ and $\delta' = D_0',\dots,D_{n-1}'$ as $C_0' = C_1 \cup D_1$,  and
% $C_i' = C_{i+1}$, $D_{i}' = D_{i+1}$ with $ 1 \leq i < n$. Then
% % following holds:
% %\begin{enumerate} 
% %\item  
% $\exists$ $P'$ such that $P \xrightarrow{\link{\silent}{\silent}} P'$ and  $P'=[|{\cal A},\gamma'|]$;
% %\item if $\exists$ $P'$, $P''$ such that $P \xrightarrow{\link{\silent}{\silent}} P'$ and $P \xrightarrow{\link{\silent}{\silent}} P''$, then $P' = P''$.
% %\end{enumerate}
% \end{proposition} 
 \begin{proof}
% {\bf Point 1.}\\
By Proposition~\ref{prop:corr1}, if there exists $P'$ such that $P\xrightarrow{\link{\silent}{\silent}\dots \link{\silent}{\silent} }P'$, then the structure of  $P'$ is deterministically computed.\\
Now, to prove that always exists $P'$, we observe that even in the case no reaction $a$ is applicable in  the interactive process $\pi$ in $A$, then process $P$ can always execute a step transition, as its subprocesses $P_a$ can always execute one of the {\em alternative code
for when reaction $a$ is not applicable} (see Definition~\ref{def:trans}, code for $P_a$ processes).

%The proof proceeds by cases on the possible evolution of each element  $s \in S$, 
%following the evolution from $C_0$ to $C_0'$.\\
%
% %Let  ${\cal A} = (S, A)$ be a reaction system and let $e$ be an element in the system.
% % Then 
%Let $\pi' = (\gamma', \delta')$ be the evolution of  $\pi$ after one step. There are four cases to be considered: \ 
% 1) \ $s \in C_0$, $s \in C_0'$; \ 2)\  $C_0 \in S$, $s \notin C_0'$; \ 3)\  $s \notin C_0$, $s \notin C_0'$; \ 4)\  $s \notin C_0$, $s \in C_0'$.\\
%
%
%Before considering each one of the four previous cases, we analyse the code structure of the processes
%involved in the c\CNA \ transition  mimicking the one step evolution from $C_0$ to $C_0'$.
%
%
%\noindent
%\underline{case  $s \in C_0$ and $s \in C_0'$}\\
%We have to prove that the process $P_s$ is a subprocess of $P'$.\\
%By hypothesis, one of the three following situations must have been took place:
%\begin{enumerate}
%\item at least one rule in $A$ producing the element $s$  has been applied, i.e. $s \in D_1$ (hence $s \in C_0'$), code of point {\bf (a)};
%\item the element $s$ has been provided by the context, i.e. $s \in C_1 $ (hence $s \in C_0'$), code of point {\bf (e)}
%\item both the previous cases 1) and 2)
%\end{enumerate}
%Also, some reaction $a \in A$, requiring the presence of the entity $s$, could have been applied, and all the rules requiring the absence of $s$ could not have been applied.\\
%%Then, let $R_i = [e';e'',e]$ be a rule producing $e$ with $e'' \neq e$, and let $i$ be its associated number. \\
%%By hypothesis, $s \in C_0'$, then it must be that at least a process with code structure of points {\bf (a)}or a process with code structure of point
%
%It follows that process $P_s$ has participated in the transition with code of point {\bf (f)} with $z \geq 0$, or with code of point {\bf (g)}. Then we have that  $P' \defeq \restrict{reacts,ctxs,ents}(P_s | B)$.\\
%
%\noindent
%The other cases can be proved in a  similar way.\\
\end{proof}

%\noindent
%We recall Corollary~\ref{corr:compl2}
%(the original appears on p.\,\pageref{corr:compl2}).
\corrcompldue*

% \begin{corollary}[completeness2]
% \label{prop:completeness}
%  Let ${\cal A} = (S, A)$ be a $rs$, and let $\pi=(\gamma,\delta)$ be a $n$-step interactive process in ${\cal A}$ with $\gamma = C_0,\dots,C_n$ and $\delta = D_0,\dots,D_n$. Let $P=[|{\cal A},\gamma|]$ its c\CNA \ translation. If we define $\pi'= (\gamma',\delta')$  with   $\gamma' = C_0',\dots,C_{n-1}'$ and $\delta' = D_0',\dots,D_{n-1}'$  as
%  $C_0'=C_w \cup D_w$, and   $C_i'=C_{i+w}$, $D_i' = D_{i+w}$ with $1 \leq i \leq n-w $, then $\exists$ $P'$ such that $P \rightarrow^w P'$ and  $P'=[|{\cal A},\gamma'|]$.  
% \end{corollary}
 \begin{proof}
 The proof proceeds by induction on the number $j$, and it is similar to the one of Corollary~\ref{corr:corr2}\\
% case $j=0$\\
% It is enough to set $C_0' = C_0 \cup D_0$ and $P' =P$.\\
% case $j=1$\\
% This case falls into the case of Proposition~\ref{prop:compl1}.\\
% case $j>1$\\
% We set $C_0'=C_w \cup D_w = C_1'' \cup D_1''$, $C_g' = C_{n-w}= C_{g+1}''$, and $D_g' = D_{n-w} = D_{g+1}''$, with $1 \leq  g < n-w$.\\
% By inductive hypothesis, it holds that $\exists$ $P''$ such that $P \xrightarrow{\link{\silent}{\silent}}^{w-1} P''$ and $P'' = [| {\cal A},\gamma'|]$.
% By Lemma~\ref{lem:compl2}.1, it exists $P'$ such that $P'' \xrightarrow{\link{\silent}{\silent}} P'$,
% and $P' = [|{\cal A},\gamma'|]$. By Lemma~\ref{lem:compl2}.2, $P'$ is unique.
%% By Lemma~\ref{lem:compl2}.2, if $\exists$ $P'$ such that $P'' \xrightarrow{\link{\silent}{\silent}} P'$, then $P'$ is 
% %unique, and by Lemma~\ref{lem:compl2}.1 $P' = [|{\cal A},\gamma'|]$. 
% And we are done. 
 \end{proof}

 \newpage

 \section{The \emph{lac} operon encoding}
 \begin{table}[h]
{\small
\[
\boxed{
\begin{array}{lcl}
P_{cAMP\textrm{-}CAP} & \defeq & %P_{cAMP\textrm{-}CAP}(cAMP\textrm{-}CAP)=\\
\sum_{h =0}^1 ( \startchain{cAMP\textrm{-}CAP_i}\chainedlink{cAMP\textrm{-}CAP_o}{\noact}\chainend{\noact})^h\,  \startchain{\widetilde{cAMP\textrm{-}CAP}_i}\chainend{\widetilde{cAMP\textrm{-}CAP}_o}.P_{cAMP\textrm{-}CAP}\\
&&+\\
&&\link{cAMP\textrm{-}CAP_i}{cAMP\textrm{-}CAP_o} .\overline{P_{cAMP\textrm{-}CAP}}\\
\\
\overline{P_{cAMP\textrm{-}CAP}} & \defeq &
\sum_{h =0}^1  ( \startchain{\overline{cAMP\textrm{-}CAP}_i}\chainedlink{\overline{cAMP\textrm{-}CAP}_o}{\noact}\chainend{\noact})^h\ \link{\widetilde{cAMP\textrm{-}CAP}_i}{\widetilde{cAMP\textrm{-}CAP}_o} .P_{cAMP\textrm{-}CAP}\\
&&+\\
&&\link{\overline{cAMP\textrm{-}CAP}_i}{\overline{cAMP\textrm{-}CAP}_o} .\overline{P_{cAMP\textrm{-}CAP}}\\
\end{array}}
\]}
\caption{The process for $cAMP\textrm{-}CAP$.}\label{tab:cAMP-CAP}
%\end{table}
%
%\begin{table}[h]
\[
\boxed{
\begin{array}{lcl}
P_{glucose} &\defeq &
\sum_{h=0}^1 (\startchain{glucose_i}\chainedlink{glucose_o}{\noact}\chainend{\noact})^h\ \link{\widehat{glucose_i}}{\widehat{glucose_o}}.P_{glucose}
\\
&& + \\
&& \sum_{h=0}^1 (\startchain{glucose_i}\chainedlink{glucose_o}{\noact}\chainend{\noact})^h\ \link{\underline{glucose_i}}{\underline{glucose_o}}.\overline{P_{glucose}}\\
\\
\overline{P_{glucose}} &\defeq &\sum_{h=0}^1 (\startchain{\overline{glucose}_i}\chainedlink{\overline{glucose}_o}{\noact}\chainend{\noact})^h\ \link{\widehat{glucose_i}}{\widehat{glucose_o}}.P_{glucose}\\
&&+\\
&& \sum_{h=0}^1 (\startchain{\overline{glucose}_i}\chainedlink{\overline{glucose}_o}{\noact}\chainend{\noact})^h\ \link{\underline{glucose_i}}{\underline{glucose_o}}.\overline{P_{glucose}}\\
\end{array}}
\]
\caption{The process for $glucose$.}\label{tab:glucose}
%\end{table}
%
%\begin{table}[h]
\[
\boxed{
\begin{array}{lcl@{\qquad}lcl}
P_A&\defeq &\startchain{\tilde{A_i}}\chainedlink{\tilde{A_o}}{\noact}\chainedlink{\noact}{\underline{A_i}}\chainend{\underline{A_o}}.P_A\ + \
\link{\underline{A_i}}{\underline{A_o}}.\overline{P_A}
&
P_y&\defeq &\startchain{\tilde{y_i}}\chainedlink{\tilde{y_o}}{\noact}\chainedlink{\noact}{\underline{y_i}}\chainend{\underline{y_o}}.P_y\ + \
\link{\underline{y_i}}{\underline{y_o}}.\overline{P_y}
\\
\overline{P_A} &\defeq &\startchain{\tilde{A_i}}\chainedlink{\tilde{A_o}}{\noact}\chainedlink{\noact}{\underline{A_i}}\chainend{\underline{A_o}}.\overline{P_A}\ + \
\link{\underline{A_i}}{\underline{A_o}}.\overline{P_A}
&
\overline{P_y} &\defeq &\startchain{\tilde{y_i}}\chainedlink{\tilde{y_o}}{\noact}\chainedlink{\noact}{\underline{y_i}}\chainend{\underline{y_o}}.\overline{P_y}\ + \
\link{\underline{y_i}}{\underline{y_o}}.\overline{P_y}
\end{array}}
\]
\caption{The processes for $y$ and $A$.}\label{tab:yA}

\end{table}

### 2020-04-02 04-CCNA.tex
changed biblio and other files
some small changes in several files

@@
  For example, for $R = \{a,b\}$ the expression $ \blockchain{c_i}{c_o}{c \in R} $\;
-denotes the block of chains ${}_{a_i}^{\noact}\chainedlink{\noact}{a_o} \backslash{}_{b_i}^{\noact}\chainedlink{\noact}{b_o}$; and the expression $\link{r_1}{} \blockchain{c_i}{c_o}{c \in R}\;\link{}{r_2} $ denotes the chain $ \startchain{r_1}\chainedlink{a_i}{\noact}\chainedlink{\noact}{a_o}\chainedlink{b_i}{\noact}\chainedlink{\noact}{b_o}\chainend{r_2}$.
+denotes the block of chains ${}_{a_i}^{\noact}\chainedlink{\noact}{a_o} \backslash{}_{b_i}^{\noact}\chainedlink{\noact}{b_o}$; and the expression $\link{r_1}{} \blockchain{c_i}{c_o}{c \in R}\;\link{}{r_2} $ denotes the chain \\
+$ \startchain{r_1}\chainedlink{a_i}{\noact}\chainedlink{\noact}{a_o}\chainedlink{b_i}{\noact}\chainedlink{\noact}{b_o}\chainend{r_2}$.

### 2020-04-02 06b-bisimulation.tex
changed biblio and other files
some small changes in several files

@@
 
-Now, we introduce a slightly modified version of the Hennessy Milner Logic~\cite{}, called bioHML; due to the reasons we explained above, we do not want to look at the complete transition labels, thus we rely on our simple assertion language to make it parametric to the assertion $F$ of interest:
+Now, we introduce a slightly modified version of the Hennessy Milner Logic~\cite{HM80}, called bioHML; due to the reasons we explained above, we do not want to look at the complete transition labels, thus we rely on our simple assertion language to make it parametric to the assertion $F$ of interest:
 \begin{definition}[bioHML]

### 2020-04-02 05-translation.tex
modificato file 05
modificato file 05

@@
 \begin{array}{lcl@{\quad}l}
-P_{aj}& \defeq & 
+P_{aj} \defeq \\ 
  ^{r_j} \backslash  \blockchain{s_i}{s_o}{s \in R_j} \, \link{}{} \blockchain{\overline{e}_i}{\overline{e}_o}{e \in I_j} \, \chainedlink{r_{j+1}}{\noact}\chainedlink{\noact}{p_j}\link{}{} \blockchain{\widetilde{c}_i}{\widetilde{c}_o}{c \in P_j}\, \backslash_{p_{j+1}}.P_{aj} & \mbox{\small  \{$aj$ is applicable\}}\\
-&& + \ \\
-&& \sum_{s\in R_j}\startchain{r_j}\chainedlink{\overline{s}_i}{\noact}\chainedlink{\noact}{\overline{s}_o} \chainedlink{r_{j+1}}{\noact}\chainedlink{\noact}{p_{j}}\chainend{p_{j+1}}.P_{aj}  & \mbox{\small  \{$aj$ is not applicable\}}\\
-&&+\\
-&& \sum_{e\in I_j}\startchain{r_j}\chainedlink{e_i}{\noact} \chainedlink{\noact}{e_o} \chainedlink{r_{i+1}}{\noact}\chainedlink{\noact}{p_{j}}\chainend{p_{j+1}}.P_{aj}  & \mbox{\small \{$aj$ is not applicable\}}\\
+ + \ \\
+ \sum_{s\in R_j}\startchain{r_j}\chainedlink{\overline{s}_i}{\noact}\chainedlink{\noact}{\overline{s}_o} \chainedlink{r_{j+1}}{\noact}\chainedlink{\noact}{p_{j}}\chainend{p_{j+1}}.P_{aj}  & \mbox{\small  \{$aj$ is not applicable\}}\\
++\\
+ \sum_{e\in I_j}\startchain{r_j}\chainedlink{e_i}{\noact} \chainedlink{\noact}{e_o} \chainedlink{r_{i+1}}{\noact}\chainedlink{\noact}{p_{j}}\chainend{p_{j+1}}.P_{aj}  & \mbox{\small \{$aj$ is not applicable\}}\\
 \end{array}
@@
  .P_{s1} + \dots; \\   
-  P_{r_2} &\defeq& \dots + \startchain{r_2}\chainedlink{\overline{s2}_i}{\noact}\chainedlink{\noact}{\overline{s2}_o}\chainedlink{\mathit{cxt}}{\noact} \chainedlink{\noact}{p_2}\chainend{\silent}.P_{r_2} + \dots; &
+  P_{r_2} &\defeq& \dots + \startchain{r_2}\chainedlink{\overline{s2}_i}{\noact}\chainedlink{\noact}{\overline{s2}_o}\chainedlink{\mathit{cxt}}{\noact} \chainedlink{\noact}{p_2}\chainend{\silent}.P_{r_2} + \dots; \\
   \overline{P}_{s2} &\defeq& \startchain{\overline{s2}_i}\chainedlink{\overline{s2}_o}{\noact}

### 2020-04-02 09-appendix.tex
updated appendix

@@
  $j \in [1,\dots,u]$, and $cxt_h$, with $h \in [1, \dots, w]$, and all the annotated versions of 
- $s_i$, $s_o$ are restricted. Also, processes $Cxt_s$ always requires the 
- interaction with $P_s$ on either on channels $\hat{s}_i$, $\hat{s}_o$ or on channels 
+ $s_i$, $s_o$ are restricted. Also, processes $\mathit{Cxt}$ always requires the 
+ interaction with $P_s$ on either on channels $\widehat{s}_i$, $\widehat{s}_o$ or on channels 
  $\underline{s}_i$, $\underline{s}_o$.
- It derives that all the processes: $P_a$  (coding the behaviour of reaction $a \in A$), $P_s$ (coding the behaviour of entity $s \in S$), and $Cxt_s$ (coding the behaviour of the context regarding the entity $s$) have been involved in the transition.
-We have the following cases:
+ It derives that all the processes: $P_a$  (coding the behaviour of reaction $a \in A$), $P_s$ (coding the behaviour of entity $s \in S$), and $\mathit{Cxt}$ (coding the behaviour of the context regarding all the entities) have been involved in the transition.
+
+For any process $P_{aj}$ encoding a reaction $aj$ we have the following cases:
 \begin{description}
-\item[(a)] processes $P_a$ encoding a reaction $a$,  with serial  number $j$, producing the entity $s$ provide a code of this type:
+\item[(a)] 
+if $aj$ is applicable and it produces the entity $s$,  
+the process $P_{aj}$ provides a code of this type:
 %By Lemma~\ref{lem:compl1}, all the processes $A_j$, and $Cxt_h$ participated in the tr
 %The code executed by any process encoding rule $R_i$ producing $e$ is 
-$P_a\defeq \link{r_j}{\dots}\chainedlink{r_{j+1}}{\noact}\chainedlink{\noact}{p_j}\link{}{\dots}\chainedlink{\tilde{s}_i}{\noact}\chainedlink{\noact}{\dots}\link{\tilde{s}_o}{\dots} \chainend{p_{j+1}}.P_a $;
-\item[(b)] processes $P_a$ encoding a reaction $a$, with serial number $j$, consuming the entity $s$
-% without producing it, 
-provide a code of this type: 
-$P_a\defeq \link{r_j}{\dots}\chainedlink{s_i}{\noact}\chainedlink{\noact}{}\link{s_o}{\dots} \chainedlink{r_{j+1}}{\noact}\chainedlink{\noact}{p_j}\link{}{\dots}\link{}{p_{j+1}}.P_a $;
-\item[(c)]  processes $P_a$ encoding a reaction $a$, with serial number $j$, requiring the absence of the entity $s$ provide a code of this type: \\
-$P_a\defeq \link{r_j}{\dots}\chainedlink{\overline{s}_i}{\noact}\chainedlink{\noact}{}\link{\overline{s}_o}{\dots} \chainedlink{r_{j+1}}{\noact}\chainedlink{\noact}{p_j}\link{}{\dots}\link{}{p_{j+1}}.P_a $;
-\item[(d)] processes $P_a$ encoding a reaction $a$ (that cannot be applied), with serial number $j$, 
-execute a code  capturing either the absence of one of its reactants (case 1), or the presence  of one of its inhibitors (case 2):
+$$P_{aj}\defeq \link{r_j}{\dots}\chainedlink{r_{j+1}}{\noact}\chainedlink{\noact}{p_j}\link{}{\dots}\chainedlink{\widetilde{s}_i}{\noact}\chainedlink{\noact}{\dots}\link{\widetilde{s}_o}{\dots} \chainend{p_{j+1}}.P_{aj} ;$$
+
+\item[(b)] 
+ if $aj$ is applicable and it  consumes the entity $s$,
+ the process $P_{aj}$ provides a code of this type: 
+$$P_{aj}\defeq \link{r_j}{\dots}\chainedlink{s_i}{\noact}\chainedlink{\noact}{}\link{s_o}{\dots} \chainedlink{r_{j+1}}{\noact}\chainedlink{\noact}{p_j}\link{}{\dots}\link{}{p_{j+1}}.P_{aj} ;$$
+
+\item[(c)]  
+ if $aj$ is applicable and it requires the absence of the entity $s$,
+ the process $P_{aj}$  provides a code of this type: 
+$$P_{aj}\defeq \link{r_j}{\dots}\chainedlink{\overline{s}_i}{\noact}\chainedlink{\noact}{}\link{\overline{s}_o}{\dots} \chainedlink{r_{j+1}}{\noact}\chainedlink{\noact}{p_j}\link{}{\dots}\link{}{p_{j+1}}.P_{aj} ;$$
+
+\item[(d)] 
+ if $aj$ is not applicable,
+the process $P_{aj}$ executes a code  capturing either the absence of one of its reactants (case 1), or the presence  of one of its inhibitors (case 2):
 \begin{enumerate}
-\item $P_a\defeq \link{r_j}{\dots}\chainedlink{\overline{s}_i}{\noact}\chainedlink{\noact}{}\link{\overline{s}_o}{\dots} \chainedlink{r_{j+1}}{\noact}\chainedlink{\noact}{p_j}\link{}{\dots}\link{}{p_{j+1}}.P_a $.
-\item $P_a\defeq \link{r_j}{\dots}\chainedlink{s_i}{\noact}\chainedlink{\noact}{}\link{s_o}{\dots} \chainedlink{r_{j+1}}{\noact}\chainedlink{\noact}{p_j}\link{}{\dots}\link{}{p_{j+1}} .P_a $.
+\item $P_{aj}\defeq \link{r_j}{\dots}\chainedlink{\overline{s}_i}{\noact}\chainedlink{\noact}{}\link{\overline{s}_o}{\dots} \chainedlink{r_{j+1}}{\noact}\chainedlink{\noact}{p_j}\link{}{\dots}\link{}{p_{j+1}}.P_{aj} $;
+\item $P_{aj}\defeq \link{r_j}{\dots}\chainedlink{s_i}{\noact}\chainedlink{\noact}{}\link{s_o}{\dots} \chainedlink{r_{j+1}}{\noact}\chainedlink{\noact}{p_j}\link{}{\dots}\link{}{p_{j+1}} .P_{aj}$.
 \end{enumerate}
@@
 %involve the entity $s$.\\
-Now, we consider the structure of the process $Cxt_s = Cxt^0_s$. By Definition~\ref{def:trans}, $Cxt^0_s$ is the unique process encoding the behaviour of the context regulating the supply of $s$. 
-The code of the process $Cxt^t_s$, with $t \geq 0$, has the following structure: 
+Now, we consider the structure of the process $\mathit{Cxt}^1$. 
+By Definition~\ref{def:trans}, $Cxt^1$ is the unique process encoding the behaviour of the context regulating the supply of any entity.
+
 \begin{description}
-\item[(e)] $Cxt^i_s \defeq  \startchain{cxt_h}\chainedlink{\hat{s}_i}{\noact}\chainedlink{\noact}{\hat{s}_o}\chainend{cxt_{h+1}}. Cxt^{i+1}_{s}$, if $s \in C_{i+1}$;
-\item[(f)] $Cxt^i_s \defeq  \startchain{cxt_h}\chainedlink{\underline{s}_i}{\noact}\chainedlink{\noact}{\underline{s}_o}\chainend{cxt_{h+1}}. Cxt^{i+1}_{s}$, if $s \notin C_{i+1}$;
-%\item[(e)] $Cxt_{s}^1 \defeq \startchain{cxt_h}\chainedlink{\hat{s}_i}{\noact}\chainedlink{\noact}{\hat{s}_o}\chainend{cxt_{h+1}}. Cxt_{s}^{2}$, if  $s \in C_1$ (and hence $s \in C_0')$;
-%\item[(f)] $Cxt_{s}^1 \defeq \startchain{cxt_h}\chainedlink{\underline{s}_i}{\noact}\chainedlink{\noact}{\underline{s}_o}\chainend{cxt_{h+1}}. Cxt_{s}^{2}$, if  $s \notin C_1$.
+\item[(e)] 
+The code of the process $Cxt^1$ that provides $s$ and not $e$ has the following structure: 
+$$
+\mathit{Cxt}^1 \defeq
+\link{\mathit{cxt}}{\dots}
+\chainedlink{\widehat{s}_i}{\noact}\chainedlink{\noact}{}\link{\widehat{s}_o}{\dots} 
+\chainedlink{\underline{e}_i}{\noact}\chainedlink{\noact}{}\link{\underline{e}_o}{\dots} 
+\chainend{p_{1}}. \mathit{Cxt}^{2} .
+$$
 \end{description}
 
-%$R_j \defeq \startchain{rule_j}\chainedlink{e_i}{\noact}\chainedlink{\noact}{e_o}\chainedlink{\overline{e'}_i}{\noact} \chainedlink{\noact}{\overline{e'}_o}\chainedlink{\tilde{e''}_i}{\noact}\chainedlink{\noact}{\tilde{e''}_o} \chainend{rule_{j+1}}.R_j + \dots$, with $e',e''\neq e $.\\
+%$R_j \defeq \startchain{rule_j}\chainedlink{e_i}{\noact}\chainedlink{\noact}{e_o}\chainedlink{\overline{e'}_i}{\noact} \chainedlink{\noact}{\overline{e'}_o}\chainedlink{\widetilde{e''}_i}{\noact}\chainedlink{\noact}{\widetilde{e''}_o} \chainend{rule_{j+1}}.R_j + \dots$, with $e',e''\neq e $.\\
 %The code executed by any rule $j$ not requiring $e$ is $R_j \defeq \startchain{rule_j}\chainedlink{e_i}{\noact}\chainedlink{\noact}{e_o} \chainend{rule_{j+1}}.R_j + \dots$.\\
@@
 \begin{description}
-\item[(g)] $P_s \defeq  \sum_{h,k\geq0}(\startchain{s_i}\chainedlink{s_o}{\noact}\chainend{\noact}) ^h\, \startchain{\hat{s}_i}\chainedlink{\hat{s}_o}{\noact}\chainend{\noact}
- (\startchain{\tilde{s}_i}\chainedlink{\tilde{s}_o}{\noact}\chainend{\noact})^k\,.P_s $, if $s \in C_{i+1}$;
-% \underbrace{ \startchain{s_i}\chainedlink{s_o}{\noact}\chainedlink{\noact}{s_i}\chainedlink{s_o}{\noact}\chainend{\noact}...}_{v\ times}\underbrace{ \startchain{\tilde{s}_i}\chainedlink{\tilde{s}_o}{\noact}\chainedlink{\noact}{\tilde{s}_i}\chainedlink{\tilde{s}_o}{\noact}\chainend{\noact}...}_{z\ times}\link{\hat{s}_i}{\hat{s}_o}.P_s $, if $s \in C_{i+1}$;
-%\item[()] $P_s \defeq  \underbrace{ \startchain{s_i}\chainedlink{s_o}{\noact}\chainedlink{\noact}{s_i}\chainedlink{s_o}{\noact}\chainend{\noact}...}_{v\ times}\link{\hat{s}_i}{\hat{s}_o}.P_s $, if $s \in C_1$ %(and hence $s \in C_0'$)
-\item[(h)] $P_s \defeq \sum_{h \geq 0,k\geq 1}(\startchain{s_i}\chainedlink{s_o}{\noact}\chainend{\noact})^h\ \startchain{\underline{s}_i}\chainedlink{\underline{s}_o}{\noact}\chainend{\noact} (\startchain{\tilde{s}_i}\chainedlink{\tilde{s}_o}{\noact}\chainend{\noact})^k
+\item[(f)] 
+$P_s \defeq  \sum_{h,k\geq0}(\startchain{s_i}\chainedlink{s_o}{\noact}\chainend{\noact}) ^h\, \startchain{\widehat{s}_i}\chainedlink{\widehat{s}_o}{\noact}\chainend{\noact}
+ (\startchain{\widetilde{s}_i}\chainedlink{\widetilde{s}_o}{\noact}\chainend{\noact})^k\,.P_s $, if $s \in C_{i+1}$;
+% \underbrace{ \startchain{s_i}\chainedlink{s_o}{\noact}\chainedlink{\noact}{s_i}\chainedlink{s_o}{\noact}\chainend{\noact}...}_{v\ times}\underbrace{ \startchain{\widetilde{s}_i}\chainedlink{\widetilde{s}_o}{\noact}\chainedlink{\noact}{\widetilde{s}_i}\chainedlink{\widetilde{s}_o}{\noact}\chainend{\noact}...}_{z\ times}\link{\widehat{s}_i}{\widehat{s}_o}.P_s $, if $s \in C_{i+1}$;
+%\item[()] $P_s \defeq  \underbrace{ \startchain{s_i}\chainedlink{s_o}{\noact}\chainedlink{\noact}{s_i}\chainedlink{s_o}{\noact}\chainend{\noact}...}_{v\ times}\link{\widehat{s}_i}{\widehat{s}_o}.P_s $, if $s \in C_1$ %(and hence $s \in C_0'$)
+\item[(g)] 
+$P_s \defeq \sum_{h \geq 0,k\geq 1}(\startchain{s_i}\chainedlink{s_o}{\noact}\chainend{\noact})^h\ \startchain{\underline{s}_i}\chainedlink{\underline{s}_o}{\noact}\chainend{\noact} (\startchain{\widetilde{s}_i}\chainedlink{\widetilde{s}_o}{\noact}\chainend{\noact})^k
 .P_s$, if $s \not\in C_{i+1}$
-% \underbrace{ \startchain{s_i}\chainedlink{s_o}{\noact}\chainedlink{\noact}{s_i}\chainedlink{s_o}{\noact}\chainend{\noact}...}_{v\ times}\underbrace{ \startchain{\tilde{s}_i}\chainedlink{\tilde{s}_o}{\noact}\chainedlink{\noact}{\tilde{s}_i}\chainedlink{\tilde{s}_o}{\noact}\chainend{\noact}...}_{z >0 \ times}\link{\underline{s}_i}{\underline{s}_o}.P_s$, if $s \in C_{i+1}$
-\item[(i)] $P_s \defeq  \sum_{h\geq0} (\startchain{s_i}\chainedlink{s_o}{\noact}\chainend{\noact})^h
+% \underbrace{ \startchain{s_i}\chainedlink{s_o}{\noact}\chainedlink{\noact}{s_i}\chainedlink{s_o}{\noact}\chainend{\noact}...}_{v\ times}\underbrace{ \startchain{\widetilde{s}_i}\chainedlink{\widetilde{s}_o}{\noact}\chainedlink{\noact}{\widetilde{s}_i}\chainedlink{\widetilde{s}_o}{\noact}\chainend{\noact}...}_{z >0 \ times}\link{\underline{s}_i}{\underline{s}_o}.P_s$, if $s \in C_{i+1}$
+\item[(h)] 
+$P_s \defeq  \sum_{h\geq0} (\startchain{s_i}\chainedlink{s_o}{\noact}\chainend{\noact})^h
 %{s_i}\chainedlink{s_o}{\noact}\chainend{\noact}...}_{v\ times}
@@
 where, by Lemma~\ref{lem:struct}.2, $h$ is the number of reactions requiring the presence of $s$ plus possibly some reactions not requiring $s$; and $k$ is the number of reactions producing $s$.
-The code executed by  $\overline{P}_s$ has the following structure:
+
+Similarly, the code executed by  $\overline{P}_s$ has the following structure:
 \begin{description}
-\item[(g')] $\overline{P}_s \defeq  \sum_{h,k\geq 0}(\startchain{\overline{s}_i}\chainedlink{\overline{s}_o}{\noact}\chainend{\noact})^h  \, \startchain{\hat{s}_i}\chainedlink{\hat{s}_o}{\noact}\chainend{\noact}
- (\startchain{\tilde{s}_i}\chainedlink{\tilde{s}_o}{\noact}\chainend{\noact})^k\,.P_s $, if $s \in C_{i+1}$;
-% \underbrace{ \startchain{\overline{s}_i}\chainedlink{\overline{s}_o}{\noact}\chainedlink{\noact}{\overline{s}_i}\chainedlink{\overline{s}_o}{\noact}\chainend{\noact}...}_{v\ times}\underbrace{ \startchain{\tilde{s}_i}\chainedlink{\tilde{s}_o}{\noact}\chainedlink{\noact}{\tilde{s}_i}\chainedlink{\tilde{s}_o}{\noact}\chainend{\noact}...}_{z\ times}\link{\hat{s}_i}{\hat{s}_o}.P_s $, if $s \in C_{i+1}$;
-%\item[()] $P_s \defeq  \underbrace{ \startchain{s_i}\chainedlink{s_o}{\noact}\chainedlink{\noact}{s_i}\chainedlink{s_o}{\noact}\chainend{\noact}...}_{v\ times}\link{\hat{s}_i}{\hat{s}_o}.P_s $, if $s \in C_1$ %(and hence $s \in C_0'$)
-\item[(h')] $\overline{P}_s \defeq  \sum_{h \geq 0,k\geq 1} (\startchain{\overline{s}_i}\chainedlink{\overline{s}_o}{\noact}\chainend{\noact})^h\ \startchain{\underline{s}_i}\chainedlink{\underline{s}_o}{\noact}\chainend{\noact} (\startchain{\tilde{s}_i}\chainedlink{\tilde{s}_o}{\noact}\chainend{\noact})^k.P_s$, if $s \notin C_{i+1}$
+\item[(f')] 
+$\overline{P}_s \defeq  \sum_{h,k\geq 0}(\startchain{\overline{s}_i}\chainedlink{\overline{s}_o}{\noact}\chainend{\noact})^h  \, \startchain{\widehat{s}_i}\chainedlink{\widehat{s}_o}{\noact}\chainend{\noact}
+ (\startchain{\widetilde{s}_i}\chainedlink{\widetilde{s}_o}{\noact}\chainend{\noact})^k\,.P_s $, if $s \in C_{i+1}$;
+% \underbrace{ \startchain{\overline{s}_i}\chainedlink{\overline{s}_o}{\noact}\chainedlink{\noact}{\overline{s}_i}\chainedlink{\overline{s}_o}{\noact}\chainend{\noact}...}_{v\ times}\underbrace{ \startchain{\widetilde{s}_i}\chainedlink{\widetilde{s}_o}{\noact}\chainedlink{\noact}{\widetilde{s}_i}\chainedlink{\widetilde{s}_o}{\noact}\chainend{\noact}...}_{z\ times}\link{\widehat{s}_i}{\widehat{s}_o}.P_s $, if $s \in C_{i+1}$;
+%\item[()] $P_s \defeq  \underbrace{ \startchain{s_i}\chainedlink{s_o}{\noact}\chainedlink{\noact}{s_i}\chainedlink{s_o}{\noact}\chainend{\noact}...}_{v\ times}\link{\widehat{s}_i}{\widehat{s}_o}.P_s $, if $s \in C_1$ %(and hence $s \in C_0'$)
+\item[(g')] 
+$\overline{P}_s \defeq  \sum_{h \geq 0,k\geq 1} (\startchain{\overline{s}_i}\chainedlink{\overline{s}_o}{\noact}\chainend{\noact})^h\ \startchain{\underline{s}_i}\chainedlink{\underline{s}_o}{\noact}\chainend{\noact} (\startchain{\widetilde{s}_i}\chainedlink{\widetilde{s}_o}{\noact}\chainend{\noact})^k.P_s$, if $s \notin C_{i+1}$
 
-% \underbrace{ \startchain{\overline{s}_i}\chainedlink{\overline{s}_o}{\noact}\chainedlink{\noact}{\overline{s}_i}\chainedlink{\overline{s}_o}{\noact}\chainend{\noact}...}_{v\ times}\underbrace{ \startchain{\tilde{s}_i}\chainedlink{\tilde{s}_o}{\noact}\chainedlink{\noact}{\tilde{s}_i}\chainedlink{\tilde{s}_o}{\noact}\chainend{\noact}...}_{z >0 \ times}\link{\underline{s}_i}{\underline{s}_o}.P_s$, if $s \in C_{i+1}$
-\item[(i')] $\overline{P}_s \defeq  \sum_{h \geq 0}(\startchain{\overline{s}_i}\chainedlink{\overline{s}_o}{\noact}\chainend{\noact})^h\,\link{\underline{s}_i}{\underline{s}_o}.\overline{P}_s$, if $s \notin C_{i+1}$
+% \underbrace{ \startchain{\overline{s}_i}\chainedlink{\overline{s}_o}{\noact}\chainedlink{\noact}{\overline{s}_i}\chainedlink{\overline{s}_o}{\noact}\chainend{\noact}...}_{v\ times}\underbrace{ \startchain{\widetilde{s}_i}\chainedlink{\widetilde{s}_o}{\noact}\chainedlink{\noact}{\widetilde{s}_i}\chainedlink{\widetilde{s}_o}{\noact}\chainend{\noact}...}_{z >0 \ times}\link{\underline{s}_i}{\underline{s}_o}.P_s$, if $s \in C_{i+1}$
+\item[(h')] 
+$\overline{P}_s \defeq  \sum_{h \geq 0}(\startchain{\overline{s}_i}\chainedlink{\overline{s}_o}{\noact}\chainend{\noact})^h\,\link{\underline{s}_i}{\underline{s}_o}.\overline{P}_s$, if $s \notin C_{i+1}$
 \end{description}
  \noindent
-where, by Lemma~\ref{lem:struct}.2, $h$ is the number of reactions requiring the absence of $s$ plus possibly some reactions  requiring $s$; and $k$ is the number of reactions producing $s$.\\
+where, by Lemma~\ref{lem:struct}.2, $h$ is the number of reactions requiring the absence of $s$ plus possibly some reactions  requiring $s$; and $k$ is the number of reactions producing $s$.
 
-\noindent 
 It is worth nothing that, depending on the presence ($P_s$) or the absence ($\overline{P}_s$) of each entity $s$, for each process $P_a$ (encoding a reaction $a$) the choice between the execution of the reaction code (points {\bf (a)}, {\bf (b)}, {\bf (c)}) or  the code expressing that reaction $a$ is not applicable (point {\bf (d)}) is deterministic.
-Also, the building of the code of processes $Cxt_s$ (points {\bf (e)}, {\bf (f)}), follows the evolution of $\gamma$.
-It derives that the trend followed by the processes $P_s$ (or $\overline{P}_s$) is also deterministic (points {\bf (g)}, {\bf (h)}, {\bf (i)} or {\bf (g')}, {\bf (h')}, {\bf (i')}), leading to $P' = \llbracket {\cal A},\gamma^1\rrbracket$.
+Also, the building of the code of process $\mathit{Cxt}$ (points {\bf (e)}, {\bf (f)}), is univocally determined by the evolution of $\gamma$.
+It derives that the trend followed by the processes $P_s$ (or $\overline{P}_s$) is also deterministic (points {\bf (f)}, {\bf (g)}, {\bf (h)} or {\bf (f')}, {\bf (g')}, {\bf (h')}), leading to $P' = \llbracket {\cal A},\gamma^1\rrbracket$.
  \end{proof}
@@
 \begin{proof}
-We proceed by induction on the transition number $j \geq 0$.\\
+We proceed by induction on the transition number $j \geq 0$.
 % case $j=0$\\
 % It is enough to set $C_0' = C_0 \cup D_0$ and $P' =P$.\\
- case $j=1$\\
- This case falls into the case of Proposition~\ref{prop:corr1}.\\
- case $j>1$\\
-By inductive hypothesis, it holds that $\exists$ $P'$ such that $P \xrightarrow{\link{\silent}{\silent}\dots \link{\silent}{\silent}}^{j-1} P'$ and $P' = \llbracket{\cal A},\gamma^{j-1}\rrbracket$.\\
+\begin{description}
+\item[base case $j=1$:]
+ This case falls into the case of Proposition~\ref{prop:corr1}.
+\item[inductive case:]
+We assume, by inductive hypothesis, that $\exists P'$ such that 
+$$P \xrightarrow{\link{\silent}{\silent}\dots \link{\silent}{\silent}}^{j-1} P'$$ 
+and $P' = \llbracket{\cal A},\gamma^{j-1}\rrbracket$.
 %, with $1 \leq  i < t-1$.\\
@@
 % and $D_i'' = D_{i+1}'$.  And we are done, as $C_0''=C_1' \cup D_1' = C_j \cup D_j$, $C_i''= C_{i+1}' = C_{i+j}$,  and $D_i'' = D_{i+1}'= D_{i+j}$.
+\end{description}
 \end{proof}
@@
 \begin{array}{lcl@{\qquad}lcl}
-P_A&\defeq &\startchain{\tilde{A_i}}\chainedlink{\tilde{A_o}}{\noact}\chainedlink{\noact}{\underline{A_i}}\chainend{\underline{A_o}}.P_A\ + \
+P_A&\defeq &\startchain{\widetilde{A_i}}\chainedlink{\widetilde{A_o}}{\noact}\chainedlink{\noact}{\underline{A_i}}\chainend{\underline{A_o}}.P_A\ + \
 \link{\underline{A_i}}{\underline{A_o}}.\overline{P_A}
 &
-P_y&\defeq &\startchain{\tilde{y_i}}\chainedlink{\tilde{y_o}}{\noact}\chainedlink{\noact}{\underline{y_i}}\chainend{\underline{y_o}}.P_y\ + \
+P_y&\defeq &\startchain{\widetilde{y_i}}\chainedlink{\widetilde{y_o}}{\noact}\chainedlink{\noact}{\underline{y_i}}\chainend{\underline{y_o}}.P_y\ + \
 \link{\underline{y_i}}{\underline{y_o}}.\overline{P_y}
 \\
-\overline{P_A} &\defeq &\startchain{\tilde{A_i}}\chainedlink{\tilde{A_o}}{\noact}\chainedlink{\noact}{\underline{A_i}}\chainend{\underline{A_o}}.\overline{P_A}\ + \
+\overline{P_A} &\defeq &\startchain{\widetilde{A_i}}\chainedlink{\widetilde{A_o}}{\noact}\chainedlink{\noact}{\underline{A_i}}\chainend{\underline{A_o}}.\overline{P_A}\ + \
 \link{\underline{A_i}}{\underline{A_o}}.\overline{P_A}
 &
-\overline{P_y} &\defeq &\startchain{\tilde{y_i}}\chainedlink{\tilde{y_o}}{\noact}\chainedlink{\noact}{\underline{y_i}}\chainend{\underline{y_o}}.\overline{P_y}\ + \
+\overline{P_y} &\defeq &\startchain{\widetilde{y_i}}\chainedlink{\widetilde{y_o}}{\noact}\chainedlink{\noact}{\underline{y_i}}\chainend{\underline{y_o}}.\overline{P_y}\ + \
 \link{\underline{y_i}}{\underline{y_o}}.\overline{P_y}

### 2020-04-02 06b-bisimulation.tex
Update 06b-bisimulation.tex

@@
 We write $\upsilon \entails F$ (read as the transition label $v$ satisfies the assertion $F$) 
-if $\mathit{flat}(\upsilon)\in \llbracket F\rrbracket$, otherwise we write $\upsilon \not\entails F$ and say that $F$ does not hold at $v$. 
+if $\mathit{flat}(\upsilon)\in \llbracket F\rrbracket$, otherwise we write $\upsilon \not\entails F$ (or also $\upsilon \entails \neg F$) and say that $F$ does not hold at $v$. 
 \end{definition}
@@
 The notion of bio-simulation builds on the above language of assertions to parameterize the induced equivalence on the property of interest. 
-Please recall that we have defined the behaviour of the contest in a non determinist way, thus 
+Please recall that we have defined the behaviour of the contest in a non deterministic way, thus 
 at each step, different possible sets of entities can be provided to the system and different sets of reaction can be enabled/disabled. 
@@
 \begin{definition}[Bio-similarity  $ \sim_F $]
-Given an assertion $F$, a \emph{bio-simulation} $\mathbf{R}_F$ that respects $F$ is a binary relation over {\tt link}-calculus processes such that, if $P \mathrel{\mathbf{R}_F} Q$ then:
+Given an assertion $F$, a \emph{bio-simulation} $\mathbf{R}_F$ that respects $F$ is a binary relation over c\CNA \  processes such that, whenever $P~\mathrel{\mathbf{R}_F}~Q$ then:
  \begin{itemize} 
@@
  for any $v,P'$ such that 
- $P \xrightarrow{v} P'$ then there exist $w,Q'$ such that $Q  \xrightarrow{w} Q'$ with $v\equiv_F w$ and $P' \mathrel{\mathbf{R}_F} Q'$.
+ $P \xRightarrow{v} P'$ then there exist $w,Q'$ such that $Q  \xRightarrow{w} Q'$ with $v\equiv_F w$ and $P' \mathrel{\mathbf{R}_F} Q'$.
 \item
  for any $w,Q'$ such that 
- $Q \xrightarrow{w} Q'$ then there exist $v,P'$ such that $P  \xrightarrow{v} P'$ with $v\equiv_F w$ and $P' \mathrel{\mathbf{R}_F} Q'$.
+ $Q \xRightarrow{w} Q'$ then there exist $v,P'$ such that $P  \xRightarrow{v} P'$ with $v\equiv_F w$ and $P' \mathrel{\mathbf{R}_F} Q'$.
 \end{itemize}
@@
 \begin{remark}
-An alternative way to look at a bio-simulation that respects $F$ is to define it as an ordinary bisimulation over the transition system labelled over $\{F,\neg F\}$ obtained by transforming each transition $P \xrightarrow{v} P'$ such that $v\entails F$ into $P \xrightarrow{F} P'$ and  each transition $P \xrightarrow{v} P'$ such that $v\not\entails F$ into $P \xrightarrow{\neg F} P'$.
+Please remember that the notation $P \xRightarrow{v} P'$ refers to ordinary transitions $P \xrightarrow{v} P'$ where $v$ is a complete chain (solid and with $\silent$ actions at the extremes).
+It should not be confused with the notation for weak transitions commonly found in the literature on process algebras.
 \end{remark}
+
+\begin{remark}
+An alternative way to look at a bio-simulation that respects $F$ is to define it as an ordinary bisimulation over the transition system labelled over $\{F,\neg F\}$ obtained by transforming each transition $P \xRightarrow{v} P'$ such that $v\entails F$ into $P \xrightarrow{F} P'$ and  each transition $P \xRightarrow{v} P'$ such that $v\not\entails F$ into $P \xrightarrow{\neg F} P'$.
+\end{remark}
+
+It can be easily shown that the identity relation is a bio-simulation and that bio-simulations are closed under (relational) inverse, composition and union and that, as a consequence, bio-similarity is an equivalence relation.
 
@@
 \begin{array}{lcl}
- \llbracket  \langle \chi \rangle G \rrbracket &\defeq& \{P \in {\cal P}: \exists v,P'.\ P\xrightarrow{v}P' \mbox{ with } v \entails \chi \mbox{ and } P' \in  \llbracket  G \rrbracket\} \\
- \llbracket [\chi]G \rrbracket &\defeq& \{P \in {\cal P}: \forall v,P'.\ P\xrightarrow{v}P' \mbox{ implies } v \entails \chi \mbox{ and } P' \in  \llbracket  G \rrbracket\}
+ \llbracket  \langle \chi \rangle G \rrbracket &\defeq& \{P \in {\cal P}: \exists v,P'.\ P\xRightarrow{v}P' \mbox{ with } v \entails \chi \mbox{ and } P' \in  \llbracket  G \rrbracket\} \\
+ \llbracket [\chi]G \rrbracket &\defeq& \{P \in {\cal P}: \forall v,P'.\ P\xRightarrow{v}P' \mbox{ implies } v \entails \chi \mbox{ and } P' \in  \llbracket  G \rrbracket\}
 \end{array}
@@
 \end{definition}
+
+Negation is not included in the syntax, but the converse $\overline{G}$ of a bioHML formula $G$ can be easily defined inductively in the same way as for HML logic.
+
+\begin{definition}[converse]
+Given a bioHML formula $G$ we define its converse $\overline{G}$ as follows:
+\[
+\begin{array}{rcl@{\hspace{1cm}}rcl@{\hspace{1cm}}rcl}
+\overline{\tt t} & \defeq & {\tt f} &
+\overline{G \wedge H} & \defeq & \overline{G} \vee \overline{H} &
+\overline{\langle \chi \rangle G} & \defeq & [\chi]\overline{G}
+\\
+\overline{\tt f} & \defeq & {\tt t} &
+\overline{G \vee H} & \defeq & \overline{G} \wedge \overline{H} &
+\overline{[\chi]G} & \defeq & \langle \chi \rangle\overline{G}
+\end{array}
+\]
+\end{definition}
+
+We observe that, as expected, for any bioHML formula $G$ and process $P$ we have $\overline{\overline{G}} = G$ and  $P \entails \overline{G}$ iff  $P\not\entails G$.
 
@@
 \end{theorem}
+\begin{proof}
+The proof is just an adaptation of the classical result. The two implications are proved separately.
+\begin{description}
+\item[$\sim_F\ \subseteq\ \equiv_{\mathcal{L}_F}$:]
+Given any two processes $P\sim_F Q$ we need to prove that for any bioHML formula $G$ we have $P\entails G$ iff $Q\entails G$. 
+Without loss of generality, we prove that $P\entails G$ implies $Q\entails G$.
+The proof is by structural induction on $G$.
+\begin{itemize}
+\item 
+if $G={\tt t}$, then $Q\entails G$.
+
+\item
+if $G={\tt f}$, then the assumption $P\entails G$ is false and the implication holds.
+
+\item
+if $G=G_1 \wedge G_2$ we take as inductive hypotheses that
+$$
+\begin{array}{l}
+\forall R,S.\ R \sim_F S \wedge R\entails G_1 \Rightarrow S \entails G_1\\
+\forall R,S.\ R \sim_F S \wedge R\entails G_2 \Rightarrow S \entails G_2
+\end{array}
+$$
+We need to prove that $Q\entails G$.
+Since $P\entails G=G_1 \wedge G_2$ we have $P\entails G_1$ and $P\entails G_2$.
+Since $P\sim_F Q$, by inductive hypotheses we get $Q\entails G_1$ and $Q\entails G_2$.
+Hence $Q\entails G_1 \wedge G_2=G$.
+
+\item
+if $G=G_1 \vee G_2$  we take as inductive hypotheses that
+$$
+\begin{array}{l}
+\forall R,S.\ R \sim_F S \wedge R\entails G_1 \Rightarrow S \entails G_1\\
+\forall R,S.\ R \sim_F S \wedge R\entails G_2 \Rightarrow S \entails G_2
+\end{array}
+$$
+We need to prove that $Q\entails G$.
+Since $P\entails G=G_1 \vee G_2$ we have $P\entails G_1$ or $P\entails G_2$.
+If $P\entails G_1$, since $P\sim_F Q$, by inductive hypotheses we get $Q\entails G_1$ and thus $Q\entails G_1 \vee G_2=G$.
+If $P\entails G_2$, since $P\sim_F Q$, by inductive hypotheses we get $Q\entails G_2$ and thus $Q\entails G_1 \vee G_2=G$.
+
+\item
+if $G=\langle\chi\rangle H$ we take as inductive hypothesis that
+$$
+\forall R,S.\ R \sim_F S \wedge R\entails H \Rightarrow S \entails H
+$$
+We need to prove that $Q\entails G$.
+Since $P\entails \langle\chi\rangle H$ it means that there exists $\upsilon, P'$ such that $P\xRightarrow{\upsilon} P'$ with $\upsilon\entails \chi$ and $P'\entails H$.
+Since $P\sim_F Q$, there exists $w,Q'$ such that $Q\xRightarrow{w} Q'$ with $w\entails \chi$ and $P'\sim_F Q'$.
+Then, by inductive hypothesis, $Q'\entails H$ and thus $Q\entails \langle\chi\rangle H=G$.
+
+\item
+if $G=[\chi] H$ we take as inductive hypothesis that
+$$
+\forall R,S.\ R \sim_F S \wedge R\entails H \Rightarrow S \entails H
+$$
+We need to prove that $Q\entails G$.
+If there is no $\upsilon\entails \chi$ such that $Q\xRightarrow{\upsilon}Q'$ for some $Q'$ then 
+ $Q\entails [\chi] H=G$ trivially.
+For any $\upsilon,Q'$ such that $Q\xRightarrow{\upsilon}Q'$ with $\upsilon\entails \chi$, then as $P\sim_F Q$ there must exist $w,P'$ such that $P\xRightarrow{w}P'$ with $w\entails \chi$ and $P'\sim_F Q'$.
+Since $P\entails G=[\chi] H$ then it must be $P'\entails H$.
+Since $P'\sim_F Q'$, by inductive hypothesis $Q'\entails H$.
+Hence $Q\entails [\chi] H = G$.
+\end{itemize}
+
+\item[$\equiv_{\mathcal{L}_F}\ \subseteq\ \sim_F$:]
+We prove that $\equiv_{\mathcal{L}_F}$ is a bio-simulation and thus included in $\sim_F$.
+Take two generic processes $P\equiv_{\mathcal{L}_F} Q$ and suppose $P\xRightarrow{\upsilon}P'$ for some $\upsilon,P'$. 
+\begin{itemize}
+\item
+If $\upsilon \entails F$ we want to prove that there exists some $w,Q'$ such that $Q\xRightarrow{w}Q'$, with $w\entails F$ and $P'\equiv_{\mathcal{L}_F} Q'$.
+
+Towards a contradiction, assume that we cannot find such $w,Q'$.
+
+If there is no transition $Q\xRightarrow{w}Q'$ such that $w\entails F$, then the bioHML formula $G \defeq \langle F\rangle{\tt t}$ is such that $P\entails G$ and $Q\not\entails G$, contradicting the assumption $P\equiv_{\mathcal{L}_F} Q$.
+
+Otherwise, let $\mathcal{Q}\defeq \{Q'~\mid~\exists w.\ Q\xRightarrow{w}Q' \ \wedge\ w\entails F\}$ be the set of processes reachable from $Q$ via a transition with a (complete) label that satisfies $F$.
+Since our processes are with guarded recursion, the set $\mathcal{Q}$ is finite.
+Let $\mathcal{Q} = \{Q'_1,...,Q'_n\}$.
+By hypothesis all processes in $\mathcal{Q}$ must not be bio-logically equivalent to $P'$, hence for any $i\in[1,n]$ there exists a bioHML formula $G_i$ such that $P'\entails G_i$ and $Q'_i\not\entails G_i$ (if it was the opposite, $P'\not\entails G_i$ and $Q'_i\entails G_i$, we can use the converse formulas).
+But then the formula $G\defeq \langle F\rangle(G_1\wedge\dots\wedge G_n)$ is such that $P\entails G$ and $Q\not\entails G$, contradicting the assumption $P\equiv_{\mathcal{L}_F} Q$.
+
+\item
+If $\upsilon \not\entails F$ then the proof is analogous to the previous case and thus omitted.
+\end{itemize}
+\end{description}
+\end{proof}

### 2020-04-03 09-appendix.tex
tolto lac operon da appendix e aggiunto file con codifica lac operon
tolto lac operon in appendice ed aggiunto file con la codifica del lac operon, per non perderla

@@
  \end{proof}
-
- \newpage
-
- \section{The \emph{lac} operon encoding}
- \begin{table}[h]
-{\small
-\[
-\boxed{
-\begin{array}{lcl}
-P_{cAMP\textrm{-}CAP} & \defeq & %P_{cAMP\textrm{-}CAP}(cAMP\textrm{-}CAP)=\\
-\sum_{h =0}^1 ( \startchain{cAMP\textrm{-}CAP_i}\chainedlink{cAMP\textrm{-}CAP_o}{\noact}\chainend{\noact})^h\,  \startchain{\widetilde{cAMP\textrm{-}CAP}_i}\chainend{\widetilde{cAMP\textrm{-}CAP}_o}.P_{cAMP\textrm{-}CAP}\\
-&&+\\
-&&\link{cAMP\textrm{-}CAP_i}{cAMP\textrm{-}CAP_o} .\overline{P_{cAMP\textrm{-}CAP}}\\
-\\
-\overline{P_{cAMP\textrm{-}CAP}} & \defeq &
-\sum_{h =0}^1  ( \startchain{\overline{cAMP\textrm{-}CAP}_i}\chainedlink{\overline{cAMP\textrm{-}CAP}_o}{\noact}\chainend{\noact})^h\ \link{\widetilde{cAMP\textrm{-}CAP}_i}{\widetilde{cAMP\textrm{-}CAP}_o} .P_{cAMP\textrm{-}CAP}\\
-&&+\\
-&&\link{\overline{cAMP\textrm{-}CAP}_i}{\overline{cAMP\textrm{-}CAP}_o} .\overline{P_{cAMP\textrm{-}CAP}}\\
-\end{array}}
-\]}
-\caption{The process for $cAMP\textrm{-}CAP$.}\label{tab:cAMP-CAP}
-%\end{table}
-%
-%\begin{table}[h]
-\[
-\boxed{
-\begin{array}{lcl}
-P_{glucose} &\defeq &
-\sum_{h=0}^1 (\startchain{glucose_i}\chainedlink{glucose_o}{\noact}\chainend{\noact})^h\ \link{\widehat{glucose_i}}{\widehat{glucose_o}}.P_{glucose}
-\\
-&& + \\
-&& \sum_{h=0}^1 (\startchain{glucose_i}\chainedlink{glucose_o}{\noact}\chainend{\noact})^h\ \link{\underline{glucose_i}}{\underline{glucose_o}}.\overline{P_{glucose}}\\
-\\
-\overline{P_{glucose}} &\defeq &\sum_{h=0}^1 (\startchain{\overline{glucose}_i}\chainedlink{\overline{glucose}_o}{\noact}\chainend{\noact})^h\ \link{\widehat{glucose_i}}{\widehat{glucose_o}}.P_{glucose}\\
-&&+\\
-&& \sum_{h=0}^1 (\startchain{\overline{glucose}_i}\chainedlink{\overline{glucose}_o}{\noact}\chainend{\noact})^h\ \link{\underline{glucose_i}}{\underline{glucose_o}}.\overline{P_{glucose}}\\
-\end{array}}
-\]
-\caption{The process for $glucose$.}\label{tab:glucose}
-%\end{table}
-%
-%\begin{table}[h]
-\[
-\boxed{
-\begin{array}{lcl@{\qquad}lcl}
-P_A&\defeq &\startchain{\widetilde{A_i}}\chainedlink{\widetilde{A_o}}{\noact}\chainedlink{\noact}{\underline{A_i}}\chainend{\underline{A_o}}.P_A\ + \
-\link{\underline{A_i}}{\underline{A_o}}.\overline{P_A}
-&
-P_y&\defeq &\startchain{\widetilde{y_i}}\chainedlink{\widetilde{y_o}}{\noact}\chainedlink{\noact}{\underline{y_i}}\chainend{\underline{y_o}}.P_y\ + \
-\link{\underline{y_i}}{\underline{y_o}}.\overline{P_y}
-\\
-\overline{P_A} &\defeq &\startchain{\widetilde{A_i}}\chainedlink{\widetilde{A_o}}{\noact}\chainedlink{\noact}{\underline{A_i}}\chainend{\underline{A_o}}.\overline{P_A}\ + \
-\link{\underline{A_i}}{\underline{A_o}}.\overline{P_A}
-&
-\overline{P_y} &\defeq &\startchain{\widetilde{y_i}}\chainedlink{\widetilde{y_o}}{\noact}\chainedlink{\noact}{\underline{y_i}}\chainend{\underline{y_o}}.\overline{P_y}\ + \
-\link{\underline{y_i}}{\underline{y_o}}.\overline{P_y}
-\end{array}}
-\]
-\caption{The processes for $y$ and $A$.}\label{tab:yA}
-
-\end{table}

### 2020-04-03 appendixLacOperonCodification.tex
tolto lac operon da appendix e aggiunto file con codifica lac operon
tolto lac operon in appendice ed aggiunto file con la codifica del lac operon, per non perderla

\newpage

 \section{The \emph{lac} operon encoding}
 \begin{table}[h]
{\small
\[
\boxed{
\begin{array}{lcl}
P_{cAMP\textrm{-}CAP} & \defeq & %P_{cAMP\textrm{-}CAP}(cAMP\textrm{-}CAP)=\\
\sum_{h =0}^1 ( \startchain{cAMP\textrm{-}CAP_i}\chainedlink{cAMP\textrm{-}CAP_o}{\noact}\chainend{\noact})^h\,  \startchain{\widetilde{cAMP\textrm{-}CAP}_i}\chainend{\widetilde{cAMP\textrm{-}CAP}_o}.P_{cAMP\textrm{-}CAP}\\
&&+\\
&&\link{cAMP\textrm{-}CAP_i}{cAMP\textrm{-}CAP_o} .\overline{P_{cAMP\textrm{-}CAP}}\\
\\
\overline{P_{cAMP\textrm{-}CAP}} & \defeq &
\sum_{h =0}^1  ( \startchain{\overline{cAMP\textrm{-}CAP}_i}\chainedlink{\overline{cAMP\textrm{-}CAP}_o}{\noact}\chainend{\noact})^h\ \link{\widetilde{cAMP\textrm{-}CAP}_i}{\widetilde{cAMP\textrm{-}CAP}_o} .P_{cAMP\textrm{-}CAP}\\
&&+\\
&&\link{\overline{cAMP\textrm{-}CAP}_i}{\overline{cAMP\textrm{-}CAP}_o} .\overline{P_{cAMP\textrm{-}CAP}}\\
\end{array}}
\]}
\caption{The process for $cAMP\textrm{-}CAP$.}\label{tab:cAMP-CAP}
%\end{table}
%
%\begin{table}[h]
\[
\boxed{
\begin{array}{lcl}
P_{glucose} &\defeq &
\sum_{h=0}^1 (\startchain{glucose_i}\chainedlink{glucose_o}{\noact}\chainend{\noact})^h\ \link{\widehat{glucose_i}}{\widehat{glucose_o}}.P_{glucose}
\\
&& + \\
&& \sum_{h=0}^1 (\startchain{glucose_i}\chainedlink{glucose_o}{\noact}\chainend{\noact})^h\ \link{\underline{glucose_i}}{\underline{glucose_o}}.\overline{P_{glucose}}\\
\\
\overline{P_{glucose}} &\defeq &\sum_{h=0}^1 (\startchain{\overline{glucose}_i}\chainedlink{\overline{glucose}_o}{\noact}\chainend{\noact})^h\ \link{\widehat{glucose_i}}{\widehat{glucose_o}}.P_{glucose}\\
&&+\\
&& \sum_{h=0}^1 (\startchain{\overline{glucose}_i}\chainedlink{\overline{glucose}_o}{\noact}\chainend{\noact})^h\ \link{\underline{glucose_i}}{\underline{glucose_o}}.\overline{P_{glucose}}\\
\end{array}}
\]
\caption{The process for $glucose$.}\label{tab:glucose}
%\end{table}
%
%\begin{table}[h]
\[
\boxed{
\begin{array}{lcl@{\qquad}lcl}
P_A&\defeq &\startchain{\widetilde{A_i}}\chainedlink{\widetilde{A_o}}{\noact}\chainedlink{\noact}{\underline{A_i}}\chainend{\underline{A_o}}.P_A\ + \
\link{\underline{A_i}}{\underline{A_o}}.\overline{P_A}
&
P_y&\defeq &\startchain{\widetilde{y_i}}\chainedlink{\widetilde{y_o}}{\noact}\chainedlink{\noact}{\underline{y_i}}\chainend{\underline{y_o}}.P_y\ + \
\link{\underline{y_i}}{\underline{y_o}}.\overline{P_y}
\\
\overline{P_A} &\defeq &\startchain{\widetilde{A_i}}\chainedlink{\widetilde{A_o}}{\noact}\chainedlink{\noact}{\underline{A_i}}\chainend{\underline{A_o}}.\overline{P_A}\ + \
\link{\underline{A_i}}{\underline{A_o}}.\overline{P_A}
&
\overline{P_y} &\defeq &\startchain{\widetilde{y_i}}\chainedlink{\widetilde{y_o}}{\noact}\chainedlink{\noact}{\underline{y_i}}\chainend{\underline{y_o}}.\overline{P_y}\ + \
\link{\underline{y_i}}{\underline{y_o}}.\overline{P_y}
\end{array}}
\]
\caption{The processes for $y$ and $A$.}\label{tab:yA}

\end{table}

### 2020-04-03 00-abstract.tex
small changes in abstract and intro

@@
 pose problems of implementation, as it is a new computation model. 
-In this paper we consider the {\tt link}-calculus, which allows to model 
+In this paper we consider a variant of the {\tt link}-calculus, which allows to model 
 multiparty interaction in concurrent systems, and show that it allows to

### 2020-04-03 01-introduction.tex
just typos
changed biblio and other files
some small changes in several files
small changes in abstract and intro

@@
 We also show that our embedding preserves the main
-features of RSs, and prove its correctness and completeness.
+features of RSs, and prove its correctness and completeness
+from the operational semantics viewpoint.
 
@@
 properties of reaction systems. The classical notion of bisimulation for 
-process algebras allows to consider equivalent two processes
+process algebras allows to consider two processes as equivalent 
 where one process can simulate all the actions executed 
-by the other one and viceversa. In this paper we define a new
+by the other one and vice versa. In this paper we define a new
 notion of bisimilarity which takes into account the characteristics
@@
 In fact, bio-similarity is parametric on a given assertion of interest.
-The we prove that the well-known logical characterization of bisimilarity in terms of Hennessy-Milner Logic~\cite{HM80} (HML)
+Then we prove that the well-known logical characterisation of bisimilarity in terms of Hennessy-Milner Logic~\cite{HM80} (HML)
 can be extended to the case of bio-similarity by tailoring HML formulas to
@@
 can be
-specified in a recursive way as an ordinary process;
+specified in a recursive way, as ordinary processes;
 \item along the same lines, non-deterministic contexts can be seamlessly expressed in our setting; 
 \item when deterministic contexts are considered, the c\CNA \
-computation is deterministic.
+computation is deterministic and mimics the evolution of the corresponding RS;
 \item we define an assertion language to specify local properties of 
 reaction systems and exploit it to define a novel notion of behavioural equivalence for reaction systems, called bio-similarity;
-\item we show that our assertions extend can be used to extend the Hennessy-Milner logic to our framework, in such a way that logical equivalence of processes coincide with bio-similarity;
+\item we show that our assertions can be used to extend the Hennessy-Milner logic to our framework, in such a way that logical equivalence of processes coincide with bio-similarity;
 \item we sketch how one can express the behaviour of entity mutation,
@@
 $s'$ can  take part to only a subset of rules requiring entity $s$;
-\item we show that with a little coding effort, two RSs can
-communicate; i.e.  a subset of those entities that the context  can
-provide, are then provided by a second RS.
+\item we show that with a little coding effort, our encoding allows two RSs to
+communicate; i.e. we can model scenarios where a subset of those entities that the context  can
+provide, are instead provided by a second RS.
 \end{itemize}
@@
 an implementation of the {\tt link}-calculus. 
+The examples that we propose also show that some optimisations are possible
+to reduce the coding when suitable assumptions are made about the provision of entities.
 
@@
 
-\paragraph{Related work.}
+\paragraph{Related work}
 
@@
 biological processes, see~\cite{BBDFH18} for a recent survey.
-We are not aware of any inference rules for the RS; it seems not the case
+We are not aware of any SOS operational semantics for the RS; it seems not the case
 that a deterministic behaviour, as the one of the RS, once the initial state is 
@@
 framework the Hennessy-Milner logic;
-\item proofs of main results are included here.
+\item we include here all proofs of main results.
 \end{itemize}
 
-\paragraph{Structure of the paper.}
+\paragraph{Structure of the paper}
 
@@
 it.
+Section~\ref{ex:examples} presents a couple of examples with computer science and biological applications.
 %Section \ref{ex:lactose} shows a more complex example taken from

### 2020-04-03 06-examples.tex
vari cambi file sforamenti linea
cambi in file con sforamento linea

@@
 \begin{array}{lcl @{\hspace{0.5cm}}lcl @{\hspace{0.5cm}}lcl @{\hspace{0.5cm}}lcl}
-P_T & \defeq & P(T,\widetilde{T})& \overline{P}_T & \defeq & P(\overline{T},\widetilde{T}) & P_{T'}& \defeq & P(T',\widetilde{T'}) & \overline{P}_{T',\widetilde{T'}}& \defeq & P(\overline{T'},\widetilde{T'})\\\
+P_T & \defeq & P(T,\widetilde{T})& \overline{P}_T & \defeq & P(\overline{T},\widetilde{T}) & P_{T'}& \defeq & P(T',\widetilde{T'}) & \\
+\overline{P}_{T',\widetilde{T'}}& \defeq & P(\overline{T'},\widetilde{T'})\\\
 \end{array}

### 2020-04-03 07-discussion.tex
vari cambi file sforamenti linea
cambi in file con sforamento linea

% !TEX root = ./main_RS2L_LNCS.tex
%\section{Discussion}

\section{Enhanced  Reaction Systems}
\label{sec:discussion}

Our encoding increases the expressivity of  RS  concerning:
the behaviour of the context, the possibility of alternative behaviour of
mutated entities, the communication between two different reaction systems.
It is important to note that our encoding guarantees that from each state, 
in the  c\CNA \ transition system, only one transition comes out, 
as the dynamics is totally deterministic.

\subsection{Recursive contexts}
In RS, the behaviour of the context is finite. For the first $n$ steps, it is specified 
 which are the entities that are provided from the context. 
Using c\CNA\ we can describe in a natural way the behaviour of the context in a recursive way. 
Then, the context behaviour would not necessarily end after $n$ steps,
and  could be infinite.
For example, in an extended interactive process, we may want that the entity $s$ is 
intermittently provided by the context every two steps: 
\[
\begin{array}{lcll}
Cxt_s & \defeq &  \startchain{cxt_j}\chainedlink{\hat{s}_i}{\noact}\chainedlink{\noact}{\hat{s}_o}\chainend{cxt_{j+1}}.Cxt_{s'}, &\mbox{ context provides $s$;} \\
Cxt_{s'} & \defeq & \startchain{cxt_j}\chainedlink{\underline{s}_i}{\noact}\chainedlink{\noact}{\underline{s}_o}\chainend{cxt_{j+1}}.Cxt_{s''},& \mbox{ context doesn't provide $s$;} \\
Cxt_{s''} & \defeq & \startchain{cxt_j}\chainedlink{\underline{s}_i}{\noact}\chainedlink{\noact}{\underline{s}_o}\chainend{cxt_{j+1}}.Cxt_{s}, & \mbox{ context doesn't provide $s$.} \\
\end{array}
\]

\subsection{Mutating entities} 
In RS, when an entity is present, it can potentially be involved in each reactions where it is required.
With a few more lines of code, in c\CNA~it is possible to describe the behaviour of a mutation of 
an entity, in a way that the mutated version of the entity can take 
part to only a subset of the rules requiring  the  \emph{normal version} of the entity.
For example, let us assume that entity $s1$ is consumed by reactions $a1$ and $a2$.
Reaction $a1$ produces also $s1$ if $s2$ is present, otherwise $a1$ produces a mutated version of $s1$, 
say $s1'$.
When $s1'$ is produced, reaction $a2$ behaves in the same way as if $s1$ would be absent, whereas $a2$
recognises the presence of $s1'$ and behaves in the same way as if $s1$ would be present.
Technically, in both cases it is enough to add one more non 
deterministic choice in the code of $P_{a1}$ and $P_{a2}$.

\subsection{Communicating reaction systems} 
We sketch how it is possible to program two RSs encodings, in a way that 
the entities that usually come from context of one RS will be provided instead from the other RS.

%Here we present a a simple example showing how it is possible to implement in c\CNA\ a 
%two RSs model where a subset of the entities of one RS is provided by the second one and not 
%provided by the context.

\begin{example}
\label{ex:twoRS}
Let  $rs1$ and $rs2$ be two RSs, defined by the rules in Table~\ref{tab:rules}.
%such that $rs2$ provides the entity $s1$ to $rs1$.

\begin{table}[t]
\centering
\begin{tabular}{|c|c|}
\hline
$rs1$  & $rs2$ \\
\hline
%$r1_1=(s1,\ , s2)$ & $r2_1=(s3,\  ,s1)$\\
$a_1=(s,\ , x)$ & $a_2=(y,\  ,s)$\\
%$r1_2=(\ , s2,s1)$ & $r2_2=(s1,\ ,s3)$\\
\hline 
\end{tabular}
\caption{The two reaction systems $rs1$ and $rs2$.}
\label{tab:rules}
\end{table}
 \noindent
Now, we set our example such that the two contexts, for  $rs1$ and $rs2$, do not provide any entities.
We also assume that entity $s$ in $rs1$ is provided by $rs2$, as $rs2$ produces a quantity of $s$ that is enough for $rs1$ and $rs2$.
For technical reasons, we can not use the same name for  $s$ in both the two RSs, then we use the name $ss$ in $rs2$.
%
We need to modify our translation technique to suite this new setting. 
As we do not model contexts, we introduce  \emph{dummy} channel names $dx$ and $dss$ to model the absence of entities. Also, thanks to the simplicity of the example, we can leave out the use of the   $p_i$ channels. This streamlining does not affect the programming technique we propose to make two RSs communicate.
%Thus, there will be no $Cxt$ processes and we manage the code in a way that $rs1$ will receive the entity $s1$ from $rs2$. 
%Please also note that, for our simple example, we do not have anymore to specify the all steps
%of a interactive system, as the communication mechanism from $rs2$ to $rs1$ will generate an infinite
%behaviour.
%\noindent
%First, we translate the rules in $rs1$:
First, we translate the reaction in $rs1$:
\[
\begin{array}{lcl}
%\llbracket r1_1\rrbracket \defeq P_{r1_1}&\defeq &\startchain{\silent}\chainedlink{s1}{\noact}\chainedlink{\noact}{s1}\chainedlink{\tilde{s2}}{\noact}\chainedlink{\noact}{\tilde{s2}}\chainend{r1_2}.P_{r1_1} +
%\startchain{\silent}\chainedlink{\overline{s1}}{\noact}\chainedlink{\noact}{\overline{s1}}\chainend{r1_2}.P_{r1_1}\\
\llbracket a_1\rrbracket \defeq P_{a_1}&\defeq &
\startchain{\silent}\chainedlink{s_i}{{\color{blue}\noact}}\chainedlink{{\color{blue}\noact}}{s_o}\chainedlink{\tilde{x}_i}{{\color{gray}\noact}}\chainedlink{{\color{gray}\noact}}{\tilde{x}_o}\chainend{a_2}.P_{a_1} +
\startchain{\silent}\chainedlink{\overline{s}_i}{\noact}\chainedlink{\noact}{\overline{s}_o}    \chainedlink{dx_i}{\noact}\chainedlink{\noact}{dx_o}\chainend{a_2}.P_{a_1}\\
%\end{array}
%\]
%\\
%
%\begin{array}{lcl}
%\llbracket r1_2\rrbracket \defeq P_{r1_2}&\defeq &\startchain{r1_2}\chainedlink{\overline{s2}}{\noact}\chainedlink{\noact}{\overline{s2}}\chainedlink{\tilde{s1}}{\noact}\chainedlink{\noact}{\tilde{s1}}\chainend{r2_1}.P_{r2_1} +
%\startchain{r1_2}\chainedlink{s2}{\noact}\chainedlink{\noact}{s2}\chainend{r2_1}.P_{r2_1}\\
\end{array}
\]

\noindent
Please note, that prefixes of process $P_{a_1}$ end with the channel name $a_2$, as the link chain is now connected with the  reaction of  $rs2$.
The translation for the entities follows.
\[
\begin{array}{lclcl}
\llbracket s \rrbracket &\defeq &P_{s} &\defeq & \startchain{{\color{blue}s_i}}\chainedlink{{\color{blue}s_o}}{\noact}\chainedlink{\noact}{{\color{blue}\hat{s}_i}}\chainend{{\color{blue}\hat{s}_o}}.P_{s} + \startchain{s_i}\chainend{s_o}.\overline{P_{s}}\\
&& \overline{P_{s}}&\defeq &\startchain{\overline{s}_i}\chainedlink{\overline{s}_o}{\noact}\chainedlink{\noact}{\hat{s}_i}\chainend{\hat{s}_o}.P_{s}+\startchain{\overline{s}_i}\chainend{\overline{s}_i}.\overline{P_{s}}\\
%\begin{array}{lclclll}
\llbracket x \rrbracket &\defeq & P_{x}& \defeq &\startchain{\tilde{x}_i}\chainend{\tilde{x}_o}.P_{x} +  \startchain{dx_i}\chainend{dx_o}.\overline{P_{x}}\\
%\\
%&&\overline{P_{s2}}&\defeq &\startchain{\overline{s2}}\chainedlink{\overline{s2}}{\noact}\chainedlink{\noact}{\tilde{s2}}\chainend{\tilde{s2}}.P_{s2} +
%\startchain{\overline{s2}}\chainend{\overline{s2}}.\overline{P_{s2}}\\
%\\
%\\
&& \overline{P_{x}}&\defeq &\startchain{{\color{gray}\tilde{x}_i}}\chainend{{\color{gray}\tilde{x}_o}}.P_{x} +  \startchain{dx_i}\chainend{dx_o}.\overline{P_{x}}\\
%\end{array}
%\]
%
%\[

\end{array}
\]

%\[
%\begin{array}{lclclll}
%\llbracket s2 \rrbracket &\defeq& P_{s2}&\defeq &\startchain{s2}\chainedlink{s2}{\noact}\chainedlink{\noact}{\tilde{s2}}\chainend{\tilde{s2}}.P_{s2} +  \startchain{s2}\chainend{s2}.\overline{P_{s2}}\\
%%\\
%&&\overline{P_{s2}}&\defeq &\startchain{\overline{s2}}\chainedlink{\overline{s2}}{\noact}\chainedlink{\noact}{\tilde{s2}}\chainend{\tilde{s2}}.P_{s2} +
%\startchain{\overline{s2}}\chainend{\overline{s2}}.\overline{P_{s2}}\\
%
%\end{array}
%\]

\noindent
The translation for the $rs2$ follows.

\[
\begin{array}{lcl}
\llbracket a_2\rrbracket \defeq P_{a_2}&\defeq &\startchain{a_2}\chainedlink{y_i}{{\color{gray}\noact}}\chainedlink{{\color{gray}\noact}}{y_o}\chainedlink{\tilde{ss}_i}{{\color{red}\noact}}\chainedlink{{\color{red}\noact}}{\tilde{ss}_o}\chainend{\silent}.P_{a_2}+
\startchain{a_2}\chainedlink{\overline{y}_i}{\noact}\chainedlink{\noact}{\overline{y}_o}    
\chainedlink{dss_i}{\noact}\chainedlink{\noact}{dss_o}\chainend{\silent}.P_{a_2}\\
%\end{array}
%\]
%
%\[
%\begin{array}{lcl}
%\llbracket r2_2\rrbracket \defeq P_{r2_2}&\defeq &\startchain{r2_2}\chainedlink{ss1}{\noact}\chainedlink{\noact}{ss1}\chainedlink{\tilde{s3}}{\noact}\chainedlink{\noact}{\tilde{s3}}\chainend{\silent}.P_{r2_2}
%+
%\startchain{r2_2}\chainedlink{\overline{ss1}}{\noact}\chainedlink{\noact}{\overline{ss1}}\chainend{\silent}.P_{r2_2}\\
\end{array}
\]
\noindent
In the translation of the entities in  $rs2$, we introduce the mechanism that allows the entity
$s$ ($ss$ in $rs2$) to be provided in $rs1$. Every time $ss$ is produced in $rs2$,  a virtual link is
created to synchronise with $rs1$   on link $\link{\hat{s}_i}{\hat{s}_o}$:
\[
\begin{array}{lclcl}
\llbracket ss \rrbracket& \defeq& P_{ss}&\defeq &\startchain{\tilde{ss}_i}\chainedlink{\hat{s}_i}{\noact}\chainedlink{\noact}{\hat{s}_o}\chainend{\tilde{ss}_o}.P_{ss}+
\startchain{dss_i}   \chainend{dss_o}.\overline{P_{ss}} \\

%\end{array}
%\]
%\[
%\begin{array}{lcllcl}
&&\overline{P_{ss}}&\defeq &\startchain{{\color{red}\tilde{ss}_i}}\chainedlink{{\color{red}\hat{s}_i}}{{\color{blue}\noact}}\chainedlink{{\color{blue}\noact}}{{\color{red}\hat{s}_o}}\chainend{{\color{red}\tilde{ss}_o}}.P_{ss}
+
\startchain{dss_i}\chainend{dss_o}.\overline{P_{ss}}\\
\llbracket y\rrbracket&\defeq & P_{y}&\defeq &\startchain{{\color{gray}y_i}}\chainend{{\color{gray}y_o}}.\overline{P_{y}} \\
&&\overline{P_{y}}&\defeq &
%\startchain{\overline{y}_i}\chainedlink{\overline{y}_o}{\noact}\chainedlink{\noact}{\tilde{y}_i}\chainend{\tilde{y}_o}.P_{y}
%+ 
\startchain{\overline{y}_i}\chainend{\overline{y}_o}.\overline{P_{y}}
\end{array}
\]
%\[
%\begin{array}{lcllcl}

%\end{array}
%\]
\end{example}

\noindent
We now assume that the initial system is $S \defeq \restrict{names}(P_{a_1}|P_{a_2}|P_s|P_y|\overline{P_x}|\overline{P_{ss}})$, i.e.
only entities $s$ and $y$ are present. Now, the only possible transition has  the following label (that we report without restriction):
\[
\startchain{\silent}\chainedlink{s_i}{{\color{blue}s_i}}\chainedlink{{\color{blue}s_o}}{s_o}\chainedlink{\tilde{x}_i}{{\color{gray}\tilde{x}_i}}\chainedlink{{\color{gray}\tilde{x}_o}}{\tilde{x}_o} \chainedlink{a_2}{a_2}\chainedlink{y_i}{ {\color{gray}y_i}}\chainedlink{{\color{gray}y_o}}{y_o}\chainedlink{\tilde{ss}_i}{{\color{red}\tilde{ss}_i}}
\chainedlink{{\color{red}\hat{s}_i}}{{\color{blue}\hat{s}_i}}
\chainedlink{{\color{blue}\hat{s}_o}}{{\color{red}\hat{s}_o}}
\chainedlink{{\color{red}\tilde{ss}_o}}{\tilde{ss}_o}\chainend{\silent},
\]
where the black links belong to the prefixes of $P_{a_1}$, and $P_{a_2}$, the blue links belong to $P_s$, the gray links belong to $P_y$, and $\overline{P_x}$ and the red links belong to $\overline{P_{ss}}$.
After the execution, the entity $s$ is still present in $rs1$ as it has been provided by $rs2$.\\

As we have briefly sketched, our model of two \emph{communicating reaction systems} 
can enable the study of the behaviour of one RS
in relation to another one.
In Example~\ref{ex:twoRS} describing  the behaviour of the \emph{lac} operon,
the two entities lactose and glucose are controlled 
non deterministically by the context.
In our framework,
instead, by exploiting the expressivity of c\CNA, 
the \emph{lac} operon system has been connected with the two systems 
producing the lactose and the glucose.
This way, the presence of these two entities in the \emph{lac} operon system 
can be regulated by realistic mechanisms.
%which can be modified to represent pathological states as well.

% Again, we stress the fact the verbosity of the {\tt link}-calculus code 
%should not be
% seen as a drawback, as it undergoes to automatic code translation.

### 2020-04-03 05-translation.tex
updated appendix
modificata freccia doppia pag 12
freccia doppia pag 12 diventa singola
cambio figura gialla

@@
 %\left \{
-^{cxt} \backslash  \blockchain{s_i}{s_o}{s \in C_n} \, \link{}{} \blockchain{\overline{e}_i}{\overline{e}_o}{e \not\in C_n} \,  \backslash_{p_1}. \mathit{Cxt}^{n+1}\\
+^{cxt} \backslash  \blockchain{s_i}{s_o}{s \in C_n} \, \link{}{} \blockchain{\overline{e}_i}{\overline{e}_o}{e \not\in C_n} \,  \backslash_{p_1}. \mathit{Cxt}^{n+1}
 %+ \\
@@
  $$
- \llparenthesis{\cal A},\gamma\rrparenthesis \xRightarrow{
+ \llparenthesis{\cal A},\gamma\rrparenthesis \xrightarrow{
  \startchain{\silent}
@@
   \draw (0.76,0.25)  node[minimum size=0.5cm,fill=yellow] {$\ $};
-  \draw (1.21,0.25)  node[minimum size=0.5cm,fill=yellow] {${\scriptscriptstyle s1_i}$};
+  \draw (1.21,0.25)  node[minimum size=0.5cm,fill=yellow] {${\scriptscriptstyle r_1 \ }$};
 
@@
 \draw (7.6,0.25)  node[minimum size=0.48cm,draw,fill=yellow, draw=yellow] {$ \ $};
-\draw (8.14,0.25)  node[minimum size=0.475cm,draw,fill=yellow,draw=yellow] {${\scriptscriptstyle \tilde{s1}_i}$};
+\draw (8.14,0.25)  node[minimum size=0.475cm,draw,fill=yellow,draw=yellow] {${\scriptscriptstyle \widetilde{s2}_i}$};
  \draw (8.14,0.75) node{$\noact$};%node[minimum size=0.5cm,draw,draw=white] {$\noact$};
@@
     % \draw (9,1) -- (10,1) -- (10,0.5) -- (10.5,0.5) --  (10.5,0) -- (9.5,0)  (9.5,0.5) -- (9,0.5) -- (9,1);
- \draw (9.24,0.75) node[minimum size=0.48cm,draw,fill=yellow,draw=yellow]{${\scriptscriptstyle \tilde{s1}_o}$};%node[minimum size=0.5cm,draw,draw=white] {$\noact$};
+ \draw (9.24,0.75) node[minimum size=0.48cm,draw,fill=yellow,draw=yellow]{${\scriptscriptstyle p_2 \; }$};%node[minimum size=0.5cm,draw,draw=white] {$\noact$};
 

### 2020-04-03 06-examples.tex
piccoli cambi
piccoli cambi

@@
 \noindent
-We consider a biological toy example where a gene $a$  codes for a protein $T$ when molecules 
+We consider a biological toy example in the style of gene's alternative splicing \ref{WBB13},
+where a gene $a$  codes for a protein $T$ when molecules 
 $G$ is present and $C$ is absent, and in the opposite situation $a$ codes for protein $T'$.

### 2020-04-03 06-examples.tex
Update 06-examples.tex
piccolo cambio
cambio
cambio
cambio

--- 2020-04-03 06-examples.tex
@@
 \noindent
-We consider a biological toy example where a gene $a$  codes for a protein $T$ when molecules 
+We consider a biological toy  example in the style of gene's alternative splicing \cite{WBB13}, 
+where a gene $a$  codes for a protein $T$ when molecules 
 $G$ is present and $C$ is absent, and in the opposite situation $a$ codes for protein $T'$.

### 2020-04-04 00-abstract.tex
typos e notazione
qualche typo e aggiunta def per complete chains e freccia doppia,
tolto tau finale da flattened sequences

--- 2020-03-30 00-abstract.tex
@@
 pose problems of implementation, as it is a new computation model. 
-In this paper we consider the {\tt link}-calculus, which allows to model 
-multiparty interaction in concurrent systems, and show that it allows to
-embed reaction systems in a modular way, by representing the behaviour of each entity and
+In this paper we consider a variant of the {\tt link}-calculus, which allows to model 
+multiparty interaction in concurrent systems, and show that we can
+represent reaction systems as processes 
+in a modular way, by representing the behaviour of each entity and
 preserving faithfully their features. 

### 2020-04-04 01-introduction.tex
typos e notazione
qualche typo e aggiunta def per complete chains e freccia doppia,
tolto tau finale da flattened sequences

@@
 Then, we present a methodology for verifying formally
-properties of reaction systems. The classical notion of bisimulation for 
+properties of Reaction Systems. The classical notion of bisimulation for 
 process algebras allows to consider two processes as equivalent 
@@
 a subset of the possible biological actions. This is useful for concentrating 
-on the sub-model that one may need to study for a specific study or application,
+on the sub-model that one may need to consider for a specific study or application,
 without getting lost in the complexity of the full biological system, or network.
 The notion of bio-simulation relies on a new and simple assertion language,
-which allows to focus on some properties of the reaction systems to be verified.
+which allows to focus on some properties of the Reaction Systems to be verified.
 In fact, bio-similarity is parametric on a given assertion of interest.
@@
 formalism for specifying properties of labeled transition systems, 
-its variant considered here looks a suitable option to specify formulas 
+its variant introduced here looks a suitable option to specify formulas 
 over complex labels by abstracting from unnecessary details.
@@
 \item we define an assertion language to specify local properties of 
-reaction systems and exploit it to define a novel notion of behavioural equivalence for reaction systems, called bio-similarity;
+Reaction Systems and exploit it to define a novel notion of behavioural equivalence for Reaction Systems, called bio-similarity;
 \item we show that our assertions can be used to extend the Hennessy-Milner logic to our framework, in such a way that logical equivalence of processes coincide with bio-similarity;
@@
 and give more detailed explanations about its use;
-\item we introduce an assertion language to specify the properties of reaction systems;
-\item we define the notion of bio-simulation to relate reaction systems and compare their behaviours;
+\item we introduce an assertion language to specify the properties of Reaction Systems;
+\item we define the notion of bio-simulation to relate Reaction Systems and compare their behaviours;
 \item we show that our assertion language allows to specify properties extending to our
@@
 Section~\ref{sec:biosimulation} presents a methodology for the formal verification of
-properties of reaction systems that are expressed in a novel assertion language.
+properties of Reaction Systems that are expressed in a novel assertion language.
 Section~\ref{sec:discussion} presents some features and 

### 2020-04-04 05-translation.tex
typos e notazione
qualche typo e aggiunta def per complete chains e freccia doppia,
tolto tau finale da flattened sequences

@@
 P_{aj} \defeq \\ 
- ^{r_j} \backslash  \blockchain{s_i}{s_o}{s \in R_j} \, \link{}{} \blockchain{\overline{e}_i}{\overline{e}_o}{e \in I_j} \, \chainedlink{r_{j+1}}{\noact}\chainedlink{\noact}{p_j}\link{}{} \blockchain{\widetilde{c}_i}{\widetilde{c}_o}{c \in P_j}\, \backslash_{p_{j+1}}.P_{aj} & \mbox{\small  \{$aj$ is applicable\}}\\
+ ^{r_j} \backslash  \blockchain{s_i}{s_o}{s \in R_j} \link{}{} \blockchain{\overline{e}_i}{\overline{e}_o}{e \in I_j} \chainedlink{r_{j+1}}{\noact}\chainedlink{\noact}{p_j}\link{}{} \blockchain{\widetilde{c}_i}{\widetilde{c}_o}{c \in P_j} \backslash_{p_{j+1}}.P_{aj} & \mbox{\small  \{$aj$ is applicable\}}\\
  + \ \\
@@
 \item
-$\mathit{reacts}\defeq \{r_1,...,r_u\}$ is the set of reaction names associated with each reaction $aj$;
+$\mathit{reacts}\defeq \{r_1,...,r_u\}$ is the set of reaction names $r_j$ associated with each reaction $aj$;
 \item 
@@
 It has the peculiarity to start and end with silent actions and to include all names in $\mathit{reacts}\cup\mathit{prods}\cup\{\mathit{cxt}\}$.
-As a matter of notation we call such chain $\upsilon$ \emph{complete} (solid and with $\silent$ actions at the extremes).
+As a matter of notation we call such chain $\upsilon$ \emph{complete}.
+
+\begin{definition}[Complete Chain]
+A chain $\upsilon$ is called \emph{complete} if it is solid (i.e., it contains no virtual link) and has silent actions $\silent$ at its extremes.
+We  write $P\xRightarrow{\upsilon} P'$ to mean that $P\xrightarrow{\upsilon} P'$ with $\upsilon$ complete.
+\end{definition}
+
 We will then use $\llparenthesis{\cal A},\gamma\rrparenthesis$ to refer to the encoding without topmost name restrictions, i.e.,
@@
 \llparenthesis{\cal A},\gamma\rrparenthesis \defeq 
-I~|~ \prod_{a\in A} P_a ~|~  \mathit{Cxt}^1~|~\prod_{s\in C_0}P_{s} ~|~ \prod_{s \notin C_0}\overline{P_s} 
+I~|~ \prod_{a\in A} P_a ~|~  \mathit{Cxt}^1~|~\prod_{s\in C_0}P_{s} ~|~ \prod_{s \notin C_0}\overline{P_s} .
 $$
-and will write $P\xRightarrow{\upsilon} P'$ to mean that $P\xrightarrow{\upsilon} P'$ with $\upsilon$ complete.
+and will focus on the complete transitions of $\llparenthesis{\cal A},\gamma\rrparenthesis$.
 
@@
 $$
-\mathit{unflat}(x_1\dots x_n) \defeq\ ^{\silent}\backslash \mathit{unflat}(x_1) \backslash \dots \backslash \mathit{unflat}(x_1) \backslash_{\silent}
+\mathit{unflat}(x_1\dots x_n) \defeq\ ^{\silent}\backslash \mathit{unflat}(x_1) \backslash \dots \backslash \mathit{unflat}(x_n) \backslash_{\silent}
 $$

### 2020-04-04 06b-bisimulation.tex
bioHML example
added bioHML formula at the end of example on bio-sim. corrected small typos
more typos and notation

@@
 and that combines regular expression operators with conjunction and disjunction.
+In the following we let $\mathit{syms} = \mathit{decs}\cup\mathit{reacts}\cup\mathit{prods}\cup\{\mathit{cxt}\}$ be the set of symbolic names used in our assertion language.
 
@@
 %\noindent where $s \in {\cal C}$, and  $\oplus \in \{\wedge, \vee, \to \}$. 
-where $\alpha \in \mathit{names}$ and $N\subseteq \mathit{names}$.
+where $\alpha \in \mathit{syms}$ and $N\subseteq \mathit{syms}$.
 %, and  $\oplus \in \{\wedge, \vee \}$. 
@@
 
-Roughly, a \emph{singleton} assertion $\zeta$ denotes either a string composed by a single symbol $\alpha$ (one of the symbols in the set $\mathit{names}$ for denoting a particular entity, rule, production or context), or
-the wildcard $?$ that stands for any symbol, or the pattern $[N]$ that stands for any of the strings composed by a single symbol in the set $N$.
-Clearly $?$  is just a shorthand for $[\mathit{names}]$.
+Roughly, a \emph{singleton} assertion $\zeta$ denotes either a string composed by a single symbol $\alpha$ (one of the symbols in the set $\mathit{syms}$ for denoting a particular entity, rule, production or context), or
+the wildcard $?$ that stands for any symbol, or the pattern $[N]$ that stands for any of the strings composed by a single symbol in the set $N$. Clearly $?$  is just a shorthand for $[\mathit{syms}]$.
 We write $\mathbf{0}$ for $[\emptyset]$ and $[s_1,...,s_n]$ instead of $[\{s_1,...,s_n\}]$.
@@
 
-An assertion denotes a set of strings over the alphabet $\mathit{names}$ as expected.
+An assertion denotes a set of strings over the alphabet $\mathit{syms}$ as expected.
 
 \begin{definition}[Semantics of assertions]
- We define $\llbracket F\rrbracket \subseteq\wp(\mathit{names}^*)$ by induction on the structure of $F$: 
+ We define $\llbracket F\rrbracket \subseteq\wp(\mathit{syms}^*)$ by induction on the structure of $F$: 
 \[
-\begin{array}{rcl}
- \llbracket \alpha \rrbracket &\defeq&  \{\alpha\} \\
- \llbracket ? \rrbracket &\defeq&   \mathit{names}\\
- \llbracket [N]\rrbracket &\defeq&   N\\
- \llbracket \epsilon \rrbracket &\defeq&  \{\epsilon\} \\
- \llbracket F_1::F_2 \rrbracket &\defeq& \{\omega_1 :: \omega_2~\mid~ \omega_1 \in \llbracket F_1 \rrbracket \wedge \omega_2 \in \llbracket F_2 \rrbracket\}  \\
- \llbracket F^+ \rrbracket &\defeq& \llbracket F \rrbracket^+  \\
- \llbracket F^* \rrbracket &\defeq& \llbracket F \rrbracket^*  \\
- \llbracket  F_1 \vee F_2 \rrbracket &\defeq& \llbracket  F_1\rrbracket \cup \llbracket  F_2\rrbracket  \\
- \llbracket  F_1 \wedge F_2 \rrbracket &\defeq& \llbracket F_1 \rrbracket \cap \llbracket F_2\rrbracket 
+\begin{array}{rcl@{\hspace{1cm}}rcl}
+ \llbracket \epsilon \rrbracket &\defeq&  \{\epsilon\}
+ & \llbracket F_1::F_2 \rrbracket &\defeq& \{\omega_1 :: \omega_2~\mid~ \omega_1 \in \llbracket F_1 \rrbracket \wedge \omega_2 \in \llbracket F_2 \rrbracket\}  \\
+ \llbracket \alpha \rrbracket &\defeq&  \{\alpha\} 
+ & \llbracket F^+ \rrbracket &\defeq& \llbracket F \rrbracket^+  \\
+ \llbracket ? \rrbracket &\defeq&   \mathit{syms}
+ & \llbracket F^* \rrbracket &\defeq& \llbracket F \rrbracket^*  \\
+ \llbracket [N]\rrbracket &\defeq&   N
+ & \llbracket  F_1 \vee F_2 \rrbracket &\defeq& \llbracket  F_1\rrbracket \cup \llbracket  F_2\rrbracket  \\
+ &&&\llbracket  F_1 \wedge F_2 \rrbracket &\defeq& \llbracket F_1 \rrbracket \cap \llbracket F_2\rrbracket 
 \end{array}
@@
 We need to prove that $Q\entails G$.
-If there is no $\upsilon\entails \chi$ such that $Q\xRightarrow{\upsilon}Q'$ for some $Q'$ then 
+If there is no $\upsilon\entails \chi$ such that $Q\xRightarrow{\upsilon}Q'$ for some $Q'$, then 
  $Q\entails [\chi] H=G$ trivially.

### 2020-04-04 07-discussion.tex
more typos and notation

@@
 the behaviour of the context, the possibility of alternative behaviour of
-mutated entities, the communication between two different reaction systems.
+mutated entities, the communication between two different Reaction Systems.
 It is important to note that our encoding guarantees that from each state, 
@@
 
-\subsection{Communicating reaction systems} 
+\subsection{Communicating Reaction Systems} 
 We sketch how it is possible to program two RSs encodings, in a way that 
@@
 \end{tabular}
-\caption{The two reaction systems $rs1$ and $rs2$.}
+\caption{The two Reaction Systems $rs1$ and $rs2$.}
 \label{tab:rules}
@@
 
-As we have briefly sketched, our model of two \emph{communicating reaction systems} 
+As we have briefly sketched, our model of two \emph{communicating Reaction Systems} 
 can enable the study of the behaviour of one RS

### 2020-04-04 08-conclusion.tex
more typos and notation

@@
 %shown 
-an elegant embedding of reaction systems,
+a faithful embedding of Reaction Systems,
 an emerging formalism to model computationally biochemical systems.
@@
 For instance, the context behaviour can also be expressed recursively. 
-Entity mutations can be expressed easily. Reaction systems can
+Entity mutations can be expressed easily. Reaction Systems can
 communicate between them.
@@
 We have then modified the classical notion (in concurrency) of bisimulation
-making it more suitable for proving properties of reaction systems in our
+making it more suitable for proving properties of Reaction Systems in our
 framework.
@@
 We believe that our embedding can contribute to extend the applications
-of reaction systems to diverse fields of computer science, and life
+of Reaction Systems to diverse fields of computer science, and life
 sciences.

### 2020-04-04 05-translation.tex
more typos and notation
widetilde widehat proof in appendice

--- 2020-04-03 05-translation.tex
@@
   \draw (4.75,0.25)  node[minimum size=0.48cm,draw,fill=yellow,yellow] {$\ $};
-  \draw (5.2,0.255)  node[minimum size=0.485cm,draw,fill=yellow,draw=yellow] {${\scriptscriptstyle  \hat{s1}_i}$};
+  \draw (5.2,0.255)  node[minimum size=0.485cm,draw,fill=yellow,draw=yellow] {${\scriptscriptstyle  \widehat{s1}_i}$};
   \draw(5,1) -- (5,0.5) -- (5.5,0.5) -- (5.5,0);
@@
  \draw (6.36,0.75) node[minimum size=0.475cm,draw,fill=yellow,draw=yellow] 
- {${\scriptscriptstyle \hat{s2}_o \ }$};
+ {${\scriptscriptstyle \widehat{s2}_o \ }$};
  %node[minimum size=0.5cm,draw,draw=white] {$\noact$};

### 2020-04-04 06b-bisimulation.tex
widetilde widehat proof in appendice

--- 2020-04-02 06b-bisimulation.tex
@@
 \item
-$\star :: p_j :: [  \mathit{ents} ]^* :: \tilde{s}_i :: \star$
+$\star :: p_j :: [  \mathit{ents} ]^* :: \widetilde{s}_i :: \star$
 \item
-$\star :: \tilde{s}_i :: \star$
+$\star :: \widetilde{s}_i :: \star$
 \item
-$\star :: \hat{s}_i :: \star$
+$\star :: \widehat{s}_i :: \star$
 \item
@@
 
-\begin{theorem}
+\begin{restatable}[Correspondence]{theorem}{theocorr}
+%\begin{theorem}
 $\sim_F\ =\ \equiv_{\mathcal{L}_F}$
-\end{theorem}
-\begin{proof}
-The proof is just an adaptation of the classical result. The two implications are proved separately.
-\begin{description}
-\item[$\sim_F\ \subseteq\ \equiv_{\mathcal{L}_F}$:]
-Given any two processes $P\sim_F Q$ we need to prove that for any bioHML formula $G$ we have $P\entails G$ iff $Q\entails G$. 
-Without loss of generality, we prove that $P\entails G$ implies $Q\entails G$.
-The proof is by structural induction on $G$.
-\begin{itemize}
-\item 
-if $G={\tt t}$, then $Q\entails G$.
-
-\item
-if $G={\tt f}$, then the assumption $P\entails G$ is false and the implication holds.
-
-\item
-if $G=G_1 \wedge G_2$ we take as inductive hypotheses that
-$$
-\begin{array}{l}
-\forall R,S.\ R \sim_F S \wedge R\entails G_1 \Rightarrow S \entails G_1\\
-\forall R,S.\ R \sim_F S \wedge R\entails G_2 \Rightarrow S \entails G_2
-\end{array}
-$$
-We need to prove that $Q\entails G$.
-Since $P\entails G=G_1 \wedge G_2$ we have $P\entails G_1$ and $P\entails G_2$.
-Since $P\sim_F Q$, by inductive hypotheses we get $Q\entails G_1$ and $Q\entails G_2$.
-Hence $Q\entails G_1 \wedge G_2=G$.
-
-\item
-if $G=G_1 \vee G_2$  we take as inductive hypotheses that
-$$
-\begin{array}{l}
-\forall R,S.\ R \sim_F S \wedge R\entails G_1 \Rightarrow S \entails G_1\\
-\forall R,S.\ R \sim_F S \wedge R\entails G_2 \Rightarrow S \entails G_2
-\end{array}
-$$
-We need to prove that $Q\entails G$.
-Since $P\entails G=G_1 \vee G_2$ we have $P\entails G_1$ or $P\entails G_2$.
-If $P\entails G_1$, since $P\sim_F Q$, by inductive hypotheses we get $Q\entails G_1$ and thus $Q\entails G_1 \vee G_2=G$.
-If $P\entails G_2$, since $P\sim_F Q$, by inductive hypotheses we get $Q\entails G_2$ and thus $Q\entails G_1 \vee G_2=G$.
-
-\item
-if $G=\langle\chi\rangle H$ we take as inductive hypothesis that
-$$
-\forall R,S.\ R \sim_F S \wedge R\entails H \Rightarrow S \entails H
-$$
-We need to prove that $Q\entails G$.
-Since $P\entails \langle\chi\rangle H$ it means that there exists $\upsilon, P'$ such that $P\xRightarrow{\upsilon} P'$ with $\upsilon\entails \chi$ and $P'\entails H$.
-Since $P\sim_F Q$, there exists $w,Q'$ such that $Q\xRightarrow{w} Q'$ with $w\entails \chi$ and $P'\sim_F Q'$.
+%\end{theorem}
+\end{restatable}
+%\begin{proof}
+%The proof is just an adaptation of the classical result. The two implications are proved separately.
+%\begin{description}
+%\item[$\sim_F\ \subseteq\ \equiv_{\mathcal{L}_F}$:]
+%Given any two processes $P\sim_F Q$ we need to prove that for any bioHML formula $G$ we have $P\entails G$ iff $Q\entails G$. 
+%Without loss of generality, we prove that $P\entails G$ implies $Q\entails G$.
+%The proof is by structural induction on $G$.
+%\begin{itemize}
+%\item 
+%if $G={\tt t}$, then $Q\entails G$.
+%
+%\item
+%if $G={\tt f}$, then the assumption $P\entails G$ is false and the implication holds.
+%
+%\item
+%if $G=G_1 \wedge G_2$ we take as inductive hypotheses that
+%$$
+%\begin{array}{l}
+%\forall R,S.\ R \sim_F S \wedge R\entails G_1 \Rightarrow S \entails G_1\\
+%\forall R,S.\ R \sim_F S \wedge R\entails G_2 \Rightarrow S \entails G_2
+%\end{array}
+%$$
+%We need to prove that $Q\entails G$.
+%Since $P\entails G=G_1 \wedge G_2$ we have $P\entails G_1$ and $P\entails G_2$.
+%Since $P\sim_F Q$, by inductive hypotheses we get $Q\entails G_1$ and $Q\entails G_2$.
+%Hence $Q\entails G_1 \wedge G_2=G$.
+%
+%
+%\item
+%if $G=G_1 \vee G_2$  we take as inductive hypotheses that
+%$$
+%\begin{array}{l}
+%\forall R,S.\ R \sim_F S \wedge R\entails G_1 \Rightarrow S \entails G_1\\
+%\forall R,S.\ R \sim_F S \wedge R\entails G_2 \Rightarrow S \entails G_2
+%\end{array}
+%$$
+%We need to prove that $Q\entails G$.
+%Since $P\entails G=G_1 \vee G_2$ we have $P\entails G_1$ or $P\entails G_2$.
+%If $P\entails G_1$, since $P\sim_F Q$, by inductive hypotheses we get $Q\entails G_1$ and thus $Q\entails G_1 \vee G_2=G$.
+%If $P\entails G_2$, since $P\sim_F Q$, by inductive hypotheses we get $Q\entails G_2$ and thus $Q\entails G_1 \vee G_2=G$.
+%
+%\item
+%if $G=\langle\chi\rangle H$ we take as inductive hypothesis that
+%$$
+%\forall R,S.\ R \sim_F S \wedge R\entails H \Rightarrow S \entails H
+%$$
+%We need to prove that $Q\entails G$.
+%Since $P\entails \langle\chi\rangle H$ it means that there exists $\upsilon, P'$ such that $P\xRightarrow{\upsilon} P'$ with $\upsilon\entails \chi$ and $P'\entails H$.
+%Since $P\sim_F Q$, there exists $w,Q'$ such that $Q\xRightarrow{w} Q'$ with $w\entails \chi$ and $P'\sim_F Q'$.
 Then, by inductive hypothesis, $Q'\entails H$ and thus $Q\entails \langle\chi\rangle H=G$.
-
-\item
-if $G=[\chi] H$ we take as inductive hypothesis that
-$$
-\forall R,S.\ R \sim_F S \wedge R\entails H \Rightarrow S \entails H
-$$
-We need to prove that $Q\entails G$.
-If there is no $\upsilon\entails \chi$ such that $Q\xRightarrow{\upsilon}Q'$ for some $Q'$ then 
- $Q\entails [\chi] H=G$ trivially.
-For any $\upsilon,Q'$ such that $Q\xRightarrow{\upsilon}Q'$ with $\upsilon\entails \chi$, then as $P\sim_F Q$ there must exist $w,P'$ such that $P\xRightarrow{w}P'$ with $w\entails \chi$ and $P'\sim_F Q'$.
+%
+%\item
+%if $G=[\chi] H$ we take as inductive hypothesis that
+%$$
+%\forall R,S.\ R \sim_F S \wedge R\entails H \Rightarrow S \entails H
+%$$
+%We need to prove that $Q\entails G$.
+%If there is no $\upsilon\entails \chi$ such that $Q\xRightarrow{\upsilon}Q'$ for some $Q'$, then 
+% $Q\entails [\chi] H=G$ trivially.
+%For any $\upsilon,Q'$ such that $Q\xRightarrow{\upsilon}Q'$ with $\upsilon\entails \chi$, then as $P\sim_F Q$ there must exist $w,P'$ such that $P\xRightarrow{w}P'$ with $w\entails \chi$ and $P'\sim_F Q'$.
 Since $P\entails G=[\chi] H$ then it must be $P'\entails H$.
@@
 Hence $Q\entails [\chi] H = G$.
-\end{itemize}
-
-\item[$\equiv_{\mathcal{L}_F}\ \subseteq\ \sim_F$:]
-We prove that $\equiv_{\mathcal{L}_F}$ is a bio-simulation and thus included in $\sim_F$.
-Take two generic processes $P\equiv_{\mathcal{L}_F} Q$ and suppose $P\xRightarrow{\upsilon}P'$ for some $\upsilon,P'$. 
-\begin{itemize}
-\item
-If $\upsilon \entails F$ we want to prove that there exists some $w,Q'$ such that $Q\xRightarrow{w}Q'$, with $w\entails F$ and $P'\equiv_{\mathcal{L}_F} Q'$.
-
-Towards a contradiction, assume that we cannot find such $w,Q'$.
-
-If there is no transition $Q\xRightarrow{w}Q'$ such that $w\entails F$, then the bioHML formula $G \defeq \langle F\rangle{\tt t}$ is such that $P\entails G$ and $Q\not\entails G$, contradicting the assumption $P\equiv_{\mathcal{L}_F} Q$.
-
-Otherwise, let $\mathcal{Q}\defeq \{Q'~\mid~\exists w.\ Q\xRightarrow{w}Q' \ \wedge\ w\entails F\}$ be the set of processes reachable from $Q$ via a transition with a (complete) label that satisfies $F$.
+%\end{itemize}
+%
+%\item[$\equiv_{\mathcal{L}_F}\ \subseteq\ \sim_F$:]
+%We prove that $\equiv_{\mathcal{L}_F}$ is a bio-simulation and thus included in $\sim_F$.
+%Take two generic processes $P\equiv_{\mathcal{L}_F} Q$ and suppose $P\xRightarrow{\upsilon}P'$ for some $\upsilon,P'$. 
+%\begin{itemize}
+%\item
+%If $\upsilon \entails F$ we want to prove that there exists some $w,Q'$ such that $Q\xRightarrow{w}Q'$, with $w\entails F$ and $P'\equiv_{\mathcal{L}_F} Q'$.
+%
+%Towards a contradiction, assume that we cannot find such $w,Q'$.
+%
+%If there is no transition $Q\xRightarrow{w}Q'$ such that $w\entails F$, then the bioHML formula $G \defeq \langle F\rangle{\tt t}$ is such that $P\entails G$ and $Q\not\entails G$, contradicting the assumption $P\equiv_{\mathcal{L}_F} Q$.
+%
+%Otherwise, let $\mathcal{Q}\defeq \{Q'~\mid~\exists w.\ Q\xRightarrow{w}Q' \ \wedge\ w\entails F\}$ be the set of processes reachable from $Q$ via a transition with a (complete) label that satisfies $F$.
 Since our processes are with guarded recursion, the set $\mathcal{Q}$ is finite.
@@
 But then the formula $G\defeq \langle F\rangle(G_1\wedge\dots\wedge G_n)$ is such that $P\entails G$ and $Q\not\entails G$, contradicting the assumption $P\equiv_{\mathcal{L}_F} Q$.
-
-\item
-If $\upsilon \not\entails F$ then the proof is analogous to the previous case and thus omitted.
-\end{itemize}
-\end{description}
-\end{proof}
+%
+%\item
+%If $\upsilon \not\entails F$ then the proof is analogous to the previous case and thus omitted.
+%\end{itemize}
+%\end{description}
+%\end{proof}

### 2020-04-04 07-discussion.tex
widetilde widehat proof in appendice

--- 2020-04-03 07-discussion.tex
@@
 Our encoding increases the expressivity of  RS  concerning:
-the behaviour of the context, the possibility of alternative behaviour of
-mutated entities, the communication between two different reaction systems.
+%the behaviour of the context, 
+the possibility of alternative behaviour of
+mutated entities, and  the communication between two different reaction systems.
 It is important to note that our encoding guarantees that from each state, 
@@
 
-\subsection{Recursive contexts}
-In RS, the behaviour of the context is finite. For the first $n$ steps, it is specified 
- which are the entities that are provided from the context. 
-Using c\CNA\ we can describe in a natural way the behaviour of the context in a recursive way. 
-Then, the context behaviour would not necessarily end after $n$ steps,
-and  could be infinite.
-For example, in an extended interactive process, we may want that the entity $s$ is 
-intermittently provided by the context every two steps: 
-\[
-\begin{array}{lcll}
-Cxt_s & \defeq &  \startchain{cxt_j}\chainedlink{\hat{s}_i}{\noact}\chainedlink{\noact}{\hat{s}_o}\chainend{cxt_{j+1}}.Cxt_{s'}, &\mbox{ context provides $s$;} \\
-Cxt_{s'} & \defeq & \startchain{cxt_j}\chainedlink{\underline{s}_i}{\noact}\chainedlink{\noact}{\underline{s}_o}\chainend{cxt_{j+1}}.Cxt_{s''},& \mbox{ context doesn't provide $s$;} \\
-Cxt_{s''} & \defeq & \startchain{cxt_j}\chainedlink{\underline{s}_i}{\noact}\chainedlink{\noact}{\underline{s}_o}\chainend{cxt_{j+1}}.Cxt_{s}, & \mbox{ context doesn't provide $s$.} \\
-\end{array}
-\]
+%\subsection{Recursive contexts}
+%In RS, the behaviour of the context is finite. For the first $n$ steps, it is specified 
+% which are the entities that are provided from the context. 
+%Using c\CNA\ we can describe in a natural way the behaviour of the context in a recursive way. 
+%Then, the context behaviour would not necessarily end after $n$ steps,
+%and  could be infinite.
+%For example, in an extended interactive process, we may want that the entity $s$ is 
+%intermittently provided by the context every two steps: 
+%\[
+%\begin{array}{lcll}
+%Cxt_s & \defeq &  \startchain{cxt_j}\chainedlink{\hat{s}_i}{\noact}\chainedlink{\noact}{\hat{s}_o}\chainend{cxt_{j+1}}.Cxt_{s'}, &\mbox{ context provides $s$;} \\
+%Cxt_{s'} & \defeq & \startchain{cxt_j}\chainedlink{\underline{s}_i}{\noact}\chainedlink{\noact}{\underline{s}_o}\chainend{cxt_{j+1}}.Cxt_{s''},& \mbox{ context doesn't provide $s$;} \\
+%Cxt_{s''} & \defeq & \startchain{cxt_j}\chainedlink{\underline{s}_i}{\noact}\chainedlink{\noact}{\underline{s}_o}\chainend{cxt_{j+1}}.Cxt_{s}, & \mbox{ context doesn't provide $s$.} \\
+%\end{array}
+%\]
 

### 2020-04-04 08-conclusion.tex
widetilde widehat proof in appendice

--- 2020-03-29 08-conclusion.tex
@@
 This translation shows several benefits.
-For instance, the context behaviour can also be expressed recursively. 
-Entity mutations can be expressed easily. Reaction systems can
+{\color{red} For instance, in our paper the context works non deterministically and it is recursively defined.}
+%  the context behaviour is expressed recursively and no be expressed recursively. 
+Also, entity mutations can be expressed easily and different reaction systems can
 communicate between them.

### 2020-04-04 09-appendix.tex
widetilde widehat proof in appendice

@@
 \section{Omitted Proofs}
-In this section we report the proofs for the results in Section~\ref{sec:trans}.
+In this section we report the proofs for the results in Section~\ref{sec:trans} and in Section~\ref{sec:biosimulation}.
 %\noindent
@@
  \begin{proof}
- The proof proceeds by induction on the number $j$, and it is similar to the one of Corollary~\ref{corr:corr2}\\
+ The proof proceeds by induction on the number $j$, and it is similar to the one of Corollary~\ref{corr:corr2}.\\
 % case $j=0$\\
@@
  \end{proof}
+
+ \theocorr*
+ \begin{proof}
+The proof is just an adaptation of the classical result. The two implications are proved separately.
+\begin{description}
+\item[$\sim_F\ \subseteq\ \equiv_{\mathcal{L}_F}$:]
+Given any two processes $P\sim_F Q$ we need to prove that for any bioHML formula $G$ we have $P\entails G$ iff $Q\entails G$. 
+Without loss of generality, we prove that $P\entails G$ implies $Q\entails G$.
+The proof is by structural induction on $G$.
+\begin{itemize}
+\item 
+if $G={\tt t}$, then $Q\entails G$.
+
+\item
+if $G={\tt f}$, then the assumption $P\entails G$ is false and the implication holds.
+
+\item
+if $G=G_1 \wedge G_2$ we take as inductive hypotheses that
+$$
+\begin{array}{l}
+\forall R,S.\ R \sim_F S \wedge R\entails G_1 \Rightarrow S \entails G_1\\
+\forall R,S.\ R \sim_F S \wedge R\entails G_2 \Rightarrow S \entails G_2
+\end{array}
+$$
+We need to prove that $Q\entails G$.
+Since $P\entails G=G_1 \wedge G_2$ we have $P\entails G_1$ and $P\entails G_2$.
+Since $P\sim_F Q$, by inductive hypotheses we get $Q\entails G_1$ and $Q\entails G_2$.
+Hence $Q\entails G_1 \wedge G_2=G$.
+
+\item
+if $G=G_1 \vee G_2$  we take as inductive hypotheses that
+$$
+\begin{array}{l}
+\forall R,S.\ R \sim_F S \wedge R\entails G_1 \Rightarrow S \entails G_1\\
+\forall R,S.\ R \sim_F S \wedge R\entails G_2 \Rightarrow S \entails G_2
+\end{array}
+$$
+We need to prove that $Q\entails G$.
+Since $P\entails G=G_1 \vee G_2$ we have $P\entails G_1$ or $P\entails G_2$.
+If $P\entails G_1$, since $P\sim_F Q$, by inductive hypotheses we get $Q\entails G_1$ and thus $Q\entails G_1 \vee G_2=G$.
+If $P\entails G_2$, since $P\sim_F Q$, by inductive hypotheses we get $Q\entails G_2$ and thus $Q\entails G_1 \vee G_2=G$.
+
+\item
+if $G=\langle\chi\rangle H$ we take as inductive hypothesis that
+$$
+\forall R,S.\ R \sim_F S \wedge R\entails H \Rightarrow S \entails H
+$$
+We need to prove that $Q\entails G$.
+Since $P\entails \langle\chi\rangle H$ it means that there exists $\upsilon, P'$ such that $P\xRightarrow{\upsilon} P'$ with $\upsilon\entails \chi$ and $P'\entails H$.
+Since $P\sim_F Q$, there exists $w,Q'$ such that $Q\xRightarrow{w} Q'$ with $w\entails \chi$ and $P'\sim_F Q'$.
+Then, by inductive hypothesis, $Q'\entails H$ and thus $Q\entails \langle\chi\rangle H=G$.
+
+\item
+if $G=[\chi] H$ we take as inductive hypothesis that
+$$
+\forall R,S.\ R \sim_F S \wedge R\entails H \Rightarrow S \entails H
+$$
+We need to prove that $Q\entails G$.
+If there is no $\upsilon\entails \chi$ such that $Q\xRightarrow{\upsilon}Q'$ for some $Q'$, then 
+ $Q\entails [\chi] H=G$ trivially.
+For any $\upsilon,Q'$ such that $Q\xRightarrow{\upsilon}Q'$ with $\upsilon\entails \chi$, then as $P\sim_F Q$ there must exist $w,P'$ such that $P\xRightarrow{w}P'$ with $w\entails \chi$ and $P'\sim_F Q'$.
+Since $P\entails G=[\chi] H$ then it must be $P'\entails H$.
+Since $P'\sim_F Q'$, by inductive hypothesis $Q'\entails H$.
+Hence $Q\entails [\chi] H = G$.
+\end{itemize}
+
+\item[$\equiv_{\mathcal{L}_F}\ \subseteq\ \sim_F$:]
+We prove that $\equiv_{\mathcal{L}_F}$ is a bio-simulation and thus included in $\sim_F$.
+Take two generic processes $P\equiv_{\mathcal{L}_F} Q$ and suppose $P\xRightarrow{\upsilon}P'$ for some $\upsilon,P'$. 
+\begin{itemize}
+\item
+If $\upsilon \entails F$ we want to prove that there exists some $w,Q'$ such that $Q\xRightarrow{w}Q'$, with $w\entails F$ and $P'\equiv_{\mathcal{L}_F} Q'$.
+
+Towards a contradiction, assume that we cannot find such $w,Q'$.
+
+If there is no transition $Q\xRightarrow{w}Q'$ such that $w\entails F$, then the bioHML formula $G \defeq \langle F\rangle{\tt t}$ is such that $P\entails G$ and $Q\not\entails G$, contradicting the assumption $P\equiv_{\mathcal{L}_F} Q$.
+
+Otherwise, let $\mathcal{Q}\defeq \{Q'~\mid~\exists w.\ Q\xRightarrow{w}Q' \ \wedge\ w\entails F\}$ be the set of processes reachable from $Q$ via a transition with a (complete) label that satisfies $F$.
+Since our processes are with guarded recursion, the set $\mathcal{Q}$ is finite.
+Let $\mathcal{Q} = \{Q'_1,...,Q'_n\}$.
+By hypothesis all processes in $\mathcal{Q}$ must not be bio-logically equivalent to $P'$, hence for any $i\in[1,n]$ there exists a bioHML formula $G_i$ such that $P'\entails G_i$ and $Q'_i\not\entails G_i$ (if it was the opposite, $P'\not\entails G_i$ and $Q'_i\entails G_i$, we can use the converse formulas).
+But then the formula $G\defeq \langle F\rangle(G_1\wedge\dots\wedge G_n)$ is such that $P\entails G$ and $Q\not\entails G$, contradicting the assumption $P\equiv_{\mathcal{L}_F} Q$.
+
+\item
+If $\upsilon \not\entails F$ then the proof is analogous to the previous case and thus omitted.
+\end{itemize}
+\end{description}
+\end{proof}

### 2020-04-04 05-translation.tex

--- 2020-04-04 05-translation.tex
@@
  \[
- \begin{array}{l@{\hspace{-0.009cm}}c@{\hspace{-0.009cm}}ll@{\hspace{-0.009cm}}c@{\hspace{-0.009cm}}l}
- P_{r_1} &\defeq& \startchain{r_1}\chainedlink{s1_i}{\noact}\chainedlink{\noact}{s1_o}\chainedlink{r_2}{\noact} \chainedlink{\noact}{p_1}\chainedlink{\widetilde{s2}_i}{\noact}\chainedlink{\noact}{\widetilde{s2}_o}\chainend{p_2}.P_{r_1} + \dots;  &
+ \begin{array}{rcl}
+ P_{r_1} &\defeq& \startchain{r_1}\chainedlink{s1_i}{\noact}\chainedlink{\noact}{s1_o}\chainedlink{r_2}{\noact} \chainedlink{\noact}{p_1}\chainedlink{\widetilde{s2}_i}{\noact}\chainedlink{\noact}{\widetilde{s2}_o}\chainend{p_2}.P_{r_1} + \dots;   \\
+  P_{r_2} &\defeq& \dots + \startchain{r_2}\chainedlink{\overline{s2}_i}{\noact}\chainedlink{\noact}{\overline{s2}_o}\chainedlink{\mathit{cxt}}{\noact} \chainedlink{\noact}{p_2}\chainend{\silent}.P_{r_2} + \dots; \\
    P_{s1} &\defeq& \startchain{s1_i}\chainedlink{s1_o}{\noact}\chainedlink{\noact}{\widehat{s1}_i}\chainend{\widehat{s1}_o}
@@
  .P_{s1} + \dots; \\
-  P_{r_2} &\defeq& \dots + \startchain{r_2}\chainedlink{\overline{s2}_i}{\noact}\chainedlink{\noact}{\overline{s2}_o}\chainedlink{\mathit{cxt}}{\noact} \chainedlink{\noact}{p_2}\chainend{\silent}.P_{r_2} + \dots; \\
   \overline{P}_{s2} &\defeq& \startchain{\overline{s2}_i}\chainedlink{\overline{s2}_o}{\noact}
@@
  \llparenthesis{\cal A},\gamma\rrparenthesis \xrightarrow{
+ \startchain{\silent}
+ \chainedlink{r_1}{r_1}
+ \chainedlink{s1_i}{s1_i}
+ \chainedlink{s1_o}{s1_o}
+ \chainedlink{r_2}{r_2}
+ \chainedlink{\overline{s2}_i}{\overline{s2}_i}
+ \chainedlink{\overline{s2}_o}{\overline{s2}_o}
+ \chainedlink{\mathit{cxt}}{\mathit{cxt}}
+ \chainedlink{\widehat{s1}_i}{\widehat{s1}_i}
+ \chainedlink{\widehat{s1}_o}{\widehat{s1}_o}
+  \chainedlink{\widehat{s2}_i}{\widehat{s2}_i}
+  \chainedlink{\widehat{s2}_o}{\widehat{s2}_o}
+  \chainedlink{p_1}{p_1}
+  \chainedlink{\widetilde{s2}_i}{\widetilde{s2}_i}
+  \chainedlink{\widetilde{s2}_o}{\widetilde{s2}_o}
+  \chainedlink{p_2}{p_2}
+  \chainend{\silent}
+ }
+ P'
+ $$
+
+ Since the chain in the transition label is complete we can also write
+  $$
+ \llparenthesis{\cal A},\gamma\rrparenthesis \xRightarrow{
  \startchain{\silent}

### 2020-04-04 06-examples.tex
cambi
cambi
typos e notazione
qualche typo e aggiunta def per complete chains e freccia doppia,
tolto tau finale da flattened sequences

@@
 This example is inspired by the example in~\cite{BEMR11}, where
-a deterministic transition system is coded in the reaction system
+a deterministic transition system is coded in the Reaction System
 framework.
@@
 \[
-r_1~\overline{a}~r_2~q~b~\overline{w}~\overline{a}~r_3~\overline{a}~r_4~\overline{w}~\mathit{cxt}~\widehat{a}~\underline{b}~p_1~p_2~\widetilde{w}~p_3~p_4~\silent\]
+r_1~\overline{a}~r_2~q~b~\overline{w}~\overline{a}~r_3~\overline{a}~r_4~\overline{w}~\mathit{cxt}~\widehat{a}~\underline{b}~p_1~p_2~\widetilde{w}~p_3~p_4\]
 
@@
 %  \chainedlink{p_2}{p_2}\chainedlink{p_3}{p_3}\chainedlink{p_4}{p_4} \chainedlink{\widetilde{G}_i}{\bf \widetilde{G}_i}\chainedlink{\bf \widetilde{G}_o}{\widetilde{G}_o} \chainend{\silent}
-r_1~ a~G~\overline{C}~ r_2~ G~ r_3~T'~r_4~T'~\overline{F}~cxt~\underline{C}~\underline{G}~\widehat{F}~p_1~\widetilde{T}~p_2~p_3~p_4~\widetilde{G}~\silent
+r_1~ a~G~\overline{C}~ r_2~ G~ r_3~T'~r_4~T'~\overline{F}~cxt~\underline{C}~\underline{G}~\widehat{F}~p_1~\widetilde{T}~p_2~p_3~p_4~\widetilde{G}
 \]
 The original label can then be reconstructed just applying the function $\mathit{unflat}(\cdot)$ to the string above.
+
+In \cite{BBF19} we have shown a more complex example, by modeling a RS 
+of a regulatory network for \emph{lac} operon, 
+presented in~\cite{CMMBM12}.

### 2020-04-04 07-discussion.tex

@@
 the possibility of alternative behaviour of
-mutated entities, and  the communication between two different reaction systems.
+mutated entities, and  the communication between two different Reaction Systems.
 It is important to note that our encoding guarantees that from each state, 
@@
 
-\subsection{Communicating reaction systems} 
+\subsection{Communicating Reaction Systems}
 We sketch how it is possible to program two RSs encodings, in a way that

### 2020-04-04 08-conclusion.tex

--- 2020-04-04 08-conclusion.tex
@@
 This translation shows several benefits.
-For instance, the context behaviour can also be expressed recursively. 
-Entity mutations can be expressed easily. Reaction Systems can
+{\color{red} For instance, in our paper the context works non deterministically and it is recursively defined.}
+%  the context behaviour is expressed recursively and no be expressed recursively. 
+Also, entity mutations can be expressed easily and different reaction systems can
 communicate between them.

### 2020-04-04 07-discussion.tex
widehat

@@
 \begin{array}{lclcl}
-\llbracket s \rrbracket &\defeq &P_{s} &\defeq & \startchain{{\color{blue}s_i}}\chainedlink{{\color{blue}s_o}}{\noact}\chainedlink{\noact}{{\color{blue}\hat{s}_i}}\chainend{{\color{blue}\hat{s}_o}}.P_{s} + \startchain{s_i}\chainend{s_o}.\overline{P_{s}}\\
-&& \overline{P_{s}}&\defeq &\startchain{\overline{s}_i}\chainedlink{\overline{s}_o}{\noact}\chainedlink{\noact}{\hat{s}_i}\chainend{\hat{s}_o}.P_{s}+\startchain{\overline{s}_i}\chainend{\overline{s}_i}.\overline{P_{s}}\\
+\llbracket s \rrbracket &\defeq &P_{s} &\defeq & \startchain{{\color{blue}s_i}}\chainedlink{{\color{blue}s_o}}{\noact}\chainedlink{\noact}{{\color{blue}\widehat{s}_i}}\chainend{{\color{blue}\widehat{s}_o}}.P_{s} + \startchain{s_i}\chainend{s_o}.\overline{P_{s}}\\
+&& \overline{P_{s}}&\defeq &\startchain{\overline{s}_i}\chainedlink{\overline{s}_o}{\noact}\chainedlink{\noact}{\widehat{s}_i}\chainend{\widehat{s}_o}.P_{s}+\startchain{\overline{s}_i}\chainend{\overline{s}_i}.\overline{P_{s}}\\
 %\begin{array}{lclclll}
@@
 $s$ ($ss$ in $rs2$) to be provided in $rs1$. Every time $ss$ is produced in $rs2$,  a virtual link is
-created to synchronise with $rs1$   on link $\link{\hat{s}_i}{\hat{s}_o}$:
+created to synchronise with $rs1$   on link $\link{\widehat{s}_i}{\widehat{s}_o}$:
 \[
 \begin{array}{lclcl}
-\llbracket ss \rrbracket& \defeq& P_{ss}&\defeq &\startchain{\tilde{ss}_i}\chainedlink{\hat{s}_i}{\noact}\chainedlink{\noact}{\hat{s}_o}\chainend{\tilde{ss}_o}.P_{ss}+
+\llbracket ss \rrbracket& \defeq& P_{ss}&\defeq &\startchain{\tilde{ss}_i}\chainedlink{\widehat{s}_i}{\noact}\chainedlink{\noact}{\widehat{s}_o}\chainend{\tilde{ss}_o}.P_{ss}+
 \startchain{dss_i}   \chainend{dss_o}.\overline{P_{ss}} \\
@@
 %\begin{array}{lcllcl}
-&&\overline{P_{ss}}&\defeq &\startchain{{\color{red}\tilde{ss}_i}}\chainedlink{{\color{red}\hat{s}_i}}{{\color{blue}\noact}}\chainedlink{{\color{blue}\noact}}{{\color{red}\hat{s}_o}}\chainend{{\color{red}\tilde{ss}_o}}.P_{ss}
+&&\overline{P_{ss}}&\defeq &\startchain{{\color{red}\tilde{ss}_i}}\chainedlink{{\color{red}\widehat{s}_i}}{{\color{blue}\noact}}\chainedlink{{\color{blue}\noact}}{{\color{red}\widehat{s}_o}}\chainend{{\color{red}\tilde{ss}_o}}.P_{ss}
 +
@@
 \startchain{\silent}\chainedlink{s_i}{{\color{blue}s_i}}\chainedlink{{\color{blue}s_o}}{s_o}\chainedlink{\tilde{x}_i}{{\color{gray}\tilde{x}_i}}\chainedlink{{\color{gray}\tilde{x}_o}}{\tilde{x}_o} \chainedlink{a_2}{a_2}\chainedlink{y_i}{ {\color{gray}y_i}}\chainedlink{{\color{gray}y_o}}{y_o}\chainedlink{\tilde{ss}_i}{{\color{red}\tilde{ss}_i}}
-\chainedlink{{\color{red}\hat{s}_i}}{{\color{blue}\hat{s}_i}}
-\chainedlink{{\color{blue}\hat{s}_o}}{{\color{red}\hat{s}_o}}
+\chainedlink{{\color{red}\widehat{s}_i}}{{\color{blue}\widehat{s}_i}}
+\chainedlink{{\color{blue}\widehat{s}_o}}{{\color{red}\widehat{s}_o}}
 \chainedlink{{\color{red}\tilde{ss}_o}}{\tilde{ss}_o}\chainend{\silent},

### 2020-04-04 04-CCNA.tex
small corrections
Update 04-CCNA.tex

@@
 of a variant of the {\tt link}-calculus~\cite{BodeiBB12}, the c\CNA\  (chained \CNA)
-where the prefixes are link chains.
+where the prefixes are link chains and not just links.
 %In this section we just focus on the basic ideas of link
@@
 Virtual links represent missing elements of a chain. 
+A chain is called \emph{solid} if it does not contain any virtual link.
 The equivalence $\blackstretcheq$ models expansion/contraction of virtual links to adjust the length of a link chain.

### 2020-04-04 06b-bisimulation.tex
cambi piccoli
cambi piccoli

@@
 
-For example we would like to express properties about each step of the simulation of a system like the following:
+For instance we would like to express properties about each step of the simulation of a system like in the following:
+
+\begin{example} \label{properties_ex}
 \begin{enumerate}
@@
 \end{enumerate}
+\end{example}
 
@@
 
-The corresponding formulas to the sample queries listed above are as follows:\\
+The formulas corresponding to the sample queries listed in Example~\ref{properties_ex} are as follows:
 

### 2020-04-04 07-discussion.tex
cambi piccoli
cambi piccoli

@@
 
-As we have briefly sketched, our model of two \emph{communicating Reaction Systems} 
-can enable the study of the behaviour of one RS
-in relation to another one.
-In Example~\ref{ex:twoRS} describing  the behaviour of the \emph{lac} operon,
-the two entities lactose and glucose are controlled 
-non deterministically by the context.
-In our framework,
-instead, by exploiting the expressivity of c\CNA, 
-the \emph{lac} operon system has been connected with the two systems 
-producing the lactose and the glucose.
-This way, the presence of these two entities in the \emph{lac} operon system 
-can be regulated by realistic mechanisms.
+As we have briefly sketched, our model of two 
+ \emph{communicating Reaction Systems} can enable the study of the behaviour of one RS in relation to another one. 
+Thus, the products of the reactions of one RS can become the input for another one. 
+This could allow for a modular approach to modeling complex systems, by composing different Reaction Systems.
+
 %which can be modified to represent pathological states as well.

### 2020-04-05 05-translation.tex
small corrections
Update 05-translation.tex
small adjustments

@@
 \item
-$\mathit{reacts}\defeq \{r_1,...,r_u\}$ is the set of reaction names $r_j$ associated with each reaction $aj$;
+$\mathit{reacts}\defeq \{r_1,...,r_{|A|}\}$ is the set of reaction names $r_j$ associated with each reaction $aj$;
 \item 
-$\mathit{prods}\defeq \{p_1,...,p_u\}$ is the set of names for production $p_j$, associated with  reaction $aj$.
+$\mathit{prods}\defeq \{p_1,...,p_{|A|}\}$ is the set of names for production $p_j$, associated with  reaction $aj$.
 \end{itemize}
@@
 %$r_1 = \silent$,
-$r_{u+1} = \mathit{cxt}$ and $p_{u+1}=\silent$,  for $u$ the cardinality of $\mathit{reacts}$.
+$r_{u+1} = \mathit{cxt}$ and $p_{u+1}=\silent$,  for $u=|A|$.
 \end{definition}
@@
 Let $P=\llbracket{\cal A},\gamma\rrbracket$ its c\CNA~translation.
-If $exists$ $P'$ such that $t= (P \xrightarrow{\restrict{\mathit{names}}\upsilon} P')$ is a transition of $P$, then
+If $exists$ $P'$ such that $P \xrightarrow{\restrict{\mathit{names}}\upsilon} P'$ is a transition of $P$, then
   \begin{enumerate}

### 2020-04-05 06b-bisimulation.tex
small adjustments

@@
 
-\begin{definition}[Assertion language]
+\begin{definition}[Assertion Language]
 \label{def:assetionl}
@@
 
-\begin{definition}[Semantics of assertions]
+\begin{definition}[Semantics of Assertions]
  We define $\llbracket F\rrbracket \subseteq\wp(\mathit{syms}^*)$ by induction on the structure of $F$: 
@@
 
-\begin{definition}[Satisfaction as membership]\label{sec:semantics}
+\begin{definition}[Satisfaction as Membership]\label{sec:semantics}
 Let  $\upsilon$ be  a transition label, and $F$ be an assertion.
@@
 Now, we introduce a slightly modified version of the Hennessy Milner Logic~\cite{HM80}, called bioHML; due to the reasons we explained above, we do not want to look at the complete transition labels, thus we rely on our simple assertion language to make it parametric to the assertion $F$ of interest:
-\begin{definition}[bioHML]
+\begin{definition}[BioHML]
 Let   $F$ be an assertion, then 
@@
 
-\begin{definition}[semantics of $G$]
+\begin{definition}[Semantics of BioHML]
  We define $\llbracket G\rrbracket \subseteq {\cal P}$ by induction on the structure of $G$:
@@
 
-\begin{definition}[converse]
+\begin{definition}[Converse]
 Given a bioHML formula $G$ we define its converse $\overline{G}$ as follows:

### 2020-04-05 07-discussion.tex
small adjustments

@@
 mutated entities, and  the communication between two different Reaction Systems.
-It is important to note that our encoding guarantees that from each state, 
-in the  c\CNA \ transition system, only one transition comes out, 
+It is important to note that, when the context is deterministic, our encoding guarantees that from each state, 
+in the  c\CNA \ transition system, only one state is reachable, 
 as the dynamics is totally deterministic.
@@
 recognises the presence of $s1'$ and behaves in the same way as if $s1$ would be present.
-Technically, in both cases it is enough to add one more non 
-deterministic choice in the code of $P_{a1}$ and $P_{a2}$.
+Technically, in both cases it is enough to add one more nondeterministic choice in the code of $P_{a1}$ and $P_{a2}$.
 
@@
 \begin{array}{lcl}
-%\llbracket r1_1\rrbracket \defeq P_{r1_1}&\defeq &\startchain{\silent}\chainedlink{s1}{\noact}\chainedlink{\noact}{s1}\chainedlink{\tilde{s2}}{\noact}\chainedlink{\noact}{\tilde{s2}}\chainend{r1_2}.P_{r1_1} +
+%\llbracket r1_1\rrbracket \defeq P_{r1_1}&\defeq &\startchain{\silent}\chainedlink{s1}{\noact}\chainedlink{\noact}{s1}\chainedlink{\widetilde{s2}}{\noact}\chainedlink{\noact}{\widetilde{s2}}\chainend{r1_2}.P_{r1_1} +
 %\startchain{\silent}\chainedlink{\overline{s1}}{\noact}\chainedlink{\noact}{\overline{s1}}\chainend{r1_2}.P_{r1_1}\\
 \llbracket a_1\rrbracket \defeq P_{a_1}&\defeq &
-\startchain{\silent}\chainedlink{s_i}{{\color{blue}\noact}}\chainedlink{{\color{blue}\noact}}{s_o}\chainedlink{\tilde{x}_i}{{\color{gray}\noact}}\chainedlink{{\color{gray}\noact}}{\tilde{x}_o}\chainend{a_2}.P_{a_1} +
+\startchain{\silent}\chainedlink{s_i}{{\color{blue}\noact}}\chainedlink{{\color{blue}\noact}}{s_o}\chainedlink{\widetilde{x}_i}{{\color{gray}\noact}}\chainedlink{{\color{gray}\noact}}{\widetilde{x}_o}\chainend{a_2}.P_{a_1} +
 \startchain{\silent}\chainedlink{\overline{s}_i}{\noact}\chainedlink{\noact}{\overline{s}_o}    \chainedlink{dx_i}{\noact}\chainedlink{\noact}{dx_o}\chainend{a_2}.P_{a_1}\\
@@
 %\begin{array}{lcl}
-%\llbracket r1_2\rrbracket \defeq P_{r1_2}&\defeq &\startchain{r1_2}\chainedlink{\overline{s2}}{\noact}\chainedlink{\noact}{\overline{s2}}\chainedlink{\tilde{s1}}{\noact}\chainedlink{\noact}{\tilde{s1}}\chainend{r2_1}.P_{r2_1} +
+%\llbracket r1_2\rrbracket \defeq P_{r1_2}&\defeq &\startchain{r1_2}\chainedlink{\overline{s2}}{\noact}\chainedlink{\noact}{\overline{s2}}\chainedlink{\widetilde{s1}}{\noact}\chainedlink{\noact}{\widetilde{s1}}\chainend{r2_1}.P_{r2_1} +
 %\startchain{r1_2}\chainedlink{s2}{\noact}\chainedlink{\noact}{s2}\chainend{r2_1}.P_{r2_1}\\
@@
 %\begin{array}{lclclll}
-\llbracket x \rrbracket &\defeq & P_{x}& \defeq &\startchain{\tilde{x}_i}\chainend{\tilde{x}_o}.P_{x} +  \startchain{dx_i}\chainend{dx_o}.\overline{P_{x}}\\
+\llbracket x \rrbracket &\defeq & P_{x}& \defeq &\startchain{\widetilde{x}_i}\chainend{\widetilde{x}_o}.P_{x} +  \startchain{dx_i}\chainend{dx_o}.\overline{P_{x}}\\
 %\\
-%&&\overline{P_{s2}}&\defeq &\startchain{\overline{s2}}\chainedlink{\overline{s2}}{\noact}\chainedlink{\noact}{\tilde{s2}}\chainend{\tilde{s2}}.P_{s2} +
+%&&\overline{P_{s2}}&\defeq &\startchain{\overline{s2}}\chainedlink{\overline{s2}}{\noact}\chainedlink{\noact}{\widetilde{s2}}\chainend{\widetilde{s2}}.P_{s2} +
 %\startchain{\overline{s2}}\chainend{\overline{s2}}.\overline{P_{s2}}\\
@@
 %\\
-&& \overline{P_{x}}&\defeq &\startchain{{\color{gray}\tilde{x}_i}}\chainend{{\color{gray}\tilde{x}_o}}.P_{x} +  \startchain{dx_i}\chainend{dx_o}.\overline{P_{x}}\\
+&& \overline{P_{x}}&\defeq &\startchain{{\color{gray}\widetilde{x}_i}}\chainend{{\color{gray}\widetilde{x}_o}}.P_{x} +  \startchain{dx_i}\chainend{dx_o}.\overline{P_{x}}\\
 %\end{array}
@@
 %\begin{array}{lclclll}
-%\llbracket s2 \rrbracket &\defeq& P_{s2}&\defeq &\startchain{s2}\chainedlink{s2}{\noact}\chainedlink{\noact}{\tilde{s2}}\chainend{\tilde{s2}}.P_{s2} +  \startchain{s2}\chainend{s2}.\overline{P_{s2}}\\
+%\llbracket s2 \rrbracket &\defeq& P_{s2}&\defeq &\startchain{s2}\chainedlink{s2}{\noact}\chainedlink{\noact}{\widetilde{s2}}\chainend{\widetilde{s2}}.P_{s2} +  \startchain{s2}\chainend{s2}.\overline{P_{s2}}\\
 %%\\
-%&&\overline{P_{s2}}&\defeq &\startchain{\overline{s2}}\chainedlink{\overline{s2}}{\noact}\chainedlink{\noact}{\tilde{s2}}\chainend{\tilde{s2}}.P_{s2} +
+%&&\overline{P_{s2}}&\defeq &\startchain{\overline{s2}}\chainedlink{\overline{s2}}{\noact}\chainedlink{\noact}{\widetilde{s2}}\chainend{\widetilde{s2}}.P_{s2} +
 %\startchain{\overline{s2}}\chainend{\overline{s2}}.\overline{P_{s2}}\\
@@
 \begin{array}{lcl}
-\llbracket a_2\rrbracket \defeq P_{a_2}&\defeq &\startchain{a_2}\chainedlink{y_i}{{\color{gray}\noact}}\chainedlink{{\color{gray}\noact}}{y_o}\chainedlink{\tilde{ss}_i}{{\color{red}\noact}}\chainedlink{{\color{red}\noact}}{\tilde{ss}_o}\chainend{\silent}.P_{a_2}+
+\llbracket a_2\rrbracket \defeq P_{a_2}&\defeq &\startchain{a_2}\chainedlink{y_i}{{\color{gray}\noact}}\chainedlink{{\color{gray}\noact}}{y_o}\chainedlink{\widetilde{ss}_i}{{\color{red}\noact}}\chainedlink{{\color{red}\noact}}{\widetilde{ss}_o}\chainend{\silent}.P_{a_2}+
 \startchain{a_2}\chainedlink{\overline{y}_i}{\noact}\chainedlink{\noact}{\overline{y}_o}    
@@
 %\begin{array}{lcl}
-%\llbracket r2_2\rrbracket \defeq P_{r2_2}&\defeq &\startchain{r2_2}\chainedlink{ss1}{\noact}\chainedlink{\noact}{ss1}\chainedlink{\tilde{s3}}{\noact}\chainedlink{\noact}{\tilde{s3}}\chainend{\silent}.P_{r2_2}
+%\llbracket r2_2\rrbracket \defeq P_{r2_2}&\defeq &\startchain{r2_2}\chainedlink{ss1}{\noact}\chainedlink{\noact}{ss1}\chainedlink{\widetilde{s3}}{\noact}\chainedlink{\noact}{\widetilde{s3}}\chainend{\silent}.P_{r2_2}
 %+
@@
 \begin{array}{lclcl}
-\llbracket ss \rrbracket& \defeq& P_{ss}&\defeq &\startchain{\tilde{ss}_i}\chainedlink{\widehat{s}_i}{\noact}\chainedlink{\noact}{\widehat{s}_o}\chainend{\tilde{ss}_o}.P_{ss}+
+\llbracket ss \rrbracket& \defeq& P_{ss}&\defeq &\startchain{\widetilde{ss}_i}\chainedlink{\widehat{s}_i}{\noact}\chainedlink{\noact}{\widehat{s}_o}\chainend{\widetilde{ss}_o}.P_{ss}+
 \startchain{dss_i}   \chainend{dss_o}.\overline{P_{ss}} \\
@@
 %\begin{array}{lcllcl}
-&&\overline{P_{ss}}&\defeq &\startchain{{\color{red}\tilde{ss}_i}}\chainedlink{{\color{red}\widehat{s}_i}}{{\color{blue}\noact}}\chainedlink{{\color{blue}\noact}}{{\color{red}\widehat{s}_o}}\chainend{{\color{red}\tilde{ss}_o}}.P_{ss}
+&&\overline{P_{ss}}&\defeq &\startchain{{\color{red}\widetilde{ss}_i}}\chainedlink{{\color{red}\widehat{s}_i}}{{\color{blue}\noact}}\chainedlink{{\color{blue}\noact}}{{\color{red}\widehat{s}_o}}\chainend{{\color{red}\widetilde{ss}_o}}.P_{ss}
 +
@@
 &&\overline{P_{y}}&\defeq &
-%\startchain{\overline{y}_i}\chainedlink{\overline{y}_o}{\noact}\chainedlink{\noact}{\tilde{y}_i}\chainend{\tilde{y}_o}.P_{y}
+%\startchain{\overline{y}_i}\chainedlink{\overline{y}_o}{\noact}\chainedlink{\noact}{\widetilde{y}_i}\chainend{\widetilde{y}_o}.P_{y}
 %+ 
@@
 \[
-\startchain{\silent}\chainedlink{s_i}{{\color{blue}s_i}}\chainedlink{{\color{blue}s_o}}{s_o}\chainedlink{\tilde{x}_i}{{\color{gray}\tilde{x}_i}}\chainedlink{{\color{gray}\tilde{x}_o}}{\tilde{x}_o} \chainedlink{a_2}{a_2}\chainedlink{y_i}{ {\color{gray}y_i}}\chainedlink{{\color{gray}y_o}}{y_o}\chainedlink{\tilde{ss}_i}{{\color{red}\tilde{ss}_i}}
+\startchain{\silent}\chainedlink{s_i}{{\color{blue}s_i}}\chainedlink{{\color{blue}s_o}}{s_o}\chainedlink{\widetilde{x}_i}{{\color{gray}\widetilde{x}_i}}\chainedlink{{\color{gray}\widetilde{x}_o}}{\widetilde{x}_o} \chainedlink{a_2}{a_2}\chainedlink{y_i}{ {\color{gray}y_i}}\chainedlink{{\color{gray}y_o}}{y_o}\chainedlink{\widetilde{ss}_i}{{\color{red}\widetilde{ss}_i}}
 \chainedlink{{\color{red}\widehat{s}_i}}{{\color{blue}\widehat{s}_i}}
 \chainedlink{{\color{blue}\widehat{s}_o}}{{\color{red}\widehat{s}_o}}
-\chainedlink{{\color{red}\tilde{ss}_o}}{\tilde{ss}_o}\chainend{\silent},
+\chainedlink{{\color{red}\widetilde{ss}_o}}{\widetilde{ss}_o}\chainend{\silent},
 \]

### 2020-04-05 09-appendix.tex
small adjustments

@@
 Let $\mathcal{Q} = \{Q'_1,...,Q'_n\}$.
-By hypothesis all processes in $\mathcal{Q}$ must not be bio-logically equivalent to $P'$, hence for any $i\in[1,n]$ there exists a bioHML formula $G_i$ such that $P'\entails G_i$ and $Q'_i\not\entails G_i$ (if it was the opposite, $P'\not\entails G_i$ and $Q'_i\entails G_i$, we can use the converse formulas).
+By hypothesis all processes in $\mathcal{Q}$ must not be bio-logically equivalent to $P'$, hence for any $i\in[1,n]$ there exists a bioHML formula $G_i$ such that $P'\entails G_i$ and $Q'_i\not\entails G_i$ (if it was the opposite, $P'\not\entails H_i$ and $Q'_i\entails H_i$ for some $H_i$, we can use the converse formula $G_i\defeq \overline{H}_i$).
 But then the formula $G\defeq \langle F\rangle(G_1\wedge\dots\wedge G_n)$ is such that $P\entails G$ and $Q\not\entails G$, contradicting the assumption $P\equiv_{\mathcal{L}_F} Q$.
@@
 \item
-If $\upsilon \not\entails F$ then the proof is analogous to the previous case and thus omitted.
+If $\upsilon \not\entails F$ then the proof is analogous to the previous case (by exploiting $\neg F$) and thus omitted.
 \end{itemize}

### 2020-04-05 06b-bisimulation.tex
notazione formule
nella sezione 6.3 si confondeva l'entità G con la formula G. Ho usato \mathsf{} per tutte le formule.

@@
 \\
-F & ::= & 
+\mathsf{F} & ::= & 
 \epsilon ~\mid ~ 
 \zeta ~\mid ~ 
-F::F ~\mid ~ 
-F^+ ~\mid ~ 
-F^* ~\mid ~ 
-F \vee F ~\mid ~ 
-F \wedge F
+\mathsf{F}::\mathsf{F} ~\mid ~ 
+\mathsf{F}^+ ~\mid ~ 
+\mathsf{F}^* ~\mid ~ 
+\mathsf{F} \vee \mathsf{F} ~\mid ~ 
+\mathsf{F} \wedge \mathsf{F}
 \end{array}
@@
 
-An \emph{assertion} $F$ is either the empty string $\epsilon$, a singleton assertion $\zeta$, the concatenation of two assertions $F_1::F_2$, the replication of $F$ for 1 or more times $F^+$, the replication of $F$ for 0 or more times $F^*$, the disjunction of two assertions $F_1 \vee F_2$ or their conjunction $F_1 \wedge F_2$.
+An \emph{assertion} $\mathsf{F}$ is either the empty string $\epsilon$, a singleton assertion $\zeta$, the concatenation of two assertions $\mathsf{F}_1::\mathsf{F}_2$, the replication of $\mathsf{F}$ for 1 or more times $\mathsf{F}^+$, the replication of $\mathsf{F}$ for 0 or more times $\mathsf{F}^*$, the disjunction of two assertions $\mathsf{F}_1 \vee \mathsf{F}_2$ or their conjunction $\mathsf{F}_1 \wedge \mathsf{F}_2$.
 We write $\star$ as an abbreviation of $?^*$.
@@
 \begin{definition}[Semantics of Assertions]
- We define $\llbracket F\rrbracket \subseteq\wp(\mathit{syms}^*)$ by induction on the structure of $F$: 
+ We define $\llbracket \mathsf{F}\rrbracket \subseteq\wp(\mathit{syms}^*)$ by induction on the structure of $\mathsf{F}$: 
 \[
@@
  \llbracket \epsilon \rrbracket &\defeq&  \{\epsilon\}
- & \llbracket F_1::F_2 \rrbracket &\defeq& \{\omega_1 :: \omega_2~\mid~ \omega_1 \in \llbracket F_1 \rrbracket \wedge \omega_2 \in \llbracket F_2 \rrbracket\}  \\
+ & \llbracket \mathsf{F}_1::\mathsf{F}_2 \rrbracket &\defeq& \{\omega_1 :: \omega_2~\mid~ \omega_1 \in \llbracket \mathsf{F}_1 \rrbracket \wedge \omega_2 \in \llbracket \mathsf{F}_2 \rrbracket\}  \\
  \llbracket \alpha \rrbracket &\defeq&  \{\alpha\} 
- & \llbracket F^+ \rrbracket &\defeq& \llbracket F \rrbracket^+  \\
+ & \llbracket \mathsf{F}^+ \rrbracket &\defeq& \llbracket \mathsf{F} \rrbracket^+  \\
  \llbracket ? \rrbracket &\defeq&   \mathit{syms}
- & \llbracket F^* \rrbracket &\defeq& \llbracket F \rrbracket^*  \\
+ & \llbracket \mathsf{F}^* \rrbracket &\defeq& \llbracket \mathsf{F} \rrbracket^*  \\
  \llbracket [N]\rrbracket &\defeq&   N
- & \llbracket  F_1 \vee F_2 \rrbracket &\defeq& \llbracket  F_1\rrbracket \cup \llbracket  F_2\rrbracket  \\
- &&&\llbracket  F_1 \wedge F_2 \rrbracket &\defeq& \llbracket F_1 \rrbracket \cap \llbracket F_2\rrbracket 
+ & \llbracket  \mathsf{F}_1 \vee \mathsf{F}_2 \rrbracket &\defeq& \llbracket  \mathsf{F}_1\rrbracket \cup \llbracket  \mathsf{F}_2\rrbracket  \\
+ &&&\llbracket  \mathsf{F}_1 \wedge \mathsf{F}_2 \rrbracket &\defeq& \llbracket \mathsf{F}_1 \rrbracket \cap \llbracket \mathsf{F}_2\rrbracket 
 \end{array}
@@
 \begin{definition}[Satisfaction as Membership]\label{sec:semantics}
-Let  $\upsilon$ be  a transition label, and $F$ be an assertion.
-We write $\upsilon \entails F$ (read as the transition label $v$ satisfies the assertion $F$) 
-if $\mathit{flat}(\upsilon)\in \llbracket F\rrbracket$, otherwise we write $\upsilon \not\entails F$ (or also $\upsilon \entails \neg F$) and say that $F$ does not hold at $v$. 
+Let  $\upsilon$ be  a transition label, and $\mathsf{F}$ be an assertion.
+We write $\upsilon \entails \mathsf{F}$ (read as the transition label $v$ satisfies the assertion $\mathsf{F}$) 
+if $\mathit{flat}(\upsilon)\in \llbracket \mathsf{F}\rrbracket$, otherwise we write $\upsilon \not\entails \mathsf{F}$ (or also $\upsilon \entails \neg \mathsf{F}$) and say that $\mathsf{F}$ does not hold at $v$.
 \end{definition}
 
-Given two transition labels $v,w$ we write $v\equiv_F w$ if $v \entails F\ \Leftrightarrow\ w \entails F$, i.e. if both $v,w$ satisfy $F$ or they do not.
+Given two transition labels $v,w$ we write $v\equiv_{\mathsf{F}} w$ if $v \entails \mathsf{F}\ \Leftrightarrow\ w \entails \mathsf{F}$, i.e. if both $v,w$ satisfy $\mathsf{F}$ or they do not.
 
@@
 
-\begin{definition}[Bio-similarity  $ \sim_F $]
-Given an assertion $F$, a \emph{bio-simulation} $\mathbf{R}_F$ that respects $F$ is a binary relation over c\CNA \  processes such that, whenever $P~\mathrel{\mathbf{R}_F}~Q$ then:
+\begin{definition}[Bio-similarity  $ \sim_{\mathsf{F}}$]
+Given an assertion $\mathsf{F}$, a \emph{bio-simulation} $\mathbf{R}_{\mathsf{F}}$ that respects $\mathsf{F}$ is a binary relation over c\CNA \  processes such that, whenever $P~\mathrel{\mathbf{R}_{\mathsf{F}}}~Q$ then:
  \begin{itemize} 
@@
  for any $v,P'$ such that 
- $P \xRightarrow{v} P'$ then there exist $w,Q'$ such that $Q  \xRightarrow{w} Q'$ with $v\equiv_F w$ and $P' \mathrel{\mathbf{R}_F} Q'$.
+ $P \xRightarrow{v} P'$ then there exist $w,Q'$ such that $Q  \xRightarrow{w} Q'$ with $v\equiv_{\mathsf{F}} w$ and $P' \mathrel{\mathbf{R}_{\mathsf{F}}} Q'$.
 \item
  for any $w,Q'$ such that 
- $Q \xRightarrow{w} Q'$ then there exist $v,P'$ such that $P  \xRightarrow{v} P'$ with $v\equiv_F w$ and $P' \mathrel{\mathbf{R}_F} Q'$.
+ $Q \xRightarrow{w} Q'$ then there exist $v,P'$ such that $P  \xRightarrow{v} P'$ with $v\equiv_{\mathsf{F}} w$ and $P' \mathrel{\mathbf{R}_{\mathsf{F}}} Q'$.
 \end{itemize}
@@
 %\end{itemize}
-We let $ \sim_F $ denote the largest  bio-simulation and we say that $P$ is \emph{bio-similar} to $Q$, with respect to $F$, if $P \sim_F Q$.
+We let $ \sim_{\mathsf{F}} $ denote the largest  bio-simulation and we say that $P$ is \emph{bio-similar} to $Q$, with respect to $\mathsf{F}$, if $P \sim_{\mathsf{F}} Q$.
 \end{definition}
@@
 \begin{remark}
-An alternative way to look at a bio-simulation that respects $F$ is to define it as an ordinary bisimulation over the transition system labelled over $\{F,\neg F\}$ obtained by transforming each transition $P \xRightarrow{v} P'$ such that $v\entails F$ into $P \xrightarrow{F} P'$ and  each transition $P \xRightarrow{v} P'$ such that $v\not\entails F$ into $P \xrightarrow{\neg F} P'$.
+An alternative way to look at a bio-simulation that respects $\mathsf{F}$ is to define it as an ordinary bisimulation over the transition system labelled over $\{\mathsf{F},\neg \mathsf{F}\}$ obtained by transforming each transition $P \xRightarrow{v} P'$ such that $v\entails \mathsf{F}$ into $P \xrightarrow{\mathsf{F}} P'$ and  each transition $P \xRightarrow{v} P'$ such that $v\not\entails \mathsf{F}$ into $P \xrightarrow{\neg \mathsf{F}} P'$.
 \end{remark}
@@
 
-Now, we introduce a slightly modified version of the Hennessy Milner Logic~\cite{HM80}, called bioHML; due to the reasons we explained above, we do not want to look at the complete transition labels, thus we rely on our simple assertion language to make it parametric to the assertion $F$ of interest:
+Now, we introduce a slightly modified version of the Hennessy Milner Logic~\cite{HM80}, called bioHML; due to the reasons we explained above, we do not want to look at the complete transition labels, thus we rely on our simple assertion language to make it parametric to the assertion $\mathsf{F}$ of interest:
+
 \begin{definition}[BioHML]
-Let   $F$ be an assertion, then 
-the set of bioHML formulas $G$ that respects $F$ are built by the following syntax:
+Let   $\mathsf{F}$ be an assertion, then 
+the set of bioHML formulas $\mathsf{G}$ that respects $\mathsf{F}$ are built by the following syntax:
 $$
 \begin{array}{rcl}
-\chi & ::= & F~\mid~ \neg F\\
-G &::= &{\tt t} ~\mid~ {\tt f} ~\mid~ G\wedge G ~\mid~ G \vee G ~\mid~ \langle \chi \rangle G ~\mid~ [\chi]G
+\chi & ::= & \mathsf{F}~\mid~ \neg \mathsf{F}\\
+\mathsf{G},\mathsf{H} &::= &{\tt t} ~\mid~ {\tt f} ~\mid~ \mathsf{G}\wedge \mathsf{G} ~\mid~ \mathsf{G} \vee \mathsf{G} ~\mid~ \langle \chi \rangle \mathsf{G} ~\mid~ [\chi]\mathsf{G}
 \end{array}
@@
 \begin{remark}
-An alternative way to look at bioHML formulas is as ordinary HML formulas over the set of labels $\{F,\neg F\}$.
+An alternative way to look at bioHML formulas is as ordinary HML formulas over the set of labels $\{\mathsf{F},\neg \mathsf{F}\}$.
 \end{remark}
@@
 \begin{definition}[Semantics of BioHML]
- We define $\llbracket G\rrbracket \subseteq {\cal P}$ by induction on the structure of $G$:
+ We define $\llbracket \mathsf{G}\rrbracket \subseteq {\cal P}$ by induction on the structure of $\mathsf{G}$:
  %(read as the transition label $v$ satisfies the formula $F$) as: 
@@
 \begin{array}{ccl@{\hspace{1cm}}ccl}
- \llbracket {\tt t}\rrbracket &\defeq& {\cal P} &  \llbracket G\wedge H\rrbracket &\defeq&  \llbracket G\rrbracket  \cap \llbracket H\rrbracket  \\
+ \llbracket {\tt t}\rrbracket &\defeq& {\cal P} &  \llbracket \mathsf{G}\wedge \mathsf{H}\rrbracket &\defeq&  \llbracket \mathsf{G}\rrbracket  \cap \llbracket \mathsf{H}\rrbracket  \\
 \llbracket {\tt f}\rrbracket& \defeq& \emptyset &
-  \llbracket G\vee H\rrbracket &\defeq&  \llbracket G\rrbracket  \cup \llbracket H\rrbracket \\
+  \llbracket \mathsf{G}\vee \mathsf{H}\rrbracket &\defeq&  \llbracket \mathsf{G}\rrbracket  \cup \llbracket \mathsf{H}\rrbracket \\
 %\item   $ v \entails F_1 \vee F_2$ iff   $v \entails F_1$  $\vee$  $v \entails F_2$.
@@
 \begin{array}{lcl}
- \llbracket  \langle \chi \rangle G \rrbracket &\defeq& \{P \in {\cal P}: \exists v,P'.\ P\xRightarrow{v}P' \mbox{ with } v \entails \chi \mbox{ and } P' \in  \llbracket  G \rrbracket\} \\
- \llbracket [\chi]G \rrbracket &\defeq& \{P \in {\cal P}: \forall v,P'.\ P\xRightarrow{v}P' \mbox{ implies } v \entails \chi \mbox{ and } P' \in  \llbracket  G \rrbracket\}
+ \llbracket  \langle \chi \rangle \mathsf{G} \rrbracket &\defeq& \{P \in {\cal P}: \exists v,P'.\ P\xRightarrow{v}P' \mbox{ with } v \entails \chi \mbox{ and } P' \in  \llbracket  \mathsf{G} \rrbracket\} \\
+ \llbracket [\chi]\mathsf{G} \rrbracket &\defeq& \{P \in {\cal P}: \forall v,P'.\ P\xRightarrow{v}P' \mbox{ implies } v \entails \chi \mbox{ and } P' \in  \llbracket  \mathsf{G} \rrbracket\}
 \end{array}
 \]
-We write $P \entails G$ ($P$ satisfies $G$) if and only if $P \in  \llbracket G \rrbracket$.
+We write $P \entails \mathsf{G}$ ($P$ satisfies $\mathsf{G}$) if and only if $P \in  \llbracket \mathsf{G} \rrbracket$.
 \end{definition}
 
-Negation is not included in the syntax, but the converse $\overline{G}$ of a bioHML formula $G$ can be easily defined inductively in the same way as for HML logic.
+Negation is not included in the syntax, but the converse $\overline{\mathsf{G}}$ of a bioHML formula $\mathsf{G}$ can be easily defined inductively in the same way as for HML logic.
 
 \begin{definition}[Converse]
-Given a bioHML formula $G$ we define its converse $\overline{G}$ as follows:
+Given a bioHML formula $\mathsf{G}$ we define its converse $\overline{\mathsf{G}}$ as follows:
 \[
@@
 \overline{\tt t} & \defeq & {\tt f} &
-\overline{G \wedge H} & \defeq & \overline{G} \vee \overline{H} &
-\overline{\langle \chi \rangle G} & \defeq & [\chi]\overline{G}
+\overline{\mathsf{G} \wedge \mathsf{H}} & \defeq & \overline{\mathsf{G}} \vee \overline{\mathsf{H}} &
+\overline{\langle \chi \rangle \mathsf{G}} & \defeq & [\chi]\overline{\mathsf{G}}
 \\
 \overline{\tt f} & \defeq & {\tt t} &
-\overline{G \vee H} & \defeq & \overline{G} \wedge \overline{H} &
-\overline{[\chi]G} & \defeq & \langle \chi \rangle\overline{G}
+\overline{\mathsf{G} \vee \mathsf{H}} & \defeq & \overline{\mathsf{G}} \wedge \overline{\mathsf{H}} &
+\overline{[\chi]\mathsf{G}} & \defeq & \langle \chi \rangle\overline{\mathsf{G}}
 \end{array}
@@
 
-We observe that, as expected, for any bioHML formula $G$ and process $P$ we have $\overline{\overline{G}} = G$ and  $P \entails \overline{G}$ iff  $P\not\entails G$.
+We observe that, as expected, for any bioHML formula $G$ and process $P$ we have $\overline{\overline{\mathsf{G}}} = \mathsf{G}$ and  $P \entails \overline{\mathsf{G}}$ iff  $P\not\entails \mathsf{G}$.
 
-We let $\mathcal{L}_F$ be the set of all bioHML formulas that respects $F$ and 
+We let $\mathcal{L}_{\mathsf{F}}$ be the set of all bioHML formulas that respects $\mathsf{F}$ and 
 we say that two processes $P,Q$ are bio-logically equivalent w.r.t.
-$F$, written $P \equiv_{\mathcal{L}_F}Q$, when $P$ and $Q$ satisfy the exactly the same bioHML formulas in $\mathcal{L}_F$, i.e.
-when for any $G\in \mathcal{L}_F$ we have $P \entails G\ \Leftrightarrow Q \entails G$.
+$\mathsf{F}$, written $P \equiv_{\mathcal{L}_{\mathsf{F}}}Q$, when $P$ and $Q$ satisfy the exactly the same bioHML formulas in $\mathcal{L}_{\mathsf{F}}$, i.e.
+when for any $\mathsf{G}\in \mathcal{L}_{\mathsf{F}}$ we have $P \entails \mathsf{G}\ \Leftrightarrow Q \entails \mathsf{G}$.
 
@@
 %\begin{theorem}
-$\sim_F\ =\ \equiv_{\mathcal{L}_F}$
+$\sim_{\mathsf{F}}\ =\ \equiv_{\mathcal{L}_{\mathsf{F}}}$
 %\end{theorem}

### 2020-04-05 06-examples.tex
small adjustments
modificati 7 files
modificati 7 files

@@
 \]
-In fact the presence/absence of $q$ and $w$ will be exploited by at least one rule and as most three rules.
+In fact the presence/absence of $q$ and $w$ will be exploited by at least one rule and at most three rules.
 
@@
  \chainedlink{q_i}{\bf q_i}\chainedlink{\bf q_o}{q_o}
- \chainedlink{b_i}{\bf b_i}\chainedlink{\bf b_o}{b_io}
+ \chainedlink{b_i}{\bf b_i}\chainedlink{\bf b_o}{b_o}
  \chainedlink{\overline{w}_i}{\bf \overline{w}_i}\chainedlink{\bf \overline{w}_o}{\overline{w}_o}
@@
 
-\subsection{A  biological toy example}\label{subsec:toy}
+\subsection{A  biological toy example of gene expression}\label{subsec:toy}
 
 \noindent
-We consider a biological toy  example in the style of gene's alternative splicing \cite{WBB13}, 
+We consider a biological toy  example in the style of gene's alternative splicing \cite{WBB13}.
+Alternative splicing is a regulated process during gene expression that results in a single gene 
+coding for multiple proteins. 
+In practice, particular exons of a gene may be included within or excluded from the final, 
+processed messenger RNA (mRNA) produced from that gene.
+
 where a gene $a$  codes for a protein $T$ when molecules 

### 2020-04-05 06b-bisimulation.tex
modificati 7 files
modificati 7 files

--- 2020-04-05 06b-bisimulation.tex
@@
 
-The classical notion of bisimulation for process algebra equates two processes
+The classical notion of bisimulation for process algebras equates two processes
 when one process can simulate all the instructions executed by the other one and viceversa.

### 2020-04-05 07-discussion.tex
widetilde

--- 2020-04-04 07-discussion.tex
@@
 \llbracket a_1\rrbracket \defeq P_{a_1}&\defeq &
-\startchain{\silent}\chainedlink{s_i}{{\color{blue}\noact}}\chainedlink{{\color{blue}\noact}}{s_o}\chainedlink{\tilde{x}_i}{{\color{gray}\noact}}\chainedlink{{\color{gray}\noact}}{\tilde{x}_o}\chainend{a_2}.P_{a_1} +
 \startchain{\silent}\chainedlink{\overline{s}_i}{\noact}\chainedlink{\noact}{\overline{s}_o}    \chainedlink{dx_i}{\noact}\chainedlink{\noact}{dx_o}\chainend{a_2}.P_{a_1}\\
@@
 %\begin{array}{lclclll}
-\llbracket x \rrbracket &\defeq & P_{x}& \defeq &\startchain{\tilde{x}_i}\chainend{\tilde{x}_o}.P_{x} +  \startchain{dx_i}\chainend{dx_o}.\overline{P_{x}}\\
 %\\
@@
 %\\
-&& \overline{P_{x}}&\defeq &\startchain{{\color{gray}\tilde{x}_i}}\chainend{{\color{gray}\tilde{x}_o}}.P_{x} +  \startchain{dx_i}\chainend{dx_o}.\overline{P_{x}}\\
 %\end{array}
@@
 \begin{array}{lcl}
-\llbracket a_2\rrbracket \defeq P_{a_2}&\defeq &\startchain{a_2}\chainedlink{y_i}{{\color{gray}\noact}}\chainedlink{{\color{gray}\noact}}{y_o}\chainedlink{\tilde{ss}_i}{{\color{red}\noact}}\chainedlink{{\color{red}\noact}}{\tilde{ss}_o}\chainend{\silent}.P_{a_2}+
 \startchain{a_2}\chainedlink{\overline{y}_i}{\noact}\chainedlink{\noact}{\overline{y}_o}    
@@
 \begin{array}{lclcl}
-\llbracket ss \rrbracket& \defeq& P_{ss}&\defeq &\startchain{\tilde{ss}_i}\chainedlink{\widehat{s}_i}{\noact}\chainedlink{\noact}{\widehat{s}_o}\chainend{\tilde{ss}_o}.P_{ss}+
 \startchain{dss_i}   \chainend{dss_o}.\overline{P_{ss}} \\
@@
 %\begin{array}{lcllcl}
-&&\overline{P_{ss}}&\defeq &\startchain{{\color{red}\tilde{ss}_i}}\chainedlink{{\color{red}\widehat{s}_i}}{{\color{blue}\noact}}\chainedlink{{\color{blue}\noact}}{{\color{red}\widehat{s}_o}}\chainend{{\color{red}\tilde{ss}_o}}.P_{ss}
 +
@@
 \[
-\startchain{\silent}\chainedlink{s_i}{{\color{blue}s_i}}\chainedlink{{\color{blue}s_o}}{s_o}\chainedlink{\tilde{x}_i}{{\color{gray}\tilde{x}_i}}\chainedlink{{\color{gray}\tilde{x}_o}}{\tilde{x}_o} \chainedlink{a_2}{a_2}\chainedlink{y_i}{ {\color{gray}y_i}}\chainedlink{{\color{gray}y_o}}{y_o}\chainedlink{\tilde{ss}_i}{{\color{red}\tilde{ss}_i}}
 \chainedlink{{\color{red}\widehat{s}_i}}{{\color{blue}\widehat{s}_i}}
 \chainedlink{{\color{blue}\widehat{s}_o}}{{\color{red}\widehat{s}_o}}
-\chainedlink{{\color{red}\tilde{ss}_o}}{\tilde{ss}_o}\chainend{\silent},
 \]

### 2020-04-05 07-discussion.tex
widetilde

@@
 \llbracket a_1\rrbracket \defeq P_{a_1}&\defeq &
+\startchain{\silent}\chainedlink{s_i}{{\color{blue}\noact}}\chainedlink{{\color{blue}\noact}}{s_o}\chainedlink{\widetilde{x}_i}{{\color{gray}\noact}}\chainedlink{{\color{gray}\noact}}{\widetilde{x}_o}\chainend{a_2}.P_{a_1} +
 \startchain{\silent}\chainedlink{\overline{s}_i}{\noact}\chainedlink{\noact}{\overline{s}_o}    \chainedlink{dx_i}{\noact}\chainedlink{\noact}{dx_o}\chainend{a_2}.P_{a_1}\\
@@
 %\begin{array}{lclclll}
+\llbracket x \rrbracket &\defeq & P_{x}& \defeq &\startchain{\widetilde{x}_i}\chainend{\widetilde{x}_o}.P_{x} +  \startchain{dx_i}\chainend{dx_o}.\overline{P_{x}}\\
 %\\
@@
 %\\
+&& \overline{P_{x}}&\defeq &\startchain{{\color{gray}\widetilde{x}_i}}\chainend{{\color{gray}\widetilde{x}_o}}.P_{x} +  \startchain{dx_i}\chainend{dx_o}.\overline{P_{x}}\\
 %\end{array}
@@
 \begin{array}{lcl}
+\llbracket a_2\rrbracket \defeq P_{a_2}&\defeq &\startchain{a_2}\chainedlink{y_i}{{\color{gray}\noact}}\chainedlink{{\color{gray}\noact}}{y_o}\chainedlink{\widetilde{ss}_i}{{\color{red}\noact}}\chainedlink{{\color{red}\noact}}{\widetilde{ss}_o}\chainend{\silent}.P_{a_2}+
 \startchain{a_2}\chainedlink{\overline{y}_i}{\noact}\chainedlink{\noact}{\overline{y}_o}    
@@
 \begin{array}{lclcl}
+\llbracket ss \rrbracket& \defeq& P_{ss}&\defeq &\startchain{\widetilde{ss}_i}\chainedlink{\widehat{s}_i}{\noact}\chainedlink{\noact}{\widehat{s}_o}\chainend{\widetilde{ss}_o}.P_{ss}+
 \startchain{dss_i}   \chainend{dss_o}.\overline{P_{ss}} \\
@@
 %\begin{array}{lcllcl}
+&&\overline{P_{ss}}&\defeq &\startchain{{\color{red}\widetilde{ss}_i}}\chainedlink{{\color{red}\widehat{s}_i}}{{\color{blue}\noact}}\chainedlink{{\color{blue}\noact}}{{\color{red}\widehat{s}_o}}\chainend{{\color{red}\widetilde{ss}_o}}.P_{ss}
 +
@@
 \[
+\startchain{\silent}\chainedlink{s_i}{{\color{blue}s_i}}\chainedlink{{\color{blue}s_o}}{s_o}\chainedlink{\widetilde{x}_i}{{\color{gray}\widetilde{x}_i}}\chainedlink{{\color{gray}\widetilde{x}_o}}{\widetilde{x}_o} \chainedlink{a_2}{a_2}\chainedlink{y_i}{ {\color{gray}y_i}}\chainedlink{{\color{gray}y_o}}{y_o}\chainedlink{\widetilde{ss}_i}{{\color{red}\widetilde{ss}_i}}
 \chainedlink{{\color{red}\widehat{s}_i}}{{\color{blue}\widehat{s}_i}}
 \chainedlink{{\color{blue}\widehat{s}_o}}{{\color{red}\widehat{s}_o}}
+\chainedlink{{\color{red}\widetilde{ss}_o}}{\widetilde{ss}_o}\chainend{\silent},
 \]

### 2020-04-05 05-translation.tex

@@
 
-In the following use the following conventions for denoting different categories of names:
+In the following we use the following conventions for denoting different categories of names:
 \begin{itemize}
@@
 We could say that these processes form the \emph{backbone} of the interaction.
-   In contrast, the processes encoding the entities, $P_{s1}$,  $\overline{P}_{s2}$, provides the
+   In contrast, the processes encoding the entities, $P_{s1}$,  $\overline{P}_{s2}$, provide the
     solid links to be merged with the virtual links of the backbone (i.e. to be plugged in the backbone).
@@
 
-Viceversa, it is immediate to define the function $\mathit{unflat}$ to rebuild the complete label from the compact string (here we exploit again the half link and block notation):
+It is then immediate to define the function $\mathit{unflat}$ to rebuild the complete label from the compact string (here we exploit again the half link and block notation):
 $$
@@
 With the next proposition, we analyse the structure of a c\CNA \ 
-process encoding of  a reactive process after one step transition.
+process encoding of  a reactive process after one transition step.
 In the following four statements, for brevity, we let ${\cal A} = (S, A)$ be a RS, and let  $\pi=(\gamma,\delta)$ be an extended interactive process in $A$, with $\gamma=\{C_i\}_{i\in\mathbb{N}}$ and $\delta=\{D_i\}_{i\in\mathbb{N}}$.

### 2020-04-05 08-conclusion.tex
small corrections
prova cambio
cambio

@@
 This translation shows several benefits.
-{\color{red} For instance, in our paper the context works non deterministically and it is recursively defined.}
+For instance, in our paper the context works non deterministically and it 
+is recursively defined.
 %  the context behaviour is expressed recursively and no be expressed recursively. 
@@
 to our framework.
+We believe that our work can make 
+possible to investigate how to integrate our methodology
+with other formal techniques to prove 
+properties of the modeled systems \cite{CFHOT15,OCHF16,BBGLBH2017}.
 
@@
 of RSs towards a model which can improve the communication
-between different RSs. We also believe that our work can make 
-possible to investigate how to apply formal techniques to prove 
-properties of the
-modeled systems \cite{CFHOT15,OCHF16,BBGLBH2017}.
+between different RSs.

### 2020-04-05 09-appendix.tex

--- 2020-04-03 09-appendix.tex
@@
  \end{proof}
+
+ \theocorr*
+ \begin{proof}
+The proof is just an adaptation of the classical result. The two implications are proved separately.
+\begin{description}
+\item[$\sim_{\mathsf{F}}\ \subseteq\ \equiv_{\mathcal{L}_{\mathsf{F}}}$:]
+Given any two processes $P\sim_{\mathsf{F}} Q$ we need to prove that for any bioHML formula $\mathsf{G}$ we have $P\entails \mathsf{G}$ iff $Q\entails \mathsf{G}$. 
+Without loss of generality, we prove that $P\entails \mathsf{G}$ implies $Q\entails \mathsf{G}$.
+The proof is by structural induction on $\mathsf{G}$.
+\begin{itemize}
+\item 
+if $\mathsf{G}={\tt t}$, then $Q\entails \mathsf{G}$.
+
+\item
+if $\mathsf{G}={\tt f}$, then the assumption $P\entails \mathsf{G}$ is false and the implication holds.
+
+\item
+if $\mathsf{G}=\mathsf{G}_1 \wedge \mathsf{G}_2$ we take as inductive hypotheses that
+$$
+\begin{array}{l}
+\forall R,S.\ R \sim_{\mathsf{F}} S \wedge R\entails \mathsf{G}_1 \Rightarrow S \entails \mathsf{G}_1\\
+\forall R,S.\ R \sim_{\mathsf{F}} S \wedge R\entails \mathsf{G}_2 \Rightarrow S \entails \mathsf{G}_2
+\end{array}
+$$
+We need to prove that $Q\entails \mathsf{G}$.
+Since $P\entails \mathsf{G}=\mathsf{G}_1 \wedge \mathsf{G}_2$ we have $P\entails \mathsf{G}_1$ and $P\entails \mathsf{G}_2$.
+Since $P\sim_{\mathsf{F}} Q$, by inductive hypotheses we get $Q\entails \mathsf{G}_1$ and $Q\entails \mathsf{G}_2$.
+Hence $Q\entails \mathsf{G}_1 \wedge \mathsf{G}_2=\mathsf{G}$.
+
+\item
+if $\mathsf{G}=\mathsf{G}_1 \vee \mathsf{G}_2$  we take as inductive hypotheses that
+$$
+\begin{array}{l}
+\forall R,S.\ R \sim_{\mathsf{F}} S \wedge R\entails \mathsf{G}_1 \Rightarrow S \entails \mathsf{G}_1\\
+\forall R,S.\ R \sim_{\mathsf{F}} S \wedge R\entails \mathsf{G}_2 \Rightarrow S \entails \mathsf{G}_2
+\end{array}
+$$
+We need to prove that $Q\entails \mathsf{G}$.
+Since $P\entails \mathsf{G}=\mathsf{G}_1 \vee \mathsf{G}_2$ we have $P\entails \mathsf{G}_1$ or $P\entails \mathsf{G}_2$.
+If $P\entails \mathsf{G}_1$, since $P\sim_{\mathsf{F}} Q$, by inductive hypotheses we get $Q\entails \mathsf{G}_1$ and thus $Q\entails \mathsf{G}_1 \vee \mathsf{G}_2=\mathsf{G}$.
+If $P\entails \mathsf{G}_2$, since $P\sim_{\mathsf{F}} Q$, by inductive hypotheses we get $Q\entails \mathsf{G}_2$ and thus $Q\entails \mathsf{G}_1 \vee \mathsf{G}_2=\mathsf{G}$.
+
+\item
+if $\mathsf{G}=\langle\chi\rangle \mathsf{H}$ we take as inductive hypothesis that
+$$
+\forall R,S.\ R \sim_{\mathsf{F}} S \wedge R\entails \mathsf{H} \Rightarrow S \entails \mathsf{H}
+$$
+We need to prove that $Q\entails \mathsf{G}$.
+Since $P\entails \langle\chi\rangle \mathsf{H}$ it means that there exists $\upsilon, P'$ such that $P\xRightarrow{\upsilon} P'$ with $\upsilon\entails \chi$ and $P'\entails \mathsf{H}$.
+Since $P\sim_{\mathsf{F}} Q$, there exists $w,Q'$ such that $Q\xRightarrow{w} Q'$ with $w\entails \chi$ and $P'\sim_{\mathsf{F}} Q'$.
+Then, by inductive hypothesis, $Q'\entails \mathsf{H}$ and thus $Q\entails \langle\chi\rangle \mathsf{H}=\mathsf{G}$.
+
+\item
+if $\mathsf{G}=[\chi] \mathsf{H}$ we take as inductive hypothesis that
+$$
+\forall R,S.\ R \sim_{\mathsf{F}} S \wedge R\entails \mathsf{H} \Rightarrow S \entails \mathsf{H}
+$$
+We need to prove that $Q\entails \mathsf{G}$.
+If there is no $\upsilon\entails \chi$ such that $Q\xRightarrow{\upsilon}Q'$ for some $Q'$, then 
+ $Q\entails [\chi] \mathsf{H}=\mathsf{G}$ trivially.
+For any $\upsilon,Q'$ such that $Q\xRightarrow{\upsilon}Q'$ with $\upsilon\entails \chi$, then as $P\sim_{\mathsf{F}} Q$ there must exist $w,P'$ such that $P\xRightarrow{w}P'$ with $w\entails \chi$ and $P'\sim_{\mathsf{F}} Q'$.
+Since $P\entails \mathsf{G}=[\chi] \mathsf{H}$ then it must be $P'\entails \mathsf{H}$.
+Since $P'\sim_{\mathsf{F}} Q'$, by inductive hypothesis $Q'\entails \mathsf{H}$.
+Hence $Q\entails [\chi] \mathsf{H} = \mathsf{G}$.
+\end{itemize}
+
+\item[$\equiv_{\mathcal{L}_{\mathsf{F}}}\ \subseteq\ \sim_{\mathsf{F}}$:]
+We prove that $\equiv_{\mathcal{L}_{\mathsf{F}}}$ is a bio-simulation and thus included in $\sim_{\mathsf{F}}$.
+Take two generic processes $P\equiv_{\mathcal{L}_{\mathsf{F}}} Q$ and suppose $P\xRightarrow{\upsilon}P'$ for some $\upsilon,P'$. 
+\begin{itemize}
+\item
+If $\upsilon \entails \mathsf{F}$ we want to prove that there exists some $w,Q'$ such that $Q\xRightarrow{w}Q'$, with $w\entails \mathsf{F}$ and $P'\equiv_{\mathcal{L}_{\mathsf{F}}} Q'$.
+
+Towards a contradiction, assume that we cannot find such $w,Q'$.
+
+If there is no transition $Q\xRightarrow{w}Q'$ such that $w\entails \mathsf{F}$, then the bioHML formula $\mathsf{G} \defeq \langle \mathsf{F}\rangle{\tt t}$ is such that $P\entails \mathsf{G}$ and $Q\not\entails \mathsf{G}$, contradicting the assumption $P\equiv_{\mathcal{L}_{\mathsf{F}}} Q$.
+
+Otherwise, let $\mathcal{Q}\defeq \{Q'~\mid~\exists w.\ Q\xRightarrow{w}Q' \ \wedge\ w\entails \mathsf{F}\}$ be the (non-empty) set of processes reachable from $Q$ via a transition with a (complete) label that satisfies $\mathsf{F}$.
+Since our processes are with guarded recursion, the set $\mathcal{Q}$ is finite.
+Let $\mathcal{Q} = \{Q'_1,...,Q'_n\}$.
+By hypothesis all processes in $\mathcal{Q}$ must not be bio-logically equivalent to $P'$, hence for any $i\in[1,n]$ there exists a bioHML formula $\mathsf{G}_i$ such that $P'\entails \mathsf{G}_i$ and $Q'_i\not\entails \mathsf{G}_i$ (if it was the opposite, $P'\not\entails \mathsf{H}_i$ and $Q'_i\entails \mathsf{H}_i$ for some $\mathsf{H}_i$, we can use the converse formula $\mathsf{G}_i\defeq \overline{\mathsf{H}}_i$).
+But then the formula $\mathsf{G}\defeq \langle \mathsf{F}\rangle(\mathsf{G}_1\wedge\dots\wedge \mathsf{G}_n)$ is such that $P\entails \mathsf{G}$ and $Q\not\entails \mathsf{G}$, contradicting the assumption $P\equiv_{\mathcal{L}_{\mathsf{F}}} Q$.
+
+\item
+If $\upsilon \not\entails \mathsf{F}$ then the proof is analogous to the previous case (by exploiting $\neg \mathsf{F}$) and thus omitted.
+\end{itemize}
+\end{description}
+\end{proof}

### 2020-04-05 06-examples.tex
piccoli cambi

@@
 					        \chainedlink{\widetilde{q}_i}{\noact}\chainedlink{\noact}{\widetilde{q}_o}
-			\chainend{p_{n+1}}\\
+			\chainend{p_{n+1}}\\[6pt]
 \pi'_n(x) &\defeq &  \startchain{r_n}\chainedlink{x_i}{\noact}\chainedlink{\noact}{x_o}\chainedlink{r_{n+1}}{\noact}\chainedlink{\noact}{p_n} \chainend{p_{n+1}} 
@@
 					        \chainedlink{\widetilde{T}_i}{\noact}\chainedlink{\noact}{\widetilde{T}_o}
-			\chainend{p_{n+1}}\\
+			\chainend{p_{n+1}}\\[6pt]
 \pi'_n(x) &\defeq &  \startchain{r_n}\chainedlink{x_i}{\noact}\chainedlink{\noact}{x_o}\chainedlink{r_{n+1}}{\noact}\chainedlink{\noact}{p_n} \chainend{p_{n+1}} 
@@
 \]
-Here, the entities that can be produced by the reactions and that can be  provided by the context:
+The entities that can be produced by the reactions and that can be  provided by the context are as follows:
 \[
@@
 \begin{array}{lcl @{\hspace{1cm}}lcl@{\hspace{0.4cm}}lcl@{\hspace{0.4cm}}lcl}
-P_C & \defeq &  P(C,\widehat{C},\underline{C},\widetilde{C})&  P_G& \defeq &   P(G,\widehat{G},\underline{G},\widetilde{G}) \\
+P_C & \defeq &  P(C,\widehat{C},\underline{C},\widetilde{C})&  P_G& \defeq &   P(G,\widehat{G},\underline{G},\widetilde{G}) \\[6pt]
 \overline{P}_C & \defeq &  P(\overline{C},\widehat{C},\underline{C},\widetilde{C}) & 
@@
 
-Here, the encoding of the entity $F$ that can be provided by the context:
+The encoding of the entity $F$ that can be provided by the context follows:
 \[
@@
 P_F&\defeq & \sum_{h=0}^1 (\startchain{F_i}\chainedlink{F_o}{\noact}\chainend{\noact})^h\ \link{\widehat{F}_i}{\widehat{F}_o}.P_F
-& +  && \sum_{h=0}^1 (\startchain{F_i}\chainedlink{F_o}{\noact}\chainend{\noact})^h. \link{\underline{F}_i}{\underline{F}_o}.\overline{P}_F \\
+& +  && \sum_{h=0}^1 (\startchain{F_i}\chainedlink{F_o}{\noact}\chainend{\noact})^h. \link{\underline{F}_i}{\underline{F}_o}.\overline{P}_F \\[6pt]
 \overline{P}_F&\defeq & \sum_{h=0}^1 (\startchain{\overline{F}_i}\chainedlink{\overline{F}_o}{\noact}\chainend{\noact})^h\ \link{\widehat{F}_i}{\widehat{F}_o}.P_F

### 2020-04-05 08-conclusion.tex
piccoli cambi

@@
 sciences.
-As we have already mentioned, the evolution of each process resulting from our embedding 
-is deterministic, thus we do not have the problem of having infinitely many transitions
-in the produced labelled transition system. In any case, we can exploit the implementation of  the symbolic semantics of the {\tt link}-calculus~\cite{BrodoO17} that can be found in~\cite{tool}.
 
@@
 with an automatic translation from RSs to  {\tt link}-calculus. 
+We can also exploit the implementation of  the symbolic semantics of 
+the {\tt link}-calculus~\cite{BrodoO17} that can be found in~\cite{tool}.
+
 We believe that our work can also help to extend the framework

### 2020-04-05 08-conclusion.tex
ultime modifiche

@@
 
-In this paper we have introduced a variant of the {\tt link}-calculus
+In this paper we have introduced c\CNA, a variant of the {\tt link}-calculus
 where prefixes are link chains and no more single links, as it was 
@@
 This translation shows several benefits.
-For instance, in our paper the context works non deterministically and it 
+For instance, in our paper the context of a RS can work non deterministically and it 
 is recursively defined.
@@
 
-We believe that our embedding can contribute to extend the applications
+We are confident that our embedding can contribute to extend the applications
 of Reaction Systems to diverse fields of computer science, and life
@@
 
-We believe that our work can also help to extend the framework
-of RSs towards a model which can improve the communication
-between different RSs.
+We also plan to build on the work in this paper to extend the framework
+of RSs towards a model where different RSs are allowed to communicate.

### 2020-04-05 00-abstract.tex
piccoli cambi e sottomissione
cambi e sottomissione a TCS

@@
 
-%Natural computing is an emerging area of research for modeling biological 
-%computations by means of computer science formal methods. Reaction systems 
-%have been defined in this area as a qualitative abstraction inspired by the 
-%functioning of living cells, suitable to model the main
-%mechanisms of biochemical reactions.
-%Natural computing is an emerging area of research for modeling biological 
-%computations by means of computer science formal methods. 
 In the area of Natural Computing, reaction systems are
@@
 systems interact with the environment represented by the context, and
-pose problems of implementation, as it is a new computation model. 
+pose challenges for further extensions as well as
+for the implementation, as it is a new computation model. 
 In this paper we consider a variant of the {\tt link}-calculus, which allows to model 
@@
 and biological applications. 
-%how to embed a \emph{lac} operon 
-%regulatory network. 
 On top of the LTS semantics for reaction systems provided by our framework, 
@@
 different reaction systems.
-%entities and the context.

### 2020-06-18 TCS_version/01-introduction.tex
reviewed directory

--- 2020-04-04 01-introduction.tex
@@
 process algebras allows to consider two processes as equivalent 
-where one process can simulate all the actions executed 
+when one process can simulate all the actions executed 
 by the other one and vice versa. In this paper we define a new
@@
 formalism for specifying properties of labeled transition systems, 
-its variant introduced here looks a suitable option to specify formulas 
+its variant introduced here is a suitable option to specify formulas 
 over complex labels by abstracting from unnecessary details.
@@
 \begin{itemize}
-\item here we present several new examples to illustrate our framework,
+\item we present several new examples to illustrate our framework,
 and give more detailed explanations about its use;

### 2020-06-18 TCS_version/02-ReactionSystem.tex
reviewed directory

--- 2020-04-01 02-ReactionSystem.tex
@@
 
-\begin{definition}[extended interactive process]
+\begin{definition}[Extended Interactive Process]
 Let ${\cal A}=(S,A)$ be a RS, and let $\pi =(\gamma,\delta)$  be an  $n$-step interactive process,
@@
 C_j & \mbox{if $j\in[0,n]$}\\
-\emptyset & \mbox{otherwise ($j>n$)}
+\emptyset & \mbox{if $j>n$}
 \end{array}
@@
 D_0 & \mbox{if $j=0$}\\
-\mathit{res}_A(D_{j-1}' \cup C_{j-1}') & \mbox{otherwise ($j\geq 1$)}
+\mathit{res}_A(D_{j-1}' \cup C_{j-1}') & \mbox{if ($j\geq 1$)}
 \end{array}

### 2020-06-18 TCS_version/05-translation.tex
reviewed directory

--- 2020-04-05 05-translation.tex
@@
-% !TEX root = ./main_RS2L_LNCS.tex
+% !TEX root = ./main_TCS.tex
 
@@
 \paragraph{Processes for entities}
-Given an entity $s \in S$, we exploit four different names for the interactions over $s$:
+Given an entity $s \in S$, we exploit five different pairs of names for the interactions over $s$:
 names $s_i$, $s_o$ are used to test the presence of $s$ in the system; names $\widehat{s}_i$, $\widehat{s}_o$ are used to test the provisioning of $s$ from the context;
@@
 \]
+
 The first line of $ E(s,\widetilde{s},\widehat{s},\underline{s})$ accounts for the case where $s$ is tested for presence by $h$ reactions and produced by $k$ reactions, while being provided by the context ($\link{\widehat{s}_i}{\widehat{s}_o}$).
 Thus, $s$ will be present at the next step (the continuation is $P_s$). Here $h$ and $k$ are not known a priori and therefore any
-combination is possible. By knowing the number of reactions that test $s$, we can bound the maximum 
+combination is possible. 
+In practice, by knowing the number of reactions that test $s$, we can bound the maximum 
 values of $h$ and $k$.
@@
 
-\noindent
 We exploit names $r_j,p_j$ to join the chains provided by the application of all the reactions. 
 Channels $r_j$ and $r_{j+1}$ enclose the enabling/disabling condition of reaction $aj$.
-Channels $p_i$ and $p_{j+1}$ enclose the links related to the entities produced by $aj$.
+Channels $p_j$ and $p_{j+1}$ enclose the links related to the entities produced by $aj$.
 We will see that all the link chain labels of transitions follow the same schema: first we find all the reactions limited to the reactants and inhibitors (chained using $r_j$ channels), then all the supplies by the contexts (chained using the channel $\mathit{cxt}$, to be introduced next), and finally the products for all the reactions (chained using $p_j$ channels).
@@
 %\left \{
-^{cxt} \backslash  \blockchain{s_i}{s_o}{s \in C_n} \, \link{}{} \blockchain{\overline{e}_i}{\overline{e}_o}{e \not\in C_n} \,  \backslash_{p_1}. \mathit{Cxt}^{n+1}
+^{cxt} \backslash  \blockchain{\hat{s}_i}{\hat{s}_o}{s\!\in \!C_n}  \link{}{} \blockchain{\underline{e}_i}{\underline{e}_o}{e \!\not\in \! C_n}   \backslash_{p_1}. \mathit{Cxt}^{n+1}
 %+ \\
@@
 $$
-and will focus on the complete transitions of $\llparenthesis{\cal A},\gamma\rrparenthesis$.
+and we will focus on the complete transitions of $\llparenthesis{\cal A},\gamma\rrparenthesis$.
 
@@
  \begin{array}{rcl}
- P_{r_1} &\defeq& \startchain{r_1}\chainedlink{s1_i}{\noact}\chainedlink{\noact}{s1_o}\chainedlink{r_2}{\noact} \chainedlink{\noact}{p_1}\chainedlink{\widetilde{s2}_i}{\noact}\chainedlink{\noact}{\widetilde{s2}_o}\chainend{p_2}.P_{r_1} + \dots;   \\
-  P_{r_2} &\defeq& \dots + \startchain{r_2}\chainedlink{\overline{s2}_i}{\noact}\chainedlink{\noact}{\overline{s2}_o}\chainedlink{\mathit{cxt}}{\noact} \chainedlink{\noact}{p_2}\chainend{\silent}.P_{r_2} + \dots; \\
+ P_{r_1} &\defeq& \startchain{r_1}\chainedlink{s1_i}{\noact}\chainedlink{\noact}{s1_o}\chainedlink{r_2}{\noact} \chainedlink{\noact}{p_1}\chainedlink{\widetilde{s2}_i}{\noact}\chainedlink{\noact}{\widetilde{s2}_o}\chainend{p_2}.P_{r_1} + \dots;   \\[6pt]
+  P_{r_2} &\defeq& \dots + \startchain{r_2}\chainedlink{\overline{s2}_i}{\noact}\chainedlink{\noact}{\overline{s2}_o}\chainedlink{\mathit{cxt}}{\noact} \chainedlink{\noact}{p_2}\chainend{\silent}.P_{r_2} + \dots; \\[6pt]
    P_{s1} &\defeq& \startchain{s1_i}\chainedlink{s1_o}{\noact}\chainedlink{\noact}{\widehat{s1}_i}\chainend{\widehat{s1}_o}
  %{\noact}\chainedlink{\noact}{\tilde{s1}_i}\chainedlink{\tilde{s1}_o}
- .P_{s1} + \dots; \\
+ .P_{s1} + \dots; \\[6pt]
   \overline{P}_{s2} &\defeq& \startchain{\overline{s2}_i}\chainedlink{\overline{s2}_o}{\noact}
-   \chainedlink{\noact} {\widehat{s2}_i}\chainedlink{\widehat{s2}_o}{\noact}\chainedlink{\noact} {\widetilde{s2}_i}\chainend{\widetilde{s2}_o}.P_{s2} + \dots;\\
+   \chainedlink{\noact} {\widehat{s2}_i}\chainedlink{\widehat{s2}_o}{\noact}\chainedlink{\noact} {\widetilde{s2}_i}\chainend{\widetilde{s2}_o}.P_{s2} + \dots;\\[6pt]
  \mathit{Cxt}& \defeq & \startchain{\mathit{cxt}}\chainedlink{\widehat{s1}_i}{\noact}\chainedlink{\noact}{\widehat{s1}_o}
@@
 \end{array}
-\]  \\
+\]
+
 For clarity of exposition, we show the code of the processes just in part, to focus on the prefixes that will be involved in the first transition of the system.
@@
     solid links to be merged with the virtual links of the backbone (i.e. to be plugged in the backbone).
+{\color{red}  In Figure~\ref{fig:backbone} we added the underline brackets, where the notation $P_1(P_2,P_3)$ means that the segment of the link chain is delimited by the process $P_1$ and it leaves "holes" where processes between the brackets insert their links.}
  Formally, we have
@@
   \end{tikzpicture}
+  $
+  \underbrace{\phantom{loremipsu}}_\text{Init}   \underbrace{\phantom{loremipsuo}}_\text{$P_{r1} (P_{s1})$} \underbrace{\phantom{loremip}}_\text{ $P_{r2}(\overline{P}_{s2})$}\underbrace{\phantom{loremipsumlore}}_\text{$Cxt (P_{s1},\overline{P}_{s2})$}
+     \underbrace{\phantom{loremipsumlo}}_\text{$P_{r1} (\overline{P}_{s2})$}
+      \underbrace{\phantom{loremip}}_\text{ $P_{r2}$}
+  $
   \caption{The link chain structure arising from reactions and context processes.}
@@
 This transformation is harmless, in the sense that it retains all the information in the chain, because it is applied to complete chains only.
-The function $\mathit{flat}$ is defined inductively as follows:
+The function $\mathit{flat}(\cdot)$ is defined inductively as follows:
 %\[
@@
 % \end{proposition}
-\noindent
+
 Now, we extend the previous result to a series of transitions.
@@
 
-\noindent
 Now, we extend the previous result to a series of transitions.

### 2020-06-18 TCS_version/06-exampleLactose.tex
reviewed directory

% !TEX root = ./main_RS2L_LNCS.tex
\section{Example: \emph{lac} operon}
\label{ex:lactose}
In this section we present the encoding of a RS example
taken from~\cite{CMMBM12}.
%, concerning the lac Operon.
\subsection{The \emph{lac} operon}
An operon is a cluster of genes under the control of a single promoter. 
The \emph{lac} operon  is involved in the metabolism of lactose in \emph{Escherichia coli} cells;
it is composed by three adjacent structural genes (plus some regulatory components):  $lacZ$, $lacY$ and $lacA$ encoding for two enzymes $Z$ and $A$, and a transporter $Y$, involved in the digestion of the $lactose$. The main regulations are:
\begin{itemize}
\item the gene $lacI$ encodes for a repressor protein $I$;
\item the DNA sequence, called \emph{promoter},  is recognised by a RNA polymerase
to iniziate the transcription  of the genes $lacZ$, $lacY$ and $lacA$;
\item a DNA segment, called  the \emph{operator} ($OP$),   obstructs the RNA polymerase functionality when  the repressor protein $I$ is bound to it forming $I\textrm{-}OP$;
\item  a short DNA sequence, called the $CAP\textrm{-}binding\ site$, when it is bound to the complex composed by  the protein $CAP$ and the signal molecule $cAMP$, acts as a promoter for the interaction between the RNA polymerase and the promoter.
\end{itemize}
The functionality of the \emph{lac} operon depends on the integration of two control mechanisms, one mediated by \emph{lactose}, and the other one mediated by \emph{glucose}.

In the first control mechanism, an effect of the absence of the \emph{lactose} is that $I$ is able to bind the operator sequence preventing the \emph{lac} operon expression. If lactose is available, $I$ is unable to bind the operator sequence, and the \emph{lac} operon can be potentially expressed.

In  the second control mechanism, when glucose is absent, the molecule $cAMP$ and the protein $CAP$ increase the \emph{lac} operon expression, thanks to the fact that  the binding between the molecular complex $cAMP\textrm{-}CAP$ and the $CAP\textrm{-}binding\ site$ increases.
In summary, the condition promoting the operon
gene expression is when the $lactose$ is present and the $glucose$ is absent.

In the following we report the description of the \emph{lac} operon mechanism in the reaction system formalism
and then show its encoding in c\CNA.

\subsection{The RS formalization}\label{subsec:RSlacoperon}
The reaction system for the \emph{lac} operon is defined as
$A_{lac} = (S, A)$, where the set S represents
the main biochemical components involved in this genetic system, while the reaction set A contains the biochemical
reactions involved in the regulation of the \emph{lac} operon expression. Formally, the \emph{lac} operon reaction system is defined as
follows:
$S$ is the set\\ $\{lac, Z, Y, A, lacI, I, I\textrm{-}OP, cya, cAMP, crp, CAP,cAMP\textrm{-}CAP, lactose,glucose\}$,
and A consists of the following 10 reactions:
{\small
\[
\begin{array}{lcllcl}
a_1 &= &(\{lac\},\{. . .\},\{lac\}), & a_6 &= &(\{cya\},\{. . .\},\{cAMP\}),\\
a_2 &= &(\{lacI\},\{. . .\},\{lacI\}), & a_7 &= &(\{crp\},\{. . .\},\{crp\}),\\
a_3 &= &(\{lacI\},\{. . .\},\{I\}), & a_8 &= &(\{crp\},\{. . .\},\{CAP\}),\\
a_4 &= &(\{I\},\{lactose\},\{I\textrm{-}OP\}), & a_9 &= &(\{cAMP, CAP\},\{glucose\},\{cAMP\textrm{-}CAP\}),\\
a_5 &= &(\{cya\},\{. . .\},\{cya\}),&
a_{10} &=& (\{lac, cAMP\textrm{-}CAP\},\{I\textrm{-}OP\},\{Z, Y, A\}).
\end{array}
\]
}
The default context ($DC$) is composed by those entities that are always present in the system
$DC = \{lac,lacI,I,cya,cAMP,crp,CAP\}$, whereas the $lactose$ and the $glucose$ are given non-deterministically by the context.
\subsection{The RS encoding}
For the sake of readability, the encoding we propose exploits the specific features of the example in hand 
to perform some simplifications:
\begin{itemize}
\item for the entities in the default context, $s \in DC$, as they are persistent, we do not provide the $\overline{P_s}$ processes and the $Cxt_s$ processes;
\item for the reactions requiring the presence of entities $s \in DC$, we do not provide the reaction alternative behaviour for when $s$ is absent;
\item the $Cxt_s$ processes are specified only for those entitiess that are really provided by the context.
\end{itemize}
Moreover, we do not model the \emph{dummy} entity that is specified by  dots ($\dots$) by the RS reactions in Section~\ref{subsec:RSlacoperon}.
Finally, we exclude the \emph{duplication reactions} ($a_1$, $a_2$, $a_5$, $a_7$), and renumber the remaining reactions :
{\small
\[
\begin{array}{cccl}
%a1 &= &(\{lac\},\{. . .\},\{lac\}),\\
%a2 &= &(\{lacI\},\{. . .\},\{lacI\}),\\
\textrm{old } &\textrm{new }&& \textrm{reactions}\\
a_3 &a_1&=&(\{lacI\},\{. . .\},\{I\}),\\
a_4 &a_2&=&(\{I\},\{lactose\},\{I\textrm{-}OP\}),\\
%a5 &= &(\{cya\},\{. . .\},\{cya\}),\\
a_6 & a_3&=&(\{cya\},\{. . .\},\{cAMP\}),\\
%a7 &= &(\{crp\},\{. . .\},\{crp\}),\\
a_8 &a_4&=&(\{crp\},\{. . .\},\{CAP\}),\\
a_9 &a_5&=&(\{cAMP, CAP\},\{glucose\},\{cAMP\textrm{-}CAP\}),\\
a_{10} & a_6&=& (\{lac, cAMP\textrm{-}CAP\},\{I\textrm{-}OP\},\{Z, Y, A\}).
\end{array}
\]
}

%\underline{Processes for the rules}\\
%We provide parametrised code duplication and for expression rules.\\
%\ \\
\paragraph{Duplication reactions.}
As explained previously, their encoding is omitted.
%%$
%\begin{array}{lcccl}
%P_{ai} & \defeq& P_d(ai,s) & = & \startchain{r_i}\chainedlink{s_i}{\noact}\chainedlink{\noact}{s_o}
%\chainedlink{\tilde{s}_i}{\noact}\chainedlink{\noact}{\tilde{s}_o}\chainend{r_{i+1}}.P_d(ai,s)
%\end{array}
%$,
%with $r_1 = \silent$.
%There are four duplication rules:\\ $P_d(a1,lac)$, $P_d(a2,lacI)$,$P_d(a5,cya)$,$P_d(a7,crp)$.\\
\paragraph{Expression reactions.}
First we define the parametric process
\[
\begin{array}{lcl}
% P_i(s1,s2) & \defeq & \startchain{r_i}\chainedlink{s1_i}{\noact}\chainedlink{\noact}{s1_o}
%\chainedlink{\tilde{s2}_i}{\noact}\chainedlink{\noact}{\tilde{s2}_o}\chainend{r_{i+1}}.P_i(s1,s2)\\
 P_i(s1,s2) & \defeq & \startchain{r_i}\chainedlink{s1_i}{\noact}\chainedlink{\noact}{s1_o}
 \chainedlink{r_{i+1}}{\noact}\chainedlink{\noact}{p_{i}}\chainedlink{\widetilde{s2}_i}{\noact}
\chainedlink{\noact}{\widetilde{s2}_o}\chainend{p_{i+1}}.P_i(s1,s2)\\
\end{array}
\]
Then, we let $P_{a1}\defeq P_1(lacI,I)$,  $P_{a3} \defeq P_3(cya,cAMP)$, and $P_{a4} \defeq P_4(crp,CAP)$.
%\[
%\begin{array}{lcl}
%P_{a1} & =& \startchain{\silent}\chainedlink{lac_i}{\noact}\chainedlink{\noact}{lac_o}\chainedlink{\tilde{lac}_i}{\noact}\chainedlink{\noact}{\tilde{lac}_o}\chainend{r_2}.P_{a1}\\
%P_{a2} & =& \startchain{r_2}\chainedlink{lacI_i}{\noact}\chainedlink{\noact}{lacI_o}\chainedlink{\tilde{lacI}_i}{\noact}\chainedlink{\noact}{\tilde{lacI}_o}\chainend{r_3}.P_{a2}\\
%P_{a3} & =& \startchain{r_3}\chainedlink{I_i}{\noact}\chainedlink{\noact}{I_o}\chainedlink{\tilde{I}_i}{\noact}\chainedlink{\noact}{\tilde{I}_o}\chainend{r_4}.P_{a3}\\
\paragraph{Regulation reactions.}

{\small
\[
\begin{array}{lcl}
%P_{a2} & \defeq& \startchain{r_2}\chainedlink{I_i}{\noact}\chainedlink{\noact}{I_o}\chainedlink{\overline{lactose}_i}{\noact}\chainedlink{\noact}{\overline{lactose}_o}
%\chainedlink{\widetilde{I\textrm{-}OP}_i}{\noact}\chainedlink{\noact}{\widetilde{I\textrm{-}OP}_o}\chainend{r_3}.P_{a2}  \\
P_{a2} & \defeq& \startchain{r_2}\chainedlink{I_i}{\noact}\chainedlink{\noact}{I_o}\chainedlink{\overline{lactose}_i}{\noact}\chainedlink{\noact}{\overline{lactose}_o}
\chainedlink{r_3}{\noact}\chainedlink{\noact}{p_2}
\chainedlink{\widetilde{I\textrm{-}OP}_i}{\noact}\chainedlink{\noact}{\widetilde{I\textrm{-}OP}_o}\chainend{p_3}.P_{a2}  \\
&&+\\
 && \startchain{r_2}\chainedlink{lactose_i}{\noact}\chainedlink{\noact}{lactose_o}\chainedlink{r_3}{\noact}\chainedlink{\noact}{p_2}\chainend{p_3}.P_{a2} \ +\  \startchain{r_2}\chainedlink{\overline{I}_i}{\noact}\chainedlink{\noact}{\overline{I}_o}\chainedlink{r_3}{\noact}\chainedlink{\noact}{p_2}\chainend{p_3}.P_{a2}\\[5pt]
%%P_{a5} & =& \startchain{r_5}\chainedlink{cya_i}{\noact}\chainedlink{\noact}{cya_o}\chainedlink{\tilde{cya}_i}{\noact}\chainedlink{\noact}{\tilde{cya}_o}\chainend{r_6}.P_{a5}\\
%P_{a6} & =& \startchain{r_6}\chainedlink{cya_i}{\noact}\chainedlink{\noact}{cya_o}\chainedlink{\tilde{cya}_i}{\noact}\chainedlink{\noact}{\tilde{cya}_o}\chainend{r_7}.P_{a6}\\
%%P_{a7} & =& \startchain{r_7}\chainedlink{crp_i}{\noact}\chainedlink{\noact}{crp_o}\chainedlink{\tilde{crp}_i}{\noact}\chainedlink{\noact}{\tilde{crp}_o}\chainend{r_8}.P_{a7}\\
%P_{a8} & =& \startchain{r_8}\chainedlink{crp_i}{\noact}\chainedlink{\noact}{crp_o}\chainedlink{\tilde{CAP}_i}{\noact}\chainedlink{\noact}{\tilde{CAP}_o}\chainend{r_9}.P_{a8}\\
P_{a5} & \defeq& \startchain{r_5}\chainedlink{cAMP_i}{\noact}\chainedlink{\noact}{cAMP_o}
\chainedlink{CAP_i}{\noact}\chainedlink{\noact}{CAP_o}\chainedlink{\overline{glucose}_i}{\noact}
\chainedlink{\noact}{\overline{glucose}_o}\chainedlink{r_6}{\noact}
\chainedlink{\noact}{p_5}
\chainedlink{\widetilde{cAMP\textrm{-}CAP}_i}{\noact}\chainedlink{\noact}{\widetilde{cAMP\textrm{-}CAP}_o}
\chainend{p_6}.P_{a5}\\
%P_{a5} & \defeq& \startchain{r_5}\chainedlink{cAMP_i}{\noact}\chainedlink{\noact}{cAMP_o}
%\chainedlink{CAP_i}{\noact}\chainedlink{\noact}{CAP_o}\chainedlink{\overline{glucose}_i}{\noact}
%\chainedlink{\noact}{\overline{glucose}_o}
%\chainedlink{\widetilde{cAMP\textrm{-}CAP}_i}{\noact}\chainedlink{\noact}{\widetilde{cAMP\textrm{-}CAP}_o}
%\chainend{r_6}.P_{a5}\\
&& +\\
&& \startchain{r_5}\chainedlink{glucose_i}{\noact}\chainedlink{\noact}{glucose_o} \chainedlink{r_6}{\noact}\chainedlink{\noact}{p_5}\chainend{p_6}.P_{a5}\\
&& + \\
&&  \startchain{r_5}\chainedlink{\overline{cAMP}_i}{\noact}\chainedlink{\noact}{\overline{cAMP}_o} \chainedlink{r_6}{\noact}\chainedlink{\noact}{p_5}\chainend{p_6}.P_{a5} \ + \ \startchain{r_5}\chainedlink{\overline{CAP}_i}{\noact}\chainedlink{\noact}{\overline{CAP}_o} \chainedlink{r_6}{\noact}\chainedlink{\noact}{p_5}\chainend{p_6}.P_{a5}\\[5pt]
%\end{array}
%\]\\
%%$lac$ operon expression.\\
%$
%\begin{array}{lcl}
P_{a6} & \defeq& \startchain{r_{6}}\chainedlink{lac_i}{\noact}\chainedlink{\noact}{lac_o}
\chainedlink{cAMP_i}{\noact}\chainedlink{\noact}{cAMP_o}\chainedlink{\overline{I\textrm{-}OP}_i}{\noact}
\chainedlink{\noact}{\overline{I\textrm{-}OP}_o}\chainedlink{cxt_1}{\noact}
\chainedlink{\noact}{p_6}
\chainedlink{\widetilde{z}_i}{\noact}\chainedlink{\noact}{\widetilde{z}_o}
\chainedlink{\widetilde{y}_i}{\noact}\chainedlink{\noact}{\widetilde{y}_o}
\chainedlink{\widetilde{A}_i}{\noact}\chainedlink{\noact}{\widetilde{A}_o}
\chainend{\silent}.P_{a6}\\
%P_{a6} & \defeq& \startchain{r_{6}}\chainedlink{lac_i}{\noact}\chainedlink{\noact}{lac_o}
%\chainedlink{cAMP_i}{\noact}\chainedlink{\noact}{cAMP_o}\chainedlink{\overline{I\textrm{-}OP}_i}{\noact}
%\chainedlink{\noact}{\overline{I\textrm{-}OP}_o}
%\chainedlink{\tilde{z}_i}{\noact}\chainedlink{\noact}{\tilde{z}_o}
%\chainedlink{\tilde{y}_i}{\noact}\chainedlink{\noact}{\tilde{y}_o}
%\chainedlink{\tilde{A}_i}{\noact}\chainedlink{\noact}{\tilde{A}_o}
%\chainend{cxt_1}.P_{a6}\\
&& +\\
&& \startchain{r_{6}}\chainedlink{I\textrm{-}OP_i}{\noact}\chainedlink{\noact}{I\textrm{-}OP_o} \chainedlink{cxt_1}{\noact}\chainedlink{\noact}{p_6}\chainend{\silent}.P_{a6} \\
&& + \\
&&  \startchain{r_{6}}\chainedlink{\overline{lac}_i}{\noact}\chainedlink{\noact}{\overline{lac}_o} \chainedlink{cxt_1}{\noact}\chainedlink{\noact}{p_6}\chainend{\silent}.P_{a6} \ + \ \chainedlink{\overline{cAMP}_i}{\noact}\chainedlink{\noact}{\overline{cAMP}_o} \chainedlink{cxt_1}{\noact}\chainedlink{\noact}{p_6}\chainend{\silent}.P_{a6}
\end{array}
\]
}
\paragraph{Processes for the entities.}
We exploit the specificity of the example in hand to optimising the code, and we specify exactly
the number of solid links that each process encoding an entity must offer.
For the always present entities we let:
\[
\begin{array}{lcl@{\qquad}lcl}
%P_{lac}&\defeq&\startchain{lac_i}\chainedlink{lac_o}{\noact}\chainedlink{\noact}{lac_i}\chainedlink{lac_o}{\noact}\chainedlink{\noact}{\tilde{lac}_i}\chainend{\tilde{lac}_o}.P_{lac} +
 %\startchain{lac_i}\chainedlink{lac_o}{\noact}\chainedlink{\noact}{\tilde{lac}_i}\chainend{\tilde{lac}_o}.P_{lac}\\
% P_{lac}&\defeq&\sum_{h=1}^2(\startchain{lac_i}\chainedlink{lac_o}{\noact}\chainend{\noact})^h\ \link{\tilde{lac}_i}{\tilde{lac}_o}.P_{lac}
%P_{cya}&\defeq&\startchain{cya_i}\chainedlink{cya_o}{\noact}\chainedlink{\noact}{cya_i}\chainedlink{cya_o}{\noact}
%\chainedlink{\noact}{\tilde{cya}_i}\chainend{\tilde{cya}_o}.P_{cya}\\
P_{cya}&\defeq&\link{cya_i}{cya_o}.P_{cya} &
%&
%P_{I}&\defeq&\sum_{h=0}^1(\startchain{I_i}\chainedlink{I_o}{\noact}\chainend{\noact})^h\ \link{\tilde{I}_i}{\tilde{I}_o}.P_{I} \\
 %{\noact}
%\chainedlink{\noact}{\cap{lac}_i}\chainend{\cap{lac}_o}.P_{lac}\\
%P_{crp}&\defeq&\startchain{crp_i}\chainedlink{crp_o}{\noact}\chainedlink{\noact}{crp_i}\chainedlink{crp_o}{\noact}\chainedlink{\noact}{\tilde{crp}_i}\chainend{\tilde{crp}_o}.P_{crp}
P_{crp}&\defeq&\link{crp_i}{crp_o}.P_{crp}\\[5pt]
%&
%P_{cAMP}&\defeq&\sum_{h=0}^1(\startchain{cAMP_i}\chainedlink{cAMP_o}{\noact}\chainend{\noact})^h\ \link{\widetilde{cAMP}_i}{\widetilde{cAMP}_o}.P_{cAMP}
P_{lacI}&\defeq& \link{lacI_i}{lacI_o}.P_{lacI} &
%&
%P_{CAP}&\defeq&\sum_{h=0}^1(\startchain{CAP_i}\chainedlink{CAP_o}{\noact}\chainend{\noact})^h\ \link{\widetilde{CAP}_i}{\widetilde{CAP}_o}.P_{CAP}\\
%{lacI_i}\chainedlink{lacI_o}{\noact}\chainedlink{\noact}{\tilde{lacI}_i}\chainend{\tilde{lacI}_o}.P_{lacI}\\
% \startchain{\tilde{I}_i}\chainend{\tilde{I}_o}.P_{I}\\
%P_{I}&\defeq&\startchain{I_i}\chainedlink{I_o}{\noact}\chainedlink{\noact}{\tilde{I}_i}\chainend{\tilde{I}_o}.P_{I} +
% \startchain{\tilde{I}_i}\chainend{\tilde{I}_o}.P_{I}\\
%P_{cya}&\defeq&\startchain{cya_i}\chainedlink{cya_o}{\noact}\chainedlink{\noact}{cya_i}\chainedlink{cya_o}{\noact}\chainedlink{\noact}{\tilde{cya}_i}\chainend{\tilde{cya}_o}.P_{cya}\\
%P_{lacI}&\defeq&\sum_{h=1}^2(\startchain{lacI_i}\chainedlink{lacI_o}{\noact}\chainend{\noact})^h\ \link{\tilde{lacI}_i}{\tilde{lacI}_o}.P_{lacI}\\
 P_{lac}&\defeq&\link{lac_i}{lac_o}.P_{lac}
%P_{cAMP}&\defeq&\startchain{cAMP_i}\chainedlink{cAMP_o}{\noact}\chainedlink{\noact}{\tilde{cAMP}_i}\chainend{\tilde{cAMP}_o}.P_{cAMP} + \startchain{\tilde{cAMP}_i}\chainend{\tilde{cAMP}_o}.P_{cAMP}\\
%P_{crp}&\defeq&\startchain{crp_i}\chainedlink{crp_o}{\noact}\chainedlink{\noact}{crp_i}\chainedlink{crp_o}{\noact}\chainedlink{\noact}{\tilde{crp}_i}\chainend{\tilde{crp}_o}.P_{crp}\\
\end{array}\]
For the entities always produced (i.e. not present only at the first step), we  provide a  parametric
definition  
$P_e(s) \defeq  \startchain{s_i}\chainedlink{s_o}{\noact}\chainedlink{\noact}{\widetilde{s}_i}\chainend{\widetilde{s}_o}.P_e(s) +  \startchain{\widetilde{s_i}}\chainend{\widetilde{s}_o} .P_e(s) $.\\
% There are four entities of the first type: $P_d(lac)$, $P_d(cya)$, $P_d(crp)$, $P_d(lacI)$.
 There are three entities of the second type:  $$P_{cAMP}\defeq P_e(cAMP) \qquad P_{CAP} \defeq P_e(CAP)\qquad P_I \defeq P_e(I).$$

The entity $I\textrm{-}OP$ can be either produced (by $a_2$) or tested for absence (by $a_6$).
Correspondingly, the process  $P_{I\textrm{-}OP}$ is defined as follows:

\[\begin{array}{lcl}
P_{I\textrm{-}OP}&\defeq & 
\sum_{h=0}^1 (\startchain{I\textrm{-}OP_i}\chainedlink{I\textrm{-}OP_o}{\noact}\chainend{\noact})^h  \, \startchain{\widetilde{I\textrm{-}OP}_i}\chainend{\widetilde{I\textrm{-}OP}_o}.P_{I\textrm{-}OP}
%\\
%&&+\\
%&&\sum_{h=0}^1(\startchain{I-OP_i}\chainedlink{I-OP_o}{\noact}\chainend{\noact})^h\ \startchain{\widetilde{I-OP}_i}\chainedlink{\widetilde{I-OP}_o}{\noact}\chainedlink{\noact}{\underline{I-OP}_i}\chainend{\underline{I-OP}_o}.P_{I-OP}\\
%&&
\  +\ %\\
%&&
\link{I\textrm{-}OP_i}{I\textrm{-}OP_o}.\overline{P_{I\textrm{-}OP}}\\
\overline{P_{I\textrm{-}OP}} & \defeq & 
% P_{I\textrm{-}OP}(\overline{I\textrm{-}OP})\\
\sum_{h=0}^1  (\startchain{\overline{I\textrm{-}OP}_i}\chainedlink{\overline{I\textrm{-}OP}_o}{\noact}\chainend{\noact})^h \,
\startchain{\widetilde{I\textrm{-}OP}_i}\chainend{\widetilde{I\textrm{-}OP}_o}.P_{I\textrm{-}OP}%\\
%&&
\ +\ %\\
%&&
\link{\overline{I\textrm{-}OP}_i}{\overline{I\textrm{-}OP}_o}.\overline{P_{I\textrm{-}OP}}\\

\end{array}
\]
\noindent
The  process $P_{cAMP\textrm{-}CAP}$ is similar to $P_{I\textrm{-}OP}$, as it is produced by $a_5$ and tested for presence by $a_6$. Its code is  in Table~\ref{tab:cAMP-CAP}, in the Appendix.
The $lactose$ is provided by the context and tested for absence by $a_2$.

\[
\begin{array}{lcllcl}
P_{lactose} &\defeq &%== \overline{P_{glucose}}&\defeq & P_{glu-lact}(\overline{glucose})\\
\sum_{h=0}^1 (\startchain{lactose_i}\chainedlink{lactose_o}{\noact}\chainend{\noact})^h\ \link{\widehat{lactose_i}}{\widehat{lactose_o}}.P_{lactose}
\\
&& + \\
&& \sum_{h=0}^1 (\startchain{lactose_i}\chainedlink{lactose_o}{\noact}\chainend{\noact})^h\ \link{\underline{lactose_i}}{\underline{lactose_o}}.\overline{P_{lactose}}\\
%P_{glucose} &\defeq &\sum_{h=0}^1 \underbrace{\link{glucose_i}{glucose_o}}_{h\ times}
%
%\[
%\begin{array}{lcl}

\overline{P_{lactose}} &\defeq &\sum_{h=0}^1 (\startchain{\overline{lactose}_i}\chainedlink{\overline{lactose}_o}{\noact}\chainend{\noact})^h\ \link{\widehat{lactose_i}}{\widehat{lactose_o}}.P_{lactose}\\
&&+\\
&& \sum_{h=0}^1 (\startchain{\overline{lactose}_i}\chainedlink{\overline{lactose}_o}{\noact}\chainend{\noact})^h\ \link{\underline{lactose_i}}{\underline{lactose_o}}.\overline{P_{lactose}}\\
\end{array}
\]
\noindent
The  process $P_{glucose}$ is similar to $P_{lactose}$ and tested for absence by $a_5$. Its code is  in Table~\ref{tab:glucose}, in the Appendix.
The entity $z$ can only be produced by rule $a_6$, while it is never provided by the context. Moreover, there is no rule for testing its presence or absence.
\[
\begin{array}{lcl@{\qquad\qquad}lcl}
P_z&\defeq &\startchain{\widetilde{z_i}}\chainedlink{\widetilde{z_o}}{\noact}\chainedlink{\noact}{\underline{z_i}}\chainend{\underline{z_o}}.P_z\ + \
\link{\underline{z_i}}{\underline{z_o}}.\overline{P_z}
&
\overline{P_z} &\defeq &\startchain{\widetilde{z_i}}\chainedlink{\widetilde{z_o}}{\noact}\chainedlink{\noact}{\underline{z_i}}\chainend{\underline{z_o}}.\overline{P_z}\ + \
\link{\underline{z_i}}{\underline{z_o}}.\overline{P_z}\\\end{array}
\]
\noindent
The entities $y$ and $A$ are treated in the same way as $z$. Their processes are in Table~\ref{tab:yA} in the Appendix.

\paragraph{Context.}
%Now, we define the behaviour of the context.
The entities in $DC$ are assumed always present by default, so no context process is needed for them.
The entities $z$, $y$, and $A$ are assumed never provided by the context. Their processes are
\[
\begin{array}{rcl@{\qquad\qquad}rcl}
Cxt_{z}&\defeq & \startchain{cxt_1}\chainedlink{\underline{z}_i}{\noact}\chainedlink{\noact}{\underline{z}_o}\chainend{cxt_2}.Cxt_{z}
&
Cxt_{y}&\defeq & \startchain{cxt_2}\chainedlink{\underline{y}_i}{\noact}\chainedlink{\noact}{\underline{y}_o}\chainend{cxt_3}.Cxt_{y}
\\
Cxt_{A}&\defeq & \startchain{cxt_3}\chainedlink{\underline{A}_i}{\noact}\chainedlink{\noact}{\underline{A}_o}\chainend{cxt_4}.Cxt_{A}
\end{array}
\]
\noindent
%Finally, we need to specify the context for  $lactose$ and $glucose$.
For the sake of presentation, we assume that the $lactose$ is always provided by the context, in contrast,
$glucose$ is never provided.
\[
\begin{array}{lcllcl}
Cxt_{lactose}&\defeq& \startchain{cxt_4}\chainedlink{\widehat{lactose}_i}{\noact}\chainedlink{\noact}{\widehat{lactose}_o}\chainend{cxt_5}.Cxt_{lactose}
\\
Cxt_{glucose}&\defeq & \startchain{cxt_5}\chainedlink{\underline{glucose}_i}{\noact}\chainedlink{\noact}{\underline{glucose}_o}\chainend{p_1}.Cxt_{glucose}
\end{array}
\]
\noindent
In the following we let $CXT \defeq Cxt_z | Cxt_y |Cxt_A | Cxt_{lactose} | Cxt_{glucose}$ be the processes for context. 
%
The whole system
%, encoding an extended interactive process, with the previous context behaviour
is as follows: 
\[ lacOp \defeq \restrict{names} (\Pi_{i=1}^6 P_{ai} | \Pi_{s \in DC} P_s |   \Pi_{s \in S \backslash DC} \overline{P}_s| CXT)
\]

\paragraph{Execution.}
Now, we show two transitions. 
After the first transition the entity $cAMP\textrm{-}CAP$ is produced due to the absence of $glucose$, while the presence of $lactose$ inhibits the production of $I\textrm{-}OP$:
$lacOp \xrightarrow{\restrict{names} v} lacOp'$
where: 
{\small
$$
\begin{array}{rcl}
v & = & \link{\silent}{lac_i}\dots \link{lactose_o}{r3}\dots
 \link{r_5}{cAMP_i}\dots
  \link{\overline{glucose}_o}{r_6}
\dots \link{p_5}{\widetilde{cAMP\textrm{-}CAP}_i}\dots \link{p_6}{\silent}\\[5pt]
%{\underline{glucose}_o}
lacOp' & \defeq & \restrict{names} (\Pi_{i =1}^6 P_{ai} | \Pi_{s \in AP} P_s |   \Pi_{s \in S \backslash AP} \overline{P}_s| CXT)
\end{array}
$$ 
}
with $AP= DC\cup \{cAMP\textrm{-}CAP\}$  the actual context.\\
After the second step the entities $z$, $y$ and $A$ are produced, due to the presence of $cAMP\textrm{-}CAP$ and the absence of $I\textrm{-}OP$, thus
$lacOp' \xrightarrow{\restrict{names}v'} lacOp''$
where:
{\small
$$
\begin{array}{rcl}
v' & = & \link{\silent}{lac_i}\dots \link{lac_o}{cAMP-CAP_i}\dots \startchain{\overline{I-OP}_o}
\chainedlink{cxt_1}{\dots}\startchain{\underline{glucose}_o}\chainedlink{p_1}{p_1}
\chainedlink{\dots}{p_6}\chainedlink{\widetilde{z}_i}{\widetilde{z}_i}\chainedlink{\widetilde{z}_o}{\widetilde{z}_o}
\chainedlink{\widetilde{y}_i}{\widetilde{y}_i}\chainedlink{\widetilde{y}_o}{\widetilde{y}_o}\chainedlink{\widetilde{A}_i}{\widetilde{A}_i}\chainedlink{\widetilde{A}_o}{\widetilde{A}_o}\chainend{\silent}\\[5pt]
%\chainend{cxt_1}\dots \link{\underline{glucose}_o}{\silent}\\[5pt]
lacOp'' & \defeq & \restrict{names} (\Pi_{i =1}^{6} P_{ai} | \Pi_{s \in AP'} P_s |   \Pi_{s \in S \backslash AP'} \overline{P}_s| CXT)
\end{array}
$$
}
with $AP'= DC\cup \{z,y,A\} $.

### 2020-06-18 TCS_version/06-examples.tex
reviewed directory

--- 2020-04-05 06-examples.tex
@@
-% !TEX root = ./main_RS2L_LNCS.tex
+% !TEX root = ./main_TCS.tex
 \section{Examples}
@@
 
-Hereafter, we shall often omit topmost restrictions but shall take into account only transitions whose labels are complete, i.e. they do not contain 
+Hereafter, we shall often omit topmost restrictions but shall take into account only transitions whose labels are complete chains, i.e. they do not contain 
 virtual links and start/end with the $\silent$ action symbol.
@@
 \begin{array}{cl@{\hspace{1cm}} cl}
-1 \ &(\{q,a\},\{w,b\},\{q\}) & 2\ &(\{q,b\},\{w,a\},\{w\})\\
-3 \  &(\{w,a\},\{q,b\},\{q\}) & 4\ &(\{w,b\},\{q,a\},\{w\})\\ 
+1 \ &(\{q,a\},\{w,b\},\{w\}) & 2\ &(\{q,b\},\{w,a\},\{q\})\\
+3 \  &(\{w,a\},\{q,b\},\{w\}) & 4\ &(\{w,b\},\{q,a\},\{q\})\\ 
 \end{array}
@@
 \]
-\noindent
+
 Then we have 
 \[
-\begin{array}{lcl @{\hspace{0.5cm}}lcl  }
-P_1 & \defeq & P_1(q,a,w,b,q) & P_3 & \defeq & P_3(w,a,q,b,w)  \\
-P_2 & \defeq & P_2(q,b,w,a,w) &P_4 & \defeq & P_4(w,b,q,a,q) 
+\begin{array}{lcl @{\hspace{1cm}}lcl  }
+P_1 & \defeq & P_1(q,a,w,b,w) & P_3 & \defeq & P_3(w,a,q,b,w)  \\
+P_2 & \defeq & P_2(q,b,w,a,q) &P_4 & \defeq & P_4(w,b,q,a,q) 
 \end{array}
 \]
-\noindent
 and we put, as usual,
@@
 \]
+
 In fact the presence/absence of $q$ and $w$ will be exploited by at least one rule and at most three rules.
@@
 
-Notice that we exploit here the capabilities of the process algebraic framework to define a nondeterministic, recursive context:
+Notice that we exploit here the capabilities of the process algebraic framework to define a nondeterministic, recursive context.
 We model the context to always offer either $a$ or $b$, but never both the entities together. The reason is that in
@@
 \]
+
 Then, only the second reaction can be applied, and the transition carries the complete label $\upsilon$ below
@@
  \chainedlink{p_2}{p_2}
- \chainedlink{\widetilde{w}_i}{\bf \widetilde{w}_i}\chainedlink{\bf \widetilde{w}_o}{\widetilde{w}_o}
+ \chainedlink{\widetilde{q}_i}{\bf \widetilde{q}_i}\chainedlink{\bf \widetilde{q}_o}{\widetilde{q}_o}
  \chainedlink{p_3}{p_3}
@@
 In the label we can read that the rules $1$ and $4$ have been not executed because the entity $a$ is absent, 
-the rule $3$ has been not applied because the entity $w$ is absent, then only rule $2$ has been applied, and it has produced the entity $w$.
+the rule $3$ has been not applied because the entity $w$ is absent, then only rule $2$ has been applied, and it has produced the entity $q$.
 Also,  the context provides entity $a$, that will be available in the next state, and not the entity $b$.
@@
 \[
-r_1~\overline{a}~r_2~q~b~\overline{w}~\overline{a}~r_3~\overline{a}~r_4~\overline{w}~\mathit{cxt}~\widehat{a}~\underline{b}~p_1~p_2~\widetilde{w}~p_3~p_4\]
+r_1~\overline{a}~r_2~q~b~\overline{w}~\overline{a}~r_3~\overline{a}~r_4~\overline{w}~\mathit{cxt}~\widehat{a}~\underline{b}~p_1~p_2~\widetilde{q}~p_3~p_4 .
+\]
 
@@
 processed messenger RNA (mRNA) produced from that gene.
-
-where a gene $a$  codes for a protein $T$ when molecules 
-$G$ is present and $C$ is absent, and in the opposite situation $a$ codes for protein $T'$.
+ {\color{red} In our example,
+%where 
+a gene $a$  codes for a protein $T$ when molecules 
+$G$ is present and $C$ is absent, and in the opposite situation $a$ codes for protein $T'$.}
 This behavior is encoded in rules $1$ and $2$.
@@
 \]
-\noindent
+
 Then we have 
@@
 \]
+
 Then, we have
@@
 P_T & \defeq & P(T,\widetilde{T})& \overline{P}_T & \defeq & P(\overline{T},\widetilde{T}) & P_{T'}& \defeq & P(T',\widetilde{T'}) & \\
-\overline{P}_{T',\widetilde{T'}}& \defeq & P(\overline{T'},\widetilde{T'})\\\
+%\overline{P}_{T',\widetilde{T'}}& \defeq & P(\overline{T'},\widetilde{T'})
+\overline{P}_{T'}& \defeq & P(\overline{T'},\widetilde{T'})
 \end{array}
 \]
+
 The entities that can be produced by the reactions and that can be  provided by the context are as follows:
@@
 \begin{array}{lcl @{\hspace{1cm}}lcl}
-P(C,\widehat{C},\underline{C},\widetilde{C}) & \defeq &  \sum_{h=0}^1 (\startchain{C_i}\chainedlink{C_o}{\noact}\chainend{\noact})^h\ \link{\widehat{C}_i}{\widehat{C}_o} (\startchain{\noact}\chainedlink{\noact}{\widetilde{C}_i}\chainend{\widetilde{C}_o})^h.
+P(C,\widehat{C},\underline{C},\widetilde{C}) & \defeq &  \sum_{h=0}^1 (\startchain{C_i}\chainedlink{C_o}{\noact}\chainend{\noact})^h \ \startchain{\widehat{C}_i}\chainedlink{\widehat{C}_o} {\noact}\chainend{\noact}( \startchain{\widetilde{C}_i}\chainedlink{\widetilde{C}_o}{\noact}\chainend{\noact})^h.
 P(C,\widehat{C},\underline{C},\widetilde{C})\\
@@
 \]
+
 In the above configuration, reactions $1$ and $4$ can be applied, and also we assume that the context will provide the entity $F$, that will be available in the target configuration.
@@
 %  \chainedlink{p_2}{p_2}\chainedlink{p_3}{p_3}\chainedlink{p_4}{p_4} \chainedlink{\widetilde{G}_i}{\bf \widetilde{G}_i}\chainedlink{\bf \widetilde{G}_o}{\widetilde{G}_o} \chainend{\silent}
-r_1~ a~G~\overline{C}~ r_2~ G~ r_3~T'~r_4~T'~\overline{F}~cxt~\underline{C}~\underline{G}~\widehat{F}~p_1~\widetilde{T}~p_2~p_3~p_4~\widetilde{G}
+r_1~ a~G~\overline{C}~ r_2~ G~ r_3~T'~r_4~T'~\overline{F}~cxt~\underline{C}~\underline{G}~\widehat{F}~p_1~\widetilde{T}~p_2~p_3~p_4~\widetilde{G} .
 \]
+
 The original label can then be reconstructed just applying the function $\mathit{unflat}(\cdot)$ to the string above.

### 2020-06-18 TCS_version/06b-bisimulation.tex
reviewer 1

--- 2020-04-05 06b-bisimulation.tex
@@
-% !TEX root = ./main_RS2L_LNCS.tex
+% !TEX root = ./main_TCS.tex
 \section{Bio-simulation}
@@
 There are many variants of the bisimulation for process algebras,  for example the
-barbed bisimulation~\cite{10.1007/3-540-55719-9_114} only consider the execution of invisible actions, and then equates two processes when the expose the same prefixes; for the mobile ambients~\cite{CardelliG00}, a process algebra equipped with a reduction semantics, a notion of
+barbed bisimulation~\cite{10.1007/3-540-55719-9_114} only consider the execution of invisible actions, and then equates two processes when they expose the same prefixes; for the mobile ambients~\cite{CardelliG00}, a process algebra equipped with a reduction semantics, a notion of
 behavioural equivalence equates two processes when they expose the same ambients~\cite{GC03}.
@@
 
-For instance we would like to express properties about each step of the simulation of a system like in the following:
-
 \begin{example} \label{properties_ex}
+For instance we would like to express properties about each step of the bio-simulation of a system like the ones below:
 \begin{enumerate}
@@
 and that combines regular expression operators with conjunction and disjunction.
-In the following we let $\mathit{syms} = \mathit{decs}\cup\mathit{reacts}\cup\mathit{prods}\cup\{\mathit{cxt}\}$ be the set of symbolic names used in our assertion language.
+We let $\mathit{syms} = \mathit{decs}\cup\mathit{reacts}\cup\mathit{prods}\cup\{\mathit{cxt}\}$ be the set of symbolic names used in our assertion language.
 
@@
 
+\begin{example}
 The formulas corresponding to the sample queries listed in Example~\ref{properties_ex} are as follows:
-
 \begin{enumerate}
@@
 Viceversa, if the reaction $r_j$ has been applied the occurrence of $r_j$ must be followed by at least one of the symbols in $\{s_1,...,s_n\}$ and then by at least one of the symbols in $\{\overline{s}_1,...,\overline{s}_n\}$.
+\end{example}
 
@@
 The notion of bio-simulation builds on the above language of assertions to parameterize the induced equivalence on the property of interest. 
-Please recall that we have defined the behaviour of the contest in a non deterministic way, thus 
+Please recall that we have defined the behaviour of the context in a non deterministic way, thus 
 at each step, different possible sets of entities can be provided to the system and different sets of reaction can be enabled/disabled. 
@@
 %\end{itemize}
+
 We let $ \sim_{\mathsf{F}} $ denote the largest  bio-simulation and we say that $P$ is \emph{bio-similar} to $Q$, with respect to $\mathsf{F}$, if $P \sim_{\mathsf{F}} Q$.
@@
 Please remember that the notation $P \xRightarrow{v} P'$ refers to ordinary transitions $P \xrightarrow{v} P'$ where $v$ is a complete chain (solid and with $\silent$ actions at the extremes).
-It should not be confused with the notation for weak transitions commonly found in the literature on process algebras.
+The double arrow notation should not be confused with the notation for weak transitions commonly found in the literature on process algebras.
 \end{remark}
@@
 \]
+
 We write $P \entails \mathsf{G}$ ($P$ satisfies $\mathsf{G}$) if and only if $P \in  \llbracket \mathsf{G} \rrbracket$.

### 2020-06-18 TCS_version/09-appendix.tex
reviewer 1

--- 2020-04-05 09-appendix.tex
@@
 by Definition~\ref{def:trans} and by Lemma~\ref{lem:struct}.1, all the  channels $r_j,p_j$, with
- $j \in [1,\dots,u]$, and $cxt_h$, with $h \in [1, \dots, w]$, and all the annotated versions of 
+ $j \in [1,\dots,u]$, and $cxt$,
+ %$cxt_h$, with $h \in [1, \dots, w]$,
+  and all the annotated versions of 
  $s_i$, $s_o$ are restricted. Also, processes $\mathit{Cxt}$ always requires the 
@@
 %The code executed by any process encoding rule $R_i$ producing $e$ is 
-$$P_{aj}\defeq \link{r_j}{\dots}\chainedlink{r_{j+1}}{\noact}\chainedlink{\noact}{p_j}\link{}{\dots}\chainedlink{\widetilde{s}_i}{\noact}\chainedlink{\noact}{\dots}\link{\widetilde{s}_o}{\dots} \chainend{p_{j+1}}.P_{aj} ;$$
+$$P_{aj}\defeq \link{r_j}{\dots}\chainedlink{r_{j+1}}{\noact}\chainedlink{\noact}{p_j}\link{}{\dots}\chainedlink{\widetilde{s}_i}{\noact}\chainedlink{\noact}{\widetilde{s}_o}\link{}{\dots} \chainend{p_{j+1}}.P_{aj} ;$$
 
@@
  the process $P_{aj}$ provides a code of this type: 
-$$P_{aj}\defeq \link{r_j}{\dots}\chainedlink{s_i}{\noact}\chainedlink{\noact}{}\link{s_o}{\dots} \chainedlink{r_{j+1}}{\noact}\chainedlink{\noact}{p_j}\link{}{\dots}\link{}{p_{j+1}}.P_{aj} ;$$
+$$P_{aj}\defeq \link{r_j}{\dots}\chainedlink{s_i}{\noact}\chainedlink{\noact}{s_o}\link{}{\dots} \chainedlink{r_{j+1}}{\noact}\chainedlink{\noact}{p_j}\link{}{\dots}\link{}{p_{j+1}}.P_{aj} ;$$
 
@@
  the process $P_{aj}$  provides a code of this type: 
-$$P_{aj}\defeq \link{r_j}{\dots}\chainedlink{\overline{s}_i}{\noact}\chainedlink{\noact}{}\link{\overline{s}_o}{\dots} \chainedlink{r_{j+1}}{\noact}\chainedlink{\noact}{p_j}\link{}{\dots}\link{}{p_{j+1}}.P_{aj} ;$$
+$$P_{aj}\defeq \link{r_j}{\dots}\chainedlink{\overline{s}_i}{\noact}\chainedlink{\noact}{\overline{s}_o}\link{}{\dots} \chainedlink{r_{j+1}}{\noact}\chainedlink{\noact}{p_j}\link{}{\dots}\link{}{p_{j+1}}.P_{aj} ;$$
 
@@
 \begin{enumerate}
-\item $P_{aj}\defeq \link{r_j}{\dots}\chainedlink{\overline{s}_i}{\noact}\chainedlink{\noact}{}\link{\overline{s}_o}{\dots} \chainedlink{r_{j+1}}{\noact}\chainedlink{\noact}{p_j}\link{}{\dots}\link{}{p_{j+1}}.P_{aj} $;
-\item $P_{aj}\defeq \link{r_j}{\dots}\chainedlink{s_i}{\noact}\chainedlink{\noact}{}\link{s_o}{\dots} \chainedlink{r_{j+1}}{\noact}\chainedlink{\noact}{p_j}\link{}{\dots}\link{}{p_{j+1}} .P_{aj}$.
+\item $P_{aj}\defeq \link{r_j}{\dots}\chainedlink{\overline{s}_i}{\noact}\chainedlink{\noact}{\overline{s}_o}\link{}{\dots} \chainedlink{r_{j+1}}{\noact}\chainedlink{\noact}{p_j}\link{}{\dots}\link{}{p_{j+1}}.P_{aj} $;
+\item $P_{aj}\defeq \link{r_j}{\dots}\chainedlink{s_i}{\noact}\chainedlink{\noact}{s_o}\link{}{\dots} \chainedlink{r_{j+1}}{\noact}\chainedlink{\noact}{p_j}\link{}{\dots}\link{}{p_{j+1}} .P_{aj}$.
 \end{enumerate}
@@
 \link{\mathit{cxt}}{\dots}
-\chainedlink{\widehat{s}_i}{\noact}\chainedlink{\noact}{}\link{\widehat{s}_o}{\dots} 
+\chainedlink{\widehat{s}_i}{\noact}\chainedlink{\noact}{\widehat{s}_o}\link{}{\dots} 
 \chainedlink{\underline{e}_i}{\noact}\chainedlink{\noact}{}\link{\underline{e}_o}{\dots} 

### 2020-06-20 TCS_version/00-abstract.tex
secondo revisore

% !TEX root =  main_TCS.tex

In the area of Natural Computing, reaction systems are
a qualitative abstraction inspired by the 
functioning of living cells, suitable to model the main
mechanisms of biochemical reactions.
This model has already been applied
and extended successfully to various areas of research. Reaction 
systems interact with the environment represented by the context, and
pose challenges for further extensions as well as
for the implementation, as it is a new computation model. 
In this paper {\color{red} we define a modular encoding of Reaction Systems as processes
in the chained Core Network Algebra (cCNA), which is a variant  of the  {\tt link}-calculus,
by representing the behaviour of each entity and
preserving faithfully their features, and we prove its correctness and completeness.}
%we consider a variant of the {\tt link}-calculus, which allows to model 
%multiparty interaction in concurrent systems, and show that we can
%represent reaction systems as processes 
%in a modular way, by representing the behaviour of each entity and
%preserving faithfully their features.
%We show the correctness and completeness 
%of our embedding.
We illustrate our framework by showing the embedding of 
a few examples expressing computer science
and biological applications. 
{\color{red}Our embedding provides a Labelled Transition System (LTS) semantics for reaction systems. Based on the LTS semantics, we adapt  the classical notion of  process bisimulation to define a bio-simulation relation for studying properties of reaction systems.}
%On top of the LTS semantics for reaction systems provided by our framework, 
%we then adapt the classical notion (in concurrency) of bisimulation
%to make it more suitable for studying properties of reaction systems.
In particular, we define a new assertion language based on regular expressions, 
which allows to specify the properties of interest, and use it to extend Hennessy-Milner logic
to our framework. 
{\color{red} We prove that our bio-simulation relation and the logical equivalence, that are defined parametrically on some assertion of interest, coincide.}
%Interestingly, the novel notion of bisimilarity and logical equivalence that are defined parametrically on some assertion of interest are proved to coincide, like in the classical case.
Finally, 
{\color{red} we claim that}
our methodology 
%can 
contribute to increase the expressiveness
of reaction systems, by exploiting the interaction among 
different reaction systems.

### 2020-06-20 TCS_version/01-introduction.tex
secondo revisore

@@
 %Formally, Reaction Systems are defined as a rewriting formalism.
-In this paper, we present an encoding from RSs, 
-to the open multiparty process algebra c\CNA,\footnote{After `chained Core Network 
-Algebra'.} a variant of the {\tt link}-calculus~\cite{BodeiBB12,BBB17} without name mobility.
+%In this paper, we present an encoding from RSs, 
+%to the  open multiparty process algebra c\CNA,\footnote{After `chained Core Network 
+%Algebra'.} a variant of the {\tt link}-calculus~\cite{BodeiBB12,BBB17} without name mobility.
+{\color{red}In this paper, we present an encoding from RSs, 
+to the chained Core Network  Algebra (c\CNA), first defined in~\cite{BBF19}, a variant of the 
+ open multiparty process algebra~\CNA~\cite{BBB17}; the \CNA~equipped with mobility is referred to as the {\tt link}-calculus~\cite{BODEI2020104587}}.
 This formalism allows several processes to synchronise and 
@@
 mechanism based on links and link chains. 
-The initial motivation for introducing this mechanism was to encode
+The initial motivation for introducing
+{\color{red} an open multiparty mechanism}
+% this mechanism 
+was to encode
 Mobile Ambients~\cite{CardelliG00}, getting a much stronger operational
@@
 notion of bisimilarity which takes into account the characteristics
-of biological systems. We define a new {\em bio-simulation} having in mind
+of biological systems. We define a new {\em bio-simulation} relation having in mind
 the possibility that two interacting systems may be compared w.r.t.
@@
 %way that a mutation $s'$ can be expressed; 
-in such a way that the mutated entity
-$s'$ can  take part to only a subset of rules requiring entity $s$;
+in such a way that {\color{red} the mutated version of the entity
+$s$ can}  take part to only a subset of rules requiring entity $s$;
 \item we show that with a little coding effort, our encoding allows two RSs to
@@
 Example~\ref{ex:twoRS} and more in general the discussion in 
-Section~\ref{sec:discussion} show that 
+Section~\ref{subsec:comRS}
+%~\ref{sec:discussion}
+ show that 
 the interaction between RSs can help to model new scenarios.
@@
 biological processes, see~\cite{BBDFH18} for a recent survey.
-We are not aware of any SOS operational semantics for the RS; it seems not the case
+We are not aware of any 
+%SOS 
+structural operational semantics for the RS; it seems not the case
 that a deterministic behaviour, as the one of the RS, once the initial state is 

### 2020-06-20 TCS_version/02-ReactionSystem.tex
secondo revisore

@@
-% !TEX root = ./main_RS2L_LNCS.tex
+% !TEX root = ./main_TCS.tex
 
-\section{Reaction Systems}
+\section{An overview of Reaction Systems}
 \label{sec:rs}
@@
 
-The sets $R, I, P$ are also written $R_a, I_a, P_a$ and called the 
+{\color{red} The set $S$ is called the entity set}, the sets 
+$R, I, P$ are also written $R_a, I_a, P_a$ and called the 
 \emph{reactant set} of $a$, the  
@@
 A Reaction System (RS) is an ordered pair ${\cal A} = (S, A)$ 
-such that $S$ is a finite set, and $A \subseteq rac(S)$.
+such that $S$ is a finite sete {\color{red} of  entities, and $A \subseteq rac(S)$ is a set over $S$}.
 %The set $S$ is called the background (set) of $A$.\\
@@
 \begin{definition}[Reaction Result]
-Let $W$ be a finite subset of $S$.
+Let $W$ be a finite subset of  entities $S$.
 \begin{enumerate}
-\item Let $a$ be a reaction in $S$. Then $a$ is enabled by $W$, denoted by $\mathit{en}_a(W)$, 
-if $R_a \subseteq W$
-and $I_a \cap W = \emptyset$. The result of $a$ on $W$, denoted by $\mathit{res}_a(W)$, 
+\item Let $a$ be a reaction in $S$. Then $a$ is enabled by the entities in the set $W$, denoted by $\mathit{en}_a(W)$, 
+if $R_a \subseteq W$, i.e. all the reactants of $a$ are in $W$,
+and $I_a \cap W = \emptyset$, i.e. none of the inhibitors of $a$ are in $W$. The result of $a$ on $W$, denoted by $\mathit{res}_a(W)$, 
 is defined by:
@@
 \begin{definition}[Interactive Process]
-Let ${\cal A} = (S, A)$ be a RS and let $n \geq 0$. 
+Let ${\cal A} = (S, A)$ be a RS and let $n$
+% \geq 0$.
+be a nonnegative integer;   
 An $n$-step

### 2020-06-20 TCS_version/04-CCNA.tex
secondo revisore

--- 2020-04-04 04-CCNA.tex
@@
 it is \emph{solid} if $\alpha,\beta \neq \ \noacts$; 
-the link $\link{\noact}{\noact}$ is called \emph{virtual}.
+{\color{red} $\alpha$ and $\beta$ are two interaction points, to fix the idea $\alpha$ is the action, and $\beta$ is the coaction.}
+The link $\link{\noact}{\noact}$ is called \emph{virtual}.
 A link is \emph{valid} if it is solid or virtual.
@@
 % Again,  all the channel names appearing in the half links or in the open blocks of chain are not required to match.
- For example, for $R = \{a,b\}$ the expression $ \blockchain{c_i}{c_o}{c \in R} $\;
-denotes the block of chains ${}_{a_i}^{\noact}\chainedlink{\noact}{a_o} \backslash{}_{b_i}^{\noact}\chainedlink{\noact}{b_o}$; and the expression $\link{r_1}{} \blockchain{c_i}{c_o}{c \in R}\;\link{}{r_2} $ denotes the chain \\
+ For example, for $R = \{a,b\}$ the expression $ \blockchain{c_i}{c_o}{c \in R} $
+denotes the block ${}_{a_i}^{\noact}\chainedlink{\noact}{a_o} \backslash{}_{b_i}^{\noact}\chainedlink{\noact}{b_o}$; and the expression $\link{r_1}{} \blockchain{c_i}{c_o}{c \in R}\link{}{r_2} $ denotes the chain
 $ \startchain{r_1}\chainedlink{a_i}{\noact}\chainedlink{\noact}{a_o}\chainedlink{b_i}{\noact}\chainedlink{\noact}{b_o}\chainend{r_2}$.

### 2020-06-20 TCS_version/05-translation.tex
secondo revisore

@@
 Given an entity $s \in S$, we exploit five different pairs of names for the interactions over $s$:
-names $s_i$, $s_o$ are used to test the presence of $s$ in the system; names $\widehat{s}_i$, $\widehat{s}_o$ are used to test the provisioning of $s$ from the context;
-names $\widetilde{s}_i$, $\widetilde{s}_o$  are used to test the production of $s$ by some reaction; names $\overline{s}_i$, $\overline{s}_o$ are used to test the absence of $s$ in the system; and names $\underline{s}_i$, $\underline{s}_o$ are used to test the absence of $s$ from the context.
+\begin{itemize}
+\item names $s_i$, $s_o$ are used to test the presence of $s$ in the system; 
+\item names $\widehat{s}_i$, $\widehat{s}_o$ are used to test the provisioning of $s$ from the context;
+\item names $\widetilde{s}_i$, $\widetilde{s}_o$  are used to test the production of $s$ by some reaction; 
+\item names $\overline{s}_i$, $\overline{s}_o$ are used to test the absence of $s$ in the system; 
+\item  names $\underline{s}_i$, $\underline{s}_o$ are used to test the absence of $s$ from the context.
+\end{itemize}
 We let $P_s$ be the process implementing the  presence of $s$ in the system, and $\overline{P_s}$ be the one for its absence.
@@
 \paragraph{Processes for reactions}
-We assume that each reaction $a$ is assigned a progressive number $j$.
+We assume that {color{red} all the reactions $a$ are ordered by assign to them a progressive  number $j$
+from $1$ to $n$.}
+%each reaction $a$ is 
+%assigned a progressive number $j$.
 The process for reaction $aj=(R_j,I_j,P_j)$ must assert either the possibility to apply the reaction or its impossibility.
@@
 %For notational convenience, we fix that $r_1 = \silent$,  $r_{u+1} = cxt_1$ for $u$ the number of reacts, and $cxt_{w+1} = \silent$ for $w$ the number of entities.
-For technical reasons, we introduce the init process $I \defeq \link{\silent}{r_1}.I$
+For technical reasons, we introduce the {\color{red} recursive} init  process $I \defeq \link{\silent}{r_1}.I$
 to allow the name $r_1$ to be matched at the start of any chain; also, 
@@
  \begin{array}{rcl}
- P_{r_1} &\defeq& \startchain{r_1}\chainedlink{s1_i}{\noact}\chainedlink{\noact}{s1_o}\chainedlink{r_2}{\noact} \chainedlink{\noact}{p_1}\chainedlink{\widetilde{s2}_i}{\noact}\chainedlink{\noact}{\widetilde{s2}_o}\chainend{p_2}.P_{r_1} + \dots;   \\[6pt]
-  P_{r_2} &\defeq& \dots + \startchain{r_2}\chainedlink{\overline{s2}_i}{\noact}\chainedlink{\noact}{\overline{s2}_o}\chainedlink{\mathit{cxt}}{\noact} \chainedlink{\noact}{p_2}\chainend{\silent}.P_{r_2} + \dots; \\[6pt]
+ P_{r_1} &\defeq& \startchain{r_1}\chainedlink{s1_i}{\noact}\chainedlink{\noact}{s1_o}\chainedlink{r_2}{\noact}\chainedlink{\noact}{p_1}\chainedlink{\widetilde{s2}_i}{\noact}\!\!\chainedlink{\noact}{\widetilde{s2}_o}\chainend{p_2}.P_{r_1} + \dots;   \\[6pt]
+  P_{r_2} &\defeq& \dots + \startchain{r_2}\chainedlink{\overline{s2}_i}{\noact}\chainedlink{\noact}{\overline{s2}_o}\chainedlink{\mathit{cxt}}{\noact}\chainedlink{\noact}{p_2}\chainend{\silent}.P_{r_2} + \dots; \\[6pt]
    P_{s1} &\defeq& \startchain{s1_i}\chainedlink{s1_o}{\noact}\chainedlink{\noact}{\widehat{s1}_i}\chainend{\widehat{s1}_o}
@@
   \overline{P}_{s2} &\defeq& \startchain{\overline{s2}_i}\chainedlink{\overline{s2}_o}{\noact}
-   \chainedlink{\noact} {\widehat{s2}_i}\chainedlink{\widehat{s2}_o}{\noact}\chainedlink{\noact} {\widetilde{s2}_i}\chainend{\widetilde{s2}_o}.P_{s2} + \dots;\\[6pt]
+   \chainedlink{\noact} {\widehat{s2}_i}\chainedlink{\widehat{s2}_o}{\noact}\!\!\chainedlink{\noact} {\widetilde{s2}_i}\chainend{\widetilde{s2}_o}.P_{s2} + \dots;\\[6pt]
  \mathit{Cxt}& \defeq & \startchain{\mathit{cxt}}\chainedlink{\widehat{s1}_i}{\noact}\chainedlink{\noact}{\widehat{s1}_o}
-  \chainedlink{\widehat{s2}_i}{\noact}\chainedlink{\noact}{\widehat{s2}_o}\chainend{p_1}.\mathit{Cxt}
+  \chainedlink{\widehat{s2}_i}{\noact}\!\chainedlink{\noact}{\widehat{s2}_o}\chainend{p_1}.\mathit{Cxt}
   %Cxt& \defeq & \startchain{cxt}\chainedlink{\hat{s1}_i}{\noact}\chainedlink{\noact}{\hat{s1}_o}

### 2020-06-20 TCS_version/06b-bisimulation.tex
secondo revisore

@@
 There are many variants of the bisimulation for process algebras,  for example the
-barbed bisimulation~\cite{10.1007/3-540-55719-9_114} only consider the execution of invisible actions, and then equates two processes when they expose the same prefixes; for the mobile ambients~\cite{CardelliG00}, a process algebra equipped with a reduction semantics, a notion of
+barbed bisimulation~\cite{10.1007/3-540-55719-9_114} only considers the execution of invisible actions, and then equates two processes when they expose the same prefixes; for the mobile ambients~\cite{CardelliG00}, a process algebra equipped with a reduction semantics, a notion of
 behavioural equivalence equates two processes when they expose the same ambients~\cite{GC03}.
@@
 
-As usual, the semantics of a bioHML formula is the set of processes that satisfy it.
+As usual, the semantics of a bioHML formula is the set of c\CNA~processes that satisfy it.
 
@@
 We observe that, as expected, for any bioHML formula $G$ and process $P$ we have $\overline{\overline{\mathsf{G}}} = \mathsf{G}$ and  $P \entails \overline{\mathsf{G}}$ iff  $P\not\entails \mathsf{G}$.
-
+\begin{definition}
 We let $\mathcal{L}_{\mathsf{F}}$ be the set of all bioHML formulas that respects $\mathsf{F}$ and 
-we say that two processes $P,Q$ are bio-logically equivalent w.r.t.
-$\mathsf{F}$, written $P \equiv_{\mathcal{L}_{\mathsf{F}}}Q$, when $P$ and $Q$ satisfy the exactly the same bioHML formulas in $\mathcal{L}_{\mathsf{F}}$, i.e.
+we say that two processes $P,Q$ are \emph{bio-logically equivalent w.r.t.
+$\mathsf{F}$}, written $P \equiv_{\mathcal{L}_{\mathsf{F}}}Q$, when $P$ and $Q$ satisfy exactly the same bioHML formulas in $\mathcal{L}_{\mathsf{F}}$, i.e.
 when for any $\mathsf{G}\in \mathcal{L}_{\mathsf{F}}$ we have $P \entails \mathsf{G}\ \Leftrightarrow Q \entails \mathsf{G}$.
+\end{definition}
 

### 2020-06-20 TCS_version/07-discussion.tex
secondo revisore

--- 2020-04-05 07-discussion.tex
@@
 \subsection{Communicating Reaction Systems}
+\label{subsec:comRS}
 We sketch how it is possible to program two RSs encodings, in a way that
-the entities that usually come from context of one RS will be provided instead from the other RS.
+the entities that usually come from the context of one RS will be provided instead from the other RS.
 
@@
 \noindent
-The translation for the $rs2$ follows.
+The translation for $rs2$ follows.
 

### 2020-06-22 TCS_version/01-introduction.tex
aggiunta biblio

@@
 the basic RS framework, as it can only happen with the context.
+{\color{red}The proposed network of reaction systems~\cite{BLR20} allows
+any reaction system in the network to receive entities 
+produced from its neighbors (that will represent its context), with the hypothesis that  
+reaction systems without neighbors will receive no entities 
+from the context. 
+}
 By exploiting recursion, the kind of interactions which can be 

### 2020-06-24 TCS_version/01-introduction.tex
altri piccoli cambiamenti

@@
 \item we define an assertion language to specify local properties of 
-Reaction Systems and exploit it to define a novel notion of behavioural equivalence for Reaction Systems, called bio-similarity;
+Reaction Systems and exploit it to define a novel notion of behavioural equivalence for 
+RSs,
+%Reaction Systems, 
+called bio-similarity;
 \item we show that our assertions can be used to extend the Hennessy-Milner logic to our framework, in such a way that logical equivalence of processes coincide with bio-similarity;
@@
 from the context. 
+In our case, a RS can receive some entities from the context and some others from a second RS, in any case we always assume contextes having a non-determinist behaviour.
 }
@@
 and give more detailed explanations about its use;
-\item we introduce an assertion language to specify the properties of Reaction Systems;
-\item we define the notion of bio-simulation to relate Reaction Systems and compare their behaviours;
+\item we introduce an assertion language to specify the properties of RSs;
+%Reaction Systems;
+\item we define the notion of bio-simulation to relate RSs
+%Reaction Systems 
+and compare their behaviours;
 \item we show that our assertion language allows to specify properties extending to our

### 2020-06-24 TCS_version/08-conclusion.tex
altri piccoli cambiamenti

--- 2020-04-05 08-conclusion.tex
@@
 %shown 
-a faithful embedding of Reaction Systems,
+a faithful embedding of 
+RSs,
+%Reaction Systems,
 an emerging formalism to model computationally biochemical systems.
@@
 We have then modified the classical notion (in concurrency) of bisimulation
-making it more suitable for proving properties of Reaction Systems in our
+making it more suitable for proving properties of 
+RSs
+%Reaction Systems 
+in our
 framework.
@@
 We are confident that our embedding can contribute to extend the applications
-of Reaction Systems to diverse fields of computer science, and life
+of 
+RSs
+%Reaction Systems
+to diverse fields of computer science, and life
 sciences.

### 2020-06-25 TCS_version/00-abstract.tex
embedding encoding translation

@@
 for the implementation, as it is a new computation model. 
-In this paper {\color{red} we define a modular encoding of Reaction Systems as processes
+In this paper {\color{red} we define a modular embedding of Reaction Systems as processes
 in the chained Core Network Algebra (cCNA), which is a variant  of the  {\tt link}-calculus,
@@
 %of our embedding.
-We illustrate our framework by showing the embedding of 
+%We illustrate our framework by 
+{\color{red} We show} the embedding of 
 a few examples expressing computer science
@@
 which allows to specify the properties of interest, and use it to extend Hennessy-Milner logic
-to our framework. 
+to our {\color{red} embedding.}
+% framework. 
 {\color{red} We prove that our bio-simulation relation and the logical equivalence, that are defined parametrically on some assertion of interest, coincide.}

### 2020-06-25 TCS_version/01-introduction.tex
embedding encoding translation

@@
 %Algebra'.} a variant of the {\tt link}-calculus~\cite{BodeiBB12,BBB17} without name mobility.
-{\color{red}In this paper, we present an encoding from RSs, 
+{\color{red}In this paper, we present an embedding
+%encoding 
+from RSs, 
 to the chained Core Network  Algebra (c\CNA), first defined in~\cite{BBF19}, a variant of the 
@@
 called bio-similarity;
-\item we show that our assertions can be used to extend the Hennessy-Milner logic to our framework, in such a way that logical equivalence of processes coincide with bio-similarity;
+\item we show that our assertions can be used to extend the Hennessy-Milner logic to our 
+%framework
+{\color{red}embedding}, in such a way that logical equivalence of processes coincide with bio-similarity;
 \item we sketch how one can express the behaviour of entity mutation,
@@
 $s$ can}  take part to only a subset of rules requiring entity $s$;
-\item we show that with a little coding effort, our encoding allows two RSs to
+\item we show that with a little coding effort, our 
+%encoding 
+{\color{red} embedding}
+allows two RSs to
 communicate; i.e. we can model scenarios where a subset of those entities that the context  can
@@
 
-The main drawback of our proposal, is that the c\CNA \ translation is
+The main drawback of our proposal, is that the {\color{red} resulting c\CNA \ 
+code}
+%translation 
+is
 verbose. Nevertheless it is clear that our
-translation can be automatised by means of a proper front-end in
+%translation 
+{\color{red} embedding}
+can be automatised by means of a proper front-end in
 an implementation of the {\tt link}-calculus. 
@@
 
-As we have remarked, in our translation, Reaction Systems 
+As we have remarked, in our 
+%translation,
+{\color{red} embedding}
+ Reaction Systems 
 get the ability to interact between them in a synchronized manner. 
@@
 \begin{itemize}
-\item we present several new examples to illustrate our framework,
+\item we present several new examples to illustrate our 
+{\color{red} embedding},
+%framework,
 and give more detailed explanations about its use;
@@
 \item we show that our assertion language allows to specify properties extending to our
-framework the Hennessy-Milner logic;
+{\color{red} embedding}
+%framework 
+the Hennessy-Milner logic;
 \item we include here all proofs of main results.

### 2020-06-25 TCS_version/04-CCNA.tex
altri piccoli cambiamenti
embedding encoding translation

@@
 Hereafter we make use of some new notations for link chains that will facilitate  
-the presentation of our translation.
+the presentation of our 
+%translation.
+{\color{red} embedding.}
 
@@
 Let $R$ be a set of names. We 
-define an \emph{open block} as $\blockchain{a_i}{a_o}{a \in R}$\ , where $a_i$ and $a_o$ are annotated version of the name $a$, as
+define an \emph{open block} as $\blockchain{a_i}{a_o}{a \in R}$, where $a_i$ and $a_o$ are annotated version of the name $a$, as
 \[

### 2020-06-25 TCS_version/05-translation.tex
embedding encoding translation

@@
 \label{sec:trans}
-Here we present a translation from Reaction Systems to c\CNA.
+Here we present {\color{red} an embedding of}
+%a translation from 
+Reaction Systems {\color{red} into} c\CNA.
 The idea is to define separated processes for representing the behaviour of each entity, each reaction, 
@@
 
-\begin{definition}[Translation]
+\begin{definition}[{\color{red} Embedding}]
+%Translation]
 \label{def:trans}
 Let ${\cal A} = (S, A)$ be a RS, and let $\pi=(\gamma,\delta)$ be an extended interactive process in ${\cal A}$, with  $\gamma=\{C_i\}_{i\in\mathbb{N}}$. 
-We define its c\CNA~translation $\llbracket{\cal A},\gamma \rrbracket$ as follows: 
+We define its c\CNA~{\color{red} embedding}
+%translation 
+$\llbracket{\cal A},\gamma \rrbracket$ as follows: 
 $$
@@
 
-It is important to observe that,   for each transition, our c\CNA \ encoding requires  all the processes 
+It is important to observe that,   for each transition, our c\CNA \ 
+%encoding 
+{\color{red} embedding}
+requires  all the processes 
 %$P_a$, with $a \in A$, and  $\mathit{Cxt}^n$ and  $P_s$, with $s \in S$,  
@@
  %\label{lem:struct}
-  Let ${\cal A} = (S, A)$ be a RS and let $\pi=(\gamma,\delta)$ be an extended  interactive process in ${\cal A}$.
-Let $P=\llbracket{\cal A},\gamma\rrbracket$ its c\CNA~translation.
+  Let ${\cal A} = (S, A)$ be a RS and let $\pi=(\gamma,\delta)$ be an extended  interactive process in ${\cal A}$. Let $P=\llbracket{\cal A},\gamma\rrbracket$ its c\CNA~{\color{red} embedding.}
+  %translation. 
   If $exists$ $P'$ such that $P \xrightarrow{\restrict{\mathit{names}}\upsilon} P'$ is a transition of $P$, then
@@
 
-We will then use $\llparenthesis{\cal A},\gamma\rrparenthesis$ to refer to the encoding without topmost name restrictions, i.e.,
+We will then use $\llparenthesis{\cal A},\gamma\rrparenthesis$ to refer to the 
+%encoding 
+{\color{red} embedding}
+without topmost name restrictions, i.e.,
 $$

### 2020-06-25 TCS_version/06-examples.tex
reviewer 1
secondo revisore
embedding encoding translation

@@
 \noindent
-In our $c\CNA$ encoding, $q$, $w$, $a$, and $b$ became the $entities$, the context can only provide $a$ and $b$ and the reaction rules are as follows:\\
+In our $c\CNA$ 
+%encoding,
+{\color{red} embedding,} $q$, $w$, $a$, and $b$ became the $entities$, the context can only provide $a$ and $b$ and the reaction rules are as follows:\\
 \[
@@
 \paragraph{Encoding of the rules}
-The encoding of the rules for reactions is  given in a parametric way:
+The encoding of the rules for reactions is  given in a parametric way, {\color{red} with $n \in \{1,2,3,4\}$}:
 \[
@@
 P (T,\widetilde{T})&\defeq & \sum_{h=0}^1 (\startchain{T_i}\chainedlink{T_o}{\noact}\chainend{\noact})^h\ \link{\widetilde{T}_i}{\widetilde{T}_o}.P (T,\widetilde{T})
-& +  && \link{T_i}{T_o}.P (\overline{T},\widetilde{T}) \\
+& +  & \link{T_i}{T_o}.P (\overline{T},\widetilde{T}) \\
 %\overline{P} (q)&\defeq &  \sum_{h=0}^1 (\startchain{\overline{T}_i}\chainedlink{\overline{T}_o}{\noact}\chainend{\noact})^h\link{\widetilde{T}_i}{\widetilde{T}_o}.P(q)\\
@@
 P_F&\defeq & \sum_{h=0}^1 (\startchain{F_i}\chainedlink{F_o}{\noact}\chainend{\noact})^h\ \link{\widehat{F}_i}{\widehat{F}_o}.P_F
-& +  && \sum_{h=0}^1 (\startchain{F_i}\chainedlink{F_o}{\noact}\chainend{\noact})^h. \link{\underline{F}_i}{\underline{F}_o}.\overline{P}_F \\[6pt]
+& +  & \sum_{h=0}^1 (\startchain{F_i}\chainedlink{F_o}{\noact}\chainend{\noact})^h\ \link{\underline{F}_i}{\underline{F}_o}.\overline{P}_F \\[6pt]
 \overline{P}_F&\defeq & \sum_{h=0}^1 (\startchain{\overline{F}_i}\chainedlink{\overline{F}_o}{\noact}\chainend{\noact})^h\ \link{\widehat{F}_i}{\widehat{F}_o}.P_F
-& +  && \sum_{h=0}^1 (\startchain{\overline{F}_i}\chainedlink{\overline{F}_o}{\noact}\chainend{\noact})^h. \link{\underline{F}_i}{\underline{F}_o}.\overline{P}_F \\
+& +  & \sum_{h=0}^1 (\startchain{\overline{F}_i}\chainedlink{\overline{F}_o}{\noact}\chainend{\noact})^h\ \link{\underline{F}_i}{\underline{F}_o}.\overline{P}_F \\
 %\overline{P} (q)&\defeq &  \sum_{h=0}^1 (\startchain{\overline{T}_i}\chainedlink{\overline{T}_o}{\noact}\chainend{\noact})^h\link{\widetilde{T}_i}{\widetilde{T}_o}.P(q)\\

### 2020-06-25 TCS_version/07-discussion.tex
altri piccoli cambiamenti
embedding encoding translation

@@
-% !TEX root = ./main_RS2L_LNCS.tex
+% !TEX root = ./main_TCS.tex
 %\section{Discussion}
@@
 the possibility of alternative behaviour of
-mutated entities, and  the communication between two different Reaction Systems.
+mutated entities, and  the communication between two different 
+RSs.
+%Reaction Systems.
 It is important to note that, when the context is deterministic, our encoding guarantees that from each state, 
@@
 %
-We need to modify our translation technique to suite this new setting. 
+We need to modify our {\color{red} embedding}
+%translation 
+technique to suite this new setting. 
 As we do not model contexts, we introduce  \emph{dummy} channel names $dx$ and $dss$ to model the absence of entities.
@@
 Please note, that prefixes of process $P_{a_1}$ end with the channel name $a_2$, as the link chain is now connected with the  reaction of  $rs2$.
-The translation for the entities follows.
+The {\color{red} encoding} 
+%translation 
+for the entities follows.
 \[
@@
 \noindent
-The translation for $rs2$ follows.
+The {\color{red} encoding}
+%translation 
+for $rs2$ follows.
 
@@
 \noindent
-In the translation of the entities in  $rs2$, we introduce the mechanism that allows the entity
+In the {\color{red} encoding}
+%translation 
+of the entities in  $rs2$, we introduce the mechanism that allows the entity
 $s$ ($ss$ in $rs2$) to be provided in $rs1$. Every time $ss$ is produced in $rs2$,  a virtual link is

### 2020-06-25 TCS_version/08-conclusion.tex
embedding encoding translation

@@
 %into multilink processes.  
-This translation shows several benefits.
+This {\color{red} embedding}
+%translation 
+shows several benefits.
 For instance, in our paper the context of a RS can work non deterministically and it 
@@
 in our
-framework.
+%framework.
+{\color{red} embedding.}
 We have defined a new assertion language, which allows us to specify
 the properties to be verified, and have shown that it extends Hennessy-Milner logic
-to our framework.
+to our 
+%framework.
+{\color{red} embedding.}
 We believe that our work can make 
@@
 
-As future work, we plan to implement a prototype of our framework,
+As future work, we plan to implement a prototype of our 
+%framework,
+{\color{red} embedding,}
 with an automatic translation from RSs to  {\tt link}-calculus. 

### 2020-06-29 TCS_version/00-abstract.tex
minor updates

@@
 
-In the area of Natural Computing, reaction systems are
+In the area of Natural Computing, Reaction Systems (RSs) are
 a qualitative abstraction inspired by the 
@@
 mechanisms of biochemical reactions.
-This model has already been applied
-and extended successfully to various areas of research. Reaction 
-systems interact with the environment represented by the context, and
-pose challenges for further extensions as well as
-for the implementation, as it is a new computation model. 
-In this paper {\color{red} we define a modular embedding of Reaction Systems as processes
-in the chained Core Network Algebra (cCNA), which is a variant  of the  {\tt link}-calculus,
-by representing the behaviour of each entity and
-preserving faithfully their features, and we prove its correctness and completeness.}
+%This model has already been applied
+%and extended successfully to various areas of research. 
+RSs interact with a context, and
+pose challenges for modularity, compositionality, extendibility and behavioural equivalence. 
+In this paper we define a modular encoding of RSs as processes
+in the chained Core Network Algebra (cCNA), which is a new variant  of the  {\tt link}-calculus.
+The encoding represents the behaviour of each entity separately and
+preserves faithfully their features, and we prove its correctness and completeness.
 %we consider a variant of the {\tt link}-calculus, which allows to model 
@@
 %We illustrate our framework by 
-{\color{red} We show} the embedding of 
-a few examples expressing computer science
-and biological applications. 
-{\color{red}Our embedding provides a Labelled Transition System (LTS) semantics for reaction systems.
-Based on the LTS semantics, we adapt  the classical notion of  process bisimulation to define a bio-simulation relation for studying properties of reaction systems.}
+%{\color{red} We show} the embedding of 
+%a few examples expressing computer science
+%and biological applications. 
+Our encoding provides a Labelled Transition System (LTS) semantics for RSs.
+Based on the LTS semantics, we adapt  the classical notion of  bisimulation to define a novel equivalence, called bio-similarity, for studying properties of RSs.
 %On top of the LTS semantics for reaction systems provided by our framework, 
@@
 which allows to specify the properties of interest, and use it to extend Hennessy-Milner logic
-to our {\color{red} embedding.}
+to our setting.
 % framework. 
-{\color{red} We prove that our bio-simulation relation and the logical equivalence, that are defined parametrically on some assertion of interest, coincide.}
+We prove that our bio-similarity relation and the logical equivalence, that are defined parametrically on some assertion of interest, coincide.
 %Interestingly, the novel notion of bisimilarity and logical equivalence that are defined parametrically on some assertion of interest are proved to coincide, like in the classical case.
 Finally, 
-{\color{red} we claim that}
-our methodology 
+we claim that
+our encoding 
 %can 
-contribute to increase the expressiveness
-of reaction systems, by exploiting the interaction among 
-different reaction systems.
+contributes to increase the expressiveness
+of RSs, by exploiting the interaction among 
+different RSs.

### 2020-06-29 TCS_version/01-introduction.tex
minor updates

@@
 
-Natural Computing is an emerging area of research which has two main 
+Natural Computing is an active area of research which builds on two main 
 aspects: human designed computing inspired by nature, and computation 
@@
 This theory has already shown to be relevant in several different 
-fields, such as computer science~\cite{MPR15}, 
-biology~\cite{ABP14,CMMBM12,Az17,BarbutiGLM16}, 
-molecular chemistry~\cite{OY16}.
+fields, 
+%such as computer science~\cite{MPR15}, 
+including biology~\cite{ABP14,CMMBM12,Az17,BarbutiGLM16}
+and molecular chemistry~\cite{OY16}.
 %,
@@
 %international meetings. 
-Reaction Systems formalise the mechanisms of biochemical systems, 
+RSs formalise the mechanisms of biochemical systems, 
 such as {\em facilitation} and {\em inhibition}. 
@@
 %Algebra'.} a variant of the {\tt link}-calculus~\cite{BodeiBB12,BBB17} without name mobility.
-{\color{red}In this paper, we present an embedding
+In this paper, which is an extended version of~\cite{BBF19},
+we present an encoding
 %encoding 
-from RSs, 
-to the chained Core Network  Algebra (c\CNA), first defined in~\cite{BBF19}, a variant of the 
- open multiparty process algebra~\CNA~\cite{BBB17}; the \CNA~equipped with mobility is referred to as the {\tt link}-calculus~\cite{BODEI2020104587}}.
+from RSs
+to the chained Core Network  Algebra (c\CNA), a variant of the 
+ open multiparty process algebra~\CNA~\cite{BBB17}; the \CNA~equipped with mobility is referred to as the {\tt link}-calculus~\cite{BodeiBB12,BODEI2020104587}.
+ Here mobility is not needed.
 This formalism allows several processes to synchronise and 
@@
 The initial motivation for introducing
-{\color{red} an open multiparty mechanism}
+an open multiparty mechanism in~\cite{BodeiBB12}
 % this mechanism 
@@
 \begin{itemize}
-\item the behaviour of the context, for each single 
-entity,
-%\footnote{Entities that belong to the `domain' of a RS.}
-can be
-specified in a recursive way, as ordinary processes;
-\item along the same lines, non-deterministic contexts can be seamlessly expressed in our setting; 
-\item when deterministic contexts are considered, the c\CNA \
-computation is deterministic and mimics the evolution of the corresponding RS;
+\item any context is represented as an ordinary c\CNA \ process, which allows us to specify recursive and non-deterministic contexts in a natural way; 
+when deterministic contexts are considered, as in ordinary RSs, the c\CNA \
+computation is also deterministic and matches the evolution of the underlying RS;
 \item we define an assertion language to specify local properties of 
-Reaction Systems and exploit it to define a novel notion of behavioural equivalence for 
-RSs,
+RSs and exploit it to define a novel notion of behavioural equivalence,
 %Reaction Systems, 
 called bio-similarity;
-\item we show that our assertions can be used to extend the Hennessy-Milner logic to our 
+\item we show that our assertions can be used to extend the Hennessy-Milner's logic to our 
 %framework
-{\color{red}embedding}, in such a way that logical equivalence of processes coincide with bio-similarity;
+encoding, in such a way that logical equivalence of processes coincide with bio-similarity.
+\end{itemize}
+
+Moreover, we sketch how the encoding can be used to enhance the expressivity of RSs along different dimensions:
+\begin{itemize}
 \item we sketch how one can express the behaviour of entity mutation,
@@
 %way that a mutation $s'$ can be expressed; 
-in such a way that {\color{red} the mutated version of the entity
-$s$ can}  take part to only a subset of rules requiring entity $s$;
-\item we show that with a little coding effort, our 
+in such a way that the mutated version of the entity
+$s$ can  take part to only a subset of rules requiring entity $s$;
+\item we sketch how with a little coding effort, our 
 %encoding 
-{\color{red} embedding}
+embedding
 allows two RSs to
@@
 
-The main drawback of our proposal, is that the {\color{red} resulting c\CNA \ 
-code}
+The main drawback of our proposal, is that the resulting c\CNA \ 
+code
 %translation 
@@
 %translation 
-{\color{red} embedding}
+encoding
 can be automatised by means of a proper front-end in
@@
 %translation,
-{\color{red} embedding}
- Reaction Systems 
-get the ability to interact between them in a synchronized manner. 
+encoding
+the entities and the reactions
+get the ability to interact in a synchronous manner. 
 This interaction is not foreseen in
-the basic RS framework, as it can only happen with the context.
-{\color{red}The proposed network of reaction systems~\cite{BLR20} allows
-any reaction system in the network to receive entities 
-produced from its neighbors (that will represent its context), with the hypothesis that  
-reaction systems without neighbors will receive no entities 
+the basic RS framework, as it can only happen once the context is provided.
+The proposed network of RSs~\cite{BLR20} allows
+any RS in the network to receive entities 
+produced from its neighbours (that will represent its context), with the hypothesis that  
+RSs without neighbours will receive no entities 
 from the context. 
-In our case, a RS can receive some entities from the context and some others from a second RS, in any case we always assume contextes having a non-determinist behaviour.
-}
+In our case, a RS can receive some entities from the context and some others from a second RS; moreover we can represent contexts having a recursive, non-determinist behaviour.
+%
 By exploiting recursion, the kind of interactions which can be 
@@
 %SOS 
-structural operational semantics for the RS; it seems not the case
+structural operational semantics for RS; it seems not the case
 that a deterministic behaviour, as the one of the RS, once the initial state is 
@@
 
-A preliminary version of this paper appeared in~\cite{BBF19}.
-There are several major differences w.r.t. the paper~\cite{BBF19}, which is here extended as follows:
+As already mentioned, a preliminary version of this paper appeared in~\cite{BBF19}.
+There are several major differences w.r.t. the conference version~\cite{BBF19}, which is here extended as follows:
 \begin{itemize}
 \item we present several new examples to illustrate our 
-{\color{red} embedding},
+ encoding,
 %framework,
@@
 \item we show that our assertion language allows to specify properties extending to our
-{\color{red} embedding}
+encoding
 %framework 

### 2020-06-29 TCS_version/02-ReactionSystem.tex
altri piccoli cambiamenti
minor updates

--- 2020-06-18 TCS_version/02-ReactionSystem.tex
@@
-% !TEX root = ./main_RS2L_LNCS.tex
+% !TEX root = ./main_TCS.tex
 
-\section{Reaction Systems}
+\section{An Overview of Reaction Systems}
 \label{sec:rs}
@@
 {\em facilitation} and {\em inhibition}. These mechanisms are 
-reflected in the basic definitions of Reaction Systems.
+reflected in the basic definitions of RSs.
 
 \begin{definition}[Reaction]
-A reaction is a triplet $a = (R,I,P)$, where $R, I, P$ are finite,  
-non empty sets  and
-$R \cap I = \emptyset$. If $S$ is a set such that  $R, I, P \subseteq S$, 
-then $a$ is a reaction in $S$.
+Let $S$ be a set of entities.
+A reaction over $S$ is a triplet $a = (R,I,P)$, where $R, I, P$ are finite,  
+non empty subsets of $S$ and
+$R \cap I = \emptyset$. 
+%If $S$ is a set such that  $R, I, P \subseteq S$, 
+%then $a$ is a reaction in $S$.
 \end{definition}
 
-The sets $R, I, P$ are also written $R_a, I_a, P_a$ and called the 
+The sets 
+$R, I, P$ are also written $R_a, I_a, P_a$ and called the 
 \emph{reactant set} of $a$, the  
@@
 A Reaction System (RS) is an ordered pair ${\cal A} = (S, A)$ 
-such that $S$ is a finite set, and $A \subseteq rac(S)$.
+such that $S$ is a finite set of  entities, and $A \subseteq rac(S)$ is a set of reactions over $S$.
 %The set $S$ is called the background (set) of $A$.\\
@@
 
-The set $S$ is called the \emph{background set} of ${\cal A}$, its elements are called \emph{entities},
-they represent molecular substances (e.g., atoms, ions, molecules) that may be present
+The set $S$ is called the \emph{background set} of ${\cal A}$; its elements represent molecular substances (e.g., atoms, ions, molecules) that may be present
 in the states of a biochemical system.
@@
 \begin{definition}[Reaction Result]
-Let $W$ be a finite subset of $S$.
+Given a set of entities $S$, let $W\subseteq S$ be a finite subset of  entities.
 \begin{enumerate}
-\item Let $a$ be a reaction in $S$. Then $a$ is enabled by $W$, denoted by $\mathit{en}_a(W)$, 
-if $R_a \subseteq W$
-and $I_a \cap W = \emptyset$. The result of $a$ on $W$, denoted by $\mathit{res}_a(W)$, 
+\item Let $a$ be a reaction over $S$. Then $a$ is enabled by the entities in the set $W$, denoted by $\mathit{en}_a(W)$, 
+if $R_a \subseteq W$ and $I_a \cap W = \emptyset$, i.e. all the reactants of $a$ are in $W$,
+ while none of the inhibitors of $a$ are in $W$. The result of $a$ on $W$, denoted by $\mathit{res}_a(W)$, 
 is defined by:
-$\mathit{res}_a(W) =P_a$, if $\mathit{en}_a(W)$, and $\mathit{res}_a(W) = \emptyset$ otherwise.
-\item Let $A$ be a finite set of reactions. The result of $A$ on $W$, denoted 
+$\mathit{res}_a(W) \defeq P_a$, if $\mathit{en}_a(W)$, and $\mathit{res}_a(W) \defeq \emptyset$ otherwise.
+\item Let $A$ be a finite set of reactions over $S$. The result of $A$ on $W$, denoted 
 by $\mathit{res}_A(W)$,  is defined
-by: $\mathit{res}_A(W) = \bigcup_{a \in A} \mathit{res}_a(W)$.
+by: $\mathit{res}_A(W) \defeq \bigcup_{a \in A} \mathit{res}_a(W)$.
 \end{enumerate}
@@
 where 
-$ C_{i}, D_{i}  \subseteq S$ for any $i\in[0,n]$, $D_0 = \emptyset$, and 
+$ C_{i}, D_{i}  \subseteq S$ are set of entities for any $i\in[0,n]$, $D_0 = \emptyset$, and 
 %$C_0,\dots, C_n, D_0,\dots, D_n  \subseteq S$, $D_0 = \emptyset$, and 
@@
 the external environment, in discrete steps. 
-The sequence $\gamma$ is the {\em context sequence} of $\pi$ and 
-represents the influence of the environment on the Reaction System.
+The sequence $\gamma$ is the {\em context sequence} of $\pi$;
+it can be arbitrarily defined and 
+represents the influence of the environment on the 
+RS.
+%Reaction System.
 The sequence $\delta$ is
@@
 
-For technical reasons, we extend the notion of an interactive process to deal with 
+Since we will be able to deal with recursively contexts, we extend the notion of an interactive process to deal with 
 infinite sequences.

### 2020-06-29 TCS_version/04-CCNA.tex
minor updates

@@
 In this section we introduce  the syntax and operational semantics
-of a variant of the {\tt link}-calculus~\cite{BodeiBB12}, the c\CNA\  (chained \CNA)
-where the prefixes are link chains and not just links.
+of the process algebra c\CNA\  (chained \CNA) to be used for encoding RSs.
+As already explained in the Introduction, c\CNA\ is a variant of \CNA~\cite{BBB17}, the non-mobile
+fragment of {\tt link}-calculus~\cite{BodeiBB12,BODEI2020104587}.
+In c\CNA\  the action prefixes are link chains and not just links.
 %In this section we just focus on the basic ideas of link

### 2020-06-29 TCS_version/01-introduction.tex
qualche progresso

@@
 Mobile Ambients~\cite{CardelliG00}, getting a much stronger operational
-correspondence than any available in the literature, such as the one in~\cite{B16}.
+correspondence than any available in the literature, such as the ones in~\cite{CZ07,B16}.
 Later it was shown that the {\tt link}-calculus allowed one to easily encode calculi for biology equipped with membranes, as in~\cite{BodeiBBC14}.

### 2020-06-29 TCS_version/02-ReactionSystem.tex
qualche progresso

@@
 Any inhibitor blocks the reaction if it is present. Products are the outcome of the reaction.
-Also,  $R_a \cup I_a$ is the set of the resources of $a$ and $ rac(S)$
+Also,  $R_a \cup I_a$ is the set of the resources of $a$ and $\mathit{rac}(S)$
 denotes the set of all reactions in $S$.
@@
 A Reaction System (RS) is an ordered pair ${\cal A} = (S, A)$ 
-such that $S$ is a finite set of  entities, and $A \subseteq rac(S)$ is a set of reactions over $S$.
+such that $S$ is a finite set of  entities, and $A \subseteq \mathit{rac}(S)$ is a set of reactions over $S$.
 %The set $S$ is called the background (set) of $A$.\\

### 2020-06-29 TCS_version/04-CCNA.tex
qualche progresso

@@
 In this section we introduce  the syntax and operational semantics
-of the process algebra c\CNA\  (chained \CNA) to be used for encoding RSs.
+of the process algebra c\CNA\  (chained \CNA)~\cite{BBF19} to be used for encoding RSs.
 As already explained in the Introduction, c\CNA\ is a variant of \CNA~\cite{BBB17}, the non-mobile
@@
 \paragraph{Link Chains}
-Let $\mathcal{C}$ be the set of channels, ranged over by $a,b,...$, and 
-let $\mathcal{A} = \mathcal{C} \cup \setof{\silent} \cup \setof{{\noacts}}$ be the set of actions, ranged over by $\alpha,\beta,...$,
+Let $\mathcal{C}$ be the set of channels, ranged over by $\mathsf{a},\mathsf{b},...$, and 
+let $\mathit{Act} \defeq \mathcal{C} \cup \setof{\silent} \cup \setof{{\noacts}}$ be the set of actions, ranged over by $\alpha,\beta,...$,
 where the symbol $\silent$ denotes a \emph{silent} action, while the symbol $\noact$ denotes a \emph{virtual} (non-specified) action.
@@
 it is \emph{solid} if $\alpha,\beta \neq \ \noacts$; 
-{\color{red} $\alpha$ and $\beta$ are two interaction points, to fix the idea $\alpha$ is the action, and $\beta$ is the coaction.}
+intuitively, $\alpha$ and $\beta$ are two interaction points, one for incoming requests and the other for outgoing requests.
 The link $\link{\noact}{\noact}$ is called \emph{virtual}.
@@
 A chain is called \emph{solid} if it does not contain any virtual link.
+The empty chain is denoted by $\epsilon$.
 The equivalence $\blackstretcheq$ models expansion/contraction of virtual links to adjust the length of a link chain.
@@
 v\link{\noact}{\noact} & \blackstretcheq &  v & \qquad &
-v_1 \startchain{\noact}\chainedlink{\noact}{\noact}\chainend{\noact}v_2 & \blackstretcheq & v_1 \link{\noact}{\noact} v_2\\
+v_1 \startchain{\noact}\chainedlink{\noact}{\noact}\chainend{\noact}v_2 & \blackstretcheq & v_1 \link{\noact}{\noact} v_2\\[5pt]
 \link{\noact}{\noact}v & \blackstretcheq & v & &
-v_1 \startchain{\alpha}\chainedlink{a}{a}\chainend{\beta} v_2
+v_1 \startchain{\alpha}\chainedlink{\mathsf{a}}{\mathsf{a}}\chainend{\beta} v_2
 & \blackstretcheq & 
-v_1 \startchain{\alpha}\chainedlink{a}{\noact}\chainedlink{\noact}{a}\chainend{\beta}v_2 
+v_1 \startchain{\alpha}\chainedlink{\mathsf{a}}{\noact}\chainedlink{\noact}{\mathsf{a}}\chainend{\beta}v_2 
 \end{array}
@@
 %
-For example, given $v_1 = \startchain{\silent}\chainedlink{a}{\noact}\chainedlink{\noact}{\noact}\chainend{\noact}$, $v_2 = \startchain{\noact}\chainedlink{\noact}{a}\chainedlink{b}{\noact}\chainend{\noact}$ and $v = \startchain{\silent}\chainedlink{a}{a}\chainedlink{b}{\noact}\chainend{\noact}$
+For example, given $v_1 = \startchain{\silent}\chainedlink{\mathsf{a}}{\noact}\chainedlink{\noact}{\noact}\chainend{\noact}$, $v_2 = \startchain{\noact}\chainedlink{\noact}{\mathsf{a}}\chainedlink{\mathsf{b}}{\noact}\chainend{\noact}$ and $v = \startchain{\silent}\chainedlink{\mathsf{a}}{\mathsf{a}}\chainedlink{\mathsf{b}}{\noact}\chainend{\noact}$
 we have $\merges{v_1}{v_2} = v$, whereas $\merges{v_1}{v}$ is not defined. Notably the merge operation is commutative and associative.
@@
 This is possible only if they are matched by some adjacent link.
-Restriction is denoted by $\restrict{a}v$.
+Restriction is denoted by $\restrict{\mathsf{a}}v$.
 %
-For example, given $v = \startchain{\silent}\chainedlink{a}{a}\chainedlink{b}{\noact}\chainend{\noact}$ as above, we have 
-$\restrict{a}v = \startchain{\silent}\chainedlink{\silent}{\silent}\chainedlink{b}{\noact}\chainend{\noact}$,
-whereas $\restrict{b}v$ is not defined.
+For example, given $v = \startchain{\silent}\chainedlink{\mathsf{a}}{\mathsf{a}}\chainedlink{\mathsf{b}}{\noact}\chainend{\noact}$ as above, we have 
+$\restrict{\mathsf{a}}v = \startchain{\silent}\chainedlink{\silent}{\silent}\chainedlink{\mathsf{b}}{\noact}\chainend{\noact}$,
+whereas $\restrict{\mathsf{b}}v$ is not defined.
 
@@
 P,Q & ::= &
-%\nil \prodsep
+\nil \prodsep
 %X \prodsep
-%s.P  \prodsep
-\sum_{i\in I} \upsilon_i.P_i \prodsep
+\upsilon.P  \prodsep
+%\sum_{i\in I} \upsilon_i.P_i \prodsep
+P+Q \prodsep
 P|Q \prodsep
-\restrict{a} P \prodsep
-P[\phi] \prodsep
+\restrict{\mathsf{a}} P \prodsep
+%P[\phi] \prodsep
 A
@@
 %with $\alpha,\alpha',\beta,\beta'\neq \ \noacts$),
-$\phi$ is a channel renaming function,
-and $A$ is a process identifier for which we assume a definition $A
-\defeq P$ is available in a given set $\Delta$ of (possibly
-recursive) process definitions. We let $\nil$, the inactive process,
-denote the empty summation.
+%$\phi$ is a channel renaming function,
+and $A$ is a process identifier.
+The syntax of c\CNA\  extends that of \CNA~\cite{BBB17} by allowing to use link chains as prefixes instead of links, i.e. we allow to write $\upsilon.P$ instead of $\ell.P$.
+For the rest it features nondeterministic choice $P+Q$ (also called sum), parallel composition $P|Q$, restriction $\restrict{\mathsf{a}} P$, and possibly recursively defined process identifiers $A$. 
+Here we do not consider name mobility, which is present instead in the {\tt link}-calculus and we omit the relabelling operator of \CNA\ (not needed in the encoding).
+
+As common in process algebras we restrict to consider prefix-guarded sums $\upsilon_1.P_1 + \upsilon_2.P_2$ and, exploiting associativity, we use the shorthand $\sum_{i\in I} \upsilon_i.P_i$ for a finite set of indexes $I=\{i_1,...,i_k\}$ instead of $\upsilon_{i_1}.P_{i_1} + \cdots + \upsilon_{i_k}.P_{i_k}$.
+The inactive process $\nil$ is thus just the empty summation.
+
+Regarding process constants,  we rely on a given set $\Delta=\{A_i\defeq P_i\}_{i\in I}$ of (possibly
+recursive) process definitions.
 %\end{definition}
-
-The syntax of c\CNA\  extends that of \CNA~\cite{BBB17} by allowing to use link chains as prefixes instead of links.
-For the rest it features nondeterministic (guarded) choice, parallel composition, restriction, relabelling and possibly recursive definitions.
-Here we do not consider name mobility, which is present instead in the {\tt link}-calculus.
 
@@
 The only difference is due to the presence of prefixes that are link chains.
-Briefly: rule (\textit{Sum}) selects one alternative and puts as label a possible contraction/expansion of the link chain in the selected prefix; rule (\textit{Ide}) selects one transition of the defining process for a constant; rule (\textit{Res}) restricts some names in the label (it cannot be applied when $\restrict{a}v$ is not defined); rules (\textit{Lpar}) and (\textit{Rpar}) account for interleaving in parallel composition; rule (\textit{Com})
+Briefly: rule (\textit{Sum}) selects one alternative and puts as label a possible contraction/expansion of the link chain in the selected prefix; rule (\textit{Ide}) selects one transition of the defining process for a constant; rule (\textit{Res}) restricts some names in the label (it cannot be applied when $\restrict{\mathsf{a}}v$ is not defined); rules (\textit{Lpar}) and (\textit{Rpar}) account for interleaving in parallel composition; rule (\textit{Com})
 synchronises interactions (it cannot be applied when $\merges{v}{v'}$ is not defined).
@@
 \begin{example}
-As some simple examples, consider the recursive process definitions $H \defeq \link{a}{b}.\link{a}{c}.H$ and $K \defeq \link{a}{b}.K + \link{a}{c}.K$: the former is some sort of forwarder that alternates between providing a link from $a$ to $b$ and to $c$; the latter is a nondeterministic forwarder that at each step can link $a$ to $b$ or to $c$.
+As some simple examples, consider the recursive process definitions $H \defeq \link{\mathsf{a}}{\mathsf{b}}.\link{\mathsf{a}}{\mathsf{c}}.H$ and $K \defeq \link{\mathsf{a}}{\mathsf{b}}.K + \link{\mathsf{a}}{\mathsf{c}}.K$: the former recursively provides a link from $\mathsf{a}$ to $\mathsf{b}$ and then, at the next step, from $\mathsf{a}$ to $\mathsf{c}$; the latter provides at each step a link from $\mathsf{a}$ to $\mathsf{b}$ or from $\mathsf{a}$ to $\mathsf{c}$, nondeterministically.
 \end{example}
@@
 \DisplayProof
-\
+\qquad
 \AxiomC{$P \xrightarrow{\upsilon} P'$}
@@
 \RightLabel{\scriptsize(\textit{Res})}
-\UnaryInfC{$\restrict{a}P \xrightarrow{\restrict{a}\upsilon} \restrict{a}P'$} 
+\UnaryInfC{$\restrict{\mathsf{a}}P \xrightarrow{\restrict{\mathsf{a}}\upsilon} \restrict{a}P'$} 
 \DisplayProof
+\quad
 %\,
@@
 \DisplayProof
+\quad
+\AxiomC{$Q \xrightarrow{v} Q'$} 
+\RightLabel{\scriptsize(\textit{Rpar})}
+\UnaryInfC{$P|Q \xrightarrow{\upsilon} P|Q'$} 
+%\DisplayProof
+\end{prooftree} 
+\end{center}
+
+\begin{center} 
+\begin{prooftree} 
 %\,
@@
 \end{center}
-\caption{SOS semantics of the c\CNA\  (rules (\textit{Rel}) and (\textit{Rpar}) omitted).}
+\caption{SOS semantics of c\CNA\  processes.}
 \label{fig:cnasos}
@@
 %translation.
-{\color{red} embedding.}
+encoding.
 
 \begin{definition}[Replication]
-Let $\upsilon$ be a link chain such that $\upsilon\upsilon$ is also a valid link chain. Its $n$ times replication $\upsilon^n$ is defined recursively by letting
+Let $\upsilon$ be a valid link chain such that $\upsilon\upsilon$ is also a valid link chain. The $n$ times replication of $\upsilon$, written $\upsilon^n$, is defined recursively by letting
 $\upsilon^0 = \epsilon$ (i.e. the empty chain) and $\upsilon^n = \upsilon\upsilon^{n-1}$.
@@
 
-For example,  the expression $(\startchain{a}\chainedlink{b}{\noact}\chainend{\noact})^3$  denotes the chain $\startchain{a}\chainedlink{b}{\noact}\chainedlink{\noact}{a}\chainedlink{b}{\noact}\chainedlink{\noact}{a}\chainedlink{b}{\noact}\chainend{\noact}$.
-Instead the expression $(\link{a}{b})^2$ is ill-defined because $a$ does not match with $b$.
+For example,  the expression $(\startchain{\mathsf{a}}\chainedlink{\mathsf{b}}{\noact}\chainend{\noact})^3$  denotes the chain $\startchain{\mathsf{a}}\chainedlink{\mathsf{b}}{\noact}\chainedlink{\noact}{\mathsf{a}}\chainedlink{\mathsf{b}}{\noact}\chainedlink{\noact}{\mathsf{a}}\chainedlink{\mathsf{b}}{\noact}\chainend{\noact}$.
+Instead the expression $(\link{\mathsf{a}}{\mathsf{b}})^2$ is ill-defined because $\mathsf{a}$ does not match with $\mathsf{b}$.
 
@@
 \begin{definition}[Half links]
-Let $a$ be a channel name, we define the \emph{half left link}  $\link{a}{}$,   and the  \emph{half right link} $\link{}{a}$.
+Let $\mathsf{a}$ be a channel name, we denote by $\link{\mathsf{a}}{}$ the \emph{half left link},  and by $\link{}{\mathsf{a}}$ the  \emph{half right link}.
 \end{definition}
 
+In the encoding of RS we will make extensive use of subscripted names:  each name $\mathsf{a}$ will come in two variants $\mathsf{a}_i$ and $\mathsf{a}_o$.
+This is just a technical issue to prevent accidental matching between links.
+To see why this is important, compare the chain
+$\startchain{\silent}\chainedlink{\mathsf{a}}{\noact}\chainedlink{\noact}{\mathsf{a}}\chainend{\silent}$ with 
+$\startchain{\silent}\chainedlink{\mathsf{a}_i}{\noact}\chainedlink{\noact}{\mathsf{a}_o}\chainend{\silent}$: the former is $\blackstretcheq$ equivalent to the solid chain $\startchain{\silent}\chainedlink{\mathsf{a}}{\mathsf{a}}\chainend{\silent}$; the latter cannot become solid unless merged with a chain that links $\mathsf{a}_i$ to $\mathsf{a}_o$, like
+$\startchain{\noact}\chainedlink{\noact}{\mathsf{a}_i}\chainedlink{\mathsf{a}_o}{\noact}\chainend{\noact}$.
+This form of subscripting is exploited in the definition of open blocks.
+
 \begin{definition}[Open block]
-Let $R$ be a set of names. We 
-define an \emph{open block} as $\blockchain{a_i}{a_o}{a \in R}$, where $a_i$ and $a_o$ are annotated version of the name $a$, as
+Let $\sigma$ be a finite sequence of names. We 
+define an \emph{open block} as $\blockchain{\mathsf{a}_i}{\mathsf{a}_o}{\mathsf{a} \in \sigma}$, where $\mathsf{a}_i$ and $\mathsf{a}_o$ are annotated version of the name $\mathsf{a}$ (as explained above), by letting
 \[
-\begin{array}{l@{\quad}l@{\quad}l}
-\mbox{set}& \mbox{block of chain}& \mbox{result}\\
-R= \emptyset & \blockchain{a_i}{a_o}{a \in R}   &\epsilon\\
-R = \{ b\}    &\blockchain{a_i}{a_o}{a \in R}  & {}_{b_i}^{\noact}\chainedlink{\noact}{b_o}\\
-R = \{b\} \cup R' & \blockchain{a_i}{a_o}{a \in R} & {}_{b_i}^{\noact}\chainedlink{\noact}{b_o}\link{}{}\blockchain{a_i}{a_o}{a \in R'}
+\blockchain{\mathsf{a}_i}{\mathsf{a}_o}{\mathsf{a} \in \sigma} \defeq
+\left\{
+\begin{array}{ll}
+\epsilon 
+& \mbox{if $\sigma=\epsilon$ is the empty sequence}\\[5pt]
+{}_{\mathsf{b}_i}^{\noact}\chainedlink{\noact}{\mathsf{b}_o}\blockchain{\mathsf{a}_i}{\mathsf{a}_o}{\mathsf{a} \in \rho}
+& \mbox{if $\sigma=\mathsf{b}\rho$}
 \end{array}
+\right.
 \]
+%\[
+%\begin{array}{l@{\quad}l@{\quad}l}
+%\mbox{set}& \mbox{block of chain}& \mbox{result}\\
+%R= \emptyset & \blockchain{a_i}{a_o}{a \in R}   &\epsilon\\
+%R = \{ b\}    &\blockchain{a_i}{a_o}{a \in R}  & {}_{b_i}^{\noact}\chainedlink{\noact}{b_o}\\
+%R = \{b\} \cup R' & \blockchain{a_i}{a_o}{a \in R} & {}_{b_i}^{\noact}\chainedlink{\noact}{b_o}\link{}{}\blockchain{a_i}{a_o}{a \in R'}
+%\end{array}
+%\]
 \end{definition}
@@
 
+Abusing the notation, we will use the open block notation for sets of names rather than sequences, assuming the names in the set are taken  according to some default order (e.g.
+the lexicographic one).
+
 %\noindent
-We then combine half links and  open blocks to form regular link chains.
+We then combine half links and  open blocks to form valid link chains.
 %
@@
 % Again,  all the channel names appearing in the half links or in the open blocks of chain are not required to match.
- For example, for $R = \{a,b\}$ the expression $ \blockchain{c_i}{c_o}{c \in R} $
-denotes the block ${}_{a_i}^{\noact}\chainedlink{\noact}{a_o} \backslash{}_{b_i}^{\noact}\chainedlink{\noact}{b_o}$; and the expression $\link{r_1}{} \blockchain{c_i}{c_o}{c \in R}\link{}{r_2} $ denotes the chain
-$ \startchain{r_1}\chainedlink{a_i}{\noact}\chainedlink{\noact}{a_o}\chainedlink{b_i}{\noact}\chainedlink{\noact}{b_o}\chainend{r_2}$.
+
+ For example, for $X = \{\mathsf{a},\mathsf{b}\}$ the expression $ \blockchain{\mathsf{c}_i}{\mathsf{c}_o}{\mathsf{c} \in X} $
+denotes the block ${}_{\mathsf{a}_i}^{\noact}\chainedlink{\noact}{\mathsf{a}_o} \backslash{}_{\mathsf{b}_i}^{\noact}\chainedlink{\noact}{\mathsf{b}_o}$; and the expression $\link{\mathsf{r}_1}{} \blockchain{\mathsf{c}_i}{\mathsf{c}_o}{\mathsf{c} \in X}\link{}{\mathsf{r}_2} $ denotes the chain
+$ \startchain{\mathsf{r}_1}\chainedlink{\mathsf{a}_i}{\noact}\chainedlink{\noact}{\mathsf{a}_o}\chainedlink{\mathsf{b}_i}{\noact}\chainedlink{\noact}{\mathsf{b}_o}\chainend{\mathsf{r}_2}$.

### 2020-06-29 TCS_version/05-translation.tex
qualche progresso

@@
 \label{sec:trans}
-Here we present {\color{red} an embedding of}
+Here we define an encoding of
 %a translation from 
-Reaction Systems {\color{red} into} c\CNA.
+Reaction Systems into c\CNA.
 The idea is to define separated processes for representing the behaviour of each entity, each reaction, 
@@
 
+In the following we refer to a given set of entities $S$ and a set of reactions $A \subseteq \mathit{rac}(S)$, i.e. that the Reaction System ${\cal A} = (S, A)$ is known.
+
 \paragraph{Processes for entities}
-Given an entity $s \in S$, we exploit five different pairs of names for the interactions over $s$:
+Given an entity $s \in S$, we exploit five different pairs of channel names for the interactions over $s$:
 \begin{itemize}
@@
 \paragraph{Processes for reactions}
-We assume that {color{red} all the reactions $a$ are ordered by assign to them a progressive  number $j$
-from $1$ to $n$.}
+We assume that all the reactions $a$ are numbered and use $j$ as an index for reactions.
+We introduce two channel names for each reaction $aj$:
+\begin{itemize}
+\item $r_j$ to mark the occurrence of the reaction;
+\item $p_j$ to mark the product set of the reaction.
+\end{itemize}
+We shall exploit names $r_j,p_j$ to join the chains provided by the application of all the reactions. 
 %each reaction $a$ is 
 %assigned a progressive number $j$.
-The process for reaction $aj=(R_j,I_j,P_j)$ must assert either the possibility to apply the reaction or its impossibility.
+The process for the $j$th reaction $aj=(R_j,I_j,P_j)$ must assert either the possibility to apply the reaction or its impossibility.
 The first case happens when all its reactants are present (the link $\link{s_i}{s_o}$ is requested for any $s\in R_j$) and all its inhibitors are absent (the link $\link{\overline{e}_i}{\overline{e}_o}$ is requested for any $e\in I_j$), then the product set is released (the link $\link{\widetilde{c}_i}{\widetilde{c}_o}$ is requested for any $c\in P_j$).
-The next case can happen for two reasons: one of the reactants is absent (the link $\link{\overline{s}_i}{\overline{s}_o}$ is requested for some $s\in R_j$) or one of the inhibitors is present (the link $\link{e_i}{e_o}$ is requested for some $e\in I_j$).
+The second case can happen for two reasons: one of the reactants is absent (the link $\link{\overline{s}_i}{\overline{s}_o}$ is requested for some $s\in R_j$) or one of the inhibitors is present (the link $\link{e_i}{e_o}$ is requested for some $e\in I_j$).
 The process is recursive so that reactions can be applied at any step.
@@
 
-We exploit names $r_j,p_j$ to join the chains provided by the application of all the reactions. 
 Channels $r_j$ and $r_{j+1}$ enclose the enabling/disabling condition of reaction $aj$.
@@
 We will see that all the link chain labels of transitions follow the same schema: first we find all the reactions limited to the reactants and inhibitors (chained using $r_j$ channels), then all the supplies by the contexts (chained using the channel $\mathit{cxt}$, to be introduced next), and finally the products for all the reactions (chained using $p_j$ channels).
+For notational convenience, we fix that 
+%$r_1 = \silent$,
+$r_{|A|+1} = \mathit{cxt}$ and $p_{|A|+1}=\silent$.
 In the following there is an example explaining this schema.
@@
 \paragraph{Processes for contexts}
-The context can vary at each step $n$ and for each entity $s \in S$, the context must say if the entity is provided or not.
-Let us denote by $C_n$ the context at step $n$, i.e.
-the set of entities that are made available.
-Correspondingly, we introduce another process $\mathit{Cxt}^n$ defined as follows:
+For marking the part of the chain provided by the context, we exploit the name $\mathit{cxt}$.
+In RSs, the context sequence $\gamma$ provides a set of entities $C_n$ at each instant of time $n$: for each entity $s \in S$, the context must say if the entity is provided or not.
+Correspondingly, we introduce another process $\mathit{Cxt}_n$ defined as follows:
 \[
 \begin{array}{lcl}
-\mathit{Cxt}^n &\defeq &
+\mathit{Cxt}_n &\defeq &
 %\left \{
-^{cxt} \backslash  \blockchain{\hat{s}_i}{\hat{s}_o}{s\!\in \!C_n}  \link{}{} \blockchain{\underline{e}_i}{\underline{e}_o}{e \!\not\in \! C_n}   \backslash_{p_1}. \mathit{Cxt}^{n+1}
+^{cxt} \backslash  \blockchain{\hat{s}_i}{\hat{s}_o}{s\!\in \!C_n}  \link{}{} \blockchain{\underline{e}_i}{\underline{e}_o}{e \!\not\in \! C_n}   \backslash_{p_1}. \mathit{Cxt}_{n+1}
 %+ \\
@@
 
-We only consider $\mathit{Cxt}^n$ with $n >0$, as the entities  that are present at step zero  are considered to be present in the initial system (if $s\in C_0$ the process $P_s$ will be present initially, otherwise $\overline{P_s}$ will be present).
+We only consider $\mathit{Cxt}_n$ with $n >0$, as the entities  that are present at step zero  are considered to be present in the initial system (if $s\in C_0$ the process $P_s$ will be present initially, otherwise $\overline{P_s}$ will be present).
 
+\paragraph{Encoding}
 In the following we use the following conventions for denoting different categories of names:
@@
 \item
-$\mathit{decs} \defeq \{ s, \overline{s}, \widetilde{s},\widehat{s},\underline{s}~|~ s\in S\}$ is the set of names for decorated entities (without subscripts $_i$ and $_o$);
+$\mathit{decs} \defeq \{ s, \overline{s}, \widetilde{s},\widehat{s},\underline{s}~|~ s\in S\}$ is the set of channel names for decorated entities (without subscripts $_i$ and $_o$);
 \item
-$\mathit{ents} \defeq \{ d_i, d_o~|~ d\in \mathit{decs}\}$ is the set names for entities;
+$\mathit{ents} \defeq \{ d_i, d_o~|~ d\in \mathit{decs}\}$ is the set of channel names for entities;
 \item
-$\mathit{reacts}\defeq \{r_1,...,r_{|A|}\}$ is the set of reaction names $r_j$ associated with each reaction $aj$;
+$\mathit{reacts}\defeq \{r_1,...,r_{|A|+1}\}$ is the set of channel names $r_j$ associated with each reaction $aj$ (we remind that $r_{|A|+1} = \mathit{cxt}$);
 \item 
-$\mathit{prods}\defeq \{p_1,...,p_{|A|}\}$ is the set of names for production $p_j$, associated with  reaction $aj$.
+$\mathit{prods}\defeq \{p_1,...,p_{|A|}\}$ is the set of channel names $p_j$ for product sets associated with each reaction $aj$ (we remind that $p_{|A|+1} = \silent$).
 \end{itemize}
 
-\begin{definition}[{\color{red} Embedding}]
+\begin{definition}[Encoding]
 %Translation]
@@
 Let ${\cal A} = (S, A)$ be a RS, and let $\pi=(\gamma,\delta)$ be an extended interactive process in ${\cal A}$, with  $\gamma=\{C_i\}_{i\in\mathbb{N}}$. 
-We define its c\CNA~{\color{red} embedding}
+We define its c\CNA\ encoding
 %translation 
@@
 to allow the name $r_1$ to be matched at the start of any chain; also, 
-for notational convenience, we fix that 
-%$r_1 = \silent$,
-$r_{u+1} = \mathit{cxt}$ and $p_{u+1}=\silent$,  for $u=|A|$.
 \end{definition}

### 2020-06-29 TCS_version/09-appendix.tex
qualche progresso

@@
 Then, all the processes $I$, $P_a$, $P_s$ (or $\overline{P_s}$), and $\mathit{Cxt}$  must participate to each transition.
- \item for each reaction $aj \in A$, the c\CNA \ code of  $P_{aj}$ leaves one virtual link between two solid links of the types $\link{r_j}{\dots} \chainedlink{s_{i}}{\noact}\chainedlink{\noact}{s_{o}}\link{}{\dots}\link{}{r_{j+1}} ^{\dots}\link{p_j}{\dots}\link{}{p_{j+1}}$.
+ \item for each reaction $aj \in A$, the c\CNA \ code of  $P_{aj}$ leaves one virtual link between two solid links of the types $\link{r_j}{\dots} \chainedlink{s_{i}}{\noact}\chainedlink{\noact}{s_{o}}\link{}{\dots}\link{}{r_{j+1}} {\dots}\link{p_j}{\dots}\link{}{p_{j+1}}$.
 Then, it derives that the process $P_{s}$, encoding the behaviour of entity $s$, can participate by filling the virtual link in the above transition by only offering one solid link of the form $\link{s_{i}}{s_{o}}$.

### 2020-06-29 TCS_version/05-translation.tex
poco

@@
 $r_{|A|+1} = \mathit{cxt}$ and $p_{|A|+1}=\silent$.
+{\color{red} Please, recall that $A$ is a subset of reactions in $S$.}
 In the following there is an example explaining this schema.
@@
 It is important to observe that,   for each transition, our c\CNA \ 
-%encoding 
-{\color{red} embedding}
+encoding 
+%{\color{red} embedding}
 requires  all the processes 
@@
  %\label{lem:struct}
-  Let ${\cal A} = (S, A)$ be a RS and let $\pi=(\gamma,\delta)$ be an extended  interactive process in ${\cal A}$. Let $P=\llbracket{\cal A},\gamma\rrbracket$ its c\CNA~{\color{red} embedding.}
+  Let ${\cal A} = (S, A)$ be a RS and let $\pi=(\gamma,\delta)$ be an extended  interactive process in ${\cal A}$. Let $P=\llbracket{\cal A},\gamma\rrbracket$ its c\CNA~encoding.
+  %{\color{red} embedding.}
   %translation. 
@@
 \begin{example}\label{ex:backbone}
- Let ${\cal A}$ be a RS whose specification contains two entities, $s1$ and $s2$, and the reactions $r_1=(s1,\ , s2)$ and $r_2 = (s2, \ , s1)$ that produce $s2$ if $s1$ is present and $s1$ if $s2$ is present.
+ Let ${\cal A}$ be a RS whose specification contains two entities, $s1$ and $s2$, and the reactions $r_1=(s1,\emptyset , s2)$ and $r_2 = (s2, \emptyset , s1)$ that produce $s2$ if $s1$ is present and $s1$ if $s2$ is present.
 For simplicity we do not consider inhibitors, but the reader can assume a void inhibitor is present in both reactions.
@@
     solid links to be merged with the virtual links of the backbone (i.e. to be plugged in the backbone).
-{\color{red}  In Figure~\ref{fig:backbone} we added the underline brackets, where the notation $P_1(P_2,P_3)$ means that the segment of the link chain is delimited by the process $P_1$ and it leaves "holes" where processes between the brackets insert their links.}
+{\color{red}  In Figure~\ref{fig:backbone} we added the underline brackets, where the notation $P_1(P_2,P_3)$ means that the segment of the link chain is delimited by the process $P_1$ and it leaves "holes" where processes between the brackets, in this case $P_2$ and $P_3$, insert their links.}
  Formally, we have
@@
 \begin{array}{lcl}
-\mathit{flat}(\link{\alpha}{\beta}\upsilon)& \defeq & \mathit{flat}(\link{\alpha}{\beta}):: \mathit{flat}(\upsilon)\\
+\mathit{flat}(\link{\alpha}{\beta}\upsilon)& \defeq & \mathit{flat}(\link{\alpha}{\beta}) \mathit{flat}(\upsilon)\\
 \end{array}\\
 \]
-where  $::$ is the usual string concatenation operator.
+{\color{red} where the juxtaposition of two string results in the   
+%$::$ is the 
+usual string concatenation operator.}
 

### 2020-06-29 TCS_version/06-examples.tex
poco

@@
 \label{ex:examples}
-
-Hereafter, we shall often omit topmost restrictions but shall take into account only transitions whose labels are complete chains, i.e. they do not contain 
-virtual links and start/end with the $\silent$ action symbol.
+{\color{red} For readability,  we shall often omit topmost restrictions.
+Please remember that, semantically, restriction allows for only interactions with no virtual links, creating a private interaction among all participants where all the channel names in the transition labels are hidden.
+%Hereafter, we shall often omit topmost restrictions
+ Nevertheless, for simplicity,
+ % but 
+ we shall take into account only transitions whose labels are complete chains, i.e. they do not contain 
+virtual links and start/end with the $\silent$ action symbol.}
 
@@
 This example is inspired by the example in~\cite{BEMR11}, where
-a deterministic transition system is coded in the Reaction System
+a deterministic transition system is 
+%coded 
+{\color{red} encoded} in the Reaction System
 framework.

### 2020-06-29 TCS_version/04-CCNA.tex
revisione

@@
 The inactive process $\nil$ is thus just the empty summation.
+
+In a prefix $\upsilon.P$ we can always assume that the links at the extremities of $\upsilon$ are solid: if $\upsilon$ needs to be used in larger chains, the operational semantics will add as many virtual links as needed by exploiting the equivalence $\blackstretcheq$ (see rule $\mathit{Sum}$ in Fig.~\ref{fig:cnasos}).
+For example, the process $\link{\mathsf{a}}{\mathsf{b}}.P$ and $\startchain{\mathsf{a}}\chainedlink{\mathsf{b}}{\noact}\chainend{\noact}.P$
+ are completely equivalent.
 

### 2020-06-29 TCS_version/05-translation.tex
revisione

@@
 \paragraph{Processes for reactions}
+Here we focus on the encoding of the set of reactions $A \subseteq \mathit{rac}(S)$.
 We assume that all the reactions $a$ are numbered and use $j$ as an index for reactions.
@@
 Channels $p_j$ and $p_{j+1}$ enclose the links related to the entities produced by $aj$.
+Each reaction defines a pattern to be satisfied, i.e. each reaction inserts as many virtual links  as  the number of reactants, inhibitors, and products, as required by the corresponding reaction.
+
 We will see that all the link chain labels of transitions follow the same schema: first we find all the reactions limited to the reactants and inhibitors (chained using $r_j$ channels), then all the supplies by the contexts (chained using the channel $\mathit{cxt}$, to be introduced next), and finally the products for all the reactions (chained using $p_j$ channels).
@@
 $r_{|A|+1} = \mathit{cxt}$ and $p_{|A|+1}=\silent$.
-{\color{red} Please, recall that $A$ is a subset of reactions in $S$.}
-In the following there is an example explaining this schema.
+This schema will be later illustrated in detail in Example~\ref{ex:backbone}.
 %\[
@@
 %\left \{
-^{cxt} \backslash  \blockchain{\hat{s}_i}{\hat{s}_o}{s\!\in \!C_n}  \link{}{} \blockchain{\underline{e}_i}{\underline{e}_o}{e \!\not\in \! C_n}   \backslash_{p_1}. \mathit{Cxt}_{n+1}
+^{cxt} \backslash  \blockchain{\widehat{s}_i}{\widehat{s}_o}{s\!\in \!C_n}  \link{}{} \blockchain{\underline{e}_i}{\underline{e}_o}{e \!\not\in \! C_n}   \backslash_{p_1}. \mathit{Cxt}_{n+1}
 %+ \\
@@
 \triangleq
- \restrict{\mathit{names}}\left(I ~|~ \prod_{a\in A} P_a ~|~  \mathit{Cxt}^1 ~|~\prod_{s\in C_0}P_{s} ~|~ \prod_{s \notin C_0}\overline{P_s}\right)
+ \restrict{\mathit{names}}\left(I ~|~ \prod_{a\in A} P_a ~|~  \mathit{Cxt}_1 ~|~\prod_{s\in C_0}P_{s} ~|~ \prod_{s \notin C_0}\overline{P_s}\right)
 $$
@@
 %For notational convenience, we fix that $r_1 = \silent$,  $r_{u+1} = cxt_1$ for $u$ the number of reacts, and $cxt_{w+1} = \silent$ for $w$ the number of entities.
-For technical reasons, we introduce the {\color{red} recursive} init  process $I \defeq \link{\silent}{r_1}.I$
-to allow the name $r_1$ to be matched at the start of any chain; also, 
+For technical reasons, we introduce the (trivially recursive) init  process $I \defeq \link{\silent}{r_1}.I$: it is needed
+to allow the name $r_1$ to be matched at the start of any chain at any instant of time.
 \end{definition}
@@
 running in parallel to  interact in that transition.
-This is due to the fact that  all the channels $r_j$, $p_j$, $\mathit{cxt}$, including those for decorated names $s_{i},s_{o},\overline{s}_i,\overline{s}_o$ are  restricted.
-Each reaction defines a pattern to be satisfied, i.e. each reaction inserts as many virtual links  as  the number of reactants, inhibitors, and products, as required by the corresponding reaction.
+This is due to the fact that  all the channel names $r_j$, $p_j$, $\mathit{cxt}$, including those for decorated names $s_{i},s_{o},\overline{s}_i,\overline{s}_o,...$ are  restricted.
 
@@
  %involved in the transition $t$;
-  \item for each reaction $aj \in A$ and each entity $s \in S$, each virtual link offered by processes $P_a$ and $\mathit{Cxt}$ is  overlapped by exactly one solid link offered by  processes representing entities.
+  \item for each reaction $aj \in A$ and each entity $s \in S$, each virtual link offered by processes $P_a$ and $\mathit{Cxt}_1$ is  overlapped by exactly one solid link offered by  processes representing entities.
   %each process $P_s $ (or its alternative version $\overline{P}_s$), with $s \in S$, participates at most  with one solid link for each process $P_a$, and with exactly one solid link for each process $Cxt_s$.%, if $s$ is a \emph{required element} of $A_j$ in $t$.
@@
 We will then use $\llparenthesis{\cal A},\gamma\rrparenthesis$ to refer to the 
-%encoding 
-{\color{red} embedding}
+encoding 
 without topmost name restrictions, i.e.,
@@
 \llparenthesis{\cal A},\gamma\rrparenthesis \defeq 
-I~|~ \prod_{a\in A} P_a ~|~  \mathit{Cxt}^1~|~\prod_{s\in C_0}P_{s} ~|~ \prod_{s \notin C_0}\overline{P_s} .
+I~|~ \prod_{a\in A} P_a ~|~  \mathit{Cxt}_1~|~\prod_{s\in C_0}P_{s} ~|~ \prod_{s \notin C_0}\overline{P_s} .
 $$
@@
  Let ${\cal A}$ be a RS whose specification contains two entities, $s1$ and $s2$, and the reactions $r_1=(s1,\emptyset , s2)$ and $r_2 = (s2, \emptyset , s1)$ that produce $s2$ if $s1$ is present and $s1$ if $s2$ is present.
-For simplicity we do not consider inhibitors, but the reader can assume a void inhibitor is present in both reactions.
+For simplicity, we consider empty sets of inhibitors, which are not allowed by Definition~\ref{def:reaction}, but the reader can assume a void inhibitor is present in both reactions.
 Then, we assume an  extended interactive process $\pi=(\gamma,\delta)$ where the context $\gamma$ provides  $s1$ and $s2$ at every step, but we assume that only $s1$ is initially present.
+Since the context sequence is constant, we omit the subscript from $\mathit{Cxt}$.
 The corresponding c\CNA \ process is $\llbracket{\cal A},\gamma\rrbracket \defeq \restrict{\mathit{names}} \llparenthesis{\cal A},\gamma\rrparenthesis$, with
@@
     solid links to be merged with the virtual links of the backbone (i.e. to be plugged in the backbone).
-{\color{red}  In Figure~\ref{fig:backbone} we added the underline brackets, where the notation $P_1(P_2,P_3)$ means that the segment of the link chain is delimited by the process $P_1$ and it leaves "holes" where processes between the brackets, in this case $P_2$ and $P_3$, insert their links.}
+In Figure~\ref{fig:backbone}, at the bottom of the chain, we have underlined with brackets the origin of the solid links that appear in the chain: the notation $P_1(P_2,P_3)$ means that the segment of the link chain is delimited by the process $P_1$ and it leaves "holes" where processes between the brackets, in this case $P_2$ and $P_3$, insert their links.
  Formally, we have
@@
 \paragraph{The flat function}
-For technical reasons, our transition labels are quite verbose; then, to simplify their processing, we introduce a function that takes a solid link chain and returns a simple string by eliminating all the channel matching pairs leaving just one placeholder for them.
+Our transition labels are quite verbose; then, to simplify their processing, we introduce a function that takes a solid link chain and returns a simple string by eliminating all the channel matching pairs and leaving just one placeholder for them.
 This transformation is harmless, in the sense that it retains all the information in the chain, because it is applied to complete chains only.
@@
 \]
-{\color{red} where the juxtaposition of two string results in the   
-%$::$ is the 
-usual string concatenation operator.}
+where the usual string concatenation if represented by juxtaposition.
 
@@
  $P=\llbracket{\cal A},\gamma\rrbracket$ with
-$$P  = \restrict{\mathit{names}} \left(I~|~\prod_{a \in A}P_a ~|~ \mathit{Cxt}^1 ~|~ \prod_{s \in C_0}P_s ~|~ \prod_{s \notin C_0}\overline{P}_s\right).
+$$P  = \restrict{\mathit{names}} \left(I~|~\prod_{a \in A}P_a ~|~ \mathit{Cxt}_1 ~|~ \prod_{s \in C_0}P_s ~|~ \prod_{s \notin C_0}\overline{P}_s\right).
 $$
@@
   \item 
- $P'= \restrict{\mathit{names}} \left(I~|~\prod_{a \in A}P_a~|~ \mathit{Cxt}^2 ~|~ \prod_{s \in C_1 \cup D_1}P_s| \prod_{s \notin C_1 \cup D_1}\overline{P}_s\right)$.
+ $P'= \restrict{\mathit{names}} \left(I~|~\prod_{a \in A}P_a~|~ \mathit{Cxt}_2 ~|~ \prod_{s \in C_1 \cup D_1}P_s| \prod_{s \notin C_1 \cup D_1}\overline{P}_s\right)$.
  \end{enumerate}

### 2020-06-29 TCS_version/06-examples.tex
revisione

--- 2020-06-25 TCS_version/06-examples.tex
@@
 \label{ex:examples}
-
-Hereafter, we shall often omit topmost restrictions but shall take into account only transitions whose labels are complete chains, i.e. they do not contain 
+Semantically, the topmost restriction $\restrict{\mathit{names}}$ filters out any interaction with  virtual links, and releases a private interaction among all participants where all the channel names in the transition labels are hidden (their occurrences are all replaced by $\silent$).
+%Hereafter, we shall often omit topmost restrictions
+This amounts to require that only complete chains are computed by the interaction.
+In this section, for simplicity,  we shall often omit topmost restrictions $\restrict{\mathit{names}}$ from our encoding, but 
+ we shall take into account only transitions whose labels are complete chains, i.e. they do not contain 
 virtual links and start/end with the $\silent$ action symbol.
+This way it is possible to observe all channel names that occur in the interaction.
 
@@
 This example is inspired by the example in~\cite{BEMR11}, where
-a deterministic transition system is coded in the Reaction System
+a deterministic transition system is 
+%coded 
+encoded in the Reaction System
 framework.
@@
 \end{figure}
-\noindent
-In our $c\CNA$ 
-%encoding,
-{\color{red} embedding,} $q$, $w$, $a$, and $b$ became the $entities$, the context can only provide $a$ and $b$ and the reaction rules are as follows:\\
+
+At the level of RSs, the set of entities to consider is the union of sets of states and of labels of the transition system.
+Moreover, there is one reaction for each transition: its reactant set consists of the source state and transition label, its inhibitor set includes every other state and label, and its product set is the singleton with the target state.
+For the transition system in Fig.~\ref{fig:lts}, we take $S = \{q,w,a,b\}$ and the reaction rules are as follows:\\
 \[
@@
 \]
+
+Next we show how the above RS is encoded in c\CNA.
+
 \paragraph{Encoding of the rules}
-The encoding of the rules for reactions is  given in a parametric way, {\color{red} with $n \in \{1,2,3,4\}$}:
+The encoding of the rules for reactions is  given in a parametric way, with $n \in \{1,2,3,4\}$:
 \[
 \begin{array}{lcl}
-P_n(q,b,w,a,q) & \defeq & \pi_n(q,b,w,a,q).P_n(q,b,w,a,q)  \\
+P_n(q,b,w,a,q) & \defeq & \upsilon_n(q,b,w,a,q).P_n(q,b,w,a,q)  \\
 			&& +\\
-			&& \sum_{x\in\{\overline{q},\overline{b},w,a\}}\ \pi'_n(x).P_n(q,b,w,a,q)\\
+			&& \sum_{x\in\{\overline{q},\overline{b},w,a\}}\ \upsilon'_n(x).P_n(q,b,w,a,q)\\
 %			&& +\ \pi'(\overline{b}).P_n(q,b,w,a,q)\\
@@
 \begin{array}{lcl}
-\pi_n(q,b,w,a,q) &\defeq & \startchain{r_n}\chainedlink{q_i}{\noact}\chainedlink{\noact}{q_o}
+\upsilon_n(q,b,w,a,q) &\defeq & \startchain{r_n}\chainedlink{q_i}{\noact}\chainedlink{\noact}{q_o}
                                                     \chainedlink{b_i}{\noact}\chainedlink{\noact}{b_o}
@@
 			\chainend{p_{n+1}}\\[6pt]
-\pi'_n(x) &\defeq &  \startchain{r_n}\chainedlink{x_i}{\noact}\chainedlink{\noact}{x_o}\chainedlink{r_{n+1}}{\noact}\chainedlink{\noact}{p_n} \chainend{p_{n+1}} 
+\upsilon'_n(x) &\defeq &  \startchain{r_n}\chainedlink{x_i}{\noact}\chainedlink{\noact}{x_o}\chainedlink{r_{n+1}}{\noact}\chainedlink{\noact}{p_n} \chainend{p_{n+1}} 
 % \ + \   \startchain{r_n}\chainedlink{w_i}{\noact}\chainedlink{\noact}{w_o}\chainedlink{r_{n+1}}{\noact}\chainedlink{\noact}{p_n} \chainend{p_{n+1}}.P_n(q,b,w,a,q) \\
@@
 
-Then we have 
+Then, we have 
 \[
@@
 processed messenger RNA (mRNA) produced from that gene.
- {\color{red} In our example,
+In our example,
 %where 
 a gene $a$  codes for a protein $T$ when molecules 
-$G$ is present and $C$ is absent, and in the opposite situation $a$ codes for protein $T'$.}
+$G$ is present and $C$ is absent, and in the opposite situation $a$ codes for protein $T'$.
 This behavior is encoded in rules $1$ and $2$.
@@
 \[
-\begin{array}{lcl @{\hspace{0.5cm}}lcl @{\hspace{0.5cm}}lcl @{\hspace{0.5cm}}lcl}
-P_T & \defeq & P(T,\widetilde{T})& \overline{P}_T & \defeq & P(\overline{T},\widetilde{T}) & P_{T'}& \defeq & P(T',\widetilde{T'}) & \\
+\begin{array}{l @{\hspace{0.5cm}}l @{\hspace{0.5cm}}l @{\hspace{0.5cm}}l @{\hspace{0.5cm}}lcl}
+P_T \defeq  P(T,\widetilde{T})& \overline{P}_T  \defeq  P(\overline{T},\widetilde{T}) & P_{T'} \defeq  P(T',\widetilde{T'}) & 
 %\overline{P}_{T',\widetilde{T'}}& \defeq & P(\overline{T'},\widetilde{T'})
-\overline{P}_{T'}& \defeq & P(\overline{T'},\widetilde{T'})
+\overline{P}_{T'} \defeq  P(\overline{T'},\widetilde{T'})
 \end{array}
@@
 
-Also in this example we account for a nondeterministic context that can (nondeterministically) provide any combination of the entities: $C$, $G$ $F$:
+Also in this example we account for a nondeterministic context that can (nondeterministically) provide any combination of the entities $C$, $G$ $F$:
 \[

### 2020-06-29 TCS_version/06b-bisimulation.tex
revisione

@@
 Bisimulation also relates species that have identical solutions at all time points when starting from the same initial conditions.
-Both the mentioned formalisms~\cite{BMMT08,CTTV15} adopt a classical approach to bisimulation. 
+Both the fore mentioned formalisms~\cite{BMMT08,CTTV15} adopt a classical approach to bisimulation. 
 Albeit the bisimulation is a powerful tool for verifying if the behaviour of two different software 
-programs is indistinguishable, in the case of biological systems the classical bisimulation seems to be inappropriate, as it considers too many details.
-In fact, in a biological soup, a high number of interactions occur every seconds, and generally, biologists
-are only interested to analyse a small subset of them.
+programs is indistinguishable, in the case of biological systems the classical bisimulation seems to be inappropriate, as the labels of the transitions systems are too concrete.
+In fact, in a biological soup, a high number of interactions occur at every computational stage, and generally, biologists
+are only interested to analyse a small subset of them and to focus just on some entities.
 
 For this reason, we propose an alternative notion of bisimulation, that hereafter we call \emph{bio-simulation},
-that allows us  to compare two biological systems by restricting the observation to only the limited events of interest.
+that allows us  to compare two biological systems by restricting the observation to only a limited set of events of interest, which can be chosen according to the property one wants to investigate in an experiment.
 This allows one to tailor the equivalence to different applications and purposes.
@@
 
-In a way, at each step of the bisimulation game, we want to query our transition labels to get only the information we care about.
+In a way, we want to query our transition labels to extract only the information we care about.
 To this goal, we introduce a simple language
@@
 \item Has the entity $s_i$ been used by rule $r_j$ as reagent?
-\item Has the entity $s_i$ blocked the application of  rule $r_j$?
+\item Has the entity $s_i$ been blocked the application of  rule $r_j$?
 \item Has the entity $s_i$ been produced by rule $r_j$?
@@
 
-Next, we introduce an assertion language that operates on simple strings
+Next, we introduce an assertion language that operates on strings
 and that combines regular expression operators with conjunction and disjunction.
-We let $\mathit{syms} = \mathit{decs}\cup\mathit{reacts}\cup\mathit{prods}\cup\{\mathit{cxt}\}$ be the set of symbolic names used in our assertion language.
+We let $\mathit{names}$ be the set of symbolic names used in our assertion language.
 
@@
 \label{def:assetionl}
-Assertions are built from: 
+Atomic assertions $\zeta$ and general assertions $\mathsf{F}$ are built from the syntax below: 
 $$
@@
 \zeta & ::= & 
-\alpha ~\mid ~ 
+\eta ~\mid ~ 
 ? ~\mid ~ 
@@
 %\noindent where $s \in {\cal C}$, and  $\oplus \in \{\wedge, \vee, \to \}$. 
-where $\alpha \in \mathit{syms}$ and $N\subseteq \mathit{syms}$.
+where $\eta \in \mathit{names}$ and $N\subseteq \mathit{names}$.
 %, and  $\oplus \in \{\wedge, \vee \}$. 
@@
 
-Roughly, a \emph{singleton} assertion $\zeta$ denotes either a string composed by a single symbol $\alpha$ (one of the symbols in the set $\mathit{syms}$ for denoting a particular entity, rule, production or context), or
-the wildcard $?$ that stands for any symbol, or the pattern $[N]$ that stands for any of the strings composed by a single symbol in the set $N$. Clearly $?$  is just a shorthand for $[\mathit{syms}]$.
+Roughly, an atomic assertion $\zeta$ denotes either a string composed by a single name $\eta$ (one of the symbols in the set $\mathit{names}$ for denoting a particular entity, rule, production or context), or
+the wildcard $?$ that stands for any symbol, or the pattern $[N]$ that stands for any string composed by a single symbol in the set $N$. Clearly $?$  is just a shorthand for $[\mathit{names}]$.
 We write $\mathbf{0}$ for $[\emptyset]$ and $[s_1,...,s_n]$ instead of $[\{s_1,...,s_n\}]$.
 
-An \emph{assertion} $\mathsf{F}$ is either the empty string $\epsilon$, a singleton assertion $\zeta$, the concatenation of two assertions $\mathsf{F}_1::\mathsf{F}_2$, the replication of $\mathsf{F}$ for 1 or more times $\mathsf{F}^+$, the replication of $\mathsf{F}$ for 0 or more times $\mathsf{F}^*$, the disjunction of two assertions $\mathsf{F}_1 \vee \mathsf{F}_2$ or their conjunction $\mathsf{F}_1 \wedge \mathsf{F}_2$.
-We write $\star$ as an abbreviation of $?^*$.
+An assertion $\mathsf{F}$ is either the empty string $\epsilon$, an atomic assertion $\zeta$, the concatenation of two assertions $\mathsf{F}_1::\mathsf{F}_2$, the replication of $\mathsf{F}$ for 1 or more times $\mathsf{F}^+$, the replication of $\mathsf{F}$ for 0 or more times $\mathsf{F}^*$, the disjunction of two assertions $\mathsf{F}_1 \vee \mathsf{F}_2$ or their conjunction $\mathsf{F}_1 \wedge \mathsf{F}_2$.
+We denote by $\star$ the assertion $?^*$.
 
-An assertion denotes a set of strings over the alphabet $\mathit{syms}$ as expected.
+An assertion denotes a set of strings over the alphabet $\mathit{names}$ as expected.
+Below we let $\wp(X)$ denote the powerset of a set $X$.
 
 \begin{definition}[Semantics of Assertions]
- We define $\llbracket \mathsf{F}\rrbracket \subseteq\wp(\mathit{syms}^*)$ by induction on the structure of $\mathsf{F}$: 
+ We define $\llbracket \mathsf{F}\rrbracket \subseteq\wp(\mathit{names}^*)$ by induction on the structure of $\mathsf{F}$: 
 \[
@@
  & \llbracket \mathsf{F}^+ \rrbracket &\defeq& \llbracket \mathsf{F} \rrbracket^+  \\
- \llbracket ? \rrbracket &\defeq&   \mathit{syms}
+ \llbracket ? \rrbracket &\defeq&   \mathit{names}
  & \llbracket \mathsf{F}^* \rrbracket &\defeq& \llbracket \mathsf{F} \rrbracket^*  \\
@@
 \begin{example}
-The formulas corresponding to the sample queries listed in Example~\ref{properties_ex} are as follows:
+The assertions corresponding to the sample queries listed in Example~\ref{properties_ex} are as follows:
 \begin{enumerate}
@@
 
-Now, we introduce a slightly modified version of the Hennessy Milner Logic~\cite{HM80}, called bioHML; due to the reasons we explained above, we do not want to look at the complete transition labels, thus we rely on our simple assertion language to make it parametric to the assertion $\mathsf{F}$ of interest:
+Now, we introduce a slightly modified version of the Hennessy Milner Logic (HML)~\cite{HM80}, called bioHML; due to the reasons we explained above, we do not want to look at the complete transition labels, thus we rely on our simple assertion language to make it parametric w.r.t the assertion $\mathsf{F}$ of interest:
 
@@
 We observe that, as expected, for any bioHML formula $G$ and process $P$ we have $\overline{\overline{\mathsf{G}}} = \mathsf{G}$ and  $P \entails \overline{\mathsf{G}}$ iff  $P\not\entails \mathsf{G}$.
-\begin{definition}
+
+\begin{definition}[Bio-logical equivalence]
 We let $\mathcal{L}_{\mathsf{F}}$ be the set of all bioHML formulas that respects $\mathsf{F}$ and 

### 2020-06-29 TCS_version/07-discussion.tex
revisione

@@
 
-\section{Enhanced  Reaction Systems}
+\section{Towards Enhanced  Reaction Systems}
 \label{sec:discussion}
@@
 \label{ex:twoRS}
-Let  $rs1$ and $rs2$ be two RSs, defined by the rules in Table~\ref{tab:rules}.
+Let  $rs1$ and $rs2$ be two RSs, defined, respectively, by the reactions 
+%in Table~\ref{tab:rules}.
 %such that $rs2$ provides the entity $s1$ to $rs1$.
-
-\begin{table}[t]
-\centering
-\begin{tabular}{|c|c|}
-\hline
-$rs1$  & $rs2$ \\
-\hline
-%$r1_1=(s1,\ , s2)$ & $r2_1=(s3,\  ,s1)$\\
-$a_1=(s,\ , x)$ & $a_2=(y,\  ,s)$\\
-%$r1_2=(\ , s2,s1)$ & $r2_2=(s1,\ ,s3)$\\
-\hline 
-\end{tabular}
-\caption{The two Reaction Systems $rs1$ and $rs2$.}
-\label{tab:rules}
-\end{table}
- \noindent
+$a_1=(s,\emptyset , x)$ and $a_2=(y,\emptyset  ,s)$.
+%and 
+%
+%
+%\begin{table}[t]
+%\centering
+%\begin{tabular}{|c|c|}
+%\hline
+%$rs1$  & $rs2$ \\
+%\hline
+%%$r1_1=(s1,\ , s2)$ & $r2_1=(s3,\  ,s1)$\\
+%$a_1=(s,\emptyset , x)$ & $a_2=(y,\emptyset  ,s)$\\
+%%$r1_2=(\ , s2,s1)$ & $r2_2=(s1,\ ,s3)$\\
+%\hline 
+%\end{tabular}
+%\caption{The two Reaction Systems $rs1$ and $rs2$.}
+%\label{tab:rules}
+%\end{table}
+% \noindent
 Now, we set our example such that the two contexts, for  $rs1$ and $rs2$, do not provide any entities.
@@
 %
-We need to modify our {\color{red} embedding}
+We need to modify our encoding
 %translation 
-technique to suite this new setting. 
+technique to suit this new setting. 
 As we do not model contexts, we introduce  \emph{dummy} channel names $dx$ and $dss$ to model the absence of entities.
@@
 %First, we translate the rules in $rs1$:
-First, we translate the reaction in $rs1$:
+First, we translate the reaction in $rs1$, by setting
+$\llbracket a_1\rrbracket \defeq P_{a_1}$ where:
 \[
@@
 %\startchain{\silent}\chainedlink{\overline{s1}}{\noact}\chainedlink{\noact}{\overline{s1}}\chainend{r1_2}.P_{r1_1}\\
-\llbracket a_1\rrbracket \defeq P_{a_1}&\defeq &
+ P_{a_1}&\defeq &
 \startchain{\silent}\chainedlink{s_i}{{\color{blue}\noact}}\chainedlink{{\color{blue}\noact}}{s_o}\chainedlink{\widetilde{x}_i}{{\color{gray}\noact}}\chainedlink{{\color{gray}\noact}}{\widetilde{x}_o}\chainend{a_2}.P_{a_1} +
@@
 Please note, that prefixes of process $P_{a_1}$ end with the channel name $a_2$, as the link chain is now connected with the  reaction of  $rs2$.
-The {\color{red} encoding} 
+The encoding
 %translation 
-for the entities follows.
+for the entities is given by setting $\llbracket s \rrbracket \defeq P_{s}$ and $\llbracket x \rrbracket \defeq  P_{x}$, where:
 \[
-\begin{array}{lclcl}
-\llbracket s \rrbracket &\defeq &P_{s} &\defeq & \startchain{{\color{blue}s_i}}\chainedlink{{\color{blue}s_o}}{\noact}\chainedlink{\noact}{{\color{blue}\widehat{s}_i}}\chainend{{\color{blue}\widehat{s}_o}}.P_{s} + \startchain{s_i}\chainend{s_o}.\overline{P_{s}}\\
-&& \overline{P_{s}}&\defeq &\startchain{\overline{s}_i}\chainedlink{\overline{s}_o}{\noact}\chainedlink{\noact}{\widehat{s}_i}\chainend{\widehat{s}_o}.P_{s}+\startchain{\overline{s}_i}\chainend{\overline{s}_i}.\overline{P_{s}}\\
+\begin{array}{lcl}
+P_{s} &\defeq & \startchain{{\color{blue}s_i}}\chainedlink{{\color{blue}s_o}}{\noact}\chainedlink{\noact}{{\color{blue}\widehat{s}_i}}\chainend{{\color{blue}\widehat{s}_o}}.P_{s} + \startchain{s_i}\chainend{s_o}.\overline{P_{s}}\\
+\overline{P_{s}}&\defeq &\startchain{\overline{s}_i}\chainedlink{\overline{s}_o}{\noact}\chainedlink{\noact}{\widehat{s}_i}\chainend{\widehat{s}_o}.P_{s}+\startchain{\overline{s}_i}\chainend{\overline{s}_i}.\overline{P_{s}}\\
 %\begin{array}{lclclll}
-\llbracket x \rrbracket &\defeq & P_{x}& \defeq &\startchain{\widetilde{x}_i}\chainend{\widetilde{x}_o}.P_{x} +  \startchain{dx_i}\chainend{dx_o}.\overline{P_{x}}\\
+P_{x}& \defeq &\startchain{\widetilde{x}_i}\chainend{\widetilde{x}_o}.P_{x} +  \startchain{dx_i}\chainend{dx_o}.\overline{P_{x}}\\
 %\\
@@
 %\\
-&& \overline{P_{x}}&\defeq &\startchain{{\color{gray}\widetilde{x}_i}}\chainend{{\color{gray}\widetilde{x}_o}}.P_{x} +  \startchain{dx_i}\chainend{dx_o}.\overline{P_{x}}\\
+\overline{P_{x}}&\defeq &\startchain{{\color{gray}\widetilde{x}_i}}\chainend{{\color{gray}\widetilde{x}_o}}.P_{x} +  \startchain{dx_i}\chainend{dx_o}.\overline{P_{x}}\\
 %\end{array}
@@
 \noindent
-The {\color{red} encoding}
+The encoding
 %translation 
-for $rs2$ follows.
+for $rs2$ is given by $\llbracket a_2\rrbracket \defeq P_{a_2}$, where:
 
@@
 \begin{array}{lcl}
-\llbracket a_2\rrbracket \defeq P_{a_2}&\defeq &\startchain{a_2}\chainedlink{y_i}{{\color{gray}\noact}}\chainedlink{{\color{gray}\noact}}{y_o}\chainedlink{\widetilde{ss}_i}{{\color{red}\noact}}\chainedlink{{\color{red}\noact}}{\widetilde{ss}_o}\chainend{\silent}.P_{a_2}+
+P_{a_2}&\defeq &\startchain{a_2}\chainedlink{y_i}{{\color{gray}\noact}}\chainedlink{{\color{gray}\noact}}{y_o}\chainedlink{\widetilde{ss}_i}{{\color{red}\noact}}\chainedlink{{\color{red}\noact}}{\widetilde{ss}_o}\chainend{\silent}.P_{a_2}+
 \startchain{a_2}\chainedlink{\overline{y}_i}{\noact}\chainedlink{\noact}{\overline{y}_o}    
@@
 \noindent
-In the {\color{red} encoding}
+In the encoding
 %translation 
@@
 $s$ ($ss$ in $rs2$) to be provided in $rs1$. Every time $ss$ is produced in $rs2$,  a virtual link is
-created to synchronise with $rs1$   on link $\link{\widehat{s}_i}{\widehat{s}_o}$:
+created to synchronise with $rs1$   on link $\link{\widehat{s}_i}{\widehat{s}_o}$. 
+To this purpose we define $\llbracket ss \rrbracket \defeq P_{ss}$ and $\llbracket y\rrbracket\defeq  P_{y}$, where:
+
 \[
-\begin{array}{lclcl}
-\llbracket ss \rrbracket& \defeq& P_{ss}&\defeq &\startchain{\widetilde{ss}_i}\chainedlink{\widehat{s}_i}{\noact}\chainedlink{\noact}{\widehat{s}_o}\chainend{\widetilde{ss}_o}.P_{ss}+
+\begin{array}{lcl}
+P_{ss}&\defeq &\startchain{\widetilde{ss}_i}\chainedlink{\widehat{s}_i}{\noact}\chainedlink{\noact}{\widehat{s}_o}\chainend{\widetilde{ss}_o}.P_{ss}+
 \startchain{dss_i}   \chainend{dss_o}.\overline{P_{ss}} \\
-
 %\end{array}
@@
 %\begin{array}{lcllcl}
-&&\overline{P_{ss}}&\defeq &\startchain{{\color{red}\widetilde{ss}_i}}\chainedlink{{\color{red}\widehat{s}_i}}{{\color{blue}\noact}}\chainedlink{{\color{blue}\noact}}{{\color{red}\widehat{s}_o}}\chainend{{\color{red}\widetilde{ss}_o}}.P_{ss}
+\overline{P_{ss}}&\defeq &\startchain{{\color{red}\widetilde{ss}_i}}\chainedlink{{\color{red}\widehat{s}_i}}{{\color{blue}\noact}}\chainedlink{{\color{blue}\noact}}{{\color{red}\widehat{s}_o}}\chainend{{\color{red}\widetilde{ss}_o}}.P_{ss}
 +
 \startchain{dss_i}\chainend{dss_o}.\overline{P_{ss}}\\
-\llbracket y\rrbracket&\defeq & P_{y}&\defeq &\startchain{{\color{gray}y_i}}\chainend{{\color{gray}y_o}}.\overline{P_{y}} \\
-&&\overline{P_{y}}&\defeq &
+P_{y}&\defeq &\startchain{{\color{gray}y_i}}\chainend{{\color{gray}y_o}}.\overline{P_{y}} \\
+\overline{P_{y}}&\defeq &
 %\startchain{\overline{y}_i}\chainedlink{\overline{y}_o}{\noact}\chainedlink{\noact}{\widetilde{y}_i}\chainend{\widetilde{y}_o}.P_{y}

### 2020-06-29 TCS_version/08-conclusion.tex
revisione

@@
 
-In this paper we have introduced c\CNA, a variant of the {\tt link}-calculus
-where prefixes are link chains and no more single links, as it was 
-briefly described in the future work section in~\cite{BBB17}.
-This variant allowed us to define 
+In this paper we have introduced c\CNA, that generalises \CNA\ by allowing the use of prefixes that are link chains and not just single links.
+This extension was initially described in the future work section of~\cite{BBB17}.
+Thanks to this enhancement, c\CNA\ allowed us to define 
 %shown 
-a faithful embedding of 
-RSs,
-%Reaction Systems,
-an emerging formalism to model computationally biochemical systems.
-%into multilink processes.  
-This {\color{red} embedding}
+a faithful encoding of 
+Reaction Systems in a process algebraic framework.  
+This encoding
 %translation 
 shows several benefits.
-For instance, in our paper the context of a RS can work non deterministically and it 
-is recursively defined.
-%  the context behaviour is expressed recursively and no be expressed recursively. 
-Also, entity mutations can be expressed easily and different reaction systems can
-communicate between them.
+First, contexts of RSs can be easily defined recursively and exhibit non deterministic behaviour.
+Second, the operational semantics is defined in a compositional way by a set of SOS inference rules.
+Third, we have defined a new assertion language, which allows us to specify
+the properties to be verified over the labels of the operational semantics.
+Assertions can be used to tailor the classical notion of bisimilarity and Hennessy-Milner logic to focus on some particular aspects or experiments.
+We have called \emph{bio-similarity} the induced notion of equivalence.
 
-We have then modified the classical notion (in concurrency) of bisimulation
-making it more suitable for proving properties of 
-RSs
-%Reaction Systems 
-in our
-%framework.
-{\color{red} embedding.}
-We have defined a new assertion language, which allows us to specify
-the properties to be verified, and have shown that it extends Hennessy-Milner logic
-to our 
-%framework.
-{\color{red} embedding.}
-We believe that our work can make 
-possible to investigate how to integrate our methodology
+We are currently investigating how to integrate our methodology
 with other formal techniques to prove 
-properties of the modeled systems \cite{CFHOT15,OCHF16,BBGLBH2017}.
-
-We are confident that our embedding can contribute to extend the applications
-of 
-RSs
-%Reaction Systems
-to diverse fields of computer science, and life
-sciences.
+properties of the modeled systems, along the lines in~\cite{CFHOT15,OCHF16,BBGLBH2017}.
+Moreover, we are considering possible enhancements of RSs based on entity mutation and on the possibility for two RSs to exchange entities.
 
@@
 %framework,
-{\color{red} embedding,}
-with an automatic translation from RSs to  {\tt link}-calculus. 
-We can also exploit the implementation of  the symbolic semantics of 
+embedding,
+with an automatic translation from RSs to  {\tt link}-calculus, so to exploit the implementation of  the symbolic semantics of 
 the {\tt link}-calculus~\cite{BrodoO17} that can be found in~\cite{tool}.
-
-We also plan to build on the work in this paper to extend the framework
-of RSs towards a model where different RSs are allowed to communicate.

### 2020-06-30 TCS_version/09-appendix.tex
appendice Cxt_1

@@
 Since the context processes is involved, then all entities processes are also involved.
-Then, all the processes $I$, $P_a$, $P_s$ (or $\overline{P_s}$), and $\mathit{Cxt}$  must participate to each transition.
+Then, all the processes $I$, $P_a$, $P_s$ (or $\overline{P_s}$), and $\mathit{Cxt}_1$  must participate to each transition.
  \item for each reaction $aj \in A$, the c\CNA \ code of  $P_{aj}$ leaves one virtual link between two solid links of the types $\link{r_j}{\dots} \chainedlink{s_{i}}{\noact}\chainedlink{\noact}{s_{o}}\link{}{\dots}\link{}{r_{j+1}} {\dots}\link{p_j}{\dots}\link{}{p_{j+1}}$.
@@
 In fact, there is no other way to generate a solid chain from $s_{i}$ to $s_{o}$.
- The same reasoning holds for the processes $\mathit{Cxt}$ and for all the decorated versions of $s_i,s_o$.
+ The same reasoning holds for the processes $\mathit{Cxt}_1$ and for all the decorated versions of $s_i,s_o$.
 % $P_{s_1}$ can not participate with a link chain, as by juxtaposing  $\link{s_{1i}}{s_{1o}}$, not valid
@@
   and all the annotated versions of 
- $s_i$, $s_o$ are restricted. Also, processes $\mathit{Cxt}$ always requires the 
+ $s_i$, $s_o$ are restricted. Also, processes $\mathit{Cxt}_1$ always requires the 
  interaction with $P_s$ on either on channels $\widehat{s}_i$, $\widehat{s}_o$ or on channels 
  $\underline{s}_i$, $\underline{s}_o$.
- It derives that all the processes: $P_a$  (coding the behaviour of reaction $a \in A$), $P_s$ (coding the behaviour of entity $s \in S$), and $\mathit{Cxt}$ (coding the behaviour of the context regarding all the entities) have been involved in the transition.
+ It derives that all the processes: $P_a$  (coding the behaviour of reaction $a \in A$), $P_s$ (coding the behaviour of entity $s \in S$), and $\mathit{Cxt}_1$ (coding the behaviour of the context regarding all the entities) have been involved in the transition.
 
@@
 %involve the entity $s$.\\
-Now, we consider the structure of the process $\mathit{Cxt}^1$. 
-By Definition~\ref{def:trans}, $Cxt^1$ is the unique process encoding the behaviour of the context regulating the supply of any entity.
+Now, we consider the structure of the process $\mathit{Cxt}_1$. 
+By Definition~\ref{def:trans}, $Cxt_1$ is the unique process encoding the behaviour of the context regulating the supply of any entity.
 
@@
 \item[(e)] 
-The code of the process $Cxt^1$ that provides $s$ and not $e$ has the following structure: 
+The code of the process $Cxt_1$ that provides $s$ and not $e$ has the following structure: 
 $$
-\mathit{Cxt}^1 \defeq
+\mathit{Cxt}_1 \defeq
 \link{\mathit{cxt}}{\dots}
@@
 \chainedlink{\underline{e}_i}{\noact}\chainedlink{\noact}{}\link{\underline{e}_o}{\dots} 
-\chainend{p_{1}}. \mathit{Cxt}^{2} .
+\chainend{p_{1}}. \mathit{Cxt}_{2} .
 $$

### 2020-07-19 TCS_version/01-introduction.tex
revisione
piccolezze

@@
 %into account the quantity of reagents effectively present. They
-consider if a necessary reagent is or not present, and likewise they
-consider if an inhibiting molecule is or not present. 
+consider if a necessary reagent is present or not, and likewise they
+consider if an inhibiting molecule is present or not. 
 The  possible reactants and inhibitors are called `entities'.
@@
 to reduce the coding when suitable assumptions are made about the provision of entities.
-
-As we have remarked, in our 
+{\color{red}
+\st{ As we have remarked, in our 
 %translation,
@@
 This interaction is not foreseen in
-the basic RS framework, as it can only happen once the context is provided.
-The proposed network of RSs~\cite{BLR20} allows
-any RS in the network to receive entities 
-produced from its neighbours (that will represent its context), with the hypothesis that  
-RSs without neighbours will receive no entities 
-from the context. 
-In our case, a RS can receive some entities from the context and some others from a second RS; moreover we can represent contexts having a recursive, non-determinist behaviour.
-%
-By exploiting recursion, the kind of interactions which can be 
-defined can be complex and expressive.
-Example~\ref{ex:twoRS} and more in general the discussion in 
-Section~\ref{subsec:comRS}
-%~\ref{sec:discussion}
- show that 
-the interaction between RSs can help to model new scenarios.
+the basic RS framework, as it can only happen once the context is provided.}}
+
+%The proposed network of RSs~\cite{BLR20} allows
+%any RS in the network to receive entities 
+%produced from its neighbours (that will represent its context), with the hypothesis that  
+%RSs without neighbours will receive no entities 
+%from the context. 
+%In our case, a RS can receive some entities from the context and some others from a second RS; moreover we can represent contexts having a recursive, non-determinist behaviour.
+%%
+%By exploiting recursion, the kind of interactions which can be 
+%defined can be complex and expressive.
+%Example~\ref{ex:twoRS} and more in general the discussion in 
+%Section~\ref{subsec:comRS}
+%%~\ref{sec:discussion}
+% show that 
+%the interaction between RSs can help to model new scenarios.
 
@@
 to exactly keep trace of those elements that  dissolve in the next 
-computation step.
+computation step.\\
+%
+The proposed network of RSs~\cite{BLR20} allows
+any RS in the network to receive entities 
+produced from its neighbours (that will represent its context), with the hypothesis that  
+RSs without neighbours will receive no entities 
+from the context. 
+%In our case
+In the idea we sketch, a RS can receive some entities from the context and some others from a second RS; moreover we can represent contexts having a recursive, non-determinist behaviour.
+%
+By exploiting recursion, the kind of interactions which can be 
+defined can be complex and expressive.
+Example~\ref{ex:twoRS} and more in general the discussion in 
+Section~\ref{subsec:comRS}
+%~\ref{sec:discussion}
+ show that 
+the interaction between RSs can help to model new scenarios.
 
@@
 \begin{itemize}
-\item we present several new examples to illustrate our 
+\item we present several new examples (those in Sections~\ref{ex:examples} and~\ref{sec:biowork}) to illustrate our 
  encoding,
@@
 describes RSs and their semantics (interactive processes).
-Section \ref{sec:ccna} describes briefly the c\CNA \ process algebra and
+Section \ref{sec:ccna} briefly describes  the c\CNA \ process algebra and
 its operational semantics.

### 2020-07-19 TCS_version/06-examples.tex
piccolezze

@@
 \]}
-The parts in bold are provided by the entity processes, the other parts are provided by the processes encoding the reactions and by the process encoding the context (starting at $\mathit{cxt}$ and ended at $p_1$.
+The parts in bold are provided by the entity processes, the other parts are provided by the processes encoding the reactions and by the process encoding the context (starting at $\mathit{cxt}$ and ending at $p_1$.
 In the label we can read that the rules $1$ and $4$ have been not executed because the entity $a$ is absent, 
@@
 coding for multiple proteins. 
-In practice, particular exons of a gene may be included within or excluded from the final, 
+In practice, particular exons of a gene may be included within or excluded from the final 
 processed messenger RNA (mRNA) produced from that gene.
@@
 
-The encoding of the entity $F$ that can be provided by the context follows:
+The encoding of the entity $F$ that can only be provided by the context follows:
 \[

### 2020-07-19 TCS_version/09-appendix.tex
piccolezze

@@
  \begin{proof}
- First, we note that all the channels in the system are restricted, see Def.~\ref{def:trans}, then
+ First, we note that all the channels in the system are restricted, see Definition~\ref{def:trans}, then
  it holds that the transition labels are of the form  $v=\link{\silent}{\silent}\dots\link{\silent}{\silent}$.\\

### 2020-07-23 TCS_version/09-appendix.tex
piccole correzioni

@@
 \item[$\sim_{\mathsf{F}}\ \subseteq\ \equiv_{\mathcal{L}_{\mathsf{F}}}$:]
-Given any two processes $P\sim_{\mathsf{F}} Q$ we need to prove that for any bioHML formula $\mathsf{G}$ we have $P\entails \mathsf{G}$ iff $Q\entails \mathsf{G}$. 
+Given any two processes $P\sim_{\mathsf{F}} Q$ we need to prove that for any bioHML formula $\mathsf{G} \in \mathcal{L}_{\mathsf{F}}$ we have $P\entails \mathsf{G}$ iff $Q\entails \mathsf{G}$. 
 Without loss of generality, we prove that $P\entails \mathsf{G}$ implies $Q\entails \mathsf{G}$.
@@
 Let $\mathcal{Q} = \{Q'_1,...,Q'_n\}$.
-By hypothesis all processes in $\mathcal{Q}$ must not be bio-logically equivalent to $P'$, hence for any $i\in[1,n]$ there exists a bioHML formula $\mathsf{G}_i$ such that $P'\entails \mathsf{G}_i$ and $Q'_i\not\entails \mathsf{G}_i$ (if it was the opposite, $P'\not\entails \mathsf{H}_i$ and $Q'_i\entails \mathsf{H}_i$ for some $\mathsf{H}_i$, we can use the converse formula $\mathsf{G}_i\defeq \overline{\mathsf{H}}_i$).
+By hypothesis all processes in $\mathcal{Q}$ must not be bio-logically equivalent to $P'$, hence for any $i\in[1,n]$ there exists a bioHML formula $\mathsf{G}_i  \in \mathcal{L}_{\mathsf{F}}$ such that $P'\entails \mathsf{G}_i$ and $Q'_i\not\entails \mathsf{G}_i$ (if it was the opposite, $P'\not\entails \mathsf{H}_i$ and $Q'_i\entails \mathsf{H}_i$ for some $\mathsf{H}_i$, we can use the converse formula $\mathsf{G}_i\defeq \overline{\mathsf{H}}_i$).
 But then the formula $\mathsf{G}\defeq \langle \mathsf{F}\rangle(\mathsf{G}_1\wedge\dots\wedge \mathsf{G}_n)$ is such that $P\entails \mathsf{G}$ and $Q\not\entails \mathsf{G}$, contradicting the assumption $P\equiv_{\mathcal{L}_{\mathsf{F}}} Q$.

### 2020-07-23 TCS_version/01-introduction.tex
piccole modifiche

@@
 \begin{itemize}
+\item 
+\begin{color}{blue}
+RSs are encoded in a modular way, in the sense that all constituents of a RS, namely entities, reactions and the context, are seen as interacting  c\CNA \ processes; in principle, this allows to study the behaviour of any constituent in isolation, contrary to what happens in the basic RS framework, where the evolution is possible only when all constituents are given;
+\end{color}
+
 \item any context is represented as an ordinary c\CNA \ process, which allows us to specify recursive and non-deterministic contexts in a natural way; 
@@
 computation is also deterministic and matches the evolution of the underlying RS;
+
 \item we define an assertion language to specify local properties of 
@@
 called bio-similarity;
+
 \item we show that our assertions can be used to extend the Hennessy-Milner's logic to our 
@@
 The examples that we propose also show that some optimisations are possible
-to reduce the coding when suitable assumptions are made about the provision of entities.
-{\color{red}
-\st{ As we have remarked, in our 
-%translation,
-encoding
-the entities and the reactions
-get the ability to interact in a synchronous manner. 
-This interaction is not foreseen in
-the basic RS framework, as it can only happen once the context is provided.}}
+to reduce the coding efforts when suitable assumptions are made about the provision of entities.
+%{\color{red}
+%\st{ As we have remarked, in our 
+%%translation,
+%encoding
+%the entities and the reactions
+%get the ability to interact in a synchronous manner. 
+%This interaction is not foreseen in
+%the basic RS framework, as it can only happen once the context is provided.}}
 
@@
 %SOS 
-structural operational semantics for RS; it seems not the case
+Structural Operational Semantics for RS; it seems not the case
 that a deterministic behaviour, as the one of the RS, once the initial state is 
@@
 to exactly keep trace of those elements that  dissolve in the next 
-computation step.\\
+computation step.
+
 %
@@
 %In our case
-In the idea we sketch, a RS can receive some entities from the context and some others from a second RS; moreover we can represent contexts having a recursive, non-determinist behaviour.
+In the idea we sketch, a RS can receive some entities from the context and some others from a second RS; moreover we can represent contexts having a recursive, non-deterministic behaviour.
 %
@@
 it.
-Section~\ref{ex:examples} presents a couple of examples with computer science and biological applications.
+Section~\ref{ex:examples} presents a couple of examples with automata theory and biological applications.
 %Section \ref{ex:lactose} shows a more complex example taken from

### 2020-07-24 TCS_version/06-examples.tex
piccole correzioni
reactions

@@
 Moreover, there is one reaction for each transition: its reactant set consists of the source state and transition label, its inhibitor set includes every other state and label, and its product set is the singleton with the target state.
-For the transition system in Fig.~\ref{fig:lts}, we take $S = \{q,w,a,b\}$ and the reaction rules are as follows:\\
+For the transition system in Fig.~\ref{fig:lts}, we take $S = \{q,w,a,b\}$ and the reactions are as follows:\\
 \[
@@
 
-\paragraph{Encoding of the rules}
-The encoding of the rules for reactions is  given in a parametric way, with $n \in \{1,2,3,4\}$:
+\paragraph{Encoding of the reactions}
+The encoding of the  reactions is  given in a parametric way, with $n \in \{1,2,3,4\}$:
 \[
@@
 
-In fact the presence/absence of $q$ and $w$ will be exploited by at least one rule and at most three rules.
+In fact the presence/absence of $q$ and $w$ will be exploited by at least one reaction and at most three reactions.
 
-Here, for the entities $a$ and $b$ that can be provided by the context but not produced by the rules, we let:
+Here, for the entities $a$ and $b$ that can be provided by the context but not produced by  reactions, we let:
 \[
@@
 The parts in bold are provided by the entity processes, the other parts are provided by the processes encoding the reactions and by the process encoding the context (starting at $\mathit{cxt}$ and ending at $p_1$.
-In the label we can read that the rules $1$ and $4$ have been not executed because the entity $a$ is absent, 
-the rule $3$ has been not applied because the entity $w$ is absent, then only rule $2$ has been applied, and it has produced the entity $q$.
+In the label we can read that reactions $1$ and $4$ have been not executed because the entity $a$ is absent, 
+the reaction $3$ has been not applied because the entity $w$ is absent, then only reaction $2$ has been applied, and it has produced the entity $q$.
 Also,  the context provides entity $a$, that will be available in the next state, and not the entity $b$.
@@
 $G$ is present and $C$ is absent, and in the opposite situation $a$ codes for protein $T'$.
-This behavior is encoded in rules $1$ and $2$.
-Then, rule 3 codes for the production of $C$ when proteins $T$ and $F$ are present, and $T'$ absent;
-rule 4 codes for the production of $G$ when  proteins $T'$ is present and $F$ is absent.
+This behavior is encoded in reactions $1$ and $2$.
+Then, reaction $3$ codes for the production of $C$ when proteins $T$ and $F$ are present, and $T'$ absent;
+reaction $4$ codes for the production of $G$ when  proteins $T'$ is present and $F$ is absent.
 
-\paragraph{Encoding of the rules}
+\paragraph{Encoding of the reactions}
 
-The encoding of the rules for reactions is  given in a parametric way:
+The encoding of reactions is  given in a parametric way:
 \[

### 2020-07-27 TCS_v1/05-translation.tex
pulizia e riorganizzazione

--- 2020-04-05 05-translation.tex
@@
 
-\noindent
 We exploit names $r_j,p_j$ to join the chains provided by the application of all the reactions. 
@@
 %\left \{
-^{cxt} \backslash  \blockchain{s_i}{s_o}{s \in C_n} \, \link{}{} \blockchain{\overline{e}_i}{\overline{e}_o}{e \not\in C_n} \,  \backslash_{p_1}. \mathit{Cxt}^{n+1}
+^{cxt} \backslash  \blockchain{s_i}{s_o}{s \in C_n}  \link{}{} \blockchain{\overline{e}_i}{\overline{e}_o}{e \not\in C_n}   \backslash_{p_1}. \mathit{Cxt}^{n+1}
 %+ \\

### 2020-07-27 TCS_v1/06-examples.tex
pulizia e riorganizzazione

--- 2020-04-05 06-examples.tex
@@
 \]
-\noindent
+
 Then we have 
 \[
-\begin{array}{lcl @{\hspace{0.5cm}}lcl  }
+\begin{array}{lcl @{\hspace{1cm}}lcl  }
 P_1 & \defeq & P_1(q,a,w,b,q) & P_3 & \defeq & P_3(w,a,q,b,w)  \\
@@
 \]
-\noindent
 and we put, as usual,
@@
 \[
-r_1~\overline{a}~r_2~q~b~\overline{w}~\overline{a}~r_3~\overline{a}~r_4~\overline{w}~\mathit{cxt}~\widehat{a}~\underline{b}~p_1~p_2~\widetilde{w}~p_3~p_4\]
+r_1~\overline{a}~r_2~q~b~\overline{w}~\overline{a}~r_3~\overline{a}~r_4~\overline{w}~\mathit{cxt}~\widehat{a}~\underline{b}~p_1~p_2~\widetilde{w}~p_3~p_4 .
+\]
 
@@
 P_T & \defeq & P(T,\widetilde{T})& \overline{P}_T & \defeq & P(\overline{T},\widetilde{T}) & P_{T'}& \defeq & P(T',\widetilde{T'}) & \\
-\overline{P}_{T',\widetilde{T'}}& \defeq & P(\overline{T'},\widetilde{T'})\\\
+\overline{P}_{T',\widetilde{T'}}& \defeq & P(\overline{T'},\widetilde{T'})
 \end{array}
 \]
+
 The entities that can be produced by the reactions and that can be  provided by the context are as follows:

### 2020-07-27 TCS_v1/07-discussion.tex
pulizia e riorganizzazione

--- 2020-04-05 07-discussion.tex
@@
 We sketch how it is possible to program two RSs encodings, in a way that
-the entities that usually come from context of one RS will be provided instead from the other RS.
+the entities that usually come from the context of one RS will be provided instead from the other RS.
 

### 2020-07-27 TCS_v2/06b-bisimulation.tex
pulizia e riorganizzazione

--- 2020-06-29 TCS_version/06b-bisimulation.tex
@@
 \begin{enumerate}
-\item Has the entity $s_i$ been used by rule $r_j$ as reagent?
-\item Has the entity $s_i$ been blocked the application of  rule $r_j$?
-\item Has the entity $s_i$ been produced by rule $r_j$?
-\item Has the entity $s_i$ been produced by some rule?
+\item Has the entity $s_i$ been used by reaction $r_j$ as reagent?
+\item Has the entity $s_i$ been blocked the application of  reaction $r_j$?
+\item Has the entity $s_i$ been produced by reaction $r_j$?
+\item Has the entity $s_i$ been produced by some reaction?
 \item Has the entity $s_i$ been provided by the context? 
-\item Has the rule $r_j$ not been applied? 
+\item Has the reaction $r_j$ not been applied? 
 %\item Has the entity $s$ been used by some rule as reagent ?
@@
 
-As detailed before, in the following we assume that: (i) the context can be non-deterministic, otherwise it makes little sense to rely on bisimulation to observe the branching structure of system dynamics; (ii) we are interested in observing the names of the entities involved in the transitions and also the rules that have been applied, thus we assume top level restrictions are absent and rely on solid transitions only (with leftmost and rightmost silent actions).
+As detailed before, in the following we assume that: (i) the context can be non-deterministic, otherwise it makes little sense to rely on bisimulation to observe the branching structure of system dynamics; (ii) we are interested in observing the names of the entities involved in the transitions and also the reactions that have been applied, thus we assume top level restrictions are absent and rely on solid transitions only (with leftmost and rightmost silent actions).
 
@@
 
-Roughly, an atomic assertion $\zeta$ denotes either a string composed by a single name $\eta$ (one of the symbols in the set $\mathit{names}$ for denoting a particular entity, rule, production or context), or
+Roughly, an atomic assertion $\zeta$ denotes either a string composed by a single name $\eta$ (one of the symbols in the set $\mathit{names}$ for denoting a particular entity, reaction, production or context), or
 the wildcard $?$ that stands for any symbol, or the pattern $[N]$ that stands for any string composed by a single symbol in the set $N$. Clearly $?$  is just a shorthand for $[\mathit{names}]$.
@@
 Please recall that we have defined the behaviour of the context in a non deterministic way, thus 
-at each step, different possible sets of entities can be provided to the system and different sets of reaction can be enabled/disabled. 
-Bio-simulation can thus be used to compare the behaviour of different systems that share some of the reactions or entities or also to compare the behaviour of the same set of reaction rules when different contexts are provided.
+at each step, different possible sets of entities can be provided to the system and different sets of reactions can be enabled/disabled. 
+Bio-simulation can thus be used to compare the behaviour of different systems that share some of the reactions or entities or also to compare the behaviour of the same set of reactions when different contexts are provided.
 

### 2020-07-27 TCS_v2/07-discussion.tex
pulizia e riorganizzazione

--- 2020-06-29 TCS_version/07-discussion.tex
@@
 an entity, in a way that the mutated version of the entity can take 
-part to only a subset of the rules requiring  the  \emph{normal version} of the entity.
+part to only a subset of the reactions requiring  the  \emph{normal version} of the entity.
 For example, let us assume that entity $s1$ is consumed by reactions $a1$ and $a2$.

### 2020-07-27 Catuscia - final/01-introduction.tex
da dropbox

--- 2020-07-27 TCS_v1/01-introduction.tex
@@
 aspects: human designed computing inspired by nature, and computation 
-performed in nature. Reaction Systems (RSs)~\cite{BEMR11} are 
+performed in nature. Reaction Systems (RSs)~\cite{BEMR11} are a 
 % model of
 % formal computation 
-a rewriting formalism
+rewriting formalism
 %model 
@@
 fields, such as computer science~\cite{MPR15}, 
-biology~\cite{ABP14,CMMBM12,Az17,BarbutiGLM16}, 
+biology~\cite{ABP14,CMMBM12,Az17,BarbutiGLM16}, and
 molecular chemistry~\cite{OY16}.
@@
 In this paper, we present an encoding from RSs, 
-to the open multiparty process algebra c\CNA,\footnote{After `chained Core Network 
-Algebra'.} a variant of the {\tt link}-calculus~\cite{BodeiBB12,BBB17} without name mobility.
+to the open multiparty process algebra c\CNA\footnote{After `chained Core Network 
+Algebra'.}, a variant of the {\tt link}-calculus~\cite{BodeiBB12,BBB17} without name mobility.
 This formalism allows several processes to synchronise and 
-communicate altogether, at the same time, with a new interaction
+communicate altogether, at the same time, with a new communicating
 mechanism based on links and link chains. 
-The initial motivation for introducing this mechanism was to encode
-Mobile Ambients~\cite{CardelliG00}, getting a much stronger operational
-correspondence than any available in the literature, such as the one in~\cite{B16}.
-Later it was shown that the {\tt link}-calculus allowed one to easily encode calculi for biology equipped with membranes, as in~\cite{BodeiBBC14}.
-
-We illustrate our embedding by means of some examples coming
-both from the computer science and the biological field.
-%and then we consider a more complex example, by modeling a RS 
-%representing a regulatory network for \emph{lac} operon, presented in~\cite{CMMBM12}. 
-We also show that our embedding preserves the main
-features of RSs, and prove its correctness and completeness
-from the operational semantics viewpoint.
-
-Then, we present a methodology for verifying formally
-properties of Reaction Systems. The classical notion of bisimulation for 
-process algebras allows to consider two processes as equivalent 
-when one process can simulate all the actions executed 
-by the other one and vice versa. In this paper we define a new
-notion of bisimilarity which takes into account the characteristics
-of biological systems. We define a new {\em bio-simulation} having in mind
-the possibility that two interacting systems may be compared w.r.t.
-a subset of the possible biological actions. This is useful for concentrating 
-on the sub-model that one may need to consider for a specific study or application,
-without getting lost in the complexity of the full biological system, or network.
-The notion of bio-simulation relies on a new and simple assertion language,
-which allows to focus on some properties of the Reaction Systems to be verified.
-In fact, bio-similarity is parametric on a given assertion of interest.
-Then we prove that the well-known logical characterisation of bisimilarity in terms of Hennessy-Milner Logic~\cite{HM80} (HML)
-can be extended to the case of bio-similarity by tailoring HML formulas to
-the same assertion of interest used in bio-similarity.  As HML is a powerful
-formalism for specifying properties of labeled transition systems, 
-its variant introduced here looks a suitable option to specify formulas 
-over complex labels by abstracting from unnecessary details.
-
+Our initial motivation for introducing this mechanism was to encode
+Mobile Ambients~\cite{CardelliG00}, obtaining a much stronger operational
+correspondence than any one available in the literature, such as the one in~\cite{B16}.
+ This allowed us to easily encode calculi for biology equipped with membranes, as in~\cite{BodeiBBC14}.
+Process calculi have been used successfully to model
+biological processes, see~\cite{BBDFH18} for a recent survey.
+We illustrate our embedding by means of some simple basic examples,
+and then we consider a more complex example, by modeling a RS 
+representing a regulatory network for \emph{lac} operon, presented in~\cite{CMMBM12}. We also show that our embedding preserves the main
+features of RSs, and prove its correctness and completeness.
 Our main contributions are as follows:
@@
 can be
-specified in a recursive way, as ordinary processes;
-\item along the same lines, non-deterministic contexts can be seamlessly expressed in our setting; 
-\item when deterministic contexts are considered, the c\CNA \
-computation is deterministic and mimics the evolution of the corresponding RS;
-\item we define an assertion language to specify local properties of 
-Reaction Systems and exploit it to define a novel notion of behavioural equivalence for Reaction Systems, called bio-similarity;
-\item we show that our assertions can be used to extend the Hennessy-Milner logic to our framework, in such a way that logical equivalence of processes coincide with bio-similarity;
-\item we sketch how one can express the behaviour of entity mutation,
+specified in a recursive way as an ordinary process;
+\item 
+%\begin{color}{red}
+our translation results in  a  c\CNA \ system where from
+each state only one transition can be generated, thus the c\CNA \
+computation is fully deterministic as that of the encoded RS;
+%\end{color}
+\item we can express the behaviour of entity mutation,
 %change the behaviour of  each single entity $s$  in a
@@
 $s'$ can  take part to only a subset of rules requiring entity $s$;
-\item we show that with a little coding effort, our encoding allows two RSs to
-communicate; i.e. we can model scenarios where a subset of those entities that the context  can
-provide, are instead provided by a second RS.
+\item with a little coding effort, two RSs can
+communicate; i.e.  a subset of those entities that the context  can
+provide, are then provided by a second RS.
 \end{itemize}
@@
 an implementation of the {\tt link}-calculus.
-The examples that we propose also show that some optimisations are possible
-to reduce the coding when suitable assumptions are made about the provision of entities.
 
@@
 
-\paragraph{Related work}
-
-Process calculi have been used successfully to model
-biological processes, see~\cite{BBDFH18} for a recent survey.
-We are not aware of any SOS operational semantics for the RS; it seems not the case
-that a deterministic behaviour, as the one of the RS, once the initial state is 
-fixed, could be defined by inference rules that classically are applied 
-to define non-deterministic transition systems.
-The reversible computation paradigm for RS extends the RS framework  by
-allowing  backward computations as well as forward computations; for this 
-goal a set of inference rules for rewriting logic has been 
-defined in~\cite{10.1007/978-3-319-73359-3_3}
-to exactly keep trace of those elements that  dissolve in the next 
-computation step.
-
-A preliminary version of this paper appeared in~\cite{BBF19}.
-There are several major differences w.r.t. the paper~\cite{BBF19}, which is here extended as follows:
-\begin{itemize}
-\item we present several new examples to illustrate our framework,
-and give more detailed explanations about its use;
-\item we introduce an assertion language to specify the properties of Reaction Systems;
-\item we define the notion of bio-simulation to relate Reaction Systems and compare their behaviours;
-\item we show that our assertion language allows to specify properties extending to our
-framework the Hennessy-Milner logic;
-\item we include here all proofs of main results.
-\end{itemize}
-
-\paragraph{Structure of the paper}
-
-Section \ref{sec:rs} 
+\paragraph{Structure of the paper.} Section \ref{sec:rs} 
 describes RSs and their semantics (interactive processes).
@@
 it.
-Section~\ref{ex:examples} presents a couple of examples with computer science and biological applications.
-%Section \ref{ex:lactose} shows a more complex example taken from
-%the literature on RSs and illustrating a \emph{lac} operon.
-Section~\ref{sec:biosimulation} presents a methodology for the formal verification of
-properties of Reaction Systems that are expressed in a novel assertion language.
+Section \ref{ex:lactose} shows a more complex example taken from
+the literature on RSs and illustrates a \emph{lac} operon.
 Section~\ref{sec:discussion} presents some features and 
@@
 Finally, Section~\ref{sec:conclusion}
-discusses future work, and concludes.
+discusses future work and concludes.

### 2020-07-27 Catuscia - final/02-ReactionSystem.tex
da dropbox

--- 2020-03-30 02-ReactionSystem.tex
@@
 not only for modeling biological phenomena, but also for
-the contributions which is giving to computer science~\cite{MPR15}, 
+the contributions given to computer science~\cite{MPR15}, 
 theory of computing, 
@@
 Sometimes artificial inhibitors are used that are never produced by any reaction.
-For the sake of simplicity, in some examples, we will allow $I$ to be empty.
+For the sake of simplicity,  and without loss of generality, in some examples, we will allow $I$ to be empty.
 %Note that since $R_a$ and $I_a$ are both nonempty, we always 
@@
 The set $S$ is called the \emph{background set} of ${\cal A}$, its elements are called \emph{entities},
-they represent molecular substances (e.g., atoms, ions, molecules) that may be present
+and they represent molecular substances (e.g., atoms, ions, molecules) that may be present
 in the states of a biochemical system.
 % modeled by $A$. 
-The set $A$ is the set of \emph{reactions} of $\mathcal{A}$.
+The set $A$ is the set of \emph{reactions} of ${\cal A}$.
 Since $S$ is finite, so is $A$: 
@@
 
+%\begin{color}{red}
 \begin{definition}[Reaction Result]
-Let $W$ be a finite subset of $S$.
+Given a finite set of entities $S$, let $T\subseteq S$.
 \begin{enumerate}
-\item Let $a$ be a reaction in $S$. Then $a$ is enabled by $W$, denoted by $\mathit{en}_a(W)$, 
-if $R_a \subseteq W$
-and $I_a \cap W = \emptyset$. The result of $a$ on $W$, denoted by $\mathit{res}_a(W)$, 
+\item Let $a\in rac(S)$ be a reaction. Then $a$ is enabled by $T$, denoted by $en_a(T)$, 
+if $R_a \subseteq T$
+and $I_a \cap T = \emptyset$. 
+\item Let $a\in rac(S)$ be a reaction. The result of $a$ on $T$, denoted by $res_a(T)$, 
 is defined by:
-$\mathit{res}_a(W) =P_a$, if $\mathit{en}_a(W)$, and $\mathit{res}_a(W) = \emptyset$ otherwise.
-\item Let $A$ be a finite set of reactions. The result of $A$ on $W$, denoted 
-by $\mathit{res}_A(W)$,  is defined
-by: $\mathit{res}_A(W) = \bigcup_{a \in A} \mathit{res}_a(W)$.
+$res_a(T) =P_a$ if $en_a(T)$, and $res_a(T) = \emptyset$ otherwise.
+\item Let $A\subseteq rac(S)$ be a finite set of reactions. The result of $A$ on $T$, denoted 
+by $res_A(T)$,  is defined
+by: $res_A(T) = \bigcup_{a \in A} res_a(T)$.
 \end{enumerate}
 \end{definition}
+%\end{color}
 
@@
 \begin{itemize}
-\item {\bf No permanency.} An entity of a set $W$ vanishes unless it is sustained 
+\item {\bf No permanency.} An entity of a set $T$ vanishes unless it is sustained 
 by a reaction. This reflects the fact that a living cell would die for lack 
@@
 %$C_0,\dots, C_n, D_0,\dots, D_n  \subseteq S$, $D_0 = \emptyset$, and 
-$D_i =  \mathit{res}_{A}(D_{i-1} \cup C_{i-1})$ for any $i \in [1,n]$.
+$D_i =  res_{\cal A}(D_{i-1} \cup C_{i-1})$ for any $i \in [1,n]$.
 % $i \in \{1,\dots, n\}$.
@@
 is the union of two sets: the context $C_{i}$
-at step $i$ and the result $D_i=\mathit{res}_{A}(W_{i-1})$ from the previous step.
+at step $i$ and the result of the previous step.
 
@@
 
-For technical reasons, we extend the notion of an interactive process to deal with 
+For technical reasons, we extend in a straightforward way the notion of an interactive process to deal with 
 infinite sequences.
@@
 Let ${\cal A}=(S,A)$ be a RS, and let $\pi =(\gamma,\delta)$  be an  $n$-step interactive process,
-with $ \gamma=\{C_i\}_{i\in[0,n]}$ and  $\delta=\{D_i\}_{i\in[0,n]}$
+with $ \gamma=\{C_i\}_{i\in[0,n]}$ and  $\delta=\{D_i\}_{i\in[0,n]}$.
 %, $\delta= D_0, \dots D_n$.
-Then, we let $\pi^\infty= (\gamma^\infty,\delta^\infty)$ be the extended interactive process of  $\pi$,
-defined as $  \gamma^\infty=\{C_i'\}_{i\in\mathbb{N}}$, $\delta^\infty=\{D_i'\}_{i\in\mathbb{N}}$, where:
-$$
-C'_j = \left\{
-\begin{array}{ll}
-C_j & \mbox{if $j\in[0,n]$}\\
-\emptyset & \mbox{otherwise ($j>n$)}
-\end{array}
-\right.
-\qquad
-D'_j = \left\{
-\begin{array}{ll}
-D_0 & \mbox{if $j=0$}\\
-\mathit{res}_A(D_{j-1}' \cup C_{j-1}') & \mbox{otherwise ($j\geq 1$)}
-\end{array}
-\right.
-$$
-% $C_j' = C_j$ for $j\in[0,n]$ and $C_j' =\emptyset$ for $j>n$,   $D_0' = D_0$ and  $D_j' = \mathit{res}_A(D_{j-1}' \cup C_{j-1}')$ for $j\geq 1$.
+Then, let $\pi'= (\gamma',\delta')$ be the extended interactive process of  $\pi =(\gamma,\delta)$,
+defined as $  \gamma'=\{C_i'\}_{i\in\mathbb{N}}$, $\delta'=\{D_i'\}_{i\in\mathbb{N}}$, where
+ $C_j' = C_j$ for $j\in[0,n]$ and $C_j' =\emptyset$ for $j>n$,   $D_0' = D_0$ and  $D_j' = res_A(D_{j-1}' \cup C_{j-1}')$ for $j\geq 1$.
 \end{definition}
+
+%\begin{color}{red}
+The next example shows that the functioning of the interacting processes is deterministic, once the context is fixed.
+
+\begin{example}
+Let  ${\cal A}=(S,A)$ be a RS with  $S=\{s_1,s_2,s_3,s_4\}$, and  $A = \{ a_1,a_2\}$, where $a_1 = (\{s_1\},\{s_3\},\{s_1\})$,  $a_2 = (\{s_2\},\{s_4\},\{s_2\})$. 
+ %It is easy to see that $r_1$ is incompatible with $r_2$ and with $r_3$, as $r_1$ requires the presence of $s_1$, instead $r_2$ and $r_3$ requires the absence of $s_1$. \\
+% Now, reaction $r_1$ is only applied  when the entity configuration is $\{s_1,s_2\}$, or $\{s_1\}$.
+ %If the configuration is $\{s_3,s_2\}$, then both $r_2$ and $r_3$ are applied.
+ %If the configuration is $\{s_2\}$, only $r_3$ is applied. The configurations $\{s_1,s_3\}$,and $\{s_3\}$ make no reaction applicable.\\
+ Now, we show the first three steps of an interacting process $\pi=(\gamma,\delta)$ where the context provides entities as follows: $\gamma = \{C_0 = \{s_1,s_2\}, C_1 =\{s_3\},C_2=\{s_1,s_4\},C_3= \emptyset, C_4,\dots, C_n\}$.
+Initally $D_0 = \emptyset$ and from
+ $C_0\cup D_0=\{s_1,s_2\}$ we get to $C_1\cup D_1=\{s_1,s_2,s_3\}$ by applying  $a_1$, $a_2$ (they both are enabled); after that  we get to $C_2\cup D_2=\{s_1,s_2,s_4\}$, by applying $a_2$ only ($a_1$ is not enabled); finally we get to $C_3\cup D_3=\{s_1\}$ by applying $a_1$ ($a_2$ is not enabled).
+Thus $\delta=\{D_0 = \emptyset, D_1=\{s_1,s_2\},D_2=\{s_2\},D_3=\{s_1\},D_4,\dots,D_n\}$.
+We remark that at every state $C_i \cup D_i$ all the reactions which are enabled are applied, so the computation is deterministic.
+\end{example}
+%\end{color}

### 2020-07-27 Catuscia - final/04-CCNA.tex
da dropbox

--- 2020-04-02 04-CCNA.tex
@@
 In this section we introduce  the syntax and operational semantics
-of a variant of the {\tt link}-calculus~\cite{BodeiBB12}, the c\CNA\  (chained \CNA)
+of a variant of the {\tt link}-calculus~\cite{BodeiBB12}, the c\CNA\  (chained \CNA),
 where the prefixes are link chains.
@@
 
-\paragraph{Link Chains}
+\paragraph{Link Chains.}
 Let $\mathcal{C}$ be the set of channels, ranged over by $a,b,...$, and 
-let $\mathcal{A} = \mathcal{C} \cup \setof{\silent} \cup \setof{{\noacts}}$ be the set of actions, ranged over by $\alpha,\beta,...$,
+let $\mathcal{N} = \mathcal{C} \cup \setof{\silent} \cup \setof{{\noacts}}$ be the set of actions, 
+ranged over by $\alpha,\beta,...$,
 where the symbol $\silent$ denotes a \emph{silent} action, while the symbol $\noact$ denotes a \emph{virtual} (non-specified) action.
@@
 it is \emph{solid} if $\alpha,\beta \neq \ \noacts$; 
-the link $\link{\noact}{\noact}$ is called \emph{virtual}.
+the link $\link{\noact}{\noact}$ is called \emph{virtual};
+%\begin{color}{red}
+the link $\link{\silent}{\silent}$ is called \emph{silent}.
+%\end{color}
 A link is \emph{valid} if it is solid or virtual.
-We let $\mathcal{L}$ be the set of valid links.
+%We let $\mathcal{L}$ be the set of valid links.
 %
 A \emph{link chain} is a 
-finite sequence $v = \ell_{1}...\ell_{n}$ of (valid) links  $\ell_{i} = \link{\alpha_{i}}{\beta_{i}}$ such that:
+finite sequence $v = \ell_{1}...\ell_{n}$ of valid links  $\ell_{i} = \link{\alpha_{i}}{\beta_{i}}$ such that:
 \begin{enumerate}
@@
 
-Virtual links represent missing elements of a chain. 
-The equivalence $\blackstretcheq$ models expansion/contraction of virtual links to adjust the length of a link chain.
+%\begin{color}{red}
+A link chain whose links are silent is also called silent.
+%\end{color}
+Virtual links $\link{\noact}{\noact}$ represent missing elements of a chain. 
+The equivalence $\blackstretcheq$ models expansion and contraction of virtual links to adjust the length of a link chain.
 
@@
 %
-For example, given $v_1 = \startchain{\silent}\chainedlink{a}{\noact}\chainedlink{\noact}{\noact}\chainend{\noact}$, $v_2 = \startchain{\noact}\chainedlink{\noact}{a}\chainedlink{b}{\noact}\chainend{\noact}$ and $v = \startchain{\silent}\chainedlink{a}{a}\chainedlink{b}{\noact}\chainend{\noact}$
-we have $\merges{v_1}{v_2} = v$, whereas $\merges{v_1}{v}$ is not defined. Notably the merge operation is commutative and associative.
+For example, given $v_1 = \startchain{a}\chainedlink{b}{\noact}\chainedlink{\noact}{\noact}\chainend{\noact}$ and $v_2 = \startchain{\noact}\chainedlink{\noact}{b}\chainedlink{c}{\noact}\chainend{\noact}$ we have $\merges{v_1}{v_2} = \startchain{a}\chainedlink{b}{b}\chainedlink{c}{\noact}\chainend{\noact}$.
 
@@
 %
-For example, given $v = \startchain{\silent}\chainedlink{a}{a}\chainedlink{b}{\noact}\chainend{\noact}$ as above, we have 
-$\restrict{a}v = \startchain{\silent}\chainedlink{\silent}{\silent}\chainedlink{b}{\noact}\chainend{\noact}$,
-whereas $\restrict{b}v$ is not defined.
+For example, given $v = \startchain{a}\chainedlink{b}{b}\chainedlink{c}{\noact}\chainend{\noact}$ we have 
+$\restrict{b}v = \startchain{a}\chainedlink{\silent}{\silent}\chainedlink{c}{\noact}\chainend{\noact}$.
 
-\paragraph{Syntax}
+\paragraph{Syntax.}
 %\begin{definition}
-The set of c\CNA\ processes, denoted as $\cal{P}$ and ranged over by $P,Q$, is defined by the following grammar:
+The c\CNA\ processes are generated by the following grammar:
 \[
@@
 
+%\begin{color}{red}
 The syntax of c\CNA\  extends that of \CNA~\cite{BBB17} by allowing to use link chains as prefixes instead of links.
-For the rest it features nondeterministic (guarded) choice, parallel composition, restriction, relabelling and possibly recursive definitions.
+This extension was already discussed in~\cite{BBB17} and it preserves all the main formal properties of \CNA. 
+For the rest it features nondeterministic choice, parallel composition, restriction, relabelling and possibly recursive definitions of the form $A\defeq P$ for some constant $A$.
 Here we do not consider name mobility, which is present instead in the {\tt link}-calculus.
+%\end{color}
 
-\paragraph{Semantics}
+\paragraph{Semantics.}
 
+%\begin{color}{red}
 The operational semantics of c\CNA\ is defined in the SOS style by the inference rules in Fig.\ref{fig:cnasos}.
-The rules are reminiscent of those for Milner's CCS and they essentially coincide with those of \CNA\ in~\cite{BBB17}.
-The only difference is due to the presence of prefixes that are link chains.
-Briefly: rule (\textit{Sum}) selects one alternative and puts as label a possible contraction/expansion of the link chain in the selected prefix; rule (\textit{Ide}) selects one transition of the defining process for a constant; rule (\textit{Res}) restricts some names in the label (it cannot be applied when $\restrict{a}v$ is not defined); rules (\textit{Lpar}) and (\textit{Rpar}) account for interleaving in parallel composition; rule (\textit{Com})
+The rules are reminiscent of those for Milner's CCS and they essentially coincide with those of \CNA\ in~\cite{BBB17}, except for the presence of prefixes that are link chains instead of single links.
+Briefly: rule (\textit{Sum}) selects one alternative and uses, as a label, a possible contraction/expansion $v$ of the link chain $v_j$ in the selected prefix; rule (\textit{Ide}) selects one transition of the process defined by a constant; rule (\textit{Rel}) renames the channels in the label as indicated by $\phi$; rule (\textit{Res}) restricts some names in the label (it cannot be applied when $\restrict{a}v$ is not defined); rules (\textit{Lpar}) and (\textit{Rpar}) account for interleaving in parallel composition; rule (\textit{Com})
 synchronises interactions (it cannot be applied when $\merges{v}{v'}$ is not defined).
-
-\begin{example}
-As some simple examples, consider the recursive process definitions $H \defeq \link{a}{b}.\link{a}{c}.H$ and $K \defeq \link{a}{b}.K + \link{a}{c}.K$: the former is some sort of forwarder that alternates between providing a link from $a$ to $b$ and to $c$; the latter is a nondeterministic forwarder that at each step can link $a$ to $b$ or to $c$.
-\end{example}
-
-Analogously to \CNA, the operational semantics of c\CNA\ satisfies the so called Accordion Lemma: whenever $P \xrightarrow{v} P'$ and $v'\blackstretcheq v$ then $P \xrightarrow{v'} P'$.
+%
+Analogously to \CNA, the operational semantics of c\CNA\ satisfies the so called Accordion Lemma: whenever $P \xrightarrow{v} Q$ and $v'\blackstretcheq v$ then $P \xrightarrow{v'} Q$.
+%
+As a matter of notation, we write $P \rightarrow Q$ when $P \xrightarrow{v} Q$ for some silent link chain $v$ and call it a \emph{silent transition}.
+Similarly, a sequence of $j$ silent transitions is denoted $P \rightarrow^j Q$.
+%\end{color}
 
@@
 \begin{prooftree} 
-\AxiomC{$\upsilon \blackstretcheq \upsilon_j\quad j\in I$} %\mbox{($\ell$ {\em only} solid link in $s$)}
+\AxiomC{$v \blackstretcheq v_j\quad j\in I$} %\mbox{($\ell$ {\em only} solid link in $s$)}
 \RightLabel{\scriptsize(\textit{Sum})}
-\UnaryInfC{$\sum_{i\in I} \upsilon_i.P_i \xrightarrow{\upsilon} P_j$} 
+\UnaryInfC{$\sum_{i\in I} \upsilon_i.P_i \xrightarrow{v} P_j$} 
 \DisplayProof
 \
-\AxiomC{$P \xrightarrow{\upsilon} P'$}
+\
+\AxiomC{$P \xrightarrow{v} P'$}
 \AxiomC{$(A \defeq P)\in\Delta$}
 \RightLabel{\scriptsize(\textit{Ide})}
-\BinaryInfC{$A \xrightarrow{\upsilon} P'$} 
+\BinaryInfC{$A \xrightarrow{v} P'$} 
+\end{prooftree} 
+\end{center}
+
+\begin{center} 
+\begin{prooftree} 
+\AxiomC{$P \xrightarrow{v} P'$} 
+\RightLabel{\scriptsize(\textit{Rel})}
+\UnaryInfC{$P[\phi] \xrightarrow{\phi(v)} P'[\phi]$} 
+\DisplayProof
+\
+\
+\AxiomC{$P \xrightarrow{v} P'$} 
+\RightLabel{\scriptsize(\textit{Res})}
+\UnaryInfC{$\restrict{a}P \xrightarrow{\restrict{a}v} \restrict{a}P'$} 
 %\AxiomC{$P \xrightarrow{v} P'$} 
@@
 \begin{prooftree} 
-\AxiomC{$P \xrightarrow{\upsilon} P'$} 
-\RightLabel{\scriptsize(\textit{Res})}
-\UnaryInfC{$\restrict{a}P \xrightarrow{\restrict{a}\upsilon} \restrict{a}P'$} 
-\DisplayProof
 %\,
@@
 \RightLabel{\scriptsize(\textit{Lpar})}
-\UnaryInfC{$P|Q \xrightarrow{\upsilon} P'|Q$} 
+\UnaryInfC{$P|Q \xrightarrow{v} P'|Q$} 
 \DisplayProof
-%\,
+\
+\
 %\AxiomC{$Q \xrightarrow{s} Q'$} 
@@
 %\begin{prooftree} 
-\AxiomC{$P \xrightarrow{\upsilon'} P'$} 
-\AxiomC{$Q \xrightarrow{\upsilon} Q'$} 
+\AxiomC{$P \xrightarrow{v'} P'$} 
+\AxiomC{$Q \xrightarrow{v} Q'$} 
 \RightLabel{\scriptsize(\textit{Com})}
-\BinaryInfC{$P|Q \xrightarrow{\merges{\upsilon}{\upsilon'}} P'|Q'$} 
+\BinaryInfC{$P|Q \xrightarrow{\merges{v}{v'}} P'|Q'$} 
 %\DisplayProof
@@
 the presentation of our translation.
-
+%First, we define the juxtaposition of link or link chains as
 \begin{definition}[Replication]
-Let $\upsilon$ be a link chain such that $\upsilon\upsilon$ is also a valid link chain. Its $n$ times replication $\upsilon^n$ is defined recursively by letting
-$\upsilon^0 = \epsilon$ (i.e. the empty chain) and $\upsilon^n = \upsilon\upsilon^{n-1}$.
+Let $\upsilon$ be a link chain. Its $n$ times replication $\upsilon^n$ is defined recursively by letting
+$\upsilon^0 = \epsilon$ (i.e. the empty chain) and $\upsilon^n = \upsilon^{n-1}\upsilon$,
 %\[
@@
 %\]
-%with the hypothesis that all the links in the resulting link chains match.
+with the hypothesis that all the links in the resulting link chains match.
 %, and that $\upsilon^0 = \epsilon$ is the empty chain.
@@
 For example,  the expression $(\startchain{a}\chainedlink{b}{\noact}\chainend{\noact})^3$  denotes the chain $\startchain{a}\chainedlink{b}{\noact}\chainedlink{\noact}{a}\chainedlink{b}{\noact}\chainedlink{\noact}{a}\chainedlink{b}{\noact}\chainend{\noact}$.
-Instead the expression $(\link{a}{b})^2$ is ill-defined because $a$ does not match with $b$.
 
-Then, we introduce the notation for \emph{half links} that will be used in conjunction with the \emph{open block of chain} to form regular link chains.
+ \noindent
+We introduce the \emph{half link} that will be used in conjunction with the \emph{open block of chain} to form regular link chains.
+\noindent
+Let  $\link{a}{}$ denote the \emph{half left link} of a  link $\link{a}{x}$, and conversely
+let  $\link{}{a}$ denote the \emph{half right link} of  $\link{x}{a}$.
 
-\begin{definition}[Half links]
-Let $a$ be a channel name, we define the \emph{half left link}  $\link{a}{}$,   and the  \emph{half right link} $\link{}{a}$.
-\end{definition}
+%\begin{definition}[Half links]
+%Let $a$ be a channel name, we define the \emph{half left link}:  $\link{a}{}$,   and the  \emph{half right link}: $\link{}{a}$.
+%\end{definition}
 
-\begin{definition}[Open block]
-Let $R$ be a set of names. We 
-define an \emph{open block} as $\blockchain{a_i}{a_o}{a \in R}$\ , where $a_i$ and $a_o$ are annotated version of the name $a$, as
+%\begin{color}{red}
+\begin{definition}[Open block]\label{def:open}
+Let $R$ be a totally ordered, finite set of names. We 
+define an \emph{open block} as $\blockchain{c_i}{c_o}{c \in R}$\ , where $c_i$ and $c_o$ are annotated version of the name $c$, as follows
+
 \[
-\begin{array}{l@{\quad}l@{\quad}l}
-\mbox{set}& \mbox{block of chain}& \mbox{result}\\
-R= \emptyset & \blockchain{a_i}{a_o}{a \in R}   &\epsilon\\
-R = \{ b\}    &\blockchain{a_i}{a_o}{a \in R}  & {}_{b_i}^{\noact}\chainedlink{\noact}{b_o}\\
-R = \{b\} \cup R' & \blockchain{a_i}{a_o}{a \in R} & {}_{b_i}^{\noact}\chainedlink{\noact}{b_o}\link{}{}\blockchain{a_i}{a_o}{a \in R'}
+\begin{array}{l@{\quad}c@{\quad}l}
+\mbox{set}& \mbox{open block expression}& \mbox{result}\\
+R= \emptyset & \blockchain{c_i}{c_o}{c \in R}   &\epsilon\\
+R = \{ a\}    &\blockchain{c_i}{c_o}{c \in R}  & {}_{a_i}^{\noact}\chainedlink{\noact}{a_o}\\
+R = \{a\} \uplus R' \mbox{ with } a=\min R& \blockchain{c_i}{c_o}{c \in R} & {}_{a_i}^{\noact}\chainedlink{\noact}{a_o}\link{}{}\blockchain{c_i}{c_o}{c \in R'}
 \end{array}
@@
 \end{definition}
+%\end{color}
 %Please note that, channel names of juxtaposed open blocks of chain are not required to match.
@@
  For example, for $R = \{a,b\}$ the expression $ \blockchain{c_i}{c_o}{c \in R} $\;
-denotes the block of chains ${}_{a_i}^{\noact}\chainedlink{\noact}{a_o} \backslash{}_{b_i}^{\noact}\chainedlink{\noact}{b_o}$; and the expression $\link{r_1}{} \blockchain{c_i}{c_o}{c \in R}\;\link{}{r_2} $ denotes the chain \\
-$ \startchain{r_1}\chainedlink{a_i}{\noact}\chainedlink{\noact}{a_o}\chainedlink{b_i}{\noact}\chainedlink{\noact}{b_o}\chainend{r_2}$.
+denotes the block of chains ${}_{a_i}^{\noact}\chainedlink{\noact}{a_o} \backslash{}_{b_i}^{\noact}\chainedlink{\noact}{b_o}$; and the expression $\link{r_1}{} \blockchain{c_i}{c_o}{c \in R}\;\link{}{r_2} $ denotes the chain $ \startchain{r_1}\chainedlink{a_i}{\noact}\chainedlink{\noact}{a_o}\chainedlink{b_i}{\noact}\chainedlink{\noact}{b_o}\chainend{r_2}$.
+
+%\noindent
+%{\color{red} In Definition~\ref{def:open}, we do not care about a specific order for the set $R$, thus we assume to have fixed a specific one.}

### 2020-07-27 Catuscia - final/05-translation.tex
da dropbox

% !TEX root = ./main_RS2L_LNCS.tex

\section{From Reaction Systems to c\CNA}
\label{sec:trans}
Here we present a translation from Reaction Systems to c\CNA.
The idea is to define separated processes for representing the behaviour of each entity, each reaction, 
and for the provisioning of each entity by the context.

\paragraph{Processes for entities.}
Given an entity $s \in S$, we exploit four different names for the interactions over $s$:
names $s_i$, $s_o$ are used to test the presence of $s$ in the system; names $\hat{s}_i$, $\hat{s}_o$ are used to test the provisioning of $s$ from the context;
names $\tilde{s}_i$, $\tilde{s}_o$  are used to test the production of $s$ by some reaction; names $\overline{s}_i$, $\overline{s}_o$ are used to test the absence of $s$  in the system; and names $\underline{s}_i$, $\underline{s}_o$ are used to test the absence of $s$ from the context.
We let $P_s$ be the process implementing the  presence of $s$ in the system, and $\overline{P_s}$ its absence.
%\begin{color}{red}
They are defined below:

%\[
%\begin{array}{lcl@{\hskip 1cm}lcl}
% P_s &\defeq & P(s,\tilde{s},\hat{s},\underline{s}) 
% &
% \overline{P_s}& \defeq & P(\overline{s},\tilde{s},\hat{s},\underline{s})
%\end{array}
%\]
%{\color{red}
%\[
%\begin{array}{rcl}
% P(s,\tilde{s},\hat{s},\underline{s}) & \defeq &\sum_{h_j,k_j\in\{0,1\}}
%  (\startchain{s_i}\chainedlink{s_o}{\noact}\chainend{\noact})^{h_j} \ (\startchain{\tilde{s}_i}\chainedlink{\tilde{s}_o}{\noact}\chainend{\noact})^{k_j}\ \link{\hat{s}_i}{\hat{s}_o}.P_s\\
%&&\ + \\
%&&
%\sum_{\scriptsize \begin{array}{l}h_j,k_j, \\ h'_j\in\{0,1\}\end{array}}
%  (\startchain{s_i}\chainedlink{s_o}{\noact}\chainend{\noact})^{h_j} \ (\startchain{\tilde{s}_i}\chainedlink{\tilde{s}_o}{\noact}\chainend{\noact})^{k_j} \  \startchain{\tilde{s}_i}\chainedlink{\tilde{s}_o}{\noact}\chainend{\noact} \    (\startchain{s_i}\chainedlink{s_o}{\noact}\chainend{\noact})^{h'_j} \ \link{\underline{s}_i}{\underline{s}_o}.P_s \\
%
%%\sum_{R_1+R_2\geq 0}(\sum_{h,k \in[0,1]} 
%%  (\startchain{s_i}\chainedlink{s_o}{\noact}\chainend{\noact})^h \ (\startchain{\tilde{s}_i}\chainedlink{\tilde{s}_o}{\noact}\chainend{\noact})^k)^{R_1} \startchain{\tilde{s}_i}\chainedlink{\tilde{s}_o}{\noact}\chainend{\noact} (\sum_{h ,k\in[0,1]}
%%  (\startchain{s_i}\chainedlink{s_o}{\noact}\chainend{\noact})^h \ (\startchain{\tilde{s}_i}\chainedlink{\tilde{s}_o}{\noact}\chainend{\noact})^k)^{R_2} \link{\underline{s}_i}{\underline{s}_o}.P_s\\
%&&\ +\\
%&&\sum_{j\geq 0}(
%  \startchain{s_i}\chainedlink{s_o}{\noact}\chainend{\noact})^j\ \ \link{\underline{s}_i}{\underline{s}_o}.\overline{P_s}
%\end{array}
%\]
%}
\[
\begin{array}{rcl}
% P(s,\tilde{s},\hat{s},\underline{s}) 
 P_s & \defeq &\sum_{h\geq 0,k\geq 0}  
  (\startchain{s_i}\chainedlink{s_o}{\noact}\chainend{\noact})^h \ \startchain{\hat{s}_i}\chainedlink{\hat{s}_o}{\noact}\chainend{\noact} (\startchain{\tilde{s}_i}\chainedlink{\tilde{s}_o}{\noact}\chainend{\noact})^k.P_s\\
&&\ + \\
&&
\sum_{h\geq 0,k\geq 1}
 (\startchain{s_i}\chainedlink{s_o}{\noact}\chainend{\noact})^h \ \startchain{\underline{s}_i}\chainedlink{\underline{s}_o}{\noact}\chainend{\noact}(\startchain{\tilde{s}_i}\chainedlink{\tilde{s}_o}{\noact}\chainend{\noact})^k.P_s\\
&&\ +\\
&&\sum_{h\geq 0}(\startchain{s_i}\chainedlink{s_o}{\noact}\chainend{\noact})^h \ \link{\underline{s}_i}{\underline{s}_o}.\overline{P_s}
\\
\\
%P(\overline{s},\tilde{s},\hat{s},\underline{s}) 
\overline{P_s}& \defeq &\sum_{h\geq 0,k\geq 0}  
  (\startchain{\overline{s}_i}\chainedlink{\overline{s}_o}{\noact}\chainend{\noact})^h \ \startchain{\hat{s}_i}\chainedlink{\hat{s}_o}{\noact}\chainend{\noact} (\startchain{\tilde{s}_i}\chainedlink{\tilde{s}_o}{\noact}\chainend{\noact})^k.P_s\\
&&\ + \\
&&
\sum_{h\geq 0,k\geq 1}
 (\startchain{\overline{s}_i}\chainedlink{\overline{s}_o}{\noact}\chainend{\noact})^h \ \startchain{\underline{s}_i}\chainedlink{\underline{s}_o}{\noact}\chainend{\noact}(\startchain{\tilde{s}_i}\chainedlink{\tilde{s}_o}{\noact}\chainend{\noact})^k.P_s\\
&&\ +\\
&&\sum_{h\geq 0}(\startchain{\overline{s}_i}\chainedlink{\overline{s}_o}{\noact}\chainend{\noact})^h \ \link{\underline{s}_i}{\underline{s}_o}.\overline{P_s}
\end{array}
\]
%\end{color}

The first line of $P_s$ accounts for the case where $s$ is tested for presence by $h$ reactions and produced by $k$ reactions, while being provided by the context ($\link{\hat{s}_i}{\hat{s}_o}$).
Thus, $s$ will be present at the next step (the continuation is $P_s$). Here $h$ and $k$ are not known a priori and therefore any
combination is possible. By knowing the number of reactions that test $s$, we can bound the maximum 
values of $h$ and $k$.
The second line accounts for the analogous case where $s$ is not provided by the context ($\link{\underline{s}_i}{\underline{s}_o}$). 
The condition $k\geq 1$ guarantees that $s$ will remain present (the continuation is $P_s$).
%Here taking
%$h = \sum_{j\geq0} h_j +h'_j$, $k = (\sum_{j \geq 0} k_j )  + 1$ guarantees that $s$ will be present in the next step. Please note that 
The third line accounts for the case where $s$ is tested for presence by $h$ reactions, but it is neither produced nor provided by  the context. Therefore, in the next step $s$ will be absent in the system (the continuation is $\overline{P_s}$).
Note that in the case of $\overline{P_s}$ the test for presence of $s$ in the system is just replaced by the test for its absence. As before, $s$ will be present again in the system if $s$ will be produced (rows $1$ and $2$ in  $\overline{P_s}$ code), or if the context will provide it (row $1$ in $\overline{P_s}$ code).
%Process $P_s$ (and $\overline{P_s}$) offers link chain prefixes for all the possible
% permutations of the solid links; nevertheless if the reactions, and their enumeration, is known
% the code of $P_s$ can be drastically simplified.

\paragraph{Processes for reactions.}
We assume that each reaction $a$ is assigned a progressive number $j$.
The process for reaction $a_j=(R_j,I_j,P_j)$ must assert either the possibility to apply the reaction or its impossibility.
The first case happens when all its reactants are present (the link $\link{s_i}{s_o}$ is requested for any $s\in R_j$) and all its inhibitors are absent (the link $\link{\overline{e}_i}{\overline{e}_o}$ is requested for any $e\in I_j$), then the product set is released (the link $\link{\tilde{c}_i}{\tilde{c}_o}$ is requested for any $c\in P_j$).
The next case can happen for two reasons: one of the reactants is absent (the link $\link{\overline{s}_i}{\overline{s}_o}$ is requested for some $s\in R_j$) or one of the inhibitors is present (the link $\link{e_i}{e_o}$ is requested for some $e\in I_j$).
The process is recursive so that reactions can be applied at any step.
\[
\begin{array}{lcl@{\quad}l}
P_{a_j}& \defeq & 
 ^{r_j} \backslash  \blockchain{s_i}{s_o}{s \in R_j} \, \link{}{} \blockchain{\overline{e}_i}{\overline{e}_o}{e \in I_j} \, \chainedlink{r_{j+1}}{\noact}\chainedlink{\noact}{p_j}\link{}{} \blockchain{\tilde{c}_i}{\tilde{c}_o}{c \in P_j}\, \backslash_{p_{j+1}}.P_{a_j} & \mbox{\small  \{$a_j$ is applicable\}}\\
&& + \ \\
&& \sum_{s\in R_j}\startchain{r_j}\chainedlink{\overline{s}_i}{\noact}\chainedlink{\noact}{\overline{s}_o} \chainedlink{r_{j+1}}{\noact}\chainedlink{\noact}{p_{j}}\chainend{p_{j+1}}.P_{a_j}  & \mbox{\small  \{$a_j$ is not applicable\}}\\
&&+\\
&& \sum_{e\in I_j}\startchain{r_j}\chainedlink{e_i}{\noact} \chainedlink{\noact}{e_o} \chainedlink{r_{i+1}}{\noact}\chainedlink{\noact}{p_{j}}\chainend{p_{j+1}}.P_{a_j}  & \mbox{\small \{$a_j$ is not applicable\}}\\
\end{array}
\]

\noindent
%\begin{color}{red}
Formally speaking, the definition of the case when $a_j$ is applicable (row 1) requires $R_j$, $I_j$ and $P_j$ to be totally ordered, because open block expressions are used. It is worth noting that the chosen orders are inessential for exploiting $P_{a_j}$ in the encoding of a RS, but they are needed to disambiguate its definition.
Without loss of generality, we  assume that all the entities are enumerated, likewise reactions, so that $R_j$, $I_j$ and $P_j$ inherit the same order. 
%\end{color}
%
We exploit names $r_j,p_j$ to join the chains provided by the application of all the reactions. 
Channels $r_j$ and $r_{j+1}$ enclose the enabling/disabling condition of reaction $a_j$.
Channels $p_i$ and $p_{j+1}$ enclose the links related to the entities produced by $a_j$.
We will see that all the link chain labels of transitions follow the same schema: first we find all the reactions limited to the reactants and inhibitors (chained using $r_j$ channels), then all the supplies by the contexts (chained using $cxt_j$ channels, to be introduced next), and finally the products for all the reactions (chained using $p_j$ channels). 
In the following there is an example explaining this schema.
%\[
%\begin{array}{lcl@{\qquad}l}
%P_{aj}& \defeq & 
% ^{r_j} \backslash  \blockchain{s_i}{s_o}{s \in R_j} \, \link{}{} \blockchain{\overline{e}_i}{\overline{e}_o}{e \in I_j} \, \link{}{}
% \blockchain{\tilde{c}_i}{\tilde{c}_o}{c \in P_j}\, \backslash_{r_{j+1}}.P_{aj} & \mbox{\small  \{$aj$ is applicable\}}\\
%&& + \ \\
%&& \sum_{s\in R_j}\startchain{r_j}\chainedlink{\overline{s}_i}{\noact}\chainedlink{\noact}{\overline{s}_o} \chainend{r_{j+1}}.P_{aj}  & \mbox{\small  \{$aj$ is not applicable\}}\\
%&&+\\
%&& \sum_{e\in I_j}\startchain{r_j}\chainedlink{e_i}{\noact} \chainedlink{\noact}{e_o} \chainend{r_{i+1}}.P_{aj}  & \mbox{\small \{$aj$ is not applicable\}}\\
%\end{array}
%\]

\paragraph{Processes for contexts.}
For each entity $s \in S$, we introduce another process $Cxt_s$,  participating in each transition and determining whether the entity $s$ is provided by the context or not.
As already said, we assume that entities are enumerated and use the names $cxt_j$ to concatenate
the chains formed by the application of all the contexts. 
For each entity $s$ with number $j$, at step $n >0$  there are two possible behaviours: 
\[
\begin{array}{lcl}
Cxt_{s}^n &\defeq &
\left \{\begin{array}{ll}
 \startchain{cxt_j}\chainedlink{\hat{s}_i}{\noact}\chainedlink{\noact}{\hat{s}_o}\chainend{cxt_{j+1}}. Cxt_{s}^{n+1} & \ \mbox{if the context provides $s$  at the $n$-th step}\\[5pt] 
 \startchain{cxt_j}\chainedlink{\underline{s}_i}{\noact}\chainedlink{\noact}{\underline{s}_o}\chainend{cxt_{j+1}}.Cxt_s^{n+1} & \ \mbox{otherwise}
\end{array}\right.\\
Cxt_s & \defeq & Cxt_s^1\\
\end{array}
\]
%\begin{color}{red}
We only consider $Cxt_s^n$ with $n >0$, as the entities such that $s \in C_0$ (resp.\ $s\not\in C_0$) are modeled by the $P_s$ processes (resp.\ $\overline{P_s}$) in the initial c\CNA  \ process. 
%Differently from RSs, where at each step the context can non deterministically provide entities, here the behaviour of the context is defined a priori.
%Nevertheless, 
The intrinsic modularity of c\CNA \ allows us to run different context behaviours
%change the context behaviour leaving unchanged the rest of the system. Thus  can be test
 for the same system by changing the definitions of $Cxt_s^n$.
%present at step zero  are considered to be present in the initial system (process $P_s$ instead of $\overline{P_s}$).
%\end{color}

\begin{definition}[Translation]
\label{def:trans}
Let ${\cal A} = (S, A)$ be a RS, and let $\pi=(\gamma,\delta)$ be an extended interactive process in ${\cal A}$, with  $\gamma=\{C_i\}_{i\in\mathbb{N}}$. 
We define its c\CNA~translation $\llbracket{\cal A},\gamma \rrbracket$ as follows: 
\begin{center}$\llbracket{\cal A},\gamma \rrbracket = \restrict{reacts,cxts,ents,prods}(\Pi_{s\in C_0}P_{s} | \Pi_{s \notin C_0}\overline{P_s} | \Pi_{a\in A} P_a| \Pi_{s\in S} Cxt_s)$,
\end{center} 
with  $reacts$ be the set of reaction names $r_j $, $cxts$ the set of context names $cxt_j$,  $ents$ the set of decorated entity names  $\{s_i,s_o,\hat{s}_i,\hat{s}_o,\tilde{s}_i,\tilde{s}_o,\overline{s}_i,\overline{s}_o,\underline{s}_i,\underline{s}_o| s \in S\}$, and  $prods$ be the set of names $p_j$ associated to each reaction. 
In the following, we set $names = reacts\, \cup\, cxts\, \cup\, ents\, \cup \, prods$.
%For notational convenience, we fix that $r_1 = \silent$,  $r_{u+1} = cxt_1$ for $u$ the number of reacts, and $cxt_{w+1} = \silent$ for $w$ the number of entities.
For notational convenience, we fix that $r_1 = \silent$,  $r_{u+1} = cxt_1$ for $u$ the number of reacts, and $cxt_{w+1} = p_1$ $p_{u+1}=\silent$ for $w$ the number of entities.
\end{definition}

%can be expressed depending on the actual behaviour of the system we want
%simulate: it could be  recursive, mutually recursive, etc..
%In the general case we can use a counter, $N$, that allows us to specify the behaviour of the context for the first $N$ steps of the system, i.e.  $CXT_s \defeq CXT_s(N)$ and the counter at each step decrements its values, until it arrive to a stable behaviour.

It is important to observe that,   for each transition, our c\CNA \ encoding requires  all the processes $P_a$, with $a \in A$, and  $Cxt_s$ and  $P_s$, with $s \in S$,  be interacting in that transition.
This is due to the fact that  all the channels $r_j$, $p_j$, $cxt_j$, and $s_{i}$, and $s_{o}$ are  restricted.
Each reaction defines a pattern to be satisfied, i.e. each reaction inserts as many virtual links  as  the number of reactants, inhibitors, and products, as required by the corresponding reaction.

 \begin{restatable}[]{lemma}{lemmastruct}
 \label{lem:struct}
%\begin{lemma}
 %\label{lem:struct}
  Let ${\cal A} = (S, A)$ be a RS and let $\pi=(\gamma,\delta)$ be an extended  interactive process in ${\cal A}$. Let $P=\llbracket{\cal A},\gamma\rrbracket$ its c\CNA~translation. If there $exists$ $P'$ such that $t= (P \xrightarrow{\restrict{names}\upsilon} P')$ is a transition of $P$, then
  \begin{enumerate}
  \item for each reaction  $a_j \in A$, the corresponding channels $r_j$ and $p_j$ appear in  $\upsilon$; for each entity $s_h \in S$ (where $h$ is the identifying number of $s$), the corresponding channel $s_h$ (suitably decorated), and the corresponding channel $cxt_h$ appear in $\upsilon$; 
  %processes $P_s$ (or $\overline{P_s}$) 
%  for each  $Cxt_s$ are
 %involved in the transition $t$;
  \item for each reaction $a \in A$ and each entity $s \in S$, each virtual link offered by processes $P_a$ and $Cxt_s$ is  overlapped by exactly one solid link offered by  processes representing entities.
  %each process $P_s $ (or its alternative version $\overline{P}_s$), with $s \in S$, participates at most  with one solid link for each process $P_a$, and with exactly one solid link for each process $Cxt_s$.%, if $s$ is a \emph{required element} of $A_j$ in $t$.
\end{enumerate}
\end{restatable}

\begin{example}\label{ex:backbone}
%\begin{color}{red}
Let ${\cal A}$ be a RS whose specification contains two entities, $s1$ and $s2$, and, among the others, the reaction $a=(\{s1\},\{\dots\} , \{s1\})$ that guarantees the persistence of the entity $s1$ once it is present in the system. Note that we use here a dummy inhibitor $(\dots)$ which will never be present. Then, we assume an  extended interactive process $\pi=(\gamma,\delta)$ where the context $\gamma$ provides  $s1$ and $s2$.
 %, for the first $n$ steps.
 Our translation includes the processes:
 %\end{color}
 \[
 \begin{array}{lcllcl}
 P_a &\defeq& \startchain{\silent}\chainedlink{s1_i}{\noact}\chainedlink{\noact}{s1_o}\chainedlink{r_2}{\noact} \chainedlink{\noact}{p_1}\chainedlink{\tilde{s1}_i}{\noact}\chainedlink{\noact}{\tilde{s1}_o}\chainend{p_2}.P_a + \dots; \\
 P_{s1} &\defeq& \startchain{s1_i}\chainedlink{s1_o}{\noact}\chainedlink{\noact}{\hat{s1}_i}\chainedlink{\hat{s1}_o}{\noact}\chainedlink{\noact}{\tilde{s1}_i}\chainedlink{\tilde{s1}_o}.P_{s1} + \dots; &
  P_{s2} &\defeq& \link{\hat{s2}_i}{\hat{s2}_o}.P_{s2} + \dots;\\
 Cxt_{s1}& \defeq & \startchain{cxt_1}\chainedlink{\hat{s1}_i}{\noact}\chainedlink{\noact}{\hat{s1}_o}\chainend{cxt_2}.Cxt_{s1} &
  Cxt _{s2} &\defeq& \startchain{cxt_2}\chainedlink{\hat{s2}_i}{\noact}\chainedlink{\noact}{\hat{s2}_o}\chainend{p_1}. Cxt_{s2}
\end{array}
\]  \\
   Now, we assume that $s1$ is in the initial state of ${\cal A}$, and in Figure~\ref{fig:backbone} we show the  structure of a link chain label related to the execution of a transition of the c\CNA \ system: $\restrict{names} (P_{s1}|P_{s2}| P_a|\dots|Cxt_{s1}|Cxt_{s2})$. The yellow blocks are referred to the processes encoding the reactions ($P_a$, in our case) and the contexts ($Cxt_{s1}$ and $Cxt_{s2}$). As the figure puts in evidence, these two kinds of processes determine the structure of the link chain, from end to end, i.e. from the left $\silent$ to the right one. We could say that these processes form the \emph{backbone} of the interaction.
   In contrast, the processes encoding the entities ($P_{s1}$, and $P_{s2}$, in our case) provides the
    solid links to overlap the virtual links   of the backbone.

  \begin{figure}[t]
  \begin{tikzpicture}
  \draw (0,1) -- (3,1) ;
  \draw (4,1) -- (5,1) ;
  \draw (0.5,0) -- (3.5,0);
  \draw (4.5,0) -- (5.5,0) ;
  \draw (0,1)--(0,0.5) -- (0.5,0.5) -- (0.5,0) -- (1.5,0) ;
  \draw (0.25,0.75)  node[minimum size=0.495cm,fill=yellow] {${ \silent}$};
  \draw (0.75,0.75)  node[minimum size=0.5cm,fill=yellow] {$\ $};
  \draw (0.76,0.25)  node[minimum size=0.5cm,fill=yellow] {$\ $};
  \draw (1.21,0.25)  node[minimum size=0.5cm,fill=yellow] {${\scriptscriptstyle s1_i}$};

  \draw (2,1) -- (2,0.5) -- (2.5,0.5) -- (2.5,0) -- (1.5,0)  -- (1.5,0.5) -- (1,0.5) -- (1,1);

 \draw (1.25,0.75) node{$\noact$};%node[minimum size=0.5cm,draw,draw=white] {$\noact$};
 \draw (2.25,0.25) node{$\noact$};
 \draw (2.31,0.75)  node[minimum size=0.47cm,draw,fill=yellow, draw=yellow] {${\scriptscriptstyle s1_o}$};
  \draw (2.75,0.75)  node[minimum size=0.48cm,draw,fill=yellow,yellow] {$\ $};
  \draw (2.75,0.26)  node[minimum size=0.49cm,draw,fill=yellow,yellow] {$\ $};

  \draw (3.245,0.25)  node[minimum size=0.475cm,draw,fill=yellow,draw=yellow] {${\scriptscriptstyle   r_2}$};

  \draw (3.75,0.5) node{$\dots$};

  \draw(3.5,0.5) -- (3.5,0);
%%     \draw (3.25,0.75)  node[minimum size=0.5cm,draw,draw=white] {$\noact$};
%%  \draw (3.75,0.75)  node[minimum size=0.5cm,draw,white] {$\ $};
%%  \draw (3.75,0.25)  node[minimum size=0.5cm,draw,white] {$\ $};
%%  \draw (4.25,0.25)  node[minimum size=0.5cm,draw,draw=white] {$\noact$};
\draw  (4,1) --  (4,0.5) -- (4.5,0.5) -- (4.5,0)   (3.5,0.5) -- (3,0.5) -- (3,1);
% \draw (3.25,0.75) node{$\noact$};%node[minimum size=0.5cm,draw,draw=white] {$\noact$};
%  \draw (4.25,0.25) node{$\noact$};
\draw (4.75,0.745)  node[minimum size=0.485cm,draw,fill=yellow,yellow] {$\ $};
  \draw (4.38,0.75) node[minimum size=0.47cm,draw,fill=yellow, draw=yellow] 
  {${\scriptscriptstyle  cxt_1}$};

  \draw (4.75,0.25)  node[minimum size=0.48cm,draw,fill=yellow,yellow] {$\ $};
  \draw (5.2,0.255)  node[minimum size=0.485cm,draw,fill=yellow,draw=yellow] {${\scriptscriptstyle  \hat{s1}_i}$};
  \draw(5,1) -- (5,0.5) -- (5.5,0.5) -- (5.5,0);

   \draw (5.75,0.5) node{$\dots$};

    \draw (6,1) -- (11.5,1) ;
    \draw (7,1) -- (7,0.5) -- (7.5,0.5) -- (7.5,0);
    \draw (6.5,0) -- (11.5,0);
    \draw   (6,1)-- (6,0.5) -- (6.5,0.5) -- (6.5,0.5) -- (6.5,0) ;
     \draw (6.75,0.75)  node[minimum size=0.47cm,fill=yellow, draw=yellow] {$\ $};
 \draw (6.36,0.75) node[minimum size=0.475cm,draw,fill=yellow,draw=yellow] 
 {${\scriptscriptstyle cxt_2}$};%node[minimum size=0.5cm,draw,draw=white] {$\noact$};
   \draw (6.75,0.26)  node[minimum size=0.485cm,fill=yellow, draw=yellow] {$\ $};

 \draw (7.20,0.25) node[minimum size=0.475cm,fill=yellow, draw=yellow] {${\scriptscriptstyle \hat{s2}_i}$};

     \draw (7.3,0.75)   node{$\noact$};
  \draw (7.75,0.75) node{$\ $};
  \draw (7.76,0.26) node {$\ $};
  \draw (8.25,0.26)  node {$\noact$ };

  \draw (8.3,0.75)  node[minimum size=0.48cm,draw,fill=yellow, draw=yellow] {${\scriptscriptstyle \hat{s2}_o}$};
  \draw (8,1)--(8,0.5)--(8.5,0.5) -- (8.5,0);
 % \draw(9.5,0.5) --(9.5,0);
  \draw (8.745,0.75)  node[minimum size=0.48cm,draw,fill=yellow,yellow] {$\ $};
  \draw (8.75,0.26)  node[minimum size=0.48cm,draw,fill=yellow,yellow] {$\ $};
 \draw (9.24,0.26)  node[minimum size=0.48cm,draw,fill=yellow,draw=yellow] {${\scriptscriptstyle p_1}$};
 %\draw(9.5,0.5) --(9.5,0);
    % \draw (9,1) -- (10,1) -- (10,0.5) -- (10.5,0.5) --  (10.5,0) -- (9.5,0)  (9.5,0.5) -- (9,0.5) -- (9,1);
 \draw (9.24,0.75) node[minimum size=0.48cm,draw,fill=yellow,draw=yellow]{${\scriptscriptstyle p_1}$};%node[minimum size=0.5cm,draw,draw=white] {$\noact$};
 \draw (10.22,0.26) node [minimum size=0.48cm,draw,fill=yellow,draw=yellow]{${\scriptscriptstyle \tilde{s1}_i}$};
 \draw (9.74,0.745) node [minimum size=0.485cm,draw,fill=yellow,draw=yellow]{$\ $};
 \draw (9.735,0.26) node [minimum size=0.48cm,draw,fill=yellow,draw=yellow]{$\ $};
 \draw (10,1) -- (10,0.5) -- (10.5,0.5)--(10.5,0);
  \draw (10.31,0.75)  node{$\noact$};
  \draw (10.75,0.75)  node {$\ $};
  \draw (10.75,0.26)  node{$\ $};
 \draw (11.25,0.25)  node {$\noact$};
   \draw (11,1) -- (11,0.5) -- (11.5,0.5) -- (11.5,0);% -- (10,0) -- (9,0) -- (9,0.5) -- (8.5,0.5) -- (8.5,1);
    \draw (11,1) -- (11.5,1) -- (12,1) -- (12,0.5) -- (12.5,0.5)--(12.5,0)--(11.5,0);
   %\draw (11.25,0.25)  node[minimum size=0.48cm,draw,fill=yellow,draw=yellow] {$\noact$};
    \draw (11.31,0.75) node[minimum size=0.48cm,draw,fill=yellow,draw=yellow]{${\scriptscriptstyle \tilde{s1}_o}$};
    \draw (11.75,0.26)node[minimum size=0.48cm,draw,fill=yellow,draw=yellow]{$\ $};
      \draw (11.75,0.75)node[minimum size=0.48cm,draw,fill=yellow,draw=yellow]{$\ $};
     \draw (12.2,0.25) node[minimum size=0.478cm,draw,fill=yellow,draw=yellow]{\;$\silent$\;  };
  \end{tikzpicture}
  \caption{The link chain structure arising from reactions and context processes.}
   \label{fig:backbone}
 \end{figure}
 \end{example}

% \begin{figure}[t]
% \begin{tikzpicture}
%   \draw (0,1) -- (5,1) ;
%    \draw (0.5,0) -- (5.5,0) ;
%   \draw (0,1)--(0,0.5) -- (0.5,0.5) -- (0.5,0) -- (1.5,0) ;
%     \draw (0.25,0.75)  node[minimum size=0.495cm,fill=yellow] {${ \silent}$};
% \draw (0.75,0.75)  node[minimum size=0.5cm,fill=yellow] {$\ $};
%  \draw (0.76,0.25)  node[minimum size=0.5cm,fill=yellow] {$\ $};
%  \draw (1.21,0.25)  node[minimum size=0.5cm,fill=yellow] {${\scriptscriptstyle s1_i}$};
%%    \draw (1.25,0.75)  node[minimum size=0.5cm,draw,draw=white] {$\noact$};
%%  \draw (1.75,0.75)  node[minimum size=0.5cm,draw,white] {$\ $};
%%  \draw (1.75,0.25)  node[minimum size=0.5cm,draw,white] {$\ $};
%%  \draw (2.25,0.25)  node[minimum size=0.5cm,draw,draw=white] {$\noact$};
%  \draw (2,1) -- (2,0.5) -- (2.5,0.5) -- (2.5,0) -- (1.5,0)  -- (1.5,0.5) -- (1,0.5) -- (1,1);
%
% \draw (1.25,0.75) node{$\noact$};%node[minimum size=0.5cm,draw,draw=white] {$\noact$};
%  \draw (2.25,0.25) node{$\noact$};
%   \draw (2.31,0.75)  node[minimum size=0.47cm,draw,fill=yellow, draw=yellow] {${\scriptscriptstyle s1_o}$};
%  \draw (2.75,0.75)  node[minimum size=0.48cm,draw,fill=yellow,yellow] {$\ $};
%  \draw (2.75,0.26)  node[minimum size=0.49cm,draw,fill=yellow,yellow] {$\ $};
%
%  \draw (3.20,0.25)  node[minimum size=0.47cm,draw,fill=yellow,draw=yellow] {${\scriptscriptstyle  \tilde{s1}_i}$};
%  
%  
%  \draw(3.5,0.5) -- (3.5,0);
%%     \draw (3.25,0.75)  node[minimum size=0.5cm,draw,draw=white] {$\noact$};
%%  \draw (3.75,0.75)  node[minimum size=0.5cm,draw,white] {$\ $};
%%  \draw (3.75,0.25)  node[minimum size=0.5cm,draw,white] {$\ $};
%%  \draw (4.25,0.25)  node[minimum size=0.5cm,draw,draw=white] {$\noact$};
%\draw (3,1) -- (4,1) -- (4,0.5) -- (4.5,0.5) -- (4.5,0) -- (3.5,0)  (3.5,0.5) -- (3,0.5) -- (3,1);
% \draw (3.25,0.75) node{$\noact$};%node[minimum size=0.5cm,draw,draw=white] {$\noact$};
%  \draw (4.25,0.25) node{$\noact$};
%  \draw (4.31,0.75)  node[minimum size=0.47cm,draw,fill=yellow, draw=yellow] {${ \scriptscriptstyle \tilde{s1}_o}$};
%  \draw (4.75,0.745)  node[minimum size=0.48cm,draw,fill=yellow,yellow] {$\ $};
%  \draw (4.75,0.25)  node[minimum size=0.48cm,draw,fill=yellow,yellow] {$\ $};
%  \draw (5.24,0.255)  node[minimum size=0.485cm,draw,fill=yellow,draw=yellow] {${\scriptscriptstyle  r_2}$};
%  \draw(5,1) -- (5,0.5) -- (5.5,0.5) -- (5.5,0);
%   \draw (6,0.5) node{$\dots$};
%    \draw (6.5,1) -- (11.5,1) ;
%    \draw (7.5,1) -- (7.5,0.5) -- (8,0.5) -- (8,0);
%    \draw (7,0) -- (12,0);
%    \draw   (6.5,1)-- (6.5,0.5) -- (7,0.5) -- (7,0.5) -- (7,0) ;
% \draw (6.77,0.75) node{$\noact$};%node[minimum size=0.5cm,draw,draw=white] {$\noact$};
% \draw (7.77,0.25) node{$\noact$};
%
%     \draw (7.80,0.75)  node[minimum size=0.485cm,draw,fill=yellow, draw=yellow] {${\scriptscriptstyle  \tilde{s1}_o}$};
%  \draw (8.25,0.75)  node[minimum size=0.485cm,draw,fill=yellow,yellow] {$\ $};
%  \draw (8.26,0.26)  node[minimum size=0.495cm,draw,fill=yellow,yellow] {$\ $};
%  \draw (8.75,0.26)  node[minimum size=0.49cm,draw,fill=yellow,draw=yellow] {${\scriptscriptstyle cxt_2}$ };
%
%  \draw (8.75,0.75)  node[minimum size=0.48cm,draw,fill=yellow, draw=yellow] {${\scriptscriptstyle cxt_2}$};
%  \draw (9.25,0.75)  node[minimum size=0.48cm,draw,fill=yellow,yellow] {$\ $};
%  \draw (9.25,0.26)  node[minimum size=0.48cm,draw,fill=yellow,yellow] {$\ $};
% \draw (9.71,0.25)  node[minimum size=0.47cm,draw,fill=yellow,draw=yellow] {${\scriptscriptstyle \tilde{s2}_i}$};
% \draw(10,0.5) --(10,0);
%     \draw (9.5,1) -- (10.5,1) -- (10.5,0.5) -- (11,0.5) --  (11,0) -- (10,0)  (10,0.5) -- (9.5,0.5) -- (9.5,1);
% \draw (9.75,0.75) node{$\noact$};%node[minimum size=0.5cm,draw,draw=white] {$\noact$};
% \draw (10.75,0.25) node{$\noact$};
%  \draw (10.81,0.75)  node[minimum size=0.47cm,draw,fill=yellow, draw=yellow] {${\scriptscriptstyle \tilde{s2}_o}$};
%  \draw (11.25,0.75)  node[minimum size=0.48cm,draw,fill=yellow,yellow] {$\ $};
%  \draw (11.25,0.26)  node[minimum size=0.49cm,draw,fill=yellow,yellow] {$\ $};
% \draw (11.75,0.25)  node[minimum size=0.48cm,draw,fill=yellow,draw=yellow] {$\silent$};
%   \draw (11.5,1) -- (11.5,0.5) -- (12,0.5) -- (12,0);% -- (10,0) -- (9,0) -- (9,0.5) -- (8.5,0.5) -- (8.5,1);
%  \end{tikzpicture}
%  \caption{The link chain structure arising from reactions and context processes.}
%   \label{fig:backbone}
% \end{figure}
% \end{example}

Example~\ref{ex:backbone}   outlines two different roles of the processes defining the translation of an interactive process: those processes encoding the reactions and the context provide the backbone of each transition, whereas the processes encoding the entities provide the resources needed for the communication to take place.

With the next proposition, we analyse the structure of a c\CNA \ 
process encoding of  a reactive process after one step transition.
In the following four statements, for brevity, we let ${\cal A} = (S, A)$ be a RS, and let  $\pi=(\gamma,\delta)$ be an extended interactive process in $A$, with $\gamma=\{C_i\}_{i\in\mathbb{N}}$ and $\delta=\{D_i\}_{i\in\mathbb{N}}$. Moreover, we denote by $\pi^j$ the shift of $\pi$ starting at the $j$-th state sequence; formally we let $\pi^j=(\gamma^j,\delta^j)$ with
$\gamma^j=\{C_i'\}_{i\in\mathbb{N}}$, $\delta^j=\{D_i'\}_{i\in\mathbb{N}}$ such that 
$C_0'=C_j \cup D_j$, 
%\begin{color}{red}
$D_0' = \emptyset$,
%\end{color}
and   $C_i'=C_{i+j}$, $D_i'= D_{i+j}$ for any $i\geq 1$.

%\begin{color}{red}
\begin{restatable}[Correctness 1]{proposition}{propcorruno}
 \label{prop:corr1}
 Let 
 $P=\llbracket{\cal A},\gamma\rrbracket$ with
$$P  = \restrict{names} (\Pi_{a \in A}P_a| \Pi_{s \in S}Cxt_s^1 | \Pi_{s \in C_0}P_s| \Pi_{s \notin C_0}\overline{P}_s).$$
 If $P \xrightarrow{v}Q$, then
  $v$ is a silent action and $Q = \llbracket {\cal A},\gamma^1\rrbracket$, namely
$$Q= \restrict{names} (\Pi_{a \in A}P_a| \Pi_{s \in S}Cxt_s^2 | \Pi_{s \in C_1 \cup D_1}P_s| \Pi_{s \notin C_1 \cup D_1}\overline{P}_s).$$
\end{restatable}

Now, we extend the previous result to a series of transitions.

 \begin{restatable}[Correctness 2]{corollary}{corrcorrdue} 
 \label{corr:corr2}
  Let $P = \llbracket{\cal A},\gamma\rrbracket$ and $j\geq 1$.
If there exists  $Q$ such that $P \rightarrow^j Q$, then 
$Q =\llbracket{\cal A},\gamma^j\rrbracket$.
\end{restatable}

Conversely, we prove that the $c\CNA$ process $\llbracket{\cal A},\gamma\rrbracket$ can simulate all the evolutions  of the underlying interactive process.

 \begin{restatable}[Completeness 1]{proposition}{propcompluno} 
 \label{prop:compl1}
  $\llbracket{\cal A},\gamma\rrbracket \rightarrow \llbracket{\cal A},\gamma^1\rrbracket$.

\end{restatable}

Now, we extend the previous result to a series of transitions.

\begin{restatable}[Completeness 2]{corollary}{corrcompldue} 
 \label{corr:compl2}
$\llbracket{\cal A},\gamma\rrbracket \rightarrow^j \llbracket{\cal A},\gamma^j\rrbracket$.  
 \end{restatable}
%\end{color}

% 
% \begin{proposition}[correctness]
% Let $P$ a c\CNA\  process such that exists a reaction system ${\cal A}$ such that $P=[|{\cal A}|]$, then if $P \rightarrow P'$, then $\exists$ ${\cal A'}$ such that ${\cal A}$  evolves in ${\cal A'}$ in one step and $P'=[|{\cal A'}|]$.
% \end{proposition}

### 2020-07-27 Catuscia - final/06-exampleLactose.tex
da dropbox

--- 2020-06-18 TCS_version/06-exampleLactose.tex
@@
 \begin{itemize}
-\item the gene $lacI$ encodes for a repressor protein $I$;
+\item 
+%\begin{color}{red}
+the gene $lacI$ encodes a repressor protein $I$;
+%\end{color}
 \item the DNA sequence, called \emph{promoter},  is recognised by a RNA polymerase
@@
 follows:
-$S$ is the set\\ $\{lac, Z, Y, A, lacI, I, I\textrm{-}OP, cya, cAMP, crp, CAP,cAMP\textrm{-}CAP, lactose,glucose\}$,
+$S$ is the set
+{\small
+$$\{lac, Z, Y, A, lacI, I, I\textrm{-}OP, cya, cAMP, crp, CAP,cAMP\textrm{-}CAP, lactose,glucose\},$$
+}
 and A consists of the following 10 reactions:
@@
 \item for the entities in the default context, $s \in DC$, as they are persistent, we do not provide the $\overline{P_s}$ processes and the $Cxt_s$ processes;
-\item for the reactions requiring the presence of entities $s \in DC$, we do not provide the reaction alternative behaviour for when $s$ is absent;
+\item for the reactions requiring the presence of entities $s \in DC$, we do not provide the reaction alternative behaviour when $s$ is absent;
 \item the $Cxt_s$ processes are specified only for those entitiess that are really provided by the context.
@@
 }
-
-%\underline{Processes for the rules}\\
-%We provide parametrised code duplication and for expression rules.\\
-%\ \\
-\paragraph{Duplication reactions.}
-As explained previously, their encoding is omitted.
+%\paragraph{Duplication reactions.}
+%As explained previously, their encoding is omitted.
 %%$
@@
  P_i(s1,s2) & \defeq & \startchain{r_i}\chainedlink{s1_i}{\noact}\chainedlink{\noact}{s1_o}
- \chainedlink{r_{i+1}}{\noact}\chainedlink{\noact}{p_{i}}\chainedlink{\widetilde{s2}_i}{\noact}
-\chainedlink{\noact}{\widetilde{s2}_o}\chainend{p_{i+1}}.P_i(s1,s2)\\
+ \chainedlink{r_{i+1}}{\noact}\chainedlink{\noact}{p_{i}}\chainedlink{\tilde{s2}_i}{\noact}
+\chainedlink{\noact}{\tilde{s2}_o}\chainend{p_{i+1}}.P_i(s1,s2)\\
 \end{array}
@@
  && \startchain{r_2}\chainedlink{lactose_i}{\noact}\chainedlink{\noact}{lactose_o}\chainedlink{r_3}{\noact}\chainedlink{\noact}{p_2}\chainend{p_3}.P_{a2} \ +\  \startchain{r_2}\chainedlink{\overline{I}_i}{\noact}\chainedlink{\noact}{\overline{I}_o}\chainedlink{r_3}{\noact}\chainedlink{\noact}{p_2}\chainend{p_3}.P_{a2}\\[5pt]
+ \\
 %%P_{a5} & =& \startchain{r_5}\chainedlink{cya_i}{\noact}\chainedlink{\noact}{cya_o}\chainedlink{\tilde{cya}_i}{\noact}\chainedlink{\noact}{\tilde{cya}_o}\chainend{r_6}.P_{a5}\\
@@
 &&  \startchain{r_5}\chainedlink{\overline{cAMP}_i}{\noact}\chainedlink{\noact}{\overline{cAMP}_o} \chainedlink{r_6}{\noact}\chainedlink{\noact}{p_5}\chainend{p_6}.P_{a5} \ + \ \startchain{r_5}\chainedlink{\overline{CAP}_i}{\noact}\chainedlink{\noact}{\overline{CAP}_o} \chainedlink{r_6}{\noact}\chainedlink{\noact}{p_5}\chainend{p_6}.P_{a5}\\[5pt]
+\\
 %\end{array}
@@
 P_{a6} & \defeq& \startchain{r_{6}}\chainedlink{lac_i}{\noact}\chainedlink{\noact}{lac_o}
-\chainedlink{cAMP_i}{\noact}\chainedlink{\noact}{cAMP_o}\chainedlink{\overline{I\textrm{-}OP}_i}{\noact}
+\chainedlink{cAMP\textrm{-}CAP_i}{\noact}\chainedlink{\noact}{cAMP\textrm{-}CAP_o}\chainedlink{\overline{I\textrm{-}OP}_i}{\noact}
 \chainedlink{\noact}{\overline{I\textrm{-}OP}_o}\chainedlink{cxt_1}{\noact}
 \chainedlink{\noact}{p_6}
-\chainedlink{\widetilde{z}_i}{\noact}\chainedlink{\noact}{\widetilde{z}_o}
-\chainedlink{\widetilde{y}_i}{\noact}\chainedlink{\noact}{\widetilde{y}_o}
-\chainedlink{\widetilde{A}_i}{\noact}\chainedlink{\noact}{\widetilde{A}_o}
+\chainedlink{\tilde{z}_i}{\noact}\chainedlink{\noact}{\tilde{z}_o}
+\chainedlink{\tilde{y}_i}{\noact}\chainedlink{\noact}{\tilde{y}_o}
+\chainedlink{\tilde{A}_i}{\noact}\chainedlink{\noact}{\tilde{A}_o}
 \chainend{\silent}.P_{a6}\\
@@
 && + \\
-&&  \startchain{r_{6}}\chainedlink{\overline{lac}_i}{\noact}\chainedlink{\noact}{\overline{lac}_o} \chainedlink{cxt_1}{\noact}\chainedlink{\noact}{p_6}\chainend{\silent}.P_{a6} \ + \ \chainedlink{\overline{cAMP}_i}{\noact}\chainedlink{\noact}{\overline{cAMP}_o} \chainedlink{cxt_1}{\noact}\chainedlink{\noact}{p_6}\chainend{\silent}.P_{a6}
+&&  \startchain{r_{6}}\chainedlink{\overline{lac}_i}{\noact}\chainedlink{\noact}{\overline{lac}_o} \chainedlink{cxt_1}{\noact}\chainedlink{\noact}{p_6}\chainend{\silent}.P_{a6} \ + \ \chainedlink{\overline{cAMP\textrm{-}CAP}_i}{\noact}\chainedlink{\noact}{\overline{cAMP\textrm{-}CAP}_o} \chainedlink{cxt_1}{\noact}\chainedlink{\noact}{p_6}\chainend{\silent}.P_{a6}
 \end{array}
@@
 \paragraph{Processes for the entities.}
-We exploit the specificity of the example in hand to optimising the code, and we specify exactly
+We exploit the specificity of the example in hand to optimise the code, and we specify exactly
 the number of solid links that each process encoding an entity must offer.
 For the always present entities  we let:
+%{\color{red} For the always present entities we define the parametric process $P(s)\defeq \link{s_i}{s_o}.P_{s}$, and we let:
+%\[
+%\begin{array}{lcl@{\qquad}lcl@{\qquad}lcl@{\qquad}lcl}
+%P_{cya}&\defeq&P(cya) &P_{crp}&\defeq&P(crp)&P_{lacI}&\defeq&P(lacI) &
+% P_{lac}&\defeq&P(lac)
+%\end{array}\]
 \[
 \begin{array}{lcl@{\qquad}lcl}
-%P_{lac}&\defeq&\startchain{lac_i}\chainedlink{lac_o}{\noact}\chainedlink{\noact}{lac_i}\chainedlink{lac_o}{\noact}\chainedlink{\noact}{\tilde{lac}_i}\chainend{\tilde{lac}_o}.P_{lac} +
- %\startchain{lac_i}\chainedlink{lac_o}{\noact}\chainedlink{\noact}{\tilde{lac}_i}\chainend{\tilde{lac}_o}.P_{lac}\\
-% P_{lac}&\defeq&\sum_{h=1}^2(\startchain{lac_i}\chainedlink{lac_o}{\noact}\chainend{\noact})^h\ \link{\tilde{lac}_i}{\tilde{lac}_o}.P_{lac}
-%P_{cya}&\defeq&\startchain{cya_i}\chainedlink{cya_o}{\noact}\chainedlink{\noact}{cya_i}\chainedlink{cya_o}{\noact}
-%\chainedlink{\noact}{\tilde{cya}_i}\chainend{\tilde{cya}_o}.P_{cya}\\
 P_{cya}&\defeq&\link{cya_i}{cya_o}.P_{cya} &
-%&
-%P_{I}&\defeq&\sum_{h=0}^1(\startchain{I_i}\chainedlink{I_o}{\noact}\chainend{\noact})^h\ \link{\tilde{I}_i}{\tilde{I}_o}.P_{I} \\
- %{\noact}
-%\chainedlink{\noact}{\cap{lac}_i}\chainend{\cap{lac}_o}.P_{lac}\\
-%P_{crp}&\defeq&\startchain{crp_i}\chainedlink{crp_o}{\noact}\chainedlink{\noact}{crp_i}\chainedlink{crp_o}{\noact}\chainedlink{\noact}{\tilde{crp}_i}\chainend{\tilde{crp}_o}.P_{crp}
 P_{crp}&\defeq&\link{crp_i}{crp_o}.P_{crp}\\[5pt]
-%&
-%P_{cAMP}&\defeq&\sum_{h=0}^1(\startchain{cAMP_i}\chainedlink{cAMP_o}{\noact}\chainend{\noact})^h\ \link{\widetilde{cAMP}_i}{\widetilde{cAMP}_o}.P_{cAMP}
 P_{lacI}&\defeq& \link{lacI_i}{lacI_o}.P_{lacI} &
-%&
-%P_{CAP}&\defeq&\sum_{h=0}^1(\startchain{CAP_i}\chainedlink{CAP_o}{\noact}\chainend{\noact})^h\ \link{\widetilde{CAP}_i}{\widetilde{CAP}_o}.P_{CAP}\\
-%{lacI_i}\chainedlink{lacI_o}{\noact}\chainedlink{\noact}{\tilde{lacI}_i}\chainend{\tilde{lacI}_o}.P_{lacI}\\
-% \startchain{\tilde{I}_i}\chainend{\tilde{I}_o}.P_{I}\\
-%P_{I}&\defeq&\startchain{I_i}\chainedlink{I_o}{\noact}\chainedlink{\noact}{\tilde{I}_i}\chainend{\tilde{I}_o}.P_{I} +
-% \startchain{\tilde{I}_i}\chainend{\tilde{I}_o}.P_{I}\\
-%P_{cya}&\defeq&\startchain{cya_i}\chainedlink{cya_o}{\noact}\chainedlink{\noact}{cya_i}\chainedlink{cya_o}{\noact}\chainedlink{\noact}{\tilde{cya}_i}\chainend{\tilde{cya}_o}.P_{cya}\\
-%P_{lacI}&\defeq&\sum_{h=1}^2(\startchain{lacI_i}\chainedlink{lacI_o}{\noact}\chainend{\noact})^h\ \link{\tilde{lacI}_i}{\tilde{lacI}_o}.P_{lacI}\\
  P_{lac}&\defeq&\link{lac_i}{lac_o}.P_{lac}
-%P_{cAMP}&\defeq&\startchain{cAMP_i}\chainedlink{cAMP_o}{\noact}\chainedlink{\noact}{\tilde{cAMP}_i}\chainend{\tilde{cAMP}_o}.P_{cAMP} + \startchain{\tilde{cAMP}_i}\chainend{\tilde{cAMP}_o}.P_{cAMP}\\
-%P_{crp}&\defeq&\startchain{crp_i}\chainedlink{crp_o}{\noact}\chainedlink{\noact}{crp_i}\chainedlink{crp_o}{\noact}\chainedlink{\noact}{\tilde{crp}_i}\chainend{\tilde{crp}_o}.P_{crp}\\
 \end{array}\]
@@
 definition  
-$P_e(s) \defeq  \startchain{s_i}\chainedlink{s_o}{\noact}\chainedlink{\noact}{\widetilde{s}_i}\chainend{\widetilde{s}_o}.P_e(s) +  \startchain{\widetilde{s_i}}\chainend{\widetilde{s}_o} .P_e(s) $.\\
+$P_e(s) \defeq  \startchain{s_i}\chainedlink{s_o}{\noact}\chainedlink{\noact}{\tilde{s}_i}\chainend{\tilde{s}_o}.P_e(s) +  \startchain{\tilde{s_i}}\chainend{\tilde{s}_o} .P_e(s) $.\\
 % There are four entities of the first type: $P_d(lac)$, $P_d(cya)$, $P_d(crp)$, $P_d(lacI)$.
@@
 Correspondingly, the process  $P_{I\textrm{-}OP}$ is defined as follows:
-
 \[\begin{array}{lcl}
@@
 \noindent
-The  process $P_{cAMP\textrm{-}CAP}$ is similar to $P_{I\textrm{-}OP}$, as it is produced by $a_5$ and tested for presence by $a_6$. Its code is  in Table~\ref{tab:cAMP-CAP}, in the Appendix.
+The  process $P_{cAMP\textrm{-}CAP}$ is similar to $P_{I\textrm{-}OP}$, as it is produced by $a_5$ and tested for presence by $a_6$. 
+Its code is omitted.
+%Its code is  in Table~\ref{tab:cAMP-CAP}, in the Appendix.
+
 The $lactose$ is provided by the context and tested for absence by $a_2$.
@@
 && \sum_{h=0}^1 (\startchain{lactose_i}\chainedlink{lactose_o}{\noact}\chainend{\noact})^h\ \link{\underline{lactose_i}}{\underline{lactose_o}}.\overline{P_{lactose}}\\
-%P_{glucose} &\defeq &\sum_{h=0}^1 \underbrace{\link{glucose_i}{glucose_o}}_{h\ times}
-%
-%\[
-%\begin{array}{lcl}
-
+\\
 \overline{P_{lactose}} &\defeq &\sum_{h=0}^1 (\startchain{\overline{lactose}_i}\chainedlink{\overline{lactose}_o}{\noact}\chainend{\noact})^h\ \link{\widehat{lactose_i}}{\widehat{lactose_o}}.P_{lactose}\\
@@
 \noindent
-The  process $P_{glucose}$ is similar to $P_{lactose}$ and tested for absence by $a_5$. Its code is  in Table~\ref{tab:glucose}, in the Appendix.
+The  process $P_{glucose}$ is similar to $P_{lactose}$ and tested for absence by $a_5$. 
+Its code is omitted.
+%Its code is  in Table~\ref{tab:glucose}, in the Appendix.
+
 The entity $z$ can only be produced by rule $a_6$, while it is never provided by the context. Moreover, there is no rule for testing its presence or absence.
@@
 \begin{array}{lcl@{\qquad\qquad}lcl}
-P_z&\defeq &\startchain{\widetilde{z_i}}\chainedlink{\widetilde{z_o}}{\noact}\chainedlink{\noact}{\underline{z_i}}\chainend{\underline{z_o}}.P_z\ + \
+P_z&\defeq &\startchain{\tilde{z_i}}\chainedlink{\tilde{z_o}}{\noact}\chainedlink{\noact}{\underline{z_i}}\chainend{\underline{z_o}}.P_z\ + \
 \link{\underline{z_i}}{\underline{z_o}}.\overline{P_z}
 &
-\overline{P_z} &\defeq &\startchain{\widetilde{z_i}}\chainedlink{\widetilde{z_o}}{\noact}\chainedlink{\noact}{\underline{z_i}}\chainend{\underline{z_o}}.\overline{P_z}\ + \
+\overline{P_z} &\defeq &\startchain{\tilde{z_i}}\chainedlink{\tilde{z_o}}{\noact}\chainedlink{\noact}{\underline{z_i}}\chainend{\underline{z_o}}.\overline{P_z}\ + \
 \link{\underline{z_i}}{\underline{z_o}}.\overline{P_z}\\\end{array}
@@
 \noindent
-The entities $y$ and $A$ are treated in the same way as $z$. Their processes are in Table~\ref{tab:yA} in the Appendix.
+The entities $y$ and $A$ are treated likewise $z$. Their processes are omitted.
+%in Table~\ref{tab:yA} in the Appendix.
 
@@
 The entities in $DC$ are assumed always present by default, so no context process is needed for them.
-The entities $z$, $y$, and $A$ are assumed never provided by the context. Their processes are
+The entities $z$, $y$, and $A$ are assumed never provided by the context:
+% Their processes are
 \[
-\begin{array}{rcl@{\qquad\qquad}rcl}
+\begin{array}{rcl@{\qquad}rcl}
 Cxt_{z}&\defeq & \startchain{cxt_1}\chainedlink{\underline{z}_i}{\noact}\chainedlink{\noact}{\underline{z}_o}\chainend{cxt_2}.Cxt_{z}
@@
 \]
-\noindent
-%Finally, we need to specify the context for  $lactose$ and $glucose$.
-For the sake of presentation, we assume that the $lactose$ is always provided by the context, in contrast,
-$glucose$ is never provided.
+ Also, for the sake of presentation, we assume that the $glucose$ is never provided and $lactose$ is always provided by the context:
 \[
-\begin{array}{lcllcl}
-Cxt_{lactose}&\defeq& \startchain{cxt_4}\chainedlink{\widehat{lactose}_i}{\noact}\chainedlink{\noact}{\widehat{lactose}_o}\chainend{cxt_5}.Cxt_{lactose}
-\\
+\begin{array}{lcl}
+Cxt_{lactose}&\defeq& \startchain{cxt_4}\chainedlink{\widehat{lactose}_i}{\noact}\chainedlink{\noact}{\widehat{lactose}_o}\chainend{cxt_5}.Cxt_{lactose}\\
 Cxt_{glucose}&\defeq & \startchain{cxt_5}\chainedlink{\underline{glucose}_i}{\noact}\chainedlink{\noact}{\underline{glucose}_o}\chainend{p_1}.Cxt_{glucose}
@@
 \]
+
+\noindent
+%For the sake of presentation, we assume that the $lactose$ is always provided by the context, in contrast,
+%$glucose$ is never provided.
+%\[
+%\begin{array}{lcllcl}
+%Cxt_{lactose}&\defeq& \startchain{cxt_4}\chainedlink{\widehat{lactose}_i}{\noact}\chainedlink{\noact}{\widehat{lactose}_o}\chainend{cxt_5}.Cxt_{lactose}
+%\\
+%Cxt_{glucose}&\defeq & \startchain{cxt_5}\chainedlink{\underline{glucose}_i}{\noact}\chainedlink{\noact}{\underline{glucose}_o}\chainend{p_1}.Cxt_{glucose}
+%\end{array}
+%\]
 \noindent
@@
 \paragraph{Execution.}
-Now, we show two transitions. 
-After the first transition the entity $cAMP\textrm{-}CAP$ is produced due to the absence of $glucose$, while the presence of $lactose$ inhibits the production of $I\textrm{-}OP$:
-$lacOp \xrightarrow{\restrict{names} v} lacOp'$
-where: 
+Now, we show  the execution of two transitions. 
+After the first transition, the entity $cAMP\textrm{-}CAP$ is produced due to the absence of  $glucose$(see $P_{a5}$), while the presence of $lactose$ inhibits the production of $I\textrm{-}OP$ (see $P_{a2}$): $lacOp \xrightarrow{\restrict{names} v} lacOp'$,
+where
 {\small
@@
 with $AP= DC\cup \{cAMP\textrm{-}CAP\}$  the actual context.\\
-After the second step the entities $z$, $y$ and $A$ are produced, due to the presence of $cAMP\textrm{-}CAP$ and the absence of $I\textrm{-}OP$, thus
+After the second step the entities $z$, $y$ and $A$ are produced, due to the presence of $cAMP\textrm{-}CAP$ and the absence of $I\textrm{-}OP$ (see $P_{a6}$), thus
 $lacOp' \xrightarrow{\restrict{names}v'} lacOp''$
@@
 \chainedlink{cxt_1}{\dots}\startchain{\underline{glucose}_o}\chainedlink{p_1}{p_1}
-\chainedlink{\dots}{p_6}\chainedlink{\widetilde{z}_i}{\widetilde{z}_i}\chainedlink{\widetilde{z}_o}{\widetilde{z}_o}
-\chainedlink{\widetilde{y}_i}{\widetilde{y}_i}\chainedlink{\widetilde{y}_o}{\widetilde{y}_o}\chainedlink{\widetilde{A}_i}{\widetilde{A}_i}\chainedlink{\widetilde{A}_o}{\widetilde{A}_o}\chainend{\silent}\\[5pt]
+\chainedlink{\dots}{p_6}\chainedlink{\tilde{z}_i}{\tilde{z}_i}\chainedlink{\tilde{z}_o}{\tilde{z}_o}
+\chainedlink{\tilde{y}_i}{\tilde{y}_i}\chainedlink{\tilde{y}_o}{\tilde{y}_o}\chainedlink{\tilde{A}_i}{\tilde{A}_i}\chainedlink{\tilde{A}_o}{\tilde{A}_o}\chainend{\silent}\\
 %\chainend{cxt_1}\dots \link{\underline{glucose}_o}{\silent}\\[5pt]

### 2020-07-27 Catuscia - final/07-discussion.tex
da dropbox

--- 2020-04-05 07-discussion.tex
@@
 Our encoding increases the expressivity of  RS  concerning:
-%the behaviour of the context,
-the possibility of alternative behaviour of
-mutated entities, and  the communication between two different Reaction Systems.
+the behaviour of the context, the possibility of alternative behaviour of
+mutated entities and the communication between two different reaction systems.
 It is important to note that our encoding guarantees that from each state, 
@@
 
-%\subsection{Recursive contexts}
-%In RS, the behaviour of the context is finite. For the first $n$ steps, it is specified 
-% which are the entities that are provided from the context. 
-%Using c\CNA\ we can describe in a natural way the behaviour of the context in a recursive way. 
-%Then, the context behaviour would not necessarily end after $n$ steps,
-%and  could be infinite.
-%For example, in an extended interactive process, we may want that the entity $s$ is 
-%intermittently provided by the context every two steps: 
-%\[
-%\begin{array}{lcll}
-%Cxt_s & \defeq &  \startchain{cxt_j}\chainedlink{\hat{s}_i}{\noact}\chainedlink{\noact}{\hat{s}_o}\chainend{cxt_{j+1}}.Cxt_{s'}, &\mbox{ context provides $s$;} \\
-%Cxt_{s'} & \defeq & \startchain{cxt_j}\chainedlink{\underline{s}_i}{\noact}\chainedlink{\noact}{\underline{s}_o}\chainend{cxt_{j+1}}.Cxt_{s''},& \mbox{ context doesn't provide $s$;} \\
-%Cxt_{s''} & \defeq & \startchain{cxt_j}\chainedlink{\underline{s}_i}{\noact}\chainedlink{\noact}{\underline{s}_o}\chainend{cxt_{j+1}}.Cxt_{s}, & \mbox{ context doesn't provide $s$.} \\
-%\end{array}
-%\]
+\subsection{Recursive contexts}
+In RS, the behaviour of the context is finite. For the first $n$ steps, it is specified 
+ which are the entities that are provided from the context. 
+Using c\CNA\ we can describe in a natural way the behaviour of the context in a recursive way. 
+Then, the context behaviour would not necessarily end after $n$ steps,
+and  could be infinite.
+For example, in an extended interactive process, we may want that the entity $s$ is 
+intermittently provided by the context every two steps: 
+\[
+\begin{array}{lcll}
+Cxt_s & \defeq &  \startchain{cxt_j}\chainedlink{\hat{s}_i}{\noact}\chainedlink{\noact}{\hat{s}_o}\chainend{cxt_{j+1}}.Cxt_{s'}, &\mbox{ context provides $s$;} \\
+Cxt_{s'} & \defeq & \startchain{cxt_j}\chainedlink{\underline{s}_i}{\noact}\chainedlink{\noact}{\underline{s}_o}\chainend{cxt_{j+1}}.Cxt_{s''},& \mbox{ context doesn't provide $s$;} \\
+Cxt_{s''} & \defeq & \startchain{cxt_j}\chainedlink{\underline{s}_i}{\noact}\chainedlink{\noact}{\underline{s}_o}\chainend{cxt_{j+1}}.Cxt_{s}, & \mbox{ context doesn't provide $s$.} \\
+\end{array}
+\]
 
 \subsection{Mutating entities} 
-In RS, when an entity is present, it can potentially be involved in each reactions where it is required.
+In RS, when an entity is present, it can potentially be involved in each reaction where it is required.
 With a few more lines of code, in c\CNA~it is possible to describe the behaviour of a mutation of 
@@
 For example, let us assume that entity $s1$ is consumed by reactions $a1$ and $a2$.
-Reaction $a1$ produces also $s1$ if $s2$ is present, otherwise $a1$ produces a mutated version of $s1$, 
+Reaction $a1$ produces $s1$ if $s2$ is present, otherwise $a1$ produces a mutated version of $s1$, 
 say $s1'$.
-When $s1'$ is produced, reaction $a2$ behaves in the same way as if $s1$ would be absent, whereas $a2$
+When $s1'$ is produced, reaction $a1$ behaves in the same way as if $s1$ would be absent, whereas $a2$
 recognises the presence of $s1'$ and behaves in the same way as if $s1$ would be present.
@@
 
-\subsection{Communicating Reaction Systems}
-We sketch how it is possible to program two RSs encodings, in a way that
+\subsection{Communicating reaction systems} 
+We sketch how it is possible to program two RS encodings, in a way that 
 the entities that usually come from context of one RS will be provided instead from the other RS.
@@
 \end{tabular}
-\caption{The two Reaction Systems $rs1$ and $rs2$.}
+\caption{The two reaction systems $rs1$ and $rs2$.}
 \label{tab:rules}
@@
 We need to modify our translation technique to suite this new setting. 
-As we do not model contexts, we introduce  \emph{dummy} channel names $dx$ and $dss$ to model the absence of entities.
-Also, thanks to the simplicity of the example, we can leave out the use of the   $p_i$ channels.
-This streamlining does not affect the programming technique we propose to make two RSs communicate.
+As we do not model contexts, we introduce  \emph{dummy} channel names $dx$ and $dss$ to model the case $x$ and $ss$ are not produced.
+%absence of entities. 
+Also, thanks to the simplicity of the example, we can leave out the use of the   $p_i$ channels. This streamlining does not affect the programming technique we propose to make two RSs communicate.
 %Thus, there will be no $Cxt$ processes and we manage the code in a way that $rs1$ will receive the entity $s1$ from $rs2$. 
@@
 \llbracket a_1\rrbracket \defeq P_{a_1}&\defeq &
+\startchain{\silent}\chainedlink{s_i}{{\color{blue}\noact}}\chainedlink{{\color{blue}\noact}}{s_o}\chainedlink{\tilde{x}_i}{{\color{gray}\noact}}\chainedlink{{\color{gray}\noact}}{\tilde{x}_o}\chainend{a_2}.P_{a_1} +
 \startchain{\silent}\chainedlink{\overline{s}_i}{\noact}\chainedlink{\noact}{\overline{s}_o}    \chainedlink{dx_i}{\noact}\chainedlink{\noact}{dx_o}\chainend{a_2}.P_{a_1}\\
@@
 \[
-\begin{array}{lclcl}
-\llbracket s \rrbracket &\defeq &P_{s} &\defeq & \startchain{{\color{blue}s_i}}\chainedlink{{\color{blue}s_o}}{\noact}\chainedlink{\noact}{{\color{blue}\widehat{s}_i}}\chainend{{\color{blue}\widehat{s}_o}}.P_{s} + \startchain{s_i}\chainend{s_o}.\overline{P_{s}}\\
-&& \overline{P_{s}}&\defeq &\startchain{\overline{s}_i}\chainedlink{\overline{s}_o}{\noact}\chainedlink{\noact}{\widehat{s}_i}\chainend{\widehat{s}_o}.P_{s}+\startchain{\overline{s}_i}\chainend{\overline{s}_i}.\overline{P_{s}}\\
+\begin{array}{lclcl@{\qquad}lclcl}
+\llbracket s \rrbracket &\defeq &P_{s}&\defeq &\startchain{{\color{blue}s_i}}\chainedlink{{\color{blue}s_o}}{\noact}\chainedlink{\noact}{{\color{blue}\hat{s}_i}}\chainend{{\color{blue}\hat{s}_o}}.P_{s} + \startchain{s_i}\chainend{s_o}.\overline{P_{s}}
+&
 %\begin{array}{lclclll}
+\llbracket x \rrbracket &\defeq& P_{x}&\defeq &\startchain{\tilde{x}_i}\chainend{\tilde{x}_o}.P_{x} +  \startchain{dx_i}\chainend{dx_o}.\overline{P_{x}}\\
 %\\
@@
 %\\
+&&\overline{P_{s}}&\defeq &\startchain{\overline{s}_i}\chainedlink{\overline{s}_o}{\noact}\chainedlink{\noact}{\hat{s}_i}\chainend{\hat{s}_o}.P_{s}+\startchain{\overline{s}_i}\chainend{\overline{s}_i}.\overline{P_{s}}
+&
+&& \overline{P_{x}}&\defeq &\startchain{{\color{gray}\tilde{x}_i}}\chainend{{\color{gray}\tilde{x}_o}}.P_{x} +  \startchain{dx_i}\chainend{dx_o}.\overline{P_{x}}\\
 %\end{array}
@@
 \begin{array}{lcl}
+\llbracket a_2\rrbracket \defeq P_{a_2}&\defeq &\startchain{a_2}\chainedlink{y_i}{{\color{gray}\noact}}\chainedlink{{\color{gray}\noact}}{y_o}\chainedlink{\tilde{ss}_i}{{\color{red}\noact}}\chainedlink{{\color{red}\noact}}{\tilde{ss}_o}\chainend{\silent}.P_{a_2}+
 \startchain{a_2}\chainedlink{\overline{y}_i}{\noact}\chainedlink{\noact}{\overline{y}_o}    
@@
 $s$ ($ss$ in $rs2$) to be provided in $rs1$. Every time $ss$ is produced in $rs2$,  a virtual link is
-created to synchronise with $rs1$   on link $\link{\widehat{s}_i}{\widehat{s}_o}$:
+created to synchronise with $rs1$   on link $\link{\hat{s}_i}{\hat{s}_o}$:
 \[
-\begin{array}{lclcl}
-\startchain{dss_i}   \chainend{dss_o}.\overline{P_{ss}} \\
-
+\begin{array}{lclcl@{\qquad} lclcl}
+\llbracket ss \rrbracket& \defeq& P_{ss}&\defeq &\startchain{\tilde{ss}_i}\chainedlink{\hat{s}_i}{\noact}\chainedlink{\noact}{\hat{s}_o}\chainend{\tilde{ss}_o}.P_{ss}+
+\startchain{dss_i}   \chainend{dss_o}.\overline{P_{ss}}
+&
+\llbracket y\rrbracket&\defeq & P_{y}&\defeq &\startchain{{\color{gray}y_i}}\chainend{{\color{gray}y_o}}.\overline{P_{y}} \\
 %\end{array}
@@
 %\begin{array}{lcllcl}
+&&\overline{P_{ss}}&\defeq &\startchain{{\color{red}\tilde{ss}_i}}\chainedlink{{\color{red}\hat{s}_i}}{{\color{blue}\noact}}\chainedlink{{\color{blue}\noact}}{{\color{red}\hat{s}_o}}\chainend{{\color{red}\tilde{ss}_o}}.P_{ss}
 +
-\startchain{dss_i}\chainend{dss_o}.\overline{P_{ss}}\\
-\llbracket y\rrbracket&\defeq & P_{y}&\defeq &\startchain{{\color{gray}y_i}}\chainend{{\color{gray}y_o}}.\overline{P_{y}} \\
+\startchain{dss_i}\chainend{dss_o}.\overline{P_{ss}}
+&
 &&\overline{P_{y}}&\defeq &
@@
 \[
-\chainedlink{{\color{red}\widehat{s}_i}}{{\color{blue}\widehat{s}_i}}
-\chainedlink{{\color{blue}\widehat{s}_o}}{{\color{red}\widehat{s}_o}}
+\startchain{\silent}\chainedlink{s_i}{{\color{blue}s_i}}\chainedlink{{\color{blue}s_o}}{s_o}\chainedlink{\tilde{x}_i}{{\color{gray}\tilde{x}_i}}\chainedlink{{\color{gray}\tilde{x}_o}}{\tilde{x}_o} \chainedlink{a_2}{a_2}\chainedlink{y_i}{ {\color{gray}y_i}}\chainedlink{{\color{gray}y_o}}{y_o}\chainedlink{\tilde{ss}_i}{{\color{red}\tilde{ss}_i}}
+\chainedlink{{\color{red}\hat{s}_i}}{{\color{blue}\hat{s}_i}}
+\chainedlink{{\color{blue}\hat{s}_o}}{{\color{red}\hat{s}_o}}
+\chainedlink{{\color{red}\tilde{ss}_o}}{\tilde{ss}_o}\chainend{\silent},
 \]
@@
 As we have briefly sketched, our model of two 
- \emph{communicating Reaction Systems} can enable the study of the behaviour of one RS in relation to another one. 
-Thus, the products of the reactions of one RS can become the input for another one. 
-This could allow for a modular approach to modeling complex systems, by composing different Reaction Systems.
-
-%which can be modified to represent pathological states as well.
-
-% Again, we stress the fact the verbosity of the {\tt link}-calculus code 
-%should not be
-% seen as a drawback, as it undergoes to automatic code translation.
+\emph{communicating reaction systems} 
+can enable the study of the behaviour of one RS
+in relation to another one.
+% In Section~\ref{ex:lactose} describing  the behaviour of the \emph{lac} operon,
+% the two entities lactose and glucose are controlled 
+% non deterministically by the context.
+% In our framework,
+% instead, by exploiting the expressivity of c\CNA, 
+% the \emph{lac} operon system can be connected with the two systems 
+% producing the lactose and the glucose.
+% This way, the presence of these two entities in the \emph{lac} operon system 
+% could be regulated by realistic mechanisms.
+Thus, the products of the reactions of one RS can become the input for
+another one. This could allow for a modular approach to modeling 
+complex systems, by composing different Reaction Systems.

### 2020-07-27 Catuscia - final/08-conclusion.tex
da dropbox

% !TEX root = ./main_RS2L_LNCS.tex

\section{Conclusion}\label{sec:conclusion}

In this paper we have exploited a variant of the {\tt link}-calculus
where prefixes are link chains and no more single links. This variant was already briefly discussed by the end of~\cite{BBB17}.
This variant allowed us to define 
%shown 
an elegant embedding of reaction systems,
an emerging formalism to model computationally biochemical systems.
%into multilink processes.  
This translation shows several benefits.
For instance, the context behaviour can also be expressed recursively;
entity mutations can be expressed easily; reaction systems can
communicate between them.

We believe that our embedding can contribute to extend the applications
of reaction systems to diverse fields of computer science, and life
sciences.
As we have already mentioned, the evolution of each process resulting from our embedding 
is deterministic, thus we do not have the problem of having infinitely many transitions
in the produced labelled transition system. 
In any case, we can exploit the implementation of  the symbolic semantics of the {\tt link}-calculus~\cite{BrodoO17} that is available at~\cite{tool}.

As future work, we plan to implement a prototype of our framework,
with an automatic translation from RSs to the {\tt link}-calculus. 
We believe that our work can also help to extend the framework
of RSs towards a model which can improve the communication
between different RSs. We also believe that our work can make 
possible to investigate how to apply formal techniques to prove 
properties of the
modeled systems 
\cite{CFHOT15,OCHF16,BBGLBH2017,FOP15,CFOP10}.

\medskip

\noindent
{\small {\bf Acknowledgments}
We thank the anonymous reviewers for their detailed and very useful 
criticisms and recommendations that helped us to improve our paper.}

### 2020-07-27 Catuscia - submitted/00-abstract.tex
da dropbox

% !TEX root =  main_RS2L_LNCS.tex

%Natural computing is an emerging area of research for modeling biological 
%computations by means of computer science formal methods. Reaction systems 
%have been defined in this area as a qualitative abstraction inspired by the 
%functioning of living cells, suitable to model the main
%mechanisms of biochemical reactions.
%Natural computing is an emerging area of research for modeling biological 
%computations by means of computer science formal methods. 
In the area of Natural Computing, reaction systems are
a qualitative abstraction inspired by the 
functioning of living cells, suitable to model the main
mechanisms of biochemical reactions.
This model has already been applied
and extended successfully to various areas of research. Reaction 
systems interact with the environment represented by the context, and
pose problems of implementation, as it is a new computation model. 
In this paper we consider the {\tt link}-calculus, which allows to model 
multiparty interaction in concurrent systems, and show that it allows to
embed reaction systems, by representing the behaviour of each entity and
preserving faithfully their features. 
We show the correctness and completeness 
of our embedding.
We illustrate our framework by showing how to embed a \emph{lac} operon 
regulatory network. 
Finally, our framework can contribute to increase the expressiveness
of reaction systems, by exploiting the interaction among 
different reaction systems.
%entities and the context.

### 2020-07-27 Catuscia - submitted/01-introduction.tex
da dropbox

--- 2020-07-27 Catuscia - final/01-introduction.tex
@@
 fields, such as computer science~\cite{MPR15}, 
-biology~\cite{ABP14,CMMBM12,Az17,BarbutiGLM16}, and
+biology~\cite{ABP14,CMMBM12,Az17,BarbutiGLM16}, 
 molecular chemistry~\cite{OY16}.
@@
 In this paper, we present an encoding from RSs, 
-to the open multiparty process algebra c\CNA\footnote{After `chained Core Network 
-Algebra'.}, a variant of the {\tt link}-calculus~\cite{BodeiBB12,BBB17} without name mobility.
+to the open multiparty process algebra c\CNA,\footnote{After `chained Core Network 
+Algebra'.} a variant of the {\tt link}-calculus~\cite{BodeiBB12,BBB17} without name mobility.
 This formalism allows several processes to synchronise and 
@@
 Our initial motivation for introducing this mechanism was to encode
-Mobile Ambients~\cite{CardelliG00}, obtaining a much stronger operational
-correspondence than any one available in the literature, such as the one in~\cite{B16}.
+Mobile Ambients~\cite{CardelliG00}, getting a much stronger operational
+correspondence than any available in the literature, such as the one in~\cite{B16}.
  This allowed us to easily encode calculi for biology equipped with membranes, as in~\cite{BodeiBBC14}.
@@
 specified in a recursive way as an ordinary process;
-\item 
-%\begin{color}{red}
-our translation results in  a  c\CNA \ system where from
-each state only one transition can be generated, thus the c\CNA \
-computation is fully deterministic as that of the encoded RS;
-%\end{color}
 \item we can express the behaviour of entity mutation,
@@
 communicate; i.e.  a subset of those entities that the context  can
-provide, are then provided by a second RS.
+provide, are then provided by a second RS;
+\item as our translation results in  a  c\CNA \ system, from
+each state only one transition can be generated, thus the c\CNA \
+computation is fully deterministic.
 \end{itemize}
@@
 Section \ref{ex:lactose} shows a more complex example taken from
-the literature on RSs and illustrates a \emph{lac} operon.
+the literature on RSs and illustrating a \emph{lac} operon.
 Section~\ref{sec:discussion} presents some features and 
@@
 Finally, Section~\ref{sec:conclusion}
-discusses future work and concludes.
+discusses future work, and concludes.

### 2020-07-27 Catuscia - submitted/02-ReactionSystem.tex
da dropbox

--- 2020-03-30 02-ReactionSystem.tex
@@
 % modeled by $A$. 
-The set $A$ is the set of \emph{reactions} of $\mathcal{A}$.
+The set $A$ is the set of \emph{reactions} of $A$.
 Since $S$ is finite, so is $A$: 
@@
 \begin{definition}[Reaction Result]
-Let $W$ be a finite subset of $S$.
+Let $T$ be a finite set.
 \begin{enumerate}
-\item Let $a$ be a reaction in $S$. Then $a$ is enabled by $W$, denoted by $\mathit{en}_a(W)$, 
-if $R_a \subseteq W$
-and $I_a \cap W = \emptyset$. The result of $a$ on $W$, denoted by $\mathit{res}_a(W)$, 
+\item Let $a$ be a reaction. Then $a$ is enabled by $T$, denoted by $en_a(T)$, 
+if $R_a \subseteq T$
+and $I_a \cap T = \emptyset$. The result of $a$ on $T$, denoted by $res_a(T)$, 
 is defined by:
-$\mathit{res}_a(W) =P_a$, if $\mathit{en}_a(W)$, and $\mathit{res}_a(W) = \emptyset$ otherwise.
-\item Let $A$ be a finite set of reactions. The result of $A$ on $W$, denoted 
-by $\mathit{res}_A(W)$,  is defined
-by: $\mathit{res}_A(W) = \bigcup_{a \in A} \mathit{res}_a(W)$.
+$res_a(T) =P_a$, if $en_a(T)$, and $res_a(T) = \emptyset$ otherwise.
+\item Let $A$ be a finite set of reactions. The result of $A$ on $T$, denoted 
+by $res_A(T)$,  is defined
+by: $res_A(T) = \bigcup_{a \in A} res_a(T)$.
 \end{enumerate}
@@
 %, $\delta= D_0, \dots D_n$.
-Then, we let $\pi^\infty= (\gamma^\infty,\delta^\infty)$ be the extended interactive process of  $\pi$,
-defined as $  \gamma^\infty=\{C_i'\}_{i\in\mathbb{N}}$, $\delta^\infty=\{D_i'\}_{i\in\mathbb{N}}$, where:
-$$
-C'_j = \left\{
-\begin{array}{ll}
-C_j & \mbox{if $j\in[0,n]$}\\
-\emptyset & \mbox{otherwise ($j>n$)}
-\end{array}
-\right.
-\qquad
-D'_j = \left\{
-\begin{array}{ll}
-D_0 & \mbox{if $j=0$}\\
-\mathit{res}_A(D_{j-1}' \cup C_{j-1}') & \mbox{otherwise ($j\geq 1$)}
-\end{array}
-\right.
-$$
-% $C_j' = C_j$ for $j\in[0,n]$ and $C_j' =\emptyset$ for $j>n$,   $D_0' = D_0$ and  $D_j' = \mathit{res}_A(D_{j-1}' \cup C_{j-1}')$ for $j\geq 1$.
+Then, let $\pi'= (\gamma',\delta')$ be the extended interactive process of  $\pi =(\gamma,\delta)$,
+defined as $  \gamma'=\{C_i'\}_{i\in\mathbb{N}}$, $\delta'=\{D_i'\}_{i\in\mathbb{N}}$, where
+ $C_j' = C_j$ for $j\in[0,n]$ and $C_j' =\emptyset$ for $j>n$,   $D_0' = D_0$ and  $D_j' = res_A(D_{j-1}' \cup C_{j-1}')$ for $j\geq 1$.
 \end{definition}

### 2020-07-27 Catuscia - submitted/04-CCNA.tex
da dropbox

--- 2020-04-02 04-CCNA.tex
@@
 
-\paragraph{Link Chains}
+\paragraph{Link Chains.}
 Let $\mathcal{C}$ be the set of channels, ranged over by $a,b,...$, and 
@@
 The syntax of c\CNA\  extends that of \CNA~\cite{BBB17} by allowing to use link chains as prefixes instead of links.
-For the rest it features nondeterministic (guarded) choice, parallel composition, restriction, relabelling and possibly recursive definitions.
+For the rest it features nondeterministic choice, parallel composition, restriction, relabelling and possibly recursive definitions.
 Here we do not consider name mobility, which is present instead in the {\tt link}-calculus.
-
-\paragraph{Semantics}
+\paragraph{Semantics.}
 
@@
 synchronises interactions (it cannot be applied when $\merges{v}{v'}$ is not defined).
-
-\begin{example}
-As some simple examples, consider the recursive process definitions $H \defeq \link{a}{b}.\link{a}{c}.H$ and $K \defeq \link{a}{b}.K + \link{a}{c}.K$: the former is some sort of forwarder that alternates between providing a link from $a$ to $b$ and to $c$; the latter is a nondeterministic forwarder that at each step can link $a$ to $b$ or to $c$.
-\end{example}
 
@@
 \begin{prooftree} 
-\AxiomC{$\upsilon \blackstretcheq \upsilon_j\quad j\in I$} %\mbox{($\ell$ {\em only} solid link in $s$)}
+\AxiomC{$v \blackstretcheq v_j\quad j\in I$} %\mbox{($\ell$ {\em only} solid link in $s$)}
 \RightLabel{\scriptsize(\textit{Sum})}
-\UnaryInfC{$\sum_{i\in I} \upsilon_i.P_i \xrightarrow{\upsilon} P_j$} 
+\UnaryInfC{$\sum_{i\in I} \upsilon_i.P_i \xrightarrow{v} P_j$} 
 \DisplayProof
 \
-\AxiomC{$P \xrightarrow{\upsilon} P'$}
+\AxiomC{$P \xrightarrow{v} P'$}
  \AxiomC{$(A \defeq P)\in\Delta$}
 \RightLabel{\scriptsize(\textit{Ide})}
-\BinaryInfC{$A \xrightarrow{\upsilon} P'$} 
+\BinaryInfC{$A \xrightarrow{v} P'$} 
 %\AxiomC{$P \xrightarrow{v} P'$} 
@@
 \begin{prooftree} 
-\AxiomC{$P \xrightarrow{\upsilon} P'$} 
+\AxiomC{$P \xrightarrow{v} P'$} 
 \RightLabel{\scriptsize(\textit{Res})}
-\UnaryInfC{$\restrict{a}P \xrightarrow{\restrict{a}\upsilon} \restrict{a}P'$} 
+\UnaryInfC{$\restrict{a}P \xrightarrow{\restrict{a}v} \restrict{a}P'$} 
 \DisplayProof
@@
 \RightLabel{\scriptsize(\textit{Lpar})}
-\UnaryInfC{$P|Q \xrightarrow{\upsilon} P'|Q$} 
+\UnaryInfC{$P|Q \xrightarrow{v} P'|Q$} 
 \DisplayProof
@@
 For example,  the expression $(\startchain{a}\chainedlink{b}{\noact}\chainend{\noact})^3$  denotes the chain $\startchain{a}\chainedlink{b}{\noact}\chainedlink{\noact}{a}\chainedlink{b}{\noact}\chainedlink{\noact}{a}\chainedlink{b}{\noact}\chainend{\noact}$.
-Instead the expression $(\link{a}{b})^2$ is ill-defined because $a$ does not match with $b$.
 
-Then, we introduce the notation for \emph{half links} that will be used in conjunction with the \emph{open block of chain} to form regular link chains.
+ \noindent
+We introduce the \emph{half link} that will be used in conjunction with the \emph{open block of chain} to form regular link chains.
 
 \begin{definition}[Half links]
-Let $a$ be a channel name, we define the \emph{half left link}  $\link{a}{}$,   and the  \emph{half right link} $\link{}{a}$.
+Let $a$ be a channel name, we define the \emph{half left link}:  $\link{a}{}$,   and the  \emph{half right link}: $\link{}{a}$.
 \end{definition}
@@
  For example, for $R = \{a,b\}$ the expression $ \blockchain{c_i}{c_o}{c \in R} $\;
-denotes the block of chains ${}_{a_i}^{\noact}\chainedlink{\noact}{a_o} \backslash{}_{b_i}^{\noact}\chainedlink{\noact}{b_o}$; and the expression $\link{r_1}{} \blockchain{c_i}{c_o}{c \in R}\;\link{}{r_2} $ denotes the chain \\
-$ \startchain{r_1}\chainedlink{a_i}{\noact}\chainedlink{\noact}{a_o}\chainedlink{b_i}{\noact}\chainedlink{\noact}{b_o}\chainend{r_2}$.
+denotes the block of chains ${}_{a_i}^{\noact}\chainedlink{\noact}{a_o} \backslash{}_{b_i}^{\noact}\chainedlink{\noact}{b_o}$; and the expression $\link{r_1}{} \blockchain{c_i}{c_o}{c \in R}\;\link{}{r_2} $ denotes the chain $ \startchain{r_1}\chainedlink{a_i}{\noact}\chainedlink{\noact}{a_o}\chainedlink{b_i}{\noact}\chainedlink{\noact}{b_o}\chainend{r_2}$.

### 2020-07-27 Catuscia - submitted/05-translation.tex
da dropbox

--- 2020-07-27 Catuscia - final/05-translation.tex
@@
 names $s_i$, $s_o$ are used to test the presence of $s$ in the system; names $\hat{s}_i$, $\hat{s}_o$ are used to test the provisioning of $s$ from the context;
-names $\tilde{s}_i$, $\tilde{s}_o$  are used to test the production of $s$ by some reaction; names $\overline{s}_i$, $\overline{s}_o$ are used to test the absence of $s$  in the system; and names $\underline{s}_i$, $\underline{s}_o$ are used to test the absence of $s$ from the context.
-We let $P_s$ be the process implementing the  presence of $s$ in the system, and $\overline{P_s}$ its absence.
-%\begin{color}{red}
-They are defined below:
+names $\tilde{s}_i$, $\tilde{s}_o$  are used to test the production of $s$ by some reaction; names $\overline{s}_i$, $\overline{s}_o$ are used to test the absence of $s$ from the context; and names $\underline{s}_i$, $\underline{s}_o$ are used to test the absence of $s$ in the system.
+We let $P_s$ be the process implementing the  presence of $s$ in the system, and $\overline{P_s}$ be the one for its absence.
+They can be seen as instances of the same template, which is given below.
 
-%\[
-%\begin{array}{lcl@{\hskip 1cm}lcl}
-% P_s &\defeq & P(s,\tilde{s},\hat{s},\underline{s}) 
-% &
-% \overline{P_s}& \defeq & P(\overline{s},\tilde{s},\hat{s},\underline{s})
-%\end{array}
-%\]
+\[
+\begin{array}{lcl@{\hskip 1cm}lcl}
+ P_s &\defeq & P(s,\tilde{s},\hat{s},\underline{s}) 
+ &
+ \overline{P_s}& \defeq & P(\overline{s},\tilde{s},\hat{s},\underline{s})
+\end{array}
+\]
 %{\color{red}
@@
 \begin{array}{rcl}
-% P(s,\tilde{s},\hat{s},\underline{s}) 
- P_s & \defeq &\sum_{h\geq 0,k\geq 0}  
+ P(s,\tilde{s},\hat{s},\underline{s}) & \defeq &\sum_{h,k\geq 0}  
   (\startchain{s_i}\chainedlink{s_o}{\noact}\chainend{\noact})^h \ \startchain{\hat{s}_i}\chainedlink{\hat{s}_o}{\noact}\chainend{\noact} (\startchain{\tilde{s}_i}\chainedlink{\tilde{s}_o}{\noact}\chainend{\noact})^k.P_s\\
@@
 &&\sum_{h\geq 0}(\startchain{s_i}\chainedlink{s_o}{\noact}\chainend{\noact})^h \ \link{\underline{s}_i}{\underline{s}_o}.\overline{P_s}
-\\
-\\
-%P(\overline{s},\tilde{s},\hat{s},\underline{s}) 
-\overline{P_s}& \defeq &\sum_{h\geq 0,k\geq 0}  
-  (\startchain{\overline{s}_i}\chainedlink{\overline{s}_o}{\noact}\chainend{\noact})^h \ \startchain{\hat{s}_i}\chainedlink{\hat{s}_o}{\noact}\chainend{\noact} (\startchain{\tilde{s}_i}\chainedlink{\tilde{s}_o}{\noact}\chainend{\noact})^k.P_s\\
-&&\ + \\
-&&
-\sum_{h\geq 0,k\geq 1}
- (\startchain{\overline{s}_i}\chainedlink{\overline{s}_o}{\noact}\chainend{\noact})^h \ \startchain{\underline{s}_i}\chainedlink{\underline{s}_o}{\noact}\chainend{\noact}(\startchain{\tilde{s}_i}\chainedlink{\tilde{s}_o}{\noact}\chainend{\noact})^k.P_s\\
-&&\ +\\
-&&\sum_{h\geq 0}(\startchain{\overline{s}_i}\chainedlink{\overline{s}_o}{\noact}\chainend{\noact})^h \ \link{\underline{s}_i}{\underline{s}_o}.\overline{P_s}
 \end{array}
 \]
-%\end{color}
-
-The first line of $P_s$ accounts for the case where $s$ is tested for presence by $h$ reactions and produced by $k$ reactions, while being provided by the context ($\link{\hat{s}_i}{\hat{s}_o}$).
+The first line of $ P(s,\tilde{s},\hat{s},\underline{s})$ accounts for the case where $s$ is tested for presence by $h$ reactions and produced by $k$ reactions, while being provided by the context ($\link{\hat{s}_i}{\hat{s}_o}$).
 Thus, $s$ will be present at the next step (the continuation is $P_s$). Here $h$ and $k$ are not known a priori and therefore any
@@
 %$h = \sum_{j\geq0} h_j +h'_j$, $k = (\sum_{j \geq 0} k_j )  + 1$ guarantees that $s$ will be present in the next step. Please note that 
-The third line accounts for the case where $s$ is tested for presence by $h$ reactions, but it is neither produced nor provided by  the context.
+The third line accounts for the case where $s$ is tested for presence, but it is neither produced nor provided by  the context.
 Therefore, in the next step $s$ will be absent in the system (the continuation is $\overline{P_s}$).
 Note that in the case of $\overline{P_s}$ the test for presence of $s$ in the system is just replaced by the test for its absence.
-As before, $s$ will be present again in the system if $s$ will be produced (rows $1$ and $2$ in  $\overline{P_s}$ code), or if the context will provide it (row $1$ in $\overline{P_s}$ code).
 %Process $P_s$ (and $\overline{P_s}$) offers link chain prefixes for all the possible
@@
 We assume that each reaction $a$ is assigned a progressive number $j$.
-The process for reaction $a_j=(R_j,I_j,P_j)$ must assert either the possibility to apply the reaction or its impossibility.
+The process for reaction $aj=(R_j,I_j,P_j)$ must assert either the possibility to apply the reaction or its impossibility.
 The first case happens when all its reactants are present (the link $\link{s_i}{s_o}$ is requested for any $s\in R_j$) and all its inhibitors are absent (the link $\link{\overline{e}_i}{\overline{e}_o}$ is requested for any $e\in I_j$), then the product set is released (the link $\link{\tilde{c}_i}{\tilde{c}_o}$ is requested for any $c\in P_j$).
@@
 \begin{array}{lcl@{\quad}l}
-P_{a_j}& \defeq & 
- ^{r_j} \backslash  \blockchain{s_i}{s_o}{s \in R_j} \, \link{}{} \blockchain{\overline{e}_i}{\overline{e}_o}{e \in I_j} \, \chainedlink{r_{j+1}}{\noact}\chainedlink{\noact}{p_j}\link{}{} \blockchain{\tilde{c}_i}{\tilde{c}_o}{c \in P_j}\, \backslash_{p_{j+1}}.P_{a_j} & \mbox{\small  \{$a_j$ is applicable\}}\\
+P_{aj}& \defeq & 
+ ^{r_j} \backslash  \blockchain{s_i}{s_o}{s \in R_j} \, \link{}{} \blockchain{\overline{e}_i}{\overline{e}_o}{e \in I_j} \, \chainedlink{r_{j+1}}{\noact}\chainedlink{\noact}{p_j}\link{}{} \blockchain{\tilde{c}_i}{\tilde{c}_o}{c \in P_j}\, \backslash_{p_{j+1}}.P_{aj} & \mbox{\small  \{$aj$ is applicable\}}\\
 && + \ \\
-&& \sum_{s\in R_j}\startchain{r_j}\chainedlink{\overline{s}_i}{\noact}\chainedlink{\noact}{\overline{s}_o} \chainedlink{r_{j+1}}{\noact}\chainedlink{\noact}{p_{j}}\chainend{p_{j+1}}.P_{a_j}  & \mbox{\small  \{$a_j$ is not applicable\}}\\
+&& \sum_{s\in R_j}\startchain{r_j}\chainedlink{\overline{s}_i}{\noact}\chainedlink{\noact}{\overline{s}_o} \chainedlink{r_{j+1}}{\noact}\chainedlink{\noact}{p_{j}}\chainend{p_{j+1}}.P_{aj}  & \mbox{\small  \{$aj$ is not applicable\}}\\
 &&+\\
-&& \sum_{e\in I_j}\startchain{r_j}\chainedlink{e_i}{\noact} \chainedlink{\noact}{e_o} \chainedlink{r_{i+1}}{\noact}\chainedlink{\noact}{p_{j}}\chainend{p_{j+1}}.P_{a_j}  & \mbox{\small \{$a_j$ is not applicable\}}\\
+&& \sum_{e\in I_j}\startchain{r_j}\chainedlink{e_i}{\noact} \chainedlink{\noact}{e_o} \chainedlink{r_{i+1}}{\noact}\chainedlink{\noact}{p_{j}}\chainend{p_{j+1}}.P_{aj}  & \mbox{\small \{$aj$ is not applicable\}}\\
 \end{array}
@@
 \noindent
-%\begin{color}{red}
-Formally speaking, the definition of the case when $a_j$ is applicable (row 1) requires $R_j$, $I_j$ and $P_j$ to be totally ordered, because open block expressions are used.
-It is worth noting that the chosen orders are inessential for exploiting $P_{a_j}$ in the encoding of a RS, but they are needed to disambiguate its definition.
-Without loss of generality, we  assume that all the entities are enumerated, likewise reactions, so that $R_j$, $I_j$ and $P_j$ inherit the same order. 
-%\end{color}
-%
 We exploit names $r_j,p_j$ to join the chains provided by the application of all the reactions. 
-Channels $r_j$ and $r_{j+1}$ enclose the enabling/disabling condition of reaction $a_j$.
-Channels $p_i$ and $p_{j+1}$ enclose the links related to the entities produced by $a_j$.
+Channels $r_j$ and $r_{j+1}$ enclose the enabling/disabling condition of reaction $aj$.
+Channels $p_i$ and $p_{j+1}$ enclose the links related to the entities produced by $aj$.
 We will see that all the link chain labels of transitions follow the same schema: first we find all the reactions limited to the reactants and inhibitors (chained using $r_j$ channels), then all the supplies by the contexts (chained using $cxt_j$ channels, to be introduced next), and finally the products for all the reactions (chained using $p_j$ channels).
@@
 \paragraph{Processes for contexts.}
-For each entity $s \in S$, we introduce another process $Cxt_s$,  participating in each transition and determining whether the entity $s$ is provided by the context or not.
-As already said, we assume that entities are enumerated and use the names $cxt_j$ to concatenate
+For each entity $s \in S$, we introduce another process $Cxt_s$,  participating in each transition and saying if, the entity $s$ is provided by the context or not.
+As done for the reactions, we assume that entities are enumerated and use the names $cxt_j$ to concatenate
 the chains formed by the application of all the contexts. 
@@
 \]
-%\begin{color}{red}
-We only consider $Cxt_s^n$ with $n >0$, as the entities such that $s \in C_0$ (resp.\ $s\not\in C_0$) are modeled by the $P_s$ processes (resp.\ $\overline{P_s}$) in the initial c\CNA  \ process. 
-%Differently from RSs, where at each step the context can non deterministically provide entities, here the behaviour of the context is defined a priori.
-%Nevertheless, 
-The intrinsic modularity of c\CNA \ allows us to run different context behaviours
-%change the context behaviour leaving unchanged the rest of the system. Thus  can be test
- for the same system by changing the definitions of $Cxt_s^n$.
-%present at step zero  are considered to be present in the initial system (process $P_s$ instead of $\overline{P_s}$).
-%\end{color}
+
+We only consider $Cxt_s^n$ with $n >0$, as the entities $s$ that are present at step zero  are considered to be present in the initial system (process $P_s$ instead of $\overline{P_s}$).
 
@@
 We define its c\CNA~translation $\llbracket{\cal A},\gamma \rrbracket$ as follows: 
-\begin{center}$\llbracket{\cal A},\gamma \rrbracket = \restrict{reacts,cxts,ents,prods}(\Pi_{s\in C_0}P_{s} | \Pi_{s \notin C_0}\overline{P_s} | \Pi_{a\in A} P_a| \Pi_{s\in S} Cxt_s)$,
+\begin{center}$\llbracket{\cal A},\gamma \rrbracket = \restrict{reacts,ctxs,ents,prods}(\Pi_{s\in C_0}P_{s} | \Pi_{s \notin C_0}\overline{P_s} | \Pi_{a\in A} P_a| \Pi_{s\in S} Cxt_s)$,
 \end{center} 
 with  $reacts$ be the set of reaction names $r_j $, $cxts$ the set of context names $cxt_j$,  $ents$ the set of decorated entity names  $\{s_i,s_o,\hat{s}_i,\hat{s}_o,\tilde{s}_i,\tilde{s}_o,\overline{s}_i,\overline{s}_o,\underline{s}_i,\underline{s}_o| s \in S\}$, and  $prods$ be the set of names $p_j$ associated to each reaction.
-In the following, we set $names = reacts\, \cup\, cxts\, \cup\, ents\, \cup \, prods$.
+In the following, we set $names = reacts\, \cup\, ctxs\, \cup\, ents\, \cup \, prods$.
 %For notational convenience, we fix that $r_1 = \silent$,  $r_{u+1} = cxt_1$ for $u$ the number of reacts, and $cxt_{w+1} = \silent$ for $w$ the number of entities.
@@
 It is important to observe that,   for each transition, our c\CNA \ encoding requires  all the processes $P_a$, with $a \in A$, and  $Cxt_s$ and  $P_s$, with $s \in S$,  be interacting in that transition.
-This is due to the fact that  all the channels $r_j$, $p_j$, $cxt_j$, and $s_{i}$, and $s_{o}$ are  restricted.
+This is due to the fact that  all the channels $r_j$, $p_j$, $cxt_h$, and $s_{hi}$, and $s_{ho}$ are  restricted.
 Each reaction defines a pattern to be satisfied, i.e. each reaction inserts as many virtual links  as  the number of reactants, inhibitors, and products, as required by the corresponding reaction.
@@
 Let $P=\llbracket{\cal A},\gamma\rrbracket$ its c\CNA~translation.
-If there $exists$ $P'$ such that $t= (P \xrightarrow{\restrict{names}\upsilon} P')$ is a transition of $P$, then
+If $exists$ $P'$ such that $t= (P \xrightarrow{\restrict{names}\upsilon} P')$ is a transition of $P$, then
   \begin{enumerate}
@@
   %each process $P_s $ (or its alternative version $\overline{P}_s$), with $s \in S$, participates at most  with one solid link for each process $P_a$, and with exactly one solid link for each process $Cxt_s$.%, if $s$ is a \emph{required element} of $A_j$ in $t$.
+
   \end{enumerate}
@@
  \begin{example}\label{ex:backbone}
-%\begin{color}{red}
-Let ${\cal A}$ be a RS whose specification contains two entities, $s1$ and $s2$, and, among the others, the reaction $a=(\{s1\},\{\dots\} , \{s1\})$ that guarantees the persistence of the entity $s1$ once it is present in the system.
-Note that we use here a dummy inhibitor $(\dots)$ which will never be present.
+ Let ${\cal A}$ be a RS whose specification contains two entities, $s1$ and $s2$, and, among the others, the reaction $a=(s1,\ , s1)$ that guarantees the presence of the  $s1$ in the system.
 Then, we assume an  extended interactive process $\pi=(\gamma,\delta)$ where the context $\gamma$ provides  $s1$ and $s2$.
  %, for the first $n$ steps.
- Our translation includes the processes:
- %\end{color}
+ Our translation includes the processes:\\
  \[
@@
 Moreover, we denote by $\pi^j$ the shift of $\pi$ starting at the $j$-th state sequence; formally we let $\pi^j=(\gamma^j,\delta^j)$ with
-$\gamma^j=\{C_i'\}_{i\in\mathbb{N}}$, $\delta^j=\{D_i'\}_{i\in\mathbb{N}}$ such that 
-$C_0'=C_j \cup D_j$, 
-%\begin{color}{red}
-$D_0' = \emptyset$,
-%\end{color}
-and   $C_i'=C_{i+j}$, $D_i'= D_{i+j}$ for any $i\geq 1$.
+$\gamma^j=\{C_i'\}_{i\in\mathbb{N}}$, $\delta^j=\{D_i'\}_{i\in\mathbb{N}}$ with 
+  $C_0'=C_j \cup D_j$, and   $C_i'=C_{i+j}$, $D_i'= D_{i+j}$ for any $i\geq 1$.
 
-%\begin{color}{red}
  \begin{restatable}[Correctness 1]{proposition}{propcorruno}
+% \begin{proposition}(correctness 1)\\
  \label{prop:corr1}
+% We define $\pi' = (\gamma',\delta')$ with $\gamma' = C_0',\dots,C_{n-1}'$ and $\delta' = D_0',\dots,D_{n-1}'$ as $C_0' = C_1 \cup D_1$,  and
+% $C_i' = C_{i+1}$, $D_{i}' = D_{i+1}$ with $ 1 \leq i < n$.
  Let 
- $P=\llbracket{\cal A},\gamma\rrbracket$ with
-$$P  = \restrict{names} (\Pi_{a \in A}P_a| \Pi_{s \in S}Cxt_s^1 | \Pi_{s \in C_0}P_s| \Pi_{s \notin C_0}\overline{P}_s).$$
- If $P \xrightarrow{v}Q$, then
-  $v$ is a silent action and $Q = \llbracket {\cal A},\gamma^1\rrbracket$, namely
-$$Q= \restrict{names} (\Pi_{a \in A}P_a| \Pi_{s \in S}Cxt_s^2 | \Pi_{s \in C_1 \cup D_1}P_s| \Pi_{s \notin C_1 \cup D_1}\overline{P}_s).$$
+ $P=\llbracket{\cal A},\gamma\rrbracket$ with\\
+$P  = \restrict{names} (\Pi_{a \in A}P_a| \Pi_{s \in S}Cxt_s | \Pi_{s \in C_0}P_s| \Pi_{s \notin C_0}\overline{P}_s)$.\\
+ If there exists  $P'$  such that $P \xrightarrow{v}P'$,  it holds that:\\
+  $v=\link{\silent}{\silent}\dots\link{\silent}{\silent}$ and \\
+ $P'= \restrict{names} (\Pi_{a \in A}P_a| \Pi_{s \in S}Cxt_s | \Pi_{s \in C_1 \cup D_1}P_s| \Pi_{s \notin C_1 \cup D_1}\overline{P}_s)$.\\
+Moreover, given $\pi^1=(\gamma^1,\delta^1)$, we have $P' = \llbracket {\cal A},\gamma^1\rrbracket$.
+ %\end{proposition}
 \end{restatable}
 
+% \begin{lemma}[correctness 1 {\color{red} Ma serve ??}]
+% \label{cor:complet1}
+% Let ${\cal A} = (S, A)$ be a rs, and let  $\pi=(\gamma,\delta)$ be an extended interactive process in $A$, with
+% $\gamma = C_0,\dots,C_t$, $\delta = D_0,\dots,D_t$, with $n \geq 0$.
+Let   $P=[|{\cal A},\gamma|]$ be  its c\CNA~translation.
+We define $\pi' = (\gamma',\delta')$ with $\gamma' = C_0',\dots,C_{n-1}'$ and $\delta' = D_0',\dots,D_{n-1}'$ as $C_0' = C_1 \cup D_1$,  and
+% $C_i' = C_{i+1}$, $D_{i}' = D_{i+1}$ with $ 1 \leq i < n$. Then,
+% if $\exists$ $P'$, $P''$ such that $P \xrightarrow{\link{\silent}{\silent}} P'$ and $P \xrightarrow{\link{\silent}{\silent}} P''$, then $P' = P''$.
+%\end{lemma}
+
+% \begin{proposition}[correctness2]
+% \label{prop:correctness1}
+% Let $P$ be a c\CNA \ process such that $\exists$ a $rs$ ${\cal A}$ and a $n$-step interactive process $\pi=(\gamma,\delta)$ with $\gamma =C_0,\dots,C_n$ and $\delta = D_0,\dots,D_n$ such that $P= [|{\cal A},\gamma|]$.
+% %We set $\pi'=(\gamma',\delta')$ with $\gamma'=C_0',\dots,C_{n-1}'$,  $\delta'=D_0',\dots,D_{n-1}$,  where $C_0'= C_1 \cup D_1$, $C_i'= C_{i+1}$, $D_i=D_{i+1}'$,  and $1 \leq i < n$.
+%Then,   $\exists$ $P'$ and $\pi'=(\gamma',\delta')$ such that $P \xrightarrow{\link{\silent}{\silent}}P'$, $P' =[|{\cal A},\gamma'|]$.
+% \end{proposition} 
+\noindent
 Now, we extend the previous result to a series of transitions.
@@
  \begin{restatable}[Correctness 2]{corollary}{corrcorrdue} 
+% \begin{corollary}[correcteness2]
  \label{corr:corr2}
   Let $P = \llbracket{\cal A},\gamma\rrbracket$ and $j\geq 1$.
-If there exists  $Q$ such that $P \rightarrow^j Q$, then 
-$Q =\llbracket{\cal A},\gamma^j\rrbracket$.
+%  Let $P$ a  c\CNA process such that exists a rs ${\cal A}$ and an extended  interactive process $\pi=(\gamma,\delta)$ with $\gamma =C_0,\dots,C_t$, $\delta = D_0,\dots,D_t$, with $t \geq 0$, such that $P= [|{\cal A},\gamma|]$.
+If there exists  $P''$ such that $P \xrightarrow{\link{\silent}{\silent}\dots \link{\silent}{\silent}}^{j}P''$, then letting  $\pi^j=(\gamma^j,\delta^j)$ we have
+$P'' =\llbracket{\cal A},\gamma^j\rrbracket$.
+%such that $P=[|{\cal A}|]$, then if $P \rightarrow^* P'$, then $\exists$ ${\cal A'}$ such that ${\cal A}$  evolves in ${\cal A'}$ and $P'=[|{\cal A'}|]$.
  \end{restatable}
 
-Conversely, we prove that the $c\CNA$ process $\llbracket{\cal A},\gamma\rrbracket$ can simulate all the evolutions  of the underlying interactive process.
+\noindent
+With the following propositions, we prove that, given a RS ${\cal A} = (S, A)$ and an extended  interactive process $\pi=(\gamma,\delta)$, then the $c\CNA$ process $\llbracket{\cal A},\gamma\rrbracket$ can simulate all the evolutions  of $\pi$.
 
  \begin{restatable}[Completeness 1]{proposition}{propcompluno} 
+% \begin{corollary}[correcteness2]
  \label{prop:compl1}
-  $\llbracket{\cal A},\gamma\rrbracket \rightarrow \llbracket{\cal A},\gamma^1\rrbracket$.
-
+% \begin{proposition}[completeness1]
+ %\label{lem:compl2}
+Let   $P=\llbracket{\cal A},\gamma\rrbracket$ and $\pi^1 = (\gamma^1,\delta^1)$.
+% be the extended interactive process starting  at the next state sequence. 
+Then, 
+ % following holds:
+ %\begin{enumerate} 
+ %\item  
+  $P \xrightarrow{\link{\silent}{\silent}\dots \link{\silent}{\silent}} P'=\llbracket{\cal A},\gamma^1\rrbracket$.
+ %\item if $\exists$ $P'$, $P''$ such that $P \xrightarrow{\link{\silent}{\silent}} P'$ and $P \xrightarrow{\link{\silent}{\silent}} P''$, then $P' = P''$.
+ %\end{enumerate}
+% \end{proposition} 
 \end{restatable}
 
+\noindent
 Now, we extend the previous result to a series of transitions.
@@
 \begin{restatable}[Completeness 2]{corollary}{corrcompldue} 
+% \begin{corollary}[correcteness2]
  \label{corr:compl2}
-$\llbracket{\cal A},\gamma\rrbracket \rightarrow^j \llbracket{\cal A},\gamma^j\rrbracket$.  
+% \begin{corollary}[completeness2]
+ %\label{prop:completeness}
+Let $P=\llbracket{\cal A},\gamma\rrbracket$ and  $\pi^j= (\gamma^j,\delta^j)$.  
+Then, $P \xrightarrow{\link{\silent}{\silent}\dots\link{\silent}{\silent}}^j P''=\llbracket{\cal A},\gamma^j\rrbracket$.  
  \end{restatable}
-%\end{color}
 

### 2020-07-27 Catuscia - submitted/07-discussion.tex
da dropbox

--- 2020-07-27 Catuscia - final/07-discussion.tex
@@
 the behaviour of the context, the possibility of alternative behaviour of
-mutated entities and the communication between two different reaction systems.
+mutated entities, the communication between two different reaction systems.
 It is important to note that our encoding guarantees that from each state, 
@@
 \subsection{Mutating entities} 
-In RS, when an entity is present, it can potentially be involved in each reaction where it is required.
+In RS, when an entity is present, it can potentially be involved in each reactions where it is required.
 With a few more lines of code, in c\CNA~it is possible to describe the behaviour of a mutation of 
@@
 For example, let us assume that entity $s1$ is consumed by reactions $a1$ and $a2$.
-Reaction $a1$ produces $s1$ if $s2$ is present, otherwise $a1$ produces a mutated version of $s1$, 
+Reaction $a1$ produces also $s1$ if $s2$ is present, otherwise $a1$ produces a mutated version of $s1$, 
 say $s1'$.
-When $s1'$ is produced, reaction $a1$ behaves in the same way as if $s1$ would be absent, whereas $a2$
+When $s1'$ is produced, reaction $a2$ behaves in the same way as if $s1$ would be absent, whereas $a2$
 recognises the presence of $s1'$ and behaves in the same way as if $s1$ would be present.
@@
 \subsection{Communicating reaction systems} 
-We sketch how it is possible to program two RS encodings, in a way that 
+We sketch how it is possible to program two RSs encodings, in a way that 
 the entities that usually come from context of one RS will be provided instead from the other RS.
@@
 We need to modify our translation technique to suite this new setting. 
-As we do not model contexts, we introduce  \emph{dummy} channel names $dx$ and $dss$ to model the case $x$ and $ss$ are not produced.
-%absence of entities. 
-Also, thanks to the simplicity of the example, we can leave out the use of the   $p_i$ channels. This streamlining does not affect the programming technique we propose to make two RSs communicate.
+As we do not model contexts, we introduce  \emph{dummy} channel names $dx$ and $dss$ to model the absence of entities.
+Also, thanks to the simplicity of the example, we can leave out the use of the   $p_i$ channels.
+This streamlining does not affect the programming technique we propose to make two RSs communicate.
 %Thus, there will be no $Cxt$ processes and we manage the code in a way that $rs1$ will receive the entity $s1$ from $rs2$. 
@@
 
-As we have briefly sketched, our model of two 
-\emph{communicating reaction systems} 
+As we have briefly sketched, our model of two \emph{communicating reaction systems} 
 can enable the study of the behaviour of one RS
 in relation to another one.
-% In Section~\ref{ex:lactose} describing  the behaviour of the \emph{lac} operon,
-% the two entities lactose and glucose are controlled 
-% non deterministically by the context.
-% In our framework,
-% instead, by exploiting the expressivity of c\CNA, 
-% the \emph{lac} operon system can be connected with the two systems 
-% producing the lactose and the glucose.
-% This way, the presence of these two entities in the \emph{lac} operon system 
-% could be regulated by realistic mechanisms.
-Thus, the products of the reactions of one RS can become the input for
-another one. This could allow for a modular approach to modeling 
-complex systems, by composing different Reaction Systems.
+In Example~\ref{ex:twoRS} describing  the behaviour of the \emph{lac} operon,
+the two entities lactose and glucose are controlled 
+non deterministically by the context.
+In our framework,
+instead, by exploiting the expressivity of c\CNA, 
+the \emph{lac} operon system has been connected with the two systems 
+producing the lactose and the glucose.
+This way, the presence of these two entities in the \emph{lac} operon system 
+can be regulated by realistic mechanisms.
+%which can be modified to represent pathological states as well.
+
+% Again, we stress the fact the verbosity of the {\tt link}-calculus code 
+%should not be
+% seen as a drawback, as it undergoes to automatic code translation.

### 2020-07-27 Catuscia - submitted/08-conclusion.tex
da dropbox

--- 2020-07-27 Catuscia - final/08-conclusion.tex
@@
 
-In this paper we have exploited a variant of the {\tt link}-calculus
-where prefixes are link chains and no more single links. This variant was already briefly discussed by the end of~\cite{BBB17}.
+In this paper we have introduced a variant of the {\tt link}-calculus
+where prefixes are link chains and no more single links, as it was 
+briefly described in the future work section in~\cite{BBB17}.
 This variant allowed us to define 
@@
 This translation shows several benefits.
-For instance, the context behaviour can also be expressed recursively;
-entity mutations can be expressed easily; reaction systems can
+For instance, the context behaviour can also be expressed recursively. 
+Entity mutations can be expressed easily. Reaction systems can
 communicate between them.
@@
 is deterministic, thus we do not have the problem of having infinitely many transitions
-in the produced labelled transition system. 
-In any case, we can exploit the implementation of  the symbolic semantics of the {\tt link}-calculus~\cite{BrodoO17} that is available at~\cite{tool}.
+in the produced labelled transition system. In any case, we can exploit the implementation of  the symbolic semantics of the {\tt link}-calculus~\cite{BrodoO17} that can be found in~\cite{tool}.
 
 As future work, we plan to implement a prototype of our framework,
-with an automatic translation from RSs to the {\tt link}-calculus. 
+with an automatic translation from RSs to  {\tt link}-calculus. 
 We believe that our work can also help to extend the framework
@@
 properties of the
-modeled systems 
-\cite{CFHOT15,OCHF16,BBGLBH2017,FOP15,CFOP10}.
-
-\medskip
-
-\noindent
-{\small {\bf Acknowledgments}
-We thank the anonymous reviewers for their detailed and very useful 
-criticisms and recommendations that helped us to improve our paper.}
+modeled systems \cite{CFHOT15,OCHF16,BBGLBH2017}.

### 2020-07-27 fragments/01-introduction.tex
da dropbox

--- 2020-07-27 Catuscia - submitted/01-introduction.tex
@@
 the interaction between RSs can help to model new scenarios.
+\paragraph{Related work}
+We are not aware of any inference rules for the RS; it seems not the case
+that a deterministic behavior, as the one of the RS, once the initial state is 
+fixed, could be defined by inference rules that classically are applied 
+to define non-deterministic transition systems.
+The reversible computation paradigm for RS extends the RS framework  by
+allowing  backward computations as well as forward computations; for this 
+goal a set of inference rules for rewriting logic has been 
+defined in~\cite{10.1007/978-3-319-73359-3_3}
+to exactly keep trace of those elements that  dissolve in the next 
+computation step.
 

### 2020-07-27 fragments/06b-bisimulation.tex
da dropbox

--- 2020-04-05 06b-bisimulation.tex
@@
 
-The classical notion of bisimulation for process algebras equates two processes
+The classical notion of bisimulation for process algebra equates two processes
 when one process can simulate all the instructions executed by the other one and viceversa.
@@
 barbed bisimulation~\cite{10.1007/3-540-55719-9_114} only consider the execution of invisible actions, and then equates two processes when the expose the same prefixes; for the mobile ambients~\cite{CardelliG00}, a process algebra equipped with a reduction semantics, a notion of
-behavioural equivalence equates two processes when they expose the same ambients~\cite{GC03}.
+behavioural equivalence equates two processes when they expose the same ambients\cite{GC03}.
 
 There are some previous works based on bisimulation applied to models for biological systems.
-Barbuti et al~\cite{BMMT08} define a classical setting for bisimulation for two formalisms: the Calculus of Looping Sequences, which is a rewriting system, and the Brane Calculi, which is based on process calculi.
+Barbuti et al \cite{BMMT08} define a classical setting for bisimulation for two formalisms: the Calculus of Looping Sequences, which is a rewriting system, and the Brane Calculi, which is based on process calculi.
 Bisimulation is used to verify properties of the regulation of lactose degradation in
 Escherichia coli and the EGF signalling pathway. These calculi allow the authors to model membranes' behaviour.
-Cardelli et al~\cite{CTTV15} present two quantitative behavioral equivalences over species of a 
+\cite{CTTV15} presents two quantitative behavioral equivalences over species of a 
 chemical reaction network with semantics based on ordinary differential equations.
@@
 Bisimulation also relates species that have identical solutions at all time points when starting from the same initial conditions.
-Both the mentioned formalisms~\cite{BMMT08,CTTV15} adopt a classical approach to bisimulation. 
+Both the mentioned formalisms \cite{BMMT08,CTTV15} adopt a classical approach to bisimulation. Moreover our formalisms is different, for different applications and purposes.
+
 Albeit the bisimulation is a powerful tool for verifying if the behaviour of two different software 
@@
 
-For this reason, we propose an alternative notion of bisimulation, that hereafter we call \emph{bio-simulation},
+For this reason, we propose an alternative notion of bisimulation, that hereafter we call biosimulation,
 that allows us  to compare two biological systems by restricting the observation to only the limited events of interest.
-This allows one to tailor the equivalence to different applications and purposes.
 
 The transition labels of our systems record detailed information about all the reactions that have been applied in one transition, about the elements that acted as reagents, as inhibitors or that have been produced, or that have been provided by the context.
-All these information are stored in the label because they are necessary to compose a transition in a modular way.
-Depending on the application, only a suitable abstraction over the label can be of interest.
 
-In a way, at each step of the bisimulation game, we want to query our transition labels to get only the information we care about.
+In a way, we want to query our transition labels to get only the information we are interested in.
 To this goal, we introduce a simple language
@@
 
-For instance we would like to express properties about each step of the simulation of a system like in the following:
+First, we need to introduce the $flat$ operator that takes a link chain and returns a simple string
+by eliminating all the channel matching pairs leaving just one channel per each pair:\\
+\[
+\begin{array}{lcl@{\hspace{0.5cm}}lcl@{\hspace{0.5cm}}lcl@{\hspace{0.5cm}}lcl}
+flat(\epsilon)&\defeq & \epsilon &
+flat(\link{\silent}{\beta}) & \defeq & \beta &
+flat(\link{\alpha}{\silent})& \defeq & \epsilon &
+flat(\link{\alpha}{\beta}) & \defeq &\beta \\
+\end{array}\]
+\[
+\begin{array}{lcl}
+flat(\link{\alpha}{\beta}\upsilon)& \defeq & \beta :: flat(\upsilon)\\
+\end{array}
+\]
 
-\begin{example} \label{properties_ex}
-\begin{enumerate}
-\item Has the entity $s_i$ been used by rule $r_j$ as reagent?
-\item Has the entity $s_i$ blocked the application of  rule $r_j$?
-\item Has the entity $s_i$ been produced by rule $r_j$?
-\item Has the entity $s_i$ been produced by some rule?
-\item Has the entity $s_i$ been provided by the context? 
-\item Has the rule $r_j$ not been applied? 
-%\item Has the entity $s$ been used by some rule as reagent ?
-%\item Has the entity $s$ blocked  the application of some rules ?
-%\item Has the entity $s$ been produced ?
-%\item Has the entity $s$ been provided by the context ?
-%\item Has the rule $i$ been applied ?
-\end{enumerate}
-\end{example}
-
-As detailed before, in the following we assume that: (i) the context can be non-deterministic, otherwise it makes little sense to rely on bisimulation to observe the branching structure of system dynamics; (ii) we are interested in observing the names of the entities involved in the transitions and also the rules that have been applied, thus we assume top level restrictions are absent and rely on solid transitions only (with leftmost and rightmost silent actions).
-
-\subsection{Assertion language}
-
-%To simplify the query language, we first introduce a function that takes a solid link chain and returns a simple string by eliminating all the channel matching pairs leaving just one channel per each pair.
-This transformation is harmless, in the sense that it retains all the information in the chain, because it is applied to solid chains only.
-Below we denote by $\mathit{decs} = \{ s, \overline{s}, \widetilde{s},\widehat{s},\underline{s}~|~ s\in S\}$ the set of decorated names (without subscripts $_i$ and $_o$).
-%The function $\mathit{flat}$ is defined inductively as follows:
-%%\[
-%%\begin{array}{lcl@{\hspace{0.5cm}}lcl@{\hspace{0.5cm}}lcl@{\hspace{0.5cm}}lcl}
-%%\mathit{flat}(\epsilon)&\defeq & \epsilon &
-%%\mathit{flat}(\link{\gamma}{\gamma'}) &  \defeq & \left \{ \begin{array}{cl} \epsilon & \mathrm{if } (\gamma = \beta_i \wedge \gamma = \beta_o)  \vee \gamma = \silent\\
-%%\gamma' & \mathrm{otw }  \end{array} \right. &
-%%%\mathit{flat}(\link{\alpha}{\silent})& \defeq & \epsilon &
-%%%\mathit{flat}(\link{\alpha}{\beta}) & \defeq &\beta \\
-%%\end{array}\]
-%
-%\[
-%\mathit{flat}(\epsilon)\defeq \epsilon 
-%\qquad
-%\mathit{flat}(\link{\alpha}{\beta}) \defeq 
-%\left \{ \begin{array}{ll} 
-%\beta & \mbox{if $\beta\in \mathit{reacts}\cup\{\mathit{cxt}\}\cup\mathit{prods}$}\\
-%\gamma & \mbox{if $\beta=\gamma_i$ with $\gamma\in\mathit{decs}$}\\
-%\epsilon & \mbox{otherwise}  
-%\end{array} \right. 
-%\]
-%
-%\[
-%\begin{array}{lcl}
-%\mathit{flat}(\link{\alpha}{\beta}\upsilon)& \defeq & \mathit{flat}(\link{\alpha}{\beta}):: \mathit{flat}(\upsilon)\\
-%\end{array}\\
-%\]
-%where  $::$ is the usual string concatenation operator.
-%
-
-Next, we introduce an assertion language that operates on simple strings
-and that combines regular expression operators with conjunction and disjunction.
-In the following we let $\mathit{syms} = \mathit{decs}\cup\mathit{reacts}\cup\mathit{prods}\cup\{\mathit{cxt}\}$ be the set of symbolic names used in our assertion language.
-
-\begin{definition}[Assertion Language]
+\begin{definition}[Assertion language]
 \label{def:assetionl}
@@
 \begin{array}{lcl}
-\zeta & ::= & 
-\alpha ~\mid ~ 
-? ~\mid ~ 
-[N]
-\\
-F & ::= & 
-\epsilon ~\mid ~ 
-\zeta ~\mid ~ 
-F::F ~\mid ~ 
-F^+ ~\mid ~ 
-F^* ~\mid ~ 
-F \vee F ~\mid ~ 
-F \wedge F
+\eta &::=& \zeta ~\mid ~  \eta \vee \eta ~\mid ~  \eta \wedge \eta\\
 \end{array}
@@
 %\noindent where $s \in {\cal C}$, and  $\oplus \in \{\wedge, \vee, \to \}$. 
-where $\alpha \in \mathit{syms}$ and $N\subseteq \mathit{syms}$.
+where $\zeta \in \{s, \tilde{s}, \hat{s}, \overline{s}, \underline{s} \}$ with  $s \in {\cal C}$.
 %, and  $\oplus \in \{\wedge, \vee \}$. 
@@
 
-Roughly, a \emph{singleton} assertion $\zeta$ denotes either a string composed by a single symbol $\alpha$ (one of the symbols in the set $\mathit{syms}$ for denoting a particular entity, rule, production or context), or
-the wildcard $?$ that stands for any symbol, or the pattern $[N]$ that stands for any of the strings composed by a single symbol in the set $N$. Clearly $?$  is just a shorthand for $[\mathit{syms}]$.
-We write $\mathbf{0}$ for $[\emptyset]$ and $[s_1,...,s_n]$ instead of $[\{s_1,...,s_n\}]$.
-
-An \emph{assertion} $F$ is either the empty string $\epsilon$, a singleton assertion $\zeta$, the concatenation of two assertions $F_1::F_2$, the replication of $F$ for 1 or more times $F^+$, the replication of $F$ for 0 or more times $F^*$, the disjunction of two assertions $F_1 \vee F_2$ or their conjunction $F_1 \wedge F_2$.
-We write $\star$ as an abbreviation of $?^*$.
-
-An assertion denotes a set of strings over the alphabet $\mathit{syms}$ as expected.
-
-\begin{definition}[Semantics of Assertions]
- We define $\llbracket F\rrbracket \subseteq\wp(\mathit{syms}^*)$ by induction on the structure of $F$: 
-\[
-\begin{array}{rcl@{\hspace{1cm}}rcl}
- \llbracket \epsilon \rrbracket &\defeq&  \{\epsilon\}
- & \llbracket F_1::F_2 \rrbracket &\defeq& \{\omega_1 :: \omega_2~\mid~ \omega_1 \in \llbracket F_1 \rrbracket \wedge \omega_2 \in \llbracket F_2 \rrbracket\}  \\
- \llbracket \alpha \rrbracket &\defeq&  \{\alpha\} 
- & \llbracket F^+ \rrbracket &\defeq& \llbracket F \rrbracket^+  \\
- \llbracket ? \rrbracket &\defeq&   \mathit{syms}
- & \llbracket F^* \rrbracket &\defeq& \llbracket F \rrbracket^*  \\
- \llbracket [N]\rrbracket &\defeq&   N
- & \llbracket  F_1 \vee F_2 \rrbracket &\defeq& \llbracket  F_1\rrbracket \cup \llbracket  F_2\rrbracket  \\
- &&&\llbracket  F_1 \wedge F_2 \rrbracket &\defeq& \llbracket F_1 \rrbracket \cap \llbracket F_2\rrbracket 
-\end{array}
-\]
+\begin{definition}[Semantics]\label{sec:semantics}
+Let  $\upsilon$ be  a transition label, and $\eta$ be an assertion.
+We define $\upsilon \entails \eta$ (read as the transition label $v$ satisfies the formula $\eta$) as: 
+\begin{itemize}
+ \item   $\upsilon \entails \zeta $ iff $\zeta \in flat(\upsilon)$. 
+ \item   $\upsilon \entails \eta_1 \wedge \eta_2$ iff   $\upsilon \entails \eta_1$  $\wedge$  $\upsilon \entails \eta_2$.
+\item   $ \upsilon \entails \eta_1 \vee \eta_2$ iff   $\upsilon \entails \eta_1$  $\vee$  $\upsilon \entails \eta_2$.
+\end{itemize}
+If it is not the case that $\upsilon \entails  \eta $, then we say that $\eta$ does not hold at $\upsilon$ and we write $\upsilon \not\entails \eta$. 
 \end{definition}
-
-\begin{definition}[Satisfaction as Membership]\label{sec:semantics}
-Let  $\upsilon$ be  a transition label, and $F$ be an assertion.
-We write $\upsilon \entails F$ (read as the transition label $v$ satisfies the assertion $F$) 
-if $\mathit{flat}(\upsilon)\in \llbracket F\rrbracket$, otherwise we write $\upsilon \not\entails F$ (or also $\upsilon \entails \neg F$) and say that $F$ does not hold at $v$. 
-\end{definition}
-
-Given two transition labels $v,w$ we write $v\equiv_F w$ if $v \entails F\ \Leftrightarrow\ w \entails F$, i.e. if both $v,w$ satisfy $F$ or they do not.
 
@@
 %\end{definition}
+Our  language allows us to check if the simulation of a system (i.e. the transition labels, or in other words the trace of a simulation) 
+satisfies the properties of the following types:% that can be expressed using our assertion language. Some simple examples  follow:
+\begin{itemize}
+\item Has the entity $s$ been used by some rule as reagent ?
+\item Has the entity $s$ blocked  the application of some rules ?
+\item Has the entity $s$ been produced ?
+\item Has the entity $s$ been provided by the context ?
+\item Has the rule $i$ been applied ?
+\end{itemize}
 
-The formulas corresponding to the sample queries listed in Example~\ref{properties_ex} are as follows:
+Now, we introduce a slight modified version of the Hennessy Milner Logic (the Hennessy Milner link logic, HM-linkL)~\cite{}; due to the reasons we explained above, we do not want to look at the complete transition labels, thus we rely on our simple assertion language.
+We consider the usual logical connectives $and$ and $or$
+and the two modalities:  $\langle F \rangle G$ and $[F] G$ where $F$ stands for any
+expression formulate with our assertion language.  $\langle F \rangle G$ says that there exists a 
+labeled transition using a link chain $v$ 
+target process satisfies the formula F . The later says that all successor states
+86 reachable with a link chain in ? must satisfy the formula F . Following the classical re-
+87 sult of Hennessy and Milner relating bisimulation of CCS and indistinguishability wrt
+88 HML formulas, we also show that network bisimilar CNA processes satisfy the same
+89 formulas in SLML.
 
-\begin{enumerate}
-\item
-$\star :: r_j :: [ s_1,...,s_n ]^* :: s_i :: [  s_1,...,s_n ]^* :: [  \overline{s}_1,...,\overline{s}_n ]^+ :: \star$
-\item
-$\star :: r_j :: [s_i, \overline{s}_i] :: r_{j+1} :: \star$
-\item
-$\star :: p_j :: [  \mathit{ents} ]^* :: \widetilde{s}_i :: \star$
-\item
-$\star :: \widetilde{s}_i :: \star$
-\item
-$\star :: \widehat{s}_i :: \star$
-\item
-$\star :: r_j :: ? :: r_{j+1} :: \star$
-\end{enumerate}
-where in 1, 2, 6 we exploit the fact that in a reaction $(R,I,P)$ the sets $R$ of reactants and  $I$ of inhibitors are non empty, so that if there is only one symbol between the occurrence of $r_j$ and $r_{j+1}$ it means the reaction $r_j$ has not been applied.
-Viceversa, if the reaction $r_j$ has been applied the occurrence of $r_j$ must be followed by at least one of the symbols in $\{s_1,...,s_n\}$ and then by at least one of the symbols in $\{\overline{s}_1,...,\overline{s}_n\}$.
+\begin{definition}[HM-linkL]
+Let   $F$ be an assertion, then 
+the set of  HM-linkL formulae ${\cal G}$are built by the following syntax:\\
+$$
+G,H ::= {\tt t} ~\mid~ {\tt f} ~\mid~ G\wedge H ~\mid~ G \vee H ~\mid~ \langle F \rangle G ~\mid~ [F]G
+$$
+\end{definition}
 
-\subsection{Bio-similarity and bio-logical equivalence}
+\begin{definition}[semantics of $G$]
+ We define $\llbracket G\rrbracket \subseteq {\cal P}$ by induction on the structure of $G \in {\cal G}$:%(read as the transition label $v$ satisfies the formula $F$) as: 
+\[
+\begin{array}{ccl@{\hspace{1cm}}ccl}
+ \llbracket {\tt t}\rrbracket &\defeq& {\cal P} &  \llbracket G\wedge H\rrbracket &\defeq&  \llbracket G\rrbracket  \cap \llbracket H\rrbracket  \\
+\llbracket {\tt f}\rrbracket& \defeq& \emptyset &
+  \llbracket G\vee H\rrbracket &\defeq&  \llbracket G\rrbracket  \cup \llbracket H\rrbracket \\
+%\item   $ v \entails F_1 \vee F_2$ iff   $v \entails F_1$  $\vee$  $v \entails F_2$.
+\end{array}
+\]
+\[
+\begin{array}{lcl}
+ \llbracket  \langle F \rangle G \rrbracket &\defeq& \{P \in {\cal P}: \exists P', v.P\xrightarrow{v}P' \mbox{ with } v \entails F \mbox{ and } P' \in  \llbracket  G \rrbracket\} \\
+ \llbracket [F]G \rrbracket &\defeq& \{P \in {\cal P}: \forall P',v.P\xrightarrow{v}P' \mbox{ implies } v \entails F \mbox{ and } P' \in  \llbracket  G \rrbracket\}
+\end{array}
+\]
+We write $P \vdash G$ ($P$ satisfies $G$) if and only if $P \in  \llbracket G \rrbracket$.
+\end{definition}
 
-The notion of bio-simulation builds on the above language of assertions to parameterize the induced equivalence on the property of interest. 
-Please recall that we have defined the behaviour of the contest in a non deterministic way, thus 
-at each step, different possible sets of entities can be provided to the system and different sets of reaction can be enabled/disabled. 
-Bio-simulation can thus be used to compare the behaviour of different systems that share some of the reactions or entities or also to compare the behaviour of the same set of reaction rules when different contexts are provided.
-
+Please recall that we have defined the behaviour of the contest in a non determinist way, thus 
+at each step, different possible sets of entities can be provided to the system. 
 %situations where in the same conditions, two alternative (mutually exclusive) behaviors can happen.
@@
 
-\begin{definition}[Bio-similarity  $ \sim_F $]
-Given an assertion $F$, a \emph{bio-simulation} $\mathbf{R}_F$ that respects $F$ is a binary relation over c\CNA \  processes such that, whenever $P~\mathrel{\mathbf{R}_F}~Q$ then:
- \begin{itemize} 
- \item
- for any $v,P'$ such that 
- $P \xRightarrow{v} P'$ then there exist $w,Q'$ such that $Q  \xRightarrow{w} Q'$ with $v\equiv_F w$ and $P' \mathrel{\mathbf{R}_F} Q'$.
-\item
- for any $w,Q'$ such that 
- $Q \xRightarrow{w} Q'$ then there exist $v,P'$ such that $P  \xRightarrow{v} P'$ with $v\equiv_F w$ and $P' \mathrel{\mathbf{R}_F} Q'$.
-\end{itemize}
+\begin{definition}[Biosimilarity  $ \sim_F $]
+A \emph{biosimulation} $\mathbf{R}_F$, with respect to the predicate $F$, is a binary relation over {\tt link}-calculus processes such that, if $P \mathrel{\mathbf{R}_F} Q$ then:
+%\begin{itemize}
+%\item 
+ \begin{center} %s$\exists \, v,v'$ s.t. 
+ $P \xrightarrow{v} P'$ with $v \entails F$ \, iff  \, $Q  \xrightarrow{v'} Q'$ with $v' \entails F$ and $P' \mathrel{\mathbf{R}_F} Q'$.
+ \end{center}
 %\item i$Q \xrightarrow{v} Q'$ with $v \entails F$ iff $\exists v',P'$ s.t.~$v \stretcheq v'$, $P  \xrightarrow{v'} P'$ with $v \entails F$. and $P' \mathrel{\mathbf{R}_{F}} Q'$.
 %\end{itemize}
-We let $ \sim_F $ denote the largest  bio-simulation and we say that $P$ is \emph{bio-similar} to $Q$, with respect to $F$, if $P \sim_F Q$.
+\noindent
+We let $ \sim_F $ denote the largest  biosimulation and we say that $P$ is \emph{biosimilar} to $Q$, with respect to $F$, if $P \sim_F Q$.
 \end{definition}
-
-\begin{remark}
-Please remember that the notation $P \xRightarrow{v} P'$ refers to ordinary transitions $P \xrightarrow{v} P'$ where $v$ is a complete chain (solid and with $\silent$ actions at the extremes).
-It should not be confused with the notation for weak transitions commonly found in the literature on process algebras.
-\end{remark}
-
-\begin{remark}
-An alternative way to look at a bio-simulation that respects $F$ is to define it as an ordinary bisimulation over the transition system labelled over $\{F,\neg F\}$ obtained by transforming each transition $P \xRightarrow{v} P'$ such that $v\entails F$ into $P \xrightarrow{F} P'$ and  each transition $P \xRightarrow{v} P'$ such that $v\not\entails F$ into $P \xrightarrow{\neg F} P'$.
-\end{remark}
-
-It can be easily shown that the identity relation is a bio-simulation and that bio-simulations are closed under (relational) inverse, composition and union and that, as a consequence, bio-similarity is an equivalence relation.
-
-Now, we introduce a slightly modified version of the Hennessy Milner Logic~\cite{HM80}, called bioHML; due to the reasons we explained above, we do not want to look at the complete transition labels, thus we rely on our simple assertion language to make it parametric to the assertion $F$ of interest:
-\begin{definition}[BioHML]
-Let   $F$ be an assertion, then 
-the set of bioHML formulas $G$ that respects $F$ are built by the following syntax:
-$$
-\begin{array}{rcl}
-\chi & ::= & F~\mid~ \neg F\\
-G &::= &{\tt t} ~\mid~ {\tt f} ~\mid~ G\wedge G ~\mid~ G \vee G ~\mid~ \langle \chi \rangle G ~\mid~ [\chi]G
-\end{array}
-$$
-\end{definition}
-
-\begin{remark}
-An alternative way to look at bioHML formulas is as ordinary HML formulas over the set of labels $\{F,\neg F\}$.
-\end{remark}
-
-As usual, the semantics of a bioHML formula is the set of processes that satisfy it.
-
-\begin{definition}[Semantics of BioHML]
- We define $\llbracket G\rrbracket \subseteq {\cal P}$ by induction on the structure of $G$:
- %(read as the transition label $v$ satisfies the formula $F$) as: 
-\[
-\begin{array}{ccl@{\hspace{1cm}}ccl}
- \llbracket {\tt t}\rrbracket &\defeq& {\cal P} &  \llbracket G\wedge H\rrbracket &\defeq&  \llbracket G\rrbracket  \cap \llbracket H\rrbracket  \\
-\llbracket {\tt f}\rrbracket& \defeq& \emptyset &
-  \llbracket G\vee H\rrbracket &\defeq&  \llbracket G\rrbracket  \cup \llbracket H\rrbracket \\
-%\item   $ v \entails F_1 \vee F_2$ iff   $v \entails F_1$  $\vee$  $v \entails F_2$.
-\end{array}
-\]
-\[
-\begin{array}{lcl}
- \llbracket  \langle \chi \rangle G \rrbracket &\defeq& \{P \in {\cal P}: \exists v,P'.\ P\xRightarrow{v}P' \mbox{ with } v \entails \chi \mbox{ and } P' \in  \llbracket  G \rrbracket\} \\
- \llbracket [\chi]G \rrbracket &\defeq& \{P \in {\cal P}: \forall v,P'.\ P\xRightarrow{v}P' \mbox{ implies } v \entails \chi \mbox{ and } P' \in  \llbracket  G \rrbracket\}
-\end{array}
-\]
-We write $P \entails G$ ($P$ satisfies $G$) if and only if $P \in  \llbracket G \rrbracket$.
-\end{definition}
-
-Negation is not included in the syntax, but the converse $\overline{G}$ of a bioHML formula $G$ can be easily defined inductively in the same way as for HML logic.
-
-\begin{definition}[Converse]
-Given a bioHML formula $G$ we define its converse $\overline{G}$ as follows:
-\[
-\begin{array}{rcl@{\hspace{1cm}}rcl@{\hspace{1cm}}rcl}
-\overline{\tt t} & \defeq & {\tt f} &
-\overline{G \wedge H} & \defeq & \overline{G} \vee \overline{H} &
-\overline{\langle \chi \rangle G} & \defeq & [\chi]\overline{G}
-\\
-\overline{\tt f} & \defeq & {\tt t} &
-\overline{G \vee H} & \defeq & \overline{G} \wedge \overline{H} &
-\overline{[\chi]G} & \defeq & \langle \chi \rangle\overline{G}
-\end{array}
-\]
-\end{definition}
-
-We observe that, as expected, for any bioHML formula $G$ and process $P$ we have $\overline{\overline{G}} = G$ and  $P \entails \overline{G}$ iff  $P\not\entails G$.
-
-We let $\mathcal{L}_F$ be the set of all bioHML formulas that respects $F$ and 
-we say that two processes $P,Q$ are bio-logically equivalent w.r.t.
-$F$, written $P \equiv_{\mathcal{L}_F}Q$, when $P$ and $Q$ satisfy the exactly the same bioHML formulas in $\mathcal{L}_F$, i.e.
-when for any $G\in \mathcal{L}_F$ we have $P \entails G\ \Leftrightarrow Q \entails G$.
-
-Finally, we extend the classical result establishing the correspondence between the logical equivalence induced by HML with bisimilarity for proving that bio-similarity coincides with bio-logical equivalence.
-
-\begin{restatable}[Correspondence]{theorem}{theocorr}
-%\begin{theorem}
-$\sim_F\ =\ \equiv_{\mathcal{L}_F}$
-%\end{theorem}
-\end{restatable}
-%\begin{proof}
-%The proof is just an adaptation of the classical result. The two implications are proved separately.
-%\begin{description}
-%\item[$\sim_F\ \subseteq\ \equiv_{\mathcal{L}_F}$:]
-%Given any two processes $P\sim_F Q$ we need to prove that for any bioHML formula $G$ we have $P\entails G$ iff $Q\entails G$. 
-%Without loss of generality, we prove that $P\entails G$ implies $Q\entails G$.
-%The proof is by structural induction on $G$.
-%\begin{itemize}
-%\item 
-%if $G={\tt t}$, then $Q\entails G$.
-%
-%\item
-%if $G={\tt f}$, then the assumption $P\entails G$ is false and the implication holds.
-%
-%\item
-%if $G=G_1 \wedge G_2$ we take as inductive hypotheses that
-%$$
-%\begin{array}{l}
-%\forall R,S.\ R \sim_F S \wedge R\entails G_1 \Rightarrow S \entails G_1\\
-%\forall R,S.\ R \sim_F S \wedge R\entails G_2 \Rightarrow S \entails G_2
-%\end{array}
-%$$
-%We need to prove that $Q\entails G$.
-%Since $P\entails G=G_1 \wedge G_2$ we have $P\entails G_1$ and $P\entails G_2$.
-%Since $P\sim_F Q$, by inductive hypotheses we get $Q\entails G_1$ and $Q\entails G_2$.
-%Hence $Q\entails G_1 \wedge G_2=G$.
-%
-%
-%\item
-%if $G=G_1 \vee G_2$  we take as inductive hypotheses that
-%$$
-%\begin{array}{l}
-%\forall R,S.\ R \sim_F S \wedge R\entails G_1 \Rightarrow S \entails G_1\\
-%\forall R,S.\ R \sim_F S \wedge R\entails G_2 \Rightarrow S \entails G_2
-%\end{array}
-%$$
-%We need to prove that $Q\entails G$.
-%Since $P\entails G=G_1 \vee G_2$ we have $P\entails G_1$ or $P\entails G_2$.
-%If $P\entails G_1$, since $P\sim_F Q$, by inductive hypotheses we get $Q\entails G_1$ and thus $Q\entails G_1 \vee G_2=G$.
-%If $P\entails G_2$, since $P\sim_F Q$, by inductive hypotheses we get $Q\entails G_2$ and thus $Q\entails G_1 \vee G_2=G$.
-%
-%\item
-%if $G=\langle\chi\rangle H$ we take as inductive hypothesis that
-%$$
-%\forall R,S.\ R \sim_F S \wedge R\entails H \Rightarrow S \entails H
-%$$
-%We need to prove that $Q\entails G$.
-%Since $P\entails \langle\chi\rangle H$ it means that there exists $\upsilon, P'$ such that $P\xRightarrow{\upsilon} P'$ with $\upsilon\entails \chi$ and $P'\entails H$.
-%Since $P\sim_F Q$, there exists $w,Q'$ such that $Q\xRightarrow{w} Q'$ with $w\entails \chi$ and $P'\sim_F Q'$.
-Then, by inductive hypothesis, $Q'\entails H$ and thus $Q\entails \langle\chi\rangle H=G$.
-%
-%\item
-%if $G=[\chi] H$ we take as inductive hypothesis that
-%$$
-%\forall R,S.\ R \sim_F S \wedge R\entails H \Rightarrow S \entails H
-%$$
-%We need to prove that $Q\entails G$.
-%If there is no $\upsilon\entails \chi$ such that $Q\xRightarrow{\upsilon}Q'$ for some $Q'$, then 
-% $Q\entails [\chi] H=G$ trivially.
-%For any $\upsilon,Q'$ such that $Q\xRightarrow{\upsilon}Q'$ with $\upsilon\entails \chi$, then as $P\sim_F Q$ there must exist $w,P'$ such that $P\xRightarrow{w}P'$ with $w\entails \chi$ and $P'\sim_F Q'$.
-Since $P\entails G=[\chi] H$ then it must be $P'\entails H$.
-Since $P'\sim_F Q'$, by inductive hypothesis $Q'\entails H$.
-Hence $Q\entails [\chi] H = G$.
-%\end{itemize}
-%
-%\item[$\equiv_{\mathcal{L}_F}\ \subseteq\ \sim_F$:]
-%We prove that $\equiv_{\mathcal{L}_F}$ is a bio-simulation and thus included in $\sim_F$.
-%Take two generic processes $P\equiv_{\mathcal{L}_F} Q$ and suppose $P\xRightarrow{\upsilon}P'$ for some $\upsilon,P'$. 
-%\begin{itemize}
-%\item
-%If $\upsilon \entails F$ we want to prove that there exists some $w,Q'$ such that $Q\xRightarrow{w}Q'$, with $w\entails F$ and $P'\equiv_{\mathcal{L}_F} Q'$.
-%
-%Towards a contradiction, assume that we cannot find such $w,Q'$.
-%
-%If there is no transition $Q\xRightarrow{w}Q'$ such that $w\entails F$, then the bioHML formula $G \defeq \langle F\rangle{\tt t}$ is such that $P\entails G$ and $Q\not\entails G$, contradicting the assumption $P\equiv_{\mathcal{L}_F} Q$.
-%
-%Otherwise, let $\mathcal{Q}\defeq \{Q'~\mid~\exists w.\ Q\xRightarrow{w}Q' \ \wedge\ w\entails F\}$ be the set of processes reachable from $Q$ via a transition with a (complete) label that satisfies $F$.
-Since our processes are with guarded recursion, the set $\mathcal{Q}$ is finite.
-Let $\mathcal{Q} = \{Q'_1,...,Q'_n\}$.
-By hypothesis all processes in $\mathcal{Q}$ must not be bio-logically equivalent to $P'$, hence for any $i\in[1,n]$ there exists a bioHML formula $G_i$ such that $P'\entails G_i$ and $Q'_i\not\entails G_i$ (if it was the opposite, $P'\not\entails G_i$ and $Q'_i\entails G_i$, we can use the converse formulas).
-But then the formula $G\defeq \langle F\rangle(G_1\wedge\dots\wedge G_n)$ is such that $P\entails G$ and $Q\not\entails G$, contradicting the assumption $P\equiv_{\mathcal{L}_F} Q$.
-%
-%\item
-%If $\upsilon \not\entails F$ then the proof is analogous to the previous case and thus omitted.
-%\end{itemize}
-%\end{description}
-%\end{proof}

### 2020-07-27 fragments/old/03-background-links-CNA.tex
da dropbox

% !TEX root = ./main_RS2L_LNCS.tex

\section{Background on the CNA}

\section{The Core Network Algebra}
\label{sec:network}

In this section we recall the theory of links and link chains to present the syntax and operational semantics of \CNA.
%In this section we just focus on the basic ideas of link interaction, while we shall enhance the model with name mobility in the next section.

\subsection{Syntax}

\begin{definition}
The \CNA\ processes are generated by the following grammar:
\[
\begin{array}{rcl}
P,Q & ::= &
\nil \prodsep
%X \prodsep
\ell. P  \prodsep
P+Q \prodsep
P|Q \prodsep
\restrict{a} P \prodsep
P[\phi] \prodsep
A
\end{array}
\]
\noindent
where $\ell$ is a solid link (i.e.\  $\ell = \link{\alpha}{\beta}$ with $\alpha,\beta\neq \ \noacts$),
$\phi$ is a channel renaming function,
and $A$ is a process identifier for which we assume a definition $A \defeq P$ is available in a given set $\Delta$ of (possibly recursive) process definitions.
\end{definition}

As usual, we write $\tilde{a}$ for tuples of channels and we allow parametric process definitions of the form $A(\tilde{a}) \defeq P$,
where $\tilde{a}$ is the set of free channels of $P$.
For brevity, in the examples, we sometimes write $A \defeq P$ leaving implicit that the free channles of $P$ are the parameters of $A$.
%that
%stand for the set of definitions $\{ A(\tilde{a}) \defeq P[\tilde{a} \slash \tilde{x}]  \;\mid\; \tilde{a}\; \text{ is a tuple of channels} \}$,
%where $\tilde{x}$ is the set of free channels of $P$.

%\begin{remark}
%The extension in which generic link chains are allowed as action prefixes instead of solid links is discussed in Section~\ref{subsec:altdef}.
%\end{remark}

It is evident that processes are built over a CCS-like syntax,
%\footnote{For brevity, we omit the CCS-like renaming operator $P[\Phi]$ because it will be later subsumed by the syntax in Section~\ref{sec:names}).}
with inactive process $\nil$, action prefix $\ell.P$, choice $P+Q$, parallel $P|Q$, restriction $\restrict{a} P$, renaming $P[\phi]$ and constant definition
 $A$, but where the underlying synchronisation algebra~\cite{DBLP:journals/tcs/Winskel84} is based on link chains.
This is made evident by the operational semantics that we present next.

As usual, $\restrict{a}P$ binds the occurrences of $a$ in $P$, the sets of free and of bound names of a process $P$ are defined in the obvious way and 
processes are taken up to alpha-conversion of bound names.
We shall sometimes omit trailing $\nil$, e.g.~by writing $\link{a}{b}$ instead of $\link{a}{b}.\nil$.

\begin{example}
\CNA\ provides us with a natural way to rephrase the communication primitives of usual process calculi, such as CCS and CSP~\cite{Hoare85}, in terms of links.
\begin{itemize}
\item
Intuitively, the output action $\overline{a}$ (resp. the input action $a$) of CCS can be seen as the link $\link{\silent}{a}$
(resp. $\link{a}{\silent}$) and the solid link chain $\startchain{\silent}\chainedlink{a}{a}\chainend{\silent}$  as a dyadic communication, analogous to the silent action $\tau$ of CCS.
% between an output on $a$ (the link $\link{\silent}{a}$) and an input on $a$ (the link $\link{a}{\silent}$).
\item
The action $a$ of CSP can be seen as the link $\link{a}{a}$ and the solid link chain $\startchain{a}\chainedlink{a}{a}\chainedlink{a}{a}\chainend{a}$ as a CSP-like communication among three peers over $a$.
\end{itemize}
\end{example}

%In particular, to implement I/O communication, each link in the whole chain simply carries the same list or tuple of arguments, 
%but with different (send/receive) capabilities. 
%In this way, we keep apart the interaction mechanism from the name-passing one that eventually fit together in synchronisations.
%We have just borrowed from $\pi$-calculus
%the name handling machinery
%(and liberated it from dyadic interaction legacy), still having input, output and extrusion mechanisms.

%In the tuple $t =\tuple{\vec{w}}$, names can be used either as values or as variables. They are distinguished because variables are underlined.
%During communication, variables are instantiated by values, while
%values are used for matching arguments, as in \cite{BodeiBDG10}.
%\begin{example}\label{ex-tuple}
%Consider e.g., the tuples in the two ``complementary'' prefixes $\link{\tau}{a}\tuple{id,n,\variable{x}}.P$ and $\link{a}{\tau}\tuple{id,\variable{y},m}.Q$: the two links can be merged, the first parameters of the tuples must match exactly,
%$n$ is assigned to $y$, while $m$ is assigned to $x$, where
%$\variable{x}$ is an input for $P$, $\variable{y}$ is an input for $Q$, and $id$ is a name known by both processes.
%\end{example}
%This mechanism allows, e.g., a form of multi-way communication, where all peers involved in the chain link can express arguments to be matched and provide actual arguments to replace the formal ones.

%
%For a tuple $t$, we let $\values{t}$ and $\variables{t}$ denote the set of values and the set of variables of $t$, respectively. 
%We say that a tuple $t$ is \emph{ground} if $\variables{t} = \emptyset$.

%We assume action names are admissible values, i.e.~ as in $\pi$-calculus we have the possibility to communicate (names of) means of communication.
%Similarly to $\restrict{a}$, the prefix $\ell t.P$ binds the occurrences of the variables $\variables{t}$ in $P$ (and the notions of free names $\fn{P}$, bound names $\bn{P}$ and alpha-conversion are updated accordingly).
%In the following, given two set of names $S$ and $T$, we write $S \# T$ as a shorthand for $S\cap T = \emptyset$.

%
%As usual, $\restrict{a}P$ binds the occurrences of $a$ in $P$, the sets of free names and of bound names of a process $P$ are defined in the obvious way and denoted, respectively, by $\fn{P}$ and $\bn{P}$, processes are taken up to alpha-conversion of bound names, and we shall often omit trailing $\nil$, e.g. by writing $\link{a}{b}$ instead of $\link{a}{b}.\nil$

% The only original operator is $\solid{P}$ that filters out incomplete communications
% (i.e.~it allows $P$ to evolve only via solid link chains).

\subsection{Operational Semantics}

The idea is that communication can be routed across several processes by combining the links they make available to form a link chain. Since the length of the link chain is not fixed {\em a priori}, an open multi-party synchronisation is realised.

% We take as observations only link chains whose extremities are silent,
% i.e.~ are in the form $\tau$, because they have not to be composed further.

%For compositionality reasons, we allow observations to contain
%virtual links $\link{\noact}{\noact}$, to be provided by the context.

\begin{figure}[t]
\begin{center} 
\begin{prooftree} 
\AxiomC{$s \blackstretcheq \ell $} %\mbox{($\ell$ {\em only} solid link in $s$)}
\RightLabel{\scriptsize(\textit{Act})}
\UnaryInfC{$\ell.P \xrightarrow{v} P$} 
\DisplayProof
\
\AxiomC{$P \xrightarrow{s} P'$} 
\RightLabel{\scriptsize(\textit{Lsum})}
\UnaryInfC{$P+Q \xrightarrow{v} P'$} 
\DisplayProof
\
\AxiomC{$P \xrightarrow{s} P'$} 
\RightLabel{\scriptsize(\textit{Res})}
\UnaryInfC{$\restrict{a}P \xrightarrow{\restrict{a}s} \restrict{a}P'$} 
%\DisplayProof
%\,
%\AxiomC{$Q \xrightarrow{s} Q'$} 
%\RightLabel{\scriptsize(Rsum)}
%\UnaryInfC{$P+Q \xrightarrow{s} Q'$} 
\end{prooftree} 
\end{center}

\begin{center} 
\begin{prooftree} 
\AxiomC{$P \xrightarrow{s} P'$} 
\RightLabel{\scriptsize(\textit{Ren})}
\UnaryInfC{$P[\phi] \xrightarrow{s[\phi]} P'[\phi]$} 
\DisplayProof
\
\AxiomC{$P \xrightarrow{s} P'$} 
\RightLabel{\scriptsize(\textit{Lpar})}
\UnaryInfC{$P|Q \xrightarrow{s} P'|Q$} 
\DisplayProof
\ 
%\AxiomC{$Q \xrightarrow{s} Q'$} 
%\RightLabel{\scriptsize(Rpar)}
%\UnaryInfC{$P|Q \xrightarrow{s} P|Q'$} 
%\end{prooftree} 
%\end{center}
%\begin{center} 
%\begin{prooftree} 
\AxiomC{$P \xrightarrow{s} P'$} 
\AxiomC{$Q \xrightarrow{s'} Q'$} 
\RightLabel{\scriptsize(\textit{Com})}
\BinaryInfC{$P|Q \xrightarrow{\merges{s}{s'}} P'|Q'$} 
\DisplayProof
\end{center}
\begin{center}
%\AxiomC{$P[\subst{\recproc{X}{P}}{X}] \xrightarrow{s} P'$} 
%\RightLabel{\scriptsize(Rec)}
%\UnaryInfC{$\recproc{X}{P} \xrightarrow{s} P'$} 
\AxiomC{$P \xrightarrow{s} P'$}
 \AxiomC{$(A \defeq P)\in\Delta$}
\RightLabel{\scriptsize(\textit{Ide})}
\BinaryInfC{$A \xrightarrow{s} P'$} 
\end{prooftree} 
\end{center}
\caption{SOS semantics of the \CNA\  (rules (\textit{Rsum}) and (\textit{Rpar}) are omitted).}
\label{fig:cnasos}
\end{figure}

The operational semantics is defined in terms of a Labelled Transition System, in which states are \CNA\ processes, labels are link chains, and transitions are generated by the SOS rules in Figure~\ref{fig:cnasos}. 
% 
%%The standard rules \textit{(Rsum)} and \textit{(Rpar)}, symmetric of \textit{(Lsum)} and \textit{(Lpar)} are omitted for brevity.
Notice that the rules are very similar to the ones of CCS, 
apart from the labels that record the link chains involved in the transitions:
moving from dyadic to \emph{linked} interaction does not introduce any complexity burden from the formal point of view.

We comment in details the rules (\textit{Act}), (\textit{Res}), and (\textit{Com}). 
In rules (\textit{Res}) and (\textit{Com}) we leave implicit the side conditions $\restrict{a}s \neq \bot$ and $\merges{s}{s'} \neq \bot$, respectively (they can be easily recovered by noting that otherwise the label of the transition in the conclusion would be undefined).

The rule (\textit{Act}) states that $\ell.P \xrightarrow{s} P$ for any link chain $s$, whose unique solid link is $\ell$,  i.e.\  any $s$ such that $s \blackstretcheq \ell$
(we recall that $s \blackstretcheq \ell$ if $s$ and $\ell$  differ only for the presence of virtual links).
Intuitively, $\ell.P$ can take part in any interaction, in any (admissible) position.
To join in a communication, $\ell.P$ should exhibit the capability to enlarge its link $\ell$ to a link chain $s \blackstretcheq \ell$, whose length is the same as the length of the chains offered by all the other participants, so to proceed with the merge operation. Following the early style, the suitable length is 
%guessed
inferred at the time of deducing the input transition.
Note that, by definition of link chain, if one site of $\ell$ is $\silent$, then $\ell$ can only appear at one of the extremes of $s$.

The rule (\textit{Res}) can serve different aims: 
(i) \emph{floating}, if $a$ does not occur in $s$, then $\restrict{a}s = s$ and  $\restrict{a}P \xrightarrow{s} \restrict{a}P'$;
(ii) \emph{hiding}, if $a$ is matched in $s$ ( i.e.\  $a$ appears as sites already matched by adjacent links), then all occurrences of $a$ in $s$ are transformed to $\silent$ in $\restrict{a}s$;
(iii) \emph{blocking}, if $a$ is pending in $s$ ( i.e.\  there are some
unmatched occurrences of $a$ in $s$), then $\restrict{a}s = \bot$ and the rule cannot be applied.

In the (\textit{Com}) rule the link chains recorded on both the premises' transitions are merged in the conclusion's transition.
This is possible only if $s$ and $s'$ are to some extent ``complementary''.
Contrary to CCS, the rule (\textit{Com}) can appear several times in the proof tree of a transition, because $\merges{s}{s'}$ can still contain virtual links (if $s$ and $s'$ had a virtual link in the same position) and can possibly be merged with other link chains.
However, when $\merges{s}{s'}$ is solid, no further synchronisation is possible (by Lemma~\ref{lemma:solid} (ii)).

%
%A \xrightarrow{a} B. 
%
%
\begin{example}
Let
% $R = \solid{P}$, 
$P = \link{\silent}{a}.P_{1}\ |\ \restrict{b} Q$ and $Q = \link{b}{\silent}.P_{2}\ |\ \link{a}{b}.\nil$, for some processes $P_1$ and $P_2$.
The process $\link{\silent}{a}.P_{1}$ can perform an output on $a$, the process $ \link{b}{\silent}.P_{2}$ can perform an input from $b$; the process $\link{a}{b}$ provides a one-shot link forwarder from $a$ to $b$. 
Their links match along the sequence and a three-party interaction can take place.
Together, these processes can indeed synchronise by agreeing to form the solid link chain
${\startchain{\silent}\chainedlink{a}{a}\chainedlink{\silent}{\silent}\chainend{\silent}}$ of length three, as follows.\footnote{Note that $b$ is restricted and therefore the matched communication on $b$ is replaced by $\tau$ in the observed chain.}

{\small
\begin{center} 
\begin{prooftree} 
\AxiomC{} 
\RightLabel{\scriptsize(\textit{Act})}
\UnaryInfC{$\link{\silent}{a}.P_{1} \xrightarrow{\startchain{\silent}\chainedlink{a}{\noact}\chainedlink{\noact}{\noact}\chainend{\noact}} P_{1}$} 
%
\! \!\!\! \!\!
%\begin{prooftree} 
      \AxiomC{} 
      \RightLabel{\scriptsize(\textit{Act})}
      \UnaryInfC{$\link{b}{\silent}.P_{2} \xrightarrow{\startchain{\noact}\chainedlink{\noact}{\noact}\chainedlink{\noact}{b}\chainend{\silent}} P_{2}$} 
      %
      \AxiomC{} 
      \RightLabel{\scriptsize(\textit{Act})}
      \UnaryInfC{$\link{a}{b}.\nil \xrightarrow{\startchain{\noact}\chainedlink{\noact}{a}\chainedlink{b}{\noact}\chainend{\noact}} \nil$} 
   \RightLabel{\scriptsize(\textit{Com})}
\BinaryInfC{$Q \xrightarrow{\startchain{\noact}\chainedlink{\noact}{a}\chainedlink{b}{b}\chainend{\silent}} P_{2}|\nil$} 
%\end{prooftree} 
%
%\AxiomC{$Q \xrightarrow{\startchain{\noact}\chainedlink{\noact}{a}\chainedlink{b}{b}\chainend{\silent}} P_{2}|\nil$} 
   \RightLabel{\scriptsize(\textit{Res})}
   \UnaryInfC{$\restrict{b}Q \xrightarrow{\startchain{\noact}\chainedlink{\noact}{a}\chainedlink{\silent}{\silent}\chainend{\silent}} \restrict{b} (P_{2}|\nil)$} 
\RightLabel{\scriptsize(\textit{Com})}
\BinaryInfC{$P \xrightarrow{\startchain{\silent}\chainedlink{a}{a}\chainedlink{\silent}{\silent}\chainend{\silent}} P_{1}|\restrict{b} (P_{2}|\nil)$} 
\end{prooftree} 
\end{center}
}

%where
%%
%{\scriptsize
%\begin{center} 
%\begin{prooftree} 
%      \AxiomC{} 
%      \RightLabel{\scriptsize(Act)}
%      \UnaryInfC{$\link{b}{\silent}.P_{2} \xrightarrow{\startchain{\noact}\chainedlink{\noact}{\noact}\chainedlink{\noact}{b}\chainend{\silent}} P_{2}$} 
%      %
%      \AxiomC{} 
%      \RightLabel{\scriptsize(Act)}
%      \UnaryInfC{$\link{a}{b}.\nil \xrightarrow{\startchain{\noact}\chainedlink{\noact}{a}\chainedlink{b}{\noact}\chainend{\noact}} \nil$} 
%   \RightLabel{\scriptsize(Com)}
%\BinaryInfC{$Q \xrightarrow{\startchain{\noact}\chainedlink{\noact}{a}\chainedlink{b}{b}\chainend{\silent}} P_{2}|\nil$} 
%\end{prooftree} 
%\end{center}
%}

%\noindent
%and the suitable chain length, $3$, is inferred and agreed on by all participants.
\end{example}

The following lemma, whose proof goes by rule induction, shows that labels behave like an accordion.
Concretely, any label $s$ in a transition is replaceable with any other chain having a different number of virtual links $\link{\noact}{\noact}$ added to $s$ according to the axioms of $\blackstretcheq$. 
% It is also replaceable with any composition where each $\link{\noact}{\noact}$ inside $s$ is replaced by one or more $\link{\noact}{\noact}$.
The result builds on the previous technical Lemmata~\ref{lem:mergestretch} and~\ref{lemma:res}.
This fact will be later exploited in Section~\ref{abs-sem}, where the abstract semantics is given.

\begin{lemma}[Accordion Lemma]\label{lem:stretchlts}
If $P\xrightarrow{s} P'$ and $s \blackstretcheq s'$,
then $P\xrightarrow{s'} P'$.
\end{lemma}

%\item[(i)]
%If $P\xrightarrow{s} P'$ and $s\link{\noact}{\noact}$ (respectively $\link{\noact}{\noact}s$) is valid,
%then $P\xrightarrow{s\link{\noact}{\noact}} P'$  (respectively $P\xrightarrow{\link{\noact}{\noact}s} P'$).
%\\
%Vice versa, if $P\xrightarrow{s'} P'$ and $s'=s\link{\noact}{\noact}$ or $s'=\link{\noact}{\noact}s$, 
%then $P\xrightarrow{s} P'$.
%
%\item[(ii)]
%If $P\xrightarrow{s_{1}\link{\noact}{\noact}s_{2}} P'$  
%then $P\xrightarrow{s_{1}\startchain{\noact}\chainedlink{\noact}{\noact}\chainend{\noact}s_{2}} P'$.
%\\
%Vice versa, if $P\xrightarrow{s_{1}\startchain{\noact}\chainedlink{\noact}{\noact}\chainend{\noact}s_{2}} P'$ 
%then $P\xrightarrow{s_{1}\link{\noact}{\noact}s_{2}} P'$.
%
%\item[(iii)]
%If $P\xrightarrow{s_{1}\startchain{\alpha}\chainedlink{a}{a}\chainend{\beta}s_{2}} P'$ 
%then $P\xrightarrow{s_{1}\startchain{\alpha}\chainedlink{a}{\noact}\chainedlink{\noact}{a}\chainend{\beta}s_{2}} P'$.
%\\
%Vice versa, if $P\xrightarrow{s_{1}\startchain{\alpha}\chainedlink{a}{\noact}\chainedlink{\noact}{a}\chainend{\beta}s_{2}} P'$
%then $P\xrightarrow{s=s_{1}\startchain{\alpha}\chainedlink{a}{a}\chainend{\beta}s_{2}} P'$.

%\begin{lemma}
%If $P\xrightarrow{s} P'$, 
%then $P\xrightarrow{s\link{\noact}{\noact}} P'$ and $P\xrightarrow{\link{\noact}{\noact}s} P'$.

%\noindent
%Vice versa, if $P\xrightarrow{s'} P'$ and $s'=s\link{\noact}{\noact}$ or $s'=\link{\noact}{\noact}s$, 
%then $P\xrightarrow{s} P'$.

%\end{lemma}

%\begin{lemma}
%If $P\xrightarrow{s_{1}\link{\noact}{\noact}s_{2}} P'$  
%then $P\xrightarrow{s_{1}\startchain{\noact}\chainedlink{\noact}{\noact}\chainend{\noact}s_{2}} P'$ .
%%
%Vice versa, if $P\xrightarrow{s_{1}\startchain{\noact}\chainedlink{\noact}{\noact}\chainend{\noact}s_{2}} P'$ 
%then $P\xrightarrow{s_{1}\link{\noact}{\noact}s_{2}} P'$.
%\end{lemma}

%
%\begin{lemma}
%If $P\xrightarrow{s_{1}\startchain{\alpha}\chainedlink{a}{a}\chainend{\beta}s_{2}} P'$ 
%then $P\xrightarrow{s_{1}\startchain{\alpha}\chainedlink{a}{\noact}\chainedlink{\noact}{a}\chainend{\beta}s_{2}} P'$.
%%
%Vice versa, if $P\xrightarrow{s_{1}\startchain{\alpha}\chainedlink{a}{\noact}\chainedlink{\noact}{a}\chainend{\beta}s_{2}} P'$
%then $P\xrightarrow{s=s_{1}\startchain{\alpha}\chainedlink{a}{a}\chainend{\beta}s_{2}} P'$.

%\end{lemma}

%\textcolor{blue}{METTIAMO QUI O ALTROVE LE SEGUENTI PRECISAZIONI MESSE IN SciAnnals?
%To better focus on the way in which the merge of link chains is performed, we present a simple
%example.
%We recall that the definition of the $\merges{}{}$ operator requires that the two involved link chains must be of the same length, and that, by rule \textit{(Act)}
%we can arbitrarily add virtual links, in any position.
%%, to get the point. \\
%%
%Let 
%$s_1 = \link{\silent}{b}$, $s_2 = \link{b}{c}$, and  $s_3 = \link{c}{\tau}$ be three links, then there are three possible combinations for applying the $\merges{}{}$ operator:
%\begin{itemize}
%\item  we can first define
% $s_1' = \startchain{\silent}\chainedlink{b}{\noact}\chainedlink{\noact}{\noact}\chainend{\noact}$, 
% $s_2' = \startchain{\noact}\chainedlink{\noact}{b}\chainedlink{c}{\noact}\chainend{\noact}$, and compute 
% $s_4 = \merges{s_1'}{s_2'}=\startchain{\tau}\chainedlink{b}{b}\chainedlink{c}{\noact}\chainend{\noact}$;
% then we can define 
% $s_3' =  \startchain{\silent}\chainedlink{b}{b}\chainedlink{c}{c}  \chainend{\silent}$, and, finally, we can compute 
% $s = \merges{s_4}{s_3'} =\startchain{\silent}\chainedlink{b}{\noact}\chainedlink{\noact}{c}
%\chainend{\silent}$; or
%%
%\item we can first define $s_1' = \startchain{\silent}\chainedlink{b}{\noact}\chainedlink{\noact}{\noact}\chainend{\noact}$, 
%$s_3' = \startchain{\noact}\chainedlink{\noact}{\noact}\chainedlink{\noact}{c}\chainend{\silent}$, and compute 
%$s_4 = \merges{s_1'}{s_3'} =  \startchain{\silent}\chainedlink{b}{\noact}\chainedlink{\noact}{c}\chainend{\silent}$; then we can define 
%$s_2' = \startchain{\noact}\chainedlink{\noact}{b}\chainedlink{c}{\noact}\chainend{\noact}$, 
%and finally we can compute 
%$s=\merges{s_2'}{s_4}= \startchain{\silent}\chainedlink{b}{b}\chainedlink{c}{c}
%\chainend{\silent}$; or
%%
%\item we can first define 
%$s_3' = \startchain{\noact}\chainedlink{\noact}{\noact}\chainedlink{\noact}{c}\chainend{\silent}$, 
%$s_2' = \startchain{\noact}\chainedlink{\noact}{b}\chainedlink{c}{\noact}\chainend{\noact}$, 
%and compute 
%$s_4 = \merges{s_2'}{s_3'}= \startchain{\noact}\chainedlink{\noact}{b}\chainedlink{c}{c}\chainend{\silent}$; 
%then we can define 
%$s_1' = \startchain{\silent}\chainedlink{b}{\noact}\chainedlink{\noact}{\noact}\chainend{\noact}$; 
%and finally we can compute 
%$s=\merges{s_1'}{s_4}= \startchain{\silent}\chainedlink{b}{b}\chainedlink{c}{c}\chainend{\silent}$.
%% $s_1' =  \link{\silent}{b}\link{*}{*}\link{*}{*}$, $s_2' = \link{*}{*}\link{b}{c}\link{*}{*}$, and compute $s_4 = \merges{s_1'}{s_2'}=\link{\silent}{b}\link{b}{c}\link{*}{*} $; then we can define $s_3' =  \link{*}{*}\link{*}{*} \link{c}{\silent}$, and, finally, we can compute $s = \merges{s_4}{s_3'} =\link{\silent}{b}\link{b}{c} \link{c}{\tau}$; or
%%\item we can first define $s_1' = \link{\silent}{b}\link{*}{*} \link{*}{*}$, $s_3' =  \link{*}{*}\link{*}{*} \link{c}{\silent}$, and compute $s_4 = \merges{s_1'}{s_3'} =  \link{\silent}{b}\link{*}{*} \link{c}{\silent}$; then we can define $s_2' = \link{*}{*}\link{b}{c}\link{*}{*}$, and finally we can compute $s=\merges{s_2'}{s_4}= \link{\silent}{b}\link{b}{c} \link{c}{\silent} $; or
%%\item we can first define $s_3' = \link{*}{*}\link{*}{*}\link{c}{\silent}$, $s_2' = \link{*}{*}\link{b}{c}\link{*}{*}$, 
%%and compute $s_4 = \merges{s_2'}{s_3'}= \link{*}{*}\link{b}{c}\link{c}{\silent}$; then we can
%%define $s_1' = \link{\silent}{b}\link{*}{*}\link{*}{*}$; and finally we can compute $s=\merges{s_1'}{s_4}= \link{\silent}{b}\link{b}{c}\link{c}{\silent}$.
%\end{itemize}
%}

\begin{example}[Forwarders]
We give a few examples to show how flexible is \CNA\ for defining ``forwarding'' policies.
We have already seen a one-shot and one-hop forwarder from $a$ to $b$ that can be written as $\link{a}{b}.\nil$.
Its persistent version is just written as $R(a,b) \defeq \link{a}{b}.R(a,b)$.
Moreover, the process $R(a,b) \ | R(b,a)$ provides a sort of name fusion, making $a$ and $b$ interchangeable.

% \textcolor{cyan}{(Hereafter, we denote a tuple $a_1,...,a_n$ with $\tilde{a}$).}
An alternating forwarder $A(a,b,c)$ from $a$ to $b$ first and then to $c$ can be defined as 
\[A(a,b,c) \defeq \link{a}{b}.\link{a}{c}.A(a,b,c).\]

A persistent non-deterministic forwarder $C(a,\widetilde{c})$ (the $C$ stands for \emph{choice}), from $a$ to $c_{1},...,c_{n}$ can be \mbox{written as} 
\[C(a,\widetilde{c}) \defeq \link{a}{c_{1}}.C(a,\widetilde{c})+ \cdots + \link{a}{c_{n}}.C(a,\widetilde{c}).\]

Similarly, $J(\widetilde{b},a)$ defined as (the $J$ stands for \emph{join}\footnote{With the term ``join'' we refer to the fact that messages from different sources are forwarded to the same channel. It has no relation with join patterns in join calculus.})
\[J(\widetilde{b},a)\defeq \link{b_{1}}{a}.J(\widetilde{b},a) + \cdots + \link{b_{m}}{a}.J(\widetilde{b},a)\] 
is a persistent non-deterministic forwarder, from $b_{1},...,b_{m}$ to $a$.

By combining the two processes above as in
\[F(\widetilde{b},\widetilde{c}) \defeq \restrict{a}(J(\widetilde{b},a)\ |\ C(a,\widetilde{c}))\]
we obtain a persistent forwarder from any of  the $b_{i}$'s  to any of the $c_{j}$'s.
The only admissible transitions have indeed the form
$
F(\widetilde{b},\widetilde{c})  \xrightarrow{s} F(\widetilde{b},\widetilde{c})
$
with $s \blackstretcheq \startchain{b_i}\chainedlink{\silent}{\silent}\chainend{c_j}$ for some $i\in[1,m]$ and $j\in[1,n]$ (because interaction on $a$ is restricted and $\restrict{a}(\startchain{b_i}\chainedlink{a}{a}\chainend{c_j}) = \startchain{b_i}\chainedlink{\silent}{\silent}\chainend{c_j}$).
%As an advantage, note that the combinatorial explosion is taken care of by the communication rules, and it does not affect the ``size'' of the processes.
%By combining the processes $\restrict{a}(P^{d,e}_{a} | A^{a}_{b,c})$, then we obtain a persistent forwarder from either $d$ or $e$ to $b$ and $c$ in alternation.
\end{example}

\begin{example}[Blind routing in \CNA]
We have already observed that CCS processes can be transformed to \CNA\ processes just by transforming action prefixes in link prefixes. Correspondingly, the blind routing example from Section~\ref{sec:CCS} (Example~\ref{blind}) can be encoded in \CNA\ just as follows, where for readability we have omitted the parameters from definitions.
%
\begin{eqnarray*}
A_i & \defeq & \link{\tau}{\mathit{req}_i}.\link{\tau}{\mathit{think}}.A_i\quad \mbox{ for $i\in[1,2]$} \\
S_j & \defeq & \link{\mathit{srv}_j}{\tau}.\link{\tau}{\mathit{exec}}.S_j + \link{\tau}{\mathit{busy}}.\link{\tau}{\tau}.S_j\quad \mbox{ for $j\in[1,2]$}\\
R & \defeq & \link{\mathit{req}_1}{\tau}.(\link{\tau}{\mathit{srv}_1}.R + \link{\tau}{\mathit{srv}_2}.R) + \link{\mathit{req}_2}{\tau}.\link{\tau}{\mathit{srv}_2}.R \\
N & \defeq &
\restrict{\widetilde{\mathit{req}}} 
\restrict{\widetilde{\mathit{srv}}}
(A_1\ |\ A_2\ |\ R\ |\ S_1\ |\ S_2)
\end{eqnarray*}

For example, the following transitions can be derived from the SOS rules accounting for the case where a request from $A_1$ is assigned to $S_2$:
\begin{eqnarray*}
N
& \xrightarrow{ \startchain{\tau}\chainedlink{\tau}{\tau}\chainend{\tau} } &
\restrict{\widetilde{\mathit{req}}} 
\restrict{\widetilde{\mathit{srv}}}
(\link{\tau}{\mathit{think}}.A_1\ |\ A_2\ |\ (\link{\tau}{\mathit{srv}_1}.R + \link{\tau}{\mathit{srv}_2}.R)\ |\ S_1\ |\ S_2)
\\
& \xrightarrow{ \startchain{\tau}\chainedlink{\tau}{\tau}\chainend{\tau} } &
\restrict{\widetilde{\mathit{req}}} 
\restrict{\widetilde{\mathit{srv}}}
(\link{\tau}{\mathit{think}}.A_1\ |\ A_2\ |\ R\ |\ S_1\ |\ \link{\tau}{\mathit{exec}}.S_2)
\\
& \xrightarrow{ \link{\tau}{\mathit{think}} } &
\restrict{\widetilde{\mathit{req}}} 
\restrict{\widetilde{\mathit{srv}}}
(A_1\ |\ A_2\ |\ R\ |\ S_1\ |\ \link{\tau}{\mathit{exec}}.S_2)
\\
& \xrightarrow{ \link{\tau}{\mathit{exec}} } &
N
\end{eqnarray*}

Note that, in the first two steps, interactions on channels $\mathit{req}_1$ and $\mathit{srv}_2$ are restricted and thus not observable, in fact $\restrict{\mathit{req}_1}(\startchain{\tau}\chainedlink{\mathit{req}_1}{\mathit{req}_1}\chainend{\tau}) = \startchain{\tau}\chainedlink{\tau}{\tau}\chainend{\tau}$ and $\restrict{\mathit{srv}_2}(\startchain{\tau}\chainedlink{\mathit{srv}_2}{\mathit{srv}_2}\chainend{\tau}) = \startchain{\tau}\chainedlink{\tau}{\tau}\chainend{\tau}$.

\end{example}

\begin{example}[Acknowledged routing in \CNA]
The features of \CNA\ become evident in our running example when we come to modelling acknowledged routing  (Example~\ref{ack}). This is because the explicit acknowledgment is no longer necessary as it can be implicitly accounted for by the ability to construct a chain of links.\footnote{Of course, one could also just encode the CCS processes for acknowledged routing just by translating their prefixes as we have done for blind routing.}
Correspondingly, we set
\begin{eqnarray*}
A_i & \defeq & \link{\tau}{\mathit{req}_i}.\link{\tau}{\mathit{think}}.A_i\quad \mbox{ for $i\in[1,2]$} \\
S_j & \defeq & \link{\mathit{srv}_j}{\tau}.\link{\tau}{\mathit{exec}}.S_j + \link{\tau}{\mathit{busy}}.\link{\tau}{\tau}.S_j\quad \mbox{ for $j\in[1,2]$}\\
R & \defeq & \link{\mathit{req}_1}{\mathit{srv}_1}.R + \link{\mathit{req}_1}{\mathit{srv}_2}.R + \link{\mathit{req}_2}{\mathit{srv}_2}.R \\
M & \defeq &
\restrict{\widetilde{\mathit{req}}} 
\restrict{\widetilde{\mathit{srv}}}
(A_1\ |\ A_2\ |\ R\ |\ S_1\ |\ S_2)
\end{eqnarray*}
Note that channels $\mathit{ack}_i$ are not needed and \CNA\ processes $A_i$ and $S_j$ are defined as in the case of blind routing; only the syntax of $R$ has been changed to link, in one single step, the requests from agents with the availability of servers.

For example, the following transitions can be derived from the SOS rules accounting for the case where a request from $A_1$ is assigned to $S_2$:
\begin{eqnarray*}
M
& \xrightarrow{ \startchain{\tau}\chainedlink{\tau}{\tau}\chainedlink{\tau}{\tau}\chainend{\tau} } &
\restrict{\widetilde{\mathit{req}}} 
\restrict{\widetilde{\mathit{srv}}}
(\link{\tau}{\mathit{think}}.A_1\ |\ A_2\ |\ R\ |\ S_1\ |\ \link{\tau}{\mathit{exec}}.S_2)
\\
& \xrightarrow{ \link{\tau}{\mathit{think}} } &
\restrict{\widetilde{\mathit{req}}} 
\restrict{\widetilde{\mathit{srv}}}
(A_1\ |\ A_2\ |\ R\ |\ S_1\ |\ \link{\tau}{\mathit{exec}}.S_2)
\\
& \xrightarrow{ \link{\tau}{\mathit{exec}} } &
M
\end{eqnarray*}

Note that, in the first step, the interaction on channels $\mathit{req}_1$ and $\mathit{srv}_2$ is restricted and thus not observable, in fact $\restrict{\mathit{req}_1}\restrict{\mathit{srv}_2}(\startchain{\tau}\chainedlink{\mathit{req}_1}{\mathit{req}_1}\chainedlink{\mathit{srv}_2}{\mathit{srv}_2}\chainend{\tau}) = \startchain{\tau}\chainedlink{\tau}{\tau}\chainedlink{\tau}{\tau}\chainend{\tau}$.

\end{example}

\begin{example}[Composite, acknowledged routing in \CNA]
Notably, the infrastructure presented in the previous example is already designed in a modular way: infrastructures can be composed without requiring any change and no dead path detection or backtracking mechanisms have to be put in place, because they are taken care by the  \CNA\ ``middleware''.

For instance, take the \CNA\ versions of the infrastructures presented in Example~\ref{composite}:
\begin{eqnarray*}
R' & \defeq & \link{\mathit{req}_1}{s_1}.R' + \link{\mathit{req}_1}{s_2}.R' + \link{\mathit{req}_2}{s_2}.R' \\
R'' & \defeq & \link{s_1}{s'_1}.R'' + \link{s_2}{s'_2}.R' \\
R''' & \defeq & \link{s'_2}{\mathit{srv}_2}.R''' \\
R & \defeq & \restrict{\widetilde{s'}}\restrict{\widetilde{s}} (R'\ |\ R''\ |\ R''')
\end{eqnarray*}

Besides looking considerably more concise than their CCS versions, they make evident that the delivery of any request to a server is atomic as any process executes one (link) action and reduces (recursively) to itself.

At a closer inspection, one may notice that the only possible transitions for $R$ are the following ones 
(up to $\blackstretcheq$, as explained by the Accordion Lemma~\ref{lem:stretchlts}):
\[
R \xrightarrow{\startchain{\mathit{req}_1}\chainedlink{\tau}{\tau}\chainedlink{\tau}{\tau}\chainend{\mathit{srv}_2}} R
\quad\mbox{ and }\quad
R \xrightarrow{\startchain{\mathit{req}_2}\chainedlink{\tau}{\tau}\chainedlink{\tau}{\tau}\chainend{\mathit{srv}_2}} R.
\]

As in the previous examples, note that, e.g.\ in the first step, the interaction on channels $s_2$ and $s'_2$ is restricted and thus not observable, in fact
$\restrict{s_2}\restrict{s'_2} \startchain{\mathit{req}_1}\chainedlink{s_2}{s_2}\chainedlink{s'_2}{s'_2}\chainend{\mathit{srv}_2}= 
\startchain{\mathit{req}_1}\chainedlink{\tau}{\tau}\chainedlink{\tau}{\tau}\chainend{\mathit{srv}_2}$.

Therefore, at a suitable level of abstraction, in which routing details are omitted, we would like to relate the composite infrastructure $R$ with the monolithic infrastructure $R_m$ defined as follows:
\begin{eqnarray*}
R_m & \defeq & \link{\mathit{req}_1}{\mathit{srv}_2}.R_m + \link{\mathit{req}_2}{\mathit{srv}_2}.R_m
\end{eqnarray*}

In the next section we show how this can be formalised.
\end{example}

### 2020-07-27 fragments/old/05-translation_OLD.tex
da dropbox

% !TEX root = ./main_RS2L.tex

\section{From Reaction Systems to c\CNA}
\label{sec:trans}

Here we present a translation from the Reaction Systems~\cite{} to a variant of the \CNA, the c\CNA, where
the prefixes are link chains and no simple links.

\begin{definition}
\label{def:trans}
Let ${\cal A} = (S, A)$ be a reaction system, then its  translation, $P=[|{\cal A}|]$ into the {\tt link}-calculus is defined as follows 
\begin{center}$P = \restrict{stop,r_1,\dots,r_m}(\Pi_{s\in S} P_s | \Pi_{i\in A} P_a| STOP)$,\end{center} 
with $m$ be the number of reactions, and where:
\begin{itemize}
\item for each biological element $s \in S$ we write:\\
\[
\begin{array}{lcl}
%P_s& \defeq &\sum_{h=1}^{{\cal N}_s}\underbrace{ \startchain{a_i}\chainedlink{a_o}{\noact}\chainedlink{\noact}{a_i}\chainedlink{a_o}{\noact}\chainend{\noact}...}_{size = h}.\nil + 
%\\
%&&| \ \\
%&& (  \sum_{h=1}^{{\cal P}_s}\underbrace{ \startchain{\tilde{a}_i}\chainedlink{\tilde{a}_o}{\noact}\chainedlink{\noact}{\tilde{a}_i}\chainedlink{\tilde{a}_o}{\noact}\chainend{\noact}...}_{size = h}\link{\hat{a}_i}{\hat{a}_o}.P_s
% \ + \ \link{\hat{a}_i}{\hat{a}_o}.\overline{P_s})
\\
P_s& \defeq &\sum_{h\geq 0} \sum_{k\geq 1}  \underbrace{ \startchain{s_i}\chainedlink{s_o}{\noact}\chainedlink{\noact}{s_i}\chainedlink{s_o}{\noact}\chainend{\noact}...}_{h\ \mathrm{times}}\underbrace{ \startchain{\tilde{s}_i}\chainedlink{\tilde{s}_o}{\noact}\chainedlink{\noact}{\tilde{s}_i}\chainedlink{\tilde{s}_o}{\noact}\chainend{\noact}...}_{k\ \mathrm{times}}\link{\hat{s}_i}{\hat{s}_o}.P_s\\
&&\ +\\
&&\sum_{h\geq 0}   \underbrace{ \startchain{s_i}\chainedlink{s_o}{\noact}\chainedlink{\noact}{s_i}\chainedlink{s_o}{\noact}\chainend{\noact}...}_{h\ \mathrm{times}}\link{\hat{s}_i}{\hat{s}_o}.\overline{P}_s

% \ + \ \link{\hat{a}_i}{\hat{a}_o}.\overline{P_s})
 \\
 \overline{P}_s& \defeq &\sum_{h\geq 0} \sum_{k\geq 1} \underbrace{ \startchain{\overline{s}_i}\chainedlink{\overline{s}_o}{\noact}\chainedlink{\noact}{\overline{s}_i}\chainedlink{\overline{s}_o}{\noact}\chainend{\noact}...}_{h\ \mathrm{times}}\underbrace{ \startchain{\tilde{s}_i}\chainedlink{\tilde{s}_o}{\noact}\chainedlink{\noact}{\tilde{s}_i}\chainedlink{\tilde{s}_o}{\noact}\chainend{\noact}...}_{k\ \mathrm{times}}.\link{\hat{s}_i}{\hat{s}_o}.P_s\\
 && \ +\\
 && \sum_{h\geq 0}  \underbrace{ \startchain{\overline{s}_i}\chainedlink{\overline{s}_o}{\noact}\chainedlink{\noact}{\overline{s}_i}\chainedlink{\overline{s}_o}{\noact}\chainend{\noact}...}_{h\ \mathrm{times}}.\link{\hat{s}_i}{\hat{s}_o}.\overline{P}_s\\
 %\ + \ \link{\hat{a}_i}{\hat{a}_o}.\overline{P_s})

\end{array}
\]
\end{itemize}
\begin{itemize}
\item we number each rule in a way each $a \in A$ has associated a progressive number $j$. Then, for each rule $a \in A$ with number $j$, we write:\\
 \[
%\begin{array}{lcl}
%R_i& \defeq & \startchain{rule_i}\chainedlink{a_i}{\noact}\chainedlink{\noact}{a_o}\chainedlink{\overline{a}_i}{\noact} \chainedlink{\noact}{\overline{a}_o}\chainedlink{\tilde{a}_i}{\noact}\chainedlink{\noact}{\tilde{a}_o} \chainend{rule_{i+1}}.R_i\\
%&& +\  \\
%&& \startchain{rule_i}\chainedlink{a_i}{\noact}\chainedlink{\noact}{a_o}\chainedlink{a_i}{\noact} \chainedlink{\noact}{a_o} \chainend{rule_{i+1}}.R_i\\
%&&+\\
%&& \startchain{rule_i}\chainedlink{\overline{a}_i}{\noact}\chainedlink{\noact}{\overline{a}_o}\chainedlink{\overline{a}_i}{\noact} \chainedlink{\noact}{\overline{a}_o} \chainend{rule_{i+1}}.R_i\\
%&& +\\
%&& \startchain{rule_i}\chainedlink{\overline{a}_i}{\noact}\chainedlink{\noact}{\overline{a}_o}\chainedlink{a_i}{\noact} \chainedlink{\noact}{a_o} \chainend{rule_{i+1}}.R_i\\
%\end{array}
\begin{array}{lcl}
A_j& \defeq & \startchain{r_j}
      \underbrace{\chainedlink{s_{1i}}{\noact}\chainedlink{\noact}{s_{1o}}\chainedlink{s_{2i}}{\noact}\ldots\chainedlink{\noact}{s_{ko}}}_{R_j=\{s_{1},...,s_{k}\}}
      \underbrace{\chainedlink{\overline{b}_{1i}}{\noact} \chainedlink{\noact}{\overline{b}_{1o}}
      \chainedlink{\overline{b}_{2i}}{\noact} \ldots \chainedlink{\noact}{\overline{b}_{ho}}}_{I_j=\{b_1,...,b_h\}}
      \underbrace{\chainedlink{\tilde{c}_{1i}}{\noact}\chainedlink{\noact}{\tilde{c}_{1o}} \chainedlink{\tilde{c}_{2i}}{\noact}\ldots \chainedlink{\noact}{\tilde{c}_{\ell o}}}_{P_j=\{c_1,...,c_\ell\}}
      \chainend{r_{j+1}}.A_j\\
&& +\  \\
&& \sum_{s\in R_j}\startchain{r_j}\chainedlink{\overline{s}_i}{\noact}\chainedlink{\noact}{\overline{s}_o} \chainend{r_{j+1}}.A_j\\
&&+\\
&& \sum_{b\in I_j}\startchain{r_j}\chainedlink{b_i}{\noact} \chainedlink{\noact}{b_o} \chainend{r_{i+1}}.A_i\\
%&& +\\
%&& \startchain{rule_i}\chainedlink{\overline{a}_i}{\noact}\chainedlink{\noact}{\overline{a}_o}\chainedlink{a_i}{\noact} \chainedlink{\noact}{a_o} \chainend{rule_{i+1}}.R_i\\
\end{array}
\]
\end{itemize}
where $s$ is the meta-variable for the elements required by the rule,  $b$ is the meta-variable for  the elements that the rule requires to be absent, and  $c$ is the metavariable for the elements produced.

For notational convenience, we fix that $r_1 = \tau$ and $r_{m+1} = stop$ for $m$ the number of reactions.

Eventually, the unique process $STOP$ is defined as\\
\[
\begin{array}{lcl}
STOP &\defeq& \startchain{stop}\chainedlink{\hat{s1}_{i}}{\noact}\chainedlink{\noact}{\hat{s1}_{o}}\chainedlink{\hat{s2}_{i}}{\noact}... \chainedlink{\noact}{\hat{sn}_{o}} \chainend{\silent}. STOP
\end{array}
\]
where the $n \in \mathbb{N}$ is the number of the elements in the system.
\end{definition}

 \begin{example}
 Let  $\mathcal{A} =(S,A) $ be a reaction system with $S= \{e1,e2,e3\}$, and $A=\{(e1|e3|e1),(e2|e1|e3),(e3|e2|e2) \}$.
 The code of the resulting {\tt link}-calculus process $ \Pi_{s\in S} P_s | \Pi_{i\in A} P_a| STOP = [|(S,A) |] $ is the following. \\
 The code for the elements:\\
%  \[
% \begin{array}{lcl}
% S1 &\defeq& (\startchain{e1_i}\chainedlink{e1_o}{\noact}\chainedlink{\noact} \\
%  +\startchain{e1_i}\chainedlink{e1_o}{\noact}\chainedlink{\noact}{e1_i}\chainend{e1_o}.\nil) \  |\ 
% (\startchain{\tilde{e1}_i}\chainedlink{\tilde{e1}_o}{\noact}\chainedlink{\noact}{\hat{e1}_i}\chainend{\hat{e1}_o}.S1 + \link{\hat{e1}_i}{\hat{e1}_o}.\overline{S1})\\
%\\
%\overline{S1} &\defeq& (\link{\overline{e1}_i}{{e1}_o}.\nil +\startchain{\overline{e1}_i}\chainedlink{\overline{e1}_o}{\noact}\chainedlink{\noact}{\overline{e1}_i}\chainend{\overline{e1}_o}.\nil) \  |\ 
% (\startchain{\tilde{e1}_i}\chainedlink{\tilde{e1}_o}{\noact}\chainedlink{\noact}{\hat{e1}_i}\chainend{\hat{e1}_o}.S1 + \link{\hat{e1}_i}{\hat{e1}_o}.\overline{S1})\\ 
% \\
% \\
%  S2 &\defeq& (\link{e2_i}{e2_o}.\nil +\startchain{e2_i}\chainedlink{e2_o}{\noact}\chainedlink{\noact}{e2_i}\chainend{e2_o}.\nil) \  |\ 
% (\startchain{\tilde{e2}_i}\chainedlink{\tilde{e2}_o}{\noact}\chainedlink{\noact}{\hat{e2}_i}\chainend{\hat{e2}_o}.S1 + \link{\hat{e2}_i}{\hat{e2}_o}.\overline{S2})\\
%\\
%\overline{S2} &\defeq& (\link{\overline{e2}_i}{{e2}_o}.\nil +\startchain{\overline{e2}_i}\chainedlink{\overline{e2}_o}{\noact}\chainedlink{\noact}{\overline{e2}_i}\chainend{\overline{e2}_o}.\nil) \  |\ 
% (\startchain{\tilde{e2}_i}\chainedlink{\tilde{e2}_o}{\noact}\chainedlink{\noact}{\hat{e2}_i}\chainend{\hat{e2}_o}.S1 + \link{\hat{e2}_i}{\hat{e2}_o}.\overline{S2})\\ 
% \\
% \\
%   S3 &\defeq& (\link{e3_i}{e3_o}.\nil +\startchain{e3_i}\chainedlink{e3_o}{\noact}\chainedlink{\noact}{e3_i}\chainend{e3_o}.\nil) \  |\ 
% (\startchain{\tilde{e3}_i}\chainedlink{\tilde{e3}_o}{\noact}\chainedlink{\noact}{\hat{e3}_i}\chainend{\hat{e3}_o}.S1 + \link{\hat{e3}_i}{\hat{e3}_o}.\overline{S3})\\
%\\
%\overline{S3} &\defeq& (\link{\overline{e3}_i}{{e3}_o}.\nil +\startchain{\overline{e3}_i}\chainedlink{\overline{e3}_o}{\noact}\chainedlink{\noact}{\overline{e3}_i}\chainend{\overline{e3}_o}.\nil) \  |\ 
% (\startchain{\tilde{e3}_i}\chainedlink{\tilde{e3}_o}{\noact}\chainedlink{\noact}{\hat{e3}_i}\chainend{\hat{e3}_o}.S1 + \link{\hat{e3}_i}{\hat{e3}_o}.\overline{S3})\\ 
% \end{array}
% \]

 \[
 \begin{array}{lcl}
 S1 &\defeq& \startchain{s1_i}\chainedlink{s1_o}{\noact}\chainedlink{\noact}{\tilde{s1}_i}\chainedlink{\tilde{s1}_o}{\noact}\chainedlink{\noact}{\hat{s1}_i}\chainend{\hat{s1}_o}.S1 \ + \
 \startchain{s1_i}\chainedlink{s1_o}{\noact}\chainedlink{\noact}{s1_i}\chainedlink{s1_o}{\noact}\chainedlink{\noact}{\tilde{s1}_i}\chainedlink{\tilde{s1}_o}{\noact}\chainedlink{\noact}{\hat{s1}_i}\chainend{\hat{s1}_o}.S1 \ + \ \startchain{\tilde{s1}_i}\chainedlink{\tilde{s1}_o}{\noact}\chainedlink{\noact}{\hat{s1}_i}\chainend{\hat{s1}_o}.S1\\
 && + \\
 &&  \startchain{s1_i}\chainedlink{s1_o}{\noact}\chainedlink{\noact}{\hat{s1}_i}\chainend{\hat{s1}_o}.\overline{S1} 
\  + \ \startchain{s1_i}\chainedlink{s1_o}{\noact}\chainedlink{\noact}{s1_i}\chainedlink{s1_o}{\noact}\chainedlink{\noact}{\hat{s1}_i}\chainend{\hat{s1}_o}.\overline{S1}\ + \  \link{\hat{s1}_i}{\hat{s1}_o}.\overline{S1}\\
\\
\overline{S1} &\defeq& \startchain{\overline{s1}_i}\chainedlink{\overline{s1}_o}{\noact}\chainedlink{\noact}{\tilde{s1}_i}\chainedlink{\tilde{s1}_o}{\noact}\chainedlink{\noact}{\hat{s1}_i}\chainend{\hat{s1}_o}.S1 \ + \
 \startchain{\overline{s1}_i}\chainedlink{\overline{s1}_o}{\noact}\chainedlink{\noact}{\overline{s1}_i}\chainedlink{\overline{s1}_o}{\noact}\chainedlink{\noact}{\tilde{s1}_i}\chainedlink{\tilde{s1}_o}{\noact}\chainedlink{\noact}{\hat{s1}_i}\chainend{\hat{s1}_o}.S1 \ + \ \startchain{\tilde{s1}_i}\chainedlink{\tilde{s1}_o}{\noact}\chainedlink{\noact}{\hat{s1}_i}\chainend{\hat{s1}_o}.S1\\
 && + \\
 &&  \startchain{\overline{s1}_i}\chainedlink{\overline{s1}_o}{\noact}\chainedlink{\noact}{\overline{s1}_i}\chainedlink{\overline{s1}_o}{\noact}\chainedlink{\noact}{\hat{s1}_i}\chainend{\hat{s1}_o}.\overline{S1} 
\  + \ \startchain{\overline{s1}_i}\chainedlink{\overline{s1}_o}{\noact}\chainedlink{\noact}{\hat{s1}_i}\chainend{\hat{s1}_o}.\overline{S1}\ + \  \link{\hat{s1}_i}{\hat{s1}_o}.\overline{S1}\\
 \\
 \\
 S2 & \defeq &\sum_{h=0}^{2} \sum_{k=0}^{2}  \underbrace{ \startchain{s2_i}\chainedlink{s2_o}{\noact}\chainedlink{\noact}{s2_i}\chainedlink{s2_o}{\noact}\chainend{\noact}...}_{size = h}\underbrace{ \startchain{\tilde{s2}_i}\chainedlink{\tilde{s2}_o}{\noact}\chainedlink{\noact}{\tilde{s2}_i}\chainedlink{\tilde{s2}_o}{\noact}\chainend{\noact}...}_{size = k}\link{\hat{s2}_i}{\hat{s2}_o}.S2\\
&&\ +\\
&&\sum_{h=0}^{2}   \underbrace{ \startchain{s2_i}\chainedlink{s2_o}{\noact}\chainedlink{\noact}{s_i}\chainedlink{s2_o}{\noact}\chainend{\noact}...}_{size = h}\link{\hat{s2}_i}{\hat{s2}_o}.\overline{S2}

% \ + \ \link{\hat{a}_i}{\hat{a}_o}.\overline{P_s})
 \\
 \overline{S2}& \defeq &\sum_{h=0}^{2} \sum_{k=1}^{2} \underbrace{ \startchain{\overline{s2}_i}\chainedlink{\overline{s2}_o}{\noact}\chainedlink{\noact}{\overline{s2}_i}\chainedlink{\overline{s2}_o}{\noact}\chainend{\noact}...}_{size = h}\underbrace{ \startchain{\tilde{s2}_i}\chainedlink{\tilde{s2}_o}{\noact}\chainedlink{\noact}{\tilde{s2}_i}\chainedlink{\tilde{s2}_o}{\noact}\chainend{\noact}...}_{size = k}.\link{\hat{s2}_i}{\hat{s2}_o}.S2\\
 && \ +\\
 && \sum_{h=0}^{2}  \underbrace{ \startchain{\overline{s2}_i}\chainedlink{\overline{s2}_o}{\noact}\chainedlink{\noact}{\overline{s2}_i}\chainedlink{\overline{s2}_o}{\noact}\chainend{\noact}...}_{size = h}.\link{\hat{s2}_i}{\hat{s2}_o}.\overline{S2}\\
% 
% 
%   S2 &\defeq& \startchain{e2_i}\chainedlink{e2_o}{\noact}\chainedlink{\noact}{\tilde{e2}_i}\chainedlink{\tilde{e2}_o}{\noact}\chainedlink{\noact}{\hat{e2}_i}\chainend{\hat{e2}_o}.S2 \ + \
% \startchain{e2_i}\chainedlink{e2_o}{\noact}\chainedlink{\noact}{e2_i}\chainedlink{e2_o}{\noact}\chainedlink{\noact}{\tilde{e2}_i}\chainedlink{\tilde{e2}_o}{\noact}\chainedlink{\noact}{\hat{e2}_i}\chainend{\hat{e2}_o}.S2 \ + \ \startchain{\tilde{e2}_i}\chainedlink{\tilde{e2}_o}{\noact}\chainedlink{\noact}{\hat{e2}_i}\chainend{\hat{e2}_o}.S2\\
% && + \\
% &&  \startchain{e2_i}\chainedlink{e2_o}{\noact}\chainedlink{\noact}{\hat{e2}_i}\chainend{\hat{e2}_o}.\overline{S2} 
%\  + \ \startchain{e2_i}\chainedlink{e2_o}{\noact}\chainedlink{\noact}{e2_i}\chainedlink{e2_o}{\noact}\chainedlink{\noact}{\hat{e2}_i}\chainend{\hat{e2}_o}.\overline{S2}\ + \  \link{\hat{e2}_i}{\hat{e2}_o}.\overline{S2}\\
%\\
%\overline{S2} &\defeq& \startchain{\overline{e2}_i}\chainedlink{\overline{e2}_o}{\noact}\chainedlink{\noact}{\tilde{e2}_i}\chainedlink{\tilde{e2}_o}{\noact}\chainedlink{\noact}{\hat{e2}_i}\chainend{\hat{e2}_o}.S2 \ + \
% \startchain{\overline{e2}_i}\chainedlink{\overline{e2}_o}{\noact}\chainedlink{\noact}{\overline{e2}_i}\chainedlink{\overline{e2}_o}{\noact}\chainedlink{\noact}{\tilde{e2}_i}\chainedlink{\tilde{e2}_o}{\noact}\chainedlink{\noact}{\hat{e2}_i}\chainend{\hat{e2}_o}.S2 \ + \ \startchain{\tilde{e2}_i}\chainedlink{\tilde{e2}_o}{\noact}\chainedlink{\noact}{\hat{e2}_i}\chainend{\hat{e2}_o}.S2\\
% && + \\
% &&  \startchain{\overline{e2}_i}\chainedlink{\overline{e2}_o}{\noact}\chainedlink{\noact}{\overline{e2}_i}\chainedlink{\overline{e2}_o}{\noact}\chainedlink{\noact}{\hat{e2}_i}\chainend{\hat{e2}_o}.\overline{S2} 
%\  + \ \startchain{\overline{e2}_i}\chainedlink{\overline{e2}_o}{\noact}\chainedlink{\noact}{\hat{e2}_i}\chainend{\hat{e2}_o}.\overline{S2}\ + \  \link{\hat{e2}_i}{\hat{e2}_o}.\overline{S2}\\
 \\
 \\
   S3 &\defeq& \startchain{s3_i}\chainedlink{s3_o}{\noact}\chainedlink{\noact}{\tilde{s3}_i}\chainedlink{\tilde{s3}_o}{\noact}\chainedlink{\noact}{\hat{s3}_i}\chainend{\hat{s3}_o}.S3 \ + \
 \startchain{s3_i}\chainedlink{s3_o}{\noact}\chainedlink{\noact}{s3_i}\chainedlink{s3_o}{\noact}\chainedlink{\noact}{\tilde{s3}_i}\chainedlink{\tilde{s3}_o}{\noact}\chainedlink{\noact}{\hat{s3}_i}\chainend{\hat{s3}_o}.S3 \ + \ \startchain{\tilde{s3}_i}\chainedlink{\tilde{s3}_o}{\noact}\chainedlink{\noact}{\hat{s3}_i}\chainend{\hat{s3}_o}.S3\\
 && + \\
 &&  \startchain{s3_i}\chainedlink{s3_o}{\noact}\chainedlink{\noact}{\hat{s3}_i}\chainend{\hat{s3}_o}.\overline{S3} 
\  + \ \startchain{s3_i}\chainedlink{s3_o}{\noact}\chainedlink{\noact}{s3_i}\chainedlink{s3_o}{\noact}\chainedlink{\noact}{\hat{s3}_i}\chainend{\hat{s3}_o}.\overline{S3}\ + \  \link{\hat{s3}_i}{\hat{s3}_o}.\overline{S3}\\
\\
\overline{S3} &\defeq& \startchain{\overline{s3}_i}\chainedlink{\overline{s3}_o}{\noact}\chainedlink{\noact}{\tilde{s3}_i}\chainedlink{\tilde{s3}_o}{\noact}\chainedlink{\noact}{\hat{s3}_i}\chainend{\hat{s3}_o}.S3 \ + \
 \startchain{\overline{s3}_i}\chainedlink{\overline{s3}_o}{\noact}\chainedlink{\noact}{\overline{s3}_i}\chainedlink{\overline{s3}_o}{\noact}\chainedlink{\noact}{\tilde{s3}_i}\chainedlink{\tilde{s3}_o}{\noact}\chainedlink{\noact}{\hat{s3}_i}\chainend{\hat{s3}_o}.S3 \ + \ \startchain{\tilde{s3}_i}\chainedlink{\tilde{s3}_o}{\noact}\chainedlink{\noact}{\hat{s3}_i}\chainend{\hat{s3}_o}.S3\\
 && + \\
 &&  \startchain{\overline{s3}_i}\chainedlink{\overline{s3}_o}{\noact}\chainedlink{\noact}{\overline{s3}_i}\chainedlink{\overline{s3}_o}{\noact}\chainedlink{\noact}{\hat{s3}_i}\chainend{\hat{s3}_o}.\overline{S3} 
\  + \ \startchain{\overline{s3}_i}\chainedlink{\overline{s3}_o}{\noact}\chainedlink{\noact}{\hat{s3}_i}\chainend{\hat{s3}_o}.\overline{S3}\ + \  \link{\hat{s3}_i}{\hat{s3}_o}.\overline{S3}\\
 \end{array}
 \]

 The code for the rules is as follow.:
 \begin{itemize}
\item Rule 1
 \[
 \begin{array}{lcl}
 R1 &\defeq & \startchain{\silent}\chainedlink{s1_i}{\noact}\chainedlink{\noact}{s1_o}\chainedlink{\overline{s3}_i}{\noact}\chainedlink{\noact}{\overline{s3}_o}\chainedlink{\tilde{s1}_i}{\noact}\chainedlink{\noact}{\tilde{s1}_o}\chainend{rule2}.R1
\ +\ 
 \startchain{\silent}\chainedlink{\overline{s1}_i}{\noact}\chainedlink{\noact}{\overline{s1}_o}\chainend{rule2}.R1
 \ +\ 
  \startchain{\silent}\chainedlink{s3_i}{\noact}\chainedlink{\noact}{s3_o}\chainend{rule2}.R1\\
 \end{array}
 \]
 \item Rule 2
  \[
 \begin{array}{lcl}
 R2 &\defeq & \startchain{rule2}\chainedlink{s2_i}{\noact}\chainedlink{\noact}{s2_o}\chainedlink{\overline{s1}_i}{\noact}\chainedlink{\noact}{\overline{s1}_o}\chainedlink{\tilde{s3}_i}{\noact}\chainedlink{\noact}{\tilde{s3}_o}\chainend{rule3}.R2\\
 &&+\\
 &&  \startchain{rule2}\chainedlink{\overline{s2}_i}{\noact}\chainedlink{\noact}{\overline{s2}_o}\chainend{rule3}.R2\\
 &&+\\
 &&  \startchain{rule2}\chainedlink{s1_i}{\noact}\chainedlink{\noact}{s1_o}\chainend{rule3}.R2\\
 \end{array}
 \]
 \item Rule 3
  \[
 \begin{array}{lcl}
 R3 &\defeq & \startchain{rule3}\chainedlink{s3_i}{\noact}\chainedlink{\noact}{s3_o}\chainedlink{\overline{s2}_i}{\noact}\chainedlink{\noact}{\overline{s2}_o}\chainedlink{\tilde{s2}_i}{\noact}\chainedlink{\noact}{\tilde{s2}_o}\chainend{stop}.R3\\
 &&+\\
 &&  \startchain{rule3}\chainedlink{\overline{s3}_i}{\noact}\chainedlink{\noact}{\overline{s3}_o}\chainend{stop}.R3\\
 &&+\\
 &&  \startchain{rule3}\chainedlink{s2_i}{\noact}\chainedlink{\noact}{s2_o}\chainend{stop}.R3\\
 \end{array}
 \]
\end{itemize}
Finally, the process $STOP$ is defined as:
\[
STOP\ \defeq \ \startchain{stop}\chainedlink{\hat{s1}_i}{\noact}\chainedlink{\noact}{\hat{s1_o}} \chainedlink{\hat{s2}_i}{\noact}\chainedlink{\noact}{\hat{s2}_o}\chainedlink{\hat{s3}_i}{\noact}\chainedlink{\noact}{\hat{s3}_o}\chainend{\silent}.STOP
\]

Now, let assume that our initial system is composed by only the element $S1$, as follow:
\[
System \ \defeq \ S1|\overline{S2}|\overline{S3}|R1|R2|R3|STOP
\]
Then, it is possible to apply only the first rule. The {\tt link}-calculus transition follows:
\[
\begin{array}{ll}
System & \rightarrow \\
&\startchain{\silent}\chainedlink{s1_i}{s1_i}\chainedlink{s1_o}{s1_o} \chainedlink{\overline{s3}_i}{\overline{s3}_i}\chainedlink{\overline{s3}_o}{\overline{s3}_o}\chainedlink{\tilde{s1}_i}{\tilde{s1}_i}\chainedlink{\tilde{s1}_o}{\tilde{s1}_o}\chainedlink{rule2}{rule2}\chainedlink{\overline{s2}_i}{\overline{s2}_i}\chainedlink{\overline{s2}_o}{\overline{s2}_o}\chainedlink{s1_i}{s1_i}\chainedlink{s1_o}{s1_o}\chainedlink{rule3}{rule3}\chainedlink{\overline{s3}_i}{\overline{s3}_i}\chainedlink{\overline{s3}_o}{\overline{s3}_o}\chainedlink{\overline{s2}_i}{\overline{s2}_i}\chainedlink{\overline{s2}_o}{\overline{s2}_o}\chainedlink{stop}{stop}\\
&
\chainedlink{\hat{s1}_i}{\hat{s1}_i}\chainedlink{\hat{s1}_o}{\hat{s1}_o}\chainedlink{\hat{s2}_i}{\hat{s2}_i}\chainedlink{\hat{s2}_o}{\hat{s2}_o}\chainedlink{\hat{s2}_o}{\hat{s2}_o}\chainedlink{\hat{s3}_i}{\hat{s3}_i}\chainedlink{\hat{s3}_i}{\hat{s3}_i}\chainedlink{\hat{s3}_o}{\hat{s3}_o}\chainend{\silent} \ System
\end{array}
\]
 \end{example}

% Let ${\cal A} = (S, A)$ be a reaction system, and let $t$ be a transition such that $t= ([|{\cal A}|] \xrightarrow{\link{\silent}{\silent}}P')$. By Definition~\ref{def:trans}, we know that
% each rule $a_j \in A$ is encoded by the process $A_j$;
% we denote as the \emph{required elements} by  $A_j$ in $t$ all the elements appearing in the link chain prefix offered by  $A_j$ in $t$.

 \begin{lemma}
 \label{lem:allrules}
  Let ${\cal A} = (S, A)$ be a reaction system, let $P=[|{\cal A}|]$ its c\CNA~translation, and let $t$
be a transition such that  if $t= (P \xrightarrow{\link{\silent}{\silent}} P')$, then
  \begin{enumerate}
  \item for each rule  $a_j \in A$, the corresponding process $A_j = [|a_j|]$,  is involved in the transition $t$;
  \item each process $P_s = [|s|]$, with $s \in S$, participates at most  with one solid link for each rule $A_j$.%, if $s$ is a \emph{required element} of $A_j$ in $t$.
  \end{enumerate}
  \end{lemma}
  \begin{proof}
  We prove the two items separately:
  \begin{enumerate}
\item by Definition~\ref{def:trans}, all the names $r_j$, where $j$ is the counter of the rules in ${\cal A}$, are restricted. Then, all the processes $A_j$ must participate to each transitions.
 \item the code of each rule leaves one virtual link between two solid links of the types $\dots\chainedlink{s_{1i}}{\noact}\chainedlink{\noact}{s_{1o}}\dots$. Then, it derives that the process $P_{s_1}$, that  codes the behaviour of element $s_1$, can participate in the transition with only one solid link $\link{s_{1i}}{s_{1o}}$. In fact, with a number $v$ of virtuals  $v > 1$, 
 $P_{s_1}$ can not participate with a link chain, as by juxtaposing  $\link{s_{1i}}{s_{1o}}$, not valid
 link chains would be generated.
    \end{enumerate}
  \end{proof}

 \begin{proposition}[completeness]
 Let ${\cal A} = (S, A)$ be a reaction system, and let $P=[|{\cal A}|]$ its c\CNA~translation, then if ${\cal A}$ evolves in ${\cal A'} = (S', A)$ in one step, then $\exists$ $P'$ such that $P \rightarrow P'$ and  $P'=[|{\cal A'}|]$.
 \end{proposition}
 \begin{proof}
 The proof proceeds by cases on the possible evolution of each element  $e \in S$, following the execution of one step in the reaction system ${\cal A} = (S, A)$.\\
 %Let  ${\cal A} = (S, A)$ be a reaction system and let $e$ be an element in the system.
 % Then 
  Let ${\cal A'} = (S', A)$ be the evolution of  ${\cal A}$ after one step. There are four cases to be considered: \ 
 1) \ $e \in S$, $e \in S'$; \ 2)\  $e \in S$, $e \notin S'$; \ 3)\  $e \notin S$, $e \notin S'$; \ 4)\  $e \notin S$, $e \in S'$.\\
 case  $e \in S$ and $e \in S'$\\
It means that have been applied some rules  in $A$ producing the element $e$, and also that some rules in $A$ requiring the element $e$ could have been applied. All the rules requiring the absence of $e$ have not been applied.\\
%Then, let $R_i = [e';e'',e]$ be a rule producing $e$ with $e'' \neq e$, and let $i$ be its associated number. \\
It is important to observe that our c\CNA encoding, for each transition, requires the application of all the rules in the system, as the channels $r_j$, with $j \in [1,m]$ are restricted.
Each rule, define a pattern to be satisfied, i.e. each rule inserts as many virtual links  as  the number of reactants, inhibitors, and products, as required by the corresponding RS rule.
It follows that

The code executed by any rule $R_i$ producing $e$ is $R_i \defeq \startchain{rule_i}\chainedlink{e'_i}{\noact}\chainedlink{\noact}{e'_o}\chainedlink{\overline{e''}_i}{\noact} \chainedlink{\noact}{\overline{e''}_o}\chainedlink{\tilde{e}_i}{\noact}\chainedlink{\noact}{\tilde{e}_o} \chainend{rule_{i+1}}.R_i + \dots$, with $e'' \neq e$.\\
The code executed by any  rules $j$ consuming $e$ without producing it is $R_j \defeq \startchain{rule_j}\chainedlink{e_i}{\noact}\chainedlink{\noact}{e_o}\chainedlink{\overline{e'}_i}{\noact} \chainedlink{\noact}{\overline{e'}_o}\chainedlink{\tilde{e''}_i}{\noact}\chainedlink{\noact}{\tilde{e''}_o} \chainend{rule_{j+1}}.R_j + \dots$, with $e',e''\neq e $.\\
The code executed by any rule $j$ not requiring $e$ is $R_j \defeq \startchain{rule_j}\chainedlink{e_i}{\noact}\chainedlink{\noact}{e_o} \chainend{rule_{j+1}}.R_j + \dots$.\\
The code executed by the element $e$ is $P_e \defeq  \underbrace{ \startchain{e_i}\chainedlink{e_o}{\noact}\chainedlink{\noact}{e_i}\chainedlink{e_o}{\noact}\chainend{\noact}...}_{size = h}\underbrace{ \startchain{\tilde{e}_i}\chainedlink{\tilde{e}_o}{\noact}\chainedlink{\noact}{\tilde{e}_i}\chainedlink{\tilde{e}_o}{\noact}\chainend{\noact}...}_{size = k}\link{\hat{e}_i}{\hat{e}_o}.P_s + \dots$, where $h$ is the number of  rules requiring $e$ and $k$ is the number of active rules producing $e$.
Then we observe that, together with the process $STOP$, the resulting processes are $R_i$, $R_j$, $P_e$ and $STOP$, i.e. the element $e$ is produced, because at least a rule $R_i$ produces $e$. \\
The other cases are similar. 
\end{proof}

 \begin{corollary}
  Let ${\cal A} = (S, A)$ be a reaction system, and let $P=[|{\cal A}|]$ its {\tt link}-calculus translation, then if ${\cal A}$ evolves in ${\cal A'} = (S', A)$  then $\exists$ $P'$ such that $P \rightarrow^* P'$ and  $P'=[|{\cal A'}|]$.  
 \end{corollary}

 \begin{proposition}[correctness]
 Let $P$ a {\tt link}-calculus process such that exists a reaction system ${\cal A}$ such that $P=[|{\cal A}|]$, then if $P \rightarrow P'$, then $\exists$ ${\cal A'}$ such that ${\cal A}$  evolves in ${\cal A'}$ in one step and $P'=[|{\cal A'}|]$.
 \end{proposition}

 \begin{corollary}
  Let $P$ a {\tt link}-calculus process such that exists a reaction system ${\cal A}$ such that $P=[|{\cal A}|]$, then if $P \rightarrow^* P'$, then $\exists$ ${\cal A'}$ such that ${\cal A}$  evolves in ${\cal A'}$ and $P'=[|{\cal A'}|]$.
 \end{corollary}

### 2020-07-27 fragments/old/06-timedComputation.tex
da dropbox

% !TEX root = ./main_RS2L.tex

\section{Timed Computation}

In the classical Reaction Systems, the actual state is the union of the elements resulting from the 
previous reaction, and of the elements that the context can provide.

The context can provides elements at certain points along the computation. Here we want to 
simulate this behaviour by introducing a technique that we call \emph{timed programmed computation}.
We introduce a two step computation: one step correspond to the transition simulating the reaction in a reaction system, the other step simulates the time passing.
Within this framework, we can introduce macros for delaying (and programming in advance) the entry
of those elements that in the Reaction Systems are provided by the context at certain points.
The macros $MACRO_j(P_s,n)$ are univocally numbered, and takes as parameters the program  $P_s$(simulating the $s$ element in the RS), and the delay $n$. Then, after $n$ computation steps  $MACRO_j(P_s,n)$  makes program $\overline{P}_s$ to turn into $P_s$, if is not already in the state $P_s$.\\
The c\CNA code for $MACRO_j(P_s,n)$ follows:
\[
\begin{array}{lcl}
MACRO_j(P_s,n) &\defeq & \link{t_j}{t_{j+1}}.MACRO_j(P_s,n-1)\\
MACRO_j(P_s,0) & \defeq & \link{t_j}{act_{s_i}}.\nil | \link{act_{s_o}}{t_{j+1}}.FWD_j\\
FWD_j & \defeq & \link{t_j}{t_{j+1}}.FWD_j\\
\end{array}
\] 
where $j = w$ and $w$ is the number of macro processes.\\
Then, we add an an additional process $START$ that decides the alternation between a RS computational step and a step simulates the time passing.

\[
\begin{array}{lcl}
START& \defeq &\link{\silent}{r_1}.\link{\silent}{t_1}.START
\end{array}
\]

Finally, we have to slightly  modify the code of $P_s$ processes, in both  active ($P_s$) or inactive ($\overline{P_s}$) state, corresponding to each element $s \in S$ of a RS system:
\[
\begin{array}{lcl@{\hspace {1cm}} lcl}
P_s &\defeq &   \dots  & \overline{P}_s &\defeq &   \dots \\
 && +  & && +\\
 && \link{act_{s_i}}{act_{s_o}}.P_s &
  && \link{act_{s_i}}{act_{s_o}}.P_s
\end{array}
\]

### 2020-08-14 TCS_v3/00-abstract.tex
nuovi file TCS v3

--- 2020-06-29 TCS_version/00-abstract.tex
@@
 In particular, we define a new assertion language based on regular expressions, 
-which allows to specify the properties of interest, and use it to extend Hennessy-Milner logic
+which allows us to specify the properties of interest, and use it to extend Hennessy-Milner logic
 to our setting.

### 2020-08-14 TCS_v3/01-introduction.tex
nuovi file TCS v3

--- 2020-07-23 TCS_version/01-introduction.tex
@@
 \item 
-\begin{color}{blue}
 RSs are encoded in a modular way, in the sense that all constituents of a RS, namely entities, reactions and the context, are seen as interacting  c\CNA \ processes; in principle, this allows to study the behaviour of any constituent in isolation, contrary to what happens in the basic RS framework, where the evolution is possible only when all constituents are given;
-\end{color}
 
@@
 in such a way that the mutated version of the entity
-$s$ can  take part to only a subset of rules requiring entity $s$;
+$s$ can  take part to only a subset of reactions requiring entity $s$;
 \item we sketch how with a little coding effort, our 

### 2020-08-14 TCS_v3/02-ReactionSystem.tex
nuovi file TCS v3

--- 2020-06-29 TCS_version/02-ReactionSystem.tex
@@
 
-\begin{definition}[Reaction]
+\begin{definition}[Reaction]\label{def:reaction}
 Let $S$ be a set of entities.
-A reaction over $S$ is a triplet $a = (R,I,P)$, where $R, I, P$ are finite,  
+A reaction over $S$ is a triple $a = (R,I,P)$, where $R, I, P$ are finite,  
 non empty subsets of $S$ and
@@
 \begin{enumerate}
-\item Let $a$ be a reaction over $S$. Then $a$ is enabled by the entities in the set $W$, denoted by $\mathit{en}_a(W)$, 
+\item Let $a\in \mathit{rac}(S)$ be a reaction over $S$. Then $a$ is enabled by the entities in the set $W$, denoted by $\mathit{en}_a(W)$, 
 if $R_a \subseteq W$ and $I_a \cap W = \emptyset$, i.e. all the reactants of $a$ are in $W$,
@@
 $\mathit{res}_a(W) \defeq P_a$, if $\mathit{en}_a(W)$, and $\mathit{res}_a(W) \defeq \emptyset$ otherwise.
-\item Let $A$ be a finite set of reactions over $S$. The result of $A$ on $W$, denoted 
+\item Let $A\subseteq \mathit{rac}(S)$ be a (finite) set of reactions over $S$. The result of $A$ on $W$, denoted 
 by $\mathit{res}_A(W)$,  is defined

### 2020-08-14 TCS_v3/05-translation.tex
nuovi file TCS v3

--- 2020-06-29 TCS_version/05-translation.tex
@@
  %involved in the transition $t$;
-  \item for each reaction $aj \in A$ and each entity $s \in S$, each virtual link offered by processes $P_a$ and $\mathit{Cxt}_1$ is  overlapped by exactly one solid link offered by  processes representing entities.
+  \item for each reaction $aj \in A$ and
+  % each entity $s \in S$, 
+  each virtual link offered by processes $P_a$ and $\mathit{Cxt}_1$ is  overlapped by exactly one solid link offered by  processes representing entities.
   %each process $P_s $ (or its alternative version $\overline{P}_s$), with $s \in S$, participates at most  with one solid link for each process $P_a$, and with exactly one solid link for each process $Cxt_s$.%, if $s$ is a \emph{required element} of $A_j$ in $t$.

### 2020-08-14 TCS_v3/AnswerToReviewer.txt
alcuni cambi

Dear Editor and Referees,

we thank you very much for your detailed comments and remarks, which greatly helped us to correct and improve our presentation. 
We updated the manuscript following the advices of the referees. Please find below a detailed description of the way in which 
the items pointed out by the referees were addressed in the revised version.

Best regards,
Linda Brodo, Roberto Bruni, Moreno Falaschi

-------------------------

Manuscript Number: TCS-D-20-00267R1
Manuscript Title: A process algebraic approach to reaction systems.
Author(s): Linda Brodo, Researcher; Roberto Bruni, Professor; Moreno Falaschi, Professor
Submitted to: Theoretical Computer Science

-------------------------
Reviewer #1: I think my original comments have been addressed satisfactorily by the authors, but I still have some remarks:

p.6, l.12 - replace "set of entities" by "sets of entities"
> OK. Done

p.7, l.42 - replace "where v_i is a link chain" by "where v is a link chain" 
as in the syntax from the previous line you use v.P and not v_i.P
> OK. Done

p.8, l.13, (Res) rule - make sure all a's have same font as it seems the one in (\nu a)P' is different
> OK. Done

p.8, l.30 - replace "Fig.1" by "Fig. 1"
> OK. Done

p.12, l.35 (item 2 of Lemma 12) - replace "for each reaction" by "each reaction"
> OK. We have corrected the formulation.

p.12, l.51 - in Def. 13 shouldn't be P \xrightarrow{(\nu names)v} P' instead of P \xrightarrow{v}P' ?  Only starting on p.15, l.52 you say "In this section, for simplicity, we shall often omit topmost restrictions (\nu names)". Thus, either move this assumption earlier or till Section 5 keep using (\nu names) on \xrightarrow

p.13, l.56 - (\nu names) missing from \xrightarrow label

p.17, l.49 - replace "(starting at cxt and ending at p1" by "(starting at cxt and ending at p1)", namely add missing bracket

p.33, l.41 - shouldn't be "j \geq 1" (as assumed on p.33, l.38) instead of "j \geq 0"?

p.33, l.47 and l.49 - in the proof of the base case "j=1", instead of "j-1" shouldn't be "j"? This is due to the fact that Prop. 16 you mention you have the case j=1 and also on p.33 on l.52 you got \gamma^j and not \gamma^{j-1}.

### 2020-08-16 TCS_v3/AnswerToReviewer.txt
cambi vari

@@
 Thus, either move this assumption earlier or till Section 5 keep using (\nu names) on \xrightarrow
+>  ***TO BE ANSWERED***
 
 p.13, l.56 - (\nu names) missing from \xrightarrow label
+>  ***TO BE ANSWERED***
 
 p.17, l.49 - replace "(starting at cxt and ending at p1" by "(starting at cxt and ending at p1)", namely add missing bracket
+> OK. Done
 
 p.33, l.41 - shouldn't be "j \geq 1" (as assumed on p.33, l.38) instead of "j \geq 0"?
+> OK. Done
 
@@
 16 you mention you have the case j=1 and also on p.33 on l.52 you got \gamma^j and not \gamma^{j-1}.
+>  ***TO BE DONE/ANSWERED***

### 2020-08-16 TCS_v3/06-examples.tex
cambi vari

--- 2020-07-24 TCS_version/06-examples.tex
@@
 \]}
-The parts in bold are provided by the entity processes, the other parts are provided by the processes encoding the reactions and by the process encoding the context (starting at $\mathit{cxt}$ and ending at $p_1$.
+The parts in bold are provided by the entity processes, the other parts are provided by the processes encoding the reactions and by the process encoding the context (starting at $\mathit{cxt}$ and ending at $p_1$).
 In the label we can read that reactions $1$ and $4$ have been not executed because the entity $a$ is absent, 

### 2020-08-27 TCS_v3/AnswerToReviewer.txt
ringraziamenti progetti e indice j nella prova

@@
 Thus, either move this assumption earlier or till Section 5 keep using (\nu names) on \xrightarrow
->  ***TO BE ANSWERED***
+>  No, Def.13 is correct: you can imagine that P contains no restrictions and therefore P \xrightarrow{v}P' is such that all the names in v are visible.
+Then (\nu names)P \xrightarrow{(\nu names)v} (\nu names)P' but in the label (\nu names)v all the names are transformed to silent actions \tau.
 
 p.13, l.56 - (\nu names) missing from \xrightarrow label
->  ***TO BE ANSWERED***
+>  No, see above.
 
@@
 16 you mention you have the case j=1 and also on p.33 on l.52 you got \gamma^j and not \gamma^{j-1}.
->  ***TO BE DONE/ANSWERED***
+>  We imagine you meant "in the proof of the inductive case". Yuo are right, we have slightly expanded the proof to make it clear the application of the inductive hypothesis.

### 2020-08-27 TCS_v3/09-appendix.tex
cambi vari
ringraziamenti progetti e indice j nella prova

--- 2020-07-23 TCS_version/09-appendix.tex
@@
 \begin{proof}
-We proceed by induction on the transition number $j \geq 0$.
+We proceed by induction on the transition number $j \geq 1$.
 % case $j=0$\\
@@
 \item[inductive case:]
-We assume, by inductive hypothesis, that $\exists P'$ such that 
-$$P \xrightarrow{\link{\silent}{\silent}\dots \link{\silent}{\silent}}^{j-1} P'$$ 
-and $P' = \llbracket{\cal A},\gamma^{j-1}\rrbracket$.
+We assume that the property holds for some generic $j\geq 1$ and we prove that it holds for $j+1$.
+Suppose that
+$$P \xrightarrow{\link{\silent}{\silent}\dots \link{\silent}{\silent}}^{j+1} P'' .$$
+Then, it must exist $P'$ such that 
+$$P \xrightarrow{\link{\silent}{\silent}\dots \link{\silent}{\silent}}^{j} P' \xrightarrow{\link{\silent}{\silent}\dots \link{\silent}{\silent}} P'' .$$ 
+By inductive hypothesis, $P' = \llbracket{\cal A},\gamma^{j}\rrbracket$.
 %, with $1 \leq  i < t-1$.\\
-As $P'$ is the encoding of an extended interactive process, by Proposition~\ref{prop:corr1}, it exists $P''$  such that $P' \xrightarrow{\link{\silent}{\silent}\dots\link{\silent}{\silent}} P''$,
- and $P'' = \llbracket{\cal A},\gamma^j\rrbracket$. 
+As $P'$ is the encoding of an extended interactive process, by Proposition~\ref{prop:corr1},  $P'' = \llbracket{\cal A},(\gamma^j)^1\rrbracket =\llbracket{\cal A},\gamma^{j+1}\rrbracket $. 
 % By Lemma~\ref{lem:compl2}.2, if $\exists$ $P'$ such that $P'' \xrightarrow{\link{\silent}{\silent}} P'$, then $P'$ is 

### 2020-08-31 TCS_v3/08-conclusion.tex
piccoli cambi

--- 2020-06-29 TCS_version/08-conclusion.tex
@@
 with other formal techniques to prove 
-properties of the modeled systems, along the lines in~\cite{CFHOT15,OCHF16,BBGLBH2017}.
+properties of the modeled systems, along the lines in~\cite{CFHOT15,OCHF16,BBGLBH2017,BBF15}.
 Moreover, we are considering possible enhancements of RSs based on entity mutation and on the possibility for two RSs to exchange entities.
@@
 the {\tt link}-calculus~\cite{BrodoO17} that can be found in~\cite{tool}.
+
+\medskip
+
+\noindent
+{\small {\bf Acknowledgments}
+We thank the anonymous reviewers for their detailed and very useful 
+criticisms and recommendations that helped us to improve our paper.}
